### 2011-11-10 joint action and knowing others minds/joint actions knowing others minds.tex
first

%!TEX TS-program = xelatex
%!TEX encoding = UTF-8 Unicode

\documentclass[12pt,a4paper]{extarticle}
% extarticle is like article but can handle 8pt, 9pt, 10pt, 11pt, 12pt, 14pt, 17pt, and 20pt text

\def \ititle {Joint Action and Knowing Others' Minds:}
\def \isubtitle {How social interaction grounds mindreading}
\def \iauthor {Stephen A. Butterfill}
\def \iemail{s.butterfill@warwick.ac.uk}
%\date{}

\input{$HOME/Documents/submissions/preamble_steve_paper}

\begin{document}

\setlength\footnotesep{1em}

\bibliographystyle{newapa} %apalike

\maketitle
%\tableofcontents

\begin{abstract}
***
\end{abstract}

\section{The Challenge}
%To say that something \emph{grounds} cognition of propositional attitudes is to say that it partially explains how such cognition emerges in evolution or development (or both).
I want to start with a challenge.  
The challenge is to explain the emergence, in evolution or development (or both), of sophisticated forms of social cognition.

By `sophisticated forms of social cognition', I mean those which involve representing perceptions, knowledge states, intentions, beliefs and other propositional attitudes.
In other words, the sort of social cognition measured by tasks such as the famous false belief task \citep{Wimmer:1983dz}.

Why think that representing beliefs and other propositional attitudes is a sophisticated form of social cognition?
Part of the reason is that representing propositional attitudes is hard in two senses.  
A body of evidence with humans suggests that representing  beliefs and other propositional attitudes requires conceptual sophistication, 
%
\begin{itemize}
\item for it has a protracted developmental course stretching over several years \citep{Wimmer:1983dz,Wellman:2001lz}, 
\item its acquisition is tied to the development of executive function \citep{Perner:1999yr,Sabbagh:2006ke} and language \citep{Astington2005ot}, things which two-year-olds, scrub jays and chimpanzees are deficient in, and   
\item development of reasoning about beliefs in humans is be facilitated by explicit training \citep{Slaughter:1996fv} and environmental influences, such as siblings \citep{Clements:2000nc,Hughes:2004zj}.  
\end{itemize}
%
Representing propositional attitudes also appears to be cognitively demanding, requiring attention and working memory in fully competent adults \citep{Apperly:2008jv,Apperly:2009cc,McKinnon:2007rr}.\footnote{
These findings, like the developmental findings mentioned in the previous paragraph, are open to challenge \citep[e.g.][]{Leslie:2005ef}.  
There is a wide variety of positions in this area; we provide indirect support for the view taken here elsewhere \citep{Apperly:2009ju,butterfill_minimal}.
}

It makes sense that representing propositional attitudes should be conceptually and cognitively demanding.  
After all, these are states which form complex causal structures, have arbitrarily nest-able contents, and are individuated by their causal and normative roles in explaining thoughts and actions.  
If anything should take years to acquire and consume scarce cognitive resources it is surely cognition of states with that combination of properties.

%\footnote{*There is a standard distinction between more basic forms of social cognition, which include joint attentional abilities and emotional engagement, and more sophisticated forms of social cognition which paradigmatically involve ascribing beliefs, knowledge states, intentions and other propositional attitudes.  Here we are concerned exclusively with the more sophisticated forms of social cognition.}

So the challenge is to explain the emergence, in evolution or development (or both), of sophisticated forms of social cognition---that is, of cognition that involves representing  perceptions, beliefs, intentions and other propositional attitudes.

\section{The Conjecture}
So much for the challenge.
How can we meet it?

Several researchers have conjectured that social interaction partially explains how sophisticated forms of cognition, including social cognition, emerge in development or evolution:
%
\begin{quote} 
`the unique aspects of human cognition ... were driven by, or even constituted by, social co-operation'
\citep[p.\ 1]{Moll:2007gu}.
\end{quote}
%
\begin{quote} 
`perception, action, and cognition are grounded in social interaction%
% … functions traditionally considered hallmarks of individual cognition originated through the need to interact with others
' \citep[p.\ 103]{Knoblich:2006bn}.
\end{quote}
%
These conjectures concern social interaction or social co-operation generally.
In fact these researchers focus' is on joint action.
And, like them, I want to focus on joint action.

So the conjecture I am interested is this:
%
\begin{quote}
The existence of abilities to engage in joint action partially explain how sophisticated forms of social cognition emerge in evolution or development (or both).%
\footnote{
[*don't say because only matters given the later argument]
Note that this is explanation \emph{how}.  
Distinguish explanation in the sense of (we have sophisticated cognition in order that we can engage in joint action) from explanation in the sense of (abilities to engage in joint action explain how sophisticated forms of cognition emerge in evolution or development).
}
\end{quote}
%
Faced with this conjecture, 
\begin{comment}
you might be wondering two things.
The first is what `sophisticated forms of social cognition' are.
I'll say more about this later; to anticipate, I will argue that ascriptions of beliefs, desires, intentions and other propositional attitudes are instances of sophisticated social cognition.
The other thing 
\end{comment}
you might be wondering is what joint action is.
That is my question too:
%
\begin{quote}
If this conjecture is correct, what could joint action be?
\end{quote}
%
Paradigm cases of joint action in development include:
%
\begin{itemize}
\item tidying up the toys together 
(Behne et al 2005)
\item cooperatively pulling handles in sequence to make a dog-puppet sing 
(Brownell et al 2006)
\item bouncing a ball on a large trampoline together 
(Tomasello and Carpenter 2007)
\item pretending to row a boat together
\end{itemize}
%
Philosophers' paradigm cases of joint action include painting the house together (Michael Bratman), lifting a heavy sofa together (David Velleman), preparing a hollandaise sauce together (John Searle), going to Chicago together (Christopher Kutz), and walking together (Margaret Gilbert).

So one constraint on any adequate answer to my question is that it must be consistent with the claim that these paradigm cases are indeed joint actions.
It is  these paradigm cases that given us an initial, intuitive anchor on what joint action is.
Beyond this I take everything to be open to question.

\section{Further Question (I may decide to cut this)}
I started with a Challenge, which was to explain how sophisticated forms of social cognition emerge in evolution or development, 
and a Conjecture, which was that joint action plays a role in explaining this.
Do the notions of collective and shared goals help us to understand how joint action could explain the emergence of sophisticated forms of social cognition?

How might abilities to engage in the sorts of joint action 
which involve collective or shared goals 
be involved in the evolution or development (or both) of sophisticated social cognition?

My suggestion concerns identifying the goals of actions. 
It is one thing to have a general ability to recognise goals and quite another to be able to recognise the goals of this particular activity.
To illustrate, consider Hare and Tomasello (2004).
The pictures stand for what participants in this experiment saw.
The participants were chimpanzees.
The question what whether the participants would be able to work out which of two containers contained a reward.
On the left there is a chimpanzee who is trying but failing to reach for the reward. 
Chimpanzees have no problem getting the reward in this case, suggesting that they understand the goal of the failed reach.
On the right there is a human pointing to the reward location.
Chimpanzees do not reliably  get the reward in this case, suggesting that they fail to understand the goal of the pointing action.
This is one illustration of how identifying the goals of particular actions can be difficult.

Some of the most plausibly unique aspects of human cognition depend on our abilities to recognise the goals of novel behaviours involving tools and gestures.  
\textbf{In particular, communicative actions depend on our abilities to recognise as the goals of behaviours goals which the behaviours can serve only because we recognise the behaviours as serving those goals [the Gricean circle].}
It is here that I think joint action can play a role in explaining aspects of human cognition.  
My suggestion will be, roughly, that abilities to engage in joint action provide a route to knowledge of others’ goals which is distinct from ordinary third-person interpretation.

To explain this suggestion in detail I first need to describe some reasons why it can be hard to identify the goals of particular actions …

\subsection{The Problem of Opaque Means}

Suppose you cannot gain knowledge of the current goals of another's actions through linguistic communication.

Suppose the goal is relatively novel (there are no stereotypical indications).

In this situation we cannot generally do better to work backwards from her behaviours to her goals: we determine which outcomes her behaviour is likely to bring about and then suppose that her goal is to bring about one or more of these outcomes (Dennett 1991).

But this method doesn’t work when: 
%
\begin{list}{*}{}
\item we don’t know which outcomes the observed behaviour is likely to bring about;
\item we, or the agent under observation, have false beliefs relevant to which outcomes this behaviour is likely to bring about; or
%\item there are many possible outcomes.
\end{list}
%
So one obstacle to identifying the goals of particular actions is the problem of opaque means …

Another problem involves false belief but I won't mention that here.\footnote{
 The interdependent roles of beliefs, desires and goals in producing action mean there will be many cases where observed behaviours are compatible with different ascriptions. To illustrate, consider Maya who is tidying shapes into two boxes. She mostly puts the squares into Leo’s box.
This indicates that the goal of her activity may be to put the squares into Leo’s box, but it also leaves open the possibility that her goal is to put the squares into Charlie’s box and she has a false belief about the owners of the boxes. In general, non-communicative behaviour indicates what an agent’s goals are only given assumptions about her beliefs, and it indicates what her beliefs are only given assumptions about her goals (Davidson 1974 [1984]).18	In some everyday situations this interdependence is a practical problem for knowing what others are doing. We could solve the problem if we had some way of getting at an agent’s goals independently of knowing what she believes.
 }

\subsection{The your-goal-is-my-goal route to knowledge}

How could abilities to engage in joint action provide us with knowledge of others’ goals?   The intuitive idea I started with was this: if you’re engaged in joint action with me, it’s easy for me to know what your goal is … because your goal is my goal.

This intuitive idea isn’t quite right as it stands.  For to be engaged in joint action requires that I already have expectations about the goals of your behaviours.  
So engaging in joint action presupposes rather than explains knowledge of others’ goals.  Or so it seems.

But there is a way around this.  For there are various cues that you can give me which signal that you are about to engage in joint action with me.  Seeing me struggling to get my twin pram on to the bus, you grab the front wheels and make eye contact, raising your eyebrows and smiling.  In this way you signal that you both disposed to help and are about to engage in joint action with me.  This makes it trivial for me to know what the goal of your behaviour is: your goal is my goal, to get the pram onto the bus.

My suggestion, then, is that the following inference characterises a route to knowledge of others’ goals:
%
\begin{enumerate}
\item We are about to engage in some joint action\footnote{
*What notion of joint action is needed here?  Any will do as long as it involves distributive goals.
}
or other (for example, because you have made eye contact with me while I was in the middle of attempting to do something).

\item I am not about to change my goal.

\end{enumerate}
%
Therefore:
%
\begin{enumerate}[resume]
%
\item The others will each individually perform actions directed to my goal.
\end{enumerate}
%
Call this the ‘your-goal-is-my-goal’ route to knowledge.  To say that this inference characterises a route to knowledge implies two things.  First, in some cases it is possible to know the three premises, 1–2, without already knowing the conclusion, 3.  Second, in some cases knowing the two premises puts one in a position to know the conclusion.  I take both points to be true.

The your-goal-is-my-goal route to knowledge is characterised by an inference.  However, exploiting this route to knowledge may not require actually making the inference or knowing the premises.  Depending on what knowing requires, it may be sufficient to believe the conclusion because one has reliably detected a situation in which the premises of the inference are true without necessarily being able to think of this as a situation where the premises are true.

\subsection{Application}
I want to suggest that your-goal-is-my-goal might give us a way to understand how joint action facilitates a transition from a simple understanding of goals to an understanding of communicative intent.

I already mentioned Hare and Call's (\citeyear{hare_chimpanzees_2004}) experiment which contrasts pointing with a failed reach as two ways of indicating which of two closed containers a reward is in.  
Chimps can easily interpret a failed reach but are stumped by the point to a closed container.

In discussing this experiment, Moll and Tomasello say:
%
\begin{quote}
`to understand pointing, the subject needs to understand more than the individual goal-directed behaviour. She needs to understand that ... the other attempts to communicate to her ...  and ... the communicative intention behind the gesture'
(Moll \& Tomsello 2007)
\end{quote}
%
Of course I don't want to question this assertion.
But I do want to suggest that in the context of joint action there is a way to respond reliably to informative pointing without understanding pointing at all.
For if one knows that one is engaged in joint action with the person producing the point, one already knows what the (long-term) goal of the pointing action.  
The goal of the pointing is my goal, which is to find the reward.
So in the context of a joint action, it should be no harder to understand the point than it is to understand the failed reach.
Both are attempts to get the reward.

The pointing action, unlike the failed reach, is an \textbf{opaque means} of getting the object.  But in the context of joint action this doesn't matter because the your-goal-is-my-goal tells you that the goal of the point is to get the object.

\textbf{The `my goal is your goal' inference enables you to treat pointing as having the same goal as the failed reach.}
This amounts to \emph{misunderstanding} pointing, of course.  
(The communicators' goal is unlikely to be your goal.)
But the misunderstanding is fruitful in the sense that it enables you to respond appropriately to the pointing, to make us of it.

So it is possible that the combination of minimal mindreading abilities with abilities to share goals is sufficient for understanding pointing actions in the context of joint actions.

As Ulf Liszkowski's has demonstrated in a series of experiments, humans are unlike chimpanzees in they can understand and produce communicative actions involving pointing to inform (*refs).
In fact human children's early abilities to understand and produce pointing gestures appear early in the second year of life.
What I'm suggesting is that the emergence of these abilities might be facilitated by joint action.
For it is possible to understand pointing gestures without \emph{already} understand communicative intent.

\textbf{
The idea is that in the context of joint action, communicative actions can be fruitfully misunderstood as ordinary goal-directed actions.
But once an action has been given a function in joint action, it can be used to serve that function outside joint action contexts.  And so it becomes genuinely communicative.
}

Note that I am not suggesting that young children might fail to understand communicative intent.
I am suggesting that they might first understand the goals of pointing actions without understanding communicative intent.
But of course once they understand the goals of pointing actions within the context of joint action, it's likely that they will be able to understand the goals of pointing actions outside the context of joint action too.\footnote{
Contrast Csibra's `two stances' idea. The referential action understanding involves a “stance” (p. 455); teleological and referential action interpretation “rely on different kinds of action understanding' \citep[p.\ 456]{Csibra:2003kp}; they are initially two distinct `action interpretation systems' (although of course they come together later in development)  \citep[p.\ 456]{Csibra:2003kp}.
In relation to Csibra's, my suggestion is not that there is no referential stance.
It's rather that the referential stance might emerge from what he calls the `teleological stance' together with abilities to engage in the sort of joint actions that are characterised by shared goals.
}

Joint action may explain how individuals starting with a simple understanding of goals end up understanding communicative intentions.

This is one illustration of how capacities for joint action, even very simple forms of joint action, might be relevant to explaining the development or evolution of richer forms of cultural cognition.

\section{Conclusion}
In conclusion I have suggested that there are forms of joint action which require only limited social cognition and which may play a role in the emergence of more sophisticated forms of social cognition, in development or evolution (or both).

[See last slide with diagram] On emergence, the idea was that abilities to engage in joint action combined with minimal social cognition enable humans to break into the Gricean circle and understand communicative intention.
This is turn is one of the foundations on which abilities to communicate by language are built,
and there is evidence that abilities to communicate by language in turn play a role in the emergence of full-blown mindreading abilities (*refs).
So this may be one route by which abilities to engage in joint action plus limited social cognition plays a role in the emergence of sophisticated forms of social cognition such as cognition of belief and other propositional attitudes.

\bibliography{$HOME/endnote/phd_biblio}

\end{document}

### 2011-11-16 joint action and knowing others minds/interacting mindreaders.tex
starting

%!TEX TS-program = xelatex
%!TEX encoding = UTF-8 Unicode

\documentclass[12pt,a4paper]{extarticle}
% extarticle is like article but can handle 8pt, 9pt, 10pt, 11pt, 12pt, 14pt, 17pt, and 20pt text

\def \ititle {Interacting Mindreaders}
\def \isubtitle {}
\def \iauthor {Stephen A. Butterfill}
\def \iemail{s.butterfill@warwick.ac.uk}
%\date{}

\input{$HOME/Documents/submissions/preamble_steve_paper}

\begin{document}

\setlength\footnotesep{1em}

\bibliographystyle{newapa} %apalike

\maketitle
%\tableofcontents

\begin{abstract}
***
\end{abstract}

\section{What is a theory of interpretation?}
What grounds ascriptions of thought and action to others?
What is the evidential basis for such ascriptions and how does the evidence support the ascriptions?
This is the  question  a theory of interpretation answers.

Answering this question is essential for understanding epistemic aspects of mindreading.  
For instance, in saying that some principle serves as a heuristic in mindreading we are implicitly relying on a theory of interpretation to identify standards relative to which it is a heuristic.

Note that the question a theory of interpretation answers is a question about justification.
It is not about about \emph{how} anyone actually ascribes thoughts or identifies actions. 
Nor is it (at least not immediately) a question about the mechanisms, processes or representations involved in mindreading.
To stress this, Davidson sometimes formulates the issue by asking what someone \emph{could} know that would put them in a position to know another's thoughts and actions \citep[e.g.][p.\ 126]{Davidson:1973jx}.
A theory of interpretation is just one part of an account of mindreading.

It is perhaps fair to say that most researchers regard

The two most sustained and elaborate attempts to provide a theory of interpretation,
Davidson's (\citeyear{Davidson:1984wh}) and Dennett's (\citeyear{Dennett:1987sf}),
have ignored the possibility of interaction.
The evidential basis they consider and the principles which they identify to link the evidence with ascriptions of thought and action are available to entirely passive observers.
We shall argue that mindreaders who are capable of interacting with the targets of ascriptions are at an advantage.
Their ascriptions can be justified by evidence and principles which would be unavailable if they were entirely passive observers.

Being poised to interact with others enables one to know things about their minds which one might not otherwise be in a position to know; or so we aim to show in what follows.
Because orthodox theories of interpretation have supposed that being able to interact makes no difference,
they have treated the evidential basis for interpretation as more narrow than it truly is.
This matters because broadening the evidential basis will not only enable us to give a more comprehensive theory of interpretation; 
it will also provide tools for explaining the emergence, in evolution or in development, of mindreading.

***Distinguish from `second person', interaction (Gallagher and Hutto?)

To develop the idea that 
interaction broadens the evidential basis for interpretation
we do not need to throw away the leading, most critically examined theories of interpretation due to Davidson and Dennett.
An entirely fresh start is unnecessary.
Of course we do not mean to assert, positively, that there are not devastating objections to these theories; perhaps there are.
But considerations about the role of interaction in theories of interpretation do not give rise to any such objections.
The best way to understand how interaction bears on epistemic and normative aspects of mindreading is to extend rather than to reject the leading theories.

Since we shall discuss the role of interaction in theories of interpretation as an extension of leading theories (without any commitment here to the correctness of these theories---any theory of interpretation will have to accommodate considerations about interaction), it will be useful to start with a brief outline of those theories.

\section{The leading theory of interpretation}

Donald Davidson argues that language is necessary for propositional thought \citep[p.\ 130]{Davidson:1999mo},
%What more is needed for thought? I think the answer is language.
and that

\bibliography{$HOME/endnote/phd_biblio}

\end{document}

### 2011-11-18 joint action and knowing others minds/interacting mindreaders.tex
minor

@@
 To stress this, Davidson sometimes formulates the issue by asking what someone \emph{could} know that would put them in a position to know another's thoughts and actions \citep[e.g.][p.\ 126]{Davidson:1973jx}.
-A theory of interpretation is just one part of an account of mindreading.
 
-It is perhaps fair to say that most researchers regard
+So a theory of interpretation is just one part of a full account of mindreading.
+A full account would need to address questions 
+about how mindreaders actually ascribe mental states, 
+about what mental states are, 
+and 
+about the evolution and development of mindreading,
+among others.
+Although a theory of interpretation will constrain and be constrained by answers to these questions,
+none of these questions are directly addressed or answered by a theory of interpretation.
 
@@
 An entirely fresh start is unnecessary.
-Of course we do not mean to assert, positively, that there are not devastating objections to these theories; perhaps there are.
-But considerations about the role of interaction in theories of interpretation do not give rise to any such objections.
-The best way to understand how interaction bears on epistemic and normative aspects of mindreading is to extend rather than to reject the leading theories.
+Of course we do not mean to assert, positively, that there are no good objections to these theories; perhaps there are.
+Our point is just that considerations about the role of interaction in theories of interpretation do not give rise to any such objections.
+A good way to understand how interaction bears on epistemic and normative aspects of mindreading is to extend rather than to reject the leading theories.
 
-Since we shall discuss the role of interaction in theories of interpretation as an extension of leading theories (without any commitment here to the correctness of these theories---any theory of interpretation will have to accommodate considerations about interaction), it will be useful to start with a brief outline of those theories.
+Since we shall discuss the role of interaction in theories of interpretation as an extension of leading theories (without any commitment here to the correctness of these theories; any theory of interpretation will have to accommodate considerations about interaction), it will be useful to start with a brief outline of those theories.
 

### 2011-11-25 joint action and knowing others minds/interacting mindreaders.2.tex
v2 started (narrower focus on goals)

--- 2011-11-16 joint action and knowing others minds/interacting mindreaders.tex
@@
 
-\section{What is a theory of interpretation?}
-What grounds ascriptions of thought and action to others?
-What is the evidential basis for such ascriptions and how does the evidence support the ascriptions?
-This is the  question  a theory of interpretation answers.
-
-Answering this question is essential for understanding epistemic aspects of mindreading.  
-For instance, in saying that some principle serves as a heuristic in mindreading we are implicitly relying on a theory of interpretation to identify standards relative to which it is a heuristic.
-
-Note that the question a theory of interpretation answers is a question about justification.
-It is not about about \emph{how} anyone actually ascribes thoughts or identifies actions. 
-Nor is it (at least not immediately) a question about the mechanisms, processes or representations involved in mindreading.
-To stress this, Davidson sometimes formulates the issue by asking what someone \emph{could} know that would put them in a position to know another's thoughts and actions \citep[e.g.][p.\ 126]{Davidson:1973jx}.
-A theory of interpretation is just one part of an account of mindreading.
-
-It is perhaps fair to say that most researchers regard
-
-The two most sustained and elaborate attempts to provide a theory of interpretation,
-Davidson's (\citeyear{Davidson:1984wh}) and Dennett's (\citeyear{Dennett:1987sf}),
-have ignored the possibility of interaction.
-The evidential basis they consider and the principles which they identify to link the evidence with ascriptions of thought and action are available to entirely passive observers.
-We shall argue that mindreaders who are capable of interacting with the targets of ascriptions are at an advantage.
-Their ascriptions can be justified by evidence and principles which would be unavailable if they were entirely passive observers.
-
-Being poised to interact with others enables one to know things about their minds which one might not otherwise be in a position to know; or so we aim to show in what follows.
-Because orthodox theories of interpretation have supposed that being able to interact makes no difference,
-they have treated the evidential basis for interpretation as more narrow than it truly is.
-This matters because broadening the evidential basis will not only enable us to give a more comprehensive theory of interpretation; 
-it will also provide tools for explaining the emergence, in evolution or in development, of mindreading.
-
-***Distinguish from `second person', interaction (Gallagher and Hutto?)
-
-To develop the idea that 
-interaction broadens the evidential basis for interpretation
-we do not need to throw away the leading, most critically examined theories of interpretation due to Davidson and Dennett.
-An entirely fresh start is unnecessary.
-Of course we do not mean to assert, positively, that there are not devastating objections to these theories; perhaps there are.
-But considerations about the role of interaction in theories of interpretation do not give rise to any such objections.
-The best way to understand how interaction bears on epistemic and normative aspects of mindreading is to extend rather than to reject the leading theories.
-
-Since we shall discuss the role of interaction in theories of interpretation as an extension of leading theories (without any commitment here to the correctness of these theories---any theory of interpretation will have to accommodate considerations about interaction), it will be useful to start with a brief outline of those theories.
-
-\section{The leading theory of interpretation}
-
-Donald Davidson argues that language is necessary for propositional thought \citep[p.\ 130]{Davidson:1999mo},
-%What more is needed for thought? I think the answer is language.
-and that
+\section{Goal ascription}
+Identifying which goals others' actions are directed to enables us to predict and manipulate their actions, and to emulate their achievements.
 

### 2011-11-25 joint action and knowing others minds/interacting mindreaders.tex
v2 started (narrower focus on goals)

@@
 
-\section{What is a theory of interpretation?}
+\section{Davidson's question: What grounds mindreading?}
+Many people expect to know what those around them think, within limits.
+If someone were to complain of a co-worker that she can never tell what she  thinking,
+the complaint might be informative, surprising even.
+And we might reasonably wonder where the fault lies.
+Maybe the co-worker really is a closed book,
+but maybe complainant is not really trying to understand her.
+For humans and perhaps other species too,
+access to the minds of those nearby,
+to some of their beliefs and desires,
+emotions, intentions and knowledge states,
+is often the norm in this sense: it is lack of access, not access, that is remarkable.
+And often enough we read others' minds spontaneously
+without having to stop and think.\footnote{
+***fluently --- narrative interpretation
+***spontaneously --- Rubio and Ferguson, Senju et al
+}
+In this respect we might compare identifying others' thoughts with identifying their masses.
+And although both can be disguised with the right props and techniques, some degree of revelation is normal.
+
+This mundane fact about minds raises many questions.
+Some concern mechanism, how one subject is able to discover facts about another's mind.
+Another set of questions focuses on the evolution of mindreading and the costs and benefits of being a mindreader or a target for mindreading.
+Our concern here is not with any of these questions.
+Instead we investigate a more narrowly epistemic question.
 What grounds ascriptions of thought and action to others?
 What is the evidential basis for such ascriptions and how does the evidence support the ascriptions?
-This is the  question  a theory of interpretation answers.
+
+To stress that this question is not directly about about \emph{how} anyone actually ascribes thoughts or identifies actions% 
+% nor directly about the mechanisms, processes or representations involved in mindreading
+, Davidson (who provided the best developed, most systematic answer to our question)
+sometimes formulates the question by asking what someone \emph{could} know that would put them in a position to identify another's thoughts and actions \citep[e.g.][p.\ 126]{Davidson:1973jx}.
+
+We can clarify this question
+
+%It can be informative, surprising even
+%to say that someone is a closed book,
+%that she seems to inhabit a world of her own
+%and that you cannot tell what she is thinking.
+%And if you did say this, we might reasonably wonder whether the blame lies with you rather than her.
+
+The two most sustained and elaborate attempts to answer this question,
+Davidson's (\citeyear{Davidson:1984wh}) and Dennett's (\citeyear{Dennett:1987sf}),
+do not exploit the possibility of interaction.
+The evidential basis they consider and the principles which they identify to link the evidence with ascriptions of thought and action are available to entirely passive observers.
+So on these theories,
+a purely passive mindreader observing from behind a one-way mirror
+is on a par with
+a mindreader who does or could interact with those she seeks to interpret.
+The two are on a par in this sense:
+in principle the same evidence could be available to each, and each can exploit the same principles in moving from evidence to ascriptions of thought.
+We shall argue that mindreaders who are capable of interacting with those they aim to interpret are at an advantage.
+Their ascriptions can be justified by evidence and principles which would be unavailable if they were entirely passive observers.
+
+Being poised to interact with others enables one to know things about their minds which one might not otherwise be in a position to know; or so we aim to show in what follows.
+Because orthodox theories have supposed that being able to interact makes no difference,
+they have treated the evidential basis for mindreading as more narrow than it truly is.
+This matters because broadening the evidential basis will not only enable us to give a more comprehensive theory of mindreading; 
+it will also provide tools for explaining the emergence, in evolution or in development, of mindreading.
+
+\section{Kinds of mindreading}
+Davidson's account of what grounds ascriptions of thought is also an account of what grounds ascriptions of meaning.
+In `Radical Interpretation'
+(\citeyear{Davidson:1973jx}), Davidson suggests that the evidential basis for mindreading 
+(he uses the term `interpretation')
+is of this form:
+%
+\begin{idescription}
+\item[(E)] At time $t$, Ayehsa holds $S$ true and $p$.
+\end{idescription}
+%
+Such evidence confirms or falsifies generalisations of this form:
+%
+\begin{idescription}
+\item[(G)] Ayehsa holds $S$ true if and only if and $p$.
+\end{idescription}
+%
+Given the principle that beliefs can be assumed to be true by default, these generalisations are evidence for the conclusion that:
+%
+\begin{idescription}
+\item[(T)] $S$ is true if and only if $p$.
+\end{idescription}
+%
+Knowing this much about what Ayesha's sentences mean enables us to infer her beliefs given that we know which sentences she holds true.
+
+Davidson refined and extended this theory in various ways.
+By switching from mere correlations between facts and attitudes to causal relations between them, so that the evidence available to the radical interpreted includes facts about \emph{why} subjects hold particular attitudes towards certain sentences,
+it is possible to draw counterfactual-supporting conclusions about truth conditions.\footnote{
+Cf. \citealp[p.\ 6]{Davidson:1980xp}: `we are to derive meaning and belief from evidence concerning what causes someone to hold sentences true'.
+%\citealp{Davidson:1990du}`it is possible to infer that the speaker is caused by certain kinds of events to hold a sentence true'
+}
+%
+And by switching the evidential basis from attitudes of holding holding a sentence true to attitudes of preferring the truth of one sentence over another sentence,
+Davidson was able to show how the evidence he identified 
+grounded inferences concerning what a subject desires as well as what she believes
+\citep{Davidson:1980xp}.
+
+One striking feature of Davidson's account is that the evidential basis is already intentional.
+The mindreader (or `interpreter') starts
+
+\section{***}
+***Distinguish from `second person', interaction (Gallagher and Hutto?)
+
+***Mention that being involved in an interaction makes no huge difference to how mindreading works?
+
+To develop the idea that 
+interaction broadens the evidential basis for mindreading
+we do not need to throw away the leading, most critically examined theories due to Davidson and Dennett.
+An entirely fresh start is unnecessary.
+Of course we do not mean to assert, positively, that there are no good objections to these theories; maybe there are grounds for making a fresh start.
+Our point is just that considerations about the role of interaction in theories of mindreading do not by themselves require a fresh start.
+A good way to understand how interaction bears on epistemic and normative aspects of mindreading is to extend rather than to reject the leading theories.
+
+Since we shall discuss the role of interaction in theories of mindreading as an extension of leading theories (without any commitment here to the correctness of these theories; any theory of interpretation will have to accommodate considerations about interaction), it will be useful to start with a brief outline of those theories.
+
+\section{Repetition}
 
@@
 For instance, in saying that some principle serves as a heuristic in mindreading we are implicitly relying on a theory of interpretation to identify standards relative to which it is a heuristic.
-
-Note that the question a theory of interpretation answers is a question about justification.
-It is not about about \emph{how} anyone actually ascribes thoughts or identifies actions. 
-Nor is it (at least not immediately) a question about the mechanisms, processes or representations involved in mindreading.
-To stress this, Davidson sometimes formulates the issue by asking what someone \emph{could} know that would put them in a position to know another's thoughts and actions \citep[e.g.][p.\ 126]{Davidson:1973jx}.
 
@@
 
-The two most sustained and elaborate attempts to provide a theory of interpretation,
-Davidson's (\citeyear{Davidson:1984wh}) and Dennett's (\citeyear{Dennett:1987sf}),
-have ignored the possibility of interaction.
-The evidential basis they consider and the principles which they identify to link the evidence with ascriptions of thought and action are available to entirely passive observers.
-We shall argue that mindreaders who are capable of interacting with the targets of ascriptions are at an advantage.
-Their ascriptions can be justified by evidence and principles which would be unavailable if they were entirely passive observers.
+\section{The leading theory of interpretation}
 
-Being poised to interact with others enables one to know things about their minds which one might not otherwise be in a position to know; or so we aim to show in what follows.
-Because orthodox theories of interpretation have supposed that being able to interact makes no difference,
-they have treated the evidential basis for interpretation as more narrow than it truly is.
-This matters because broadening the evidential basis will not only enable us to give a more comprehensive theory of interpretation; 
-it will also provide tools for explaining the emergence, in evolution or in development, of mindreading.
-
-***Distinguish from `second person', interaction (Gallagher and Hutto?)
-
-To develop the idea that 
-interaction broadens the evidential basis for interpretation
-we do not need to throw away the leading, most critically examined theories of interpretation due to Davidson and Dennett.
-An entirely fresh start is unnecessary.
-Of course we do not mean to assert, positively, that there are no good objections to these theories; perhaps there are.
-Our point is just that considerations about the role of interaction in theories of interpretation do not give rise to any such objections.
-A good way to understand how interaction bears on epistemic and normative aspects of mindreading is to extend rather than to reject the leading theories.
-
-Since we shall discuss the role of interaction in theories of interpretation as an extension of leading theories (without any commitment here to the correctness of these theories; any theory of interpretation will have to accommodate considerations about interaction), it will be useful to start with a brief outline of those theories.
-
-\section{The leading theory of interpretation}
+*Not just language -- distinguish intentional action from goal-directed in teleological sense; question of interpretation just for how goals are identified.
+The basic principles proposed by both Davidson and Dennett (or versions of them) apply even to the non-linguistic case.
 

### 2011-11-25 joint action and knowing others minds/interacting mindreaders.2.tex
working

@@
 \section{Goal ascription}
-Identifying which goals others' actions are directed to enables us to predict and manipulate their actions, and to emulate their achievements.
+Purposive action is action directed to the realisation of one or more outcomes.
+Identifying which outcomes others' purposive actions are directed to enables one
+to predict and manipulate their actions, and 
+to emulate their achievements.
+It is also instrumental for mindreading:
+knowing which outcomes an action is directed to may constrain hypotheses about what an agent intends 
+as well as
+potentially providing information concerning what the agent knows, believes or desires.
+For example,
+if we know that the goal of an agent's action is to retrieve the key to her house,
+and if we also know where her key is,
+we may be able to infer that she does not know where her key is,
+or that she falsely believes that her key is in her purse.
+
+***Examples.
+
+These simple facts about goal ascription raise many questions.
+Some concern mechanism, how in fact one subject is able to discover facts about which outcomes another agent's actions are directed to.
+Another set of questions focuses on the evolution of goal ascription and the costs and benefits of being able to ascribe goals and of being a potential target of goal ascription.
+Our concern here is not directly with any of these questions.
+Instead we shall focus on a more narrowly epistemic question.
+What evidence could support hypotheses about the outcomes to which actions are directed?
+And how would the evidence support the hypotheses?%
+\footnote{
+These questions are versions of those Davidson constructs a theory of interpretation to answer.
+While what follows draws freely on Davidson's insights,
+our aims here are more modest than his.
+For we are concerned only with a  fraction of the problem of ascribing mental states and meanings;
+and, unlike Davidson, we are not concerned with larger claims about the nature of mind.
+See
+\citet{Davidson:1973jx,Davidson:1990du,lepore_donald_2005}.
+}
+
+Of special interest is evidence available independently of any knowledge of mind or language.
+We want to know how it is possible to identify goals even without knowing what an agent believes or desires and even without understanding their communicative actions.
+Accordingly we will adopt the perspective of a goal ascriber who knows nothing about the mental states of her target agent that would distinguish this agent from any other.
+We will also stipulate that there is initially no common ground, shared culture or conventions.
+And we will stipulate that the goal ascriber is initially unable to understand any communicative actions.
+
+There are two sorts of motivation for these restriction on the evidential basis.
+One is simply that developmental and comparative research indicates that goal ascription does appear to take place  in such circumstances.
+This makes it important to understand the grounds on which such ascriptions may be based (this may not tell us very much about how goal ascriptions are actually made, but it will be necessary to understand reliable success of mechanisms for goal ascription).
+Another source of motivation is the conjecture that goal ascription is a prerequisite for the more sophisticated mindreading activities which reveal mental states and meanings.
+The coherence of this conjecture depends on the possibility of knowing something about which outcomes an agent's actions are directed to independently of knowing what she believes or desires and independently of understanding her communicative actions.%
+\footnote{
+*Compare and contrast Davidson?
+He did think relational attitudes (holding true) are the foundation for interpretation.
+But he also thought that interpretation had to happen all at once.)
+
+\citet[pp.48--50]{Bennett:1976rg} suggests that a theory of goals ascription has to be developed together with a theory of (proto-)belief ascription: `An animal’s behaviour does not show what it registers unless we know what it seeks; but how can we learn what it seeks before we know what it registers?' [p.
+48]
+
+\citet[p.\ 17]{Dennett:1987sf} 'Here is how it works: first you decide to treat the object whose behavior is to be predicted as a rational agent; then you figure out what beliefs that agent ought to have, given its place in the world and its purpose.
+Then you figure out what desires it ought to have, on the same considerations, and finally your predict that this rational agent will act to further its goals in the light of its beliefs.
+A little practical reasoning from the chosen set of beliefs and desires will in many—but not in all—instances yield a decision about what the agent ought to do; that is what you predict the agent will do.'
+}
+
+So what evidence could support goal ascription by someone who knows nothing discriminating about her targets' mental states or communicative actions?
+The standard view, simplified, is this.%
+\footnote{
+*ref? Dennett?
+}
+Faced with an action,
+the would-be goal ascriber first asks which outcomes this action could be a means to realising.
+She then considers which of these outcomes are potentially beneficial for, or desirable to, the agent.
+Any such outcomes are identified as goals to which the action is directed.
+So the fact that an action is a means to realising some outcome which is potentially beneficial or desirable is evidence for the conclusion that this outcome is one to which the action is directed.
+%
+\begin{idescription}
+\item[(E$_1$)] Action $a$ is a means to realising outcome $G$.
+\item[(E$_2$)] The occurrence of outcome $G$ is potentially beneficial for, or desirable to, the agent of $a$.
+\end{idescription}
+%
+Are collectively evidence for concluding that:
+%
+\begin{idescription}
+\item[(C)] $G$ is a goal to which action $a$ is directed.
+\end{idescription}
+%
 

### 2011-12-01 joint action and knowing others minds/interacting mindreaders.2.tex
working

@@
 
+\section{Mindreading: the normative project}
+
+The question is what constitutes evidence.
+
+Mostly people have supposed that the evidence is available to pure observers --- being able to interact with the targets of mindreading confers no in principle advantage.
+
+In challenging this thesis we want to start with goal ascription, which is either a precursor to or a component of mindreading.
+Recent interest in non-propositional precursors to mindreading makes this appropriate.
+
 \section{Goal ascription}
 Purposive action is action directed to the realisation of one or more outcomes.
-Identifying which outcomes others' purposive actions are directed to enables one
-to predict and manipulate their actions, and 
-to emulate their achievements.
-It is also instrumental for mindreading:
+Goal ascription is the process of identifying which outcomes others' purposive actions are directed to.
+
+Goal ascription enables one
+to predict and manipulate others' actions, and 
+to learn from their failures and successes.
+%This applies even to goal ascribers with no further mindreading abilities: 
+% even those unable to identify beliefs, desires and other propositional attitudes have much to gain from identifying the outcomes to which other agents' actions are directed.
+For example, suppose that while you are searching for some peanuts 
+another agent attempts but fails to reach for a closed container.
+In some circumstances,
+if you know that the goal of the agent's action was to obtain the peanuts
+then you now have evidence as to where they might be.
+This is one illustration of how goal ascription could in principle enable us to learn from others' failures.
+
+Goal ascription is also instrumental for mindreading:
 knowing which outcomes an action is directed to may constrain hypotheses about what an agent intends 
@@
 For example,
-if we know that the goal of an agent's action is to retrieve the key to her house,
-and if we also know where her key is,
-we may be able to infer that she does not know where her key is,
-or that she falsely believes that her key is in her purse.
+if we know that the goal of an agent's action is to retrieve some peanuts,
+and if we also know where all the peanuts are,
+we may be able to infer that she does not know where the peanuts are,
+or that she falsely believes that some of the peanuts are over there.%
+\footnote{
+Of course this can also work the other way:
+belief- and goal-ascriptions are mutually constraining.
+}
 
-***Examples.
+Despite the close connection between goal ascription and mindreading,
+goal ascription does not necessarily involve representing representations.
+To see why not we first need to be careful about the term `goal'.
+***
+[do the detour here]
+
+The fact that goal ascription does not involve metarepresentation raises the possibility that goal ascription is possible independently of mindreading.
+Consider an agent who has no communicative skills and no metarepresentational abilities.
+Nothing prevents her from \emph{representing} goals, of course.
+But could she actually ascribe goals?
+Is there any evidence which could be available to her and which would support goal ascriptions?
+Some philosophers have argued\footnote{*Bennett} or implied\footnote{*Davidson} that there could not be.
+On their view, the goal ascription and mindreading are interdependent in this sense: 
+there is no evidence for hypotheses narrowly about goals, only evidence for more complex hypotheses concerning both goals and mental states (such as beliefs).
+So, on this view, those without metarepresentational abilities are not in a position to know the goals of other agents' actions.
+
+\section{OLD}
 
@@
 These questions are versions of those Davidson constructs a theory of interpretation to answer.
-While what follows draws freely on Davidson's insights,
+While what follows draws  on Davidson's insights,
 our aims here are more modest than his.
@@
 There are two sorts of motivation for these restriction on the evidential basis.
-One is simply that developmental and comparative research indicates that goal ascription does appear to take place  in such circumstances.
-This makes it important to understand the grounds on which such ascriptions may be based (this may not tell us very much about how goal ascriptions are actually made, but it will be necessary to understand reliable success of mechanisms for goal ascription).
+One is simply that developmental and comparative research indicates that goal ascription does appear to take place  in such circumstances.%
+\footnote{
+*refs
+}
+This makes it important to understand the evidence on which such ascriptions could be based.
+(Of course 
+identifying evidence that could support such ascriptions 
+would not all by itself enable us to explain how goal ascriptions are actually made, 
+but identifying evidence is necessary if we are ever to explain the reliable success of mechanisms for goal ascription.)
 Another source of motivation is the conjecture that goal ascription is a prerequisite for the more sophisticated mindreading activities which reveal mental states and meanings.
@@
 So what evidence could support goal ascription by someone who knows nothing discriminating about her targets' mental states or communicative actions?
-The standard view, simplified, is this.%
+Ordinary third-person goal ascription, simplified and idealized, works like this.%
 \footnote{
@@
 So the fact that an action is a means to realising some outcome which is potentially beneficial or desirable is evidence for the conclusion that this outcome is one to which the action is directed.
+Schematically, the proposal is that:
 %
 \begin{idescription}
-\item[(E$_1$)] Action $a$ is a means to realising outcome $G$.
-\item[(E$_2$)] The occurrence of outcome $G$ is potentially beneficial for, or desirable to, the agent of $a$.
+\item[(E$_1$)] Action $a$ is a means of realising outcome $G$.
 \end{idescription}
 %
-Are collectively evidence for concluding that:
+and:
+%
+\begin{idescription}
+\item[(E$_2$)] The occurrence of outcome $G$ is potentially beneficial for, or desirable to, the agent of $a$.
+(And there is no other outcome, $G'$, which action $a$ is a means of realising and which would be more beneficial for, or more desirable to, the agent of $a$.)
+\end{idescription}
+%
+jointly constitute evidence for the conclusion that:
 %
@@
 \end{idescription}
+
+This proposal might be extended in various ways.
+For instance, Southgate, Johnson and Csibra offer a `principle of efficiency' according to which:
 %
+\begin{quote}
+`goal attribution requires that agents expend the least possible amount of energy within their motor constraints to achieve a certain end.'
+\citep[p.\ *]{Southgate:2008el}
+\end{quote}
+%
+If this is a correct principle of goal attribution, we could extend the proposal above to incorporate it:
+%
+\begin{idescription}
+\item[(E$_3$)] No alternative action, $a'$, is a means to realising outcome $G$ and would involve expending less energy than $a$.
+\end{idescription}
+%
+Now the proposal is that (E$_1$) to (E$_3$) are jointly evidence for (C).
+
+In at least some cases goal attribution is likely to be more complicated than this proposal allows.
+To illustrate, note that some agents may weigh the efficiency of alternative actions against their possible side effects and how reliable they would be as a means to realising an outcome.
+Where this is true, 
+identifying the evidential basis for goal ascription may require  a similar weighing of these factors in inferring backwards from actions to their goals.%
+\footnote{
+This is loosely related to what Csibra and Gergely call `the principle of rational action'.
+As they formulate the principle,
+`an action can be explained by a goal state if, and only if, it is seen as the most justifiable action towards that goal state that is available within the constraints of reality'
+(\citealp[p.\ *]{Csibra:1998cx}; cf.\ \citealp{Csibra:2003jv}).
+}
+Specifying exactly what should be weighed and how is beyond the scope of this paper,
+(and may also be something which varies between species of agent).
+We can mark the gap with an alternative to (E$_3$) which uses an unspecified notion of `better' as a placeholder:
+%
+\begin{idescription}
+\item[(E$_{3'}$)] No alternative action, $a'$, is a better means to realising outcome $G$.
+\end{idescription}
+%
+This, then, is the standard approach to answering our question about goal attribution: (E$_1$), (E$_2$) and (E$_{3'}$) jointly constitute evidence for (C) given that these approximate conditions under which it would be rational to perform $a$ in order to realise $G$ and given that agents approximate to performing $a$ in order to realise $G$ rather than any other outcome under these conditions.
 

### 2011-12-14 joint action and knowing others minds/interacting mindreaders.2.tex
still getting started

@@
 Goal ascription is the process of identifying which outcomes others' purposive actions are directed to.
+Because goal ascription has received little attention,
+we  first review some of the benefits that being able to identify others' goals brings.
 
-Goal ascription enables one
-to predict and manipulate others' actions, and 
-to learn from their failures and successes.
-%This applies even to goal ascribers with no further mindreading abilities: 
-% even those unable to identify beliefs, desires and other propositional attitudes have much to gain from identifying the outcomes to which other agents' actions are directed.
+It is a familiar idea that goal ascription enables one to learn from others' successes.
+For example,
+if you know or can guess that another agent's actions are directed to opening a nut,
+you may then be in a position to infer that the unfamiliar pattern of actions she is performing constitute a means to open nuts.\footnote{
+***
+\citep{Horner:2005pj}
+}
+Or (in a different case) knowing the goal of another agent's actions may enable you to discover a new use for a tool.\footnote{
+There is currently much 
+}
+A slightly less familiar idea is that goal ascription enables one to learn from others' failures as well as their successes.
 For example, suppose that while you are searching for some peanuts 
@@
 if you know that the goal of the agent's action was to obtain the peanuts
-then you now have evidence as to where they might be.
+then you now have evidence as to where they might be.\footnote{
+\citet{hare_chimpanzees_2004} exploit this principle in testing chimpanzees' abilities to ascribe goals.
+}
 This is one illustration of how goal ascription could in principle enable us to learn from others' failures.
+
+Goal ascription enables one
+to predict and manipulate others' actions.
+***if goal involved manipulating an object, you can predict where agent will go at some point; you can also manipulate the agents' action by revealling the location of the object.
+Eg.
+Liszowski pointing
+%This applies even to goal ascribers with no further mindreading abilities: 
+% even those unable to identify beliefs, desires and other propositional attitudes have much to gain from identifying the outcomes to which other agents' actions are directed.
 
@@
 \footnote{
-Of course this can also work the other way:
-belief- and goal-ascriptions are mutually constraining.
+\citet{Wimmer:1998kx} exploit this possibility in testing children's abilities to ascribe false beliefs.
 }
+(Of course this can also work the other way:
+information about an agent's beliefs or other mental states may support conclusions about her goals.
+Belief- and goal-ascriptions are mutually constraining.)
 
@@
 So, on this view, those without metarepresentational abilities are not in a position to know the goals of other agents' actions.
+
+\section{The problem of opaque means}
+
+While we lack a detailed theory of the evidential basis of goal ascription,
+it is certain that the evidence for goal ascription sometimes includes considerations about which ends actions are means to.
+Suppose an observer faces an action but cannot identify which end (or ends) it could be a means to.
+This may limit her ability to identify the action's goal (or goals) by depriving her of evidence.
+To illustrate, contrast two cases of tool use.
+In one case, someone uses a reamer to  juice  a lime; in the other, someone else scores shag with a lame to prevent a loaf from cracking.
+Without testimony, 
+an observer familiar with reamers but not lames 
+is less likely to be able to identify the goals of this second action than the first.
+As this illustrates, ignorance about to which ends actions are means can be an obstacle to goal ascription.%
+\footnote{
+This is not to say that no observer could ever identify the goal of any action she fails to recognise as a means to achieving that goal.
+Surely opaque means are not in every case an insurmountable obstacle to goal ascription.
+Our point is that opaque means are sometimes an obstacle to goal ascription because they deprive observers of evidence.
+}
+Call this the problem of opaque means.
+
+Some of the most plausibly unique aspects of human cognition depend on our abilities to recognise the goals of novel behaviours involving tools, and of communicative gestures.
+The problem of opaque means is likely to arise in both cases 
+if goal ascription is based on entirely observation 
+(so that the possibility of interaction is ignored).
+We have just seen an illustration of how the problem of opaque means arises where tools are used to unfamiliar ends.
+Relatedly, it is also likely to arise where actions involve multiple steps that do not form a familiar sequence, can occur in various orders and can be interspersed among other activities;
+as in preparing spirit from grain, for example.
+%\footnote{
+%*cut because 'plausibly unique aspects of human cognition'*
+%one example is food preparation in some groups of mountain gorillas
+%\citep[p.\ 531]{Byrne:2003wx}.
+%*This is slightly tricky because Byrne wants a pure behaviour reading account with no understanding of intention.  
+%Stress that this is consistent with supposing that goals are ascribed.
+%}
+ The problem of opaque means also affects communicative actions 
+ because these are generally  directed to goals which the actions are means to realising only because others recognise them as means to realising those goals (that is, they involve a Gricean circle). 
+To illustrate, consider an experiment by
+***ref whose two main conditions are depicted in figure *fig.
+The pictures in the figure stand for what participants, who were chimpanzees, saw.
+The question was whether participants would be able to work out which of two containers contained a reward.
+In the condition depicted in the left panel, participants saw a chimpanzee trying but failing to reach for the reward. 
+Participants had no problem getting the reward in this case, suggesting that they understood the goal of the failed reach.
+In the condition depicted in the right panel, a human pointed to the reward location.
+Participants did not reliably  get the reward in this case, suggesting that they failed to understand the goal of the pointing action.
+This may be because of the problem of opaque means: 
+participants could identify to what end a failed reach might be a means, but not to what end a communicative gesture might be a means.
+
+***FIGURE***
+
+In what follows we shall suggest that abilities to engage in joint action with others provides a route to knowledge of the goals of other agents' actions  that avoids the problem of opaque means.
+This is one way in which moving away from a purely observational model of interpretation yields a richer evidential base for ascriptions.
+But first we consider a second, related problem.
+
+\section{The problem of false belief}
+We have just seen that 
+failure to identify to which ends actions are means can impair goal ascription; this was the problem of opaque means.
+Another problem for purely observational goal ascription arises from the interdependence of beliefs and goals.
+To illustrate, 
+imagine sitting at a table on which two closed opaque boxes each contain an object; one contains an owl, the other a cat.
+If the goal of your action is to retrieve the owl, 
+and you believe that the owl is in the north-most box,
+then 
+(unless things are going very badly) 
+you will reach into the north-most box.
+But of course if you had believed instead that the owl was in the other box,
+then, in acting on the same goal, you would have reached into the other box.
+Now consider someone observing your actions.
+If she is has good reason to believe, falsely, that you know the owl is in the south-most box,
+then she may be justified in supposing, incorrectly, that when you reach into the north-most box the the goal of your action is to retrieve the cat.
+As this illustrates,
+differences in belief between observers and protagonists can be an obstacle to goal ascription.
+Call this the problem of false belief.
+
+There is disagreement over whether it is possible to knowledgably identify the goals of an agent's actions without also ascribing some beliefs to that agent.  
+\citet[pp.48--50]{Bennett:1976rg}
+%\citet[pp.48--50]{Bennett:1976rg} suggests that a theory of goals ascription has to be developed together with a theory of (proto-)belief ascription: `An animal’s behaviour does not show what it registers unless we know what it seeks; but how can we learn what it seeks before we know what it registers?' [p.
+48]
+and 
+\citet{Davidson:1974gh}
+both appear to hold that this is impossible,
+that identifying goals cannot be done independently of ascribing beliefs.
+By contrast,
+\citet{Gergely:1995sq},
+*Meltzoff?,
+\citet{Baillargeon:gx}
+and 
+\citet{Woodward:1998dm}
+(among many others)
+appear to assume the opposite, 
+that it is possible to identify goals without even being able to ascribe beliefs.
+For what it is worth, we tentatively favour this latter position.%
+\footnote{
+It is striking that, as far as we can tell, neither Bennett nor Davidson offers an argument for this claim.
+They do note that beliefs and goals make an interdependent contribution to observed action.
+But this by itself does not show that goal ascription cannot in some cases involve justifiably ignoring the possibility of differences in belief between interpreters and their targets.
+For instance,
+suppose that two people are sitting opposite each other at a low table
+ which is 
+sparsely populated with objects.
+The objects are all out in the open; manifestly, both can clearly see them.
+If one reaches to grasp one of these objects (the duck, say), 
+must the other ascribe beliefs in order to knowledgeably identify the goal of her action?
+On the face of it, she need not.  
+Even if she had no ability to ascribe beliefs, she might nevertheless be in a position to acquire knowledge of the goal of the other's action.
+}
+However, this debate is not directly relevant to our concerns here.
+Both sides can agree that 
+differences in belief between observer and protagonist
+are sometimes an obstacle to goal ascription.
+This is all that the problem of false belief requires.
+
+In what follows we shall argue that abilities to engage in joint action provide an interpreter with a route to knowledge of the goals of other agents' actions which does not depend on her abilities to ascribe beliefs.
+This is not because abilities to engage in joint action provide a way to avoid the problem of false beliefs altogether.
+Rather, as we shall see, abilities to engage in joint action provide a way to shift the burden of resolving the problem of false belief from an interpreter to her target.
 
@@
 
-\citet[pp.48--50]{Bennett:1976rg} suggests that a theory of goals ascription has to be developed together with a theory of (proto-)belief ascription: `An animal’s behaviour does not show what it registers unless we know what it seeks; but how can we learn what it seeks before we know what it registers?' [p.
-48]
-
-\citet[p.\ 17]{Dennett:1987sf} 'Here is how it works: first you decide to treat the object whose behavior is to be predicted as a rational agent; then you figure out what beliefs that agent ought to have, given its place in the world and its purpose.
+\citet[p.\ 17]{Dennett:1987sf} `Here is how it works: first you decide to treat the object whose behavior is to be predicted as a rational agent; then you figure out what beliefs that agent ought to have, given its place in the world and its purpose.
 Then you figure out what desires it ought to have, on the same considerations, and finally your predict that this rational agent will act to further its goals in the light of its beliefs.

### 2011-12-14 joint action and knowing others minds/interacting mindreaders.2.tex
working

@@
 
-\input{$HOME/Documents/submissions/preamble_steve_paper}
+\input{$HOME/Documents/submissions/preamble_steve_paper2}
 
@@
 it is certain that the evidence for goal ascription sometimes includes considerations about which ends actions are means to.
-Suppose an observer faces an action but cannot identify which end (or ends) it could be a means to.
-This may limit her ability to identify the action's goal (or goals) by depriving her of evidence.
+Suppose an observer faces an action but cannot identify ends to which it could be a means.
+This may prevent her from recognizing the action's goal%
+\footnote{
+It is possible that some actions have more than one goal.
+To reduce parenthetical qualifications we shall write as if actions had only one goal.  
+All of our key claims and arguments are consistent with the possibility of actions with more than one goal.
+} 
+ by depriving her of evidence.
 To illustrate, contrast two cases of tool use.
 In one case, someone uses a reamer to  juice  a lime; in the other, someone else scores shag with a lame to prevent a loaf from cracking.
-Without testimony, 
+Without communication, repetition or convention,
 an observer familiar with reamers but not lames 
-is less likely to be able to identify the goals of this second action than the first.
+may be able to identify the goal of the first action only.
 As this illustrates, ignorance about to which ends actions are means can be an obstacle to goal ascription.%
@@
 Surely opaque means are not in every case an insurmountable obstacle to goal ascription.
-Our point is that opaque means are sometimes an obstacle to goal ascription because they deprive observers of evidence.
+Our point is just that opaque means \emph{sometimes}  deprive observers of evidence and so prevent goal ascription.
 }
@@
 Some of the most plausibly unique aspects of human cognition depend on our abilities to recognise the goals of novel behaviours involving tools, and of communicative gestures.
-The problem of opaque means is likely to arise in both cases 
-if goal ascription is based on entirely observation 
-(so that the possibility of interaction is ignored).
+If goal ascription is based on entirely observation 
+(so that the possibility of interaction is ignored),
+the problem of opaque means is likely to arise in both cases.
 We have just seen an illustration of how the problem of opaque means arises where tools are used to unfamiliar ends.
@@
 %}
+
 The problem of opaque means also affects communicative actions 
- because these are generally  directed to goals which the actions are means to realising only because others recognise them as means to realising those goals (that is, they involve a Gricean circle). 
+ because these characteristically have  goals which the actions are means to realising only because others recognise them as means to realising those goals (a Gricean circle). 
 To illustrate, consider an experiment by
@@
 The pictures in the figure stand for what participants, who were chimpanzees, saw.
-The question was whether participants would be able to work out which of two containers contained a reward.
-In the condition depicted in the left panel, participants saw a chimpanzee trying but failing to reach for the reward. 
+The question was whether participants would be able to work out which of two containers concealed a reward.
+In the condition depicted in the left panel, participants saw a chimpanzee trying but failing to reach for the correct container. 
 Participants had no problem getting the reward in this case, suggesting that they understood the goal of the failed reach.
-In the condition depicted in the right panel, a human pointed to the reward location.
+In the condition depicted in the right panel, a human pointed at the correct container.
 Participants did not reliably  get the reward in this case, suggesting that they failed to understand the goal of the pointing action.
-This may be because of the problem of opaque means: 
-participants could identify to what end a failed reach might be a means, but not to what end a communicative gesture might be a means.
+This may be because of the problem of opaque means.
+One theoretically possible explanation of these findings is that the participants could identify to which end a failed reach might be a means, but not to which end a communicative gesture might be a means.%
+\footnote{
+*quote Moll \& Tomasello
+}
+Whatever the truth about the chimpanzees' performance,
+this possibility illustrates how the problem of opaque means can be an obstacle to exploiting communicative gestures.
 
@@
 
-In what follows we shall suggest that abilities to engage in joint action with others provides a route to knowledge of the goals of other agents' actions  that avoids the problem of opaque means.
+This, then, is the problem of opaque means:
+failures to identify to which ends actions are means can impair goal ascription.
+The problem is potentially a problem for interpreters given the standard, purely observational models of interpretation.
+It is not our intention to suggest that the problem of opaque means is a problem \emph{for models of interpretation}.
+Rather, it is a potential problem \emph{for interpreters}.
+The existence of this problem shows, we shall argue, that models of interpretation based on pure observation are less powerful than models taking into account the possibility of interaction.
+They are less powerful in this sense: 
+some routes to knowledge of the goals of actions are available only where a model of interpretation takes the possibility of interaction into account.
+In what follows we shall suggest that abilities to engage in joint action with others provides a route to knowledge of the goals of other agents' actions  which avoids the problem of opaque means.
 This is one way in which moving away from a purely observational model of interpretation yields a richer evidential base for ascriptions.
-But first we consider a second, related problem.
 
-\section{The problem of false belief}
+\section{Joint action}
+\label{joint_action}
+Our overall aim is to defend this claim:
+there are routes to knowledge of the goals of others' actions
+which are closed to
+interpreters who merely observe their targets
+but open to 
+interpreters capable of interacting with their targets.
+As a preliminary to defending this claim,
+we need to specify the types of interaction which are relevant.
+We shall focus on joint action.
+
+***Much disagreement about what it is.
+Necessary condition is distributive goal.
+If people want to disagree, this is only terminological.
+
+\section{Your-goal-is-my-goal: a route to knowledge}
+If an interpreter is able to interact with her targets,
+if she is not limited to merely observing them,
+how might this provide her with a route to knowledge of the goals of their actions?
+The intuitive idea we started with hinges on joint action, a particular form of interaction in which the actions of two or more agents are directed to a single goal (see section \vref{joint_action}).
+Our intuitive idea was this:
+if an interpreter is engaged in joint action with her target, 
+it's easy for the interpreter to know what the goal of her target's actions is because this goal is the goal of her own actions.
+So if she knows the goal of her own actions and she knows that she is engaged in joint action with her target,
+then she already knows what the goal of her target's actions are.
+%Roughly speaking, the interpreter can reason about her target thus: your goal is my goal.
+
+Of course this intuitive idea won't work as it stands.  
+For the inference it captures relies on the premise that the interpreter and her target are engaged in joint action.
+But for the interpreter to know this premise---for her to know that they are engaged in joint action---it seems she must   already know which goal her target's actions are directed to.
+Worse,
+on many accounts of joint action the truth of the premise  entails that the interpreter knows the goal of her target's actions.%
+\footnote{
+*ref}
+Apparently, then, engaging in joint action presupposes, and therefore cannot be a source of, knowledge of others' goals.
+
+Fortunately there is a way around this.  
+For there are various cues which signal that one agent is prepared to engage in joint action with another.  
+Seeing a new parent struggling to get his twin pram on to the bus, you grab the front wheels and make eye contact, raising your eyebrows and smiling.  
+In this way you signal both that you are disposed to help 
+and 
+that are about to engage in joint action with the stranger.  
+Since the stranger is fully committed to getting his pram onto the bus,
+he knows what the goals of his own actions will be.
+This makes enables him to infer the goal of your immanent actions: 
+your goal is his goal, to get the pram onto the bus.
+
+Our suggestion, then, is that the following inference characterises a route to knowledge of others’ goals:
+%
+\begin{enumerate}
+\item We are about to engage in some joint action\footnote{
+*What notion of joint action is needed here?  Any will do as long as it involves distributive goals.
+}
+or other 
+%(for example, because you have made eye contact with me while I was in the middle of attempting to do something).
+
+\item I am not about to change which goal my actions will be directed to.
+
+\end{enumerate}
+%
+Therefore:
+%
+\begin{enumerate}[resume]
+%
+\item The goal of your actions will be my goal.
+\end{enumerate}
+%
+Call this the \emph{your-goal-is-my-goal inference}.  
+To say that this inference characterises a route to knowledge implies two things.  
+First, in some cases it is possible to know the premises, 1–2, without already knowing the conclusion, 3.  
+Second, in some cases knowing the two premises puts one in a position to know the conclusion.  
+We shall consider these points in turn
+
+Is it ever possible to know the premises without first knowing the conclusion?  
+Consider the first premise.  
+Sometimes 
+in the right contexts
+you can see in others’ facial expressions, engaging gestures or synchronized bodily movements that they are about to engage in joint action with you.  
+Exploiting these indicators does not typically depend on knowing the particular contents of any of their beliefs, desires or goals.  
+Expressions, gestures and movements can naturally indicate imminent jointness in much the way that can also naturally indicate emotions.\footnote{
+Ideas along these lines are suggested by the discussion of \emph{emergent coordination} in \citet{Knoblich:2010fk}.
+}  
+Of course these indicators provide no guarantee that others are about to helpfully engage in joint action with you.  
+But they are sufficiently reliable to ground knowledge in some cases.  
+So knowing the first two premises of the above inference does not require already knowing which particular goals others have.
+
+On this point not everything needs to rest on indicators of imminent jointness.
+It is sometimes possible to know that others are about to engage in joint action with you even without relying on such indicators.  
+Thanks to widespread dispositions to act jointly, in some situations it is reasonable to take for granted that others are disposed to be helpful and will act jointly with you.  
+For example, this is often so for children surrounded by family or familiar adults who are having difficulty with a task.  
+And in some cultures parents with struggling small children on public transport (say) can reasonably take for granted that those around them will act jointly with them when the need is clear.
+
+Does knowing the premises of the your-goal-is-my-goal inference sometimes put one in a position to know the conclusion?  
+If you know that others are about to engage in joint action with you, it is sometimes reasonable to infer that the outcome to which the joint action will be directed is the very outcome you are currently concerned with.
+Given the account of joint action in Section 6 above, this means that the others each have an individual goal whose fulfilment requires this outcome to occur.  
+So knowing the outcome to which a joint action will be directed puts one in a position to know something about which goals the others have.  
+Of course the inference this rests on is not deductive and will only be correct when certain background conditions are met.  
+These background conditions include the others having largely true beliefs concerning which outcome you are concerned with; I return to this point below.
+
+\section{The problem of false belief [move to later]}
 We have just seen that 
 failure to identify to which ends actions are means can impair goal ascription; this was the problem of opaque means.
-Another problem for purely observational goal ascription arises from the interdependence of beliefs and goals.
+Another problem 
+affecting goal ascription given purely observational models of interpretation
+ arises from the interdependence of beliefs and goals.
 To illustrate, 

### 2011-12-16 joint action and knowing others minds/interacting mindreaders.2.tex
slow progress

@@
 
-\documentclass[12pt,a4paper]{extarticle}
+\documentclass[14pt,a4paper]{extarticle}
 % extarticle is like article but can handle 8pt, 9pt, 10pt, 11pt, 12pt, 14pt, 17pt, and 20pt text
@@
 \input{$HOME/Documents/submissions/preamble_steve_paper2}
+
+%\setromanfont[Mapping=tex-text]{Sabon LT Std}
 
@@
 \section{The problem of opaque means}
+\label{sec:opaque_means}
 
@@
  because these characteristically have  goals which the actions are means to realising only because others recognise them as means to realising those goals (a Gricean circle). 
-To illustrate, consider an experiment by
-***ref whose two main conditions are depicted in figure *fig.
+To illustrate, consider an experiment from 
+\citet[][experiment 3]{hare_chimpanzees_2004}
+ whose two main conditions are depicted in figure \vref{fig:reach_point}.
 The pictures in the figure stand for what participants, who were chimpanzees, saw.
@@
 In the condition depicted in the right panel, a human pointed at the correct container.
-Participants did not reliably  get the reward in this case, suggesting that they failed to understand the goal of the pointing action.
+Participants did not reliably  get the reward in this case, suggesting that they failed to understand the goal of the pointing action.%
+\footnote{
+The contrast between the two conditions is not due merely to the fact that one involves a human and the other a chimpanzee.
+Participants were also successful when the failed reach was executed by a human rather than another chimpanzee \citep[][experiment 1]{hare_chimpanzees_2004}. 
+}
 This may be because of the problem of opaque means.
@@
 \footnote{
-*quote Moll \& Tomasello
+Hare and Tomasello 
+consider several explanations for their findings including `the hypothesis that chimpanzees do not understand the communicative intent of a cooperative-communicative experimenter' (\citeyear[p.\ 580]{hare_chimpanzees_2004}).
+\citet[pp.\ 5--7]{Moll:2007gu} argue for a hypothesis along these lines by appeal to a range of related findings.
 }
@@
 
-***FIGURE***
+\begin{figure}
+\begin{center}
+\includegraphics[width=12cm]{figure_hare_toma_2004_e3.png}
+\caption{
+\label{fig:reach_point}
+	A failed reach (left) and a helpful point (right).
+	Reproduced from \citet[p.\ 557, figure 4]{hare_chimpanzees_2004}.
+}
+\end{center}
+\end{figure}
 
@@
 Worse,
-on many accounts of joint action the truth of the premise  entails that the interpreter knows the goal of her target's actions.%
+on many accounts of joint action the mere truth of the premise  would require the interpreter to know the goal of her target's actions.%
 \footnote{
@@
 For there are various cues which signal that one agent is prepared to engage in joint action with another.  
-Seeing a new parent struggling to get his twin pram on to the bus, you grab the front wheels and make eye contact, raising your eyebrows and smiling.  
-In this way you signal both that you are disposed to help 
-and 
-that are about to engage in joint action with the stranger.  
+\label{twin_pram}
+Seeing a stranger with the haggard look of a new parent struggling to get his twin pram on to the bus, you grab the front wheels and make eye contact, raising your eyebrows and smiling.
+(The noise of the street rules out talking, 
+and you can barely speak the local language anyway.)   
+In this way you signal that you are about to act jointly with the stranger, to lift the pram with him.   
 Since the stranger is fully committed to getting his pram onto the bus,
 he knows what the goals of his own actions will be.
-This makes enables him to infer the goal of your immanent actions: 
+This enables him to infer the goal of your imminent actions: 
 your goal is his goal, to get the pram onto the bus.
@@
 \begin{enumerate}
-\item We are about to engage in some joint action\footnote{
+\item You are willing to engage in some joint action\footnote{
 *What notion of joint action is needed here?  Any will do as long as it involves distributive goals.
 }
-or other 
+or other with me
 %(for example, because you have made eye contact with me while I was in the middle of attempting to do something).
@@
 %
-\item The goal of your actions will be my goal.
+\item A goal of your actions will be my goal, the goal I now envisage that my actions will be directed to.
 \end{enumerate}
@@
 First, in some cases it is possible to know the premises, 1–2, without already knowing the conclusion, 3.  
-Second, in some cases knowing the two premises puts one in a position to know the conclusion.  
+Second, in some of those cases knowing the premises would put one in a position to know the conclusion.  
 We shall consider these points in turn
@@
 in the right contexts
-you can see in others’ facial expressions, engaging gestures or synchronized bodily movements that they are about to engage in joint action with you.  
+an individual can recognize in another's facial expressions, engaging gestures or synchronized bodily movements that she are about to engage in joint action with him.
 Exploiting these indicators does not typically depend on knowing the particular contents of any of their beliefs, desires or goals.  
-Expressions, gestures and movements can naturally indicate imminent jointness in much the way that can also naturally indicate emotions.\footnote{
+Expressions, gestures and movements can naturally indicate imminent jointness in much the way thay can also naturally indicate emotions.\footnote{
 Ideas along these lines are suggested by the discussion of \emph{emergent coordination} in \citet{Knoblich:2010fk}.
 }  
-Of course these indicators provide no guarantee that others are about to helpfully engage in joint action with you.  
+Of course these indicators provide no guarantee that others are genuinely willing to engage in joint action.  
 But they are sufficiently reliable to ground knowledge in some cases.  
-So knowing the first two premises of the above inference does not require already knowing which particular goals others have.
+The existence of such indicators shows that knowing the first  premise of the above inference does not require already knowing which particular goals the other has.
 
-On this point not everything needs to rest on indicators of imminent jointness.
-It is sometimes possible to know that others are about to engage in joint action with you even without relying on such indicators.  
-Thanks to widespread dispositions to act jointly, in some situations it is reasonable to take for granted that others are disposed to be helpful and will act jointly with you.  
-For example, this is often so for children surrounded by family or familiar adults who are having difficulty with a task.  
-And in some cultures parents with struggling small children on public transport (say) can reasonably take for granted that those around them will act jointly with them when the need is clear.
+Not everything needs to rest on indicators, however.
+It is sometimes possible to know that others are about to willing in joint action with you even without relying on such indicators.  
+Thanks to widespread dispositions to act jointly, in some situations it is reasonable to take for granted that others are willing to act jointly.  
+For example, this is often so for children surrounded by family or familiar adults who are struggling with a coat.  
+And in at least some subcultures people using public transport can reasonably take for granted that, within limits, those around them will act jointly with them when the need is clear.
+Of course dispositions to engage in joint action may vary between cultural groups and situations.
+This may be fatal for the peripatetic cosmopolitan,
+but for others what matters is not whether the dispositions are universal but only that they are sufficiently widespread to be predictable.
 
-Does knowing the premises of the your-goal-is-my-goal inference sometimes put one in a position to know the conclusion?  
-If you know that others are about to engage in joint action with you, it is sometimes reasonable to infer that the outcome to which the joint action will be directed is the very outcome you are currently concerned with.
-Given the account of joint action in Section 6 above, this means that the others each have an individual goal whose fulfilment requires this outcome to occur.  
-So knowing the outcome to which a joint action will be directed puts one in a position to know something about which goals the others have.  
-Of course the inference this rests on is not deductive and will only be correct when certain background conditions are met.  
-These background conditions include the others having largely true beliefs concerning which outcome you are concerned with; I return to this point below.
+Turning to the second point,
+does knowing the premises of the your-goal-is-my-goal inference sometimes put one in a position to know the conclusion?  
+Of course the inference is not deductive and will only work when certain background conditions are met.  
+These background conditions include the other having largely true beliefs concerning which goal your actions are or will be directed to.
+After all, the other agent may be willing to act jointly with you while being entirely mistaken about the goals to which your actions will be directed.
+Where this happens, the premises of the inference might be true but the conclusion false.
+(We return to this point in section \vref{sec:false_belief}.)  
+But there are situations in which it is reasonable  to ignore  this possibility, as for instance when stereotypes, conventions or simplicity should and do make the goal of your actions obvious to the other agent.
+Given the right conditions,
+that others are willing to engage in joint action with you 
+is sometimes sufficient reason to hold that
+you will end up acting jointly with them
+ even where the goal to which your actions will be directed is already fixed.
 
-\section{The problem of false belief [move to later]}
+In short, then, the two requirements for the your-goal-is-my-goal inference to characterise a route to knowledge are met.  
+In some cases it is possible to know the premises without already knowing the conclusion thanks to natural expressions of willingness to engage in joint action.  
+And knowing the premises sometimes puts one in a position to know the conclusion thanks to the fact that, when things are going well, another's willingness to engage in joint action with you can be based on an accurate assessment of the goals of your actions.
+
+The your-goal-is-my-goal route to knowledge is characterised by an inference.  
+However, exploiting this route to knowledge may not require actually making the inference or knowing the premises.  
+Depending on what knowing requires,
+it may be sufficient to believe the conclusion because one has reliably detected a situation in which the premises of the inference are true;
+it may not be necessary to think of this situation as a situation where the premises are true,
+nor even to be able to think of it in this way.
+
+Our aim in identifying the your-goal-is-my-goal is not to 
+defend a detailed hypothesis about the mechanisms and processes involved in mindreading.
+Instead our present concerns are limited to a normative question about the evidential basis of mindreading.
+The your-goal-is-my-goal inference matters
+not because it describes how interpreters actually assign goals
+but because 
+its existence reveals that interacting mindreaders can have access to evidence unavailable to those relying entirely on observation.
+
+\section{Avoiding the problem of opaque means}
+
+The problem of opaque means was this:
+ failures to identify to which ends actions are means can impair goal ascription (see section \vref{sec:opaque_means}).
+%As we saw, this problem is particularly likely to arise in cases involving novel tools and communicative gestures.
+Showing how your-goal-is-my-goal makes  avoiding this problem possible is a way of demonstrating the potential value  to mindreaders of interaction.
+
+In our earlier example 
+a novice parent is struggling to lift his heavy twin pram onto a bus when a stranger joins in and they lift the pram together (\vpageref*{twin_pram}).
+Suppose the stranger starts tipping the pram in a way that the inexperienced parent fails to recognise as a means, indeed the only means, of getting it onto the bus.
+Outside the context of joint action this might give the parent sufficient evidence to reject the idea that the goal of the stranger's actions is to get the pram onto this bus.
+Or suppose the stranger starts by pulling the pram away from the bus in order to better position it for entry,
+and the novice parent does not realise that this is necessary.
+Outside the context of joint action, 
+his evidence might on balance support the conclusion that the goal of the stranger's actions is to take the pram off the bus.
+(Perhaps the stranger is impatient to get onto the bus herself.)
+But in the context of joint action,
+your-goal-is-my goal gives the parent  additional evidence for supposing that, 
+even though the stranger's actions do not seem to him to be a means to getting the pram onto the bus, 
+this really is the goal of her actions.
+Of course this additional evidence will not necessarily trump   other evidence.
+But it does provide a possible way around the problem of opaque means by providing evidence for goal ascriptions that is independent of a mindreader's understanding of which ends actions are means to.
+
+We saw earlier that the problem of opaque means  may impair goal ascription where actions involve novel uses for tools.
+How could your-goal-is-my-goal mitigate the problem in such cases?
+Imagine we are interacting with a young child,
+Katarina,
+and want her to understand how a new tool is used.  
+It is difficult to convey this to her directly.  
+So we first get her interested in achieving an outcome that would require the new tool,
+knowing that she will perform actions directed to achieving this outcome.  
+We then signal to Katarina that we are willing to act jointly with her.  
+Now she is in a position to know what the goal of our action will be when we deploy the tool.  
+She is able to identify this goal 
+despite being unable to recognize it as an end to which our tool-using action is a means.
+She is able to identify this goal 
+because she knows that this is her goal
+and that we were willing to engage in joint action with her.
+This is one illustration of how 
+interacting mindreaders have at their disposal ways of
+identifying the goals of actions involving novel uses of tools
+which are unavailable to mindreaders who can only observe.
+
+As this example indicates,
+exploiting your-goal-is-my-goal 
+can shift the burden of identifying goals from a mindreader to her target.
+In the example Katarina is the focal mindreader and we are her target;
+but her success in identifying the goal of our actions
+depends on this,
+that our willingness to act jointly with her is based on \emph{our} knowledge of the goals of \emph{her} actions.
+In purely observational mindreading, the target's beliefs about the goals of the mindreader's actions are not normally relevant (except, of course, when the mindreader is ascribing such beliefs).
+But interacting mindreaders who rely on your-goal-is-my-goal
+thereby rely on their targets' having correctly identified the goals of their actions.
+Of course this is sometimes a reason not to rely on your-goal-is-my-goal.
+But where the target has better understanding of relevant means-ends relations,
+such as actions involving novel tools,
+the your-goal-is-my-goal route to knowledge of other's goals is sometimes the only option.
+
+The distant promise of all this is that 
+understanding how interaction widens the evidential basis for mindreading
+may eventually enable us to explain
+the origins (in evolution or development, and ideally both) of abilities to learn novel and opaque uses for tools from other agents
+without assuming that 
+rich communicative skills
+or 
+sophisticated forms of mindreading 
+%(such as the ability to ascribe propositional attitudes)
+must already be  present.
+
+\section{Communicative gestures}
+When introducing the problem of opaque means 
+(in section \vref{sec:opaque_means})
+we suggested that it could sometimes affect
+communicative actions.
+To illustrate this suggestion we drew on 
+an experiment by \citet{hare_chimpanzees_2004}
+in which participants had to find a reward 
+and were helped either by a failed reach or by a helpful point to the target location.
+Strikingly, for chimpanzees the helpful point is no help at all---even though it superficially resembles the failed reach, which occurred first and did help.
+Taking this paradigm as our case study, 
+we want to suggest that your-goal-is-my-goal might 
+enable us to understand how abilities to engage in joint action 
+could be part of what enables mindreaders to make the transation from 
+a simple understanding of goals 
+to 
+understanding of communicative intent.
+
+\section{The problem of false belief}
+\label{sec:false_belief}
 We have just seen that 

### 2011-12-16 joint action and knowing others minds/interacting mindreaders.2.tex
slow progress

@@
 \footnote{
+  \label{fn:communicative_intention}
 Hare and Tomasello 
@@
 \section{Joint action}
-\label{joint_action}
+\label{sec:joint_action}
 Our overall aim is to defend this claim:
@@
 how might this provide her with a route to knowledge of the goals of their actions?
-The intuitive idea we started with hinges on joint action, a particular form of interaction in which the actions of two or more agents are directed to a single goal (see section \vref{joint_action}).
+The intuitive idea we started with hinges on joint action, a particular form of interaction in which the actions of two or more agents are directed to a single goal (see section \vref{sec:joint_action}).
 Our intuitive idea was this:
@@
 an experiment by \citet{hare_chimpanzees_2004}
-in which participants had to find a reward 
-and were helped either by a failed reach or by a helpful point to the target location.
-Strikingly, for chimpanzees the helpful point is no help at all---even though it superficially resembles the failed reach, which occurred first and did help.
+in which the chimpanzee participants had to find a reward 
+and were helped by being shown either a failed reach or  a helpful point to the target location.
+Strikingly, for chimpanzees the helpful point is no help at all---even though it superficially resembles the failed reach, which  did help.
 Taking this paradigm as our case study, 
@@
 enable us to understand how abilities to engage in joint action 
-could be part of what enables mindreaders to make the transation from 
+could be part of what enables mindreaders to make the transition from 
 a simple understanding of goals 
 to 
-understanding of communicative intent.
+an early understanding communicative actions.
+
+Let us put ourselves in the chimpanzee's position for a moment.
+We witness the pointing action.
+With our eyes we follow the point to a container \citep[see][p.\ 6]{Moll:2007gu}.
+But in selecting a container we are no more likely to choose this one than the other.
+So we associate the pointing action with its target, but we fail to ascribe any goal which would clue us in to the relevance of this target for our own actions.
+Now suppose 
+that 
+before pointing the agent had used engaging facial gestures to signal willingness to engage in joint action with us,
+that
+we were able to think of retrieving the food as a possible distributive goal%
+\footnote{
+    On \emph{distributive goal} see section \vref{sec:joint_action}.
+  }
+of our actions,
+and 
+that we had exploited the my-goal-is-your-goal inference.
+Then we would believe,
+perhaps mistakenly,
+that a goal of the pointing action was to retrieve the food.
+In which case the pointing action would have been no less helpful in enabling us to succeed than the failed reach (which, as you may recall, was very helpful).
+
+It is natural to suppose that the difficulty chimpanzees have in Hare and Tomasello's experiments with responding appropriately to helpful pointing but not to failed reaching
+is due to a failure to understand communicative intention.%
+\footnote{
+See also footnote \vref{fn:communicative_intention}.
+Relatedly,
+in their discussion of these findings
+Moll and Tomasello suggest that
+`to understand pointing, the subject needs to understand more than the individual goal-directed behaviour. 
+She needs to understand that by pointing towards a location, the other attempts to communicate to her where a desired object is located; that the other tries to inform her about something that is relevant for her'
+\citep[p.\ 6]{Moll:2007gu}.
+}
+What we are suggesting is not that this must be wrong, 
+but only that participants must also have been unable to 
+exploit the your-goal-is-my-goal inference.
+
+Consider a related experiment by \citet{leekam_adults_2010}.
+Again the participants, 
+who were two- and three-year-old children,
+had to retrieve a reward from one of several closed containers.
+This time they were shown an adult holding up a replica of the target container.
+Leekam and colleagues found that 
+when this action was accompanied by an engaging facial expression
+three-year-old children were significantly better at identifying the correct container compared to
+when the the action was accompanied by a neutral facial expression (p.\ 116).
+Why did the engaging facial expression enhance performance?
+The authors consider the idea that 
+`the adult’s social cues conveyed her communicative intent, which in turn encouraged the child to ‘see through the sign’ ... helping them to take a dual stance to it' 
+\citep[p.\ 118]{leekam_adults_2010}.
+But an alternative possibility is that children succeeded without understand the replica as a sign at all.
+Instead they may 
+have associated the replica with the container it resembled
+(which by itself was not enough to motivate them to select this container, of course)
+and 
+have exploiting my-goal-is-your-goal to infer that a goal of the action of holding up the replica was to find the reward.
+In this way they might have misunderstood the action 
+as like the failed pointing
+in being a failed attempt to retrieve the reward.
+
+*Point is not to defend conjecture about actual performance but to ...
+
+*productive misunderstandings can result in understanding communicative gesture.s
+
+***
+
+Note that I am not suggesting that young children might fail to understand communicative intent.
+I am suggesting that they might first understand the goals of pointing actions without understanding communicative intent.
+But of course once they understand the goals of pointing actions within the context of joint action, it's likely that they will be able to understand the goals of pointing actions outside the context of joint action too.\footnote{
+Contrast Csibra's `two stances' idea.
+The referential action understanding involves a “stance” (p.
+455); teleological and referential action interpretation “rely on different kinds of action understanding' \citep[p.\ 456]{Csibra:2003kp}; they are initially two distinct `action interpretation systems' (although of course they come together later in development)  \citep[p.\ 456]{Csibra:2003kp}.
+In relation to Csibra's, my suggestion is not that there is no referential stance.
+It's rather that the referential stance might emerge from what he calls the `teleological stance' together with abilities to engage in the sort of joint actions that are characterised by shared goals.
+}
+
+Joint action may explain how individuals starting with a simple understanding of goals end up understanding communicative intentions.
+
+This is one illustration of how capacities for joint action, even very simple forms of joint action, might be relevant to explaining the development or evolution of richer forms of cultural cognition.
 

### 2011-12-19 joint action and knowing others minds/interacting mindreaders.2.tex
communication bit first attempt done

--- 2011-12-16 joint action and knowing others minds/interacting mindreaders.2.tex
@@
 such as actions involving novel tools,
-the your-goal-is-my-goal route to knowledge of other's goals is sometimes the only option.
+the your-goal-is-my-goal route to knowledge of other's goals may sometimes be the only option.
 
@@
 (in section \vref{sec:opaque_means})
-we suggested that it could sometimes affect
+we saw that it could  affect
 communicative actions.
@@
 an experiment by \citet{hare_chimpanzees_2004}
-in which participants had to find a reward 
-and were helped either by a failed reach or by a helpful point to the target location.
-Strikingly, for chimpanzees the helpful point is no help at all---even though it superficially resembles the failed reach, which occurred first and did help.
-Taking this paradigm as our case study, 
+in which chimpanzees had to find a reward 
+and were helped by being shown either a failed reach or  a helpful point to the target location.
+Strikingly, for chimpanzees the helpful point is no help at all---even though it superficially resembles the failed reach, which  did help.
+Taking this paradigm as a case study, 
 we want to suggest that your-goal-is-my-goal might 
 enable us to understand how abilities to engage in joint action 
-could be part of what enables mindreaders to make the transation from 
+could be part of what enables mindreaders to make the transition from 
 a simple understanding of goals 
 to 
-understanding of communicative intent.
+an early understanding communicative actions.%
+\footnote{
+Please note that, although our discussion borrows an experimental paradigm
+our aim is not to argue for empirical hypotheses about chimpanzee social cognition.
+Our aim is only to argue for the theoretical significance of interaction for mindreading by showing 
+that your-goal-is-my-goal
+could \emph{in principle} enable
+individuals to make the transition 
+from
+ a simple understanding of goals
+to
+ an early understanding of communicative actions.
+Of course it would strengthen our argument if we could provide evidence to show that this actually happens.
+But for now we are concerned with more narrowly conceptual issues.
+}
+
+Let us imagine ourselves as the chimpanzee  for a moment.
+We witness the pointing action.
+With our eyes we follow the point to a container \citep[see][p.\ 6]{Moll:2007gu}.
+So we do associate the pointing action with its target.
+But we are no more likely to choose this container than the other in seeking the reward.
+So we probably do not think of the pointing action as having any goal which would clue us in to the relevance of the container it indicates.
+(In principle we might perfectly understand the pointing action while failing to react to it in any systematic way because we are uncertain about the agent's integrity; but let us discount this possibility for the sake of illustration.)
+Now suppose 
+that, 
+before pointing, the agent had used engaging facial gestures to signal willingness to engage in joint action with us; and
+that
+we were able to think of retrieving the food as a possible distributive goal%
+\footnote{
+    On \emph{distributive goal} see section \vref{sec:joint_action}.
+  }
+of our actions;
+and 
+that we had exploited the my-goal-is-your-goal inference.
+Then we would believe,
+perhaps mistakenly,
+that a goal of the pointing action was to retrieve the food.
+In which case the pointing action would have been no less helpful in enabling us to succeed than the failed reach---which, as you may recall, was very helpful.
+So the your-goal-is-my-goal inference can 
+enable interpreters
+to  
+misunderstand pointing actions as something like  failed reaches rather than as communicative gestures.
+This means that,
+even without any understanding of communication,
+they can respond appropriately to helpful pointing actions in the context of joint action.
+
+It is natural to suppose that the difficulty chimpanzees have in Hare and Tomasello's experiments with responding appropriately to helpful pointing but not to failed reaching
+is due to a failure to understand communicative intention.%
+\footnote{
+See  footnote \vref{fn:communicative_intention}.
+Relatedly,
+in their discussion of these findings
+Moll and Tomasello suggest that
+`to understand pointing, the subject needs to understand more than the individual goal-directed behaviour. 
+She needs to understand that by pointing towards a location, the other attempts to communicate to her where a desired object is located; that the other tries to inform her about something that is relevant for her'
+\citep[p.\ 6]{Moll:2007gu}.
+Assuming this is right, our suggestion is that individuals could reliably  respond  appropriately to pointing actions in the context of joint action without understanding pointing.
+}
+What we are suggesting is that participants must also have been unable or unwilling to 
+exploit the your-goal-is-my-goal inference.
+
+Consider a related experiment by \citet{leekam_adults_2010}.
+Again participants had to retrieve a reward from one of several closed containers, but this time they were two- and three-year-old children.
+In one condition participants were shown an adult holding up a replica of the target container.
+Leekam and colleagues found that 
+when this action was accompanied by an engaging facial expression,
+three-year-old children were significantly better at identifying the correct container compared to
+when the the action was accompanied by a neutral facial expression (p.\ 116).
+Why did the engaging facial expression enhance performance?
+The authors consider the idea that 
+engaging facial gestures somehow help children to understand communicative intentions.%
+\footnote{
+\citet[p.\ 118]{leekam_adults_2010}: `the adult’s social cues conveyed her communicative intent, which in turn encouraged the child to ‘see through the sign’ ...
+helping them  to take a dual stance to it.'
+}
+An alternative possibility is that children succeeded without understand the replica as a sign at all.
+Instead they may have 
+associated the replica with the container it resembled
+(which by itself is not enough to motivate selecting this container, of course),
+regarded the engaging facial gestures as expressing willingness to engage in joint action,
+and 
+exploited your-goal-is-my-goal 
+	to infer that a goal of the action of holding up the replica was to find the reward.
+In this way they might have understood (or misunderstood) the action of holding up the replica as 
+like a failed reach
+in being an attempt to retrieve the reward.
+
+So far we have illustrated how your-goal-is-my-goal enables responding appropriately to communicative gestures with two examples, pointing and holding up a replica.
+The pattern of reasoning generalises to a wider range of communicative gestures including single-word utterances.
+The basic requirement is this: in a particular context, the interpreter must associate a communicative gesture with its referent.
+For instance, she must associate the pointing gesture with the object indicated; or, if (say) she is looking for an object she must associate an utterance of `cupboard' with the nearby cupboard.
+As we saw, 
+outside the context of joint action,
+merely associating a gesture with its referent falls short of being able to respond appropriately.
+After all, the goals of the gesture may well be unrelated to the goals of the interpreter's actions.
+But if an interpreter supposes that her target is willing to engage in joint action with her,
+then she may infer that the goal of her target's action is her goal
+and so be motivated to treat the thing associated with a communicative gestures as relevant to the goal of her own actions.
+This will reliably (but not always) enable her to respond appropriately to the communicative gesture even without understanding it as a communicative gesture.
+And once she has experienced how that communicative gesture works as a tool for guiding others' actions in the context of joint action,
+she may be in a position to realise, further, that the same tool can be used in other contexts.
+
+This, in barest outline, is 
+how
+abilities to engage in joint action
+mean that 
+a mindreader with an ability to ascribe simple goals only (and no understanding of communicative intent)
+might 
+nevertheless reliably respond appropriately to some communicative gestures,
+and so be in a position to understand how such gestures can be used to guide others' actions.%
+\footnote{
+Contrast Csibra's claim that,
+early in human development,
+goal ascription (`teleological understanding' in his terms)
+and 
+identifying the referents of communicative gestures (`referential understanding')
+`rely on different kinds of action understanding' %\citep[p.\ 456]{Csibra:2003kp}
+and
+are initially two distinct `action interpretation systems'   \citep[p.\ 456]{Csibra:2003kp}.
+%(although of course they come together later in development)
+We have not shown that this hypothesis wrong.
+But we have shown that there is another possibility:
+the referential stance may 
+emerge from
+the teleological stance
+ together with abilities to engage in simple forms of joint action.
+}
+Of course we have only argued that this transition is theoretically possible;
+we have not attempted to defend any hypothesis about evolution or development.
+Our point is just to illustrate one of several ways in which
+interaction may matter for mindreading.
+Widening the evidential basis for goal ascription
+to include evidence available only to interpreters capable of joint action with their targets
+makes possible 
+novel hypotheses about the emergence, in evolution or development (and perhaps even both), of communicative gestures.
+
+%This is one illustration of how capacities for joint action, even very simple forms of joint action, might be relevant to explaining the development or evolution of richer forms of cultural cognition.
 

### 2011-12-19 joint action and knowing others minds/interacting mindreaders.2.tex
small changes only

@@
 \begin{abstract}
-***
+%Goal ascription is fundamental for mindreading.
+What evidence grounds ascriptions of thoughts and actions,
+and how does the evidence support the ascriptions?
+In answering this question, 
+philosophers sometimes ignore interaction and focus on mere observation.
+The only evidence considered is 
+evidence that would be available to mindreaders who observe their targets but are not necessarily in a position to interact with them.
+The present paper identifies evidence 
+that is available only to mindreaders capable of interacting with their targets.
+Being poised to interact with others may enable one to know things about their minds which one might not otherwise be in a position to know.
+
+%What evidence grounds ascriptions of goals to other agents' actions,
+%and how does the evidence support the ascriptions?
+%It is usually assumed that, in answering this question, 
+%we can focus on observation and ignore interaction.
+%We can restrict ourselves to evidence that would be available to mindreaders who observe their targets but are not necessarily in a position to interact with them.
+%The present paper identifies evidence 
+%which supports goal ascriptions
+%and 
+%is available only to mindreaders who are capable of interacting with their targets.
+
 \end{abstract}
@@
 Recent interest in non-propositional precursors to mindreading makes this appropriate.
+
+***Goal ascription essential for verifying predictions of action.
 
@@
 
+In principle,
+exploiting your-goal-is-my-goal does not require that the interpreter actually be in a position to interact with her target.
+It is sufficient (and would be necessary but for some special cases) that the target takes the interpreter to be in a position to interact with him. 
+Our concern, however,
+is with cases that are likely to be important for understanding
+development or evolution (and ideally both).
+This mandates a focus on interpreters of limited sophistication
+who lack both deep insight into others' minds and fully-fledged communicative abilities.
+Such interpreters are unlikely to be able to contrive or exploit situations in which they only appear to be in a position to interact with their targets.
+So the explanatorily relevant cases are those 
+in which an interpreter is manifestly in a position to interact with her target.
+
 Our aim in identifying the your-goal-is-my-goal is not to 
@@
 but because 
-its existence reveals that interacting mindreaders can have access to evidence unavailable to those relying entirely on observation.
+its existence reveals that interacting mindreaders can have evidence that would be unavailable to them if they were relying   entirely on observation.%
 

### 2011-12-20 joint action and knowing others minds/interacting mindreaders.2.tex
problem of false belief section first draft done, and first draft conclusion

@@
 As this illustrates, ignorance about to which ends actions are means can be an obstacle to goal ascription.%
-\footnote{
-This is not to say that no observer could ever identify the goal of any action she fails to recognise as a means to achieving that goal.
-Surely opaque means are not in every case an insurmountable obstacle to goal ascription.
-Our point is just that opaque means \emph{sometimes}  deprive observers of evidence and so prevent goal ascription.
-}
 Call this the problem of opaque means.
+
+We are not suggesting that no observer could ever identify the goal of any action she fails to recognise as a means to achieving that goal, of course.
+Of course opaque means are not in every case an insurmountable obstacle to goal ascription.
+Our point is pure commonsense and already widely appreciated:
+opaque means sometimes  deprive interpreters of evidence and so prevent goal ascription.
 
@@
 Rather, it is a potential problem \emph{for interpreters}.
-The existence of this problem shows, we shall argue, that models of interpretation based on pure observation are less powerful than models taking into account the possibility of interaction.
+
+That the problem of opaque means exists
+is the first premise in our argument
+that models of interpretation based on pure observation are less powerful than models taking into account the possibility of interaction.
 They are less powerful in this sense: 
@@
 but because 
-its existence reveals that interacting mindreaders can have evidence that would be unavailable to them if they were relying   entirely on observation.%
+it characterises a route to knowledge
+that is closed to mere observers
+but open to interacting mindreaders.
 
@@
 Imagine we are interacting with a young child,
-Katarina,
+Ayesha,
 and want her to understand how a new tool is used.  
@@
 knowing that she will perform actions directed to achieving this outcome.  
-We then signal to Katarina that we are willing to act jointly with her.  
+We then signal to Ayesha that we are willing to act jointly with her.  
 Now she is in a position to know what the goal of our action will be when we deploy the tool.  
@@
 can shift the burden of identifying goals from a mindreader to her target.
-In the example Katarina is the focal mindreader and we are her target;
+In the example Ayesha is the focal mindreader and we are her target;
 but her success in identifying the goal of our actions
@@
 mean that 
-a mindreader with an ability to ascribe simple goals only (and no understanding of communicative intent)
+a mindreader with an ability to ascribe simple goals only and no understanding of communicative intent
 might 
 nevertheless reliably respond appropriately to some communicative gestures,
-and so be in a position to understand how such gestures can be used to guide others' actions.%
+and so come be in a position to understand how such gestures can be used to guide others' actions.%
 \footnote{
@@
 Of course we have only argued that this transition is theoretically possible;
-we have not attempted to defend any hypothesis about evolution or development.
+we have not attempted to defend any hypothesis about 
+anything's evolution or development.
 Our point is just to illustrate one of several ways in which
@@
 \label{sec:false_belief}
-We have just seen that 
-failure to identify to which ends actions are means can impair goal ascription; this was the problem of opaque means.
-Another problem 
-affecting goal ascription given purely observational models of interpretation
+So far we have been arguing for the value of interaction to interpreters 
+on the grounds that 
+interacting interpreters, unlike mere observers, have 
+a route to knowledge of the goals of others' actions
+which avoids the problem of opaque means.
+There is another reason for supposing that interaction matters for interpretation.
+To introduce this reason we must first describe another problem affecting interpreters, the problem of false belief.
+
+The problem of opaque means occurs when interpreters  must rely entirely on observation and cannot identify to which ends actions are means.
+A yet more familiar problem affecting goal ascription 
 arises from the interdependence of beliefs and goals.
 To illustrate, 
-imagine sitting at a table on which two closed opaque boxes each contain an object; one contains an owl, the other a cat.
-If the goal of your action is to retrieve the owl, 
-and you believe that the owl is in the north-most box,
+imagine sitting at a table.
+On the table are two closed opaque boxes.
+One box contains an owl, the other a cat.
+If the goal of your action is to retrieve the cat, 
+and you believe that the cat is in the north box,
 then 
 (unless things are going very badly) 
-you will reach into the north-most box.
-But of course if you had believed instead that the owl was in the other box,
-then, in acting on the same goal, you would have reached into the other box.
-Now consider someone observing your actions.
-If she is has good reason to believe, falsely, that you know the owl is in the south-most box,
-then she may be justified in supposing, incorrectly, that when you reach into the north-most box the the goal of your action is to retrieve the cat.
+you will reach for the north box.
+But of course if you had believed instead that the cat was in the south box,
+then, in acting on the same goal, you would have reached for the south box.
+Now consider Ayesha who is observing your actions.
+Suppose Ayesha has sufficient reason to believe, falsely, that you know the cat is in the south box.
+Then she may be justified in supposing, incorrectly, that 
+the goal of your action, in reaching for the north box,
+is to retrieve the owl.
 As this illustrates,
-differences in belief between observers and protagonists can be an obstacle to goal ascription.
+differences in belief between observers and protagonists can 
+impair goal ascription
+when the interpreter is unaware of those differences.
 Call this the problem of false belief.
+
+The your-goal-is-my-goal
+route to knowledge
+sometimes enables interpreters to avoid the problem of false belief.
+To illustrate consider a counterfactual alternative to the above example.
+Ayesha dislikes the owl and is concerned with retrieving the cat.
+As you reach for the north box,
+your facial gestures signal willingness to engage in joint action with Ayesha.
+She then concludes, correctly, that your goal is her goal, to retrieve the cat.
+So despite the difference in belief,
+the possibility of interaction means that
+Ayesha can knowledgably identify the goal of your action.
+
+Exploiting your-goal-is-my-goal does not make it possible to  avoid the problem of false belief entirely,
+it only shifts  the problem.
+To illustrate,
+in the above example Ayesha's ability to correctly identify the goal of your action
+depends on your correctly anticipating the goals of her actions---on your knowing that she is concerned with retrieving the cat.
+If you incorrectly anticipated that her actions would be directed to retrieving the owl,
+Ayesha would have been mistaken in taking your actions to be directed to retrieving the cat.
+As this indicates,
+exploiting your-goal-is-my-goal 
+can shift 
+both 
+\emph{which} differences in belief have the potential to impair  goal ascription
+and also 
+\emph{who} needs to be aware of those differences.
+
+In short, then, abilities to engage in joint action provide an interpreter with a route to knowledge of the goals of other agents' actions which does not depend on her knowledge of what her target believes.
+This is not because abilities to engage in joint action provide a way to avoid the problem of false beliefs altogether.
+Rather they shift the burden of resolving the problem of false belief from an interpreter to her target.
+This is potentially valuable for interpreters who have limited insight into their targets' beliefs,
+or who lack abilities to track differences in belief altogether.
+
+On a purely observational model of interpretation,
+it seems inevitable that an interpreter who is unable to track others' beliefs 
+will have to proceed as if there were no differences in beliefs.
+As several philosophers' arguments suggest,
+it is plausible that 
+the interdependence of beliefs and goals 
+means that
+an interpreter will always be able to assign goals to others' actions 
+in a way that is consistent with their beliefs not differing from her own
+(*refs: Davidson; \citealp[p.\ 49]{Bennett:1976rg}).
+If so,
+on a purely observational model of interpretation
+there is no straightforward way in which
+the need to track differences in belief 
+will become apparent to an interpreter.
+Things are different
+if we extend our model of interpretation
+to take into account the possibility of interaction.
+We can then see how
+an interpreter who is unable to track others' beliefs
+will encounter anomalies.
+For some differences in belief will mean that
+ ordinary, purely observational routes to knowledge will result in ascribing one goal
+ whereas
+ exploiting your-goal-is-my-goal would result in a ascribing a different goal.
+That is, the evidence available to an interpreter may inexplicably support two incompatible ascriptions. 
+In this way,
+differences in belief have the potential to cause puzzling anomalies
+even before the interpreter is able to track such differences.
+This 
+raises the possibility
+that abilities to engage in joint action
+may play a role in explaining how  awareness of belief emerges.
+
+\section{***}
+
+***MOVE:
 
@@
 
-In what follows we shall argue that abilities to engage in joint action provide an interpreter with a route to knowledge of the goals of other agents' actions which does not depend on her abilities to ascribe beliefs.
-This is not because abilities to engage in joint action provide a way to avoid the problem of false beliefs altogether.
-Rather, as we shall see, abilities to engage in joint action provide a way to shift the burden of resolving the problem of false belief from an interpreter to her target.
+\section{Conclusion}
+Our aim was to show that interaction benefits mindreading in this sense:
+some routes to knowledge are closed to mindreaders who rely exclusively on observation
+but
+open to interacting mindreaders.
+In pursuing this aim we focused on interactions which are joint actions.
+Our suggestion was this.
+Where an interpreter recognizes that her target is willing to engage in joint action
+and is unwilling to change which goals her own actions will be directed to,
+she may be in a position to know
+that the goals of her target's actions will be her goals.
+This in outline is the `your-goal-is-my-goal' route to knowledge.
+
+To show that this route to knowledge is potentially valuable,
+we argued that it enables interpreters to overcome two problems,
+the problem of opaque means (section *)
+and the problem of false belief (section *).
+These problems may rarely arise for human adults thanks to sophisticated linguistic communication and extensive knowledge of how things can be achieved.  
+But our focus has been on 
+  very young humans,
+  or other species,
+  or earlier stages of evolution---%
+on cases where there are 
+  absent or fledgling communicative skills,
+  limited insight into others' minds,
+  and narrow knowledge of how things work.
+Where this is the case,
+the problems of false belief and opaque means may easily arise,
+particularly in cases involving novel tools 
+and 
+communicative gestures.
+In principle these interpreters 
+could overcome the two problems 
+by expliting your-goal-is-my-goal.
+This demonstrates that 
+expertise with 
+  tools 
+  and
+  communicative gestures
+does not presuppose rich insights into others intentions and beliefs.
+It is consistent to suppose that 
+such expertise,
+far from  presupposing sophisticated forms of mindreading,
+might instead
+play a role in explaining their emergence in evolution or development (or both).
 

### 2011-12-20 joint action and knowing others minds/interacting mindreaders.2.tex
small fixes

@@
 that is available only to mindreaders capable of interacting with their targets.
-Being poised to interact with others may enable one to know things about their minds which one might not otherwise be in a position to know.
+Being able to interact with others may enable one to know things about their minds which one might not otherwise be in a position to know.
+This matters for understanding possible roles for interaction in the evolution and development of mindreading.
 
@@
 \section{Conclusion}
-Our aim was to show that interaction benefits mindreading in this sense:
+Our aim was to show that interaction can facilitate mindreading in this sense:
 some routes to knowledge are closed to mindreaders who rely exclusively on observation
@@
 we argued that it enables interpreters to overcome two problems,
-the problem of opaque means (section *)
-and the problem of false belief (section *).
+the problem of opaque means
+and the problem of false belief.
 These problems may rarely arise for human adults thanks to sophisticated linguistic communication and extensive knowledge of how things can be achieved.  
-But our focus has been on 
-  very young humans,
-  or other species,
-  or earlier stages of evolution---%
-on cases where there are 
+But interest in the evolution or development of mindreading
+motivates focusing on 
+interpreters with
   absent or fledgling communicative skills,
-  limited insight into others' minds,
+  limited insight into others' minds
   and narrow knowledge of how things work.
-Where this is the case,
+For such interpreters
 the problems of false belief and opaque means may easily arise,
@@
 communicative gestures.
-In principle these interpreters 
+Interpreters 
 could overcome the two problems 
-by expliting your-goal-is-my-goal.
-This demonstrates that 
+by exploiting the your-goal-is-my-goal route to knowledge.
+It follows that 
 expertise with 
@@
   communicative gestures
-does not presuppose rich insights into others intentions and beliefs.
-It is consistent to suppose that 
-such expertise,
+does not presuppose rich insights into others' intentions and beliefs.
+Such expertise,
 far from  presupposing sophisticated forms of mindreading,

### 2011-12-21 joint action and knowing others minds/interacting mindreaders.2.tex
some progress with introduction and second section on goal ascription

@@
 In answering this question, 
-philosophers sometimes ignore interaction and focus on mere observation.
+philosophers sometimes focus on mere observation, ignoring interaction.
 The only evidence considered is 
-evidence that would be available to mindreaders who observe their targets but are not necessarily in a position to interact with them.
-The present paper identifies evidence 
-that is available only to mindreaders capable of interacting with their targets.
-Being able to interact with others may enable one to know things about their minds which one might not otherwise be in a position to know.
-This matters for understanding possible roles for interaction in the evolution and development of mindreading.
+evidence that would be available to mindreaders who observe their targets but are in no position to interact with them.
+The present paper,
+which focuses on goal ascription,
+argues that this is a mistake
+and
+identifies evidence 
+available only to mindreaders capable of interacting with their targets.
+Being poised to interact with others may make it possible to know things about their thoughts and actions which one might not otherwise be in a position to know.
+This has consequences for the possible roles of interaction in explaining the evolution and development of mindreading.
 
@@
 
-The question is what constitutes evidence.
+Many people expect to know what those around them are thinking and doing, within limits.
+If someone were to complain of a co-worker that she can never tell what he is thinking or doing,
+the complaint might be informative, surprising even.
+And we might reasonably wonder where the fault lies.
+Maybe the co-worker really is a closed book,
+but maybe the complainant is not really trying to understand him.
+For humans and perhaps other species too,
+access to the actions and minds of those nearby,
+to some of their beliefs and desires,
+emotions, intentions and knowledge states,
+is often normal in this sense: it is lack of access, not access, that is remarkable.
+In this respect we might compare identifying others' thoughts and actions with identifying their ages or masses.
+Although all of these can be disguised with the right props and techniques, some degree of revelation is normal.
 
-Mostly people have supposed that the evidence is available to pure observers --- being able to interact with the targets of mindreading confers no in principle advantage.
+The normalcy of mindreading,
+together with the fact that it often occurs spontaneously (if imperfectly),%
+%
+\footnote{
+On the spontaneity of mindreading in human adults see 
+\citet{senju_mindblind_2009} and 
+\citet{ferguson_eye_2011}.
+On imperfections in human adults' mindreading, 
+see \citet{Keysar:2003xu}.
+Both points come together in \citet{ferguson_listeners_2012}.
+}
+%
+make it easy to 
+overlook questions about its evidential basis.
+Indeed,
+while philosophers have engaged with questions about mechanisms (such as whether mindreading involves a process of simulation or of theorizing or some combination of the two),
+comparatively little effort has been devoted to issues about what evidence could ground mindreading.
+In this paper we pursue one such issue.
 
-In challenging this thesis we want to start with goal ascription, which is either a precursor to or a component of mindreading.
-Recent interest in non-propositional precursors to mindreading makes this appropriate.
+%In everyday life it is sometimes so obvious what others are thinking and doing
+%that it is easy to forget that mindreaders begin life with no knowledge of the specific contents of particular individuals' thoughts 
+%and that to acquire this knowledge
+%there must be
 
-***Goal ascription essential for verifying predictions of action.
+What is the evidential basis for ascriptions of thought and action and how does the evidence support the ascriptions?
+To stress that this question is not directly about about \emph{how} anyone actually ascribes thoughts or identifies actions% 
+% nor directly about the mechanisms, processes or representations involved in mindreading
+, Davidson 
+sometimes formulates the question by asking what someone \emph{could} know that would put them in a position to identify another's thoughts and actions \citep[e.g.][p.\ 126]{Davidson:1973jx}.
+
+The two most sustained and elaborate attempts to answer this question,
+Davidson's (\citeyear{Davidson:1984wh}) and Dennett's (\citeyear{Dennett:1987sf}),
+do not exploit the possibility of interaction.
+The evidential bases they consider and the principles which they identify to link the evidence with ascriptions of thought and action are available to entirely passive observers.
+So on their theories,
+a purely passive mindreader observing from behind a one-way mirror
+is on a par with
+a mindreader who does or could interact with those she seeks to interpret.
+The two are on a par in this sense:
+in principle the same evidence could be available to each, and each can exploit the same principles in moving from evidence to ascriptions of thought.
+In this paper we aim to show
+that mindreaders who are capable of interacting with their targets are at an advantage.
+Their ascriptions can be justified by evidence and principles which would be unavailable if they were entirely passive observers.
+
+Being poised to interact with others enables one to know things about their minds which one might not otherwise be in a position to know; or so we aim to show in what follows.
+Because the leading theories have supposed that being able to interact makes no difference,
+they have treated the evidential basis for mindreading as more narrow than it truly is.
+This matters because broadening the evidential basis will enable us to give a more accurate and comprehensive theory of mindreading.
+And, as we shall argue, this in turn matters because it may bear on the possibility of explaining the emergence, in evolution or in development, of mindreading.
+Several researchers have offered quite general conjectures about how interaction might explain the emergence of sophisticated forms of cognition.%
+\footnote{
+\citet[p.\ 1]{Moll:2007gu}
+argue for the `Vygotskian Intelligence Hypothesis' according to which `the unique aspects of human cognition ... were driven by, or even constituted by, social co-operation.'
+\citet[p.\ 103]{Knoblich:2006bn} offer a different conjecture:
+`perception, action, and cognition are grounded in social interaction  … functions traditionally considered hallmarks of individual cognition originated through the need to interact with others.'
+}
+Studying how interaction broadens the evidential basis of mindreading will eventually point to one way of filling in some details.
+
+%This, then, is our pitch.
+%To formulate (and, eventually, test) hypotheses about how mindreading develops,
+%or about how it evolves,
+%it is useful to understand not only which mechanisms mindreading involves
+%but also what evidence grounds mindreading.
+%And it since it turns out
 
@@
 Goal ascription is the process of identifying which outcomes others' purposive actions are directed to.
+Some might argue that goal ascription is not mindreading because 
+identifying relations between actions and the outcomes to which they are directed
+does not necessarily involve ascribing mental states.
+Recognizing the force of this argument,
+we wish to remain neutral on whether or not goal ascription is actually mindreading.
+Either way,
+goal ascription is essential in nearly all mindreading.
+For mindreading typically involves or depends on predictions about how individuals will act,
+and those predictions can rarely be verified without the mindreader identifying which goals the actions are directed to.
+
+We shall focus on goal ascription.
+This is partly because doing so  greatly simplifies our argument,
+but mainly because goal ascription is widely thought to be among the very earliest components of mindreading to emerge (or, if goal ascription is not mindreading, then it is a late precursor).%
+\footnote{
+See, for example,
+\citet{Gergely:1995sq} and 
+\citet{Woodward:1998dm}.
+See also
+\citet[p.\ 111, Box 1]{Baillargeon:gx}
+on two subsystems
+and 
+\citet{Povinelli:2001jf} on `behavioural regularities'.
+}
+By showing that
+  interacting mindreaders may have access to 
+  evidence for goal ascriptions 
+  which is unavailable to those who merely observe
+we may be able to indicate ways in which interaction could facilitate the emergence, in evolution or development, of more sophisticated forms of mindreading.%
+\footnote{
+Of course if goal ascription is not mindreading,
+then in showing that ***
+... indirectly ***
+}
+
+***use from below (and fn 22ish):
+What evidence could support hypotheses about the outcomes to which actions are directed? And how would the evidence support the hypotheses?22
+***
+
 Because goal ascription has received little attention,
 we  first review some of the benefits that being able to identify others' goals brings.
+
+***Recent interest in non-propositional precursors to mindreading makes focus on goal ascription this appropriate.
 
@@
 So, on this view, those without metarepresentational abilities are not in a position to know the goals of other agents' actions.
+
+***
+
+There is disagreement over whether it is possible to knowledgably identify the goals of an agent's actions without also ascribing some beliefs to that agent.  
+\citet[pp.48--50]{Bennett:1976rg}
+%\citet[pp.48--50]{Bennett:1976rg} suggests that a theory of goals ascription has to be developed together with a theory of (proto-)belief ascription: `An animal’s behaviour does not show what it registers unless we know what it seeks; but how can we learn what it seeks before we know what it registers?' [p.
+48]
+and 
+\citet{Davidson:1974gh}
+both appear to hold that this is impossible,
+that identifying goals cannot be done independently of ascribing beliefs.
+By contrast,
+\citet{Gergely:1995sq},
+*Meltzoff?,
+\citet{Baillargeon:gx}
+and 
+\citet{Woodward:1998dm}
+(among many others)
+appear to assume the opposite, 
+that it is possible to identify goals without even being able to ascribe beliefs.
+For what it is worth, we tentatively favour this latter position.%
+\footnote{
+It is striking that, as far as we can tell, neither Bennett nor Davidson offers an argument for this claim.
+They do note that beliefs and goals make an interdependent contribution to observed action.
+But this by itself does not show that goal ascription cannot in some cases involve justifiably ignoring the possibility of differences in belief between interpreters and their targets.
+For instance,
+suppose that two people are sitting opposite each other at a low table
+ which is 
+sparsely populated with objects.
+The objects are all out in the open; manifestly, both can clearly see them.
+If one reaches to grasp one of these objects (the duck, say), 
+must the other ascribe beliefs in order to knowledgeably identify the goal of her action?
+On the face of it, she need not.  
+Even if she had no ability to ascribe beliefs, she might nevertheless be in a position to acquire knowledge of the goal of the other's action.
+}
+However, this debate is not directly relevant to our concerns here.
+Both sides can agree that 
+differences in belief between observer and protagonist
+are sometimes an obstacle to goal ascription.
+This is all that the problem of false belief requires.
 
@@
 
-\section{***}
-
-***MOVE:
-
-There is disagreement over whether it is possible to knowledgably identify the goals of an agent's actions without also ascribing some beliefs to that agent.  
-\citet[pp.48--50]{Bennett:1976rg}
-%\citet[pp.48--50]{Bennett:1976rg} suggests that a theory of goals ascription has to be developed together with a theory of (proto-)belief ascription: `An animal’s behaviour does not show what it registers unless we know what it seeks; but how can we learn what it seeks before we know what it registers?' [p.
-48]
-and 
-\citet{Davidson:1974gh}
-both appear to hold that this is impossible,
-that identifying goals cannot be done independently of ascribing beliefs.
-By contrast,
-\citet{Gergely:1995sq},
-*Meltzoff?,
-\citet{Baillargeon:gx}
-and 
-\citet{Woodward:1998dm}
-(among many others)
-appear to assume the opposite, 
-that it is possible to identify goals without even being able to ascribe beliefs.
-For what it is worth, we tentatively favour this latter position.%
-\footnote{
-It is striking that, as far as we can tell, neither Bennett nor Davidson offers an argument for this claim.
-They do note that beliefs and goals make an interdependent contribution to observed action.
-But this by itself does not show that goal ascription cannot in some cases involve justifiably ignoring the possibility of differences in belief between interpreters and their targets.
-For instance,
-suppose that two people are sitting opposite each other at a low table
- which is 
-sparsely populated with objects.
-The objects are all out in the open; manifestly, both can clearly see them.
-If one reaches to grasp one of these objects (the duck, say), 
-must the other ascribe beliefs in order to knowledgeably identify the goal of her action?
-On the face of it, she need not.  
-Even if she had no ability to ascribe beliefs, she might nevertheless be in a position to acquire knowledge of the goal of the other's action.
-}
-However, this debate is not directly relevant to our concerns here.
-Both sides can agree that 
-differences in belief between observer and protagonist
-are sometimes an obstacle to goal ascription.
-This is all that the problem of false belief requires.
-
 \section{Conclusion}
@@
 This in outline is the `your-goal-is-my-goal' route to knowledge.
+
+***Return to the leading theories?
+We do not take this to show that the leading theories must be entirely discarded.
+If we are right about interaction making evidence available, 
+either leading theory can be extended accordingly.
+***
 

### 2011-12-21 joint action and knowing others minds/interacting mindreaders.tex
some progress with introduction and second section on goal ascription

@@
 \section{Davidson's question: What grounds mindreading?}
-Many people expect to know what those around them think, within limits.
-If someone were to complain of a co-worker that she can never tell what she  thinking,
+Many people expect to know what those around them think and do, within limits.
+If someone were to complain of a co-worker that she can never tell what he is thinking or doing,
 the complaint might be informative, surprising even.
@@
 Maybe the co-worker really is a closed book,
-but maybe complainant is not really trying to understand her.
+but maybe complainant is not really trying to understand him.
 For humans and perhaps other species too,
@@
 is often the norm in this sense: it is lack of access, not access, that is remarkable.
+In this respect we might compare identifying others' thoughts with identifying their ages or masses.
+Although all of these can be disguised with the right props and techniques, some degree of revelation is normal.
+
 And often enough we read others' minds spontaneously
@@
 }
-In this respect we might compare identifying others' thoughts with identifying their masses.
-And although both can be disguised with the right props and techniques, some degree of revelation is normal.
 
@@
 
-We can clarify this question
+We can clarify this question ***
 
@@
 do not exploit the possibility of interaction.
-The evidential basis they consider and the principles which they identify to link the evidence with ascriptions of thought and action are available to entirely passive observers.
+The evidential bases they consider and the principles which they identify to link the evidence with ascriptions of thought and action are available to entirely passive observers.
 So on these theories,

### 2011-12-21 joint action and knowing others minds/interacting mindreaders.2.tex
small changes

@@
 The only evidence considered is 
-evidence that would be available to mindreaders who observe their targets but are in no position to interact with them.
+ evidence that would be available to mindreaders 
+observing their targets but in no position to interact with them.
 The present paper,
@@
 , Davidson 
-sometimes formulates the question by asking what someone \emph{could} know that would put them in a position to identify another's thoughts and actions \citep[e.g.][p.\ 126]{Davidson:1973jx}.
-
-The two most sustained and elaborate attempts to answer this question,
-Davidson's (\citeyear{Davidson:1984wh}) and Dennett's (\citeyear{Dennett:1987sf}),
+sometimes formulates the question by asking what someone \emph{could} know that would put them in a position to identify another's thoughts and actions.%
+\footnote{
+See, e.g., \citet[p.\ 126]{Davidson:1973jx}.
+While what follows draws  on Davidson's questions and insights,
+our aims here are more modest than his.
+For here we are concerned only with a  fraction of the problem of ascribing mental states and meanings,
+and we are not concerned with larger claims about mind, meaning and truth 
+(on which see 
+\citet{Davidson:1990du} and 
+\citet{lepore_donald_2005}).
+}
+
+The most sustained attempts to answer this question,
+Davidson's (\citeyear{Davidson:1984wh}; \citeyear{Davidson:1990du}), Lewis' (\citeyear{lewis:1974ri}) and Dennett's (\citeyear{Dennett:1987sf}),
 do not exploit the possibility of interaction.
@@
 In this paper we aim to show
-that mindreaders who are capable of interacting with their targets are at an advantage.
-Their ascriptions can be justified by evidence and principles which would be unavailable if they were entirely passive observers.
-
-Being poised to interact with others enables one to know things about their minds which one might not otherwise be in a position to know; or so we aim to show in what follows.
+that mindreaders  capable of interacting with their targets are at an advantage.
+Their ascriptions can rely on evidence and principles which would be unavailable if they were entirely passive observers.
+
+Being poised to interact with others enables one to know things about their minds which one might not otherwise be in a position to know, or so we aim to show in what follows.
 Because the leading theories have supposed that being able to interact makes no difference,
@@
 Goal ascription is the process of identifying which outcomes others' purposive actions are directed to.
-Some might argue that goal ascription is not mindreading because 
-identifying relations between actions and the outcomes to which they are directed
-does not necessarily involve ascribing mental states.
-Recognizing the force of this argument,
-we wish to remain neutral on whether or not goal ascription is actually mindreading.
-Either way,
-goal ascription is essential in nearly all mindreading.
-For mindreading typically involves or depends on predictions about how individuals will act,
-and those predictions can rarely be verified without the mindreader identifying which goals the actions are directed to.
-
-We shall focus on goal ascription.
-This is partly because doing so  greatly simplifies our argument,
+To illustrate, suppose that
+Hannah kicks a ball thereby both preventing her sisters from scoring and also breaking a window.
+Asked about episode,
+Hannah might protest that the goal of her action was not to break the window but only to avoid losing a game.
+As this illustrates,
+among the actual and possible outcomes of an action,
+only some are outcomes to which the action was directed.
+Goal ascription is the process of identifying from among an action's actual and possible outcomes
+those to which it is directed.
+We shall focus on goal ascription rather than the ascription of beliefs and mental states that has captured most interest.
+Our immediate question is therefore, 
+What evidence could support hypotheses about the outcomes to which actions are directed, 
+and how would the evidence support the hypotheses?
+
+We focus on goal ascription 
+partly to simplify our argument,
 but mainly because goal ascription is widely thought to be among the very earliest components of mindreading to emerge (or, if goal ascription is not mindreading, then it is a late precursor).%
@@
   evidence for goal ascriptions 
-  which is unavailable to those who merely observe
+  which is unavailable to those who merely observe,
 we may be able to indicate ways in which interaction could facilitate the emergence, in evolution or development, of more sophisticated forms of mindreading.%
-\footnote{
-Of course if goal ascription is not mindreading,
-then in showing that ***
-... indirectly ***
-}
-
-***use from below (and fn 22ish):
-What evidence could support hypotheses about the outcomes to which actions are directed? And how would the evidence support the hypotheses?22
-***
 
 Because goal ascription has received little attention,
-we  first review some of the benefits that being able to identify others' goals brings.
-
-***Recent interest in non-propositional precursors to mindreading makes focus on goal ascription this appropriate.
-
+we shall briefly review some of the benefits that being able to identify others' goals brings.
 It is a familiar idea that goal ascription enables one to learn from others' successes.
@@
 if you know or can guess that another agent's actions are directed to opening a nut,
-you may then be in a position to infer that the unfamiliar pattern of actions she is performing constitute a means to open nuts.\footnote{
-***
-\citep{Horner:2005pj}
-}
-Or (in a different case) knowing the goal of another agent's actions may enable you to discover a new use for a tool.\footnote{
-There is currently much 
-}
+you may then be in a position to infer that the unfamiliar pattern of actions she is performing constitute a means to open nuts.
+%\citep{Horner:2005pj}
+In a different case, knowing the goal of another agent's actions may enable you to discover a new use for a tool.
 A slightly less familiar idea is that goal ascription enables one to learn from others' failures as well as their successes.
@@
 }
-This is one illustration of how goal ascription could in principle enable us to learn from others' failures.
+This is one illustration of how goal ascription could in principle enable us to learn from others' failures 
+(see \citealp{Want:2001hp} offer another).
 
@@
 
-Despite the close connection between goal ascription and mindreading,
-goal ascription does not necessarily involve representing representations.
-To see why not we first need to be careful about the term `goal'.
-***
-[do the detour here]
-
-The fact that goal ascription does not involve metarepresentation raises the possibility that goal ascription is possible independently of mindreading.
-Consider an agent who has no communicative skills and no metarepresentational abilities.
-Nothing prevents her from \emph{representing} goals, of course.
-But could she actually ascribe goals?
-Is there any evidence which could be available to her and which would support goal ascriptions?
-Some philosophers have argued\footnote{*Bennett} or implied\footnote{*Davidson} that there could not be.
-On their view, the goal ascription and mindreading are interdependent in this sense: 
-there is no evidence for hypotheses narrowly about goals, only evidence for more complex hypotheses concerning both goals and mental states (such as beliefs).
-So, on this view, those without metarepresentational abilities are not in a position to know the goals of other agents' actions.
-
-***
-
-There is disagreement over whether it is possible to knowledgably identify the goals of an agent's actions without also ascribing some beliefs to that agent.  
+Some might argue that goal ascription is not mindreading because 
+identifying relations between actions and the outcomes to which they are directed
+does not necessarily involve ascribing mental states.
+Recognizing the force of this argument,
+we wish to
+duck the complex issues it raises by
+remaining neutral on whether or not goal ascription is actually mindreading.
+Either way,
+goal ascription is essential in nearly all mindreading.
+For mindreading typically involves or depends on predictions about how individuals will act,
+and those predictions can rarely be verified without the mindreader identifying to which goals the actions are directed.
+So even if goal ascription is not mindreading,
+our mission---which is to show that
+being poised to interact with others enables one to know things about their minds which one might not otherwise be in a position to know---%
+can be achieved by focusing on goal ascription.
+If we can show that
+  some evidence for goal ascription is only available to mindreaders who can interact with their targets
+then,
+	given the role of goal ascription in supporting mindreading,
+we can conclude that
+ some evidence which is indirectly relevant for mindreading is available only to mindreaders who can interact with their targets.
+
+There is potential for disagreement over whether it is possible to knowledgably identify the goals of an agent's actions without also ascribing some beliefs to that agent.  
 \citet[pp.48--50]{Bennett:1976rg}
@@
 \citet{Davidson:1974gh}
-both appear to hold that this is impossible,
-that identifying goals cannot be done independently of ascribing beliefs.
+both come close to suggesting that this is impossible,
+that ascribing goals cannot be done independently of ascribing beliefs.
+On this view (whether or not it is theirs), 
+goal ascription is interdependent with propositional attitude ascription in this sense: 
+there is no evidence for hypotheses narrowly about goals, only evidence for more complex hypotheses concerning both goals and mental states (such as beliefs).
 By contrast,

### 2011-12-27 joint action and knowing others minds/interacting mindreaders.2.tex
revised intro

@@
 
-Many people expect to know what those around them are thinking and doing, within limits.
-If someone were to complain of a co-worker that she can never tell what he is thinking or doing,
-the complaint might be informative, surprising even.
-And we might reasonably wonder where the fault lies.
-Maybe the co-worker really is a closed book,
-but maybe the complainant is not really trying to understand him.
-For humans and perhaps other species too,
-access to the actions and minds of those nearby,
-to some of their beliefs and desires,
-emotions, intentions and knowledge states,
-is often normal in this sense: it is lack of access, not access, that is remarkable.
-In this respect we might compare identifying others' thoughts and actions with identifying their ages or masses.
-Although all of these can be disguised with the right props and techniques, some degree of revelation is normal.
-
-The normalcy of mindreading,
-together with the fact that it often occurs spontaneously (if imperfectly),%
-%
-\footnote{
-On the spontaneity of mindreading in human adults see 
-\citet{senju_mindblind_2009} and 
-\citet{ferguson_eye_2011}.
-On imperfections in human adults' mindreading, 
-see \citet{Keysar:2003xu}.
-Both points come together in \citet{ferguson_listeners_2012}.
-}
-%
-make it easy to 
-overlook questions about its evidential basis.
-Indeed,
-while philosophers have engaged with questions about mechanisms (such as whether mindreading involves a process of simulation or of theorizing or some combination of the two),
-comparatively little effort has been devoted to issues about what evidence could ground mindreading.
-In this paper we pursue one such issue.
-
-%In everyday life it is sometimes so obvious what others are thinking and doing
-%that it is easy to forget that mindreaders begin life with no knowledge of the specific contents of particular individuals' thoughts 
-%and that to acquire this knowledge
-%there must be
-
+Mindreading is
+the process of identifying thoughts and actions on the basis of bodily movements,
+somewhat as reading is the process of identifying propositions on the basis of inscriptions \citep[p.\ 4]{Apperly:2010kx}.
+Contrast
+a mindreader who is, or appears to be, capable of interacting with her targets 
+and
+a mindreader who can manifestly only observe.
+Is it possible that the interacting mindreader is in a position to know things which she would be unable to know if she were unable to interact with her targets?
+Our aim in this paper is to argue that the answer is a qualified `yes'.
+
+The question is about the evidential basis of mindreading, 
+not about what sorts of mechanisms are involved.
+While philosophers have engaged with questions about mechanisms (such as whether mindreading involves a process of simulation or of theorizing or some combination of the two),
+comparatively little effort has recently been devoted to issues about what evidence could ground mindreading.
+
+Our question is part of a broader question,
 What is the evidential basis for ascriptions of thought and action and how does the evidence support the ascriptions?
-To stress that this question is not directly about about \emph{how} anyone actually ascribes thoughts or identifies actions% 
-% nor directly about the mechanisms, processes or representations involved in mindreading
-, Davidson 
-sometimes formulates the question by asking what someone \emph{could} know that would put them in a position to identify another's thoughts and actions.%
-\footnote{
-See, e.g., \citet[p.\ 126]{Davidson:1973jx}.
-While what follows draws  on Davidson's questions and insights,
-our aims here are more modest than his.
-For here we are concerned only with a  fraction of the problem of ascribing mental states and meanings,
-and we are not concerned with larger claims about mind, meaning and truth 
-(on which see 
-\citet{Davidson:1990du} and 
-\citet{lepore_donald_2005}).
-}
-
+%\footnote{
+%See e.g.\ \citet[pp.\ 126-7]{Davidson:1973jx}.
+%}
+%To stress that this question is not directly about about \emph{how} anyone actually ascribes thoughts or identifies actions% 
+%% nor directly about the mechanisms, processes or representations involved in mindreading
+%, Davidson 
+%sometimes formulates the question by asking what someone \emph{could} know that \emph{would} put them in a position to identify another's thoughts and actions.%
+%\footnote{
+%See, e.g., \citet[p.\ 126]{Davidson:1973jx}.
+%While what follows draws  on Davidson's questions and insights,
+%our aims here are more modest than his.
+%For here we are concerned only with a  fraction of the problem of ascribing mental states and meanings,
+%and we are not concerned with larger claims about mind, meaning and truth 
+%(on which see 
+%\citet{Davidson:1990du} and 
+%\citet{lepore_donald_2005}).
+%}
 The most sustained attempts to answer this question,
@@
 do not exploit the possibility of interaction.
-The evidential bases they consider and the principles which they identify to link the evidence with ascriptions of thought and action are available to entirely passive observers.
+The evidence and principles they consider are available to manifestly passive observers.
 So on their theories,
@@
 is on a par with
-a mindreader who does or could interact with those she seeks to interpret.
+a mindreader who, actually or apparently, could interact with those she seeks to interpret.
 The two are on a par in this sense:
 in principle the same evidence could be available to each, and each can exploit the same principles in moving from evidence to ascriptions of thought.
+Of course none of these theories is incompatible with the idea that interaction might be useful for mindreading in practice.  
+The point is just that, on these theories, interaction makes no difference to the theory.
 In this paper we aim to show
-that mindreaders  capable of interacting with their targets are at an advantage.
-Their ascriptions can rely on evidence and principles which would be unavailable if they were entirely passive observers.
-
-Being poised to interact with others enables one to know things about their minds which one might not otherwise be in a position to know, or so we aim to show in what follows.
-Because the leading theories have supposed that being able to interact makes no difference,
-they have treated the evidential basis for mindreading as more narrow than it truly is.
-This matters because broadening the evidential basis will enable us to give a more accurate and comprehensive theory of mindreading.
-And, as we shall argue, this in turn matters because it may bear on the possibility of explaining the emergence, in evolution or in development, of mindreading.
-Several researchers have offered quite general conjectures about how interaction might explain the emergence of sophisticated forms of cognition.%
+that mindreaders actually or apparently capable of interacting with their targets are at an advantage not only in practice but also in theory.
+Their ascriptions could rely on evidence and principles which would be unavailable if they were entirely passive observers.
+
+Why investigate this claim?
+One reason is simply because the investigation will enable us to revise and extend existing accounts of the evidential basis of mindreading  in ways that make them more accurate and comprehensive.
+Another motive concerns the emergence, in evolution and in development, of mindreading.
+Several researchers have offered quite general conjectures about how interaction might explain the emergence of sophisticated forms of cognition.
+One view is that needs to interact with others have driven and shaped some aspects of cognition.%
+\footnote{
+E.g.\ \citet[p.\ 103]{Knoblich:2006bn} suggest that 
+`functions traditionally considered hallmarks of individual cognition originated through the need to interact with others'
+and that
+`perception, action, and cognition are grounded in social interaction.'
+}
+We shall not speak to this view here.
+Another view is that abilities to interact with others
+may have fostered the emergence in evolution---or, on a related view, in development---%
+of sophisticated forms of cognition including social cognition and, in particular, mindreading.%
 \footnote{
@@
 argue for the `Vygotskian Intelligence Hypothesis' according to which `the unique aspects of human cognition ... were driven by, or even constituted by, social co-operation.'
-\citet[p.\ 103]{Knoblich:2006bn} offer a different conjecture:
-`perception, action, and cognition are grounded in social interaction  … functions traditionally considered hallmarks of individual cognition originated through the need to interact with others.'
-}
-Studying how interaction broadens the evidential basis of mindreading will eventually point to one way of filling in some details.
-
-%This, then, is our pitch.
-%To formulate (and, eventually, test) hypotheses about how mindreading develops,
-%or about how it evolves,
-%it is useful to understand not only which mechanisms mindreading involves
-%but also what evidence grounds mindreading.
-%And it since it turns out
+See also
+	\citet{Hughes:2004zj},
+	\citet{Hughes:2006fu},
+	\citet{Tomasello:2007gl} and
+	\citet{tomasello:2008origins}.
+}
+Studying how interaction broadens the evidential basis of mindreading will eventually point to one way of filling in some details on just how interaction might facilitate the emergence of mindreading.
+
+Our claim needs to be qualified in several ways.
+First, we should highlight something already explicit that has not yet been emphasised:
+the claim applies not just to 
+ mindreaders who actually could interact with their targets
+but also to
+ mindreaders who appear to their targets to be potential interaction partners (even if they are not).
+Trading the risk of a minor misunderstanding for concision we use the term `interacting mindreader' to refer to both groups;
+similarly, references to mindreaders who `only observe' must be understood as excluding those who appear to their targets to be interaction partners.
+Only cases involving the first group---mindreaders who actually could interact---are likely to be of wider interest.
+Cases involving misleading appearances might be regarded as artefacts of the model.
+Any application of our claim to understanding evolution or development is bound to focus on cases where mindreaders actually could interact with their targets.
+
+A second qualification concerns the scope of mindreading.
+In most discussions of mindreading, the focus is on  ascription of beliefs and other mental states to individuals.
+But in what follows we shall focus on the ascription of goals to actions.
+As we explain below, 
+some may deny that goal ascription is mindreading.
+Our view is consistent with this denial.
+Evidence for ascriptions of beliefs and other mental states includes the occurrence (and non-occurrence) of goal-directed actions,
+and identifying these typically requires goal ascription.
+In some cases, then, 
+evidence for the ascription of a goal
+will indirectly support the ascription of a belief or other mental state.
+Accordingly, even those who deny that goal ascription is mindreading should agree that sometimes evidence for a goal ascription is indirectly evidence for the ascription of a belief or other mental state.
+So whether or not goal ascription is deemed to be mindreading,
+a mindreader's access to evidence for mindreading  depends in part on her access to evidence for goal ascription.
+Our plan is to show that interacting mindreaders could make use, in ascribing goals, of evidence and principles which are not available to mere observers (who do not even appear to be in a position to interact with their targets).
+By showing this we will have shown that
+being (or appearing) poised to interact with others enables one to know things about their minds which one might not otherwise be in a position to know.
+
+%Because the leading theories have supposed that being able to interact makes no difference,
+%they have treated the evidential basis for mindreading as more narrow than it truly is.
+%This matters
+%because broadening the evidential basis will enable us to give a more accurate and comprehensive theory of mindreading.
+%And, as we shall argue, this in turn matters because it may bear on the possibility of explaining the emergence, in evolution or in development, of mindreading.
 
@@
 We shall focus on goal ascription rather than the ascription of beliefs and mental states that has captured most interest.
-Our immediate question is therefore, 
-What evidence could support hypotheses about the outcomes to which actions are directed, 
-and how would the evidence support the hypotheses?
-
-We focus on goal ascription 
-partly to simplify our argument,
+%Our immediate question is therefore, 
+%What evidence could support hypotheses about the outcomes to which actions are directed, 
+%and how would the evidence support the hypotheses?
+
+We narrow our focus to goal ascription 
+partly because this simplifies our argument,
 but mainly because goal ascription is widely thought to be among the very earliest components of mindreading to emerge (or, if goal ascription is not mindreading, then it is a late precursor).%
@@
 and 
-\citet{Povinelli:2001jf} on `behavioural regularities'.
+\citet{Povinelli:2001jf} on `behavioural regularities' whose specification sometimes appears to involve goal-directed action.
 }
@@
 to predict and manipulate others' actions.
-***if goal involved manipulating an object, you can predict where agent will go at some point; you can also manipulate the agents' action by revealling the location of the object.
-Eg.
-Liszowski pointing
-%This applies even to goal ascribers with no further mindreading abilities: 
-% even those unable to identify beliefs, desires and other propositional attitudes have much to gain from identifying the outcomes to which other agents' actions are directed.
+If you know that an agent is engaged in a sequence of actions whose eventual goal involves retrieving some object,
+you may be able to predict that the agent will go to where the object is at some point.\footnote{
+For a practical application of this prediction, see \citet{Hare:2001ph}.
+}
+Equally, in some cases knowing this much about the goal of another's actions many enable you to assist them,
+either 
+by retrieving the object for them \citep[e.g.][]{warneken:2007sa}
+or
+by revealing the object's location to them \citep[e.g.][]{Liszkowski:2008al}.
 
@@
 our mission---which is to show that
-being poised to interact with others enables one to know things about their minds which one might not otherwise be in a position to know---%
+being (or appearing) poised to interact with others enables one to know things about their minds which one might not otherwise be in a position to know---%
 can be achieved by focusing on goal ascription.
@@
 Of course opaque means are not in every case an insurmountable obstacle to goal ascription.
-Our point is pure commonsense and already widely appreciated:
+This point is neither novel nor surprising:
 opaque means sometimes  deprive interpreters of evidence and so prevent goal ascription.

### 2011-12-27 joint action and knowing others minds/interacting mindreaders.2.tex
working on section 2 goal ascription

@@
 \section{Mindreading: the normative project}
+\label{sec:intro}
 
@@
 But in what follows we shall focus on the ascription of goals to actions.
-As we explain below, 
-some may deny that goal ascription is mindreading.
-Our view is consistent with this denial.
-Evidence for ascriptions of beliefs and other mental states includes the occurrence (and non-occurrence) of goal-directed actions,
-and identifying these typically requires goal ascription.
+Some might claim that goal ascription is not mindreading, perhaps because 
+identifying relations between actions and the outcomes to which they are directed
+does not necessarily involve ascribing mental states.
+Our view is consistent with this claim (and with its negation).
+\label{goal_ascription_verify}
+Evidence for ascriptions of belief and of other mental states often includes the occurrence (and non-occurrence) of goal-directed actions.
+Identifying these actions typically requires goal ascription.
 In some cases, then, 
 evidence for the ascription of a goal
-will indirectly support the ascription of a belief or other mental state.
+will indirectly support the ascription of a belief or other mental state---%
+the evidence will support the ascription of a goal to an action,
+and this ascription will in turn support the ascription of a belief (say) to the agent of the action.
 Accordingly, even those who deny that goal ascription is mindreading should agree that sometimes evidence for a goal ascription is indirectly evidence for the ascription of a belief or other mental state.
@@
 Our plan is to show that interacting mindreaders could make use, in ascribing goals, of evidence and principles which are not available to mere observers (who do not even appear to be in a position to interact with their targets).
-By showing this we will have shown that
-being (or appearing) poised to interact with others enables one to know things about their minds which one might not otherwise be in a position to know.
+In doing this we will be showing that
+mindreaders who are (or appear) poised to interact with others 
+could know things about their minds 
+which they might not otherwise be in a position to know.
 
@@
 Purposive action is action directed to the realisation of one or more outcomes.
-Goal ascription is the process of identifying which outcomes others' purposive actions are directed to.
+Goal ascription is the process of identifying to which outcomes others' purposive actions are directed.
 To illustrate, suppose that
@@
 Asked about episode,
-Hannah might protest that the goal of her action was not to break the window but only to avoid losing a game.
+Hannah might protest that the goal of her action was not to break the window but only to reverse the others' advance.
 As this illustrates,
@@
 those to which it is directed.
-We shall focus on goal ascription rather than the ascription of beliefs and mental states that has captured most interest.
 %Our immediate question is therefore, 
@@
 
-We narrow our focus to goal ascription 
+We focus on goal ascription 
 partly because this simplifies our argument,
@@
 we may be able to indicate ways in which interaction could facilitate the emergence, in evolution or development, of more sophisticated forms of mindreading.%
+\footnote{
+There are various ways of understanding the idea that mindreading may come in several forms, some less conceptually or cognitively demanding than others.
+For a range of views, see
+ \citet{Apperly:2009ju},
+ \citet{Call:2005qe},
+ \citet{Doherty:2006wz},
+ \citet{ONeill:2005ff}
+ and
+ \citet{Wellman:2001if}.
+}
 
 Because goal ascription has received little attention,
-we shall briefly review some of the benefits that being able to identify others' goals brings.
+we shall briefly review some potential benefits of being able to identify goals.
 It is a familiar idea that goal ascription enables one to learn from others' successes.
@@
 This is one illustration of how goal ascription could in principle enable us to learn from others' failures 
-(see \citealp{Want:2001hp} offer another).
+(\citealp{Want:2001hp} offer another).
 
@@
 you may be able to predict that the agent will go to where the object is at some point.\footnote{
-For a practical application of this prediction, see \citet{Hare:2001ph}.
+For an application see \citet{Hare:2001ph}.
 }
@@
 either 
-by retrieving the object for them \citep[e.g.][]{warneken:2007sa}
-or
-by revealing the object's location to them \citep[e.g.][]{Liszkowski:2008al}.
-
-Goal ascription is also instrumental for mindreading:
+by retrieving the object for them 
+or else
+by revealing the object's location to them.%
+\footnote{
+For a paradigm involving the former see \citet{warneken:2007sa};
+on the latter, see \citet{Liszkowski:2008al}.
+}
+
+Goal ascription is also instrumental for ascribing propositional attitudes.
+We have already mentioned that such ascriptions often involve 
+confirming predictions about actions 
+which in turn typically involves
+identify the goals of those actions (\vpageref*{goal_ascription_verify}).
+In addition,
 knowing which outcomes an action is directed to may constrain hypotheses about what an agent intends 
@@
 
-Some might argue that goal ascription is not mindreading because 
-identifying relations between actions and the outcomes to which they are directed
-does not necessarily involve ascribing mental states.
-Recognizing the force of this argument,
-we wish to
-duck the complex issues it raises by
-remaining neutral on whether or not goal ascription is actually mindreading.
-Either way,
-goal ascription is essential in nearly all mindreading.
-For mindreading typically involves or depends on predictions about how individuals will act,
-and those predictions can rarely be verified without the mindreader identifying to which goals the actions are directed.
-So even if goal ascription is not mindreading,
-our mission---which is to show that
-being (or appearing) poised to interact with others enables one to know things about their minds which one might not otherwise be in a position to know---%
-can be achieved by focusing on goal ascription.
-If we can show that
-  some evidence for goal ascription is only available to mindreaders who can interact with their targets
-then,
-	given the role of goal ascription in supporting mindreading,
-we can conclude that
- some evidence which is indirectly relevant for mindreading is available only to mindreaders who can interact with their targets.
-
-There is potential for disagreement over whether it is possible to knowledgably identify the goals of an agent's actions without also ascribing some beliefs to that agent.  
-\citet[pp.48--50]{Bennett:1976rg}
-%\citet[pp.48--50]{Bennett:1976rg} suggests that a theory of goals ascription has to be developed together with a theory of (proto-)belief ascription: `An animal’s behaviour does not show what it registers unless we know what it seeks; but how can we learn what it seeks before we know what it registers?' [p.
-48]
-and 
-\citet{Davidson:1974gh}
-both come close to suggesting that this is impossible,
-that ascribing goals cannot be done independently of ascribing beliefs.
-On this view (whether or not it is theirs), 
-goal ascription is interdependent with propositional attitude ascription in this sense: 
-there is no evidence for hypotheses narrowly about goals, only evidence for more complex hypotheses concerning both goals and mental states (such as beliefs).
-By contrast,
-\citet{Gergely:1995sq},
-*Meltzoff?,
-\citet{Baillargeon:gx}
+Now that we have reviewed some of the benefits goal ascription can bring,
+what evidence grounds ascriptions of goals to actions?
+Consider the claim that 
+knowledgeably identifying the goals to which actions are directed
+depends on 
+knowledge of the agents' intentions
+which in turn depends 
+on knowledge of their beliefs, desires and other mental states.
+If this claim were true,
+there would be no such thing as a theory of the evidential basis of goal ascription,
+ only a theory of the evidential basis of mindreading more generally.
+
+The falsity of this claim seems to be assumed in developmental research on goal ascription.%
+\footnote{ 
+Compare \citet{Gergely:1995sq}
 and 
 \citet{Woodward:1998dm}
-(among many others)
-appear to assume the opposite, 
-that it is possible to identify goals without even being able to ascribe beliefs.
-For what it is worth, we tentatively favour this latter position.%
-\footnote{
-It is striking that, as far as we can tell, neither Bennett nor Davidson offers an argument for this claim.
-They do note that beliefs and goals make an interdependent contribution to observed action.
-But this by itself does not show that goal ascription cannot in some cases involve justifiably ignoring the possibility of differences in belief between interpreters and their targets.
+among many others.
+}
+We know of no compelling reason for taking this claim to be true.
+After all,
+it is consistent with rejecting the claim to recognize that
+ ascriptions of goals to actions
+ constrain, and are constrained by,
+ ascriptions of intentions and other mental states.
+The existence of such constraints does not show that one could 
+not know something about the goals of actions 
+while not knowing anything about the agents' mental states.
+Further,
+there do seem to be situations where knowledge of agents' goals does not require any knowledge of their mental states.
 For instance,
@@
 The objects are all out in the open; manifestly, both can clearly see them.
-If one reaches to grasp one of these objects (the duck, say), 
-must the other ascribe beliefs in order to knowledgeably identify the goal of her action?
+If one person reaches to grasp one of these objects (the duck, say), 
+must the other ascribe beliefs or other mental states in order to knowledgeably identify the goal of her action as that of grasping the duck?
 On the face of it, she need not.  
-Even if she had no ability to ascribe beliefs, she might nevertheless be in a position to acquire knowledge of the goal of the other's action.
-}
-However, this debate is not directly relevant to our concerns here.
-Both sides can agree that 
-differences in belief between observer and protagonist
-are sometimes an obstacle to goal ascription.
-This is all that the problem of false belief requires.
+Even if she had no ability to ascribe mental states, it seems she might nevertheless be in a position to acquire knowledge of the goal of the other's action.
+
+While not decisive,
+these considerations are perhaps sufficient to motivate 
+exploring what evidence might ground goal ascription.
+According to 
+what Csibra and Gergely call `the principle of rational action':
+%
+\begin{quote}
+`an action can be explained by a goal state if, and only if, it is seen as the most justifiable action towards that goal state that is available within the constraints of reality.'%
+\footnote{
+\citet[p.\ 255]{Csibra:1998cx}; cf.\ \citet{Csibra:2003jv}.
+A related but different `principle of efficiency' has been formulated by \citet[p.\ 1061]{Southgate:2008el}:
+`goal attribution requires that agents expend the least possible amount of energy within their motor constraints to achieve a certain end.'
+}
+\end{quote}
+%
+Taking this as a rough starting point,
+we propose that
+goal ascription can involve 
+identifying which potentially desirable or beneficial outcomes 
+an observed action is the best available means to achieving.
+More carefully, our proposal is that:
+%
+\begin{idescription}
+\item[(E$_1$)] actions of type $a$ are normally a means of realising outcome $G$ in situations with the salient features of this situation;
+\end{idescription}
+%
+and:
+%
+\begin{idescription}
+\item[(E$_{2}$)] no alternative type of action, $a'$, is available to the agent and such that actions of this type would be normally be better\footnotemark \ means of realising outcome $G$ in situations which have the salient features of this situation;
+\footnotetext{
+An action of type $a'$ is a better means of realising outcome $G$ in a given situation than an action of type $a$ if, for instance, actions of type $a'$ normally involve less effort than actions of type $a$
+in situations with the salient features of this situation 
+and everything else is equal; 
+or if, for example, actions of type $a'$ are normally more likely to realise outcome $G$ than actions of type $a$
+in situations with the salient features of this situation 
+and everything else is equal.
+}
+\end{idescription}
+%
+and:
+%
+\begin{idescription}
+\item[(E$_3$)] the occurrence of outcome $G$ is potentially beneficial for the agent or typically desirable to agents of this type,
+and there is no other outcome, $G'$, which would be more beneficial to the agent or typically more desirable to agents of this type
+where  (E$_1$) and (E$_2$) both hold of $G'$ and $a$
+\end{idescription}
+%
+jointly constitute defeasible evidence for the conclusion that:
+%
+\begin{idescription}
+\item[(C)] $G$ is a goal to which action $a$ is directed.
+\end{idescription}
+%
+We further suggest that the above inference from (E$_1$)-(E$_3$) to (C) is a route to knowledge of the goals of actions in this sense:
+it would be possible to know the premises without already knowing the conclusion
+and 
+in some cases knowing the premises might put one in a position to know the conclusion.
+
+This proposal might be extended in various ways.
+
+Fact that this route to knowledge exists shows that goal ascription doesn't depend on mental state ascription.
+
+Don't suggest this is the only route to knowledge.
+Not claiming to be offering a comprehensive theory of the evidential basis of goal ascription.
 
@@
 
-\section{OLD}
-
-These simple facts about goal ascription raise many questions.
-Some concern mechanism, how in fact one subject is able to discover facts about which outcomes another agent's actions are directed to.
-Another set of questions focuses on the evolution of goal ascription and the costs and benefits of being able to ascribe goals and of being a potential target of goal ascription.
-Our concern here is not directly with any of these questions.
-Instead we shall focus on a more narrowly epistemic question.
-What evidence could support hypotheses about the outcomes to which actions are directed?
-And how would the evidence support the hypotheses?%
-\footnote{
-These questions are versions of those Davidson constructs a theory of interpretation to answer.
-While what follows draws  on Davidson's insights,
-our aims here are more modest than his.
-For we are concerned only with a  fraction of the problem of ascribing mental states and meanings;
-and, unlike Davidson, we are not concerned with larger claims about the nature of mind.
-See
-\citet{Davidson:1973jx,Davidson:1990du,lepore_donald_2005}.
-}
-
-Of special interest is evidence available independently of any knowledge of mind or language.
-We want to know how it is possible to identify goals even without knowing what an agent believes or desires and even without understanding their communicative actions.
-Accordingly we will adopt the perspective of a goal ascriber who knows nothing about the mental states of her target agent that would distinguish this agent from any other.
-We will also stipulate that there is initially no common ground, shared culture or conventions.
-And we will stipulate that the goal ascriber is initially unable to understand any communicative actions.
-
-There are two sorts of motivation for these restriction on the evidential basis.
-One is simply that developmental and comparative research indicates that goal ascription does appear to take place  in such circumstances.%
-\footnote{
-*refs
-}
-This makes it important to understand the evidence on which such ascriptions could be based.
-(Of course 
-identifying evidence that could support such ascriptions 
-would not all by itself enable us to explain how goal ascriptions are actually made, 
-but identifying evidence is necessary if we are ever to explain the reliable success of mechanisms for goal ascription.)
-Another source of motivation is the conjecture that goal ascription is a prerequisite for the more sophisticated mindreading activities which reveal mental states and meanings.
-The coherence of this conjecture depends on the possibility of knowing something about which outcomes an agent's actions are directed to independently of knowing what she believes or desires and independently of understanding her communicative actions.%
-\footnote{
-*Compare and contrast Davidson?
-He did think relational attitudes (holding true) are the foundation for interpretation.
-But he also thought that interpretation had to happen all at once.)
-
-\citet[p.\ 17]{Dennett:1987sf} `Here is how it works: first you decide to treat the object whose behavior is to be predicted as a rational agent; then you figure out what beliefs that agent ought to have, given its place in the world and its purpose.
-Then you figure out what desires it ought to have, on the same considerations, and finally your predict that this rational agent will act to further its goals in the light of its beliefs.
-A little practical reasoning from the chosen set of beliefs and desires will in many—but not in all—instances yield a decision about what the agent ought to do; that is what you predict the agent will do.'
-}
-
-So what evidence could support goal ascription by someone who knows nothing discriminating about her targets' mental states or communicative actions?
-Ordinary third-person goal ascription, simplified and idealized, works like this.%
-\footnote{
-*ref? Dennett?
-}
-Faced with an action,
-the would-be goal ascriber first asks which outcomes this action could be a means to realising.
-She then considers which of these outcomes are potentially beneficial for, or desirable to, the agent.
-Any such outcomes are identified as goals to which the action is directed.
-So the fact that an action is a means to realising some outcome which is potentially beneficial or desirable is evidence for the conclusion that this outcome is one to which the action is directed.
-Schematically, the proposal is that:
-%
-\begin{idescription}
-\item[(E$_1$)] Action $a$ is a means of realising outcome $G$.
-\end{idescription}
-%
-and:
-%
-\begin{idescription}
-\item[(E$_2$)] The occurrence of outcome $G$ is potentially beneficial for, or desirable to, the agent of $a$.
-(And there is no other outcome, $G'$, which action $a$ is a means of realising and which would be more beneficial for, or more desirable to, the agent of $a$.)
-\end{idescription}
-%
-jointly constitute evidence for the conclusion that:
-%
-\begin{idescription}
-\item[(C)] $G$ is a goal to which action $a$ is directed.
-\end{idescription}
-
-This proposal might be extended in various ways.
-For instance, Southgate, Johnson and Csibra offer a `principle of efficiency' according to which:
-%
-\begin{quote}
-`goal attribution requires that agents expend the least possible amount of energy within their motor constraints to achieve a certain end.'
-\citep[p.\ *]{Southgate:2008el}
-\end{quote}
-%
-If this is a correct principle of goal attribution, we could extend the proposal above to incorporate it:
-%
-\begin{idescription}
-\item[(E$_3$)] No alternative action, $a'$, is a means to realising outcome $G$ and would involve expending less energy than $a$.
-\end{idescription}
-%
-Now the proposal is that (E$_1$) to (E$_3$) are jointly evidence for (C).
-
-In at least some cases goal attribution is likely to be more complicated than this proposal allows.
-To illustrate, note that some agents may weigh the efficiency of alternative actions against their possible side effects and how reliable they would be as a means to realising an outcome.
-Where this is true, 
-identifying the evidential basis for goal ascription may require  a similar weighing of these factors in inferring backwards from actions to their goals.%
-\footnote{
-This is loosely related to what Csibra and Gergely call `the principle of rational action'.
-As they formulate the principle,
-`an action can be explained by a goal state if, and only if, it is seen as the most justifiable action towards that goal state that is available within the constraints of reality'
-(\citealp[p.\ *]{Csibra:1998cx}; cf.\ \citealp{Csibra:2003jv}).
-}
-Specifying exactly what should be weighed and how is beyond the scope of this paper,
-(and may also be something which varies between species of agent).
-We can mark the gap with an alternative to (E$_3$) which uses an unspecified notion of `better' as a placeholder:
-%
-\begin{idescription}
-\item[(E$_{3'}$)] No alternative action, $a'$, is a better means to realising outcome $G$.
-\end{idescription}
-%
-This, then, is the standard approach to answering our question about goal attribution: (E$_1$), (E$_2$) and (E$_{3'}$) jointly constitute evidence for (C) given that these approximate conditions under which it would be rational to perform $a$ in order to realise $G$ and given that agents approximate to performing $a$ in order to realise $G$ rather than any other outcome under these conditions.
-
 \bibliography{$HOME/endnote/phd_biblio}

### 2011-12-27 joint action and knowing others minds/interacting mindreaders.2.tex
second section drafted, also some work on conclusion (incomplete)

@@
 Because goal ascription has received little attention,
-we shall briefly review some potential benefits of being able to identify goals.
+in this section
+we shall 
+briefly review some potential benefits of being able to identify goals
+and then 
+consider what sort of evidence might support goal ascription.
+
 It is a familiar idea that goal ascription enables one to learn from others' successes.
@@
 Now that we have reviewed some of the benefits goal ascription can bring,
-what evidence grounds ascriptions of goals to actions?
+what evidence could support ascriptions of goals to actions?
 Consider the claim that 
@@
 
-The falsity of this claim seems to be assumed in developmental research on goal ascription.%
-\footnote{ 
-Compare \citet{Gergely:1995sq}
-and 
-\citet{Woodward:1998dm}
+The falsity of this claim seems to be assumed in both developmental and comparative research on goal ascription.%
+\footnote{ 
+Compare \citet{Gergely:1995sq},
+	\citet{Woodward:1998dm} and
+	\citet{Penn:2007ey}
 among many others.
@@
 not know something about the goals of actions 
-while not knowing anything about the agents' mental states.
+while  knowing nothing about the agents' mental states.
 Further,
@@
 On the face of it, she need not.  
-Even if she had no ability to ascribe mental states, it seems she might nevertheless be in a position to acquire knowledge of the goal of the other's action.
+Even if she had no ability to ascribe mental states, it seems she might nevertheless be in a position to identify the goal of the other's action.
 
@@
 %
-Taking this as a rough starting point,
+Taking this idea as a rough starting point,
 we propose that
-goal ascription can involve 
-identifying which potentially desirable or beneficial outcomes 
-an observed action is the best available means to achieving.
-More carefully, our proposal is that:
-%
-\begin{idescription}
-\item[(E$_1$)] actions of type $a$ are normally a means of realising outcome $G$ in situations with the salient features of this situation;
-\end{idescription}
-%
-and:
-%
-\begin{idescription}
-\item[(E$_{2}$)] no alternative type of action, $a'$, is available to the agent and such that actions of this type would be normally be better\footnotemark \ means of realising outcome $G$ in situations which have the salient features of this situation;
+%goal ascription could sometimes involve 
+%identifying which potentially desirable outcomes 
+%an observed action is the best available means to achieving.
+%More carefully, our proposal is that 
+these facts:
+%
+\begin{enumerate}
+%
+\item action $a$ is directed to some goal;
+%
+\item actions of $a$'s type are normally capable of being means of realising outcomes of $G$'s type in situations with the salient (to any concerned) features of this situation;
+% 
+\item no alternative type of action is both 
+typically available to agents of this type 
+and also 
+such that actions of this type would be normally be significantly better\footnotemark \ means of realising outcome $G$ in situations with the salient features of this situation;
 \footnotetext{
@@
 }
-\end{idescription}
-%
-and:
-%
-\begin{idescription}
-\item[(E$_3$)] the occurrence of outcome $G$ is potentially beneficial for the agent or typically desirable to agents of this type,
-and there is no other outcome, $G'$, which would be more beneficial to the agent or typically more desirable to agents of this type
-where  (E$_1$) and (E$_2$) both hold of $G'$ and $a$
-\end{idescription}
-%
-jointly constitute defeasible evidence for the conclusion that:
-%
-\begin{idescription}
-\item[(C)] $G$ is a goal to which action $a$ is directed.
-\end{idescription}
-%
-We further suggest that the above inference from (E$_1$)-(E$_3$) to (C) is a route to knowledge of the goals of actions in this sense:
-it would be possible to know the premises without already knowing the conclusion
+%
+\item the occurrence of outcome $G$ is typically desirable for agents of this type;
+%
+\end{enumerate}
+%
 and
-in some cases knowing the premises might put one in a position to know the conclusion.
-
-This proposal might be extended in various ways.
-
-Fact that this route to knowledge exists shows that goal ascription doesn't depend on mental state ascription.
-
-Don't suggest this is the only route to knowledge.
-Not claiming to be offering a comprehensive theory of the evidential basis of goal ascription.
+%
+\begin{enumerate}[resume]
+\item there is no other outcome, $G'$, 
+the occurrence of which would be at least comparably desirable for agents of this type 
+and where (2) and (3) both hold of $G'$ and $a$
+%
+%where
+%	the occurrence of $G'$ would be at least comparably desirable for agents of this type 
+%%where
+%%	$G$'s occurrence is not a means to $G'$'s nor conversely,
+%and
+%where  
+%	(2) and (3) both hold of $G'$ and $a$
+\end{enumerate}
+%
+may jointly constitute defeasible evidence for the conclusion that:
+%
+\begin{enumerate}[resume]
+\item $G$ is a goal to which action $a$ is directed.
+\end{enumerate}
+%
+We suggest that the above inference, from (1)-(5) to (6), is a route to knowledge of the goals of actions in this sense:
+it would be possible to know the premises without already knowing the conclusion;
+and, 
+in some cases knowing the premises could put one in a position to know the conclusion.%
+\footnote{
+Knowledge of the premises may not require explicit knowledge of what others find obvious, what is salient to them or what makes actions better for them.
+It is arguably sufficient that
+the individual ascribing a goal 
+is entitled to 
+rely on being sufficiently similar to the target of her ascription 
+with respect to 
+	what she finds obvious, 
+	what makes one action a better or worse means of realising an outcome than another action for her,  
+	what is desirable to her,
+	and what is salient to her.
+}
+
+Why accept this?
+Suppose that 
+	(1)-(5) are facts 
+and that
+	$G$ is not a goal to which action $a$ is directed.
+Then from (1) we know that $a$ has a goal, call it $G'$.
+And from (5) we can infer that either 
+(i) 
+the action, $a$, was not of a type that is normally capable 
+of being a means to the outcome to which it was directed,
+or
+(ii) 
+a significantly better (for instance, a more reliable or less effortful) means of achieving $G'$ is typically available to agents of this type,
+or
+(iii)
+$G'$'s occurrence would be significantly less desirable than $G$'s occurrence for agents of this type.
+So the agent's action fell short of being appropriate or optimal in some way,
+or the agent has atypical preferences or capabilities,
+or the situation is not normal.
+While any of these is possible---%
+and perhaps in some cases even to be expected---%
+there are also circumstances in which it is reasonable to suppose that none obtains,
+that the situation will be normal in relevant respects,
+that agents will have typical preferences and capabilities 
+and that they will act in ways that are appropriate and (in the limited sense in play here) optimal.
+In such circumstances, the above inference can serve as a route to knowledge of the goals of actions.
+The existence of this route to knowledge shows that goal ascription doesn't invariably depend on mental state ascription.
+
+Note that we are not suggesting that the above inference captures the only route to knowledge of the goals of actions,
+or that it provides anything like a comprehensive theory of the evidential basis of goal ascription.
+This route to knowledge has only limited applications.
+For example, it cannot be used when an action is the best available means of achieving two or more comparably desirable outcomes.
+The point of this brief discussion was only to 
+consider the sort of evidence that can bear on goal ascrpition
+and to 
+show that 
+even those incapable of ascribing mental states like belief 
+might nevertheless have evidence sufficient for goal ascription.
+
+Our narrow aim in what follows is to show that some routes to knowledge of the goals of actions 
+are available only to interacting mindreaders
+and not to those who merely observe.
+Having in this section introduced goal ascription as a topic,
+the next step is to identify an obstacle to acquiring knowledge of the goals of actions,
+one which we will eventually show can be avoided where mindreaders can (or appear to be able to) interact with their targets.
 
@@
 
-***Return to the leading theories?
-We do not take this to show that the leading theories must be entirely discarded.
-If we are right about interaction making evidence available, 
-either leading theory can be extended accordingly.
-***
+%***Return to the leading theories?
+%We do not take this to show that the leading theories must be entirely discarded.
+%If we are right about interaction making evidence available, 
+%either leading theory can be extended accordingly.
+%***
 
@@
   communicative gestures
-does not presuppose rich insights into others' intentions and beliefs.
+does not presuppose knowledge of others' intentions and beliefs.
 Such expertise,
@@
 
+How might this work?
+***sketch the story***
+Clearly this is a wild conjecture,
+and not one that we have given any reason for accepting.
+Our point in mentioning it is merely to illustrate how
+ better understanding the role of interaction in goal ascription and mindreading generally
+may eventually lead to discoveries
+on how
+ sophisticated forms of cognition
+ emerge from
+ conceptually and cognitively undemanding forms of interaction.
+If the culture of tool use,
+the simpler forms of communicative gesture
+or the sharing of knowledge and other mental states
+originates in abilities to interact,
+then we need to know how interaction could lead to the emergence of these things.
+Our  suggestion has been that
+understanding how interacting mindreaders 
+could know things about their minds which mere observers might not be in a position to know
+will provide some of the details.
+
 \bibliography{$HOME/endnote/phd_biblio}

### 2011-12-28 joint action and knowing others minds/interacting mindreaders.2.tex
section 4 on distributive goals drafted, some work on section 5 started

@@
 or that it provides anything like a comprehensive theory of the evidential basis of goal ascription.
-This route to knowledge has only limited applications.
+Clearly it does not.
+The inference has only limited applications.
 For example, it cannot be used when an action is the best available means of achieving two or more comparably desirable outcomes. 
+Fortunately the argument that follows does not depend on having a comprehensive theory of evidence for goal ascription.
 The point of this brief discussion was only to 
-consider the sort of evidence that can bear on goal ascrpition
+consider the sort of evidence that can bear on goal ascription
 and to 
@@
 They are less powerful in this sense: 
-some routes to knowledge of the goals of actions are available only where a model of interpretation takes the possibility of interaction into account.
-In what follows we shall suggest that abilities to engage in joint action with others provides a route to knowledge of the goals of other agents' actions  which avoids the problem of opaque means.
-This is one way in which moving away from a purely observational model of interpretation yields a richer evidential base for ascriptions.
-
-\section{Joint action}
+some routes to knowledge of the goals of actions are available only where a theory of the evidential basis of mindreading 
+ takes the possibility of interaction into account.
+In what follows we shall explain how abilities to interact with others provide a route to knowledge of the goals of other agents' actions  which avoids the problem of opaque means.
+This is one way in which moving away from a purely observational model of mindreading yields a richer evidential base for ascriptions.
+
+\section{Interactions involving distributive goals}
 \label{sec:joint_action}
@@
 which are closed to
-interpreters who merely observe their targets
+mindreaders who merely observe their targets
 but open to 
-interpreters capable of interacting with their targets.
+mindreaders who are, or appear to be, capable of interacting with their targets.
 As a preliminary to defending this claim,
-we need to specify the types of interaction which are relevant.
-We shall focus on joint action.
-
-***Much disagreement about what it is.
-Necessary condition is distributive goal.
-If people want to disagree, this is only terminological.
+we need to specify which types of interaction are relevant.
+That is the aim of this section.
+
+Let us stipulate that an outcome is a \emph{distributive goal} of two or more agents' actions just if two conditions are met.
+First, this outcome is a one to which each agent's actions are individually directed.
+Second, each agent's actions are related to the outcome in such a way that it is possible for all the agents (not just any agent, all of them together) to succeed in bringing about this outcome.
+
+To illustrate,
+suppose that, while doing some gardening,
+we find a turnip too big for any of us easily pull out of the ground alone.
+We each individually intend that we all pull up the enormous turnip
+and we act on this intention.%
+\footnote{
+For an argument that it is possible for each of us individually to intend that we do something, see \citet{Bratman:1999fr} and \citet[Chapter 2]{Bratman:2012fk}.
+}
+There is a single outcome,
+freeing the turnip,
+to which each of our actions is individually directed.
+And it is possible for all of us to succeed in bringing about this outcome.
+So this outcome is a distributive goal of our actions.
+
+%Contrast this episode with an alternative where no distributive goal is involved.
+%Instead of intending that we all pull up the enormous turnip,
+%each of us intends that she herself pull it up.
+%While all of our actions are directed to similar outcomes (they all involve pulling up the turnip),
+%these are clearly different outcomes.
+%They are clearly different because one could occur while another did not;
+%one of us might have broken an arm and been unable to pullwhile the others succeeded.
+%So in this case, freeing the turnip is not a distributive goal of our actions.
+%For some actions to have a distributive goal,
+%it is necessary that there be a single outcome (and not merely similar outcomes) to which they are directed.
+
+In defining `distributive goal' we stipulated that it must be possible for all of the agents of actions with a distributive goal to succeed relative to that outcome.
+Consider a variation on our first turnip-pulling story.
+Instead of each intending that we pull up the turnip,
+we each simply intend  to pull up the turnip.
+In this case,
+some might claim that
+this amounts to us each intending that he or she pulls up the turnip.
+And this claim might be taken to support the further claim that
+there is no outcome to which each of our actions are directed
+in virtue of our having these intentions
+such that we could all succeed relative to this outcome.
+%, given our intentions in this alternative case, we could not all succeed relative to the outcome of pulling up the turnip---%
+%that if your actions were to succeed relative to the turnip's extraction,
+%then ours would fail.
+We take no stand on whether either claim is correct.
+For our purposes all that matters is that there are distributive goals.
+And the existence of distributive goals follows from the possibility of each of us 
+intending that we pull up the turnip
+(rather than simply
+intending to pull it up).
+
+Note that our actions could have a distributive goal even without us each intending that we do something.
+As another variation on the turnip-pulling story,
+suppose that we each individually intend to 
+	contribute to pulling up the turnip or pull it up.
+Such intentions are agent neutral in this sense:
+relative to such an intention an agent 
+	could succeed by acting alone 
+and
+  she also could succeed by acting with unspecified other agents.
+Plausibly our having these intentions is sufficient for our actions to have a distributive goal.
+For in virtue of these intentions each of our actions is directed to the turnip's extraction,
+and this is an outcome relative to which it is possible that all of our actions succeed.
+
+For two agents' actions to have a distributive goal it is sufficient that their actions constitute a joint action 
+(at least this is true on almost any account of joint action).%
+\footnote{
+We know of two definitions of joint action on which it is not straightforward that all joint actions involve distributive goals.
+One is provided 
+ is provided by \citet[p.\ 366]{ludwig_collective_2007} and discussed in footnote \vref{fn:ludwig}; the other is due to 
+\citet[p.\ 70]{Sebanz:2006yq}.
+According to Sebanz and colleagues' `working definition', 
+`joint action can be regarded as any form of social interaction whereby two or more individuals coordinate their actions in space and time to bring about a change in the environment'
+(\citeyear[p.\ 70]{Sebanz:2006yq}).
+On one way of reading this, the latter part is equivalent to `a change in the environment is such that: to bring it about two or more individuals coordinate their actions';
+ this does suggest that the actions have a distributive goal.
+On other standard accounts it is straightforward that joint action involves distributive goals.
+Compare 
+	\citet[pp.\ 329-31]{Bratman:1992mi},
+%	\citet[p.\ 40]{Butterfill:2011fk},
+	\citet[pp.\ 96-7]{Searle:1990em} and
+	\citet[pp.\ 168-9]{gilbert:2009shared}.
+}
+However, the converse does not hold.
+As this the turnip-pulling variation  indicates,
+two or more agents' actions may have a distributive goal 
+even though the agents do not know about each others' intentions or actions.
+In fact our actions might have a distributive goal even though none of us is aware of this, or even of the others' existence. 
+(We shall refrain from illustrating this point by contriving a truly gigantic turnip and a completely dark, very windy night, but we are sure readers will be able to imagine how this would go.)
+One consequence of this is that 
+two or more agents' actions may have a distributive goal
+even though they are not engaged in joint action
+(at least not on any standard account of joint action).%
+\footnote{
+\label{fn:ludwig}
+According to 
+	 \citet[p.\ 366]{ludwig_collective_2007}, 
+`The concept of a joint action as such is just that of an event of which there are multiple agents.'
+Depending on what events are and what it is to be the agent of an event,
+it may turn out that,
+on Ludwig's definition,
+our actions' having a distributive goal 
+is sufficient for
+us to be engaged in joint action.
+However, other accounts of joint action have the consequence that each agent involved in a joint action must believe or know something about the jointness of her action.
+See 
+	\citet[p.\ 103]{Bratman:1993je}, %common knowledge
+	\citet[p.\ 40]{Butterfill:2011fk},
+	\citet[p.\ 10]{Kutz:2000si}, %Agents acting on a shared intention have `a conception of themselves as contributors to a collective end.'
+	\citet[p. 56]{miller_social_2001} % requires that each agent believes her actions are interdependent with the other agent's.
+	and
+	\citet[p.\ 361]{Roth:2004ki}.
+%: `each participant ...
+can answer the question of what he is doing or will be doing by saying for example ``We are walking together'' or ``We will/intend to walk together.'''
+}
+
+The notion of a distributive goal is a 
+narrowly technical one.
+In introducing this notion by stipulation we are not aiming to match anyone's intentions about anything.
+What matters for now is just that the definition is theoretically coherent and that it is possible for actions to have distributive goals.
+Our aim in introducing distributive goals 
+ is to capture a minimal requirement on the sort of interactions 
+relevant to characterising a route to knowledge of the goals of actions
+available only to interacting mindreaders.
+% 
+% relevant to explaining how 
+%interacting mindreaders 
+%might be able to exploit routes to knowledge of the goals of actions
+%which would be unavailable  were  they mere observers.
 
 \section{Your-goal-is-my-goal: a route to knowledge}
-If an interpreter is able to interact with her targets,
+If a mindreader is able to interact with her targets,
 if she is not limited to merely observing them,
 how might this provide her with a route to knowledge of the goals of their actions?
-The intuitive idea we started with hinges on joint action, a particular form of interaction in which the actions of two or more agents are directed to a single goal (see section \vref{sec:joint_action}).
+The intuitive idea we started with hinges on interactions involving distributive goals
+(see \vref{sec:joint_action}).
 Our intuitive idea was this:
-if an interpreter is engaged in joint action with her target, 
-it's easy for the interpreter to know what the goal of her target's actions is because this goal is the goal of her own actions.
-So if she knows the goal of her own actions and she knows that she is engaged in joint action with her target,
+if a mindreader is engaged in an interaction with her target that involves a distributive goal, 
+it may be easy for the interpreter to know what the goal of her target's actions is because this goal is the goal of her own actions.
+So if she knows the goal of her own actions and she knows that she is engaged in an interaction involving a distributive goal  with her target,
 then she already knows what the goal of her target's actions are.
@@
 Of course this intuitive idea won't work as it stands.  
-For the inference it captures relies on the premise that the interpreter and her target are engaged in joint action.
-But for the interpreter to know this premise---for her to know that they are engaged in joint action---it seems she must   already know which goal her target's actions are directed to.
-Worse,
-on many accounts of joint action the mere truth of the premise  would require the interpreter to know the goal of her target's actions.%
-\footnote{
-*ref}
-Apparently, then, engaging in joint action presupposes, and therefore cannot be a source of, knowledge of others' goals.
+For the inference it captures relies on the premise that the mindreader and her target are engaged in actions with a distributive goal.
+But for the mindreader to know this premise it seems she must   already know which goal her target's actions are directed to.
 
 Fortunately there is a way around this.  
-For there are various cues which signal that one agent is prepared to engage in joint action with another.  
+For there are various cues which signal that one agent is prepared to engage in joint action with another,
+and joint actions involve distributive goals.  
 \label{twin_pram}
@@
 Since the stranger is fully committed to getting his pram onto the bus,
-he knows what the goals of his own actions will be.
+he knows what the sole goal of his own actions will be.
+But he also knows that you will engage in joint action with him,
+which means that your actions and his actions will have a distributive goal.
 This enables him to infer the goal of your imminent actions: 
@@
 \item You are willing to engage in some joint action\footnote{
-*What notion of joint action is needed here?  Any will do as long as it involves distributive goals.
+We leave open the issue of how joint action is to be characterised subject only to the 
+requirement that all joint actions must involve distributive goals.
+Attempts to characterise joint action in ways relevant to explaining development include
+	\citet{Tollefsen:2005vh},
+	\citet{pacherie_framing_2011} and
+	\citet{Butterfill:2011fk}.
 }

### 2011-12-29 joint action and knowing others minds/interacting mindreaders.2.tex
getting there

@@
 
-Why investigate this claim?
-One reason is simply because the investigation will enable us to revise and extend existing accounts of the evidential basis of mindreading  in ways that make them more accurate and comprehensive.
+Why suppose, in advance of considering the details, that interacting mindreaders might know more?
+A mindreader's target is often also a mindreader and may sometimes reciprocate by taking the mindreader as a target for mindreading.
+%Where this happens we have mindreaders mindreading and being mindread by mindreaders.
+%The mindreader is her target's target for mindreading.
+It ought to be possible, in mindreading, to make use of this reciprocity.
+%of the fact that the mindreader's target reciprocates by taking the mindreader to be a target for mindreading.
+But how could such reciprocity facilitate mindreading?
+If we assume the mindreader is merely observing her target,
+that there is no potential for interaction,
+then it seems that any way of exploiting reciprocity would involve higher-order ascriptions. 
+The mindreader would ascribe to her target beliefs (say) about the mindreader's own beliefs and other mental states.
+And if her target reciprocates, she might escalate by ascribing to the target beliefs about her own beliefs about the target's beliefs about her beliefs.
+While this might be useful in some situations, 
+the nesting this approach requires quickly becomes dauntingly complex.
+And the basic intuition about reciprocal mindreading goes unsatisfied.
+Reciprocal mindreading should sometimes result in 
+something like a meeting of minds
+rather than
+an escalation of higher-order ascriptions.
+Perhaps, then, fully exploiting reciprocity in mindreading
+requires the mindreaders to be in a position to interact with each other.
+Perhaps in some cases 
+being or appearing poised to interact
+can somehow enable mindreaders to exploit reciprocity without first having to ascribe of higher-order mental states.
+This is the hunch we develop in what follows.
+
+Even if the hunch turns out to be right, why investigate it?
+One reason is that the investigation will enable us to revise and extend existing accounts of the evidential basis of mindreading  in ways that make them more accurate and comprehensive.
 Another motive concerns the emergence, in evolution and in development, of mindreading.
@@
 Our claim needs to be qualified in several ways.
-First, we should highlight something already explicit that has not yet been emphasised:
+First, we should highlight something already explicit but not yet emphasised:
 the claim applies not just to 
@@
 %And, as we shall argue, this in turn matters because it may bear on the possibility of explaining the emergence, in evolution or in development, of mindreading.
+
+Developing these ideas requires us to fill in some background on goal ascription and its limits, as well as on notions of goal-directed interactions.
+Readers impatient to get to the central idea should skip to section \vref{sec:your_goal_is_my_goal}.
 
@@
 
+%*** remove this line: to avoid annoying error only
+\
+
 \section{Your-goal-is-my-goal: a route to knowledge}
+\label{sec:your_goal_is_my_goal}
 If a mindreader is able to interact with her targets,
@@
 how might this provide her with a route to knowledge of the goals of their actions?
-The intuitive idea we started with hinges on interactions involving distributive goals
+The answer hinges on interactions involving distributive goals
 (see \vref{sec:joint_action}).
-Our intuitive idea was this:
+
+Here is an intuitive idea that doesn't quite work:
 if a mindreader is engaged in an interaction with her target that involves a distributive goal, 
 it may be easy for the interpreter to know what the goal of her target's actions is because this goal is the goal of her own actions.
-So if she knows the goal of her own actions and she knows that she is engaged in an interaction involving a distributive goal  with her target,
+So if she knows the goal of her own actions and she knows that she is engaged with her target in an interaction involving a distributive goal,
 then she already knows what the goal of her target's actions are.
@@
 	\citet{Tollefsen:2005vh},
+	\citet{Carpenter:2009wq},
 	\citet{pacherie_framing_2011} and
 	\citet{Butterfill:2011fk}.
+The last of these 
+not only shows how a primitive form of joint action can be characterised without ascriptions of higher-order mental states
+but also shares our focus on relations between actions and goals rather than between agents and intentions.
 }
@@
 
-\item I am not about to change which goal my actions will be directed to.
+\item I am not about to change the single goal to which my actions will be directed.
 
@@
 %
-Call this the \emph{your-goal-is-my-goal inference}.  
-To say that this inference characterises a route to knowledge implies two things.  
+Call this inference \emph{your-goal-is-my-goal}.  
+To say that it characterises a route to knowledge implies two things.  
 First, in some cases it is possible to know the premises, 1–2, without already knowing the conclusion, 3.  
@@
 in the right contexts
-an individual can recognize in another's facial expressions, engaging gestures or synchronized bodily movements that she are about to engage in joint action with him.
-Exploiting these indicators does not typically depend on knowing the particular contents of any of their beliefs, desires or goals.  
+an individual can recognize in another's facial expressions, engaging gestures or synchronized bodily movements that she is about to engage in joint action with him.
+Exploiting these indicators does not typically depend on knowing the particular contents of any beliefs, desires or goals.  
 Expressions, gestures and movements can naturally indicate imminent jointness in much the way thay can also naturally indicate emotions.\footnote{
-Ideas along these lines are suggested by the discussion of \emph{emergent coordination} in \citet{Knoblich:2010fk}.
+Ideas along these lines are suggested by the discussion of {emergent coordination} in \citet{Knoblich:2010fk}.
 }  
@@
 Not everything needs to rest on indicators, however.
-It is sometimes possible to know that others are about to willing in joint action with you even without relying on such indicators.  
+It is sometimes possible to know that others are willing to engage in joint action with you even without relying on such indicators.  
 Thanks to widespread dispositions to act jointly, in some situations it is reasonable to take for granted that others are willing to act jointly.  
@@
 Turning to the second point,
-does knowing the premises of the your-goal-is-my-goal inference sometimes put one in a position to know the conclusion?  
+could knowing the premises of the your-goal-is-my-goal inference ever put one in a position to know the conclusion?  
 Of course the inference is not deductive and will only work when certain background conditions are met.  
@@
 In principle,
-exploiting your-goal-is-my-goal does not require that the interpreter actually be in a position to interact with her target.
-It is sufficient (and would be necessary but for some special cases) that the target takes the interpreter to be in a position to interact with him. 
+exploiting your-goal-is-my-goal does not require actually being in a position to interact with the target of goal ascription.
+It is sufficient (and would be necessary but for some special cases) to be taken by the target to be in a position to interact. 
 Our concern, however,
@@
 development or evolution (and ideally both).
-This mandates a focus on interpreters of limited sophistication
+This motivates a focus on goal ascribers of limited sophistication
 who lack both deep insight into others' minds and fully-fledged communicative abilities.
 Such interpreters are unlikely to be able to contrive or exploit situations in which they only appear to be in a position to interact with their targets.
-So the explanatorily relevant cases are those 
-in which an interpreter is manifestly in a position to interact with her target.
-
-Our aim in identifying the your-goal-is-my-goal is not to 
-defend a detailed hypothesis about the mechanisms and processes involved in mindreading.
+When we come to applications of your-goal-is-my-goal, the relevant cases will be those 
+in which a goal ascriber is manifestly in a position to interact with her target.
+
+Our aim in identifying the your-goal-is-my-goal inference is not to 
+defend a detailed hypothesis about  mechanisms and processes supporting mindreading.
 Instead our present concerns are limited to a normative question about the evidential basis of mindreading.
 The your-goal-is-my-goal inference matters
-not because it describes how interpreters actually assign goals
+not because it describes exactly how anyone actually assigns goals (maybe it doesn't)
 but because 
@@
 that is closed to mere observers
-but open to interacting mindreaders.
+and open to interacting mindreaders.
 
@@
 %As we saw, this problem is particularly likely to arise in cases involving novel tools and communicative gestures.
-Showing how your-goal-is-my-goal makes  avoiding this problem possible is a way of demonstrating the potential value  to mindreaders of interaction.
+Showing how your-goal-is-my-goal makes  avoiding this problem possible is a way of demonstrating the potential value   of interaction for mindreading.
 
@@
 Of course this additional evidence will not necessarily trump   other evidence.
-But it does provide a possible way around the problem of opaque means by providing evidence for goal ascriptions that is independent of a mindreader's understanding of which ends actions are means to.
+But it does provide a possible way around the problem of opaque means by providing evidence for goal ascriptions that is independent of a goal ascriber's understanding of which ends actions are means to.
 
@@
 Of course this is sometimes a reason not to rely on your-goal-is-my-goal.
-But where the target has better understanding of relevant means-ends relations,
+But where the target understands relevant means-ends relations,
 such as actions involving novel tools,
@@
 The distant promise of all this is that 
-understanding how interaction widens the evidential basis for mindreading
+understanding how interaction widens the evidential basis for goal ascription and mindreading more generally 
 may eventually enable us to explain
-the origins (in evolution or development, and ideally both) of abilities to learn novel and opaque uses for tools from other agents
+the origins, in evolution or development (and ideally both), of abilities to learn novel and opaque uses for tools from others 
 without assuming that 
@@
 \footnote{
-Please note that, although our discussion borrows an experimental paradigm
+Note that although our discussion borrows an experimental paradigm,
 our aim is not to argue for empirical hypotheses about chimpanzee social cognition.
@@
  an early understanding of communicative actions.
-Of course it would strengthen our argument if we could provide evidence to show that this actually happens.
+Of course it would powerfully demonstrate the relevance of our theoretical argument if we could provide evidence to show that this actually happens.
 But for now we are concerned with more narrowly conceptual issues.
@@
 that, 
-before pointing, the agent had used engaging facial gestures to signal willingness to engage in joint action with us; and
-that
-we were able to think of retrieving the food as a possible distributive goal%
-\footnote{
-    On \emph{distributive goal} see section \vref{sec:joint_action}.
-  }
-of our actions;
+before pointing, the agent had used facial gestures to signal willingness to engage in joint action with us
+%; and that we were able to think of retrieving the food as a possible distributive goal%
+%\footnote{
+%    On \emph{distributive goal} see section \vref{sec:joint_action}.
+%  }
+%of our actions;
 and 
@@
 So the your-goal-is-my-goal inference can 
-enable interpreters
+enable a goal ascriber
 to  
-misunderstand pointing actions as something like  failed reaches rather than as communicative gestures.
+misunderstand pointing actions as something like  failed reaches.
 This means that,
-even without any understanding of communication,
-they can respond appropriately to helpful pointing actions in the context of joint action.
+even without any deeper understanding of communication,
+goal ascribers can respond appropriately to helpful pointing actions in the context of joint action.
 
@@
 The pattern of reasoning generalises to a wider range of communicative gestures including single-word utterances.
-The basic requirement is this: in a particular context, the interpreter must associate a communicative gesture with its referent.
-For instance, she must associate the pointing gesture with the object indicated; or, if (say) she is looking for an object she must associate an utterance of `cupboard' with the nearby cupboard.
+The basic requirement is this: in a particular context, the goal ascriber must associate a communicative gesture with its referent.
+For instance, she must associate the pointing gesture with the object indicated; or, if (say) she is looking for an individual she must associate an utterance of `daddy' with the appropriate person.%
+\footnote{
+Such associations appear to be in place by six months of age or earlier \citep{tincoff:1999_beginnings_,tincoff:2011}.
+There is debate on when infants are first able to rapidly associate words with novel objects \citep{werker:1998_acquisition,friedrich:2011_word_}.
+Interestingly,
+a possible developmental decalage separating 
+two measures of whether infants have formed a word-object association,
+preferential looking and comprehension as measured by reaching in response to a request  \citep{gurteen:2011_rapid},
+may suggest that
+infants can form a word-object association before they can comprehend communicative actions involving the word.
+}
 As we saw, 

### 2011-12-29 joint action and knowing others minds/interacting mindreaders.2.tex
first draft done incl abstract

@@
 \begin{abstract}
-%Goal ascription is fundamental for mindreading.
-What evidence grounds ascriptions of thoughts and actions,
-and how does the evidence support the ascriptions?
-In answering this question, 
-philosophers sometimes focus on mere observation, ignoring interaction.
-The only evidence considered is 
- evidence that would be available to mindreaders 
-observing their targets but in no position to interact with them.
-The present paper,
-which focuses on goal ascription,
-argues that this is a mistake
-and
-identifies evidence 
-available only to mindreaders capable of interacting with their targets.
-Being poised to interact with others may make it possible to know things about their thoughts and actions which one might not otherwise be in a position to know.
-This has consequences for the possible roles of interaction in explaining the evolution and development of mindreading.
-
+Could interacting mindreaders be in a position to know things which they would be unable to know if they were manifestly passive observers?
+This paper defends a positive answer.
+Mindreading is sometimes reciprocal: 
+the mindreader's target reciprocates by taking the mindreader as a target for mindreading.
+The paper explains how such reciprocity can significantly narrow the range of possible interpretations of behaviour 
+where mindreaders are, or appear to be, in a position to interact.
+It follows that
+revisions and extensions 
+are needed for 
+standard theories of mindreading
+on which
+all the evidence relevant to mindreading 
+is in principle available to manifestly passive mindreaders.
+The view also has consequences for 
+understanding 
+how abilities to interact combined with comparatively simple forms of mindreading
+may explain the emergence,
+in evolution or development,
+of sophisticated forms of social cognition.
+
+%
+%What evidence grounds ascriptions of thoughts and actions,
+%and how does the evidence support the ascriptions?
+%In answering this question, 
+%philosophers sometimes focus on mere observation, ignoring interaction.
+%The only evidence considered is 
+% evidence that would be available to mindreaders 
+%observing their targets but in no position to interact with them.
+%The present paper,
+%which focuses on goal ascription,
+%argues that this is a mistake
+%and
+%identifies evidence 
+%available only to mindreaders capable of interacting with their targets.
+%Being poised to interact with others may make it possible to know things about their thoughts and actions which one might not otherwise be in a position to know.
+%This has consequences for the possible roles of interaction in explaining the evolution and development of mindreading. 
+%
 %What evidence grounds ascriptions of goals to other agents' actions,
@@
 }
-Studying how interaction broadens the evidential basis of mindreading will eventually point to one way of filling in some details on just how interaction might facilitate the emergence of mindreading.
+Studying how interaction broadens the evidential basis of mindreading will eventually point to one way of filling in some details on just how interaction might facilitate the emergence of sophisticated forms of mindreading.
 
@@
 
-%*** remove this line: to avoid annoying error only
-\
-
 \section{Your-goal-is-my-goal: a route to knowledge}
@@
 \section{Communicative gestures}
+\label{sec:communicative_gestures}
 When introducing the problem of opaque means 
@@
 The basic requirement is this: in a particular context, the goal ascriber must associate a communicative gesture with its referent.
-For instance, she must associate the pointing gesture with the object indicated; or, if (say) she is looking for an individual she must associate an utterance of `daddy' with the appropriate person.%
+For instance, she must associate the pointing gesture with the object indicated; or, if (say) she is looking to see who has an object she must associate an utterance of `daddy' with the daddy.%
 \footnote{
@@
 merely associating a gesture with its referent falls short of being able to respond appropriately.
-After all, the goals of the gesture may well be unrelated to the goals of the interpreter's actions.
 But if an interpreter supposes that her target is willing to engage in joint action with her,
 then she may infer that the goal of her target's action is her goal
-and so be motivated to treat the thing associated with a communicative gestures as relevant to the goal of her own actions.
+and so be motivated to treat the thing associated with a communicative gesture as relevant to the goal of her own actions.
 This will reliably (but not always) enable her to respond appropriately to the communicative gesture even without understanding it as a communicative gesture.
@@
 how
-abilities to engage in joint action
+possessing abilities to engage in joint action
 mean that 
-a mindreader with an ability to ascribe simple goals only and no understanding of communicative intent
+an individual with an ability to ascribe simple goals only and no understanding of communicative intent
 might 
@@
 %(although of course they come together later in development)
-We have not shown that this hypothesis wrong.
+We have not shown that this claim wrong.
 But we have shown that there is another possibility:
@@
 Widening the evidential basis for goal ascription
-to include evidence available only to interpreters capable of joint action with their targets
+to include evidence available only to mindreaders actually or apparently capable of joint action with their targets
 makes possible 
-novel hypotheses about the emergence, in evolution or development (and perhaps even both), of communicative gestures.
+novel hypotheses about the emergence, in evolution or development (and perhaps both), of communicative gestures.
 
@@
 \label{sec:false_belief}
-So far we have been arguing for the value of interaction to interpreters 
+So far we have been arguing for the value of interaction for goal ascription
 on the grounds that 
-interacting interpreters, unlike mere observers, have 
+interacting mindreaders, unlike mere observers, have 
 a route to knowledge of the goals of others' actions
 which avoids the problem of opaque means.
-There is another reason for supposing that interaction matters for interpretation.
-To introduce this reason we must first describe another problem affecting interpreters, the problem of false belief.
-
-The problem of opaque means occurs when interpreters  must rely entirely on observation and cannot identify to which ends actions are means.
+There is another reason for supposing that interaction matters for goal ascription and (at least indirectly) for mindreading more generally.
+To introduce this reason we must first describe another problem affecting goal ascription, the problem of false belief.
+
+Recall that the problem of opaque means occurs when interpreters  must rely entirely on observation and cannot identify to which ends actions are means 
+(see section \vref{sec:opaque_means}).
 A yet more familiar problem affecting goal ascription 
@@
 \emph{who} needs to be aware of those differences.
-
-In short, then, abilities to engage in joint action provide an interpreter with a route to knowledge of the goals of other agents' actions which does not depend on her knowledge of what her target believes.
+This is a second illustration of how the possibility of interaction 
+may enable 
+	even individuals with only limited insights into others' minds and actions
+ 	to exploit reciprocity in mindreading---to exploit the fact that their targets are sometimes themselves mindreading the mindreaders.
+
+In short, then, abilities to engage in joint action open a route to knowledge of the goals of other agents' actions which does not depend on the goal ascriber knowing what her target believes even when their beliefs relevantly differ.
 This is not because abilities to engage in joint action provide a way to avoid the problem of false beliefs altogether.
@@
 
-On a purely observational model of interpretation,
-it seems inevitable that an interpreter who is unable to track others' beliefs 
-will have to proceed as if there were no differences in beliefs.
-As several philosophers' arguments suggest,
-it is plausible that 
-the interdependence of beliefs and goals 
-means that
-an interpreter will always be able to assign goals to others' actions 
-in a way that is consistent with their beliefs not differing from her own
-(*refs: Davidson; \citealp[p.\ 49]{Bennett:1976rg}).
-If so,
-on a purely observational model of interpretation
-there is no straightforward way in which
+On a purely observational model of mindreading,
+it seems inevitable,
+given that observed behaviours will be consistent with many different ascriptions of beliefs and goals,
+that a goal ascriber who was unable to track others' beliefs 
+would often have to proceed as if there were no relevant differences in beliefs.%
+\footnote{
+Compare \citet[p.\ 49]{Bennett:1976rg}.  For a related argument concerning beliefs and desires rather than beliefs and goals, see 
+	\citet[p.\ 17]{Davidson:1984pr}
+%what can fairly be treated as evidence for any one of the propositional attitudes can count as evidence only if untested assumptions are made about others.”  [1984d, p. 17]
+and
+	(\citeyear[pp.\ 159--167]{Davidson:1984wh}).
+%`This creates a problem, for it means that behaviour, which is the main evidential basis for attributions of belief and desire, is reckoned the result of two forces less open to public observation.
+Thus where one constellation of beliefs and desires will rationalize an action, it is always possible to find a quite different constellation that will do as well.
+Even a generous sample of actions threatens to leave open an unacceptably large number of alternative explanations.'
+%`of course it cannot be assumed that speakers never have false beliefs.
+Error is what gives belief its point.
+We can, however, take it as given that most beliefs are correct.
+The reason for this is that a belief is identified by its location in a pattern of beliefs; it is this pattern that determines the subject matter of the belief, what the belief is about.
+Before some object in, or aspect of, the world can become part of the subject matter of a belief (true of false) there must be endless true beliefs about the subject matter.
+False beliefs tend to undermine the identification of the subject matter; to undermine, therefore, the validity of a description of the belief as being about that subject.'
+}
+So on a purely observational model of interpretation
+there may fail to be any straightforward way in which
 the need to track differences in belief 
-will become apparent to an interpreter.
+could become apparent to a mindreader who did not already understand that beliefs can be false.
 Things are different
-if we extend our model of interpretation
+if we extend our model of mindreading
 to take into account the possibility of interaction.
 We can then see how
-an interpreter who is unable to track others' beliefs
+a goal ascriber who is unable to track others' beliefs
 will encounter anomalies.
-For some differences in belief will mean that
+For some differences in belief 
+(the mere fact of such differences, not anyone's awareness of them)
+will mean that
  ordinary, purely observational routes to knowledge will result in ascribing one goal
@@
  exploiting your-goal-is-my-goal would result in a ascribing a different goal.
-That is, the evidence available to an interpreter may inexplicably support two incompatible ascriptions. 
+That is, the evidence available to a goal ascriber may inexplicably support two incompatible ascriptions. 
 In this way,
 differences in belief have the potential to cause puzzling anomalies
-even before the interpreter is able to track such differences.
+even before a mindreader is able to track such differences.
 This 
@@
 that abilities to engage in joint action
-may play a role in explaining how  awareness of belief emerges.
+may play a role in explaining how  awareness of belief emerges.%
+\footnote{
+Some researchers have suggested that awareness of belief is innate in humans
+\citep{Leslie:1998nq,Leslie:2005ef,Baillargeon:gx},
+although this view is currently controversial
+\citep{Apperly:2009ju,Ruffman:2005yl}.
+But whether or not the emergence in development of awareness of belief stands in need of explanation,
+few would doubt that its emergence in evolution does need explaining.
+}
 
@@
 Our suggestion was this.
-Where an interpreter recognizes that her target is willing to engage in joint action
+Where a mindreader recognizes that her target is willing to engage in joint action
 and is unwilling to change which goals her own actions will be directed to,
@@
 that the goals of her target's actions will be her goals.
-This in outline is the `your-goal-is-my-goal' route to knowledge.
+This in outline is the `your-goal-is-my-goal' route to  knowledge (see further section \vref{sec:your_goal_is_my_goal}).
 
@@
 To show that this route to knowledge is potentially valuable,
-we argued that it enables interpreters to overcome two problems,
+we argued that it enables mindreaders to overcome two problems,
 the problem of opaque means
@@
 motivates focusing on 
-interpreters with
+those with
   absent or fledgling communicative skills,
@@
   and narrow knowledge of how things work.
-For such interpreters
-the problems of false belief and opaque means may easily arise,
+For such individuals
+the problems of opaque means and false beliefs may easily arise,
 particularly in cases involving novel tools 
@@
 communicative gestures.
-Interpreters 
-could overcome the two problems 
+Both problems can be overcome 
 by exploiting the your-goal-is-my-goal route to knowledge.
@@
 does not presuppose knowledge of others' intentions and beliefs.
-Such expertise,
+This shows that it is coherent to suppose that 
+ expertise with tools or communicative gestures (or both),
 far from  presupposing sophisticated forms of mindreading,
 might instead
-play a role in explaining their emergence in evolution or development (or both).
+play a role in explaining their emergence in evolution or development.
 
 How might this work?
-***sketch the story***
-Clearly this is a wild conjecture,
-and not one that we have given any reason for accepting.
+Here is a wild conjecture about
+a path from simple forms of interaction to sophisticated forms of mindreading.
+We have seen that 
+	abilities to ascribe goals to actions 
+ together with
+	abilities to engage in simple forms of joint action involving distributive goals
+ may be sufficient to explain how individuals come to productively misunderstand communicative gestures in the context of joint action (see section \vref{sec:communicative_gestures}).
+This might  be a step towards responding appropriately to, and using, communicative gestures in other contexts,
+which might then play a role in explaining how early forms of communication emerge.
+But we know that early, non-linguistic forms of communication are essential for the later appearance of linguistic communication,
+and abilities to communicate by language in turn appear to play an important role in the development (at least) of sophisticated forms of mindreading.%
+\footnote{
+A range of evidence is reviewed in contributions to \citet{Astington2005ot}.
+}
+Now clearly this wild conjecture is not one that we have provided any reason for accepting.
 Our point in mentioning it is merely to illustrate how
- better understanding the role of interaction in goal ascription and mindreading generally
-may eventually lead to discoveries
+ better understanding the role of interaction in goal ascription and  mindreading more generally
+might eventually lead to discoveries
 on how
@@
  conceptually and cognitively undemanding forms of interaction.
-If the culture of tool use,
-the simpler forms of communicative gesture
-or the sharing of knowledge and other mental states
-originates in abilities to interact,
-then we need to know how interaction could lead to the emergence of these things.
-Our  suggestion has been that
+A culture of tool use,
+	the flexible use of communicative gestures,
+	or insight into interpersonal differences in belief and other mental states:
+if any or all of these 
+originate in abilities to interact,
+then we need to know how interaction could lead to their emergence.
+And one small step towards acquiring this knowledge might just be 
 understanding how interacting mindreaders 
-could know things about their minds which mere observers might not be in a position to know
-will provide some of the details.
+could 
+	come to know things 
+	and
+	encounter anomalies
+which they might not be able to know or encounter if they were mere observers.
+% matters because it enables us to better understand the evidential basis of mindreading---%
+%and it might just also matter because it plays a role in demonstrating the theoretical coherence of hypotheses about mindreading's emergence.
 

### 2011-12-30 joint action and knowing others minds/interacting mindreaders.2.tex
part way through proofing

@@
 
-We are not suggesting that no observer could ever identify the goal of any action she fails to recognise as a means to achieving that goal, of course.
-Of course opaque means are not in every case an insurmountable obstacle to goal ascription.
-This point is neither novel nor surprising:
-opaque means sometimes  deprive interpreters of evidence and so prevent goal ascription.
+We are not suggesting that no observer could ever identify the goal of any action she fails to recognise as a means to achieving that goal.
+Of course opaque means are not in every case an insurmountable obstacle to goal ascription, and
+they may only rarely be a problem for human adults with sophisticated social skills.
+Our point is neither novel nor surprising:
+opaque means sometimes  deprive mindreaders of evidence and so prevent goal ascription.
+This is more likely to happen where  goal ascribers lack sophistication in mindreading, communication and culture.
 
 Some of the most plausibly unique aspects of human cognition depend on our abilities to recognise the goals of novel behaviours involving tools, and of communicative gestures.
-If goal ascription is based on entirely observation 
-(so that the possibility of interaction is ignored),
-the problem of opaque means is likely to arise in both cases.
+The problem of opaque means is likely to arise in both cases
+if goal ascription is based  entirely on observation 
+(so that the possibility of interaction is ignored)
+and if the goal ascriber lacks sophisticated social skills.
 We have just seen an illustration of how the problem of opaque means arises where tools are used to unfamiliar ends.
@@
 failures to identify to which ends actions are means can impair goal ascription.
-The problem is potentially a problem for interpreters given the standard, purely observational models of interpretation.
-It is not our intention to suggest that the problem of opaque means is a problem \emph{for models of interpretation}.
-Rather, it is a potential problem \emph{for interpreters}.
+The problem is potentially a problem for mindreaders given the standard, purely observational models of mindreading.
+It is not our intention to suggest that the problem of opaque means is a problem \emph{for models of mindreading}.
+Rather, it is a potential problem \emph{for mindreaders}.
 
@@
 is the first premise in our argument
-that models of interpretation based on pure observation are less powerful than models taking into account the possibility of interaction.
+that models of mindreading based on pure observation are less powerful than models taking into account the possibility of interaction.
 They are less powerful in this sense: 
@@
 Let us stipulate that an outcome is a \emph{distributive goal} of two or more agents' actions just if two conditions are met.
+\label{df:distributive_goal}
 First, this outcome is a one to which each agent's actions are individually directed.
@@
 %it is necessary that there be a single outcome (and not merely similar outcomes) to which they are directed.
+
+Note that our actions could have a distributive goal even without us each intending that we do something.
+As another variation on the turnip-pulling story,
+suppose that we each individually intend to 
+	contribute to pulling up the turnip or pull it up.
+Such intentions are agent neutral in this sense:
+relative to such an intention an agent 
+	could succeed by acting alone 
+and
+  she also could succeed by acting with unspecified other agents.
+Plausibly our having these intentions is sufficient for our actions to have a distributive goal.
+For in virtue of these intentions each of our actions is directed to the turnip's extraction,
+and this is an outcome relative to which it is possible that all of our actions succeed.
 
@@
 (rather than simply
-intending to pull it up).
-
-Note that our actions could have a distributive goal even without us each intending that we do something.
-As another variation on the turnip-pulling story,
-suppose that we each individually intend to 
-	contribute to pulling up the turnip or pull it up.
-Such intentions are agent neutral in this sense:
-relative to such an intention an agent 
-	could succeed by acting alone 
-and
-  she also could succeed by acting with unspecified other agents.
-Plausibly our having these intentions is sufficient for our actions to have a distributive goal.
-For in virtue of these intentions each of our actions is directed to the turnip's extraction,
-and this is an outcome relative to which it is possible that all of our actions succeed.
+intending to pull it up),
+and from the possibility of each of us
+intending to contribute to pulling up the turnip or pull it up.
 
@@
 \footnote{
+\label{fn:joint_action_distributive_goal}
 We know of two definitions of joint action on which it is not straightforward that all joint actions involve distributive goals.
 One is provided 
- is provided by \citet[p.\ 366]{ludwig_collective_2007} and discussed in footnote \vref{fn:ludwig}; the other is due to 
+ is provided by \citet[p.\ 366]{ludwig_collective_2007} and quoted in footnote \vref{fn:ludwig}; the other is due to 
 \citet[p.\ 70]{Sebanz:2006yq}.
@@
 	\citet[pp.\ 329-31]{Bratman:1992mi},
-%	\citet[p.\ 40]{Butterfill:2011fk},
 	\citet[pp.\ 96-7]{Searle:1990em} and
@@
 However, the converse does not hold.
-As this the turnip-pulling variation  indicates,
+As the first variation of the turnip-pulling story indicates,
 two or more agents' actions may have a distributive goal 
@@
 In fact our actions might have a distributive goal even though none of us is aware of this, or even of the others' existence. 
-(We shall refrain from illustrating this point by contriving a truly gigantic turnip and a completely dark, very windy night, but we are sure readers will be able to imagine how this would go.)
+(We shall refrain from illustrating this point by contriving a truly gigantic turnip and a completely dark, very stormy night; readers can probably guess how this would go.)
 One consequence of this is that 
@@
 \footnote{
-\label{fn:ludwig}
-According to 
-	 \citet[p.\ 366]{ludwig_collective_2007}, 
-`The concept of a joint action as such is just that of an event of which there are multiple agents.'
-Depending on what events are and what it is to be the agent of an event,
-it may turn out that,
-on Ludwig's definition,
-our actions' having a distributive goal 
-is sufficient for
-us to be engaged in joint action.
-However, other accounts of joint action have the consequence that each agent involved in a joint action must believe or know something about the jointness of her action.
+On many accounts of joint action, each agent involved in a joint action must believe, expect or know something about the jointness of her action.
 See 
@@
 can answer the question of what he is doing or will be doing by saying for example ``We are walking together'' or ``We will/intend to walk together.'''
+Exceptions include  \citet{pacherie_framing_2011} and \citet{ludwig_collective_2007}.
+\label{fn:ludwig}
+According to Ludwig,
+`The concept of a joint action as such is just that of an event of which there are multiple agents' (\citeyear[p.\ 366]{ludwig_collective_2007}).
+Depending on what events are and what it is to be the agent of an event,
+it may turn out that,
+on Ludwig's definition,
+our actions' having a distributive goal 
+is sufficient for
+us to be engaged in joint action.
 }
@@
 narrowly technical one.
-In introducing this notion by stipulation we are not aiming to match anyone's intentions about anything.
+In introducing this notion by stipulation we are not aiming to match anyone's intuitions about anything.
 What matters for now is just that the definition is theoretically coherent and that it is possible for actions to have distributive goals.
@@
 
-\section{Your-goal-is-my-goal: a route to knowledge}
+\section{Your-goal-is-my-goal}
 \label{sec:your_goal_is_my_goal}
@@
 The answer hinges on interactions involving distributive goals
-(see \vref{sec:joint_action}).
+(as defined \vpageref{df:distributive_goal}).
 
@@
 if a mindreader is engaged in an interaction with her target that involves a distributive goal, 
-it may be easy for the interpreter to know what the goal of her target's actions is because this goal is the goal of her own actions.
+it may be easy for the mindreader to know what the goal of her target's actions is because this goal is the goal of her own actions.
 So if she knows the goal of her own actions and she knows that she is engaged with her target in an interaction involving a distributive goal,
 then she already knows what the goal of her target's actions are.
-%Roughly speaking, the interpreter can reason about her target thus: your goal is my goal.
-
-Of course this intuitive idea won't work as it stands.  
+%Roughly speaking, the mindreader can reason about her target thus: your goal is my goal.
+
+Of course this intuitive idea is no use it stands.  
 For the inference it captures relies on the premise that the mindreader and her target are engaged in actions with a distributive goal.
@@
 Seeing a stranger with the haggard look of a new parent struggling to get his twin pram on to the bus, you grab the front wheels and make eye contact, raising your eyebrows and smiling.
-(The noise of the street rules out talking, 
-and you can barely speak the local language anyway.)   
+(The noise of the street rules out talking.)   
 In this way you signal that you are about to act jointly with the stranger, to lift the pram with him.   
@@
 which means that your actions and his actions will have a distributive goal.
-This enables him to infer the goal of your imminent actions: 
+This may enable him to infer the goal of your imminent actions: 
 your goal is his goal, to get the pram onto the bus.
@@
 We leave open the issue of how joint action is to be characterised subject only to the 
-requirement that all joint actions must involve distributive goals.
+requirement that all joint actions must involve distributive goals (see further footnote \vref{fn:joint_action_distributive_goal}).
 Attempts to characterise joint action in ways relevant to explaining development include
@@
 Thanks to widespread dispositions to act jointly, in some situations it is reasonable to take for granted that others are willing to act jointly.  
-For example, this is often so for children surrounded by family or familiar adults who are struggling with a coat.  
+For example, this is often so for children struggling with a coat while surrounded by family or familiar adults.  
 And in at least some subcultures people using public transport can reasonably take for granted that, within limits, those around them will act jointly with them when the need is clear.
@@
 (We return to this point in section \vref{sec:false_belief}.)  
-But there are situations in which it is reasonable  to ignore  this possibility, as for instance when stereotypes, conventions or simplicity should and do make the goal of your actions obvious to the other agent.
+But in some situations there is no requirement  to consider  this possibility, as for instance when stereotypes, conventions or simplicity should and do make the goal of your actions obvious to the other agent.
 Given the right conditions,
@@
 And knowing the premises sometimes puts one in a position to know the conclusion thanks to the fact that, when things are going well, another's willingness to engage in joint action with you can be based on an accurate assessment of the goals of your actions.
+
+In principle,
+exploiting your-goal-is-my-goal does not require actually being in a position to interact with the target of goal ascription.
+It is sufficient (and would be necessary but for some special cases) to be taken by the target to be in a position to interact. 
+Our concern, however,
+is with cases that are likely to be important for understanding
+development or evolution.
+This motivates a focus on goal ascribers of limited sophistication
+who lack both deep insight into others' minds and fully-fledged communicative abilities.
+Such individuals are unlikely to be able to contrive or exploit situations in which they only appear to be in a position to interact with their targets.
+When we come to applications of your-goal-is-my-goal, the relevant cases will be those 
+in which a goal ascriber is manifestly in a position to interact with her target.
 
@@
 nor even to be able to think of it in this way.
-
-In principle,
-exploiting your-goal-is-my-goal does not require actually being in a position to interact with the target of goal ascription.
-It is sufficient (and would be necessary but for some special cases) to be taken by the target to be in a position to interact. 
-Our concern, however,
-is with cases that are likely to be important for understanding
-development or evolution (and ideally both).
-This motivates a focus on goal ascribers of limited sophistication
-who lack both deep insight into others' minds and fully-fledged communicative abilities.
-Such interpreters are unlikely to be able to contrive or exploit situations in which they only appear to be in a position to interact with their targets.
-When we come to applications of your-goal-is-my-goal, the relevant cases will be those 
-in which a goal ascriber is manifestly in a position to interact with her target.
-
 Our aim in identifying the your-goal-is-my-goal inference is not to 
-defend a detailed hypothesis about  mechanisms and processes supporting mindreading.
+defend a detailed hypothesis about  mechanisms of mindreading.
 Instead our present concerns are limited to a normative question about the evidential basis of mindreading.
@@
 Suppose the stranger starts tipping the pram in a way that the inexperienced parent fails to recognise as a means, indeed the only means, of getting it onto the bus.
-Outside the context of joint action this might give the parent sufficient evidence to reject the idea that the goal of the stranger's actions is to get the pram onto this bus.
-Or suppose the stranger starts by pulling the pram away from the bus in order to better position it for entry,
-and the novice parent does not realise that this is necessary.
 Outside the context of joint action, 
-his evidence might on balance support the conclusion that the goal of the stranger's actions is to take the pram off the bus.
+failure to realise that the action is a means to getting the pram onto the bus
+might have the consequence that
+the parent's evidence on balance supports the conclusion that the goal of the stranger's actions is to take the pram off the bus.
 (Perhaps the stranger is impatient to get onto the bus herself.)
+%Or suppose the stranger starts by pulling the pram away from the bus in order to better position it for entry,
+%and the novice parent does not realise that this is necessary.
+%Outside the context of joint action, 
+%his evidence might on balance support the conclusion that the goal of the stranger's actions is to take the pram off the bus.
+%(Perhaps the stranger is impatient to get onto the bus herself.)
 But in the context of joint action,
@@
 this really is the goal of her actions.
-Of course this additional evidence will not necessarily trump   other evidence.
-But it does provide a possible way around the problem of opaque means by providing evidence for goal ascriptions that is independent of a goal ascriber's understanding of which ends actions are means to.
+Of course this additional evidence will not always outweigh   other evidence.
+But it will do sometimes, and this is all we need.
+We have identified evidence for goal ascription that is available independently of a goal ascriber's knowing which ends actions are means to, and we have illustrated how this will sometimes enable mindreaders to escape the problem of opaque means
 
@@
 such as actions involving novel tools,
-the your-goal-is-my-goal route to knowledge of other's goals may sometimes be the only option.
+the your-goal-is-my-goal route to knowledge of others' goals may sometimes be the only option.
+
+Here we have a first illustration of how
+interacting mindreaders can exploit reciprocity 
+without relying on ascriptions of higher-order mental states.
+The mindreader's knowledge of the goal of her target's actions depends on the target's knowledge of the goal of the mindreader's actions.
+What makes this dependence possible is not each knowing that the other knows something about the goal of her actions;
+ it is  the actual or apparent  possibility of interaction.
 
@@
 that your-goal-is-my-goal
-could \emph{in principle} enable
+could {in principle} enable
 individuals to make the transition 
@@
 merely associating a gesture with its referent falls short of being able to respond appropriately.
-But if an interpreter supposes that her target is willing to engage in joint action with her,
+But if a mindreader supposes that her target is willing to engage in joint action with her,
 then she may infer that the goal of her target's action is her goal
@@
 possessing abilities to engage in joint action
-mean that 
+means that 
 an individual with an ability to ascribe simple goals only and no understanding of communicative intent
@@
 we have not attempted to defend any hypothesis about 
-anything's evolution or development.
-Our point is just to illustrate one of several ways in which
-interaction may matter for mindreading.
-Widening the evidential basis for goal ascription
-to include evidence available only to mindreaders actually or apparently capable of joint action with their targets
-makes possible 
-novel hypotheses about the emergence, in evolution or development (and perhaps both), of communicative gestures.
+anything's actual evolution or development.
+%Our point is just to illustrate one of several ways in which interaction may matter for mindreading.
+But widening the evidential basis for goal ascription
+to include evidence available only to mindreaders actually or apparently capable of interacting with their targets
+does reveal the coherence of  
+novel hypotheses about a role for interaction in the emergence  of communicative gestures.
 
@@
 
-Recall that the problem of opaque means occurs when interpreters  must rely entirely on observation and cannot identify to which ends actions are means 
+Recall that the problem of opaque means occurs when mindreaders  must rely entirely on observation and cannot identify to which ends actions are means 
 (see section \vref{sec:opaque_means}).
@@
 impair goal ascription
-when the interpreter is unaware of those differences.
+when the goal ascriber is unaware of those differences.
 Call this the problem of false belief.
@@
 route to knowledge
-sometimes enables interpreters to avoid the problem of false belief.
+sometimes enables mindreaders to avoid the problem of false belief.
 To illustrate consider a counterfactual alternative to the above example.
@@
 This is not because abilities to engage in joint action provide a way to avoid the problem of false beliefs altogether.
-Rather they shift the burden of resolving the problem of false belief from an interpreter to her target.
-This is potentially valuable for interpreters who have limited insight into their targets' beliefs,
+Rather they shift the burden of resolving the problem of false belief from a goal ascriber to her target.
+This is potentially valuable for those who have limited insight into their targets' beliefs,
 or who lack abilities to track differences in belief altogether.
 
-On a purely observational model of mindreading,
-it seems inevitable,
-given that observed behaviours will be consistent with many different ascriptions of beliefs and goals,
-that a goal ascriber who was unable to track others' beliefs 
-would often have to proceed as if there were no relevant differences in beliefs.%
-\footnote{
-Compare \citet[p.\ 49]{Bennett:1976rg}.  For a related argument concerning beliefs and desires rather than beliefs and goals, see 
-	\citet[p.\ 17]{Davidson:1984pr}
-%what can fairly be treated as evidence for any one of the propositional attitudes can count as evidence only if untested assumptions are made about others.”  [1984d, p. 17]
-and
-	(\citeyear[pp.\ 159--167]{Davidson:1984wh}).
-%`This creates a problem, for it means that behaviour, which is the main evidential basis for attributions of belief and desire, is reckoned the result of two forces less open to public observation.
+%On a purely observational model of mindreading,
+%it seems inevitable,
+%given that observed behaviours will be consistent with many different ascriptions of beliefs and goals,
+%that a goal ascriber who was unable to track others' beliefs 
+%would often have to proceed as if there were no relevant differences in beliefs \citep[compare][p.\ 49]{Bennett:1976rg}. 
+%%\footnote{
+%%Compare \citet[p.\ 49]{Bennett:1976rg}.  For a related argument concerning beliefs and desires rather than beliefs and goals, see 
+%%	\citet[p.\ 17]{Davidson:1984pr}
+%%%what can fairly be treated as evidence for any one of the propositional attitudes can count as evidence only if untested assumptions are made about others.”  [1984d, p. 17]
+%%and
+%%	(\citeyear[pp.\ 159--167]{Davidson:1984wh}).
+%%%`This creates a problem, for it means that behaviour, which is the main evidential basis for attributions of belief and desire, is reckoned the result of two forces less open to public observation.
 Thus where one constellation of beliefs and desires will rationalize an action, it is always possible to find a quite different constellation that will do as well.
 Even a generous sample of actions threatens to leave open an unacceptably large number of alternative explanations.'
-%`of course it cannot be assumed that speakers never have false beliefs.
+%%%`of course it cannot be assumed that speakers never have false beliefs.
 Error is what gives belief its point.
@@
 False beliefs tend to undermine the identification of the subject matter; to undermine, therefore, the validity of a description of the belief as being about that subject.'
-}
-So on a purely observational model of interpretation
-there may fail to be any straightforward way in which
-the need to track differences in belief 
-could become apparent to a mindreader who did not already understand that beliefs can be false.
-Things are different
-if we extend our model of mindreading
-to take into account the possibility of interaction.
-We can then see how
-a goal ascriber who is unable to track others' beliefs
-will encounter anomalies.
-For some differences in belief 
-(the mere fact of such differences, not anyone's awareness of them)
-will mean that
- ordinary, purely observational routes to knowledge will result in ascribing one goal
- whereas
- exploiting your-goal-is-my-goal would result in a ascribing a different goal.
-That is, the evidence available to a goal ascriber may inexplicably support two incompatible ascriptions. 
-In this way,
-differences in belief have the potential to cause puzzling anomalies
-even before a mindreader is able to track such differences.
-This 
-raises the possibility
-that abilities to engage in joint action
-may play a role in explaining how  awareness of belief emerges.%
-\footnote{
-Some researchers have suggested that awareness of belief is innate in humans
-\citep{Leslie:1998nq,Leslie:2005ef,Baillargeon:gx},
-although this view is currently controversial
-\citep{Apperly:2009ju,Ruffman:2005yl}.
-But whether or not the emergence in development of awareness of belief stands in need of explanation,
-few would doubt that its emergence in evolution does need explaining.
-}
+%%}
+% So on a purely observational model of mindreading
+%there may fail to be any straightforward way in which
+%the need to track differences in belief 
+%could become apparent to a mindreader who did not already understand that beliefs can be false.
+%Things are different
+%if we extend our model of mindreading
+%to take into account the possibility of interaction.
+%We can then see how
+%a goal ascriber who is unable to track others' beliefs
+%will encounter anomalies.
+%For some differences in belief 
+%(the mere fact of such differences, not anyone's awareness of them)
+%will mean that
+% ordinary, purely observational routes to knowledge will result in ascribing one goal
+% whereas
+% exploiting your-goal-is-my-goal would result in a ascribing a different goal.
+%That is, the evidence available to a goal ascriber may inexplicably support two incompatible ascriptions. 
+%In this way,
+%differences in belief have the potential to cause puzzling anomalies
+%even before a mindreader is able to track such differences.
+%This 
+%raises the possibility
+%that abilities to engage in joint action
+%may play a role in explaining how  awareness of belief emerges.%
+%\footnote{
+%Some researchers have suggested that awareness of belief is innate in humans
+%\citep{Leslie:1998nq,Leslie:2005ef,Baillargeon:gx},
+%although this view is currently controversial
+%\citep{Apperly:2009ju,Ruffman:2005yl}.
+%But whether or not the emergence in development of awareness of belief stands in need of explanation,
+%few would doubt that its emergence in evolution does need explaining.
+%}
 
@@
 Where a mindreader recognizes that her target is willing to engage in joint action
-and is unwilling to change which goals her own actions will be directed to,
+and where the mindreader is unwilling to change which goals her own actions will be directed to,
 she may be in a position to know
-that the goals of her target's actions will be her goals.
-This in outline is the `your-goal-is-my-goal' route to  knowledge (see further section \vref{sec:your_goal_is_my_goal}).
+that the goals of her target's actions will be the goals of her own actions.
+This in outline is the `your-goal-is-my-goal' route to  knowledge (see  section \vref{sec:your_goal_is_my_goal}).
 
@@
  conceptually and cognitively undemanding forms of interaction.
-A culture of tool use,
+Take cultures of tool use,
 	the flexible use of communicative gestures,
@@
 	come to know things 
-	and
-	encounter anomalies
-which they might not be able to know or encounter if they were mere observers.
+%	and
+%	encounter anomalies
+which they might not be able to know 
+%or encounter
+ as mere observers.
 % matters because it enables us to better understand the evidential basis of mindreading---%

### 2011-12-30 joint action and knowing others minds/interacting mindreaders.2.tex
proofd

@@
 Could interacting mindreaders be in a position to know things which they would be unable to know if they were manifestly passive observers?
-This paper defends a positive answer.
+This paper argues that they could.
 Mindreading is sometimes reciprocal: 
@@
 revisions and extensions 
-are needed for 
+are needed to 
 standard theories of mindreading
 on which
-all the evidence relevant to mindreading 
+all the evidence 
 is in principle available to manifestly passive mindreaders.
@@
 of sophisticated forms of social cognition.
+
+\
+
+\noindent
+Word count: 10 200
 
@@
 
-\section{Mindreading: the normative project}
+\section{A question about the evidential basis of mindreading}
 \label{sec:intro}
@@
 The two are on a par in this sense:
-in principle the same evidence could be available to each, and each can exploit the same principles in moving from evidence to ascriptions of thought.
-Of course none of these theories is incompatible with the idea that interaction might be useful for mindreading in practice.  
-The point is just that, on these theories, interaction makes no difference to the theory.
+in principle the same evidence could be available to each, and each can exploit the same principles in moving from evidence to ascriptions of thought and action.
+Of course these theories are all compatible with the idea that interaction might be useful for mindreading in practice.  
+But on these theories interaction makes no difference to what can in principle be known.
 In this paper we aim to show
@@
 If we assume the mindreader is merely observing her target,
-that there is no potential for interaction,
+that there is manifestly no potential for interaction,
 then it seems that any way of exploiting reciprocity would involve higher-order ascriptions. 
@@
 an escalation of higher-order ascriptions.
-Perhaps, then, fully exploiting reciprocity in mindreading
-requires the mindreaders to be in a position to interact with each other.
-Perhaps in some cases 
+Perhaps fully exploiting reciprocity in mindreading
+requires the mindreaders to be in a position to interact with each other;
+perhaps in some cases 
 being or appearing poised to interact
-can somehow enable mindreaders to exploit reciprocity without first having to ascribe of higher-order mental states.
+can somehow enable mindreaders to exploit reciprocity without first having to ascribe higher-order mental states.
 This is the hunch we develop in what follows.
@@
 One reason is that the investigation will enable us to revise and extend existing accounts of the evidential basis of mindreading  in ways that make them more accurate and comprehensive.
-Another motive concerns the emergence, in evolution and in development, of mindreading.
+Another motive concerns the emergence, in evolution or in development, of mindreading.
 Several researchers have offered quite general conjectures about how interaction might explain the emergence of sophisticated forms of cognition.
@@
 Another view is that abilities to interact with others
-may have fostered the emergence in evolution---or, on a related view, in development---%
+may have fostered the emergence in evolution---or, on a different view, in development---%
 of sophisticated forms of cognition including social cognition and, in particular, mindreading.%
@@
 Only cases involving the first group---mindreaders who actually could interact---are likely to be of wider interest.
-Cases involving misleading appearances might be regarded as artefacts of the model.
+%Cases involving misleading appearances might be regarded as artefacts of the model.
 Any application of our claim to understanding evolution or development is bound to focus on cases where mindreaders actually could interact with their targets.
@@
 will indirectly support the ascription of a belief or other mental state---%
-the evidence will support the ascription of a goal to an action,
-and this ascription will in turn support the ascription of a belief (say) to the agent of the action.
-Accordingly, even those who deny that goal ascription is mindreading should agree that sometimes evidence for a goal ascription is indirectly evidence for the ascription of a belief or other mental state.
+the evidence will support  a goal ascription which will in turn support the ascription of a belief (say) to the agent of the action.
+Accordingly, even those who deny that goal ascription is mindreading should agree that sometimes evidence for goal ascription is indirectly evidence for ascription of belief and other mental states.
 So whether or not goal ascription is deemed to be mindreading,
-a mindreader's access to evidence for mindreading  depends in part on her access to evidence for goal ascription.
+an individual's access to evidence for mindreading  depends in part on her access to evidence for goal ascription.
 Our plan is to show that interacting mindreaders could make use, in ascribing goals, of evidence and principles which are not available to mere observers (who do not even appear to be in a position to interact with their targets).
@@
 
-Developing these ideas requires us to fill in some background on goal ascription and its limits, as well as on notions of goal-directed interactions.
-Readers impatient to get to the central idea should skip to section \vref{sec:your_goal_is_my_goal}.
+Developing these ideas requires us to fill in some background on goal ascription and its limits, as well as on notions of goal-directed interaction.
+Readers impatient to get to the central idea might skip to section \vref{sec:your_goal_is_my_goal}.
 
@@
 Hannah kicks a ball thereby both preventing her sisters from scoring and also breaking a window.
-Asked about episode,
-Hannah might protest that the goal of her action was not to break the window but only to reverse the others' advance.
+Asked about the episode,
+Hannah might protest, truthfully, that the goal of her action was not to break the window but only to reverse the others' advance.
 As this illustrates,
 among the actual and possible outcomes of an action,
-only some are outcomes to which the action was directed.
-Goal ascription is the process of identifying from among an action's actual and possible outcomes
-those to which it is directed.
+only some are outcomes to which the action is directed.
+Goal ascription is the process of identifying 
+ those outcomes.
+%, among an action's actual and possible outcomes,
+%those to which it is directed.
 %Our immediate question is therefore, 
@@
   which is unavailable to those who merely observe,
-we may be able to indicate ways in which interaction could facilitate the emergence, in evolution or development, of more sophisticated forms of mindreading.%
+we will eventually be able to indicate ways in which interaction could facilitate the emergence, in evolution or development, of more sophisticated forms of mindreading.%
 \footnote{
@@
 %\citep{Horner:2005pj}
-In a different case, knowing the goal of another agent's actions may enable you to discover a new use for a tool.
+%In a different case, knowing the goal of another agent's actions may enable you to discover a new use for a tool.
 A slightly less familiar idea is that goal ascription enables one to learn from others' failures as well as their successes.
@@
 then you now have evidence as to where they might be.\footnote{
-\citet{hare_chimpanzees_2004} exploit this principle in testing chimpanzees' abilities to ascribe goals.
+\citet{hare_chimpanzees_2004} exploit this fact in testing chimpanzees' abilities to ascribe goals.
 }
@@
 Goal ascription is also instrumental for ascribing propositional attitudes.
-We have already mentioned that such ascriptions often involve 
+We have already mentioned 
+   (\vpageref*{goal_ascription_verify})
+that such ascriptions often involve 
 confirming predictions about actions 
-which in turn typically involves
-identify the goals of those actions (\vpageref*{goal_ascription_verify}).
+which in turn typically requires
+identifying the goals of those actions.
 In addition,
@@
 If this claim were true,
-there would be no such thing as a theory of the evidential basis of goal ascription,
- only a theory of the evidential basis of mindreading more generally.
-
-The falsity of this claim seems to be assumed in both developmental and comparative research on goal ascription.%
+it would make no sense to discuss the evidential basis of goal ascription except in discussing the evidential basis of mindreading more generally.
+However,
+the falsity of this claim  is presupposed  in both developmental and comparative research on goal ascription.%
 \footnote{ 
@@
 }
-We know of no compelling reason for taking this claim to be true.
+And we know of no compelling reason for taking this claim to be true.
 After all,
@@
 these considerations are perhaps sufficient to motivate 
-exploring what evidence might ground goal ascription.
+exploring what evidence might support goal ascription.
 According to 
@@
 We suggest that the above inference, from (1)-(5) to (6), is a route to knowledge of the goals of actions in this sense:
+in some cases
 it would be possible to know the premises without already knowing the conclusion;
 and, 
-in some cases knowing the premises could put one in a position to know the conclusion.%
-\footnote{
-Knowledge of the premises may not require explicit knowledge of what others find obvious, what is salient to them or what makes actions better for them.
+in some of those cases,
+knowing the premises could put one in a position to know the conclusion.%
+\footnote{
+Knowledge of the premises may not require explicit knowledge of 
+	what is salient to others,
+	what is desirable to them,
+	or what makes actions better for them.
 It is arguably sufficient that
@@
 with respect to 
-	what she finds obvious, 
-	what makes one action a better or worse means of realising an outcome than another action for her,  
-	what is desirable to her,
-	and what is salient to her.
+these things.
+%	what makes one action a better or worse means of realising an outcome than another action for her,  
+%	what is desirable to her,
+%	and what is salient to her.
 }
@@
 may be able to identify the goal of the first action only.
-As this illustrates, ignorance about to which ends actions are means can be an obstacle to goal ascription.%
+As this illustrates, ignorance about to which ends actions are means can be an obstacle to goal ascription.
 Call this the problem of opaque means.
@@
 failures to identify to which ends actions are means can impair goal ascription.
-The problem is potentially a problem for mindreaders given the standard, purely observational models of mindreading.
-It is not our intention to suggest that the problem of opaque means is a problem \emph{for models of mindreading}.
-Rather, it is a potential problem \emph{for mindreaders}.
+The problem is potentially a problem for mindreaders given the standard, purely observational theories of mindreading.
+Note that it is not our intention to suggest that the problem of opaque means is a problem \emph{for theories of mindreading};
+what matters for our purposes is only that it is a potential problem \emph{for mindreaders}.
 
 That the problem of opaque means exists
-is the first premise in our argument
-that models of mindreading based on pure observation are less powerful than models taking into account the possibility of interaction.
+is the first step in our argument
+that 
+theories of the evidential basis of mindreading based on pure observation 
+are less powerful than theories taking into account the possibility of interaction.
 They are less powerful in this sense: 
@@
 \label{df:distributive_goal}
-First, this outcome is a one to which each agent's actions are individually directed.
+First, this outcome is  one to which each agent's actions are individually directed.
 Second, each agent's actions are related to the outcome in such a way that it is possible for all the agents (not just any agent, all of them together) to succeed in bringing about this outcome.
@@
 in the right contexts
-an individual can recognize in another's facial expressions, engaging gestures or synchronized bodily movements that she is about to engage in joint action with him.
+an individual can recognize in another's facial expressions, engaging gestures or synchronized bodily movements that she is about to engage in joint action with her.
 Exploiting these indicators does not typically depend on knowing the particular contents of any beliefs, desires or goals.  
@@
 But they are sufficiently reliable to ground knowledge in some cases.  
-The existence of such indicators shows that knowing the first  premise of the above inference does not require already knowing which particular goals the other has.
+The existence of such indicators shows that knowing the first  premise of the above inference does not require already knowing which particular goals the other's actions will be directed to.
 

### 2012-01-07 intention motor representation with corrado sinigaglia/intention and motor action.tex
added intention motor action (was svn, older revisions not imported but still in svn)

%!TEX TS-program = xelatex
%!TEX encoding = UTF-8 Unicode

\def \papersize {a4paper}

\documentclass[12pt,\papersize]{extarticle}
% extarticle is like article but can handle 8pt, 9pt, 10pt, 11pt, 12pt, 14pt, 17pt, and 20pt text

\def \ititle {Intention and motor representation:}
\def \isubtitle {how to resist temptation}
\def \iauthor {S. Butterfill* \& C. Sinigaglia**
\\ 
**Department of Philosophy, University of Warwick
\\ 
***Dipartimento di Filosofia, Università degli Studi di Milano}
\def \iemail{s.butterfill@warwick.ac.uk}
%\date{}

\input{$HOME/Documents/submissions/preamble_steve_paper2}
%\author{}
%\date{}

%\setromanfont[Mapping=tex-text]{Sabon LT Std}

\begin{document}

\setlength\footnotesep{1em}

\bibliographystyle{newapa} %apalike

\maketitle
%\tableofcontents
\title{}

\begin{abstract}
Are there distinctive roles for intention and motor representation in explaining the purposiveness of action? Standard accounts of action assign a role to intention but are silent on motor representation. The temptation is to suppose that nothing need be said here because motor representation is either only an enabling condition for purposive action or else merely a variety of intention. This paper provides reasons for resisting temptation. Some motor representations, like intentions, coordinate actions in virtue of representing outcomes; but, unlike intentions, motor representations cannot feature as premises or conclusions in practical reasoning.  This implies that motor representation has a distinctive role in explaining the purposiveness of action. It also gives rise to a problem: were the roles of intention and motor representation entirely independent, this would impair effective action. Resisting temptation therefore requires explaining how intentions interlock with motor representations. The solution, we argue, is to recognise that the contents of intention can be partially determined by the contents of motor representations. Understanding this content-determining relation enables better understanding how intentions are related to actions.
\end{abstract}

\tolerance=5000

\section{Introduction}

What is the relation between a purposive action and the outcome or outcomes to which it is directed? The standard way of answering this question appeals to intention, a propositional attitude which plays a characteristic role in planning and coordinating action, is linked to practical reasoning and is subject to characteristic norms (\citealp{Bratman:1987xw}).  On the standard view, an action is directed to an outcome in virtue of the action's being appropriately related to an intention which represents this outcome or some related outcome. As this view is usually expounded, the relation between actions and outcomes to which they are directed is treated as largely independent of the motor processes and representations underpinning action execution. Motor representations are usually considered either as philosophically irrelevant enabling conditions, or else as merely filling in additional details of the basic schema provided by the standard story. Our aim is to show that this is a mistake. We shall argue that, once some basic features of motor representations are properly understood, the standard view must be  refined and extended in ways that allow us to recognize distinctive roles for intention and motor representation in explaining the purposiveness of action.

Twin temptations stand in our way. The first temptation is to suppose that motor representation, although significant  for action execution, has no role in grounding the outcome-directedness of action. Surrendering to this temptation might be reasonable if all motor representations were about merely kinematic or dynamic features of actions. But some motor representations do represent action outcomes and some actions are directed to those outcomes in virtue of the guiding role of motor representations, or so we shall argue. This reason to resist the first temptation throws us straight into the arms of a second temptation, the temptation to suppose that where motor representations represent action outcomes, they are intentions. Surrendering to this second temptation would be reasonable if motor representations were like intentions in that they could feature as premises or conclusions in practical reasoning, or if there were some other planning processes in which both intention and motor representation could feature. However, as we shall argue, this is impossible due to differences between the propositional format of intentions and the distinctively motor, non-propositional format of motor representations.  In short, we must resist both temptations because some motor representations are like intentions in representing action outcomes while also remaining sufficiently unlike intentions in that no single planning process can integrate both intention and motor representation.

While there are good reasons to resist both temptations (or so we shall argue), the recognition that motor representation can ground the outcome-directedness of purposive actions independently of intention gives rise to a problem. For a single action, which outcomes it is directed to may be multiply determined by an intention and, seemingly independently, by a motor representation. Unless such intentions and motor representations are to pull an agent in incompatible directions, which would typically impair action execution, there are requirements concerning how the outcomes they represent must be related to each other. The problem is to explain how any such requirements could be met non-accidentally; we call this `the interface problem' . The key to solving this problem, we shall suggest, is to recognise that intentions can have constituents which refer to outcomes by deferring to motor representations of those outcomes. Effective action sometimes requires that the contents of intentions and of motor representations interlock, and this interlocking occurs when the contents of intentions are determined in part by the contents of motor representations.  So whereas discussions of intention frequently ignore motor aspects of action, it turns out that understanding how intentions interface with motor representations is required for fully understanding how intentions are related to actions.

\section{Motor representations link actions to outcomes}

Our overall aim is to show that intention and motor representation play distinctive roles in explaining the purposiveness of action, that is, in relating actions to outcomes to which they are directed. The first temptation standing in our way is the temptation to suppose that motor representation, although perhaps important for action execution in some agents, has no bearing at all on our question about the relation between actions and the outcomes to which they are directed. Just as it would be an error to suppose that details of musculoskeletal structure are relevant to this question, so equally---so the temptation---it would be an error to suppose that facts about motor representation are relevant here.

Surrendering to this temptation might be reasonable if all motor representations represented only kinematic or dynamic features of actions, such as mere joint displacements or muscle contractions. However, we shall argue in this section that some motor representations represent action outcomes such as grasping, tearing or throwing. Furthermore, as we shall go on to argue (also in this section), such representations ground purposive actions. This in outline is why the first temptation should be resisted.

How does representing an outcome differ from representing merely kinematic or dynamic features of action? How, for instance, does a motor representation of grasping (assuming for now that such a thing exists) differ from a representation of a sequence of joint displacements? First, a motor representation of grasping captures something common to many different sequences of joint displacements and postures involving a variety of effectors. To illustrate, an agent might grasp an object with her hand, with her mouth, with normal pliers (where grasping requires closing the hand), or with reverse pliers (where grasping requires opening the hand). A motor representation of grasping identifies something common to all these cases. Second, and conversely, a motor representation of grasping potentially distinguishes between the same sequence of joint displacements in different contexts. For instance, how a grasping action is represented may depend on a relatively distal outcome: on whether (say) it results in the object grasped being eaten or placed. Also the joint displacements which realise grasping in one context might in another context realise a different action, such as scratching; or they might fail to realise any action at all (because there is no target object, say). Third, a motor representation of an action outcome locates the action within a hierarchical structure. To illustrate, a left-handed precision grip and a right-handed whole hand grip are different action outcomes and both instances of grasping.\footnote{ On the notion of action outcome in motor representation, see \citet{jeannerod_motor_2006} and \citet{rizzolatti_mirrors_2008}.}

Why accept that there are motor representations of action outcomes? The first step is to consider evidence that motor processes carry information about action outcomes. For any given marker of motor processing (such as a pattern of neuronal discharge or motor-evoked potentials), how can we test whether that marker carries information about action outcomes? The basic principle is straightforward: vary kinematic and dynamic features while holding constant an action outcome; and, conversely, vary action outcomes while holding kinematic and dynamic features constant. In practice researchers have devised many ingenious ways to achieve this. In order to vary kinematic and dynamic features while holding action outcomes constant, in some studies a single action outcome is achieved using different effectors, hand, mouth or foot, say \citep{rizzolatti:1988_functional,Rizzolatti:2001ug,cattaneo:2010_state-dependent}. A variation on this approach is to contrast performing a grasping action with different tools, so that the same action outcome might require closing or opening the hand depending on the tool used \citep{umilta:2008pliers,cattaneo:2009_representation,rochat:2010_responses}. In order to vary action outcome while holding kinematic and dynamic features constant, researchers have contrasted grasping movements with different distal outcomes such as eating and placing \citep{Fogassi:2005nf,bonini:2010_ventral,cattaneo:2007_impairment}. Another approach is to contrast the same grasping movements performed in the presence or manifest absence of a target object \citep{Umilta:2001zr,villiger:2010_activity}. A related alternative is to contrast the same grasping movements in the presence of objects which could, or manifestly could not, be grasped by means of such movements \citep{koch:2010_resonance}. In each of these cases there is evidence that some markers of motor processing are correlated with action outcomes rather than narrowly kinematic or dynamic features of action.\footnote{ Of course some researchers have raised doubts concerning some of the evidence that there are markers of motor processing which carry information about action outcomes (e.g.\ \citealp{cavallo_grasping_2011,borroni_mirroring_2011}).  On balance, however, the evidence supports this conclusion.}

To say that motor processing involves information about action outcomes is not, of course, to say that there are motor representations of action outcomes.  To make the step from information to representation we have to show that information about action outcomes guides processing \citep[compare][]{Dretske:1988sq}. To this end we shall now consider how information about action outcomes is relevant in motor planning and monitoring, which are two of the functional roles of motor representation.

First take motor planning. This involves satisfying a variety of requirements. For example, in grasping a mug it is necessary for the hand to prefigure the shape of the mug, to move towards it avoiding potential obstacles and to reach it at a velocity that is both compatible with achieving the type of grip to be used and also suitable given features of the mug such as its fragility and weight \citep{Jeannerod:1995bb,jeannerod:1998nbo}. The need to plan sequences of actions, which may overlap, imposes further requirements. How one should reach for and then grasp a heavy frying pan (say) depends on what one will then do with it. One way of grasping it might be ideal for safely transporting its contents, another for emptying it. The many requirements on motor planning cannot normally be met by explicit practical reasoning, especially given the rapid and fluid transitions involved in many action sequences. Rather they require motor processes and representations.

Second, motor representations function in monitoring action. They provide inputs to internal predictive models that estimate likely effects of actions. Sensory feedback provides information about the actual course of the action which can be compared against the predictions. Adjustments are made in order to minimise the discrepancy between these \citep{wolpert:1995internal,miall:1996_forward}. Monitoring and control of this kind is a second functional role of motor representation.

These facts about the functional roles of motor representation in planning and monitoring action reveal that it would be advantageous for some such representations to represent action outcomes rather than merely patterns of joint displacement or muscle activation. In the case of planning, many of the requirements to be satisfied are partially dependent on action outcomes and not only on more narrowly kinematic or dynamic features of action. Representation of action outcomes could therefore play a role in simplifying planning processes.  Equally, monitoring involves predicting relatively distal effects of bodily movements including facts about the locations of objects. To make such predictions exclusively in terms of individual joint displacements or muscle activations would be computationally demanding as the human hand alone has over 20 degrees of freedom. Efficient prediction plausibly demands several higher-level representations of action including representations of action outcomes \citep{arbib:1985_coordinated,mason:2001_hand,santello:2002_patterns}. In short, it is clear that motor representations of action outcomes could be useful: in planning, because some requirements for efficient motor planning concern outcomes to which actions are directed; and in monitoring, because representations of outcomes could simplify prediction. Given that, as we saw, markers of motor processes carry information about action outcomes, it is reasonable to conclude from the relevance of action outcomes to the functional roles of motor representation that some motor representations are representations of action outcomes.

So far we have argued that motor processes involve representations of action outcomes. It remains to show that such representations ground purposive actions. But this is hardly a further step. How do intentions ground the purposiveness of actions?  On any standard view, an intention represents an outcome, causes an action, and does so in such a way as to facilitate the outcome's occurrence. Similarly, motor representations of outcomes represent action outcomes, play a role in generating actions, and do this in a way that normally facilitates the occurrence of the outcomes represented. To say that motor representations do all this is one way of making precise the metaphor involved in saying that purposive actions are directed to outcomes.  Moreover, there is a clear resemblance between the natural way of understanding intentions as grounding outcome-directedness and the way in which motor representations ground outcome-directedness.  So given the functional roles and representational objects of motor representations, there is little room for doubt that motor representations can ground the outcome-directedness of purposive actions.

To sum up, the first temptation was to suppose that motor representation has no bearing at all on our question about the relation between actions and the outcomes to which they are directed. One reason to resist this temptation is the fact that, as we have just argued, motor representations not only represent action outcomes but also ground the directedness of actions to outcomes.  This argument might be taken as grounds for giving in to a second temptation.

\section{A motor format for representation}

The second temptation is to suppose that motor representation, or some species of it, is a variety of intention, where intention is understood in the standard way as a propositional attitude with a characteristic role in practical reasoning \citep{Bratman:1987xw}. In this section we explain why this temptation should also be resisted.

As background we first need a generic distinction between content and format. Imagine you are in an unfamiliar city and are trying to get to the central station. A stranger offers you two routes. Each route could be represented by a distinct line on a paper map. The difference between the two lines is a difference in content. Each of the routes could alternatively have been represented by a distinct series of instructions written on the same piece of paper; these cartographic and propositional representations differ in format. The format of a representation constrains its possible contents. For example, a representation with a cartographic format cannot represent what is represented by sentences such as `There could not be a mountain whose summit is inaccessible.'\footnote{ Note that the distinction between content and format is orthogonal to issues about representational medium. The maps in our illustration may be paper map or electronic maps, and the instructions may be spoken, signed or written. This difference is one of medium.} The distinction between content and format is necessary because, as our illustration shows, each can be varied independently of the other.

Our aim in this section is to show that motor representations differ from intentions with respect to their format. This is why the second temptation must be resisted. That motor representations differ in format from intentions shows that they are genuinely distinct phenomena.\footnote{ Readers already convinced that motor representation differs from intention in being non-conceptual will not need the following argument in order to conclude that they are distinct phenomena. However the following considerations also show that motor and perceptual representations differ in format, which will be relevant in later when we consider how motor representations and intentions jointly lead to action.}

How can we distinguish representational formats? Because a representational format is typically associated with characteristic processing profiles, it is sometimes possible to infer similarities and differences in representational format from similarities and differences in the processes in which representations feature. This works both for artifactual and mental representations.  To illustrate in terms of our earlier example, suppose that two people have representations of the same route but for one person the route is represented by a line on a map (so in a cartographic format) whereas the other person has a propositional representation of the route. Some transformations are likely to be easier for the person with the cartographic route representation (depending on the projection used, of course); examples include reversing the route, determining how many times a certain river is crossed and transforming the route into a sequence of compass bearings. Other transformations, such as turning the route into a list of salient landmarks, may be easier for the person with the propositional route description. So some patterns of difference in the two people's performances may be explained by the difference in the formats of their representations. If we did not already know that the two people's route representations involved different formats, we might infer this from the facility with which each performed various transformations of the route.

Cognitive neuroscience frequently depends on inferences of just this type. Contrast imagining seeing an object moving with imagining moving an object. There is evidence that imagining seeing and imagining acting involve different processes. For instance, each can be selectively impaired \citep{sirigu:2011_motor}; and factors such as limb amputation or hand posture can selectively interfere with imagining performing an action \citep{nico:2004_left,vargas:2004s_influence,fourkas:2006_influence}. Importantly for our purposes, imagining seeing an object move and imagining moving an object also differ in ways that indicate differences in representational format. For instance, the way imagining acting unfolds in time is quite different from the way imagining seeing unfolds: how long the former takes is closely related to the time it would take to perform the action, but this is not true of the latter \citep{decety:1989_timing,decety:1996_imagined,Jeannerod:1994oz}. In addition, for object-related actions such as grasping the handle of a cup, manipulating the object in ways that make the action harder (such as orienting the handle to make it less convenient for you to grasp) make a corresponding difference to the effort involved in imagining performing the action \citep{parsons:1994_temporal,frak:2001_orientation}, whereas such manipulations would make no systematic difference to the effort involving in imaging seeing. The explanation for these differences between imagining acting and imagining seeing plausibly involves a difference in the formats of the representations involved.

How do these findings bear on motor representation? Contrast imagining an action with actually performing that action. There is substantial evidence that imagining an action involves many of the same processes which would be involved in performing that action up to the actual muscle contractions \citep{jeannerod:1995_mental,jeannerod:2003_mechanism}. There is also independent evidence that the representations underpinning imagining acting have the same format as those involved in actually acting. This evidence takes the form of similarities in processing profiles. Just as some similarities between imagining seeing something and actually seeing something can be explained on the hypothesis that the representations involved in imagining seeing and actually seeing have a common format \citep{kosslyn:1994_image}, so also some similarities between imagining grasping a cup and actually grasping it are evidence that the representations involved in imagining grasping are of the same format as those involved in actually grasping the cup. We can conclude, then, that imagining and executing actions are both underpinned by motor representations, and that motor representations differ in format from representations involved in both seeing and imagining seeing.

So far we have been arguing that motor and visual representations differ in format.  Why suppose that motor representations also differ in format from intentions?  As we noted earlier, the format of a representation constrains its possible contents. Just as a representation's having a cartographic format restricts its possible contents, so the motor format would imply restrictions on representations having it. In imagining executing actions such as grasping a cup, there are bio-mechanical, dynamical and postural constraints, among others, on what can be imagined. These constraints are closely related to those which govern actually performing such actions \citep{Jeannerod:2001yb} . Such constraints are plausibly explained by the representational format of motor representations. Since no such constraints would be expected to apply where propositional representations are involved, we can conclude (as is widely assumed) that the format of motor representations is not propositional.

To conclude that intention and motor representation are genuinely distinct phenomena it is not quite enough to know that motor representations are non-propositional, of course. In addition we must know that intentions are propositional. We take this claim to be a consequence of the role of intention in practical reasoning and of the fact that one can have intentions involving quantification and identity; for example, one can intend that one cross seven distinct bridges in 48 hours without yet specifying which bridges or hours.\footnote{ Of course some use the term `intention' for non-propositional representations involved in the execution and control of action.  This is a narrowly terminological issue.}

\section{The Interface Problem}

We have just argued for three claims. First, some motor representations represent outcomes (rather than, say, only bodily movements). Second, there are actions whose directedness to an outcome is grounded in motor representation. And third, motor representation differs from intention with respect to representational format. A consequence of these claims is that a single purposive action may involve representations of the outcomes to which it is directed in at least two different representational formats, motor and propositional. This contributes to a problem we call \emph{the interface problem}; in this section we explain how it arises.

Imagine that you are strapped to a spinning wheel facing near certain death as it plunges you into freezing water. To your right you can see a lever and to your left there is a button. In deciding that pulling the lever offers you a better chance of survival than pushing the button, you form an intention to pull the lever, hoping that this will stop the wheel. If things go well, and if intentions are not mere epiphenomena, this intention will result in your reaching for, grasping and pulling the lever. These actions---reaching, grasping and pulling---may be directed to specific outcomes in virtue of motor representations which guide their execution. It shouldn't be an accident that, in your situation, you both intend to pull a lever and you end up with motor representations of reaching for, grasping and pulling that very lever, so that the outcomes specified by your intention match those specified by motor representations. If this match between outcomes variously specified by intentions and by motor representations is not to be accidental, what could explain it?

The interface problem is the problem of answering this question, of explaining how intentions and motor representations, with their distinct representational formats, are related in such a way that, in at least some cases, the outcomes they specify non-accidentally match. But why think that this question poses a problem at all?

Let us start by putting the question more precisely.  First we should define the relevant notion of matching. Two collections of outcomes, A and B, \emph{match} in a particular context just if, in that context, either the occurrence of the A-outcomes would normally constitute or cause, at least partially, the occurrence of the B-outcomes or vice versa. To illustrate, one way of matching is for the B-outcomes to be the A-outcomes. Another way of matching is for the B-outcomes to stand to the A-outcomes as elements of a more detailed plan stand to those of a less detailed one.

Now we can put the question more generally. There are cases in which a particular action is guided both by one or more intentions and by one or more motor representations. In at least some such cases, the outcomes specified by the intentions match the outcomes specified by the motor representations. Furthermore, this match is not always accidental. How do non-accidental matches come about?

In principle one might try to explain the match by supposing that intentions and motor representations have a common cause. If the mere presence (or the mere perception) of a lever invariably triggered intentions and motor representations specifying grasping (say), it might be possible to explain matching in this way. But this sort of consideration cannot provide a full explanation of non-accidental matching for two reasons. First, intentions are not always triggered in straightforward ways by agents' environments or perceptions; to suppose otherwise is to ignore the very phenomena, decision and planning, which make intention so interesting.  Second, motor representations are also not always triggered in any straightforward way by agents' environments or perceptions either.  For these reasons there seems to be no hope of fully explaining matching by postulating common causes for intentions and motor representations.

If common cause explanations are ruled out, another natural approach is to  appeal to content-respecting causal processes. Perhaps, for example, intentions with certain contents (concerning grasping, say) reliably cause motor representations with corresponding contents (also concerning grasping, say). Alternatively we might suppose, very crudely, that some comparator process checks that the contents of motor representations are appropriate given what the agent intends.  Either way, the idea is to explain matches between outcomes specified by the contents of certain states in terms of content-respecting causal processes linking those states.

This type of explanation is arguably appropriate where the states in question have the same representational format.  For example, this type of explanation would arguably be appropriate if our aim were to explain matches between  large-scale intentions and the smaller-scale, more detailed intentions which serve as building blocks for them. But in fact we are concerned with intentions and motor representations which, as argued above, have different representational formats. This creates a potential difficulty.

In general, when two representations differ in format, postulating reliably content-respecting causal processes linking them requires us to explain how their contents are coordinated. To illustrate, suppose you are given some verbal instructions describing a route. You are then shown a representation of a route on a map and asked whether this is the same route that was verbally described. You are not allowed to find out by following the routes or by imagining following them. This puts you in something like the position of the comparator process envisaged above. Special cases aside, answering the question will involve a process of translation because two distinct representational formats are involved, propositional and cartographic. It is not be enough that you could follow either representation of the route. You will also need to be able to translate from at least one representational format into at least one other format. Similarly, for there to be reliably content-respecting causal processes linking intentions with motor representations there would have to be some process of translation.

But why is this a potential difficulty? What is wrong with postulating a process of translation? The difficulty is that nothing at all is known about this hypothetical translation between intention and motor representation, nor about how it might be achieved, nor even about how it might be investigated. Of course this doesn't show that we couldn't fully explain matching by appeal to content-respecting causal processes. But it does show that no such explanation is currently available.

This, then, is why our question about the interface between intentions and motor representations amounts to a problem. It is a problem because of two natural routes to answering the question, the first (appealing to common causes of intentions and motor representations) is a non-starter and the second (appealing to content-respecting causal processes) amounts to no more than a stab in the dark.  Our aim in what follows is to solve the interface problem without postulating either common causes or translation processes.

\section{Demonstrative and deferential action concepts}

The interface problem is the problem of explaining how, some of the time, there could be reliable matches between outcomes variously specified by intentions and motor representations. As the previous section explained, the problem arises because intentions and motor representations have different representational formats.

There is a way to link representations with different formats that requires neither common causes nor translations. To illustrate, imagine once again that we have two representations of a route, one propositional the other cartographic. But this time suppose that the propositional representation is simply `Follow this route!' where the demonstrative phrase `this route' refers to the route marked on the map. This instruction does not describe the route but merely defers to another representation of it. Because the representation deferred to is cartographic, comparing the instruction with the map no longer requires translation between representational formats. We shall suggest that something analogous holds concerning the relation between intention and motor representation. To anticipate what will be explained below, some intentions involve demonstrative concepts (or other constituents) which refer to actions by deferring to motor representations.

Let us first step back and consider thought generally. The existence of demonstrative thoughts and concepts is quite widely accepted. For instance, some philosophers have proposed that there are demonstrative colour concepts \citep{McDowell:1996yi,Brewer:1999ud}.  While there are objections to uses to which these have been put (e.g.\ \citealp{heck_jr._nonconceptual_2000,dokic_shades_2001}), the claim that there might be such concepts is barely controversial.

Consider the possibility that some action concepts are demonstrative. Someone who says to herself, `I wish I could do that too' may be entertaining a proposition which involves a demonstrative concept, \emph{that action}. Clearly this demonstrative concept cannot refer to a token action.  The wish was not to perform another agent's action but to perform an action of a certain type. Now the mere fact that the demonstrative concept must refer to a type is not a problem. As the case of colour concepts indicates, it is plausible that propositional attitudes can involve elements which demonstrate types and not just individuals \citep[see also][§3.4]{levine:2010_demonstrative}.

How could the existence of demonstrative concepts for actions be relevant  to solving the interface problem? In our earlier illustration, the instruction `follow this route!' succeeds in referring to a route by deferring to another representation with a different format. The instruction is about a route not a representation, but it succeeds in referring to a route by deferring to a representation of that route. Similarly, some demonstrative concepts may refer to types of action such as grasping or throwing by deferring to motor representations (see \citet{levine:2010_demonstrative} for one way of developing this idea). These demonstrative concepts would be concepts of actions not of motor representations, but they would succeed in being concepts of actions by deferring to motor representations. For any such concept, it is a motor representation which ultimately determines what it is a concept of.

The idea that some demonstrative concepts refer to actions by deferring to motor representations immediately raises a question. Could a demonstrative concept really defer to a motor representation? It seems clear that we can't select a motor representation to defer to in the same way that we can select a map when we say `Follow this route!'. After all, motor representations are not things we can point to with our hands. Nevertheless, it does seem that motor representations are available in some sense. To start with an analogy, consider pantomiming an action to yourself. You are rehearsing part of an operation which involves precisely grasping a delicate structure with some tweezers. Just as someone might point to a map and say `Follow that route!', so also they could point to your pantomime and say `Do that!'. In our analogy, the pantomime stands in for a motor representation of the action.  The demonstrative in `Do that!' refers to an action by deferring to the pantomime. Of course this analogy doesn't show that demonstrative concepts can defer to motor representations (at least not unless pantomimes are motor representations of actions). But now consider purely mental pantomime---that is, motor imagery. One might use motor imagery to explore different ways of completing a task and then, having hit on a good solution, think to oneself `Do that!' It seems possible that in some such cases the demonstrative refers by deferring to a motor representation of action involved in mental motor imagery.

But is this really possible? Someone might object that in appealing to motor imagery we are sneaking intentions in through the back door. How can we be sure that it is ever really a motor representation rather than an intention that one defers to in thinking `Do that!'?  Contrast two cases of motor imagery, both involving a tool. In the first case, imagine grasping an object with the tool; and in the second case, imagine releasing the same object with the tool. The two cases involve different imagery. But in what does this difference consist? Could it be a difference in intention that explains how the two cases of motor imagery differ?  Since it is possible to grasp without intentionally grasping, it is surely also possible to imagine grasping without imagining intentionally grasping.  And note that to imagine grasping without imagining intentionally grasping does not necessarily require one to imagine non-intentionally grasping. (It is easy to miss this point by confusing it with the issue of whether the imagining must itself be intentional. Whether or not it is possible to imagine grasping or releasing without \emph{intending to imagine} so acting, one can certainly imagine grasping or releasing without \emph{imagining intentionally} so acting.) So to rule out the possibility that the two cases of motor imagery differ with respect to intention, let the motor imagery be neutral with respect to intention. Is the difference between the two cases then due to differing movements or patterns of muscle contraction? To rule this possibility out, let the movements and muscle contractions involved in both cases be as similar as possible: let the difference between grasping and releasing the object be a matter of how the tool is configured rather than of how your body moves. With the contrast cases elaborated by these two stipulations, it is plausible that the difference in imagery between the two cases will be due to the different motor representations involved in grasping and releasing. So in these cases, the thought `Do that!' will refer to grasping (or releasing) by deferring to a motor image of grasping (or of releasing). Since motor images are a species of motor representation, it follows that demonstrative concepts can refer to actions by deferring to motor representations.

So far we have focused on motor imagery. We are not committed to claiming that demonstrative concepts can only refer to actions by deferring to motor images. Given the parallels between motor imagery and motor preparation \citep{Jeannerod:2001yb}, it is plausible that motor representations feature in experiences of action in such a way that motor actions are available for demonstrative reference. If so, the possibility of demonstrative concepts which refer to actions by deferring to motor representations is supported not just by motor imagery but also by experiences associated with motor activity more generally.

We have been arguing for the existence of demonstrative concepts which refer to actions by deferring to motor representations in order to solve the interface problem. As already stated, the interface problem is the problem of explaining how outcomes variously specified by intentions and motor representations, despite their different formats, sometimes non-accidentally match. As long as we think of intentions and motor representations as having logically independent contents, it seems that fully solving the problem would require appeal to processes of translation linking intention with motor representation. But where intentions involve demonstrative action concepts, their contents are not necessarily logically independent of the contents of motor representations. For a demonstrative component of an intention may refer to an action by deferring to a motor representation. Where this happens, which actions the intention specifies is partially or wholly determined by the motor representation, and so the interface problem is solved.

Or is it? So far we have not distinguished between two aspects of action outcomes.  Action outcomes often specify both a way of acting---whether to grasp or release, say---and also what to act on---on the mug or the pen, say.  Since we have been focusing on ways of acting rather than on what is to be acted on, it would be consistent with our arguments to accept that there are demonstrative concepts of ways of acting (such as grasping and reaching, perhaps) which defer by referring to motor representations while denying that any such demonstrative concepts are also about what is to be acted on. If this combination of views were correct, we would have only a partial solution to the interface problem. We would only have explained how the outcomes variously specified by intentions and motor representations non-accidentally match with respect to ways of acting, not with respect to what is to be acted on. However a range of behavioural and neurophysiological evidence shows that motor representations represent not only ways of acting but also objects on which actions might be performed and some of their features related to possible action outcomes involving them (for a review see \citealp{Gallese:2011uq}; for discussion see \citealp[pp.\ 410-3]{pacherie:2000_content}). For example visually encountering a mug sometimes involves representing features such as the orientation and shape of its handle in motor terms \citep{buccino:2009_broken,costantini:2010_where,cardellicchio:2011_space,Tucker:1998,tucker:2001_potentiation}. Reference by deference to motor representation could therefore explain how action outcomes match with respect not only to ways of acting but also with respect to what is to be acted on.\footnote{ In fact, there may be another way of explaining non-accidental matching with respect to what is to be acted on due to \citet[pp.\ 36-8, 44-5, 48-57]{Campbell:2002ge}. In outline, Campbell's idea is to note that intentions can refer to objects by means of perceptual demonstrative elements, and that perceptual and motor representations of objects may be sufficiently commensurate for a perceptually represented object to be reliably selected as a target for motor processes.  It would be consistent to hold both that a view like Campbell's correctly explains non-accidental matching of action outcomes with respect to what is to be acted on, and also that non-accidental matching with respect to ways of acting is explained by demonstrative action concepts which defer to motor representations. We shall leave open the question of whether non-accidental matching involves elements additional to those involved in the view we have been developing.  Our claim is just that a solution to the interface problem need not involve anything other than components of intentions which refer to actions by deferring to motor representations of action outcomes.}

Solving the interface problem may not always involve demonstrative concepts. What matters for solving the interface problem is deference, not demonstration. Suppose that someone acts on an intention to grasp the handle of a mug.  Suppose also that the outcome to which her action is directed, grasping the mug's handle, is specified by motor representations. As long as the concept of grasp involved in her intention refers by deferring to a motor representation of grasping, the two specifications of action outcomes (in intention and in motor representation) will not be independent and so the interface problem will be solved.

\section{Conclusion}

While most philosophers would probably agree that, as a matter of fact, intentional action often involves motor representation, this is typically treated as only an enabling condition for intentional action or as merely a variety of intention. In fact most philosophical theories of action apply indifferently to agents whose actions involve motor planning and monitoring and to (imaginary) agents who need explicit practical reasoning for each muscle contraction. In our view this is a defect of those theories. To fully understand intention, we need to understand how intentions relate to motor representations.  Or so we have argued.

The need to understand how intentions relate to motor representations follows from three claims defended in early sections of this paper. First, some motor representations represent outcomes (rather than, say, only bodily movements). Second, there are actions whose directedness to an outcome is grounded in motor representations. And, third, motor representations differ from intentions with respect to their representational format. These claims reveal that a single purposive action may involve representations of the outcomes to which it is directed in at least two different representational formats, motor and propositional. It is necessary to understand something of how intentions relate to motor representations in order to explain how the various representations involved in a single action sometimes non-accidentally specify matching outcomes.

The problem of explaining this we called \emph{the interface problem}. Theoretically uncomplicated approaches to solving it would involve appeal to common causes or to processes of translation. The interface problem is a problem because no such approach is viable given the present state of knowledge. However, we have shown that there is another way of solving the interface problem. This solution hinges on the existence of concepts which refer to actions by deferring to motor representations. Where intentions involve such deferential concepts, their contents are logically linked to those of motor representations and so the interface problem can be solved. In short, then, we have shown that there is a theoretically coherent and empirically plausible solution to the interface problem, one which requires neither common causes nor processes of translation to link intention and motor representation.

A further conjecture is that whenever there is a non-accidental match between the contents of intentions and of motor representations, this is due to deferential action concepts. The plausibility of this conjecture rests in part on the absence of any other viable solution to the interface problem, and of course we do not claim to have shown that no alternatives exists. If the conjecture is right, if there is no way of solving the interface problem other than by means of deferential action concepts, then concepts which refer by deferring to motor representations are what tie explicit practical reasoning to motor processes.

More speculatively still, we suggest that where an intention properly and reliably produces bodily movement, either acting on that intention involves a further intention or else the intention involves concepts which refer to actions by deferring to motor representations. If so, it is deferential action concepts that ultimately connect intentions to bodily movements. Only by recognising how intentions interlock with motor representations can we hope to understand how our intentions ever make a difference to the world around us.

On this view experience of action plays a novel role. Action experiences in which motor representations feature, such as those associated with motor imagery and those associated with really acting, are arguably necessary for there to be concepts which are constituents of intentions and refer to actions by deferring to motor representations. But if, as we conjecture, such deference is necessary for intentions to properly and reliably result in bodily movements, it may turn out that intentionally acting in the world depends on action experiences featuring motor representation. Much as on some views thought about objects depends on perceptual experience \citep[e.g.][]{Campbell:2002ge}, so also intending actions may depend on motor experience.

\bibliography{$HOME/endnote/phd_biblio}

\end{document}

### 2012-01-07 intention motor representation with corrado sinigaglia/with corrado motor intention action.tex
added intention motor action (was svn, older revisions not imported but still in svn)

%!TEX TS-program = xelatex
%!TEX encoding = UTF-8 Unicode

\documentclass[12pt,a4paper]{extarticle}
% extarticle is like article but can handle 8pt, 9pt, 10pt, 11pt, 12pt, 14pt, 17pt, and 20pt text

\def \ititle {Grasping a cup of coffee}
\def \isubtitle { }
\def \iauthor {}
\def \iemail{}
%\date{}

\input{$HOME/Documents/submissions/preamble_steve_paper}

\begin{document}

\setlength\footnotesep{1em}

\bibliographystyle{newapa} %apalike

\maketitle
%\tableofcontents

\begin{quote}
`our primitive actions, the ones we do not by doing something else, mere movements of the  body---these are all the actions there are.
We never do more than move our bodies.'
\citep[p.\ 59]{Davidson:1971fz}
\end{quote}

\section{Claims and terms}
\begin{enumerate}
\item A \emph{motor goal} is an action outcome that can be represented by the motor system as such.  For example, a motor goal might be to grasp the object there by its handle.

\item A \emph{motor-like act} is a unit of action corresponding to an act described by a motor goal.  For example, grasping the object there by its handle may be a motor-like act.

\item \emph{Motor-like acts are the basic elements in practical reasoning}

(i) correspondence---Motor-like acts are the basic units of action for practical reasoning in this sense: where acting on an intention involves bodily movement, acting on the intention requires agents to intend to perform motor-like acts which will realise the intention. (ii) representation---humans and other primates typically represent these basic units of action not as raw bodily movements but as activities with a goal.

Note: How to understand the relation between a description of an action and a description of an action in terms of motor goals?
What does one need to be able to represent the action in non-motor terms given that the agent is able to represent the action in motor terms?
Can a monkey plan in Bratman's sense of planning?

\item \emph{Motor-like acts are the basic elements in perceiving action}

(i) correspondence---Motor-like acts are the basic units of action for perceiving and understanding action in approximately the sense that physical objects are the basic units for perceiving.  (ii) representation---these basic units of action in terms of which actions are perceived and understood are not represented as raw bodily movements; they are represented as directed to objects and as having a goal.

(This conflicts with Csibra \& Gergeley on teleological stance as essence of action representation (although not that efficiency is entirely irrelevant, just that it's not basic).)

\item Relation to knowledge of own action (see O'Brien)

\end{enumerate}

\section{Problems}
\begin{enumerate}

\item encoding intentions for the motor system: how could practical reasoning result in bodily movements?  
Since agents often lack detailed knowledge of the bodily movements required for an action (such as the sequence of finger movements needed to grasp and post a letter), their intentions cannot directly specify the bodily movements.  
Equally, since motor cognition does not involve representations of actions under descriptions which are significant for practical reasoning (paying a bill, walking home, impressing Leslie), these cannot be transformed into bodily movements by the motor system.

\item parameters: how is it possible to rapidly identify perceived actions (e.g.\ a hand grasping a cup behind a screen) given that, construed as the movement of an articulated object, there are so many parameters to track?

\item imagination: How is it possible to imagine an action such as a hand grasping a cup behind a screen given that when construed as the movement of an articulated object this is analogous to imagining the individual movements of bees in a swarm?

\item segmentation: How is it possible to segment  actions into intention-relevant units, and to recover some of the structure of intentional action?  How are observers able to distinguish which movements are part of an action and which are extraneous to it?

\end{enumerate}

\section{*TODO}

Currently phrased in terms of motor goal is least specific representation.
Important that motor system represents actions at different levels of specificity.
From the point of view of the perceiver or agent, the representation of the action might be a composite of these many levels.
The number of parameters in terms of which an action can be described depends on the concurrent multi-level description of the action.  (E.g. grasp ON; left hand grasp OFF; right hand grasp ON; gives a pattern.)
NOT just a description at the least specific level.

READ `Cognition in action. A new look at the cortical motor system'

\section{Action from the point of view of practical reasoning}

Intentions are basic elements in practical reasoning which individuate actions.
Sam intends to get a sandwich.
His knowledge that the only nearby sandwich shop will shortly close together with his intention gives him a reason to extract himself from a conversation and walk to the shop.
At the shop he orders, pays for and collects a sandwich.
The whole episode, starting with Sam's extracting himself from the conversation and ending with Sam holding the sandwich, constitutes a single intentional action.
The episode constitutes a single intentional action because a single intention is responsible for the whole episode.  
The same pattern of movements might have occurred due to the execution of two distinct intentions (if, for example, Sam had first intended just to leave the building and only later decided to get a sandwich).
The intentional action can naturally be broken into five component episodes, the extraction, the walk, the order, the payment and the collection.  
These component episodes may overlap with each other (the walk might start before the extraction has ended).  
Each component episode is likely to be an intentional action in its own right; this is at least part of what makes it natural to break the whole into these five component episodes, and what explains why it would not be natural to break the action into sequences of bodily movement each lasting for exactly three seconds (say).
Some or all of these episodes might have further components which are themselves intentional actions.
Of course, the hierarchy of intentions can only have finite depth.  So given any intentional action, there will be a set of intentional actions where none of them has proper components which are themselves intentional actions and which jointly constitute the whole intentional action.
This set of intentional actions represents the most specific description of action that is directly involved in practical reasoning.

Let a \emph{minimally intentional action} be an intention action with no proper components which are themselves intentional actions.

\section{Action from the point of view of the motor system}

One of the motor system's roles is to control the activation of muscles which produce the bodily movements that are the ultimate physical realisation of many intentional actions.
Achieving this involves representing actions at different levels of specificity.
In the course of paying for his sandwich, Sam has to grasp one of several coins resting in his left hand and pass it to the vendor.
The grasping episode is represented at several levels of specificity.
A relatively specific representation specifies that the right hand will grasp the coin with a precision grip.
A less specific representation specifies that the coin will be grasped with a precision grip but not which effector will be used.  
If Sam were using the other hand, another effector with which he can execute precision grips or a pair of pliers, the less specific representation of the grasping episode would be the same.
The least specific representation involved in motor cognition specifies that the coin will be grasped but not how it will be grasped.
Simplifying, we can think of these least specific representations as individuated by three parameters: a type of action such as tearing, grasping and pressing; zero or more targets; and features of each target (if any) relevant to executing the action such as its size, rigidity and mass.

[This picture is complicated by representations of grasping-to-eat and grasping-to-place, which are less specific that representations of precision-grip-grasping with respect to the motor act type but more specific with respect to the motor action type.  For this reason we need to avoid talking about `\emph{the} least specific representations']

A \emph{motor goal} is an action outcome that can be represented by the motor system as such. 
To describe actions in terms of motor goals is to describe actions as the motor system might represent them.
A \emph{motor intention} is a representation of a motor goal by the motor system.
A \emph{motor act} is an action resulting from the execution of motor intention.

Two points are vital for what follows.
First, the motor system represents actions as graspings, tearings and pullings.  These representations are effector agnostic even to the extent of failing to specify whether the action will be performed with a hand, mouth or tool.  
They are highly abstract relative to both patterns of bodily movement which realise these actions and to the patterns of muscle activation responsible for producing them.
Second, the motor system does not represent actions in terms of many of the properties that are significant to agents.  An agent's intention may be to get to her office, or to his home; the motor system cares about places but it doesn't care whose office a place is.
Apparently, then, it is possible for there to be a gap between what is specified by the most specific intentions involved in practical reasoning and what is specified by the least specific motor intentions.

\section{From intentions to motor intentions}
For many intentions, their execution will involve motor actions (possible exceptions include intending to abstain from eating).  
This means that the intentions formed through processes of practical reasoning will have to cause the execution of appropriate motor intentions.
It is easy to understand how this could happen where intentions are very closely related to possible motor intentions.  
For example, an intention to grasp a coin at a particular perceived location already contains most of the information that would be needed to specify the content a motor intention: it contains the action type (grasp) and an identifier for the target (its location).
By contrast, much more is required if we compare an intention that is less closely related to any possible motor intention.
For example, there is no simple way to transform an intention to order a sandwich into motor intentions whose execution could realise the intention.

So how are intentions translated into motor intentions?  
Two possibilities:
\begin{enumerate}
\item Practical reasoning continues to generate means until it reaches intentions whose contents can be directly mapped to motor intentions.  So minimal intentional actions involve intentions whose contents are, or can be simply mapped onto, motor goals.
\item There is a process distinct from both practical reasoning and from motor cognition which normally transforms intentions into motor intentions.
\end{enumerate}
%
Against the first alternative: on the face of it, it seems that we can intend simply to pay for a sandwich without having more fine-grained intentions.

In favour of the first alternative: (a) novice actors do have to intend each motor act; (b) when conditions are not normal it also becomes necessary to intend each motor act; (c) the existence of chained motor representations (as involved in grasping-to-eat) means that experienced actors can intend whole sequences of motor acts; (d) describing an action in terms of episodes corresponding to motor acts comes naturally.  
This suggests that the intentions concerning motor acts may be present in many cases even though they are not particularly salient.

Compare the idea that in uttering a sentence it is necessary to have an intention concerning each word.
This is arguably not how it always seems when speaking, but it sometimes seems that we utter each word intentionally and this is largely how novice speakers begin (familiar phrases and songs aside).
But aside from the use of well-worn phrases, it is plausible that each word is intended.
This is why malaprop and related phenomena are so striking.
They are failures to act in accordance with an intention

Conjecture:
\begin{quote}
There is a normative requirement on intention: an intention to do something commits one to intending a course of motor-like actions which realise the intention.  (The intuitive idea was: Motor goals are the interface between deliberation and motor cognition.  This is an attempt to get part of that idea.)
\end{quote}
%

\section{Next}

Analogy between motor goals and words; the contents of intentions and sentences

Motor goals are the atoms in terms of which we perceive others' actions

Alt.: perceiving actions in terms of bodily movements.  How to characterise these?  Abstract view of body.  Just take hand and fingers.  23 parameters to track.  25 dimensional space to track movement of hand.  Description in terms of motor actions vastly reduces number of parameters that need to be tracked.  Plus there is a connection between the affordance of the object and the shape of the hand.
I.e. it's possible to represent an action very economically by identifying a small number of motor parameters.

Key point is that there is not just intentional action and continuously bodily movement.
Rather there is an (several) intermediate description.

\section{Notes}

Mele: How deeply does the action representation need to go into the motor details?  
This is just an empirical question. 
Corrado: It's not just an empirical question.  
The question has to do with the fact that we can individuate components of action which allow us to represent a goal as such.  
Non-motor goals can be described in terms of movements only if (a) we describe movement, which is practically impossible; or (b) we describe them as composites of motor goals.

Imagining acting: it's impossible to imagine grasping an object, but possible to imagine grasping it with a hand.   
Imagine that you are grasping my cup of coffee.
Motor imagery immediately characterises the action.
It's immediately obvious both that precision grip and whole-hand grip have something in common.  
E.g. You say grasp the cup.
I grasp it with a precision grip.
You say, Not in this way.
I grasp it with a whole-hand grip.
The pattern of possible changes is immediately obvious to me.
How can that be?
Correspondence between category structure implicit in motor representations of action and category structure implicit in ordinary thinking about action.
I.e. motor system representations of action have hierarchical structure in the sense that a precision grip is represented as a grasp.

Violation of expectations.

Distinguish motor repertoire from being in a position to act or interact. (...)

\bibliography{$HOME/endnote/phd_biblio}

\end{document}

### 2012-01-07 joint action and knowing others minds/interacting mindreaders.2.tex
added intention motor action (was svn, older revisions not imported but still in svn)

@@
 
-\documentclass[14pt,a4paper]{extarticle}
+\documentclass[12pt,a4paper]{extarticle}
 % extarticle is like article but can handle 8pt, 9pt, 10pt, 11pt, 12pt, 14pt, 17pt, and 20pt text
@@
 \def \isubtitle {}
-\def \iauthor {Stephen A. Butterfill}
-\def \iemail{s.butterfill@warwick.ac.uk}
-%\date{}
+%\def \iauthor {Stephen A. Butterfill}
+%\def \iemail{s.butterfill@warwick.ac.uk}
+\def \iauthor {}
+\def \iemail{}
+\date{}
 
 \input{$HOME/Documents/submissions/preamble_steve_paper2}
+\author{}
 
@@
 
-\noindent
-Word count: 10 200
+%\noindent
+%Word count: 10 100
 
@@
 
-\section{A question about the evidential basis of mindreading}
+\section{On the evidential basis of mindreading}
 \label{sec:intro}
@@
 %They are clearly different because one could occur while another did not;
-%one of us might have broken an arm and been unable to pullwhile the others succeeded.
+%one of us might have broken an arm and been unable to pull while the others succeeded.
 %So in this case, freeing the turnip is not a distributive goal of our actions.
@@
 Exploiting these indicators does not typically depend on knowing the particular contents of any beliefs, desires or goals.  
-Expressions, gestures and movements can naturally indicate imminent jointness in much the way thay can also naturally indicate emotions.\footnote{
+Expressions, gestures and movements can naturally indicate imminent jointness in much the way they can also naturally indicate emotions.\footnote{
 Ideas along these lines are suggested by the discussion of {emergent coordination} in \citet{Knoblich:2010fk}.
@@
 the possibility of interaction means that
-Ayesha can knowledgably identify the goal of your action.
+Ayesha can knowledgeably identify the goal of your action.
 

### 2012-02-23 intention motor representation with corrado sinigaglia/joint action version/intention and motor representation joint action version.tex
started on talk

--- 2012-01-07 intention motor representation with corrado sinigaglia/intention and motor action.tex
@@
 
+%a4paper or letterpaper (also used in preamble_steve_paper2
 \def \papersize {a4paper}
@@
 
-\def \ititle {Intention and motor representation:}
-\def \isubtitle {how to resist temptation}
-\def \iauthor {S. Butterfill* \& C. Sinigaglia**
+\def \ititle {Shared Intention and Social Aspects of Motor Representation, or How to Avoid Exoticism}
+\def \isubtitle {}
+\def \iauthor {
+S. Butterfill* \& C. Sinigaglia**
 \\ 
@@
 \\ 
-***Dipartimento di Filosofia, Università degli Studi di Milano}
+***Dipartimento di Filosofia, Università degli Studi di Milano
+}
 \def \iemail{s.butterfill@warwick.ac.uk}
-%\date{}
 
@@
 \begin{abstract}
-Are there distinctive roles for intention and motor representation in explaining the purposiveness of action?
-Standard accounts of action assign a role to intention but are silent on motor representation.
-The temptation is to suppose that nothing need be said here because motor representation is either only an enabling condition for purposive action or else merely a variety of intention.
-This paper provides reasons for resisting temptation.
-Some motor representations, like intentions, coordinate actions in virtue of representing outcomes; but, unlike intentions, motor representations cannot feature as premises or conclusions in practical reasoning.
-This implies that motor representation has a distinctive role in explaining the purposiveness of action.
-It also gives rise to a problem: were the roles of intention and motor representation entirely independent, this would impair effective action.
-Resisting temptation therefore requires explaining how intentions interlock with motor representations.
-The solution, we argue, is to recognise that the contents of intention can be partially determined by the contents of motor representations.
-Understanding this content-determining relation enables better understanding how intentions are related to actions.
+\noindent
+Are there distinctive roles for shared intention and social motor representation in explaining the purposiveness of joint action?
+This question needs some background.
+A joint action is a purposive action, or something resembling one, comprising two or more agents' actions.
+What is the relation between a joint action and the outcome (or outcomes) to which it is directed?
+Standard answers to this question appeal to a notion of shared intention and ignore social motor representation.
+It is tempting (for some philosophers, at least) to suppose that social motor representations are only enabling conditions for joint action, or else to suppose that social motor representation is merely a variety of shared intention.
+The talk will provide reasons for rejecting both suppositions.
+The truth may be that shared intention and social motor representation have distinctive roles in explaining the purposiveness of joint action.
+This gives rise to a challenge.
+On one hand, effective joint action---imagine erecting a tent in a gale together---sometimes requires both shared intentions and social motor representations plus a certain kind of harmony between the two.
+On the other hand, recognizing their distinctive roles precludes the existence of direct inferential links between shared intentions and social motor representations.
+The challenge is to explain how these two kinds of representation could sometimes harmoniously contribute to effective joint action despite the lack of inferential integration.
+
 \end{abstract}
@@
 
-What is the relation between a purposive action and the outcome or outcomes to which it is directed?
-The standard way of answering this question appeals to intention, a propositional attitude which plays a characteristic role in planning and coordinating action, is linked to practical reasoning and is subject to characteristic norms (\citealp{Bratman:1987xw}).
-On the standard view, an action is directed to an outcome in virtue of the action's being appropriately related to an intention which represents this outcome or some related outcome.
-As this view is usually expounded, the relation between actions and outcomes to which they are directed is treated as largely independent of the motor processes and representations underpinning action execution.
+A purposive joint action is an event involving two or more agents' actions where the actions are collectively directed to some outcome.
+What is the relation between a joint action and the outcome or outcomes to which it is directed? 
+The standard way of answering this question appeals to shared intention,
+a structure of states which stands to joint action roughly as an ordinary, individual intention stands to an ordinary, individual action.  
+On the standard view, a joint action is directed to an outcome in virtue of the joint action's being appropriately related to a shared intention which represents this outcome or some related outcome. 
+As this view is usually expounded, the relation between joint actions and outcomes to which they are directed is treated as largely independent of the motor processes and representations underpinning action execution.
 Motor representations are usually considered either as philosophically irrelevant enabling conditions, or else as merely filling in additional details of the basic schema provided by the standard story.
-Our aim is to show that this is a mistake.
-We shall argue that, once some basic features of motor representations are properly understood, the standard view must be  refined and extended in ways that allow us to recognize distinctive roles for intention and motor representation in explaining the purposiveness of action.
-
-Twin temptations stand in our way.
-The first temptation is to suppose that motor representation, although significant  for action execution, has no role in grounding the outcome-directedness of action.
-Surrendering to this temptation might be reasonable if all motor representations were about merely kinematic or dynamic features of actions.
-But some motor representations do represent action outcomes and some actions are directed to those outcomes in virtue of the guiding role of motor representations, or so we shall argue.
-This reason to resist the first temptation throws us straight into the arms of a second temptation, the temptation to suppose that where motor representations represent action outcomes, they are intentions.
-Surrendering to this second temptation would be reasonable if motor representations were like intentions in that they could feature as premises or conclusions in practical reasoning, or if there were some other planning processes in which both intention and motor representation could feature.
-However, as we shall argue, this is impossible due to differences between the propositional format of intentions and the distinctively motor, non-propositional format of motor representations.
-In short, we must resist both temptations because some motor representations are like intentions in representing action outcomes while also remaining sufficiently unlike intentions in that no single planning process can integrate both intention and motor representation.
-
-While there are good reasons to resist both temptations (or so we shall argue), the recognition that motor representation can ground the outcome-directedness of purposive actions independently of intention gives rise to a problem.
-For a single action, which outcomes it is directed to may be multiply determined by an intention and, seemingly independently, by a motor representation.
-Unless such intentions and motor representations are to pull an agent in incompatible directions, which would typically impair action execution, there are requirements concerning how the outcomes they represent must be related to each other.
-The problem is to explain how any such requirements could be met non-accidentally; we call this `the interface problem' .
-The key to solving this problem, we shall suggest, is to recognise that intentions can have constituents which refer to outcomes by deferring to motor representations of those outcomes.
-Effective action sometimes requires that the contents of intentions and of motor representations interlock, and this interlocking occurs when the contents of intentions are determined in part by the contents of motor representations.
-So whereas discussions of intention frequently ignore motor aspects of action, it turns out that understanding how intentions interface with motor representations is required for fully understanding how intentions are related to actions.
+Our aim is to show that this is a mistake:
+the standard view must be  refined and extended  to accommodate distinctive roles for intention and motor representation in explaining the purposiveness of joint action.
 
@@
 
-Our overall aim is to show that intention and motor representation play distinctive roles in explaining the purposiveness of action, that is, in relating actions to outcomes to which they are directed.
-The first temptation standing in our way is the temptation to suppose that motor representation, although perhaps important for action execution in some agents, has no bearing at all on our question about the relation between actions and the outcomes to which they are directed.
+Twin temptations stand in our way.
+The first  temptation is to suppose that motor representation has no bearing at all on our question about the relation between actions and the outcomes to which they are directed.
 Just as it would be an error to suppose that details of musculoskeletal structure are relevant to this question, so equally---so the temptation---it would be an error to suppose that facts about motor representation are relevant here.
@@
 Surrendering to this temptation might be reasonable if all motor representations represented only kinematic or dynamic features of actions, such as mere joint displacements or muscle contractions.
-However, we shall argue in this section that some motor representations represent action outcomes such as grasping, tearing or throwing.
-Furthermore, as we shall go on to argue (also in this section), such representations ground purposive actions.
-This in outline is why the first temptation should be resisted.
-
-How does representing an outcome differ from representing merely kinematic or dynamic features of action?
-How, for instance, does a motor representation of grasping (assuming for now that such a thing exists) differ from a representation of a sequence of joint displacements?
-First, a motor representation of grasping captures something common to many different sequences of joint displacements and postures involving a variety of effectors.
-To illustrate, an agent might grasp an object with her hand, with her mouth, with normal pliers (where grasping requires closing the hand), or with reverse pliers (where grasping requires opening the hand).
-A motor representation of grasping identifies something common to all these cases.
-Second, and conversely, a motor representation of grasping potentially distinguishes between the same sequence of joint displacements in different contexts.
-For instance, how a grasping action is represented may depend on a relatively distal outcome: on whether (say) it results in the object grasped being eaten or placed.
-Also the joint displacements which realise grasping in one context might in another context realise a different action, such as scratching; or they might fail to realise any action at all (because there is no target object, say).
-Third, a motor representation of an action outcome locates the action within a hierarchical structure.
-To illustrate, a left-handed precision grip and a right-handed whole hand grip are different action outcomes and both instances of grasping.\footnote{ On the notion of action outcome in motor representation, see \citet{jeannerod_motor_2006} and \citet{rizzolatti_mirrors_2008}.}
+However,
+ some motor representations represent action outcomes such as grasping, tearing or throwing.\footnote{ On the notion of action outcome in motor representation, see \citet{jeannerod_motor_2006} and \citet{rizzolatti_mirrors_2008}.}
+ Furthermore, as we shall go on to argue, some such representations ground purposive actions. This in outline is why the first temptation should be resisted.
 
@@
 The basic principle is straightforward: vary kinematic and dynamic features while holding constant an action outcome; and, conversely, vary action outcomes while holding kinematic and dynamic features constant.
-In practice researchers have devised many ingenious ways to achieve this.
 In order to vary kinematic and dynamic features while holding action outcomes constant, in some studies a single action outcome is achieved using different effectors, hand, mouth or foot, say \citep{rizzolatti:1988_functional,Rizzolatti:2001ug,cattaneo:2010_state-dependent}.
@@
 Another approach is to contrast the same grasping movements performed in the presence or manifest absence of a target object \citep{Umilta:2001zr,villiger:2010_activity}.
-A related alternative is to contrast the same grasping movements in the presence of objects which could, or manifestly could not, be grasped by means of such movements \citep{koch:2010_resonance}.
-In each of these cases there is evidence that some markers of motor processing are correlated with action outcomes rather than narrowly kinematic or dynamic features of action.\footnote{ Of course some researchers have raised doubts concerning some of the evidence that there are markers of motor processing which carry information about action outcomes (e.g.\ \citealp{cavallo_grasping_2011,borroni_mirroring_2011}).
+%A related alternative is to contrast the same grasping movements in the presence of objects which could, or manifestly could not, be grasped by means of such movements \citep{koch:2010_resonance}. 
+In each of these cases some markers of motor processing are correlated with action outcomes rather than narrowly kinematic or dynamic features of action.\footnote{ Of course some researchers have raised doubts  (e.g.\ \citealp{cavallo_grasping_2011,borroni_mirroring_2011}).
 On balance, however, the evidence supports this conclusion.}
 
-To say that motor processing involves information about action outcomes is not, of course, to say that there are motor representations of action outcomes.
+So far we have not explicitly distinguished two aspects of action outcomes.  
+Action outcomes often specify both a way of acting---whether to grasp or release, say---and also what to act on---on the mug or the pen, say.  
+We have been focusing on ways of acting. 
+ However a range of behavioural and neurophysiological evidence shows that motor processes also carry information about objects on which actions might be performed.\footnote{
+See \citet{buccino:2009_broken,costantini:2010_where,cardellicchio:2011_space,Tucker:1998,tucker:2001_potentiation}.
+For a review see \citet{Gallese:2011uq}; for discussion see \citet[pp.\ 410-3]{pacherie:2000_content}. 
+}
+
+To say that motor processes involve \emph{information} about action outcomes is not, of course, to say that there are motor \emph{representations} of action outcomes.
 To make the step from information to representation we have to show that information about action outcomes guides processing \citep[compare][]{Dretske:1988sq}.
-To this end we shall now consider how information about action outcomes is relevant in motor planning and monitoring, which are two of the functional roles of motor representation.
 
-First take motor planning.
-This involves satisfying a variety of requirements.
+To this end  consider how information about action outcomes is relevant in motor planning, a functional role of motor representation.
+Motor planning involves satisfying a variety of requirements.
 For example, in grasping a mug it is necessary for the hand to prefigure the shape of the mug, to move towards it avoiding potential obstacles and to reach it at a velocity that is both compatible with achieving the type of grip to be used and also suitable given features of the mug such as its fragility and weight \citep{Jeannerod:1995bb,jeannerod:1998nbo}.
@@
 One way of grasping it might be ideal for safely transporting its contents, another for emptying it.
-The many requirements on motor planning cannot normally be met by explicit practical reasoning, especially given the rapid and fluid transitions involved in many action sequences.
+Requirements on motor planning could not normally be met by explicit practical reasoning, especially given the rapid and fluid transitions involved in many action sequences. 
 Rather they require motor processes and representations.
-
-Second, motor representations function in monitoring action.
-They provide inputs to internal predictive models that estimate likely effects of actions.
-Sensory feedback provides information about the actual course of the action which can be compared against the predictions.
-Adjustments are made in order to minimise the discrepancy between these \citep{wolpert:1995internal,miall:1996_forward}.
-Monitoring and control of this kind is a second functional role of motor representation.
-
-These facts about the functional roles of motor representation in planning and monitoring action reveal that it would be advantageous for some such representations to represent action outcomes rather than merely patterns of joint displacement or muscle activation.
-In the case of planning, many of the requirements to be satisfied are partially dependent on action outcomes and not only on more narrowly kinematic or dynamic features of action.
-Representation of action outcomes could therefore play a role in simplifying planning processes.
-Equally, monitoring involves predicting relatively distal effects of bodily movements including facts about the locations of objects.
-To make such predictions exclusively in terms of individual joint displacements or muscle activations would be computationally demanding as the human hand alone has over 20 degrees of freedom.
-Efficient prediction plausibly demands several higher-level representations of action including representations of action outcomes \citep{arbib:1985_coordinated,mason:2001_hand,santello:2002_patterns}.
-In short, it is clear that motor representations of action outcomes could be useful: in planning, because some requirements for efficient motor planning concern outcomes to which actions are directed; and in monitoring, because representations of outcomes could simplify prediction.
-Given that, as we saw, markers of motor processes carry information about action outcomes, it is reasonable to conclude from the relevance of action outcomes to the functional roles of motor representation that some motor representations are representations of action outcomes.
+Since 
+many of the requirements to be satisfied by motor planning
+are partially dependent on action outcomes and not only on  narrowly kinematic or dynamic features of action,
+it would be advantageous for some  motor representations to represent action outcomes. 
+Given that, as we saw, markers of motor processes carry information about action outcomes, it is now reasonable to conclude that some motor representations are representations of action outcomes.
 
@@
 On any standard view, an intention represents an outcome, causes an action, and does so in such a way as to facilitate the outcome's occurrence.
-Similarly, motor representations of outcomes represent action outcomes, play a role in generating actions, and do this in a way that normally facilitates the occurrence of the outcomes represented.
+Similarly, some motor representations  represent action outcomes, play a role in generating actions, and do this in a way that normally facilitates the occurrence of the outcomes represented.
 To say that motor representations do all this is one way of making precise the metaphor involved in saying that purposive actions are directed to outcomes.
-Moreover, there is a clear resemblance between the natural way of understanding intentions as grounding outcome-directedness and the way in which motor representations ground outcome-directedness.
-So given the functional roles and representational objects of motor representations, there is little room for doubt that motor representations can ground the outcome-directedness of purposive actions.
 
-To sum up, the first temptation was to suppose that motor representation has no bearing at all on our question about the relation between actions and the outcomes to which they are directed.
-One reason to resist this temptation is the fact that, as we have just argued, motor representations not only represent action outcomes but also ground the directedness of actions to outcomes.
-This argument might be taken as grounds for giving in to a second temptation.
+To sum up so far, the first temptation was to suppose that motor representation has no bearing at all on our question about the relation between actions and the outcomes to which they are directed. 
+This temptation must be resisted because
+motor representations represent action outcomes and ground the directedness of actions to outcomes.  
+This fact about motor representation is why a second, complementary temptation is appealing ...
 
@@
 The second temptation is to suppose that motor representation, or some species of it, is a variety of intention, where intention is understood in the standard way as a propositional attitude with a characteristic role in practical reasoning \citep{Bratman:1987xw}.
-In this section we explain why this temptation should also be resisted.
+Because our claim is that intention and motor representation have distinctive roles in explaining action (§1 on your handout),
+we need to show that this temptation should also be resisted.
 
@@
 Each of the routes could alternatively have been represented by a distinct series of instructions written on the same piece of paper; these cartographic and propositional representations differ in format.
-The format of a representation constrains its possible contents.
-For example, a representation with a cartographic format cannot represent what is represented by sentences such as `There could not be a mountain whose summit is inaccessible.'\footnote{ Note that the distinction between content and format is orthogonal to issues about representational medium.
-The maps in our illustration may be paper map or electronic maps, and the instructions may be spoken, signed or written.
-This difference is one of medium.} The distinction between content and format is necessary because, as our illustration shows, each can be varied independently of the other.
+The distinction between content and format is necessary because, as our illustration shows, each can be varied independently of the other.
 
-Our aim in this section is to show that motor representations differ from intentions with respect to their format.
+Our aim now is to show that motor representations differ from intentions with respect to their format.
 This is why the second temptation must be resisted.
-That motor representations differ in format from intentions shows that they are genuinely distinct phenomena.\footnote{ Readers already convinced that motor representation differs from intention in being non-conceptual will not need the following argument in order to conclude that they are distinct phenomena.
-However the following considerations also show that motor and perceptual representations differ in format, which will be relevant in later when we consider how motor representations and intentions jointly lead to action.}
+That motor representations differ in format from intentions shows that they are genuinely distinct phenomena.
 
@@
 Because a representational format is typically associated with characteristic processing profiles, it is sometimes possible to infer similarities and differences in representational format from similarities and differences in the processes in which representations feature.
-This works both for artifactual and mental representations.
-To illustrate in terms of our earlier example, suppose that two people have representations of the same route but for one person the route is represented by a line on a map (so in a cartographic format) whereas the other person has a propositional representation of the route.
-Some transformations are likely to be easier for the person with the cartographic route representation (depending on the projection used, of course); examples include reversing the route, determining how many times a certain river is crossed and transforming the route into a sequence of compass bearings.
-Other transformations, such as turning the route into a list of salient landmarks, may be easier for the person with the propositional route description.
-So some patterns of difference in the two people's performances may be explained by the difference in the formats of their representations.
-If we did not already know that the two people's route representations involved different formats, we might infer this from the facility with which each performed various transformations of the route.
-
-Cognitive neuroscience frequently depends on inferences of just this type.
-Contrast imagining seeing an object moving with imagining moving an object.
-There is evidence that imagining seeing and imagining acting involve different processes.
-For instance, each can be selectively impaired \citep{sirigu:2011_motor}; and factors such as limb amputation or hand posture can selectively interfere with imagining performing an action \citep{nico:2004_left,vargas:2004s_influence,fourkas:2006_influence}.
+Cognitive neuroscience frequently depends on inferences of just this type. Contrast imagining seeing an object moving with imagining moving an object. 
+Imagining seeing and imagining acting involve different processes
+ \citep{sirigu:2011_motor,nico:2004_left,vargas:2004s_influence,fourkas:2006_influence}.
 Importantly for our purposes, imagining seeing an object move and imagining moving an object also differ in ways that indicate differences in representational format.
@@
 There is also independent evidence that the representations underpinning imagining acting have the same format as those involved in actually acting. 
-This evidence takes the form of similarities in processing profiles.
-Just as some similarities between imagining seeing something and actually seeing something can be explained on the hypothesis that the representations involved in imagining seeing and actually seeing have a common format \citep{kosslyn:1994_image}, so also some similarities between imagining grasping a cup and actually grasping it are evidence that the representations involved in imagining grasping are of the same format as those involved in actually grasping the cup.
 We can conclude, then, that imagining and executing actions are both underpinned by motor representations, and that motor representations differ in format from representations involved in both seeing and imagining seeing.
@@
 So far we have been arguing that motor and visual representations differ in format.  
-Why suppose that motor representations also differ in format from intentions?
-As we noted earlier, the format of a representation constrains its possible contents.
-Just as a representation's having a cartographic format restricts its possible contents, so the motor format would imply restrictions on representations having it.
-In imagining executing actions such as grasping a cup, there are bio-mechanical, dynamical and postural constraints, among others, on what can be imagined.
-These constraints are closely related to those which govern actually performing such actions \citep{Jeannerod:2001yb} .
-Such constraints are plausibly explained by the representational format of motor representations.
-Since no such constraints would be expected to apply where propositional representations are involved, we can conclude (as is widely assumed) that the format of motor representations is not propositional.
-
-To conclude that intention and motor representation are genuinely distinct phenomena it is not quite enough to know that motor representations are non-propositional, of course.
-In addition we must know that intentions are propositional.
-We take this claim to be a consequence of the role of intention in practical reasoning and of the fact that one can have intentions involving quantification and identity; for example, one can intend that one cross seven distinct bridges in 48 hours without yet specifying which bridges or hours.\footnote{ Of course some use the term `intention' for non-propositional representations involved in the execution and control of action.
+But what we need to show is that motor representations differ in format from intentions.
+Why accept this?
+We suppose that intentions are propositional. 
+This is because of their role in practical reasoning and of the fact that one can have intentions involving quantification and identity; for example, one can intend that one cross seven distinct bridges in 48 hours without yet specifying which bridges or hours.\footnote{ Of course some use the term `intention' for non-propositional representations involved in the execution and control of action.
 This is a narrowly terminological issue.}
+But we can also be sure that motor representations differ from propositional representations at least as much as they differ from visual representations given what we have just seen of the characteristic processing profile associated with them.
+So no motor representation could be an intention.
 
@@
 And third, motor representation differs from intention with respect to representational format.
-A consequence of these claims is that a single purposive action may involve representations of the outcomes to which it is directed in at least two different representational formats, motor and propositional.
-This contributes to a problem we call \emph{the interface problem}; in this section we explain how it arises.
+A consequence is that a single purposive action may involve representations of the outcomes to which it is directed in at least two different representational formats, motor and propositional.
+This contributes to a problem we call \emph{the interface problem} ...
 
-Imagine that you are strapped to a spinning wheel facing near certain death as it plunges you into freezing water.
-To your right you can see a lever and to your left there is a button.
-In deciding that pulling the lever offers you a better chance of survival than pushing the button, you form an intention to pull the lever, hoping that this will stop the wheel.
+Imagine that you are strapped to a spinning wheel facing near certain death as it plunges you into freezing water. To your right you can see a lever and to your left there is a button. 
+You form an intention to pull the lever, hoping that this will stop the wheel.
 If things go well, and if intentions are not mere epiphenomena, this intention will result in your reaching for, grasping and pulling the lever.
 These actions---reaching, grasping and pulling---may be directed to specific outcomes in virtue of motor representations which guide their execution.
-It shouldn't be an accident that, in your situation, you both intend to pull a lever and you end up with motor representations of reaching for, grasping and pulling that very lever, so that the outcomes specified by your intention match those specified by motor representations.
-If this match between outcomes variously specified by intentions and by motor representations is not to be accidental, what could explain it?
+It might not be an accident that you both intend to pull a lever and you end up with motor representations of reaching for, grasping and pulling that very lever, so that the outcomes specified by your intention match those specified by motor representations.
+If this match between outcomes variously specified by intentions and by motor representations is not  accidental, what explains it?
 
 The interface problem is the problem of answering this question, of explaining how intentions and motor representations, with their distinct representational formats, are related in such a way that, in at least some cases, the outcomes they specify non-accidentally match.
-But why think that this question poses a problem at all?
+(This question is more precisely stated on your handout.\footnote{Let us say that two collections of outcomes, A and B, \emph{match} in a particular context just if, in that context, either the occurrence of the A-outcomes would normally constitute or cause, at least partially, the occurrence of the B-outcomes or vice versa.
+In some cases (i) a particular action is guided both by one or more intentions and by one or more motor representations,
+(ii) the outcomes specified by the intentions match the outcomes specified by the motor representations, and
+(iii) this match is not accidental. 
+The question is, How do non-accidental matches come about?
+})
 
-Let us start by putting the question more precisely.
-First we should define the relevant notion of matching.
-Two collections of outcomes, A and B, \emph{match} in a particular context just if, in that context, either the occurrence of the A-outcomes would normally constitute or cause, at least partially, the occurrence of the B-outcomes or vice versa.
-To illustrate, one way of matching is for the B-outcomes to be the A-outcomes.
-Another way of matching is for the B-outcomes to stand to the A-outcomes as elements of a more detailed plan stand to those of a less detailed one.
+Why think that this question poses a problem at all?  
+To see why, consider how the question might be answered.
 
-Now we can put the question more generally.
-There are cases in which a particular action is guided both by one or more intentions and by one or more motor representations.
-In at least some such cases, the outcomes specified by the intentions match the outcomes specified by the motor representations.
-Furthermore, this match is not always accidental.
-How do non-accidental matches come about?
+One might try to explain the match by supposing that intentions and motor representations have a common cause.
+If the mere presence (or the mere perception) of a lever invariably triggered intentions and motor representations specifying grasping (say), it might be possible to explain matching in this way.
+But this sort of consideration cannot provide a full explanation of non-accidental matching
+because
+intentions are not always triggered in straightforward ways by agents' environments or perceptions.
+To suppose otherwise is to ignore the very phenomena, decision and planning, which make intention so interesting.
 
-In principle one might try to explain the match by supposing that intentions and motor representations have a common cause.
-If the mere presence (or the mere perception) of a lever invariably triggered intentions and motor representations specifying grasping (say), it might be possible to explain matching in this way.
-But this sort of consideration cannot provide a full explanation of non-accidental matching for two reasons.
-First, intentions are not always triggered in straightforward ways by agents' environments or perceptions; to suppose otherwise is to ignore the very phenomena, decision and planning, which make intention so interesting.
-Second, motor representations are also not always triggered in any straightforward way by agents' environments or perceptions either.
-For these reasons there seems to be no hope of fully explaining matching by postulating common causes for intentions and motor representations.
-
-If common cause explanations are ruled out, another natural approach is to  appeal to content-respecting causal processes.
+Another natural approach is to  appeal to content-respecting causal processes.
 Perhaps, for example, intentions with certain contents (concerning grasping, say) reliably cause motor representations with corresponding contents (also concerning grasping, say).
 Alternatively we might suppose, very crudely, that some comparator process checks that the contents of motor representations are appropriate given what the agent intends.
-Either way, the idea is to explain matches between outcomes specified by the contents of certain states in terms of content-respecting causal processes linking those states.
 
-This type of explanation is arguably appropriate where the states in question have the same representational format.
-For example, this type of explanation would arguably be appropriate if our aim were to explain matches between  large-scale intentions and the smaller-scale, more detailed intentions which serve as building blocks for them.
-But in fact we are concerned with intentions and motor representations which, as argued above, have different representational formats.
-This creates a potential difficulty.
-
-In general, when two representations differ in format, postulating reliably content-respecting causal processes linking them requires us to explain how their contents are coordinated.
+There is a difficulty for any  explanation along these lines.
+As argued,
+intention and motor representation involve different representational formats. 
+In general, when any two representations differ in format, postulating reliably content-respecting causal processes linking them requires us to explain how their contents are coordinated. 
 To illustrate, suppose you are given some verbal instructions describing a route.
@@
 Special cases aside, answering the question will involve a process of translation because two distinct representational formats are involved, propositional and cartographic.
-It is not be enough that you could follow either representation of the route.
-You will also need to be able to translate from at least one representational format into at least one other format.
 Similarly, for there to be reliably content-respecting causal processes linking intentions with motor representations there would have to be some process of translation.
 
-But why is this a potential difficulty?
-What is wrong with postulating a process of translation?
+But what is wrong with postulating a process of translation?
 The difficulty is that nothing at all is known about this hypothetical translation between intention and motor representation, nor about how it might be achieved, nor even about how it might be investigated.
-Of course this doesn't show that we couldn't fully explain matching by appeal to content-respecting causal processes.
+Of course this doesn't \emph{show} that we couldn't fully explain matching by appeal to content-respecting causal processes.
 But it does show that no such explanation is currently available.
@@
 It is a problem because of two natural routes to answering the question, the first (appealing to common causes of intentions and motor representations) is a non-starter and the second (appealing to content-respecting causal processes) amounts to no more than a stab in the dark.
-Our aim in what follows is to solve the interface problem without postulating either common causes or translation processes.
+Our aim for the rest of this talk is to solve the interface problem without postulating either common causes or translation processes.
 
@@
 
-The interface problem is the problem of explaining how, some of the time, there could be reliable matches between outcomes variously specified by intentions and motor representations.
-As the previous section explained, the problem arises because intentions and motor representations have different representational formats.
-
+The interface problem  arises because intentions and motor representations have different representational formats.
 There is a way to link representations with different formats that requires neither common causes nor translations.
@@
 Because the representation deferred to is cartographic, comparing the instruction with the map no longer requires translation between representational formats.
+
 We shall suggest that something analogous holds concerning the relation between intention and motor representation.
-To anticipate what will be explained below, some intentions involve demonstrative concepts (or other constituents) which refer to actions by deferring to motor representations.
+To anticipate what will be explained below, some intentions involve demonstrative concepts which refer to actions by deferring to motor representations.
+These demonstrative concepts would be concepts of actions not of motor representations, but they would succeed in being concepts of actions by deferring to motor representations.
 
-Let us first step back and consider thought generally.
-The existence of demonstrative thoughts and concepts is quite widely accepted.
-For instance, some philosophers have proposed that there are demonstrative colour concepts \citep{McDowell:1996yi,Brewer:1999ud}.
-While there are objections to uses to which these have been put (e.g.\ \citealp{heck_jr._nonconceptual_2000,dokic_shades_2001}), the claim that there might be such concepts is barely controversial.
+That 
+there are demonstrative  thoughts and concepts is quite widely accepted
+\citep[e.g.][]{McDowell:1996yi,Brewer:1999ud},
+ for both individuals and types  \citep[§3.4]{levine:2010_demonstrative}.
+And it seems necessary to suppose that there are demonstrative concepts of actions
+given that someone might think to herself, `I wish I could do that too'.
+But the further claim that some demonstrative concepts refer to actions \emph{by deferring to motor representations} immediately raises a question. 
+Could a demonstrative concept really defer to a motor representation?
 
-Consider the possibility that some action concepts are demonstrative.
-Someone who says to herself, `I wish I could do that too' may be entertaining a proposition which involves a demonstrative concept, \emph{that action}.
-Clearly this demonstrative concept cannot refer to a token action.
-The wish was not to perform another agent's action but to perform an action of a certain type.
-Now the mere fact that the demonstrative concept must refer to a type is not a problem.
-As the case of colour concepts indicates, it is plausible that propositional attitudes can involve elements which demonstrate types and not just individuals \citep[see also][§3.4]{levine:2010_demonstrative}.
-
-How could the existence of demonstrative concepts for actions be relevant  to solving the interface problem?
-In our earlier illustration, the instruction `follow this route!' succeeds in referring to a route by deferring to another representation with a different format.
-The instruction is about a route not a representation, but it succeeds in referring to a route by deferring to a representation of that route.
-Similarly, some demonstrative concepts may refer to types of action such as grasping or throwing by deferring to motor representations (see \citet{levine:2010_demonstrative} for one way of developing this idea).
-These demonstrative concepts would be concepts of actions not of motor representations, but they would succeed in being concepts of actions by deferring to motor representations.
-For any such concept, it is a motor representation which ultimately determines what it is a concept of.
-
-The idea that some demonstrative concepts refer to actions by deferring to motor representations immediately raises a question.
-Could a demonstrative concept really defer to a motor representation?
-It seems clear that we can't select a motor representation to defer to in the same way that we can select a map when we say `Follow this route!'.
-After all, motor representations are not things we can point to with our hands.
-Nevertheless, it does seem that motor representations are available in some sense.
+It seems clear that we can't select a motor representation to defer to in the same way that we can select a map when we say `Follow this route!' 
+We shall argue, nevertheless, that motor representations are available. 
 To start with an analogy, consider pantomiming an action to yourself.
 You are rehearsing part of an operation which involves precisely grasping a delicate structure with some tweezers.
-Just as someone might point to a map and say `Follow that route!', so also they could point to your pantomime and say `Do that!'.
+Just as someone might point to a map and say `Follow that route!', so also they could point to your pantomime and say `Do that!'
  In our analogy, the pantomime stands in for a motor representation of the action.
 The demonstrative in `Do that!' refers to an action by deferring to the pantomime.
-Of course this analogy doesn't show that demonstrative concepts can defer to motor representations (at least not unless pantomimes are motor representations of actions).
+Of course this analogy doesn't show that demonstrative concepts can defer to motor representations.
 But now consider purely mental pantomime---that is, motor imagery.
@@
 
-But is this really possible?
-Someone might object that in appealing to motor imagery we are sneaking intentions in through the back door.
-How can we be sure that it is ever really a motor representation rather than an intention that one defers to in thinking `Do that!'?
-Contrast two cases of motor imagery, both involving a tool.
-In the first case, imagine grasping an object with the tool; and in the second case, imagine releasing the same object with the tool.
-The two cases involve different imagery.
-But in what does this difference consist?
-Could it be a difference in intention that explains how the two cases of motor imagery differ?
-Since it is possible to grasp without intentionally grasping, it is surely also possible to imagine grasping without imagining intentionally grasping.
-And note that to imagine grasping without imagining intentionally grasping does not necessarily require one to imagine non-intentionally grasping.
-(It is easy to miss this point by confusing it with the issue of whether the imagining must itself be intentional.
-Whether or not it is possible to imagine grasping or releasing without \emph{intending to imagine} so acting, one can certainly imagine grasping or releasing without \emph{imagining intentionally} so acting.) So to rule out the possibility that the two cases of motor imagery differ with respect to intention, let the motor imagery be neutral with respect to intention.
-Is the difference between the two cases then due to differing movements or patterns of muscle contraction?
-To rule this possibility out, let the movements and muscle contractions involved in both cases be as similar as possible: let the difference between grasping and releasing the object be a matter of how the tool is configured rather than of how your body moves.
-With the contrast cases elaborated by these two stipulations, it is plausible that the difference in imagery between the two cases will be due to the different motor representations involved in grasping and releasing.
-So in these cases, the thought `Do that!' will refer to grasping (or releasing) by deferring to a motor image of grasping (or of releasing).
-Since motor images are a species of motor representation, it follows that demonstrative concepts can refer to actions by deferring to motor representations.
+Here we are assuming that 
+it is possible to imagine acting purposively with imagining acting intentionally.
+We lack time to defend that assumption now but 
+there's plenty more to say, perhaps in the question time.
 
-So far we have focused on motor imagery.
-We are not committed to claiming that demonstrative concepts can only refer to actions by deferring to motor images.
-Given the parallels between motor imagery and motor preparation \citep{Jeannerod:2001yb}, it is plausible that motor representations feature in experiences of action in such a way that motor actions are available for demonstrative reference.
-If so, the possibility of demonstrative concepts which refer to actions by deferring to motor representations is supported not just by motor imagery but also by experiences associated with motor activity more generally.
+Until now we have focused on motor imagery. 
+Given the parallels between motor imagery and motor preparation \citep{Jeannerod:2001yb}, it is plausible that motor representations also feature in experiences of action in such a way that motor actions are available for demonstrative reference.
 
-We have been arguing for the existence of demonstrative concepts which refer to actions by deferring to motor representations in order to solve the interface problem.
-As already stated, the interface problem is the problem of explaining how outcomes variously specified by intentions and motor representations, despite their different formats, sometimes non-accidentally match.
-As long as we think of intentions and motor representations as having logically independent contents, it seems that fully solving the problem would require appeal to processes of translation linking intention with motor representation.
-But where intentions involve demonstrative action concepts, their contents are not necessarily logically independent of the contents of motor representations.
-For a demonstrative component of an intention may refer to an action by deferring to a motor representation.
+Recall that the interface problem is the problem of explaining how outcomes variously specified by intentions and motor representations, despite their different formats, sometimes non-accidentally match.
+As long as we think of intentions and motor representations as having logically independent contents, it seems that fully solving the problem would require appeal to
+common causes or 
+processes of translatio. But intentions can involve demonstrative action concepts 
+which refer to  actions by deferring to motor representations. 
 Where this happens, which actions the intention specifies is partially or wholly determined by the motor representation, and so the interface problem is solved.
-
-Or is it?
-So far we have not distinguished between two aspects of action outcomes.
-Action outcomes often specify both a way of acting---whether to grasp or release, say---and also what to act on---on the mug or the pen, say.
-Since we have been focusing on ways of acting rather than on what is to be acted on, it would be consistent with our arguments to accept that there are demonstrative concepts of ways of acting (such as grasping and reaching, perhaps) which defer by referring to motor representations while denying that any such demonstrative concepts are also about what is to be acted on.
-If this combination of views were correct, we would have only a partial solution to the interface problem.
-We would only have explained how the outcomes variously specified by intentions and motor representations non-accidentally match with respect to ways of acting, not with respect to what is to be acted on.
-However a range of behavioural and neurophysiological evidence shows that motor representations represent not only ways of acting but also objects on which actions might be performed and some of their features related to possible action outcomes involving them (for a review see \citealp{Gallese:2011uq}; for discussion see \citealp[pp.\ 410-3]{pacherie:2000_content}).
-For example visually encountering a mug sometimes involves representing features such as the orientation and shape of its handle in motor terms \citep{buccino:2009_broken,costantini:2010_where,cardellicchio:2011_space,Tucker:1998,tucker:2001_potentiation}.
-Reference by deference to motor representation could therefore explain how action outcomes match with respect not only to ways of acting but also with respect to what is to be acted on.\footnote{ In fact, there may be another way of explaining non-accidental matching with respect to what is to be acted on due to \citet[pp.\ 36-8, 44-5, 48-57]{Campbell:2002ge}.
-In outline, Campbell's idea is to note that intentions can refer to objects by means of perceptual demonstrative elements, and that perceptual and motor representations of objects may be sufficiently commensurate for a perceptually represented object to be reliably selected as a target for motor processes.
-It would be consistent to hold both that a view like Campbell's correctly explains non-accidental matching of action outcomes with respect to what is to be acted on, and also that non-accidental matching with respect to ways of acting is explained by demonstrative action concepts which defer to motor representations.
-We shall leave open the question of whether non-accidental matching involves elements additional to those involved in the view we have been developing.
-Our claim is just that a solution to the interface problem need not involve anything other than components of intentions which refer to actions by deferring to motor representations of action outcomes.}
-
-Solving the interface problem may not always involve demonstrative concepts.
-What matters for solving the interface problem is deference, not demonstration.
-Suppose that someone acts on an intention to grasp the handle of a mug.
-Suppose also that the outcome to which her action is directed, grasping the mug's handle, is specified by motor representations.
-As long as the concept of grasp involved in her intention refers by deferring to a motor representation of grasping, the two specifications of action outcomes (in intention and in motor representation) will not be independent and so the interface problem will be solved.
 
@@
 
-While most philosophers would probably agree that, as a matter of fact, intentional action often involves motor representation, this is typically treated as only an enabling condition for intentional action or as merely a variety of intention.
-In fact most philosophical theories of action apply indifferently to agents whose actions involve motor planning and monitoring and to (imaginary) agents who need explicit practical reasoning for each muscle contraction.
-In our view this is a defect of those theories.
-To fully understand intention, we need to understand how intentions relate to motor representations.
-Or so we have argued.
+Philosophers typically treat motor representation either as as only an enabling condition for intentional action (the first temptation) or else as merely a variety of intention (the second temptation).
+In fact most philosophical theories of action  apply indifferently to 
+(imaginary) agents lacking motor systems
+who would need explicit practical reasoning for each muscle contraction. 
+This is a mistake if,
+as we have argued,
+ motor representation has a distinctive role in explaining the purposiveness of action.
 
-The need to understand how intentions relate to motor representations follows from three claims defended in early sections of this paper.
-First, some motor representations represent outcomes (rather than, say, only bodily movements).
-Second, there are actions whose directedness to an outcome is grounded in motor representations.
-And, third, motor representations differ from intentions with respect to their representational format.
-These claims reveal that a single purposive action may involve representations of the outcomes to which it is directed in at least two different representational formats, motor and propositional.
-It is necessary to understand something of how intentions relate to motor representations in order to explain how the various representations involved in a single action sometimes non-accidentally specify matching outcomes.
+Our argument was this.
+Some motor representations, like intentions, coordinate actions in virtue of representing outcomes; but motor representations and intentions differ in representational format.  
+This implies that motor representation has a distinctive role in explaining the purposiveness of action. 
+It also leads to what we called \emph{the interface problem}.
+A single purposive action may involve representations of the outcomes to which it is directed in at least two different representational formats, motor and propositional;
+and these outcomes sometimes non-accidentally match.
+The interface problem is the problem of explaining how non-accidental matching is possible, of
+explaining how intentions interlock with motor representations. 
+We have shown that there is a theoretically coherent and empirically plausible solution to the interface problem that requires neither common causes nor processes of translation.
+Both can be avoided providing that intentions can involve demonstrative concepts which refer to actions by deferring to motor representations.
+Such demonstrative concepts tie intention to motor representation.
 
-The problem of explaining this we called \emph{the interface problem}.
-Theoretically uncomplicated approaches to solving it would involve appeal to common causes or to processes of translation.
-The interface problem is a problem because no such approach is viable given the present state of knowledge.
-However, we have shown that there is another way of solving the interface problem.
-This solution hinges on the existence of concepts which refer to actions by deferring to motor representations.
-Where intentions involve such deferential concepts, their contents are logically linked to those of motor representations and so the interface problem can be solved.
-In short, then, we have shown that there is a theoretically coherent and empirically plausible solution to the interface problem, one which requires neither common causes nor processes of translation to link intention and motor representation.
-
-A further conjecture is that whenever there is a non-accidental match between the contents of intentions and of motor representations, this is due to deferential action concepts.
-The plausibility of this conjecture rests in part on the absence of any other viable solution to the interface problem, and of course we do not claim to have shown that no alternatives exists.
-If the conjecture is right, if there is no way of solving the interface problem other than by means of deferential action concepts, then concepts which refer by deferring to motor representations are what tie explicit practical reasoning to motor processes.
-
-More speculatively still, we suggest that where an intention properly and reliably produces bodily movement, either acting on that intention involves a further intention or else the intention involves concepts which refer to actions by deferring to motor representations.
+More speculatively, we suggest that where an intention properly and reliably produces bodily movement, either acting on that intention involves a further intention or else the intention involves concepts which refer to actions by deferring to motor representations.
 If so, it is deferential action concepts that ultimately connect intentions to bodily movements.
-Only by recognising how intentions interlock with motor representations can we hope to understand how our intentions ever make a difference to the world around us.
+Only by recognising how intention is tied to motor representation can we hope to understand how our intentions ever make a difference to the world around us.
 
@@
 But if, as we conjecture, such deference is necessary for intentions to properly and reliably result in bodily movements, it may turn out that intentionally acting in the world depends on action experiences featuring motor representation.
-Much as on some views thought about objects depends on perceptual experience \citep[e.g.][]{Campbell:2002ge}, so also intending actions may depend on motor experience.
+Much as, on some views, thinking about objects depends on perceptual experience \citep[e.g.][]{Campbell:2002ge}, so also intending actions may depend ultimately on motor experience.
 

### 2012-02-23 mindreading/talk ceu bcccd 2012-01 with ian apperly/how do humans develop two systems bcccd talk my parts.tex
started on talk

%!TEX TS-program = xelatex
%!TEX encoding = UTF-8 Unicode

\def \papersize {a4paper}
\documentclass[14pt,\papersize]{extarticle}
% extarticle is like article but can handle 8pt, 9pt, 10pt, 11pt, 12pt, 14pt, 17pt, and 20pt text

\def \ititle {How do humans develop two systems for mindreading?}
\def \isubtitle {}
\def \iauthor {Ian Apperly \& Stephen A. Butterfill}
\def \iemail{s.butterfill@warwick.ac.uk}
%\def \iauthor {}
%\def \iemail{}
\date{}

\input{$HOME/Documents/submissions/preamble_steve_paper2}
%\author{}

%\setromanfont[Mapping=tex-text]{Sabon LT Std}

\begin{document}

\setlength\footnotesep{1em}

\bibliographystyle{newapa} %apalike

\maketitle
%\tableofcontents

\begin{abstract}
Experimental evidence, and reflection on everyday examples suggest that adults’ mindreading is both flexible and sophisticated (as when judging someone’s guilt in a court of law) and fast and efficient (as when taking part in a competitive sport). However, conceptual analysis, and evidence from analogous cognitive domains, suggests that a single cognitive process for “theory of mind” is unlikely to be able to meet these diverse demands. This leads to the proposition that adults have “two systems” for mindreading that make different processing trade-offs between flexibility and efficiency (Apperly \& Butterfill, 2009).

How are adults’ two systems related to the mindreading abilities observed in infants? One possibility is that the infant system grows up: although simple and cognitively efficient at the outset the infant system becomes increasingly sophisticated as children gain conceptual, linguistic and executive resources. On this view, the cognitively efficient abilities of adults must have some other origin, perhaps in automatisation of abilities that were previously effortful. A second possibility is that the infant system persists: the system that explains efficient mindreading in infants also explains efficient mindreading in adults. On this view the flexible and sophisticated abilities of adults must have some other origin, perhaps in the protracted developments charted in traditional studies of children’s “theory of mind”. We will draw inspiration from Susan Carey’s work on “signature limits” in other aspects of core cognition to identify evidence that might distinguish between these accounts.

\end{abstract}

\section{Introduction}
How do humans develop two systems for mindreading?
Our question presupposes that human adults (at least) do have two systems for mindreading.
Actually this is something we want to argue for rather than assume. 
So let me start by making a distinction and an observation ...

\section{slide}
The distinction is between processes which are automatic 
and those which are not.

Roughly speaking,
a process is *automatic* if it occurs irrespective of its relevance to the particulars of an agent's motives and aims.

\section{slide}
The puzzle starts with a question.
Is adult mindreading automatic or not?

\section{slide}
Some evidence suggests that it is.
For example,
Kovacs and colleagues (2010) asked adults to identify the location of a ball.
They found that adults were significantly slower to identify the ball's location when an onlooker had a false belief about the location of the ball,
even though the onlooker's belief was not relevant to the task at all.
This and related findings seem to suggest that, at least in adults, mindreading is automatic.

\section{slide}
By contrast,
Back \& Apperly (2010) found that subjects are significantly slower to answer an unexpected question about another's belief when that belief is false compared to when it is true.
This suggests that, at least in adults, mindreading is not automatic.

So my question is whether adult mindreading is automatic.
And this question is puzzling because there seems to be evidence for both positive and negative answers.

\section{Ian's first bit}
***

\section{Steve's second bit}
So the question is how we develop two systems for mindreading, one automatic the other not. Two hypotheses are:

a. automaticity through automatization

and

b. original automaticity.

(Define 'originally automatic' mindreading as mindreading that is automatic and not a consequence of automatization.)

Our suggestion is that 
(i) automatic mindreading systems are subject to arbitrary limits (because automaticity involves trading flexibility for efficiency),
and 
(ii) we can distinguish the two hypotheses by finding arbitrary limits on automatic mindreading in infants and adults.
If the limits match, this is evidence for original automaticity.
If the limits differ, this is evidence for automaticity through automatization.

What sorts of limits might we expect to find on originally automatic mindreading?

Full-blown, explicit mindreading involves representing beliefs, desires and other propositional attitudes.
Propositional attitudes---including beliefs---are states which form complex causal structures, have arbitrarily nest-able contents, and are individuated by their causal and normative roles in explaining thoughts and actions.  
If anything should make a cognitive process effortful, it is surely representing states with that combination of properties

So there are strong theoretical reasons to suppose that originally automatic mindreading could not involve representing propositional attitudes as such.

\section{slide}
Elsewhere Ian and I have suggested, in effect, that originally automatic mindreading involves representing relational attitudes ...

\section{slide}
Paradigmatic examples of relational attitudes are things like 
being excited by an event,
encountering an object,
or wanting some type of thing.

\section{slide}
whereas propositional attitudes have arbitrarily nestable contents,
relational attitudes have no contents

whereas propositional attitudes have uncodifiably complex effects on action, 
relational attitudes set parameters for action

and whereas representing relational attitudes would permit a mindreader to track false beliefs involving mistakes about 
   appearances,
   identity
 and 
    existence,
representing relational attitudes would allow one to track only a limited range of false beliefs about properties such as location, colour and shape.

So consider this hypothesis:
whereas non-automatic mindreading involves representing propositional attitudes,
automatic mindreading involves representing relational attitudes only.

This hypothesis, if true, would explain why automatic mindreading is less demanding than full-blown mindreading.

\section{slide}

This hypothesis also has the virtue of indicating what sort of limits could distinguish automatic from full-blown mindreading ...

\section{slide}
for the hypothesis suggests that

- automatic mindreading should be good for level-1 perspective taking tasks,
but not for level-2 perspective taking tasks.

Some work on the level-1/level-2 distinction has already been published,
and this does suggest that for both children and adults,
automatic mindreading may be sufficient for success on level-1 but not level-2 perspective taking tasks.

But since this has already been published I don't want to cover that here.

\section{slide}

The hypothesis about relational attitudes also suggests that

- automatic mindreading should be good for tracking false beliefs with a simple predicate-object structure, but not for tracking false beliefs which essentially involve the non-existence of something.

\section{slide}

and, finally, the hypothesis suggests that

-automatic mindreading should be good for tracking false beliefs about the location of something,
but not for tracking false beliefs which essentially involve mistakes about identity.

So one nice feature of this hypothesis is that it makes several independently testable predictions about the limits of automatic mindreading.

If we want to find out whether   adult automatic mindreading and infant mindreading are subject to the same limits,
it makes sense to start by looking at these limits.

I want to finish by focusing on the issue of false beliefs about location vs. false beliefs about identity.

What do we mean by a mistake about identity?

Here's an illustration.
Suppose that a single person has two roles that don't often go together,
say she is both a school teacher  by day and an assassin by night.
If you shot the assassin, you might be surprised that the school teacher doesn't turn up to work.
They are one and the same person but you didn't know that.
This would be a mistake about identity: in mistakenly expecting the school teacher to turn up, you have failed to realise that she is the assassin.
Your mistake could be predicted by subjects who can represent propositional attitudes as such.
But it cannot be predicted by subjects who represent relational attitudes only.

\section{slide}
Here's how this idea might work in a format suitable for infants and adults (without the shooting).

On the screen you can see the back of the subject's head.
He is sitting opposite the protagonist.
On the table there is a screen.
Behind the screen there is a puppet.
The puppet will serve as the school teacher-assassin.
The protagonist can't see behind the barrier.
She can't see the puppet right now.
But the subject can see everything.

\section{slide}
This is the same thing, just smaller so that I can show you how the story unfolds.

I should say that the puppet has a face and costume with two aspects.
Seen from one side she looks like an assassin.
Seen from the other she looks like a school teacher.

\section{slide}
This is the same thing, just smaller so that I can show you how the story unfolds.

I should say that the puppet has a face and costume with two aspects.
Seen from one side she looks like an assassin.
Seen from the other she looks like a school teacher.

\section{slide}
The puppet now rotates behind the barrier.
The subject can see this but the protagonist can't.

\section{slide}
Now the puppet goes out from behind the barrier.
This time she looks to the protagonist as if she is an assassin.
The protagonist shoots her and she falls off the table.
So to the protagonist it looks like the assassin has been shot.
But the subject knows that the assassin is the school teacher.

\section{slide}
In the final phase,
the protagonist reaches around the barrier.

If the subject can represent propositional attitudes as such this makes perfect sense:
the protagonist wants to  retrieve the school teacher.

But if the subject is representing relational attitudes only,
this reaching doesn't make any sense.
For the assassin is the school teacher,
and so one's disappearance is the other's disappearance.

So the protagonist's reach should be expected to subjects who represent propositional attitudes but unexpected to subjects who represent relational attitudes only.

[Don't mention control conditions]

\section{slide}
So this was our idea.
We want to know whether adults' automatic mindreading is (a) automatic through automatization
or
(b) originally automatic.

Our suggestion is that if adult automatic mindreading is originally automatic, then it should be subject to the same arbitrary limits as infant mindreading.

And there are theoretical reasons to suppose that originally automatic mindreading would involve representing relational attitudes only, not propositional attitudes.

Someone who represents relational attitudes only
should have no problem tracking false beliefs about location
but should fail to track false beliefs about identity.

And, as I've just tried to illustrate, it would be possible to test for abilities to track false beliefs about identity using designs very similar to those used to track false beliefs about location.

So here is a way to test for original automaticity.
If we find that infant mindreading is like adult automatic mindreading in that neither enables them to track  false beliefs about identity,
then we have evidence for original automaticity.
But if it turns out that infants and adults differ in this respect,
then we have evidence for automaticity by automatization.

\section{Ian's conclusion}
***

\bibliography{$HOME/endnote/phd_biblio}

\end{document}

### 2012-02-26 conf espp 2012 symposium false belief counterfactuals perspective schurz rafetseder/false beliefs, perspectives and counterfactuals espp 2012 symposium submission.tex
working on ESPP 2012 symposium submission

%!TEX TS-program = xelatex
%!TEX encoding = UTF-8 Unicode

%a4paper or letterpaper (also used in preamble_steve_paper2
\def \papersize {a4paper}

\documentclass[12pt,\papersize]{extarticle}
% extarticle is like article but can handle 8pt, 9pt, 10pt, 11pt, 12pt, 14pt, 17pt, and 20pt text

\def \ititle {False Beliefs, Perspectives and Counterfactuals}
\def \isubtitle {}
\def \iauthor {Stephen A. Butterfill}
\def \iemail{s.butterfill@warwick.ac.uk}

\input{$HOME/Documents/submissions/preamble_steve_paper2}
%\author{}
%\date{}

%\setromanfont[Mapping=tex-text]{Sabon LT Std}

\begin{document}

\setlength\footnotesep{1em}

\bibliographystyle{newapa} %apalike

\maketitle
%\tableofcontents
\title{}

\begin{abstract}
\noindent
***

\end{abstract}

\section{Introduction}
The challenge is to understand what mindreading is, how it develops and which cognitive processes underpin it.
One approach is to identify components of mindreading.
As I understand it, Eva's and Matthias' research does just this.
Eva argues that mindreading, at least insofar as it involves representing beliefs, may depend on counterfactual reasoning too.
And Mattias supports the claim that mindreading involves a separable ability to process perspective differences by showing that the neural correlates of this ability are found in both mindreading and non-mindreading tasks.

There appears to be a conflict between Eva's and Matthias' interpretations of their findings and a widely held view about infant mindreading (as Eva notes in the paper on which her talk is based).
Here are three claims which are collectively inconsistent:
%
\begin{enumerate}
\item Infants can represent false beliefs from around their first birthday or earlier. \label{infant_fb}

\item Being able to represent false beliefs involves being able to  (i) process perspective differences or (ii) reason counterfactually (or both). \label{fb_is_perspectives}

\item Infants cannot (i) process perspective differences nor (ii) engage in counterfactual reasoning until they are at least one year old.  \label{infant_perspectives}

\end{enumerate}
%
Since these claims are collectively inconsistent, we must reject at least one.  But which?

\section{The puzzle}
Can we reject (\ref{infant_fb}), the claim that infants can represent false beliefs?  
This is what  Eva suggests in her paper.
But there is a growing body of evidence which appears to support  (\ref{infant_fb}).
From around their first birthday infants  predict actions of agents with false beliefs about the locations of objects \citep[]{en_1092, en_1208} and choose different ways of interacting with others depending on whether their beliefs are true or false \citep[]{en_1783,Knudsen:2011fk,southgate:2010fb}.  
And in much the way that irrelevant facts about the contents of others’ beliefs modulate adult subjects’ response times, such facts also affect how long 7-month-old infants look at some stimuli \citep[]{en_1821}.
The variety of paradigms measures---looking time, anticipatory looking, pointing and helping---makes it hard to dismiss these findings on methodological grounds.
And while some have proposed that infants might be tracking behaviour only \citet{en_1168, en_1169},
this proposal faces several objections \citep[e.g.][]{Song:2008qo}.
The core issue is not whether hypothetical behavioural strategies might in principle explain theory of mind abilities; it is whether all of these subjects’ actual behaviour reading capacities are flexible enough to explain the full range of their theory of mind abilities \citep[]{en_1686}.
In my view this is unlikely.
So straightforward rejection of  (\ref{infant_fb}) is not an option.

Can we reject (\ref{fb_is_perspectives})?
Since Matthias' and Eva's talks are (in part) arguments for this claim, let me come back to it after first considering whether we can reject 
claim (\ref{infant_perspectives}).

Can we reject (\ref{infant_perspectives}), the claim that infants neither process perspective differences nor engage in counterfactual reasoning until they are at least one year old?
One difficulty is that current developmental evidence almost uniformly supports this claim  \citep{rafetseder:2010_counterfactual,beck:2011_almost}.
A second problem is that  abilities to reason counterfactually,%
\footnote{Here and throughout I follow Eva Rafetseder in using the term `counterfactual reasoning'  to refer to reasoning which essentially involves comparing two possible situations; the term is sometimes used more broadly to include the sorts of representations involved in pretence.} 
like abilities to represent false belief,
appear to involve working memory and inhibitory control 
(on counterfactuals: 
	\citealp{drayton:2011_counterfactual,beck:2011_supporting};
on false beliefs: \citealp{en_1412, en_1698, lin:2010_reflexively, en_1547} experiments 4-5).
As even committed nativists are likely to agree, capacities for inhibitory control and working memory develop over several years and are limited in infants \citep[e.g.][]{carlson:2005_developmentally}.
So we cannot straightforwardly reject (\ref{infant_perspectives}).

Since we can't straightforwardly reject (\ref{infant_fb}) or (\ref{infant_perspectives}),
let's consider whether (\ref{fb_is_perspectives}) can be rejected.
This is the claim that being able to represent false beliefs involves being able to process perspective differences or to reason counterfactually.
One might attempt to reject this claim on the grounds that the link between false belief, perspective differences and counterfactual reasoning 
is spurious and a consequence of \emph{extraneous demands} accompanying standard false belief tasks.
Here the idea is that, although certain false belief \emph{tasks} impose demand processing perspective differences or core components of counterfactual reasoning,
these demands are extraneous in the sense that they are not a consequence merely of being required to represent false beliefs
(proponents of this view include 
	\citealp[p.\ 417]{Carpenter:2002gc},
	\citealp{Bloom:2000bt}, and
	\citealp{Leslie:1998nq}).
I want to go slowly here and consider two versions of this idea.

Sometimes it is taken for granted that representing false beliefs does not necessarily and intrinsically involve inhibition;
and so the fact that success on many false belief tasks depends on inhibitory control is interpreted as revealing a defect of the task.
Someone who holds this view will interpret the evidence Eva and Matthias present as showing that the false belief tasks they consider are defective measures of the ability to represent false beliefs.
But I think it is a mistake to assume that representing false belief---or, indeed, representing any propositional attitude---does not intrinsically demand cognitive resources such as inhibitory control.
On any standard view, propositional attitudes such as beliefs form complex causal structures, have arbitrarily nest-able contents, interact with each other in uncodifiably complex ways and are individuated by their causal and normative roles in explaining thoughts and actions \citep[]{en_809, en_249}.  
If there is anything representing which should consume scarce cognitive resources it is surely states with this combination of properties.
So there are sound theoretical reasons to suppose that representing false beliefs could intrinsically require inhibitory control and working memory \citep[see also][]{Russell:1999vr}.
And of course Eva and Matthias can provide theoretical reasons for supposing that representing false beliefs involves processing perspective differences and a core component of genuine counterfactual reasoning \citep{perner:2007_objects}. 
This is why we should not take for granted that representing belief can be separated from processing perspective differences and counterfactual reasoning; there is no good theoretical reason to assume that the connections must reflect extraneous task demands.

A less obviously flawed approach to rejecting (\ref{fb_is_perspectives}) involves analysing particular false belief tasks.
Two explore this idea it will be useful to distinguish between two types of false belief task, call them Category A and Category B.
By stipulation Category A tasks have these features:
\begin{itemize}
\item Children tend to pass them some time after their third birthday.
\item Abilities to pass these tasks has a protracted developmental course stretching over months if not years.
\item Success on these tasks is correlated with developments in executive function \citep[]{en_410, en_1130} and language \citep[]{en_1209}.
\item Success on these tasks is facilitated by explicit training \citep[]{en_85} and environmental factors such as siblings \citep[]{en_507, en_1299}.  
\item Abilities to succeed on these tasks typically emerge from extensive participation in social interactions (as \citealp{en_1300} suggest).
\end{itemize}
All the other false belief tasks are in Category B.
It is these Category B tasks on which infants succeed.
The proposal we are considering (and which I shall reject) is this.  
\begin{quote}
	All Category A tasks impose a requirement (or set of requirements) other than the requirement to represent a false belief;
call this the \emph{Extraneous Requirement}.
	The connections between success on Category A tasks and processing perspective differences and counterfactual reasoning are a consequence just of the extraneous requirement.
\end{quote}
I am going to argue that this proposal should be rejected.
First note that the proposal is harder to defend than sometimes assumed.
For instance, philosophers sometimes mistakenly assume that all Category A tasks involve language or communication, whereas Category B tasks do not.
But some Category B tasks involve language and communication 
\citep{Knudsen:2011fk,Song:2008qo}
and there are non-verbal tasks in Category A
(\citealp{Call:1999co}; \citealp{low:2010_preschoolers} Study 2).
We should also note that children who fail Category A tasks can answer questions about perception or pretence that are word-for-word identical with the questions about false belief that they cannot answer correctly (\citealp{Gopnik:1994cb}; see also\citealp{Cluster:1996ht}).
So the Extraneous Requirement cannot be straightforwardly linked to language or communication.
There is a more general problem facing attempts to identify the Extraneous Requirement.
There is much variation among the tasks in Category A.
It is not just that some are verbal whereas others are nonverbal.
It is also that some involve prediction whereas others involve retrodiction or justification \citep[e.g.]{Wimmer:1998kx},
some concern the first-person perspective, whereas others involve a second- or third-person perspective \citep[e.g.]{Gopnik:1991db},
some involve interaction whereas in others the subject is a mere observer \citep[e.g.]{Chandler:1989qa},
and some involve prediction actions whereas others involve predicting desires \citep{Astington:1991kk} or selecting an argument appropriate for someone with a false belief \citep{Bartsch:2000es}.
Despite all this variation and more, the Category A tasks all appear to measure a developmental transition \citep{Wellman:2001lz}.
This is a significant obstacle to identifying the Extraneous Requirement.
In fact, my view the currently available evidence supports the conclusion that there is no Extraneous Requirement.
And as far as I know the only plausible candidate for the transition  Category A tasks measure is a transition in the ability to represent beliefs.
This is why I don't think that we can straightforwardly reject  (\ref{fb_is_perspectives}).

To sum up so far, 
I have suggested that there are good reasons for each of (1)-(3) and 
argued that none of these claims can be straightforwardly rejected.
This leads to a puzzle.
Because (1)-(3) are collectively inconsistent, one of them must be rejected.
The puzzle is: Which should we reject? 
None of the arguments I have offered are decisive; what have aimed to establish is just that the question of which claim to reject is to a significant degree puzzling.
And although various researchers have taken one side or another,
I think it's fair to say that no one has yet succeeded in resolving the puzzle.

\section{How to resolve the puzzle}

\bibliography{$HOME/endnote/phd_biblio_en_record_num_keys}

\end{document}

### 2012-02-27 conf espp 2012 symposium false belief counterfactuals perspective schurz rafetseder/false beliefs, perspectives and counterfactuals espp 2012 symposium submission.tex
espp submit sent to matthias

@@
 
-\def \ititle {False Beliefs, Perspectives and Counterfactuals}
+\def \ititle {Representing Beliefs, Perspectives and Counterfactuals: A Puzzle}
 \def \isubtitle {}
@@
 \noindent
-***
+What is involved in representing belief?
+This talk starts with three claims that are separately defensible but appear to be jointly inconsistent: 
+%
+\begin{enumerate}
+\item Infants can represent false beliefs from around their first birthday or earlier. \label{infant_fb}
+%
+\item  \label{fb_is_perspectives}  Being able to represent false beliefs involves being able to  (i) process perspective differences or (ii) reason counterfactually (or both).
+%
+\item Infants cannot (i) process perspective differences nor (ii) engage in counterfactual reasoning until they are at least one year old.  \label{infant_perspectives}
+%
+\end{enumerate}
+%
+The first part of the talk draws on research presented elsewhere in this symposium to  defend each claim in turn.
+If each of the three claims (1)-(3) is true, they cannot be inconsistent after all.
+The second part of the talk explores ways of avoiding inconsistency by distinguishing measurement schemes and representational systems. 
+There may be multiple ways of representing belief, 
+and different ways of representing belief may vary in how flexible they are and what their implementation requires.
 
@@
 \section{Introduction}
-The challenge is to understand what mindreading is, how it develops and which cognitive processes underpin it.
+The broader challenge is to understand what mindreading is, how it develops and which cognitive processes underpin it.
 One approach is to identify components of mindreading.
@@
 Eva argues that mindreading, at least insofar as it involves representing beliefs, may depend on counterfactual reasoning too.
-And Mattias supports the claim that mindreading involves a separable ability to process perspective differences by showing that the neural correlates of this ability are found in both mindreading and non-mindreading tasks.
+And Matthias supports the claim that mindreading involves a separable ability to process perspective differences by showing that the neural correlates of this ability are found in both mindreading and non-mindreading tasks.
 
-There appears to be a conflict between Eva's and Matthias' interpretations of their findings and a widely held view about infant mindreading (as Eva notes in the paper on which her talk is based).
-Here are three claims which are collectively inconsistent:
+Eva's and Matthias' interpretations of their findings each appear to be in conflict with 
+ a widely held view about infant mindreading, 
+  as Eva notes in the paper on which her talk is based.
+These three claims appear to be jointly inconsistent:
 %
@@
 
-\item Being able to represent false beliefs involves being able to  (i) process perspective differences or (ii) reason counterfactually (or both). \label{fb_is_perspectives}
+\item  \label{fb_is_perspectives}  Being able to represent false beliefs involves being able to  (i) process perspective differences or (ii) reason counterfactually (or both).\footnote{
+Here and below I'm save words by not carefully distinguishing 
+ developmental claims about what acquiring an ability to represent false belief requires (Eva's main concern)
+ from
+ cognitive claims about what exercising an ability to represent false belief requires (Matthias' main concern).
+}
 
@@
 %
-Since these claims are collectively inconsistent, we must reject at least one.  But which?
+Since these claims appear to be jointly inconsistent, it seems we must reject at least one.  But which?
 
@@
 this proposal faces several objections \citep[e.g.][]{Song:2008qo}.
-The core issue is not whether hypothetical behavioural strategies might in principle explain theory of mind abilities; it is whether all of these subjects’ actual behaviour reading capacities are flexible enough to explain the full range of their theory of mind abilities \citep[]{en_1686}.
+The core issue is not whether hypothetical behavioural strategies might in principle explain theory of mind abilities; it is whether all of these subjects’ actual behaviour reading capacities are flexible enough to explain the full range of their theory of mind abilities \citep[]{Apperly:2009ju}
 In my view this is unlikely.
@@
 One difficulty is that current developmental evidence almost uniformly supports this claim  \citep{rafetseder:2010_counterfactual,beck:2011_almost}.
-A second problem is that  abilities to reason counterfactually,%
+A second problem is that the development of abilities to reason counterfactually,%
 \footnote{Here and throughout I follow Eva Rafetseder in using the term `counterfactual reasoning'  to refer to reasoning which essentially involves comparing two possible situations; the term is sometimes used more broadly to include the sorts of representations involved in pretence.}
-like abilities to represent false belief,
-appear to involve working memory and inhibitory control 
+like the development of abilities to represent false belief,
+appears to involve working memory and inhibitory control 
 (on counterfactuals: 
 	\citealp{drayton:2011_counterfactual,beck:2011_supporting};
-on false beliefs: \citealp{en_1412, en_1698, lin:2010_reflexively, en_1547} experiments 4-5).
+on false beliefs: \citealp{en_1412, en_1698}; \citealp{lin:2010_reflexively, en_1547} experiments 4-5; \citealp{saxe_reading_2006}).
 As even committed nativists are likely to agree, capacities for inhibitory control and working memory develop over several years and are limited in infants \citep[e.g.][]{carlson:2005_developmentally}.
@@
 It is not just that some are verbal whereas others are nonverbal.
-It is also that some involve prediction whereas others involve retrodiction or justification \citep[e.g.]{Wimmer:1998kx},
-some concern the first-person perspective, whereas others involve a second- or third-person perspective \citep[e.g.]{Gopnik:1991db},
-some involve interaction whereas in others the subject is a mere observer \citep[e.g.]{Chandler:1989qa},
+It is also that some involve prediction whereas others involve retrodiction or justification \citep[e.g.][]{Wimmer:1998kx},
+some concern the first-person perspective, whereas others involve a second- or third-person perspective \citep[e.g.][]{Gopnik:1991db},
+some involve interaction whereas in others the subject is a mere observer \citep[e.g.][]{Chandler:1989qa},
 and some involve prediction actions whereas others involve predicting desires \citep{Astington:1991kk} or selecting an argument appropriate for someone with a false belief \citep{Bartsch:2000es}.
@@
 This leads to a puzzle.
-Because (1)-(3) are collectively inconsistent, one of them must be rejected.
-The puzzle is: Which should we reject? 
+Because (1)-(3) appear to be jointly inconsistent, it seems that one of them must be rejected.
+This is puzzling because (1)-(3) also appear to be individually defensible.
 None of the arguments I have offered are decisive; what have aimed to establish is just that the question of which claim to reject is to a significant degree puzzling.
@@
 
-\section{How to resolve the puzzle}
+\section{Distinguishing implicit and explicit isn't enough}
+Could we resolve the puzzle by distinguishing between implicit and explicit representations of false belief?
+Given such a distinction we could remove the contradiction by making the following modifications:
+%
+\begin{quote}
+\begin{description}
+\item[1$'$.] Infants can represent false beliefs \emph{implicitly} from around their first birthday or earlier. \label{infant_fb_prime}
+
+\item[2$'$.] Being able to represent false beliefs \emph{explicitly}  involves being able to  (i) process perspective differences or (ii) reason counterfactually (or both).
+\label{fb_is_perspectives_prime}
+
+\item[3$'$.] Infants cannot (i) process perspective differences nor (ii) engage in counterfactual reasoning until they are at least one year old.  \label{infant_perspectives}
+\end{description}
+\end{quote}
+%
+When the primary measure of early false belief understanding was  looking times or anticipatory eye movements, this seemed promising \citep{Clements:1994cw,Clements:2000nc,Garnham:2001ql,Ruffman:2001ng}.
+However, as already mentioned, we now know that one-year-old infants succeed on false belief tasks which involve actively helping others, interpreting their utterances and pointing to provide information \citep{en_1783,Knudsen:2011fk,southgate:2010fb},
+and that older children still fail some nonverbal false belief tasks \citep{Call:1999co}.
+All other things being equal, it would be natural to suppose that success on these sorts of task reflects explicit representations of false belief.
+Certainly we lack a theoretically coherent and independently empirically motivated account of the distinction between implicit and explicit representation that associates exactly those tasks infants succeed on (strictly: Category B tasks) with implicit representations of false belief.
+So we don't know that distinguishing implicit from explicit representations does enable us to avoid contradiction.
+
+Of course I am not denying that, considered as a purely formal move, distinguishing between implicit and explicit representations could in principle enable us to avoid contradiction.
+The problem is to understand what this distinction amounts to clearly enough to explain the existing data and generate novel predictions \citep[p.\ 154]{Apperly:2010kx}.
+
+\section{How to resolve the puzzle: first step}
+\label{sec:first_step}
+
+I can't fully resolve the puzzle but I do think we can make some progress towards resolving it.  
+This section describes the first of two steps towards resolution.
+
+Let me start with an observation.
+When philosophers discuss what belief is, they generally disagree.
+They disagree about what sorts of contents beliefs have (for instance about whether their contents are propositional or map-like), 
+about which norms and causal roles characterise the attitude of believing (for instance about whether part of being a belief is being a state that ought to be true) 
+and about what it is for a belief to be true or false (for instance about whether truth or falsity is merely a matter of the belief's value as a guide to action).
+But when researchers focus on representing belief, it seems that no notice is taken of these disagreements.
+I think this is a mistake.
+Different views on what belief is are related to different accounts of what it is to represent belief.
+After all, what is involved in accurately representing a false belief depends in part on what a belief is and what it is for a belief to be false.
+
+A second observation is that there is no a priori reason to assume that all subjects who represent beliefs represent them with perfect accuracy.
+Suppose, just for the sake of argument, that beliefs are relations to full-blown Fregean propositions.
+A particular subject might nevertheless  represent beliefs as  relations to mere singular propositions.
+This subject's abilities to track others' beliefs, and to predict their actions, would be limited in various ways \citep{Salmon:1986fk}.
+But despite these limits, having such an ability to represent beliefs may still be useful in a wide range of everyday circumstances.
+In fact, in many cases this subject might be indistinguishable from another subject with a more accurate conception of belief.
+An analogy may help \citep[this is based on][]{matthews:2007_measure}.
+There are various ways to represent the efficiency of your car.
+If you were very exact, you might use a measure relating force to fuel consumption.
+But many people will instead measure distance in relation to fuel consumption,
+or even just fuel consumption in relation to periods of time (perhaps monthly).
+The point is that we have different ways of representing (measuring) efficiency.
+Some of these are highly flexible in the sense that they can cope with variations in landscape and even gravity but they are also more costly to implement in terms of the time and background knowledge they demand;
+while others are less costly to implement but also more limited in the sense that they only work in a more limited range of circumstances (perhaps only when the car is used in roughly the same way each month).
+What I am suggesting is that there is a parallel for representing (measuring) belief.
+On rich conceptions of belief, accurately representing belief will enable one to make fine-grained distinctions and to track complex contents, and this will demand conceptual sophistication and impose cognitive demands.
+On other, simpler conceptions of belief, accurately representing belief will yield less flexibility in distinguishing what others belief and in predicting their actions but the reduced flexibility in greater cognitive efficiency.
+
+Whatever belief actually is, 
+there may be multiple ways of representing it,
+just as there are multiple ways of measuring fuel efficiency.
+Different ways of representing belief will vary in how flexible they are and how demanding they are to implement.
+
+Saying that a subject can represent beliefs is like saying that she can represent fuel efficiency.
+It leaves open the question of how she represents beliefs.
+For instance, does she represent the contents of beliefs as simple tuples, as propositions of some kind, as images or maps, or what?
+The answer may vary for different subjects.
+And of course a single subject may use multiple ways of representing beliefs in different contexts (much as a single subject might rely on different ways of measuring fuel efficiency in different contexts).
+It is striking that almost no research has directly addressed the issue of how beliefs are represented in infants, children or adults.
+
+At this point it is perhaps tempting to suppose that we already have everything we need to resolve the puzzle.
+The idea would be this:
+%
+\begin{quote}
+\begin{description}
+\item[1$''$.] Infants can represent false beliefs \emph{using a comparatively simple measure} from around their first birthday or earlier.
+%
+\item[2$''$.] Being able to represent false beliefs \emph{using a comparatively sophisticated measure}  involves being able to  (i) process perspective differences or (ii) reason counterfactually (or both).
+%
+\item[3$''$.] Infants cannot (i) process perspective differences nor (ii) engage in counterfactual reasoning until they are at least one year old.  
+\end{description}
+\end{quote}
+%
+I think all these claims are true, and with Ian Apperly I have developed in detail a conjecture about the comparatively simple measure infants and others may use in representing beliefs \citep{butterfill_minimal}.
+It turns out that different measures of belief generate testable and novel predictions about the limits of subjects' abilities to track beliefs, so although we haven't tested the conjecture yet it is a testable conjecture.
+
+But although this is my view and I stand by it, I don't think it resolves the puzzle all by itself.
+The problem is this.
+Success on many standard false belief tasks in Category A is in principle possible using only a very simple measure for representing beliefs.
+Further, Category A tasks do not seem to differ from Category B tasks with respect to which measures for representing beliefs could in principle explain success on these tasks.
+So the view raises and does not answer the question of why children continue to systematically fail false belief tasks in Category A years after they can pass false belief tasks in Category B.
+
+\section{How to resolve the puzzle: second step}
+
+Here is another observation.
+In many domains it is widely recognised that patterns of performance in human subjects are best explained on the hypothesis that multiple representational systems are involved.
+This is true for representations of faces \citep{Morton:1991lr},
+of phonemes 
+and
+of causes \citep{Butterfill:2009vs}, 
+for instance.
+It is also true for representations of number \citep[e.g.][]{feigenson_tracking_2003}, agency and space \citep{Carey:1996hl}.
+In these domains and others, it seems we have to distinguish  
+	representations that paradigmatically feature in reflective thought and talk 
+from
+	representations that feature in modules or core systems.
+These modules are typically functional early in development and throughout most of the life course, 
+make few or no demands on scarce cognitive resources like working memory,
+and are limited and inflexible with respect to the information they can process and output.
+
+There are some striking parallels between patterns of performance which seem to license distinguishing multiple systems for representing number 
+and 
+patterns of performance in the case of reasoning about false beliefs and mental states more generally \citep{Apperly:2009ju}.
+This suggests that, in humans at least, representations of belief may feature in multiple cognitive systems.
+
+When a subject performing a task is described as representing a number or a belief, 
+this leaves open which system is involved.
+Is it a relatively efficient system or one that places demands on working memory or inhibitory control?  
+When does the system appear in development?
+In short, descriptions of subjects as representing beliefs need to be relativised in a second way.
+They need to be relativised not just to  a measurement scheme (see section \vref{sec:first_step}) but also to a system.
+
+The two kinds of relativisation---to a measurement scheme and to a system---are plausibly related.
+A representational system can only be cognitively efficient by avoiding sophisticated measurement schemes, which will result in limits on flexibility;
+equally, a representational system can only achieve flexibility in representing beliefs by implementing a sophisticated measurement scheme, and this will plausibly impose cognitive overheads in many situations.
+
+So here is a final refinement of the claims which make up the puzzle that I started with:
+%
+\begin{quote}
+\begin{description}
+\item[1$'''$.] Infants can represent false beliefs \emph{using a comparatively simple measure} and \emph{in a modular process} from around their first birthday or earlier.
+%
+\item[2$'''$.] Being able to represent false beliefs \emph{using a comparatively sophisticated measure} and  \emph{in a non-modular process}  involves being able to  (i) process perspective differences or (ii) reason counterfactually (or both).
+%
+\item[3$'''$.] Infants cannot (i) process perspective differences nor (ii) engage in counterfactual reasoning until they are at least one year old.  
+\end{description}
+\end{quote}
+%
+At this point the claims are no longer contradictory.
+But it remains an open question whether we now have all the ingredients necessary to explain both early success on Category B false belief tasks and late failure on Category A tasks or whether  some further complexity in the notion of representing belief also needs to be added to the mix.\footnote{
+\citet[p.\ 155-6]{Apperly:2010kx} appears to suggest a solution additionally involving egocentric bias, but I'm not yet fully convinced.
+}
+
+\section{Conclusion}
+In conclusion,
+I have argued that  Eva's and Matthias' findings generate a puzzle  resolving which may require confronting the question, What is a representation of belief?
 

### 2012-03-04 intention motor representation with corrado sinigaglia/FORMAT SECTION WIP intention and motor action.tex
new files

%!TEX TS-program = xelatex
%!TEX encoding = UTF-8 Unicode

\def \papersize {a4paper}

\documentclass[12pt,\papersize]{extarticle}
% extarticle is like article but can handle 8pt, 9pt, 10pt, 11pt, 12pt, 14pt, 17pt, and 20pt text

\def \ititle {Intention and motor representation:}
\def \isubtitle {how to resist temptation}
\def \iauthor {S. Butterfill* \& C. Sinigaglia**
\\ 
**Department of Philosophy, University of Warwick
\\ 
***Dipartimento di Filosofia, Università degli Studi di Milano}
\def \iemail{s.butterfill@warwick.ac.uk}
%\date{}

\input{$HOME/Documents/submissions/preamble_steve_paper2}
%\author{}
%\date{}

%\setromanfont[Mapping=tex-text]{Sabon LT Std}

\begin{document}

\setlength\footnotesep{1em}

\bibliographystyle{newapa} %apalike

\maketitle
%\tableofcontents
\title{}

\tolerance=5000

\section{A motor format for representation}

The second temptation is to suppose that motor representation, or some species of it, is a variety of intention, where intention is understood in the standard way as a propositional attitude with a characteristic role in practical reasoning \citep{Bratman:1987xw}. In this section we explain why this temptation should also be resisted.

As background we first need a generic distinction between content and format. Imagine you are in an unfamiliar city and are trying to get to the central station. A stranger offers you two routes. Each route could be represented by a distinct line on a paper map. The difference between the two lines is a difference in content. Each of the routes could alternatively have been represented by a distinct series of instructions written on the same piece of paper; these cartographic and propositional representations differ in format. The format of a representation constrains its possible contents. For example, a representation with a cartographic format cannot represent what is represented by sentences such as `There could not be a mountain whose summit is inaccessible.'\footnote{Note that the distinction between content and format is orthogonal to issues about representational medium. The maps in our illustration may be paper map or electronic maps, and the instructions may be spoken, signed or written. This difference is one of medium.} The distinction between content and format is necessary because, as our illustration shows, each can be varied independently of the other.

Our aim in this section is to show that motor representations differ from intentions with respect to their format. This is why the second temptation must be resisted. That motor representations differ in format from intentions shows that they are genuinely distinct phenomena.\footnote{ Readers already convinced that motor representation differs from intention in being non-conceptual will not need the following argument in order to conclude that they are distinct phenomena. However the following considerations also indicate that motor and perceptual representations differ in format, which will be relevant later when we consider how motor representations and intentions jointly lead to action.}

How in general can we identify or distinguish representational formats? Because a representational format is typically associated with characteristic processing profiles, it is sometimes possible to infer similarities and differences in representational format from similarities and differences in the processes in which representations feature. This works both for artefactual and mental representations.  To illustrate in terms of our earlier example, suppose that two people have representations of the same route but for one person the route is represented by a line on a map (so in a cartographic format) whereas the other person has a propositional representation of the route. Some transformations are likely to be easier for the person with the cartographic route representation (depending on the projection used, of course); examples include reversing the route, determining how many times a certain river is crossed and transforming the route into a sequence of compass bearings. Other transformations, such as turning the route into a list of salient landmarks, may be easier for the person with the propositional route description. So some patterns of difference in the two people's performances may be explained by the difference in the formats of their representations. If we did not already know that the two people's route representations involved different formats, we might infer this from the facility with which each performed various transformations of the route.

Cognitive neuroscience frequently depends on inferences of just this type. To illustrate, compare imagining seeing an object moving and with actually seeing it move. For this comparison we need to distinguish two ways of imagining seeing. There is a way of imagining seeing which phenomenologically is something like seeing except that it does not necessarily involve being receptive to stimuli. This way of imagining seeing, sometimes called 'sensory imagining', is commonly distinguished from more cognitive ways of imagining seeing which might for example involve thinking about seeing \citep[§2.1]{Gendler:2011_imagination}.%
\footnote{Note that we define sensory imagining in terms of phenomenology and make no stipulations about the processes and representations involved.  This is essential for our purposes, since we wish to consider how evidence bears on a conjecture about the format of representations involved in sensory imagining.}
It is sensorily imagining seeing an object move that we wish to compare with actually seeing an object move.  These two share some processing characteristics.  For instance, whether an object can be seen all at once depends on its size and distance from the perceiver; strikingly, when subjects imagine seeing an object, whether they can imagine seeing it all at once depends in the same way on size and distance (\citealp{kosslyn:1978_measuring}; \citealp[p.\ 99ff]{kosslyn:1994_image}.  Also, how long it takes to imagine looking over an object depends on the object's subjective size in the same way that how long it would take to actually look over that object would depend on its subjective size \citep{kosslyn:1978_visual}.%
\footnote{These and further examples are discussed by \citet[p.\ 165]{currie:1997_mental}.} Further, imagining seeing something (for example, imagining seeing a visual mask) can modulate and interfere with actually seeing in much the way that actually seeing the thing imagined would \citep{pearson:2008_functional,ishai:1995_common}. The similarities in processing profile and the particular patterns of interference are good (if non-decisive) reasons to conjecture that imagining seeing and actually seeing involve representations with a common format.  This conjecture is indirectly supported by evidence that sensorily imagining seeing and actually seeing not only have a common neural basis but also involve similar patterns of cortical activation \citep[e.g.][]{page:2011_erp}.

Let us turn to motor representation. Compare imagining moving a ball with actually moving a ball.  To fully specify the comparison we intend, it is again necessary to distinguish two ways of imagining. One way of imagining action is phenomenologically something like acting except that such imaginings are not necessarily responsive to the features of actual objects and do not necessarily result in bodily movements. To illustrate suppose you are about to dive into a pool and, standing at the edge, mentally pantomime launching yourself from the bank. Some phenomenal characteristics of this imaginative exercise may be barely distinguishable from some of those associated with actually launching yourself into the pool. This way of imagining action can be distinguished from more cognitive ways of imagining action which might involve thinking about an action.% 
\footnote{
On distinguishing these two ways of imagining action, see \citet[p.\ 161]{currie:1997_mental}, \citet[p.\ 727]{jeannerod:1995_mental}, and \citet[p.\ 638-9]{kosslyn:2001_neural}. The former, phenomenologically action-like imagining is sometimes labelled 'motor' or 'internal' and occasionally identified by its links to motor processes or by features of the format or content of the representations involved (\citealp[p.\ 1400]{annett:1995_motor}). We avoid these labels because we introduce the distinction by appeal to phenomenology only and do not stipulate that motor representations are involved. It is essential for what follows that the involvement of motor representations in the phenomenologically action-like way of imagining action is a discovery rather than a terminological stipulation.}
%
The comparison we intend is between imagining moving a ball in the former, phenomenologically-action like way and actually moving a ball. There is evidence that the way imagining performing an action unfolds in time is similar in some respects to the way actually performing an action of the same type would unfold.  For instance, how long it takes to imagine moving an object is closely related to how long it would take to actually move that object \citep{decety:1989_timing, decety:1996_imagined, Jeannerod:1994oz}. In addition, for object-related actions such as grasping the handle of a cup, manipulating the object in ways that make the action harder (such as orienting the handle to make it less convenient for you to grasp) make a corresponding difference to the effort involved in imagining performing the action \citep{parsons:1994_temporal, frak:2001_orientation}. Further, imagining performing an action can selectively interfere with performance of a related action. For example, suppose you are faced with an array of objects one of which---the target---you will shortly be required to grasp.  Subjects who imagine grasping an object other than the target object tend to be slower in subsequently grasping the target object than subjects who do not imagine acting or subjects who imagine grasping the target object \citep{ramsey:2010_incongruent_}. Just as the similarities between imagining seeing and actually seeing are evidence for the hypothesis that the representations involved in imagining seeing and actually seeing have a common format, so also the similarities in processing profile between imagining acting and actually acting, together with the particular patterns of interference between the two, suggest that imagining acting and actually acting involve a common representational format. And also much as in the case of seeing and imagining seeing, acting and imagining acting involve many of the same processes up to the actual muscle contractions \citep{jeannerod:1995_mental, jeannerod:2003_mechanism}.

The fact that motor representation is involved in one way of imagining acting means that we can use imagining acting to investigate the format of motor representations. This enables us to examine characteristics of processes involving motor representation which are are not a direct consequence of limits on agents' abilities to act.

Differences in processing profile provide reasons to distinguish the formats of visual and motor representations. To see why, contrast the way of imagining seeing that is phenomenologically something like seeing with the way of imagining action that is phenomenologically something like acting. To make things concrete, contrast imagining seeing a ball rotating with imagining rotating a ball.%
\footnote{It may be difficult in practice to imagine acting without also imagining seeing, and conversely; it may also sometimes be difficult to distinguish imagine acting from imagining seeing \citep[as][p.\ 170 note]{currie:1997_mental}. However ordinary subjects can separate the two well enough to confirm predictions about their differences \citep[see, e.g.,][]{kosslyn:2001_imagining}.
}
We know that the way imagining acting unfolds in time is quite different from the way imagining seeing unfolds because, as already mentioned, how quickly the former can be done is a function of how long it would take the agent to rotate the ball, whereas how quickly the latter can be done depends on at how rapidly the ball can rotate and still be perceived as rotating. Similarly, factors making actually acting more effortful also make imagining acting more effortful whereas such factors are unlikely to affect the effort involved in imagining seeing. The explanation for these differences in characteristic processing profile between imagining acting and imagining seeing plausibly involves a difference in the formats of the representations involved. That the two involve different representational formats is consistent with  evidence that these two imaginings also involve different processes \citep{kosslyn:2001_imagining}. For instance, each can be selectively impaired \citep{sirigu:2011_motor}; and factors such as limb amputation or hand posture can selectively interfere with imagining performing an action \citep{nico:2004_left, vargas:2004s_influence,  fourkas:2006_influence}.

***HERE***

In principle someone might attempt to explain differences in characteristic processing profile between imagining seeing a ball move and imagining moving the ball by appeal to differences in what is imagined rather than in format. We do not dispute that some differences might be explained by the fact that an action is being imagined in one case only.  But could this fully explain differences in characteristic processing profile?  To see that it could not, note that there are limits on what one can imagine doing which are not limits on what one can imagined seeing being done.  For instance, an agent that cannot bark/reach the ceiling will be unable to imagine---in the way of imagining that is phenomenologically action-like---barking/reaching the ceiling, but will of course be able to imagine seeing and hearing herself barking/reaching the ceiling.  The limits cannot be explained by differences in what is imagined but must be explained by differences in representational format.  ***does this work?

*** OR contrast imagine seeing an action vs. imagining performing an action. 
I.e. need same thing represented, once in motor and once in perceptual format.  If there are limits on specifically motor representation, these cannot be explained by content alone; but they can be explained by format.

*** Conson et al 2010

***NEXT : show that motor representation is not propositional in format.

-----------------------------

Recall that our aim in this section is to show that motor representations differ from intentions with respect to their format.  
\bibliography{$HOME/endnote/phd_biblio}

\end{document}

### 2012-03-07 conf joint action --- sharing goals vs sharing intentions cambridge 2012 talk/intention and motor representation in joint action.tex
talk put into tex

%!TEX TS-program = xelatex
%!TEX encoding = UTF-8 Unicode

\def \papersize {a4paper}

\documentclass[12pt,\papersize]{extarticle}
% extarticle is like article but can handle 8pt, 9pt, 10pt, 11pt, 12pt, 14pt, 17pt, and 20pt text

\def \ititle {Intention and motor representation:}
\def \isubtitle {how to resist temptation}
\def \iauthor {S. Butterfill* \& C. Sinigaglia**
\\ 
**Department of Philosophy, University of Warwick
\\ 
***Dipartimento di Filosofia, Università degli Studi di Milano}
\def \iemail{s.butterfill@warwick.ac.uk}
%\date{}

\input{$HOME/Documents/submissions/preamble_steve_paper2}
%\author{}
%\date{}

%\setromanfont[Mapping=tex-text]{Sabon LT Std}

\begin{document}

\setlength\footnotesep{1em}

\bibliographystyle{newapa} %apalike

\maketitle
%\tableofcontents
\title{}

\begin{abstract}
On the assumption that social motor representation plays a role in explaining how effective joint action is possible, do we also need motor representation to explain what joint action is?  Philosophers tend to assume that motor representation is only an enabling condition for joint action and of no direct interest to narrowly philosophical theories of joint action and shared intention.  In this talk I shall argue that social motor representation and shared intention  have distinctive roles in explaining the purposiveness of joint action.  This gives rise to a challenge.  On the one hand, effective joint action---imagine two people erecting a tent in a gale together---sometimes requires both shared intentions and social motor representations plus a certain kind of harmony between the two.  On the other hand, recognizing their distinctive roles precludes the existence of direct inferential links between shared intentions and social motor representations.  The challenge is to explain how these two kinds of representation could sometimes harmoniously contribute to effective joint action despite the lack of inferential integration.
\end{abstract}

\section{Slide}
As a philosopher I’m going to start from an empirically controversial premise.
I don’t think this premise has been established, but I do think it’s a reasonable bet.

\section{Slide}
A motor representation is \emph{agent-neutral} if it concerns an action which is not one’s one or, in the case of joint action, not entirely one’s own.

\section{Slide}
Two or more motor representations are \emph{reciprocal} just if there is a single outcome which each motor representation represents.

\section{Slide}
Sometimes I’ll using the term \emph{social} motor representation for reciprocal agent-neutral motor representations.
I have a slightly bad conscience about this because I’m not sure they’re actually social in any ordinary sense.

It is almost uncontroversial that agent-neutral motor representations exist, and I see no reason to doubt that they could be reciprocal.
The controversial part of the premise is the claim that such representations can enable joint action.

I should say, by the way, that I’m not suggesting that reciprocal agent-neutral motor representation is involved in every joint action.
Perhaps some joint actions do not involve social motor representation.
The premise is just that social motor representation is among the factors which enable some joint actions.

\section{Slide}
I’m not going to defend the premise, although I will try to explain the premise in more detail in a moment.  But first let me outline where I’m going with this.

\section{Slide}
The premise leads to a question and, relatedly, to a challenge.

\section{Slide}
The question is whether social motor representation is not only an enabling condition for joint action but also plays a role in explaining what joint action is. I want to argue that it does.  In fact I want to argue that it plays a role similar and complementary to that of shared intention.

\section{Slide}
After this, in the last part of my talk, I also want to point to a challenge raised by the existence of social motor representation.  The challenge is roughly this: planning for joint action involves representations of target outcomes in two different formats, a motor format and a propositional format.  Given that representations of these different formats cannot straightforwardly be inferentially integrated, what could ever insure that there is sometimes non-accidental harmony between motor representations and shared intentions?

\section{Slide}
Let me start by explaining the premise.

\section{Slide}
Why think this?
A direct way to test the premise would be to selectively intervene on social motor representation and see how it affects agents’ performance of joint actions.
As far as I know no one has done this.
But there are some indirect findings.
I shall mention just one, a finding by Dimitris Kourtis and colleagues.

\section{Slide}
They found that `the social relation between individuals modulates action simulation' \citep[p.\ 1]{kourtis:2010_favoritism}.  
In slightly more detail:
\begin{quote}
`motor activation during action anticipation depends on the ... relation between the actor and the observer ... Simulation of another person’s action, as reﬂected in the activation of motor cortices, gets stronger the more the other is perceived as an interaction partner.’  \citep[p.\ 4]{kourtis:2010_favoritism}
\end{quote}
I realise this doesn’t conclusively support the premise.  But I think it is evidence in its favour.

\section{Slide}
Here’s the experiment they did.
The task was simple: two people sat opposite each other.  Sometimes they acted alone, picking up and replacing an object.  And sometimes had to act together, passing an object between them.  Also present was the ‘loner’ who always acted alone.

\section{Slide}
EEG measurements of motor activation were recorded.

\section{Slide}
[image shrinks]

\section{Slide}
The researchers found that patterns of activation for self and joint action partner were similar, and different from patterns of activation for the actions of the loner which were similar to patterns in a ‘no go’ condition where no one was to act.

\section{Slide}
So here’s what I take from this paper.
If you are engaged in a joint action with someone, one which involves moving an object by passing it between you, then each of you has motor representations of the other’s actions and these motor representations are functionally equivalent to motor representations of your own actions in the sense that they are just the sorts of representation that might have caused you to do what the other is doing (if you were in her position).

\section{Slide}
Suppose that Kourtis et al are right that social motor representation is more likely to occur in joint action than when one is merely observing.
How could this facilitate joint action?

\section{Slide}
Let’s step back and consider an individual action.
An agent moves a mug from one place to another, passing in from her left hand to her right hand half way [*demonstrate].
It’s a familiar idea that motor representations for planning and monitoring action involve an hierarchical structure,
where there is a relatively abstract representation of an outcome that is progressively filled in.

\section{Slide}
I don’t suppose that this attempt to depict the hierarchy of motor representations involved in moving a mug with two hands is accurate.
I’m only trying to illustrate two familiar ideas.
One is that motor planning involves starting with relatively abstract representations of outcomes and filling in details.

\section{Slide}
The other is that there is a need, even for a single agent, to synchronize the exchange between the two hands.

How is this relevant to the case of joint action?

\section{Slide}
In joint action the agents have the same goal, to move the object from there to here.

\section{Slide}
They also face a similar coordination problem, requiring a precisely timed swap

\section{Slide}
And Koutis et al’s findings (and others’ findings) suggest that the same planning in involved in the joint action case, almost up to the actual muscle contractions.
That is, in the joint action situation each agent plans both agents’ actions as if they were the actions of a single agent.
This may be what enables them to coordinate so well : each is able to plan her own actions in a way that meshes with the other agent’s actions because each agent is planning (and monitoring) both their actions almost as if a single agent were going to execute the whole action.
And of course this is exactly what we want for small-scale joint action---we want two or more agents to act as one.

\section{Slide}
So what is the difference between the individual and the joint case?  From the point of view of motor representation, the primary difference may be that in joint action there is a need to inhibit execution of the parts of the action which are not one’s own.

Here then is the basic idea I take to be guiding Kourtis and others.
The idea is that coordination is sometimes achieved by having each agent’s motor system plan all of their actions; 
given some assumptions, this could be a way of making it likely that each will execute their part in the joint action in a way that meshes with the way the other agents execute their parts.

Now so far I’ve only been considering a possible role for social motor representation in enabling joint action.  
But I think we the details already give us grounds for holding that motor representation has a role to play in explaining what joint action is.
How does this work?
Let’s go back to individual action for a moment again.

\section{Slide}
A basic question about ordinary, individual action is:
What is the relation between a purposive action and the outcome or outcomes to which it is directed?

Many ordinary purposive actions have many different outcomes.
Grabbing little Isabel by the hands I swing her around, causing her to laugh and, simultaneously, breaking a vase.
In fact the outcome to which this purposive action was directed might not be among its actual outcomes; after all, actions can fail.

\section{Slide}
So among all the actual and possible outcomes of my action, one or some are singled out as specially related to this action.
One aspect of the question concerns what singles out the outcome or outcomes, actual or merely possible, to which a particular purposive is directed.
But there is also a second aspect ...

\section{Slide}
Ordinary purposive actions are sometimes composed of more than one motor action.  My swinging Isabel around starts with my reaching for her wrists, grasping them and then spinning us around ... and my action doesn’t include other things which I might be doing simultaneously, like refusing a cup of tea with my eyes orSo among all the actual and possible outcomes of my action, one or some are singled out as specially related to this action.
One aspect of the question concerns what singles out the outcome or outcomes, actual or merely possible, to which a particular purposive is directed.
But there is also a second aspect ...
 trying to determine whether that smell is coming from Isabel’s sister Hannah’s nappy.

\section{Slide}
So another aspect of our question is what determines which activities comprise the purposive action and which do not.

\section{Slide}
The standard answer to this question involves intention.
An intention (1) represents an outcome, (2) coordinates the one or several activities which comprise the action; and (3) coordinate these activities in a way that would normally facilitate the outcome’s occurrence.

What binds particular component actions together into larger purposive actions?  It is the fact that these actions are all parts of plans involving a single intention.
What singles out an actual or possible outcome as one to which the component actions are collectively directed?  It is the fact that this outcome is represented by the intention.

So the intention is what binds component actions together into purposive actions and links the action taken as a whole to the outcomes to which they are directed.

\section{Slide}
Now as Elisabeth Pacherie has argued \citep[pp.\ 189-90]{pacherie:2008_action} (and I’ve had a go at arguing this in joint work with Corrado Sinigaglia recently too),
motor representations are relevantly similar to intentions.

Of course motor representations differ from intentions in some important ways (as Pacherie also notes).

But they are similar in the respects that matter for explaining the purposiveness of action.
(1) Like intentions, some motor representations represent outcomes (and not merely patters of joint displacement, say).
(2) Like intentions, some motor representations play a role in coordinating multiple more component activities by virtue of their role as elements in hierarchically structured plans.
(3) And, like intentions, some motor representations coordinate these activities in a way that would normally facilitate the outcome’s occurrence.

So in the individual case, it seems to me quite straightforward that there is a role for motor representation to play in explaining the purposiveness of action [*explaining the possibility of purposive action?].

The claim is not that \emph{all} purposive actions are linked to outcomes by motor representations, just that some are.
In some cases, the purposiveness of an action is grounded in a motor representation of an outcome; in other cases it is grounded in an intention.
And of course in many cases it may be that both intention and motor representation are involved.

\section{Slide}
Now let’s turn to joint action.

\section{Slide}
The same question we asked about ordinary, individual action also arises for joint action.
What is the relation between a purposive joint action and the outcome or outcomes to which it is directed?

\section{Slide}
And again the question has two aspects.
What singles out the outcome or outcomes to which a purposive joint action is directed?

\section{Slide}
And what binds together the various activities that make up the joint action?
The difference in the case of joint action is, of course, that these activities are not necessarily activities of a single agent.

\section{Slide}
The answer to this question for the case of joint action is also superficially similar in the answer we gave in the case of ordinary, individual action.

A shared intention is what relates purposive joint actions to the outcomes to which they are directed.
For the shared intention
(1) involves a representation, on the part of each agent, of an outcome
(2) coordinates the several agents’ activities
and 
(3) coordinates the several agents’ activities in such a that would normally facilitate the occurrence of the represented outcome.

It is in this sense that a shared intention can ground the purposiveness of a joint action.

\section{Slide}
But what we saw earlier, what the research by Kourtis et al and others indicates, is that social motor representation can play a similar role to joint action.

Return to the example of a two agents moving an object in a way that involves passing it between them.
the motor representation
(1) involves a representation, on the part of each agent, of an outcome
In this case the outcome is the whole movement of the object
(2) coordinates the several agents’ activities
and 
(3) coordinates the several agents’ activities in such a that would normally facilitate the occurrence of the represented outcome.
Because each agent represents the whole movement and plans all of its implementation irrespective of which parts she will actually perform, each agent plans the action in a way that should coordinate with the other agent’s plans providing they use similar planning procedures

[*What I’m saying here, in effect, is that both shared intention and social motor representation can yield a COLLECTIVE GOAL]

\section{Slide}
What I’m suggesting is very simple.

If you think that in ordinary, individual action, the purposiveness of actions can be grounded by motor representations
(and you should think this because it’s true),
then you should also think the same about actions involving two or more agents---the purposiveness of a joint action can be grounded in motor representations as well as in shared intentions.

\section{Slide}
Let me try another way of presenting the same idea.
Here are Michael Bratman’s sufficient conditions for shared intention.
I want to suggest that social motor representation provides a parallel

\section{Slide}
There is a direct parallel with the first condition: in the case of motor representation, each agent represents the outcome (e.g. the movement of the object).
The key claim here is that some motor representations (i) represent outcomes, and (ii) represent the outcomes of actions not all of whose components will be executed by the agent whose motor representation it is.

\section{Slide}
Here there is clearly no \emph{direct} parallel. 
I don’t think motor representations can represent motor representations in the way that intentions can represent intentions.
But I do think there is a parallel of sorts.
Each agent’s having a motor representation of the distributed goal of their action does ensure meshing of subplans.
What ensures this meshing is not the fact that each agent represents the others’ plans.
Rather in the case where joint actions is grounded in social motor representation, what ensures meshing of subplans is two facts (i) each agent plans all of the agents’ actions, and (ii) the agents rely on similar planning strategies (planning strategies that are sufficiently similar to ensure meshing subplans).

\section{Slide}
I’m less sure about a parallel to the common knowledge condition.
As I see things, the justification for supposing that shared intention involves common knowledge concerns a normative link between intention and reasons.
In acting on intentions, one should be acting for reasons.
And a consideration can only be among your reasons if you know that consideration.
So I think the need for common knowledge arises from the need to explain how another person’s intentions could be among your reasons for acting.
I don’t think this need arises in the case of motor representation because it seems to me that the sort of planning of which motor representation is an element does not involve acting for reasons in the same sense.  
(In motor action, there are reasons why we do things (of course!) but these are not  reasons for which we act.)
What motor joint action requires is not that your motor plans provide reasons for mine.
There just has to be a good chance that this is true relative to the costs and benefits of joint action and the alternatives to joint action.
So I think that instead of common *knowledge*, in the case of social motor representation there is a common *background* of dispositions, habits and expectations.%
\footnote{
[***CUT but one thing that might do the work of common knowledge is a custom or habit 
that would allow the agents, in their particular social context, to rely on each other’s cooperation.
In some countries this sort of thing works on public transport; 
it is reasonable to take for granted that, if you are obviously struggling with a pram or suitcase, then someone nearby will help.]
}

\section{Slide}
If this is right, if social motor representations play a role analogous to the structure of intentions and knowledge which Bratman identifies as sufficient for shared intention, then this is another reason to think that motor representations can ground the purposiveness of joint action.

\section{Slide}
Recap: the question was: Does social motor representation also play a role in explaining what joint is?
I have just been arguing for a positive answer.
My thesis is this:
\textbf{
Reciprocal social motor representations coordinate multiple agents’ actions by virtue of representing an outcome to which each agent’s actions are directed.
}
That is, reciprocal social motor representations can ground the purposiveness of joint action.
This is why I think that fully understanding what joint action is requires understanding the coordinating role of social motor representation and not only understand shared intention.

I don’t mean to suggest that all joint actions involve social motor representation.  Surely some joint actions do not.  My claim is not that all joint actions involve social motor representation.

But, equally, \emph{there could be purposive joint actions which do not involve shared intentions},
just as there can be purposive actions which do not involve intentions at all.
Let me go slowly in explaining why I think there could be joint action without shared intention and start by returning to the case of ordinary, individual action.

\section{Slide}
So what are intentions for?

I’m going to assume that intention is something over and above basic beliefs and desires; that an intention is not, for instance, merely a strongest desire.

There is a temptation to assume that intention is involved in every case of purposive action.
But it’s hard to see what the argument for this could be.
In many cases it seems that beliefs, desires and motor representations are all that is needed to explain purposive action. 
You offer me a biscuit.  I want one, and I believe I can get one by reaching out for it.  So I do reach for it.  As far as I can see, there’s no need to suppose that, in addition to the belief and desire, it must be the case that I also intend to take a biscuit.  
(At least not unless we take ‘intention’ to mean ‘strongest desire’, which it does not.)  
Maybe I do intend this.  
But it’s possible for an agent to take and eat a biscuit, and to do so purposively, without having any intentions at all.  
Beliefs desires and motor representations are sufficient.

So (again), what are intentions for?

\section{Slide}
Motor representations enable quite sophisticated planning over short periods of time and sequences of action; for example, how you grasp a pointer will depend on what you are about to do with it \citep{zhang:2007_planning}.

This sort of planning does not need intentions at all.  So what are intentions for?

\section{Slide}
\emph{Intentions are for planning multiple separate actions over longer periods of time; and for planning multiple separate actions whose execution is mutually constraining where the outcomes cannot be represented motorically.}

This is a case where intentions are really needed --- can’t act on strongest desire (for the big reward) if want to maximise rewards by collecting the small and the large reward.
And can’t rely on motor representation because the motor system doesn’t care about things that cannot be represented in motor terms.

Not all purposive actions involve any planning of this sort.

Now you could imagine a two-person version of this task where we are rewarded for what we collectively achieve.  In this case it’s optimal if one of us goes for the small reward and the other goes for the large reward.  I think it’s this kind of planning that shared intention is really for.

\section{Slide}
By contrast, in many ordinary cases of joint action there is no need for planning of this sort and so no need for shared intention.  Actions such as these \emph{might} involve shared intention but they do not \emph{necessarily} involve shared intention.

I’m suggesting that some joint actions---like the one two people move an object in a way that involves passing it between them---don’t require this kind of planning and so don’t necessarily involve shared intentions.
In some cases, social motor representation alone is sufficient for purposive joint action.

\section{Slide}
This goes against a widely shared view in the literature

\section{Slide}
To focus on just one case, Alonso says ‘the key property of joint action lies in its internal component [...] in the participants’ having a “collective” or “shared” intention.’ \citep[pp.\ 444-5]{alonso_shared_2009}

I want to spend a bit of time on this claim because it matters for my second theme, which is a problem concerning how motor representation and shared intention interface in joint action.

As far as I know, this claim is not explicitly argued for.  From conversation with philosophers I think the central argument is supposed to contrast cases.
Contrast cases are pairs of actions which are similar in terms of behaviour and coordination but where one is joint but the other isn’t....

\section{Slide}
[nothing]

\section{Slide}
[Gilbert's Jack and Sue walking together case]

\section{Slide}
[Searle's park vistor contrast case]

These sorts of contrast case invite the question, 
How do joint actions differ from individual actions which may occur in parallel? 
What is the difference between Jack and Sue walking together and their walking side-by-side?  
Gilbert’s example shows that the difference can’t just be a matter of coordination, because people merely walking alongside each other also need to coordinate their actions in order to avoid colliding with each other.  
And Searle’s example shows that the difference between joint action and parallel individual action can’t just be that the actions have a common effect because merely parallel actions can have common effects too.

How might this lead someone to think that all joint action involves shared intention?
I think the idea is supposed to be this.
One difference between the genuinely joint actions and their merely parallel counterparts is that the genuinely joint actions involve shared intentions.
And there is no further difference that enables one systematically to distinguish the two cases.

I think this argument is mistaken.
First because there are contrast cases in which the joint action does not involve shared intention (***bouncing a cube on a trampoline---might be done by two people merely shaking the tramoline with no interest in, or awareness of, each other).
It follows that appeal to shared intention does not provide a sufficiently general way of explaining the difference illustrated by the contrast cases.
Second because there is a more general notion appeal to which enables us to distinguish the contrast cases.
Genuine joint action differs from merely parallel action because the former involves each agent’s representing an outcome to which all of their actions are directed where these outcome representations coordinate their actions in a way that would normally facilitate their collective success in bring about this outcome.

\section{Slide}
(Actually I don’t think this is quite general enough because I think some joint actions involve non-representational coordinative structures only; I am also doubtful that there is a sharp distinction between merely parallel and genuinely joint action and I think it is possible to see the difference as a matter of degree. But I don’t want to get into that here.  It’s sufficient that we have moved away from the bare shared intention account. [*WHAT WE REALLY NEED IS A COLLECTIVE OR SHARED GOAL, and the possibility of gradual construction shows that there’s no magic moment separating joint from parallel action.)

To sum up so far, I reject the claim that reflection on the contrast cases provides any support for the idea that all joint action involves shared intention.

\section{Slide}
So far I have been assuming that there is an inconsistency between two claims:
(1) my claim that reciprocal reciprocal agent-neutral motor representation can ground purposive joint action
(2) many philosophers’ claim that all joint actions involve shared intention.

But maybe it is a mistake to think that these are inconsistent.
Why not suppose that some social motor representations are a variety of shared intentions?

\section{Slide}
If some reciprocal agent-neutral motor representations are shared intentions, then there is no inconsistency.
And what I’ve just been arguing is that agent-neutral motor representations resemble shared intentions in that both play a role in coordinating agents’ actions by virtue of representing outcomes.  Isn’t that enough to justify identifying them as shared intentions?

This issue might easily seem narrowly conceptual or terminological.  
At the end of the day it doesn’t much matter if we want to call some motor representations ‘shared intentions’.  
After all, on most accounts shared intentions are neither shared nor intentions so we would hardly be doing more violence to the term than is already being done.  
So insofar as labeling some social motor representations shared intentions might help to avoid unnecessarily philosophical discussion, I’m all in favour.
However, there is an important differences between the states normally regarded as  shared intentions and any motor representations.
And this difference matters for understanding the interface between shared intention and motor representation.

[*PLAN for what follows:*]

Difference in format.
Leads to the interface problem.  
Shared intentions can be inferentially integrated with other shared intentions; but not they cannot be inferentially integrated with social motor intentions (two disjoint planning processes).  
Elisabeth Pacherie’s proposal: shared intentions set outcomes to be achieved by social motor representations.  
I borrow this idea from her, but it raises a further problem.
The problem is how the one sets outcomes for the other given the difference in representational format.

\section{Slide}
As background we first need a generic distinction between content and format. Imagine you are in an unfamiliar city and are trying to get to the central station. A stranger offers you two routes. Each route could be represented by a distinct line on a paper map. The difference between the two lines is a difference in content. Each of the routes could alternatively have been represented by a distinct series of instructions written on the same piece of paper; these cartographic and propositional representations differ in format. The format of a representation constrains its possible contents. For example, a representation with a cartographic format cannot represent what is represented by sentences such as `There could not be a mountain whose summit is inaccessible.'\footnote{ Note that the distinction between content and format is orthogonal to issues about representational medium. The maps in our illustration may be paper map or electronic maps, and the instructions may be spoken, signed or written. This difference is one of medium.} The distinction between content and format is necessary because, as our illustration shows, each can be varied independently of the other.

\section{Slide}
Format matters because only where two representations have the same format can they be straightforwardly inferentially integrated.

To illustrate, let’s stay with representations of routes.  
Suppose you are given some verbal instructions describing a route. You are then shown a representation of a route on a map and asked whether this is the same route that was verbally described. You are not allowed to find out by following the routes or by imagining following them. 
Special cases aside, answering the question will involve a process of translation because two distinct representational formats are involved, propositional and cartographic. It is not be enough that you could follow either representation of the route. You will also need to be able to translate from at least one representational format into at least one other format.

\section{Slide}
This brings me to the argument ...

\begin{enumerate}
\item Only representations with a common format can be inferentially integrated.

\item Any two intentions can be inferentially integrated in practical reasoning.

\item My intention that I visit Paris on Friday is a propositional attitude.
\end{enumerate}

\section{Slide}
It follows that 
All intentions are propositional attitudes

\section{Slide}
But it is also widely agreed that:
No motor representations are propositional attitudes.

\section{Slide}
So we can conclude that:
 No motor representations are intentions

\section{Slide}
This step---Any two intentions can be inferentially integrated in practical reasoning---is questionable.  I don’t have an argument for this and I’m not sure it isn’t terminological.  
What I care about is that we distinguish attitudes according to the processes in which they feature.
So if you like we could distinguish two kinds of intention, one propositional the other motor.
As long as we distinguish representations of different formats I don’t see that it matters too much whether we call them all intentions or whether we use that term for only some of them.

\section{Slide}
So where does this leave us?
The question was whether reciprocal agent-neutral motor intentions could count as shared intentions.

\section{Slide}
In answer to that question, I think this:
(a)
IF you agree motor representations are not intentions,
THEN reciprocal agent-neutral motor representations are not shared intentions.

And
(b)
IF motor representations are a non-propositional variety of intention,
THEN reciprocal agent-neutral motor representations are a non-standard variety of shared intentions.

The key thing is that, either way, reciprocal agent-neutral motor representations cannot be inferentially integrated with shared intentions in practical reasoning.
This leads to what I’ll call ‘The Interface Problem’

\section{Slide}
It will take me a moment to explain what the problem is.

\section{Slide}
The first step is to note that:
Some joint actions involve both shared  intention and reciprocal agent-neutral motor representation.
Imagine two people setting up a tend in a gale together, for example.
It seems success will often require many ingredients including shared intention for the large-scale planning and reciprocal agent-neutral motor representation for passing objects and bending the poles together.

\section{Slide}
Earlier I argued that reciprocal agent-neutral motor representations:

\begin{itemize}
\item represent outcomes;

\item ground the purposiveness of some 
joint actions; and
\end{itemize}

\section{Slide}
This means, I think, that in at least some cases of effective joint action there will be a certain kind of harmony between the contents of 
 reciprocal agent-neutral motor representations
 and the contents of 
 shared intentions.
Further, I think it's probable that, in some cases, the harmony is non-accidental.

To be clearer about what `harmony' means, we need the notion of matching:
\begin{quote}
Two  outcomes, A and B, \emph{match} in a particular context just if, in that context, either the occurrence of A would normally constitute or cause, at least partially, the occurrence of B or vice versa. 
\end{quote}
In some cases of joint action, the outcomes each kind of representation specifies non-accidentally match.

\section{Slide}
This leads to a question: how are non-accidental matches between the outcomes specified by shared intentions and by reciprocal agent-neutral motor representations possible?

A natural suggestion is that matching is achieved through practical reasoning.
It is in this way that the contents of shared intentions could partially determine the contents of reciprocal agent-neutral motor representations.
But there is \textbf{an obstacle} to this idea.

\section{Slide}
For as we saw earlier, reciprocal agent-neutral motor representations
\begin{itemize}
\item differ in format from (the constituent attitudes of) shared intentions.
\end{itemize}

Given that only representations with a common format can be inferentially integrated in practical reasoning,
we cannot appeal to reasoning to explain how the outcomes specified by shared intentions match the outcomes specified by reciprocal agent-neutral motor representations.
This is why the interface problem is a problem.

\section{Slide}
In conclusion,
I have suggested that 
reciprocal agent-neutral motor representations
can 
ground purposive joint action
for they
can coordinate two or more agents' actions in virtue of representing a single outcome to which each agent's actions are directed.

I also suggested, further, that
reciprocal agent-neutral motor representation
and
shared intention  
have distinctive roles in explaining the purposiveness of joint action.

And this suggestion has two consequences.
One is that we cannot properly understand what joint action is if we focus only on shared intention.
The other is that we face a challenge.
The challenge is to explain how these two kinds of representation could sometimes harmoniously contribute to effective joint action despite not being inferentially integrated in practical reasoning.

\bibliography{$HOME/endnote/phd_biblio}

\end{document}

### 2012-03-07 intention motor representation with corrado sinigaglia/FORMAT SECTION WIP intention and motor action.tex
wip

@@
 Differences in processing profile provide reasons to distinguish the formats of visual and motor representations.
-To see why, contrast the way of imagining seeing that is phenomenologically something like seeing with the way of imagining action that is phenomenologically something like acting.
+To see why, contrast the way of imagining seeing that is phenomenologically something like seeing with the way of imagining acting that is phenomenologically something like acting.
 To make things concrete, contrast imagining seeing a ball rotating with imagining rotating a ball.%
@@
 }
-We know that the way imagining acting unfolds in time is quite different from the way imagining seeing unfolds because, as already mentioned, how quickly the former can be done is a function of how long it would take the agent to rotate the ball, whereas how quickly the latter can be done depends on at how rapidly the ball can rotate and still be perceived as rotating.
-Similarly, factors making actually acting more effortful also make imagining acting more effortful whereas such factors are unlikely to affect the effort involved in imagining seeing.
+We know that the way imagining acting unfolds in time is quite different from the way imagining seeing unfolds because, as already mentioned, how quickly the former can be done is a function of how long it would take the agent to rotate the ball, whereas how quickly the latter can be done depends on how rapidly the ball can rotate and still be perceived as rotating.
+Similarly, we mentioned that factors making actually acting more effortful also make imagining acting more effortful: such factors are unlikely to affect the effort involved in imagining seeing.
 The explanation for these differences in characteristic processing profile between imagining acting and imagining seeing plausibly involves a difference in the formats of the representations involved.
@@
 
-***HERE***
-
-In principle someone might attempt to explain differences in characteristic processing profile between imagining seeing a ball move and imagining moving the ball by appeal to differences in what is imagined rather than in format.
-We do not dispute that some differences might be explained by the fact that an action is being imagined in one case only.
-But could this fully explain differences in characteristic processing profile?
-To see that it could not, note that there are limits on what one can imagine doing which are not limits on what one can imagined seeing being done.
-For instance, an agent that cannot bark/reach the ceiling will be unable to imagine---in the way of imagining that is phenomenologically action-like---barking/reaching the ceiling, but will of course be able to imagine seeing and hearing herself barking/reaching the ceiling.
-The limits cannot be explained by differences in what is imagined but must be explained by differences in representational format.
-***does this work?
-
-*** OR contrast imagine seeing an action vs. imagining performing an action. 
-I.e.
-need same thing represented, once in motor and once in perceptual format.
-If there are limits on specifically motor representation, these cannot be explained by content alone; but they can be explained by format.
-
-*** Conson et al 2010
-
-***NEXT : show that motor representation is not propositional in format.
+But are we too hasty in drawing this conclusion?
+It might be supposed that imagining seeing a ball rotating differs from imagining moving a ball with respect to the contents involved---after all, one involves an action, the other does not.
+In principle someone might attempt to explain differences in characteristic processing profile between imagining seeing a ball move and imagining moving the ball by appeal to differences in content rather than in format.
+We do not dispute that \emph{some} differences might be explained in this way.
+But could any differences in content fully explain differences in characteristic processing profile?
+To see why not, consider a further case.
+Ordinary subjects who are asked to judged the laterality of a hand rotated to various degrees are slower and less accurate when the hand's position is maximally biomechanically awkward (rotated 180 degrees, middle finger pointing down).
+By contrast, no such effect occurs for comparable tasks involving letters rather than hands.
+This difference in performance has been explained by appeal to the hypothesis that the hand task involves the phenomenologically action-like form of imagination whereas the letter task involves sensory visual imagination (**ref).
+But could this performance difference instead be explained by a difference in content (one task involves representing hands, the other letters)?
+Subjects with Amyotrophic Lateral Sclerosis (ALS) which impairs motor representation retain the ability to judge the laterality of hands (and letters) but their performance is unaffected by whether the hand is presented in a biomechanically awkward position \citep{Fiori:2012fk}.
+The difference between these subjects and ordinary, control subjects is not plausibly a matter of content since both are representing rotated hands; rather, the difference in the processing profile of their laterality judgements is plausibly a reflection of a difference in the formats in which they represent hands.
+This is why we hold that some differences between imagining seeing and imagining acting indicate differences in representational format.
 

### 2012-03-07 conf joint action --- sharing goals vs sharing intentions cambridge 2012 talk/intention and motor representation in joint action.tex
wip
am i done: no handout

@@
 
-\def \ititle {Intention and motor representation:}
-\def \isubtitle {how to resist temptation}
+\def \ititle {Intention and Motor Representation in Joint Action}
+\def \isubtitle {}
 \def \iauthor {S. Butterfill* \& C. Sinigaglia**
@@
 \end{abstract}
+
+\section{Slide}
+[hello]
 
@@
 The challenge is roughly this: planning for joint action involves representations of target outcomes in two different formats, a motor format and a propositional format.  
-Given that representations of these different formats cannot straightforwardly be inferentially integrated, what could ever insure that there is sometimes non-accidental harmony between motor representations and shared intentions?
+Given that representations in these different formats cannot straightforwardly be inferentially integrated in practical reasoning, what could ever insure that there is sometimes non-accidental harmony between motor representations and shared intentions?
 
 \section{Slide}
-Let me start by explaining the premise.
-
-\section{Slide}
+Let me start by explaining the premise: Reciprocal agent-neutral motor representation enables joint action.
 Why think this?
@@
 They found that `the social relation between individuals modulates action simulation' \citep[p.\ 1]{kourtis:2010_favoritism}.
+
+\section{Slide}
 In slightly more detail:
@@
 Now so far I’ve only been considering a possible role for social motor representation in enabling joint action.  
-But I think we the details already give us grounds for holding that motor representation has a role to play in explaining what joint action is.
-How does this work?
-Let’s go back to individual action for a moment again.
+But I think the details already give us grounds for holding that motor representation has a role to play in explaining what joint action is.
+To see why,
+let’s go back to individual action for a moment again.
 
@@
 My swinging Isabel around starts with my reaching for her wrists, grasping them and then spinning us around ...
-and my action doesn’t include other things which I might be doing simultaneously, like refusing a cup of tea with my eyes orSo among all the actual and possible outcomes of my action, one or some are singled out as specially related to this action.
-One aspect of the question concerns what singles out the outcome or outcomes, actual or merely possible, to which a particular purposive is directed.
-But there is also a second aspect ...
- trying to determine whether that smell is coming from Isabel’s sister Hannah’s nappy.
+and my action doesn’t include other things which I might be doing simultaneously, like refusing a cup of tea with my eyes or  trying to determine whether that smell is coming from Isabel’s sister Hannah’s nappy.
 
@@
 But they are similar in the respects that matter for explaining the purposiveness of action.
-(1) Like intentions, some motor representations represent outcomes (and not merely patters of joint displacement, say).
-(2) Like intentions, some motor representations play a role in coordinating multiple more component activities by virtue of their role as elements in hierarchically structured plans.
+(1) Like intentions, some motor representations represent outcomes (and not merely patters of joint displacements, say).
+(2) Like intentions, some motor representations play a role in coordinating multiple  component activities by virtue of their role as elements in hierarchically structured plans.
 (3) And, like intentions, some motor representations coordinate these activities in a way that would normally facilitate the outcome’s occurrence.
@@
 \section{Slide}
-But what we saw earlier, what the research by Kourtis et al and others indicates, is that social motor representation can play a similar role to joint action.
+But what we saw earlier, what the research by Kourtis et al and others indicates, is that social motor representation can play a  role similar to that of shared intention.
 
-Return to the example of a two agents moving an object in a way that involves passing it between them.
-the motor representation
-(1) involves a representation, on the part of each agent, of an outcome
-In this case the outcome is the whole movement of the object
-(2) coordinates the several agents’ activities
+Return to the example of two agents moving an object in a way that involves passing it between them.
+Suppose that their passing involves reciprocal agent-neutral motor representations of the outcome, which is the movement of the object. 
+These motor representations
+(1) represent an outcome to which the joint action is directed,
+(2) coordinate the several agents’ activities
 and 
-(3) coordinates the several agents’ activities in such a that would normally facilitate the occurrence of the represented outcome.
+(3) coordinate the several agents’ activities in a way that would normally facilitate the occurrence of the represented outcome.
 Because each agent represents the whole movement and plans all of its implementation irrespective of which parts she will actually perform, each agent plans the action in a way that should coordinate with the other agent’s plans providing they use similar planning procedures
@@
 Each agent’s having a motor representation of the distributed goal of their action does ensure meshing of subplans.
-What ensures this meshing is not the fact that each agent represents the others’ plans.
-Rather in the case where joint actions is grounded in social motor representation, what ensures meshing of subplans is two facts (i) each agent plans all of the agents’ actions, and (ii) the agents rely on similar planning strategies (planning strategies that are sufficiently similar to ensure meshing subplans).
+What ensures this meshing is not the fact that each agent represents the other's plans \emph{as the other's plans}.
+Rather in the case where joint action is grounded in social motor representation, what ensures meshing of subplans is two facts (i) each agent plans all of the agents’ actions, and (ii) the agents rely on similar planning strategies (planning strategies that are sufficiently similar to ensure meshing subplans).
 
@@
 There just has to be a good chance that this is true relative to the costs and benefits of joint action and the alternatives to joint action.
-So I think that instead of common *knowledge*, in the case of social motor representation there is a common *background* of dispositions, habits and expectations.%
+So \textbf{I think that instead of common \emph{knowledge}, in the case of social motor representation there is a common \emph{background} of dispositions, habits and expectations.}%
 \footnote{
@@
 \section{Slide}
-Recap: the question was: Does social motor representation also play a role in explaining what joint is?
+Recap: the question was: Does social motor representation  play a role in explaining what joint is?
 I have just been arguing for a positive answer.
 My thesis is this:
-\textbf{
-Reciprocal social motor representations coordinate multiple agents’ actions by virtue of representing an outcome to which each agent’s actions are directed.
-}
+\textbf{Reciprocal agent-neutral motor representations coordinate multiple agents’ actions around an outcome in part by virtue of representing that outcome.}
 That is, reciprocal social motor representations can ground the purposiveness of joint action.
-This is why I think that fully understanding what joint action is requires understanding the coordinating role of social motor representation and not only understand shared intention.
+This is why I think that fully understanding what joint action is requires understanding the coordinating role of social motor representation and not only understanding shared intention.
 
-I don’t mean to suggest that all joint actions involve social motor representation.  Surely some joint actions do not.  My claim is not that all joint actions involve social motor representation.
+I don’t mean to suggest that all joint actions involve social motor representation.  Surely some joint actions do not.
 
@@
 
-I’m going to assume that intention is something over and above basic beliefs and desires; that an intention is not, for instance, merely a strongest desire.
+I’m going to assume that intentions are something over and above basic beliefs and desires; that an intention is not, for instance, merely a strongest desire or, as Donald Davidson held at an early stage of his thinking, merely a belief-desire pair.
+Intention involves more than this.
 
 There is a temptation to assume that intention is involved in every case of purposive action.
-But it’s hard to see what the argument for this could be.
+But it’s hard to see what the argument for this assumption could be.
 In many cases it seems that beliefs, desires and motor representations are all that is needed to explain purposive action. 
@@
 But it’s possible for an agent to take and eat a biscuit, and to do so purposively, without having any intentions at all.  
-Beliefs desires and motor representations are sufficient.
+Beliefs, desires and motor representations are sufficient.
 
-So (again), what are intentions for?
+So if we don't need intentions merely to perform a purposive action, what are intentions for?
 
 \section{Slide}
-Motor representations enable quite sophisticated planning over short periods of time and sequences of action; for example, how you grasp a pointer will depend on what you are about to do with it \citep{zhang:2007_planning}.
+This question becomes more pressing if you consider that motor representations enable quite sophisticated planning over short periods of time and sequences of action; for example, how you grasp a pointer will depend on what you are about to do with it \citep{zhang:2007_planning}.
 
-This sort of planning does not need intentions at all.  So what are intentions for?
+This sort of planning does not need intentions at all.  So (again) what are intentions for?
 
 \section{Slide}
+Michael Bratman suggests that 
 \emph{Intentions are for planning multiple separate actions over longer periods of time; and for planning multiple separate actions whose execution is mutually constraining where the outcomes cannot be represented motorically.}
 
-This is a case where intentions are really needed --- can’t act on strongest desire (for the big reward) if want to maximise rewards by collecting the small and the large reward.
-And can’t rely on motor representation because the motor system doesn’t care about things that cannot be represented in motor terms.
+This is a case where intentions are really needed [see figure in slide] --- here one can’t act on strongest desire (for the big reward) if want to maximise rewards by collecting the small and the large reward.
+And one can’t rely on motor representation because the motor system doesn’t care about things that cannot be represented in motor terms.
+
+I don't think, of course, that intentions are only involved in actions which require planning of this sort.
+But I do think it's only in such actions that absolutely require intentions.
 
@@
 
-As far as I know, this claim is not explicitly argued for.  From conversation with philosophers I think the central argument is supposed to contrast cases.
+As far as I know, this claim is not explicitly argued for.  From conversation with philosophers I think the central argument hinges on  contrast cases.
 Contrast cases are pairs of actions which are similar in terms of behaviour and coordination but where one is joint but the other isn’t....
@@
 
-I think this argument is mistaken.
-First because there are contrast cases in which the joint action does not involve shared intention (***bouncing a cube on a trampoline---might be done by two people merely shaking the tramoline with no interest in, or awareness of, each other).
-It follows that appeal to shared intention does not provide a sufficiently general way of explaining the difference illustrated by the contrast cases.
-Second because there is a more general notion appeal to which enables us to distinguish the contrast cases.
+I think this argument is mistaken for two reasons.
+
+\section{Slide}
+First because there are contrast cases in which the joint action does not involve shared intention.
+For instance, contrast our lifting a sofa together with us each individually raising either end of the sofa, coincidentally at the same time.
+The former is a joint action whereas the latter is arguably not.
+But I don't think it's right to assume that lifting a sofa together necessarily involves shared intention (although our \emph{buying} a sofa together might).
+So it seems to me that the contrast cases give us no reason at all to suppose that all joint actions involve shared intention.
+
+In fact the existence of cases like the sofa indicate that appeal to shared intention does not provide a sufficiently general way of explaining the difference illustrated by the contrast cases.
+
+This brings me to a second consideration.
+There is a more general notion appeal to which enables us to distinguish the contrast cases.
 Genuine joint action differs from merely parallel action because the former involves each agent’s representing an outcome to which all of their actions are directed where these outcome representations coordinate their actions in a way that would normally facilitate their collective success in bring about this outcome.
@@
 So far I have been assuming that there is an inconsistency between two claims:
-(1) my claim that reciprocal reciprocal agent-neutral motor representation can ground purposive joint action
+(1) my claim that reciprocal  agent-neutral motor representation can ground purposive joint action
 (2) many philosophers’ claim that all joint actions involve shared intention.
@@
 But maybe it is a mistake to think that these are inconsistent.
-Why not suppose that some social motor representations are a variety of shared intentions?
+Why not suppose that some social motor representations are  shared intentions?
 
@@
 So insofar as labeling some social motor representations shared intentions might help to avoid unnecessarily philosophical discussion, I’m all in favour.
-However, there is an important differences between the states normally regarded as  shared intentions and any motor representations.
+However, there is an important difference between the states normally regarded as  shared intentions and any motor representations.
 And this difference matters for understanding the interface between shared intention and motor representation.
@@
 \section{Slide}
+In what respects do reciprocal agent-neutral motor representations differ from shared intentions?
+
+\section{Slide}
 As background we first need a generic distinction between content and format.
@@
 The difference between the two lines is a difference in content.
-Each of the routes could alternatively have been represented by a distinct series of instructions written on the same piece of paper; these cartographic and propositional representations differ in format.
-The format of a representation constrains its possible contents.
-For example, a representation with a cartographic format cannot represent what is represented by sentences such as `There could not be a mountain whose summit is inaccessible.'\footnote{ Note that the distinction between content and format is orthogonal to issues about representational medium.
-The maps in our illustration may be paper map or electronic maps, and the instructions may be spoken, signed or written.
-This difference is one of medium.} The distinction between content and format is necessary because, as our illustration shows, each can be varied independently of the other.
 
 \section{Slide}
+Each of the routes could alternatively have been represented by a distinct series of instructions written on the same piece of paper; these cartographic and propositional representations differ in format.
+
 Format matters because only where two representations have the same format can they be straightforwardly inferentially integrated.
 
+\section{Slide}
 To illustrate, let’s stay with representations of routes.  
@@
 Some joint actions involve both shared  intention and reciprocal agent-neutral motor representation.
-Imagine two people setting up a tend in a gale together, for example.
+Imagine two people setting up a tent in a gale together, for example.
 It seems success will often require many ingredients including shared intention for the large-scale planning and reciprocal agent-neutral motor representation for passing objects and bending the poles together.

### 2012-03-08 conf joint action --- sharing goals vs sharing intentions cambridge 2012 talk/intention and motor representation in joint action HANDOUT.tex
handout

%!TEX TS-program = xelatex
%!TEX encoding = UTF-8 Unicode

\documentclass[11pt]{extarticle}
% extarticle is like article but can handle 8pt, 9pt, 10pt, 11pt, 12pt, 14pt, 17pt, and 20pt text

\def \ititle {Intention and Motor Representation in Joint Action}
\def \isubtitle {}
\def \iauthor {Stephen A. Butterfill}
\def \iemail{s.butterfill@warwick.ac.uk}
\date{}

%for strikethrough
\usepackage[normalem]{ulem}

\input{$HOME/Documents/submissions/preamble_steve_handout}

\bibpunct{}{}{,}{s}{}{,}  %use superscript TICS style bib
%remove hanging indent for TICS style bib
%TODO doesnt work
\setlength{\bibhang}{0em}
%\setlength{\bibsep}{0.5em}

%itemize bullet should be dash
\renewcommand{\labelitemi}{$-$}

\begin{document}

\begin{multicols}{3}

\setlength\footnotesep{1em}

\bibliographystyle{newapa} %apalike

%\maketitle
%\tableofcontents

\

\begin{center}
{\Large
\textbf{Intention and Motor Representation in Joint Action}
}

%Stephen A. Butterfill

<s.butterfill@warwick.ac.uk>

\end{center}
%
%
%\textbf{Abstract}
%On the assumption that social motor representation plays a role in explaining how effective joint action is possible, do we also need motor representation to explain what joint action is?  Philosophers tend to assume that motor representation is only an enabling condition for joint action and of no direct interest to narrowly philosophical theories of joint action and shared intention.  In this talk I shall argue that social motor representation and shared intention  have distinctive roles in explaining the purposiveness of joint action.  This gives rise to a challenge.  On the one hand, effective joint action---imagine two people erecting a tent in a gale together---sometimes requires both shared intentions and social motor representations plus a certain kind of harmony between the two.  On the other hand, recognizing their distinctive roles precludes the existence of direct inferential links between shared intentions and social motor representations.  The challenge is to explain how these two kinds of representation could sometimes harmoniously contribute to effective joint action despite the lack of inferential integration.
%

\section{Introduction}
A motor representation is \emph{agent-neutral} if it concerns an action which is not one’s own or, in the case of joint action, not entirely one’s own.

Two or more motor representations are \emph{reciprocal} just if there is a single outcome which each motor representation represents.

\textbf{Premise} 
Reciprocal agent-neutral motor representation enables joint action

\begin{quote}
`the social relation between individuals modulates action simulation ...\ motor activation during action anticipation depends on the ...\ relation between the actor and the observer ...\ Simulation of another person’s action, as reﬂected in the activation of motor cortices, gets stronger the more the other is perceived as an interaction partner.’  \citep{kourtis:2010_favoritism}
\end{quote}

\textbf{Question} Does reciprocal agent-neutral motor representation also play a role in explaining what joint action is?

\textbf{Challenge}  How could social motor representation and shared intention harmoniously contribute to joint action?

\section{The possibility of purposive joint action}
What is the relation between a purposive joint action and the outcome or outcomes to which it is directed?

Reciprocal agent-neutral motor representations can
(1) involve a representation, on the part of each agent, of an outcome; 
(2) coordinate the several agents’ activities;
and 
(3) coordinate the several agents’ activities in such a that would normally facilitate the occurrence of the represented outcome.

For you and I to have a shared intention that we J it is sufficient that: `(1)(a) I intend that we J and (b) you intend that we J; (2) I intend that we J in accordance with and because of la, lb, and meshing subplans of la and lb; you intend that we J in accordance with and because of la, lb, and meshing subplans of la and lb; (3) 1 and 2 are common knowledge between us.'\citep%[View 4]
{Bratman:1993je}

\section{An objection}
‘the key property of joint action lies in its internal component [...] in the participants’ having a “collective” or “shared” intention.’ \citep[pp.\ 444-5]{alonso_shared_2009}

But could some reciprocal agent-neutral motor representations be shared intentions?  No ...
\begin{enumerate}
\item Only representations with a common format can be inferentially integrated.

\item Any two intentions can be inferentially integrated in practical reasoning.

\item My intention that I visit Paris on Friday is a propositional attitude.

\item All intentions are propositional attitudes.

\item No motor representations are propositional attitudes.

\item No motor representations are intentions.
\end{enumerate}

\section{The Interface Problem}
Two  outcomes, A and B, \emph{match} in a particular context just if, in that context, either the occurrence of A would normally constitute or cause, at least partially, the occurrence of B or vice versa.

Some joint actions involve both shared  intention and reciprocal agent-neutral motor representation.

How are non-accidental matches between the outcomes specified by shared intentions and by reciprocal agent-neutral motor representations possible?

`motor imagery could play a crucial role in bridging the gap'\citep{pacherie:2000_content}

\footnotesize 
\bibliography{$HOME/endnote/phd_biblio}

\end{multicols}

\end{document}

### 2012-03-08 conf joint action --- sharing goals vs sharing intentions cambridge 2012 talk/intention and motor representation in joint action.tex
handout

@@
 \section{Slide}
-A motor representation is \emph{agent-neutral} if it concerns an action which is not one’s one or, in the case of joint action, not entirely one’s own.
+A motor representation is \emph{agent-neutral} if it concerns an action which is not one’s own or, in the case of joint action, not entirely one’s own.
 

### 2012-03-11 intention motor representation with corrado sinigaglia/intention and motor action.tex
cambridge added

@@
 Unless such intentions and motor representations are to pull an agent in incompatible directions, which would typically impair action execution, there are requirements concerning how the outcomes they represent must be related to each other.
-The problem is to explain how any such requirements could be met non-accidentally; we call this `the interface problem' .
+The problem is to explain how any such requirements could be met non-accidentally; we call this `the interface problem'.
 The key to solving this problem, we shall suggest, is to recognise that intentions can have constituents which refer to outcomes by deferring to motor representations of those outcomes.
@@
 In imagining executing actions such as grasping a cup, there are bio-mechanical, dynamical and postural constraints, among others, on what can be imagined.
-These constraints are closely related to those which govern actually performing such actions \citep{Jeannerod:2001yb} .
+These constraints are closely related to those which govern actually performing such actions \citep{Jeannerod:2001yb}.
 Such constraints are plausibly explained by the representational format of motor representations.

### 2012-03-11 intention motor representation joint action (paper 3)/intention and motor representation in joint action.tex
just started first draft of paper version, not much done

--- 2012-03-08 conf joint action --- sharing goals vs sharing intentions cambridge 2012 talk/intention and motor representation in joint action.tex
@@
 \begin{abstract}
-On the assumption that social motor representation plays a role in explaining how effective joint action is possible, do we also need motor representation to explain what joint action is?
+\noindent
+On the assumption that social motor representation plays a role in explaining how effective joint action is possible, is there also a role for motor representation in explaining what joint action is?
 Philosophers tend to assume that motor representation is only an enabling condition for joint action and of no direct interest to narrowly philosophical theories of joint action and shared intention.
@@
 
-\section{Slide}
-[hello]
+\section{Introduction}
+Which events are joint actions?
+We shall argue that fully answering this question involves appeal to structures of motor representation.
+By contrast, 
+most philosophers and many (but not all) psychologists appear to assume that this question can be fully answered just in terms of a notion of shared (or `collective') intention.
+For an event to be a joint action is for it to be appropriately related to a shared intention.
+For comparison, one might also hold that for an event to be an (ordinary, individual) action is for it to be appropriately related to an intention.
+This assumption immediately raises the question of what shared intention is.
+On this question several philosophers offer competing theories.
+It is striking, however, that few have challenged the assumption that for an event to be a joint action it must be appropriately related to a shared intention.
+This is striking because the parallel assumption about ordinary, individual action is not as widely held among leading theorists.
+On several leading views, there could be agents whose actions are purposive---and, in some cases, even intentional---even though the agents have no intentions.
+If any such view is correct, it is impossible to fully explain what it is for an event to be an action by appeal to intention alone.
+Our aim is to defend a parallel claim about joint action.
+More specifically, we shall defend these claims: 
+first, 
+that fully explaining which events are joint actions requires appeal to structures of motor representation and not only to shared intention;
+and, 
+second, 
+that recognising a constitutive role for motor representation in joint action makes possible an account of shared intention which is simpler and more elegant than any of the leading alternatives.
 
-\section{Slide}
-As a philosopher I’m going to start from an empirically controversial premise.
-I don’t think this premise has been established, but I do think it’s a reasonable bet.
-
-\section{Slide}
+\section{Social motor representation}
+We start from the premise that reciprocal agent-neutral motor representation sometimes enables joint action.
+This needs unpacking.
 A motor representation is \emph{agent-neutral} if it concerns an action which is not one’s own or, in the case of joint action, not entirely one’s own.
-
-\section{Slide}
 Two or more motor representations are \emph{reciprocal} just if there is a single outcome which each motor representation represents.
-
-\section{Slide}
-Sometimes I’ll using the term \emph{social} motor representation for reciprocal agent-neutral motor representations.
-I have a slightly bad conscience about this because I’m not sure they’re actually social in any ordinary sense.
-
-It is almost uncontroversial that agent-neutral motor representations exist, and I see no reason to doubt that they could be reciprocal.
+We shall use the term \emph{social} motor representation as an abbreviation for reciprocal agent-neutral motor representation.
+(Of course we cannot assume in advance of argument that reciprocal agent-neutral motor representations are social in any interesting sense;
+but we are using the term `social' in a non-standard way as an abbreviation, one that reflects our aim.)
+It is almost uncontroversial that social motor representations exist.
 The controversial part of the premise is the claim that such representations can enable joint action.
 
-I should say, by the way, that I’m not suggesting that reciprocal agent-neutral motor representation is involved in every joint action.
-Perhaps some joint actions do not involve social motor representation.
-The premise is just that social motor representation is among the factors which enable some joint actions.
+Our premise should not be interpreted as implying that social motor representation is involved in every joint action,
+nor that social motor representation is sufficient for joint action.
+The premise is just that social motor representation is among the factors enabling some joint actions.
 
-\section{Slide}
-I’m not going to defend the premise, although I will try to explain the premise in more detail in a moment.  But first let me outline where I’m going with this.
+Why accept the premise that social motor representation enables joint action?
+A direct way to test the premise would be
+measure how (if at all) selective interventions on social motor representation affected agents’ performance of joint actions.
+As far as we know no one has done this and it is probably impossible.
+But there are some indirect findings.
+We shall focus on just one, a finding by Kourtis and colleagues that `Simulation of another person’s action, as reﬂected in the activation of motor cortices, gets stronger the more the other is perceived as an interaction partner’  \citep[p.\ 4]{kourtis:2010_favoritism}.
+This finding falls well short of providing conclusively support for the premise; we mention it as an illustration of the sort of evidence is relevant to the premise.
 
-\section{Slide}
-The premise leads to a question and, relatedly, to a challenge.
-
-\section{Slide}
-The question is whether social motor representation is not only an enabling condition for joint action but also plays a role in explaining what joint action is.
-I want to argue that it does.
-In fact I want to argue that it plays a role similar and complementary to that of shared intention.
-
-\section{Slide}
-After this, in the last part of my talk, I also want to point to a challenge raised by the existence of social motor representation.  
-The challenge is roughly this: planning for joint action involves representations of target outcomes in two different formats, a motor format and a propositional format.  
-Given that representations in these different formats cannot straightforwardly be inferentially integrated in practical reasoning, what could ever insure that there is sometimes non-accidental harmony between motor representations and shared intentions?
-
-\section{Slide}
-Let me start by explaining the premise: Reciprocal agent-neutral motor representation enables joint action.
-Why think this?
-A direct way to test the premise would be to selectively intervene on social motor representation and see how it affects agents’ performance of joint actions.
-As far as I know no one has done this.
-But there are some indirect findings.
-I shall mention just one, a finding by Dimitris Kourtis and colleagues.
-
-\section{Slide}
-They found that `the social relation between individuals modulates action simulation' \citep[p.\ 1]{kourtis:2010_favoritism}.
-
-\section{Slide}
-In slightly more detail:
-\begin{quote}
-`motor activation during action anticipation depends on the ...
-relation between the actor and the observer ...
-Simulation of another person’s action, as reﬂected in the activation of motor cortices, gets stronger the more the other is perceived as an interaction partner.’  \citep[p.\ 4]{kourtis:2010_favoritism}
-\end{quote}
-I realise this doesn’t conclusively support the premise.  But I think it is evidence in its favour.
-
-\section{Slide}
 Here’s the experiment they did.
@@
 Also present was the ‘loner’ who always acted alone.
-
-\section{Slide}
 EEG measurements of motor activation were recorded.
-
-\section{Slide}
-[image shrinks]
-
-\section{Slide}
 The researchers found that patterns of activation for self and joint action partner were similar, and different from patterns of activation for the actions of the loner which were similar to patterns in a ‘no go’ condition where no one was to act.
-
-\section{Slide}
-So here’s what I take from this paper.
+Since the researchers did not include a condition where there was a competitive, non-joint interaction, we cannot conclude that the modulation in action simulation they found is specific to joint action as opposed to other kinds of interaction.
+What we do take from this paper,
+together with a body of other research on motor simulation of action,
+is this.
 If you are engaged in a joint action with someone, one which involves moving an object by passing it between you, then each of you has motor representations of the other’s actions and these motor representations are functionally equivalent to motor representations of your own actions in the sense that they are just the sorts of representation that might have caused you to do what the other is doing (if you were in her position).
 
-\section{Slide}
+\section{How could social motor representation enable joint action?}
 Suppose that Kourtis et al are right that social motor representation is more likely to occur in joint action than when one is merely observing.
-How could this facilitate joint action?
+It doesn't follow, of course, that social motor representation enables some joint actions (and nor do these researchers claim that it does).
+But let us consider just the possibility that it might.
+How could social motor representation enable joint action even in principle?
 
-\section{Slide}
-Let’s step back and consider an individual action.
-An agent moves a mug from one place to another, passing in from her left hand to her right hand half way [*demonstrate].
-It’s a familiar idea that motor representations for planning and monitoring action involve an hierarchical structure,
+To answer this question
+let us take a step back and consider an individual action.
+Suppose an agent moves a mug from one place to another, passing in from her left hand to her right hand half way.
+It is a familiar idea that motor representations for planning and monitoring action involve an hierarchical structure,
 where there is a relatively abstract representation of an outcome that is progressively filled in.
@@
 But they are similar in the respects that matter for explaining the purposiveness of action.
-(1) Like intentions, some motor representations represent outcomes (and not merely patters of joint displacements, say).
+(1) Like intentions, some motor representations represent outcomes (and not merely patterns of joint displacements, say).
 (2) Like intentions, some motor representations play a role in coordinating multiple  component activities by virtue of their role as elements in hierarchically structured plans.

### 2012-03-15 intention motor representation joint action (paper 3)/intention and motor representation in joint action.tex
a

@@
 By contrast, 
-most philosophers and many (but not all) psychologists appear to assume that this question can be fully answered just in terms of a notion of shared (or `collective') intention.
-For an event to be a joint action is for it to be appropriately related to a shared intention.
-For comparison, one might also hold that for an event to be an (ordinary, individual) action is for it to be appropriately related to an intention.
+most philosophers and many (but not all) psychologists appear to assume that this question can be fully answered just in terms of a special kind of intention or structure of intention and knowledge often called a shared (or `collective') intention.
+On this widely held view, for an event to be a joint action is for it to be appropriately related to a shared intention.
+For comparison, in the case of ordinary, individual action 
+one might hold that for an event to be an action is for it to be appropriately related to an intention.
 This assumption immediately raises the question of what shared intention is.
@@
 \section{Social motor representation}
-We start from the premise that reciprocal agent-neutral motor representation sometimes enables joint action.
+We start from the premise that reciprocal agent-neutral motor representation enables some joint actions.
 This needs unpacking.
-A motor representation is \emph{agent-neutral} if it concerns an action which is not one’s own or, in the case of joint action, not entirely one’s own.
-Two or more motor representations are \emph{reciprocal} just if there is a single outcome which each motor representation represents.
+A representation (motor or not) is \emph{agent-neutral} if its content does not specify any agent or agents.
+To illustrate, agent neutral representations are sometimes found at the early stages of planning.
+Imagine that you and some friends are tasked with preparing a holiday.  
+You might first write down a plan of action without specifying who will act; the plan simply describes what is to be done.
+The plan  will eventually be implemented by you and your friends but the plan itself does not specify this and so it is agent-neutral.
+Of course the fact that this plan is your collective plan may be represented elsewhere; this fact may also be implicit in  the plan's being stapled to the door of your communal kitchen.
+The agent-neutrality of a representation does not require that the agents are nowhere specified, only that they are not specified in the content of the representation.\footnote{
+Strictly the following argument does not hinge on the agent-neutrality of representations.
+It is sufficient for our purposes that there are reciprocal motor representations which concern actions which are partly but not entirely the agent's own.
+In principle such representations could have contents which specify other agents or multiple agents.
+We focus on agent-neutral representations to simplify exposition.
+}
+Two or more agents have \emph{reciprocal} motor representations  just if there is a single outcome and each agent has a motor representation of that outcome.
 We shall use the term \emph{social} motor representation as an abbreviation for reciprocal agent-neutral motor representation.
@@
 but we are using the term `social' in a non-standard way as an abbreviation, one that reflects our aim.)
-It is almost uncontroversial that social motor representations exist.
-The controversial part of the premise is the claim that such representations can enable joint action.
+So for two or more agents to have social motor representation is just for there to be a single outcome of which each agent has an agent-neutral motor representation.
 
-Our premise should not be interpreted as implying that social motor representation is involved in every joint action,
-nor that social motor representation is sufficient for joint action.
-The premise is just that social motor representation is among the factors enabling some joint actions.
+Our premise is just that social motor representation is among the factors enabling some joint actions.
+It does not follow that social motor representation is involved in every joint action,
+nor that the existence of social motor representation would be sufficient for there to be joint action.
+
+This premise is doubly controversial.
+While it is now well established both that there are motor representations of action outcomes (and not merely patterns of joint displacement, say), and also that there are motor representations of the outcomes of others' actions rather than only of one's own actions,
+less research has so far considered the conjecture that motor representations concern outcomes of actions which are partly but not entirely one's own.
+There is also relatively little evidence for the further conjecture that such representations enable some joint actions.
+We shall not argue for either conjecture; they are empirical premises.  
+What follows is speculative philosophy: an attempt to explain the consequences and significance of the conjecture that social motor representation enables some joint actions.
 
 Why accept the premise that social motor representation enables joint action?
-A direct way to test the premise would be
+A direct way to test the premise would be to 
 measure how (if at all) selective interventions on social motor representation affected agents’ performance of joint actions.
-As far as we know no one has done this and it is probably impossible.
+As far as we know no one has done this and it may be impractical.
 But there are some indirect findings.
-We shall focus on just one, a finding by Kourtis and colleagues that `Simulation of another person’s action, as reﬂected in the activation of motor cortices, gets stronger the more the other is perceived as an interaction partner’  \citep[p.\ 4]{kourtis:2010_favoritism}.
-This finding falls well short of providing conclusively support for the premise; we mention it as an illustration of the sort of evidence is relevant to the premise.
+We shall focus on just one, which will serve to illustrate the premise (but does not provide a decisive reason to accept it).
+Kourtis and colleagues report that `[s]imulation of another person’s action, as reflected in the activation of motor cortices, gets stronger the more the other is perceived as an interaction partner’  \citep[p.\ 4]{kourtis:2010_favoritism}.
+To establish this they had two subjects sit opposite each other at a table; there was also a confederate at the table, the `loner'.  
+Sometimes the subjects acted alone, picking up and replacing an object.  And sometimes they had to act together, passing an object between them.  
+The loner always acted alone.
+EEG measurements of motor activation were recorded.
+The researchers compared patterns of motor activation in four conditions: when the first subject acted, when the second subject acted, when the loner acted and when there was a `no go' signal so that no one acted.
+They found that when one of the subjects acted, patterns of activation in each of the two subjects were similar regardless of which was acting and which was observing.
+By contrast, when the loner acted the subjects' patterns of activation resembled those found in the `no go' condition when there was no action.
+Since the researchers did not include a condition where there was a competitive, non-joint interaction, we cannot conclude that the effect they found is specific to joint action as opposed to other kinds of interaction.
+What we do take this paper to indicate, 
+together with a body of other research on motor representation of observed action,
+is this.
+If you are engaged in a joint action with someone, one which involves moving an object by passing it between you, then each of you has motor representations of actions whose (perhaps improper) components include actions which will be performed by the other  and these motor representations are functionally equivalent to motor representations of your own actions in the sense that they are just the sorts of representation that might have caused you to do what the other is doing (if you were in her position).
 
-Here’s the experiment they did.
-The task was simple: two people sat opposite each other.
-Sometimes they acted alone, picking up and replacing an object.
-And sometimes had to act together, passing an object between them.
-Also present was the ‘loner’ who always acted alone.
-EEG measurements of motor activation were recorded.
-The researchers found that patterns of activation for self and joint action partner were similar, and different from patterns of activation for the actions of the loner which were similar to patterns in a ‘no go’ condition where no one was to act.
-Since the researchers did not include a condition where there was a competitive, non-joint interaction, we cannot conclude that the modulation in action simulation they found is specific to joint action as opposed to other kinds of interaction.
-What we do take from this paper,
-together with a body of other research on motor simulation of action,
-is this.
-If you are engaged in a joint action with someone, one which involves moving an object by passing it between you, then each of you has motor representations of the other’s actions and these motor representations are functionally equivalent to motor representations of your own actions in the sense that they are just the sorts of representation that might have caused you to do what the other is doing (if you were in her position).
+For each of the agents involved in a joint action to represent motorically actions whose (perhaps improper) components include the others' action is not yet for there to be social motor representation.
+For there to be social motor representation there must be an outcome (such as the movement of an object from one place to another) which each agent has an agent-neutral motor representation of.
+The experiment by \citet{kourtis:2010_favoritism} was not designed to distinguish
+a case in which each agent represents the outcome to which the  whole joint action is directed (in this case, the entire movement of the object)
+from
+a case in which 
+each agent represents separately outcomes to which each of the individual actions comprising the joint action (in this case, the several movements of the object each of which just one agent is responsible for).
+However we are making a leap and assuming (without argument) that in cases like that studied by Kourtis and colleagues, there is sometimes social motor representation.
 
 \section{How could social motor representation enable joint action?}
-Suppose that Kourtis et al are right that social motor representation is more likely to occur in joint action than when one is merely observing.
-It doesn't follow, of course, that social motor representation enables some joint actions (and nor do these researchers claim that it does).
+Suppose that social motor representation is present in some joint action contexts, and, inspired by \citet{kourtis:2010_favoritism}, that 
+social motor representations are more likely to occur in joint action than when one is merely observing.
+It doesn't follow, of course, that social motor representation enables some joint actions.
 But let us consider just the possibility that it might.
@@
 Suppose an agent moves a mug from one place to another, passing in from her left hand to her right hand half way.
-It is a familiar idea that motor representations for planning and monitoring action involve an hierarchical structure,
-where there is a relatively abstract representation of an outcome that is progressively filled in.
-
-\section{Slide}
-I don’t suppose that this attempt to depict the hierarchy of motor representations involved in moving a mug with two hands is accurate.
-I’m only trying to illustrate two familiar ideas.
-One is that motor planning involves starting with relatively abstract representations of outcomes and filling in details.
-
-\section{Slide}
-The other is that there is a need, even for a single agent, to synchronize the exchange between the two hands.
+It is a familiar idea that motor planning, like planning generally, involves starting with relatively abstract representations of outcomes and gradually filling in details.
+We can capture this by supposing that 
+motor representations for planning and monitoring action involve a hierarchical structure of representations.
+At the top we might find a relatively abstract representation of an outcome, in this case of the movement of the object from one location to another.
+Action-relevant details are progressively filled in by representations at lower stages of the hierarchy. 
+Now in the action we are considering there is a need, even for the single agent, to coordinate the exchange between the two hands.
+How is this achieved? 
+We suppose that part of the answer involves the fact that planning for the movements of each hand is not done entirely independently.
+Rather there is a plan for the whole action
+and plans for the movements of each hand are components of this larger plan.
+It is in part because they are parts of a larger plan that the plan for one hand constrains and is constrained by the plan for the other hand.
 
 How is this relevant to the case of joint action?
+In joint action the agents have the same goal, to move the object from one place to another.
+They also face a similar coordination problem, requiring a precisely timed swap from one hand to another.
+Now suppose (inspired by Koutis et al’s and others' findings) that the same planning is involved in the individual case (where one agent performs the whole action) and in the joint action case (where the action is distributed between two agents).
+The planning is the same almost up to the actual muscle contractions.
+How could this be helpful?
+Suppose the agents' planning processes are similar enough that, for a given context and problem, they will produce approximately the same plans.
+Then having each agent plan the whole joint action means that (i) each agent plans the other agent's action,
+(ii) each agent's plan for the other agent's action is approximately the same as that agent's plan for her own action,
+and 
+(iii) each  agent's plans for their own action are constrained by their plans for the other agent's action.
+So what enables the two agents' plans to mesh is not that they represent each other's plans but more simply that they plan each other's actions as well as their own actions as if they were each about to do the whole thing themselves.
+Each agent is planning (and monitoring) both their actions almost as if a single agent were going to execute the whole action.
+And of course this is exactly what we want for small-scale joint action---we want two or more agents to act as one.
+This may be why the performance of dyads in joint actions often resembles the performance of individuals tasked with performing the whole action alone (*refs: Knoblich circles following task).
 
-\section{Slide}
-In joint action the agents have the same goal, to move the object from there to here.
+So what is the difference between the individual and the joint case?
+From the point of view of motor representation, the primary difference may be that in joint action there is a need to prevent execution of the parts of the action which are not one’s own.
 
-\section{Slide}
-They also face a similar coordination problem, requiring a precisely timed swap
+%Here then is the basic idea I take to be guiding Kourtis and others.
+%Coordination is sometimes achieved by having each agent’s motor system plan all of their actions; 
+%given some assumptions, this could be a way of making it likely that each will execute their part in the joint action in a way that meshes with the way the other agents execute their parts.
 
-\section{Slide}
-And Koutis et al’s findings (and others’ findings) suggest that the same planning in involved in the joint action case, almost up to the actual muscle contractions.
-That is, in the joint action situation each agent plans both agents’ actions as if they were the actions of a single agent.
-This may be what enables them to coordinate so well : each is able to plan her own actions in a way that meshes with the other agent’s actions because each agent is planning (and monitoring) both their actions almost as if a single agent were going to execute the whole action.
-And of course this is exactly what we want for small-scale joint action---we want two or more agents to act as one.
-
-\section{Slide}
-So what is the difference between the individual and the joint case?
-From the point of view of motor representation, the primary difference may be that in joint action there is a need to inhibit execution of the parts of the action which are not one’s own.
-
-Here then is the basic idea I take to be guiding Kourtis and others.
-The idea is that coordination is sometimes achieved by having each agent’s motor system plan all of their actions; 
-given some assumptions, this could be a way of making it likely that each will execute their part in the joint action in a way that meshes with the way the other agents execute their parts.
-
-Now so far I’ve only been considering a possible role for social motor representation in enabling joint action.  
+\section{Grounding the purposiveness of joint action}
+So far we have only been considering a possible role for social motor representation in enabling joint action.  
 But I think the details already give us grounds for holding that motor representation has a role to play in explaining what joint action is.

### 2012-03-20 intention motor representation with corrado sinigaglia/conf spp submit/intention and motor representation SPP submit.tex
before lyon

--- 2012-02-23 intention motor representation with corrado sinigaglia/joint action version/intention and motor representation joint action version.tex
@@
 %a4paper or letterpaper (also used in preamble_steve_paper2
-\def \papersize {a4paper}
+\def \papersize {letterpaper}
 
@@
 
-\def \ititle {Shared Intention and Social Aspects of Motor Representation, or How to Avoid Exoticism}
-\def \isubtitle {}
+\def \ititle {Intention and motor representation:}
+\def \isubtitle {how to resist temptation}
 \def \iauthor {
-S. Butterfill* \& C. Sinigaglia**
-\\ 
-**Department of Philosophy, University of Warwick
-\\ 
-***Dipartimento di Filosofia, Università degli Studi di Milano
+%S. Butterfill* \& C. Sinigaglia**
+%\\ 
+%**Department of Philosophy, University of Warwick
+%\\ 
+%***Dipartimento di Filosofia, Università degli Studi di Milano
 }
-\def \iemail{s.butterfill@warwick.ac.uk}
+\def \iemail{
+%	s.butterfill@warwick.ac.uk
+}
+%\date{}
 
 \input{$HOME/Documents/submissions/preamble_steve_paper2}
-%\author{}
-%\date{}
+\author{}
+\date{}
 
@@
 \noindent
-Are there distinctive roles for shared intention and social motor representation in explaining the purposiveness of joint action?
-This question needs some background.
-A joint action is a purposive action, or something resembling one, comprising two or more agents' actions.
-What is the relation between a joint action and the outcome (or outcomes) to which it is directed?
-Standard answers to this question appeal to a notion of shared intention and ignore social motor representation.
-It is tempting (for some philosophers, at least) to suppose that social motor representations are only enabling conditions for joint action, or else to suppose that social motor representation is merely a variety of shared intention.
-The talk will provide reasons for rejecting both suppositions.
-The truth may be that shared intention and social motor representation have distinctive roles in explaining the purposiveness of joint action.
-This gives rise to a challenge.
-On one hand, effective joint action---imagine erecting a tent in a gale together---sometimes requires both shared intentions and social motor representations plus a certain kind of harmony between the two.
-On the other hand, recognizing their distinctive roles precludes the existence of direct inferential links between shared intentions and social motor representations.
-The challenge is to explain how these two kinds of representation could sometimes harmoniously contribute to effective joint action despite the lack of inferential integration.
+Are there distinctive roles for intention and motor representation in explaining the purposiveness of action?
+Standard philosophical accounts of action assign a role to intention but are silent on motor representation.
+The temptation is to suppose that nothing need be said here because motor representation is either only an enabling condition for purposive action or else merely a variety of intention.
+Drawing on cognitive and neuroscientific research, this talk identifies reasons for resisting temptation.
+Some motor representations, like intentions, coordinate actions in virtue of representing outcomes; but, unlike intentions, motor representations cannot feature as premises or conclusions in practical reasoning.
+This implies that motor representation has a distinctive role in explaining the purposiveness of action.
+It also gives rise to a problem: were the roles of intention and motor representation entirely independent, this would impair effective action.
+Resisting temptation therefore requires explaining how intentions interlock with motor representations.
+The solution, we argue, is to recognise that the contents of intention can be partially determined by the contents of motor representations.
+Understanding this content-determining relation enables better understanding how intention is related to action.
+
+\
+
+\noindent
+Word count: 3731 including notes, excluding references
 
@@
 
-A purposive joint action is an event involving two or more agents' actions where the actions are collectively directed to some outcome.
-What is the relation between a joint action and the outcome or outcomes to which it is directed? 
-The standard way of answering this question appeals to shared intention,
-a structure of states which stands to joint action roughly as an ordinary, individual intention stands to an ordinary, individual action.  
-On the standard view, a joint action is directed to an outcome in virtue of the joint action's being appropriately related to a shared intention which represents this outcome or some related outcome. 
-As this view is usually expounded, the relation between joint actions and outcomes to which they are directed is treated as largely independent of the motor processes and representations underpinning action execution.
+What is the relation between a purposive action and the outcome or outcomes to which it is directed?
+The standard way of answering this question appeals to intention, a propositional attitude which plays a characteristic role in planning and coordinating action, is linked to practical reasoning and is subject to characteristic norms (\citealp{Bratman:1987xw}).
+On the standard view, an action is directed to an outcome in virtue of the action's being appropriately related to an intention which represents this outcome or some related outcome.
+As this view is usually expounded, the relation between actions and outcomes to which they are directed is treated as largely independent of the motor processes and representations underpinning action execution.
 Motor representations are usually considered either as philosophically irrelevant enabling conditions, or else as merely filling in additional details of the basic schema provided by the standard story.
 Our aim is to show that this is a mistake:
-the standard view must be  refined and extended  to accommodate distinctive roles for intention and motor representation in explaining the purposiveness of joint action.
+the standard view must be  refined and extended  to accommodate distinctive roles for intention and motor representation in explaining the purposiveness of action.
 
@@
 are partially dependent on action outcomes and not only on  narrowly kinematic or dynamic features of action,
-it would be advantageous for some  motor representations to represent action outcomes. 
+it would be advantageous for some  motor representations to represent action outcomes . 
 Given that, as we saw, markers of motor processes carry information about action outcomes, it is now reasonable to conclude that some motor representations are representations of action outcomes.

### 2012-03-20 intention motor representation with corrado sinigaglia/conf spp submit/intention and motor representation HANDOUT.tex
lyon with handout

%!TEX TS-program = xelatex
%!TEX encoding = UTF-8 Unicode

\documentclass[11pt]{extarticle}
% extarticle is like article but can handle 8pt, 9pt, 10pt, 11pt, 12pt, 14pt, 17pt, and 20pt text

\def \ititle {Intention and Motor Representation in Action Explanation}
\def \isubtitle {}
\def \iauthor {Stephen A. Butterfill}
\def \iemail{s.butterfill@warwick.ac.uk}
\date{}

%for strikethrough
\usepackage[normalem]{ulem}

\input{$HOME/Documents/submissions/preamble_steve_handout}

\bibpunct{}{}{,}{s}{}{,}  %use superscript TICS style bib
%remove hanging indent for TICS style bib
%TODO doesnt work
\setlength{\bibhang}{0em}
%\setlength{\bibsep}{0.5em}

%itemize bullet should be dash
\renewcommand{\labelitemi}{$-$}

\begin{document}

\begin{multicols}{3}

\setlength\footnotesep{1em}

\bibliographystyle{newapa} %apalike

%\maketitle
%\tableofcontents

\

\begin{center}
{\Large
\textbf{Intention and Motor Representation in Action Explanation}
}

Stephen A. Butterfill \& Corrado Sinigaglia

<s.butterfill@warwick.ac.uk>

\end{center}

%\textbf{Abstract}
%Are there distinctive roles for intention and motor representation in explaining the purposiveness of action? Standard accounts of action assign a role to intention but are silent on motor representation. The temptation is to suppose that nothing need be said here because motor representation is either only an enabling condition for purposive action or else merely a variety of intention. This paper provides reasons for resisting temptation. Some motor representations, like intentions, coordinate actions in virtue of representing outcomes; but, unlike intentions, motor representations cannot feature as premises or conclusions in practical reasoning.  This implies that motor representation has a distinctive role in explaining the purposiveness of action. It also gives rise to a problem: were the roles of intention and motor representation entirely independent, this would impair effective action. Resisting temptation therefore requires explaining how intentions interlock with motor representations. The solution, we argue, is to recognise that the contents of intention can be partially determined by the contents of motor representations. Understanding this content-determining relation enables better understanding how intentions are related to actions.

\section{The Possibility of Purposive Action}
What is the relation between a purposive joint action and the outcome or outcomes to which it is directed?

\section{In outline}
[A] Like intentions, motor representations (i) represent outcomes, (ii) coordinate actions and (iii) do so in ways that would normally facilitate the occurrence of the represented outcomes.

[B] Unlike intentions, motor representations do not have a propositional format.

[C] Motor representation has a distinctive role in explaining the purposiveness of action (from [A] and [B]).

[D] There is a problem about how intention and motor representation can be harmoniously integrated (`The Interface Problem').

[E] The solution to this problem involves constituents of intentions which refer to actions by deferring to motor representations.

[F] Maybe where an intention properly and reliably produces bodily movement, either acting on that intention involves a further intention or else the intention involves concepts which refer to actions by deferring to motor representations.

[G] Intentionally acting in the world may ultimately depend on motor experience.

\section{Motor representation can ground purposive action (§2.A)}

Motor representations carry information about outcomes.

Action planning depends on information about outcomes.

Therefore, motor representations represent outcomes.

\section{Motor representations aren't intentions (§2.B)}
\begin{enumerate}
\item Only representations with a common format can be inferentially integrated.

\item Any two intentions can be inferentially integrated in practical reasoning.

\item My intention that I visit Paris on Friday is a propositional attitude.

\item All intentions are propositional attitudes (from 1--3).

\item No motor representations are propositional attitudes.

\item No motor representations are intentions (from 4, 5).
\end{enumerate}

\section{The Interface Problem (§2.D)}
Two  outcomes, A and B, \emph{match} in a particular context just if, in that context, either the occurrence of A would normally constitute or cause, at least partially, the occurrence of B or vice versa.

Two representations of outcomes are \emph{in harmony} in a particular context if the outcomes they represent match in that context.

Some actions involve both  intention and motor representation.

In some cases, an intention and a motor representation are non-accidentally in harmony.

How is non-accidental harmony ever possible?

A natural way to answer this question would be by appeal to a process of planning or practical reasoning.
But intention and motor representation cannot be inferentially integrated (because they differ in format).

\section{Demonstrative action concepts (§2.E)}
There are demonstrative concepts of actions.

A demonstrative concept can refer to an action by deferring to a pantomime of an action.

A demonstrative concept can refer to an action by deferring to a \emph{mental} pantomime of an action.

Experiences involved in actually acting, like those involved in mentally pantomiming an action, can ground the possibility of demonstrative reference to action by deference to motor representation.

Some concepts are constituents of intentions and refer to actions by deferring to motor representations.

\end{multicols}

\end{document}

### 2012-03-21 intention motor representation with corrado sinigaglia/intention and motor action.ppr.r1.tex
intention motor representation r1 before check

--- 2012-03-11 intention motor representation with corrado sinigaglia/intention and motor action.tex
@@
 
-\def \ititle {Intention and motor representation:}
-\def \isubtitle {how to resist temptation}
+\def \ititle {Intention and motor representation in explaining action}
+\def \isubtitle {}
 \def \iauthor {S. Butterfill* \& C. Sinigaglia**
@@
 \section{A motor format for representation}
-
 The second temptation is to suppose that motor representation, or some species of it, is a variety of intention, where intention is understood in the standard way as a propositional attitude with a characteristic role in practical reasoning \citep{Bratman:1987xw}.
@@
 The format of a representation constrains its possible contents.
-For example, a representation with a cartographic format cannot represent what is represented by sentences such as `There could not be a mountain whose summit is inaccessible.'\footnote{ Note that the distinction between content and format is orthogonal to issues about representational medium.
+For example, a representation with a cartographic format cannot represent what is represented by sentences such as `There could not be a mountain whose summit is inaccessible.'\footnote{Note that the distinction between content and format is orthogonal to issues about representational medium.
 The maps in our illustration may be paper map or electronic maps, and the instructions may be spoken, signed or written.
@@
 That motor representations differ in format from intentions shows that they are genuinely distinct phenomena.\footnote{ Readers already convinced that motor representation differs from intention in being non-conceptual will not need the following argument in order to conclude that they are distinct phenomena.
-However the following considerations also show that motor and perceptual representations differ in format, which will be relevant in later when we consider how motor representations and intentions jointly lead to action.}
+However the following considerations also indicate that motor and perceptual representations differ in format, which will be relevant later when we consider how motor representations and intentions jointly lead to action.}
 
-How can we distinguish representational formats?
-Because a representational format is typically associated with characteristic processing profiles, it is sometimes possible to infer similarities and differences in representational format from similarities and differences in the processes in which representations feature.
-This works both for artifactual and mental representations.
+How in general can we identify or distinguish representational formats?
+Because a representational format is typically associated with characteristic performance profiles, it is sometimes possible to infer similarities and differences in representational format from similarities and differences in the processes in which representations feature.
+This works both for artefactual and mental representations.
 To illustrate in terms of our earlier example, suppose that two people have representations of the same route but for one person the route is represented by a line on a map (so in a cartographic format) whereas the other person has a propositional representation of the route.
@@
 Cognitive neuroscience frequently depends on inferences of just this type.
-Contrast imagining seeing an object moving with imagining moving an object.
-There is evidence that imagining seeing and imagining acting involve different processes.
-For instance, each can be selectively impaired \citep{sirigu:2011_motor}; and factors such as limb amputation or hand posture can selectively interfere with imagining performing an action \citep{nico:2004_left,vargas:2004s_influence,fourkas:2006_influence}.
-Importantly for our purposes, imagining seeing an object move and imagining moving an object also differ in ways that indicate differences in representational format.
-For instance, the way imagining acting unfolds in time is quite different from the way imagining seeing unfolds: how long the former takes is closely related to the time it would take to perform the action, but this is not true of the latter \citep{decety:1989_timing,decety:1996_imagined,Jeannerod:1994oz}.
-In addition, for object-related actions such as grasping the handle of a cup, manipulating the object in ways that make the action harder (such as orienting the handle to make it less convenient for you to grasp) make a corresponding difference to the effort involved in imagining performing the action \citep{parsons:1994_temporal,frak:2001_orientation}, whereas such manipulations would make no systematic difference to the effort involving in imaging seeing.
+To illustrate, compare imagining seeing an object moving and with actually seeing it move.
+For this comparison we need to distinguish two ways of imagining seeing.
+There is a way of imagining seeing which phenomenologically is something like seeing except that it does not necessarily involve being receptive to stimuli.
+This way of imagining seeing, sometimes called 'sensory imagining', is commonly distinguished from cognitive ways of imagining seeing which might for example involve thinking about seeing \citep[§2.1]{Gendler:2011_imagination}.%
+\footnote{Note that we define sensory imagining in terms of phenomenology and make no stipulations about the processes and representations involved.
+This is essential for our purposes, since we wish to consider how evidence bears on a conjecture about the format of representations involved in sensory imagining.}
+It is sensorily imagining seeing an object move that we wish to compare with actually seeing an object move.
+These two have similarities in characteristic performance profile.
+For instance, whether an object can be seen all at once depends on its size and distance from the perceiver; strikingly, when subjects imagine seeing an object, whether they can imagine seeing it all at once depends in the same way on size and distance (\citealp{kosslyn:1978_measuring}; \citealp[p.\ 99ff]{kosslyn:1994_image}.
+Also, how long it takes to imagine looking over an object depends on the object's subjective size in the same way that how long it would take to actually look over that object would depend on its subjective size \citep{kosslyn:1978_visual}.%
+\footnote{These and further examples are discussed by \citet[p.\ 165]{currie:1997_mental}.} Further, imagining seeing something (for example, imagining seeing a visual mask) can modulate and interfere with actually seeing in much the way that actually seeing the thing imagined would \citep{pearson:2008_functional,ishai:1995_common}.
+The similarities in characteristic performance profile and the particular patterns of interference are good (if non-decisive) reasons to conjecture that imagining seeing and actually seeing involve representations with a common format.
+This conjecture is indirectly supported by evidence that sensorily imagining seeing and actually seeing not only have a common neural basis but also involve similar patterns of cortical activation \citep[e.g.][]{page:2011_erp}.
+
+Let us turn to motor representation.
+Compare imagining moving a ball with actually moving a ball.
+To fully specify the comparison we intend, it is again necessary to distinguish two ways of imagining.
+One way of imagining action is phenomenologically something like acting except that such imaginings are not necessarily responsive to the features of actual objects and do not necessarily result in bodily movements.
+To illustrate suppose you are about to dive into a pool and, standing at the edge, mentally pantomime launching yourself from the bank.
+Some phenomenal characteristics of this imaginative exercise may be barely distinguishable from some of those associated with actually launching yourself into the pool.
+This way of imagining action can be distinguished from cognitive ways of imagining action which might involve thinking about an action.%
+\footnote{
+On distinguishing these two ways of imagining action, see \citet[p.\ 161]{currie:1997_mental}, \citet[p.\ 727]{jeannerod:1995_mental}, and \citet[p.\ 638-9]{kosslyn:2001_neural}.
+The former, phenomenologically action-like imagining is sometimes labelled 'motor' or 'internal' and occasionally identified by its links to motor processes or by features of the format or content of the representations involved (\citealp[p.\ 1400]{annett:1995_motor}).
+We avoid these labels because we introduce the distinction by appeal to phenomenology only and do not stipulate that motor representations are involved.
+It is essential for what follows that the involvement of motor representations in the phenomenologically action-like way of imagining action is a discovery rather than a terminological stipulation.}
+%
+The comparison we intend is between imagining moving a ball in the former, phenomenologically-action like way and actually moving a ball.
+There is evidence that the way imagining performing an action unfolds in time is similar in some respects to the way actually performing an action of the same type would unfold.
+For instance, how long it takes to imagine moving an object is closely related to how long it would take to actually move that object \citep{decety:1989_timing, decety:1996_imagined, Jeannerod:1994oz}.
+In addition, for actions such as grasping the handle of a cup, manipulating the target object in ways that make the action harder (such as orienting the cup's handle to make it less convenient for you to grasp) make a corresponding difference to the effort involved in imagining performing the action \citep{parsons:1994_temporal, frak:2001_orientation}.
+Further, imagining performing an action can selectively interfere with performance of a related action.
+For example, suppose you are faced with an array of objects one of which---the target---you will shortly be required to grasp.
+Subjects who imagine grasping an object other than the target object tend to be slower in subsequently grasping the target object than subjects who do not imagine acting or subjects who imagine grasping the target object \citep{ramsey:2010_incongruent_}.
+Just as the similarities between imagining seeing and actually seeing are evidence for the hypothesis that the representations involved in imagining seeing and actually seeing have a common format, so also the similarities in characteristic performance profile between imagining acting and actually acting, together with the particular patterns of interference between the two, suggest that imagining acting and actually acting involve a common representational format.
+And much as in the case of seeing and imagining seeing, acting and imagining acting involve many of the same processes almost up to the actual muscle contractions \citep{jeannerod:1995_mental, jeannerod:2003_mechanism}.
+
+The fact that motor representation is involved in one way of imagining acting means that we can use imagining acting to investigate the format of motor representations.
+This enables us to examine characteristics of processes involving motor representation which are not directly consequences of an agent's actual body or actual environment.
+
+Differences in performance provide reasons to distinguish the formats of visual and motor representations.
+To see why, consider two tasks involving mental rotation.
+Judging the laterality of a rotated letter is thought to involve the phenomenologically seeing-like form of imagination \citep{jordan:2001_cortical}, whereas judging the laterality of a rotated hand is thought to involve the phenomenologically action-like form of imagination \citep{parsons:1987_imagined, gentilucci:1998_right}.
+Ordinary subjects who are asked to judge the laterality of a hand rotated to various degrees are slower and less accurate when the hand's position is maximally biomechanically awkward, when the hand is rotated so that the middle finger points downwards.
+By contrast, no such effect occurs for comparable tasks involving letters rather than hands.
+Can this difference in performance in imagining hands and letters be entirely explained by a difference in what is imagined (hands vs.
+letters)?
+There are subjects who can perform both tasks but whose performance is not different for hands and letters \citep{Fiori:2012fk}.
+These are subjects suffering Amyotrophic Lateral Sclerosis (ALS), which impairs motor representation \citep{parsons:1998_cerebrally}.
+If the hand-letter difference in performance were entirely explained by a difference in what is imagined, we would expect ALS and ordinary subjects to exhibit the same difference in performance since both are imagining hands and letters.
+But this is not the case.
+So some differences in performance linked to visual and motor representation cannot be entirely explained by differences in content.
+This suggests that the differences in performance are explained by a difference in format.
+
+Given that we need to posit differences in the formats of motor and visual representations in order to explain performance differences in mental rotation, it is plausible that these differences in format also explain some of the differences in performance linked to the timing of, and effort required for, imagining acting and imagining seeing.
+For instance, contrast imagining rotating a ball with imagining seeing a ball rotating.%
+\footnote{It may be difficult in practice to imagine acting without also imagining seeing, and conversely; it may also sometimes be difficult to distinguish imagine acting from imagining seeing \citep[as][p.\ 170 note]{currie:1997_mental}.
+However ordinary subjects can separate the two well enough to confirm predictions about their differences \citep[see, e.g.,][]{kosslyn:2001_imagining}.
+}
+As already mentioned, how quickly the former can be done is a function of how long it would take the agent to rotate the ball, whereas how quickly the latter can be done depends on how rapidly the ball can rotate and still be perceived as rotating.
+Similarly, we mentioned that factors making actually acting more effortful also make imagining acting more effortful: such factors are unlikely to affect the effort involved in imagining seeing.
 The explanation for these differences between imagining acting and imagining seeing plausibly involves a difference in the formats of the representations involved.
-
-How do these findings bear on motor representation?
-Contrast imagining an action with actually performing that action.
-There is substantial evidence that imagining an action involves many of the same processes which would be involved in performing that action up to the actual muscle contractions \citep{jeannerod:1995_mental,jeannerod:2003_mechanism}.
-There is also independent evidence that the representations underpinning imagining acting have the same format as those involved in actually acting.
-This evidence takes the form of similarities in processing profiles.
-Just as some similarities between imagining seeing something and actually seeing something can be explained on the hypothesis that the representations involved in imagining seeing and actually seeing have a common format \citep{kosslyn:1994_image}, so also some similarities between imagining grasping a cup and actually grasping it are evidence that the representations involved in imagining grasping are of the same format as those involved in actually grasping the cup.
-We can conclude, then, that imagining and executing actions are both underpinned by motor representations, and that motor representations differ in format from representations involved in both seeing and imagining seeing.
+That the two involve different representational formats is consistent with evidence that these two imaginings also involve different processes \citep{kosslyn:2001_imagining}.
+For instance, each can be selectively impaired \citep{sirigu:2011_motor}; and factors such as limb amputation or hand posture can interfere with imagining acting without interfering with  imagining seeing \citep{nico:2004_left, vargas:2004s_influence,  fourkas:2006_influence}.
 
@@
 Why suppose that motor representations also differ in format from intentions?
-As we noted earlier, the format of a representation constrains its possible contents.
-Just as a representation's having a cartographic format restricts its possible contents, so the motor format would imply restrictions on representations having it.
-In imagining executing actions such as grasping a cup, there are bio-mechanical, dynamical and postural constraints, among others, on what can be imagined.
-These constraints are closely related to those which govern actually performing such actions \citep{Jeannerod:2001yb}.
-Such constraints are plausibly explained by the representational format of motor representations.
-Since no such constraints would be expected to apply where propositional representations are involved, we can conclude (as is widely assumed) that the format of motor representations is not propositional.
+Contrast two ways of imagining taking a shot in basketball, one involving the phenomenologically action-like kind of imagination and the other involving a cognitive kind of imagination.
+The contrast we require is roughly between the way a former player might imagine this and the way that someone who has only ever read about basketball might imagine it.
+As we have seen, the way the phenomenologically action-like kind of imagination unfolds in time and the amount of effort it involves will depend on bio-mechanical, dynamical and postural constraints, among others.
+These constraints are closely related to those which govern actually performing such actions \citep{Jeannerod:2001yb}, and some can be altered by acquiring or losing motor expertise.
+By contrast no such constraints would be expected always to apply where a cognitive kind of imagination is involved.
+In line with the general strategy of inferring differences in format from differences in characteristic performance profile, we conclude that motor representations differ in format from those involved in cognitive kinds of imagination, which are plausibly propositional.
+
+But are we too hasty in concluding that motor and propositional representations differ in format?
+It might be objected that their characteristic performance profiles are not so different, for in cognitive imagination the fact that an agent is imagining herself acting will mean that how she imagines the action unfolding will reflect constraints on what she can do.
+But while this will sometimes be the case, a cognitive kind of imagining need not involve imagining an action unfolding in a way consistent with one's actual abilities.
+What distinguishes the phenomenologically action-like form of imagination is that some bio-mechanical, dynamical and postural constraints are inescapable.
+To make this vivid consider imagining reaching for a distant object.
+If the object is manifestly far out of reach it will not normally be possible to do this using the phenomenologically action-like kind of imagination, whereas no such difficulty need occur where a cognitive kind of imagination is involved.
+After all, where a cognitive kind of imagination is involved one can imagine having much longer limbs (or an entirely different body) whereas this cannot be achieved at will where the phenomenologically action-like kind of imagination is involved.
+Finally, where there are constraints on a cognitive kind of imagination these are generally mediated by beliefs or suppositions about one's own abilities; it seems unlikely that this is true of the phenomenologically action-like kind of imagination.
 
@@
 Of course this analogy doesn't show that demonstrative concepts can defer to motor representations (at least not unless pantomimes are motor representations of actions).
-But now consider purely mental pantomime---that is, motor imagery.
-One might use motor imagery to explore different ways of completing a task and then, having hit on a good solution, think to oneself `Do that!' It seems possible that in some such cases the demonstrative refers by deferring to a motor representation of action involved in mental motor imagery.
+But now consider purely mental pantomime---that is, the phenomenologically-action like kind of imagination.
+One might use this kind of imagination to explore different ways of completing a task and then, having hit on a good solution, think to oneself `Do that!' It seems possible that in some such cases the demonstrative refers by deferring to a motor representation of action involved in imagining acting.
 
 But is this really possible?
-Someone might object that in appealing to motor imagery we are sneaking intentions in through the back door.
+Someone might object that in appealing to imagining acting we are sneaking intentions in through the back door.
 How can we be sure that it is ever really a motor representation rather than an intention that one defers to in thinking `Do that!'?
-Contrast two cases of motor imagery, both involving a tool.
+Contrast two cases of imagining acting, both involving a tool.
 In the first case, imagine grasping an object with the tool; and in the second case, imagine releasing the same object with the tool.
-The two cases involve different imagery.
-But in what does this difference consist?
-Could it be a difference in intention that explains how the two cases of motor imagery differ?
+The two cases are different but in what does their difference consist?
+Could it be a difference in intention that explains how they differ?
 Since it is possible to grasp without intentionally grasping, it is surely also possible to imagine grasping without imagining intentionally grasping.
@@
 (It is easy to miss this point by confusing it with the issue of whether the imagining must itself be intentional.
-Whether or not it is possible to imagine grasping or releasing without \emph{intending to imagine} so acting, one can certainly imagine grasping or releasing without \emph{imagining intentionally} so acting.) So to rule out the possibility that the two cases of motor imagery differ with respect to intention, let the motor imagery be neutral with respect to intention.
+Whether or not it is possible to imagine grasping or releasing without \emph{intending to imagine} so acting, one can certainly imagine grasping or releasing without \emph{imagining intentionally} so acting.) So to rule out the possibility that the two imaginings differ with respect to intention, let both be neutral with respect to intention.
 Is the difference between the two cases then due to differing movements or patterns of muscle contraction?
 To rule this possibility out, let the movements and muscle contractions involved in both cases be as similar as possible: let the difference between grasping and releasing the object be a matter of how the tool is configured rather than of how your body moves.
-With the contrast cases elaborated by these two stipulations, it is plausible that the difference in imagery between the two cases will be due to the different motor representations involved in grasping and releasing.
-So in these cases, the thought `Do that!' will refer to grasping (or releasing) by deferring to a motor image of grasping (or of releasing).
-Since motor images are a species of motor representation, it follows that demonstrative concepts can refer to actions by deferring to motor representations.
+With the contrast cases elaborated by these two stipulations, it is plausible that the difference between the two imaginings will be due to the different motor representations involved in grasping and releasing.
+So in these cases, the thought `Do that!' will refer to grasping (or releasing) by deferring to a motor representation of grasping (or of releasing).
+It follows that demonstrative concepts can refer to actions by deferring to motor representations.
 
-So far we have focused on motor imagery.
-We are not committed to claiming that demonstrative concepts can only refer to actions by deferring to motor images.
-Given the parallels between motor imagery and motor preparation \citep{Jeannerod:2001yb}, it is plausible that motor representations feature in experiences of action in such a way that motor actions are available for demonstrative reference.
-If so, the possibility of demonstrative concepts which refer to actions by deferring to motor representations is supported not just by motor imagery but also by experiences associated with motor activity more generally.
+So far we have focused on imagining acting.
+We are not committed to claiming that demonstrative concepts can only refer to actions by deferring to motor representations involved in imagination.
+Given the parallels between the phenomenologically action-like kind of imagination and motor preparation \citep{Jeannerod:2001yb}, it is plausible that motor representations feature in experiences of action in such a way that motor actions are available for demonstrative reference.
+If so, the possibility of demonstrative concepts which refer to actions by deferring to motor representations is supported not just by experiences associated with imagining acting but also by experiences associated with actually acting.
+
+But how do demonstrative concepts feature in practical reasoning, which does not appear always to involve imagining or experiencing acting?
+Our proposal is similar to what one might say about demonstrative route concepts.
+Someone encounters a map with a route marked on it.
+Her experience of this route is necessary for her to acquire a demonstrative concept which refers to the route by deferring to the cartographic representation of it.
+But once she has this demonstrative concept, she can use it on future occasions without fresh experiences of the route (although there may be some dependence on memory); and her use of this concept does not depend on the continued existence of the original representation of the route.
+Similarly, on our view experience of action is necessary for the acquisition of demonstrative concepts of action such as concepts of grasping and reaching but, perhaps subject to requirements on memory, not for their continued use in practical thought.
 
@@
 For example visually encountering a mug sometimes involves representing features such as the orientation and shape of its handle in motor terms \citep{buccino:2009_broken,costantini:2010_where,cardellicchio:2011_space,Tucker:1998,tucker:2001_potentiation}.
-Reference by deference to motor representation could therefore explain how action outcomes match with respect not only to ways of acting but also with respect to what is to be acted on.\footnote{ In fact, there may be another way of explaining non-accidental matching with respect to what is to be acted on due to \citet[pp.\ 36-8, 44-5, 48-57]{Campbell:2002ge}.
+Reference by deference to motor representation could therefore explain how action outcomes match with respect not only to ways of acting but also with respect to what is to be acted on.%
+\footnote{ In fact, there may be another way of explaining non-accidental matching with respect to what is to be acted on due to \citet[pp.\ 36-8, 44-5, 48-57]{Campbell:2002ge}.
 In outline, Campbell's idea is to note that intentions can refer to objects by means of perceptual demonstrative elements, and that perceptual and motor representations of objects may be sufficiently commensurate for a perceptually represented object to be reliably selected as a target for motor processes.
@@
 We shall leave open the question of whether non-accidental matching involves elements additional to those involved in the view we have been developing.
-Our claim is just that a solution to the interface problem need not involve anything other than components of intentions which refer to actions by deferring to motor representations of action outcomes.}
+Our claim is just that a solution to the interface problem need not involve anything other than components of intentions which refer to actions by deferring to motor representations of action outcomes.
+}
+
+At this point we should acknowledge a complication.
+Reference by deference to motor representation depends on motor experience.
+Above we suggested that motor experience is needed for the acquisition of demonstrative concepts of ways of acting but not for the continued use of such concepts in practical reasoning.
+However, things are more complicated where constituents of intentions refer to particular things to be acted on.
+Agents will not always have experienced these things in the past.
+Accordingly fully resolving the interface problem by appeal to demonstratives requires that motor representations of particular things are generally associated with experiences of those things.
+When we spontaneously intend to act on an object we have no previous experience of, our experience of that object would need to have motor as well as perhaps visual, tactile or acoustic aspects.
+There is indeed some evidence for the ubiquity of motor aspects of experience (and also evidence that motor representation does not always modulate how objects are experienced, but that is not relevant here).
+For example, temporarily changing subjects' motor abilities by artificially extending their reach systematically affects their judgements of how far away objects are \citep{linkenauger:2009_effects, costantini2011tool}.
+Given that such judgements are based on experience,\footnote{
+It may be objected that these judgements could reflect non-perceptual expectations; \citet[pp.\ 203-4]{witt:2011_action} review evidence against this possibility.} 
+this illustrates one way in which experience may be shaped in part by motor representation of objects.
+There is also evidence that perceived size can be influenced by ability \citep[e.g.][]{witt2009kicking}.
+In short, then, it is plausible that motor representation of objects modulates experiences of them.
+This indicates that it may be possible to refer not only to ways of acting by also to what to act on by deferring to motor representations.
 
@@
 \end{document}
+
+\bibliography{$HOME/endnote/phd_biblio}
+
+\end{document}

### 2012-03-27 intention motor representation with corrado sinigaglia/intention and motor action.ppr.r1.tex
after PPR resubmit (first revision)

@@
 
-\def \papersize {a4paper}
+%\def \papersize {a4paper}
+\def \papersize {letterpaper}
 
@@
 
-\def \ititle {Intention and motor representation in explaining action}
+\def \ititle {Intention and motor representation in purposive action}
 \def \isubtitle {}
-\def \iauthor {S. Butterfill* \& C. Sinigaglia**
-\\ 
-**Department of Philosophy, University of Warwick
-\\ 
-***Dipartimento di Filosofia, Università degli Studi di Milano}
+\def \iauthor {}
+%\def \iauthor {S. Butterfill* \& C. Sinigaglia**
+%\\ 
+%**Department of Philosophy, University of Warwick
+%\\ 
+%***Dipartimento di Filosofia, Università degli Studi di Milano}
 \def \iemail{s.butterfill@warwick.ac.uk}
@@
 \input{$HOME/Documents/submissions/preamble_steve_paper2}
-%\author{}
-%\date{}
-
-%\setromanfont[Mapping=tex-text]{Sabon LT Std}
+\author{}
+\date{}
 
@@
 \begin{abstract}
-Are there distinctive roles for intention and motor representation in explaining the purposiveness of action?
+\noindent
+Are there distinct roles for intention and motor representation in explaining the purposiveness of action?
 Standard accounts of action assign a role to intention but are silent on motor representation.
 The temptation is to suppose that nothing need be said here because motor representation is either only an enabling condition for purposive action or else merely a variety of intention.
-This paper provides reasons for resisting temptation.
+This paper provides reasons for resisting that temptation.
 Some motor representations, like intentions, coordinate actions in virtue of representing outcomes; but, unlike intentions, motor representations cannot feature as premises or conclusions in practical reasoning.
@@
 It also gives rise to a problem: were the roles of intention and motor representation entirely independent, this would impair effective action.
-Resisting temptation therefore requires explaining how intentions interlock with motor representations.
-The solution, we argue, is to recognise that the contents of intention can be partially determined by the contents of motor representations.
-Understanding this content-determining relation enables better understanding how intentions are related to actions.
+It is therefore necessary to explain how intentions interlock with motor representations.
+The solution, we argue, is to recognise that the contents of intentions can be partially determined by the contents of motor representations.
+Understanding this content-determining relation enables better understanding how intentions relate to actions.
 \end{abstract}
@@
 Our aim is to show that this is a mistake.
-We shall argue that, once some basic features of motor representations are properly understood, the standard view must be  refined and extended in ways that allow us to recognize distinctive roles for intention and motor representation in explaining the purposiveness of action.
+We shall argue that, once some basic features of motor representations are properly understood, the standard view must be  refined and extended in ways that allow us to recognize distinct roles for intention and motor representation in explaining the purposiveness of action.
 
@@
 
-Our overall aim is to show that intention and motor representation play distinctive roles in explaining the purposiveness of action, that is, in relating actions to outcomes to which they are directed.
+Our overall aim is to show that intention and motor representation play distinct roles in explaining the purposiveness of action, that is, in relating actions to outcomes to which they are directed.
 The first temptation standing in our way is the temptation to suppose that motor representation, although perhaps important for action execution in some agents, has no bearing at all on our question about the relation between actions and the outcomes to which they are directed.
@@
 Third, a motor representation of an action outcome locates the action within a hierarchical structure.
-To illustrate, a left-handed precision grip and a right-handed whole hand grip are different action outcomes and both instances of grasping.\footnote{ On the notion of action outcome in motor representation, see \citet{jeannerod_motor_2006} and \citet{rizzolatti_mirrors_2008}.}
+To illustrate, a left-handed precision grip and a right-handed whole hand grip are different action outcomes and both instances of grasping.\footnote{On the notion of action outcome in motor representation, see \citet{gallese:2000_inner}, \citet{jeannerod_motor_2006} and \citet{rizzolatti_mirrors_2008}.}
 
@@
 Sensory feedback provides information about the actual course of the action which can be compared against the predictions.
-Adjustments are made in order to minimise the discrepancy between these \citep{wolpert:1995internal,miall:1996_forward}.
-Monitoring and control of this kind is a second functional role of motor representation.
+Adjustments are made in order to minimise the discrepancy between these \citep{wolpert:1995internal, miall:1996_forward}.
 
@@
 How do intentions ground the purposiveness of actions?
-On any standard view, an intention represents an outcome, causes an action, and does so in such a way as to facilitate the outcome's occurrence.
+On any standard view, an intention represents an outcome, causes an action, and does so in a way that would normally facilitate the outcome's occurrence.
 Similarly, motor representations of outcomes represent action outcomes, play a role in generating actions, and do this in a way that normally facilitates the occurrence of the outcomes represented.
 To say that motor representations do all this is one way of making precise the metaphor involved in saying that purposive actions are directed to outcomes.
-Moreover, there is a clear resemblance between the natural way of understanding intentions as grounding outcome-directedness and the way in which motor representations ground outcome-directedness.
+Moreover, there is a clear resemblance between the natural way of understanding intentions as grounding outcome-directedness and the way in which motor representations ground outcome-directedness (as \citealp{pacherie:2008_action}, pp.\ 189-90 has also argued).
 So given the functional roles and representational objects of motor representations, there is little room for doubt that motor representations can ground the outcome-directedness of purposive actions.
@@
 The difference between the two lines is a difference in content.
-Each of the routes could alternatively have been represented by a distinct series of instructions written on the same piece of paper; these cartographic and propositional representations differ in format.
+Each of the routes could alternatively have been represented by a distinct series of instructions written on the same piece of paper; these cartographic and propositional representations differ in format.%
+\footnote{Note that the distinction between content and format is orthogonal to issues about representational medium.
+The maps in our illustration may be paper map or electronic maps, and the instructions may be spoken, signed or written.
+This difference is one of medium.
+} 
 The format of a representation constrains its possible contents.
-For example, a representation with a cartographic format cannot represent what is represented by sentences such as `There could not be a mountain whose summit is inaccessible.'\footnote{Note that the distinction between content and format is orthogonal to issues about representational medium.
-The maps in our illustration may be paper map or electronic maps, and the instructions may be spoken, signed or written.
-This difference is one of medium.} The distinction between content and format is necessary because, as our illustration shows, each can be varied independently of the other.
+For example, a representation with a cartographic format cannot represent what is represented by sentences such as `There could not be a mountain whose summit is inaccessible.'  The distinction between content and format is necessary because, as our illustration shows, each can be varied independently of the other.
 
 Our aim in this section is to show that motor representations differ from intentions with respect to their format.
-This is why the second temptation must be resisted.
+This is why the second temptation should be resisted.
 That motor representations differ in format from intentions shows that they are genuinely distinct phenomena.\footnote{ Readers already convinced that motor representation differs from intention in being non-conceptual will not need the following argument in order to conclude that they are distinct phenomena.
@@
 How in general can we identify or distinguish representational formats?
-Because a representational format is typically associated with characteristic performance profiles, it is sometimes possible to infer similarities and differences in representational format from similarities and differences in the processes in which representations feature.
+Because representational formats are typically associated with characteristic performance profiles, it is sometimes possible to infer similarities and differences in representational format from similarities and differences in the processes in which representations feature.
 This works both for artefactual and mental representations.
@@
 Other transformations, such as turning the route into a list of salient landmarks, may be easier for the person with the propositional route description.
-So some patterns of difference in the two people's performances may be explained by the difference in the formats of their representations.
+So some patterns of difference in the two people's performances may be explained by the difference in the formats of their representations; and some similarities in performance profile may likewise be explained by sameness of format.
 If we did not already know that the two people's route representations involved different formats, we might infer this from the facility with which each performed various transformations of the route.
@@
 Cognitive neuroscience frequently depends on inferences of just this type.
-To illustrate, compare imagining seeing an object moving and with actually seeing it move.
+\bf
+To illustrate, compare imagining seeing an object moving with actually seeing it move.
 For this comparison we need to distinguish two ways of imagining seeing.
 There is a way of imagining seeing which phenomenologically is something like seeing except that it does not necessarily involve being receptive to stimuli.
-This way of imagining seeing, sometimes called 'sensory imagining', is commonly distinguished from cognitive ways of imagining seeing which might for example involve thinking about seeing \citep[§2.1]{Gendler:2011_imagination}.%
-\footnote{Note that we define sensory imagining in terms of phenomenology and make no stipulations about the processes and representations involved.
-This is essential for our purposes, since we wish to consider how evidence bears on a conjecture about the format of representations involved in sensory imagining.}
-It is sensorily imagining seeing an object move that we wish to compare with actually seeing an object move.
+This way of imagining seeing, sometimes called `sensory imagining', is commonly distinguished from cognitive ways of imagining seeing which might for example involve thinking about seeing \citep[§2.1]{Gendler:2011_imagination}.%
+\footnote{Note that we define this way of imagining seeing in terms of phenomenology and stipulate nothing about the processes and representations involved.
+This is essential for our purposes, since we wish to consider evidence for conjectures about the format of representations it involves.}
+It is this way of imagining seeing an object move that we wish to compare with actually seeing an object move.
 These two have similarities in characteristic performance profile.
-For instance, whether an object can be seen all at once depends on its size and distance from the perceiver; strikingly, when subjects imagine seeing an object, whether they can imagine seeing it all at once depends in the same way on size and distance (\citealp{kosslyn:1978_measuring}; \citealp[p.\ 99ff]{kosslyn:1994_image}.
+For instance, whether an object can be seen all at once depends on its size and distance from the perceiver; strikingly, when subjects imagine seeing an object, whether they can imagine seeing it all at once depends in the same way on size and distance (\citealp{kosslyn:1978_measuring}; \citealp[p.\ 99ff]{kosslyn:1994_image}).
 Also, how long it takes to imagine looking over an object depends on the object's subjective size in the same way that how long it would take to actually look over that object would depend on its subjective size \citep{kosslyn:1978_visual}.%
-\footnote{These and further examples are discussed by \citet[p.\ 165]{currie:1997_mental}.} Further, imagining seeing something (for example, imagining seeing a visual mask) can modulate and interfere with actually seeing in much the way that actually seeing the thing imagined would \citep{pearson:2008_functional,ishai:1995_common}.
+\footnote{These and further examples are discussed by \citet[p.\ 165]{currie:1997_mental}.
+} 
+Further, imagining seeing something (for example, imagining seeing a visual mask) can modulate and interfere with actually seeing in much the way that actually seeing the thing imagined would \citep{pearson:2008_functional, ishai:1995_common}.
 The similarities in characteristic performance profile and the particular patterns of interference are good (if non-decisive) reasons to conjecture that imagining seeing and actually seeing involve representations with a common format.
-This conjecture is indirectly supported by evidence that sensorily imagining seeing and actually seeing not only have a common neural basis but also involve similar patterns of cortical activation \citep[e.g.][]{page:2011_erp}.
+This conjecture is indirectly supported by evidence that imagining seeing and actually seeing not only have a common neural basis but also involve similar patterns of cortical activation \citep[e.g.][]{page:2011_erp}.
 
@@
 To illustrate suppose you are about to dive into a pool and, standing at the edge, mentally pantomime launching yourself from the bank.
-Some phenomenal characteristics of this imaginative exercise may be barely distinguishable from some of those associated with actually launching yourself into the pool.
+In some respects the experiences involved in this imaginative exercise may be barely distinguishable from experiences that might be involved in actually launching yourself into the pool.
 This way of imagining action can be distinguished from cognitive ways of imagining action which might involve thinking about an action.%
@@
 On distinguishing these two ways of imagining action, see \citet[p.\ 161]{currie:1997_mental}, \citet[p.\ 727]{jeannerod:1995_mental}, and \citet[p.\ 638-9]{kosslyn:2001_neural}.
-The former, phenomenologically action-like imagining is sometimes labelled 'motor' or 'internal' and occasionally identified by its links to motor processes or by features of the format or content of the representations involved (\citealp[p.\ 1400]{annett:1995_motor}).
+The former, phenomenologically action-like imagining is sometimes labelled `motor' or `internal' and occasionally identified by its links to motor processes or by features of the format or content of the representations involved (\citealp[p.\ 1400]{annett:1995_motor}).
 We avoid these labels because we introduce the distinction by appeal to phenomenology only and do not stipulate that motor representations are involved.
-It is essential for what follows that the involvement of motor representations in the phenomenologically action-like way of imagining action is a discovery rather than a terminological stipulation.}
-%
-The comparison we intend is between imagining moving a ball in the former, phenomenologically-action like way and actually moving a ball.
+It is essential for what follows that the involvement of motor representations in the phenomenologically action-like way of imagining action is a discovery rather than a stipulation.
+}
+The comparison we intend is between imagining moving a ball in the former, phenomenologically action-like way and actually moving a ball.
 There is evidence that the way imagining performing an action unfolds in time is similar in some respects to the way actually performing an action of the same type would unfold.
@@
 
-The fact that motor representation is involved in one way of imagining acting means that we can use imagining acting to investigate the format of motor representations.
-This enables us to examine characteristics of processes involving motor representation which are not directly consequences of an agent's actual body or actual environment.
-
-Differences in performance provide reasons to distinguish the formats of visual and motor representations.
-To see why, consider two tasks involving mental rotation.
-Judging the laterality of a rotated letter is thought to involve the phenomenologically seeing-like form of imagination \citep{jordan:2001_cortical}, whereas judging the laterality of a rotated hand is thought to involve the phenomenologically action-like form of imagination \citep{parsons:1987_imagined, gentilucci:1998_right}.
-Ordinary subjects who are asked to judge the laterality of a hand rotated to various degrees are slower and less accurate when the hand's position is maximally biomechanically awkward, when the hand is rotated so that the middle finger points downwards.
-By contrast, no such effect occurs for comparable tasks involving letters rather than hands.
-Can this difference in performance in imagining hands and letters be entirely explained by a difference in what is imagined (hands vs.
-letters)?
-There are subjects who can perform both tasks but whose performance is not different for hands and letters \citep{Fiori:2012fk}.
-These are subjects suffering Amyotrophic Lateral Sclerosis (ALS), which impairs motor representation \citep{parsons:1998_cerebrally}.
-If the hand-letter difference in performance were entirely explained by a difference in what is imagined, we would expect ALS and ordinary subjects to exhibit the same difference in performance since both are imagining hands and letters.
-But this is not the case.
-So some differences in performance linked to visual and motor representation cannot be entirely explained by differences in content.
-This suggests that the differences in performance are explained by a difference in format.
-
-Given that we need to posit differences in the formats of motor and visual representations in order to explain performance differences in mental rotation, it is plausible that these differences in format also explain some of the differences in performance linked to the timing of, and effort required for, imagining acting and imagining seeing.
-For instance, contrast imagining rotating a ball with imagining seeing a ball rotating.%
-\footnote{It may be difficult in practice to imagine acting without also imagining seeing, and conversely; it may also sometimes be difficult to distinguish imagine acting from imagining seeing \citep[as][p.\ 170 note]{currie:1997_mental}.
+We have been comparing actually seeing with imagining seeing and actually acting with imagining acting as a first step towards arguing that visual and motor representations differ in format (which in turn will be background for an argument that motor representation and intention differ in format).
+This claim could not easily be established by comparing actually seeing with actually acting because performance differences between these might be explained by bodily or environmental factors only distantly related to the representations involved.
+By contrast, comparing performance in imagining seeing and imagining acting does provide reasons to conclude that visual and motor representations differ in format.
+To see why, contrast imagining rotating a ball with imagining seeing a ball rotating.%
+\footnote{Imagining acting without also imagining seeing may be difficult in practice, and conversely; it may also sometimes be difficult to distinguish imagining acting from imagining seeing \citep[as][p.\ 170 suggest]{currie:1997_mental}.
 However ordinary subjects can separate the two well enough to confirm predictions about their differences \citep[see, e.g.,][]{kosslyn:2001_imagining}.
@@
 As already mentioned, how quickly the former can be done is a function of how long it would take the agent to rotate the ball, whereas how quickly the latter can be done depends on how rapidly the ball can rotate and still be perceived as rotating.
-Similarly, we mentioned that factors making actually acting more effortful also make imagining acting more effortful: such factors are unlikely to affect the effort involved in imagining seeing.
-The explanation for these differences between imagining acting and imagining seeing plausibly involves a difference in the formats of the representations involved.
-That the two involve different representational formats is consistent with evidence that these two imaginings also involve different processes \citep{kosslyn:2001_imagining}.
+Similarly, we mentioned that factors making actually acting more effortful also make imagining acting more effortful.
+For instance, in some cases rotating a ball clockwise is easier than rotating it anti-clockwise, and so is imagining a ball rotate.
+By contrast, the effort involved in actually seeing or imagining seeing a ball rotate does not similarly differ depending on direction.
+These and other performance differences are plausibly a consequence of a difference in format between motor and visual representations.
+
+It may be objected that performance differences such as these can be explained without appealing to a difference in format.
+After all, rotating a ball involves an action whereas a ball rotating does not; consequently, imagining the former may be thought to differ from imagining the latter with respect to the contents of the representations involved.
+Supposing that there are differences in content here and in other cases, could these fully explain differences in performance profile?
+To see why not, consider two tasks involving mental rotation.
+Judging the laterality of a rotated letter is thought to involve phenomenologically vision-like imagination \citep{jordan:2001_cortical}, whereas judging the laterality of a rotated hand is thought to involve phenomenologically action-like imagination \citep{parsons:1987_imagined, gentilucci:1998_right}.
+Ordinary subjects who are asked to judge the laterality of a hand rotated to various degrees are less accurate when the hand's position is biomechanically awkward.
+By contrast, no such effect occurs for comparable tasks involving letters rather than hands.
+How could this difference in performance in imagining hands and letters be explained?
+Consider the claim that the difference in performance can be fully explained by a difference in the content of the representations involved.
+Initially this might seem plausible because one task involves hands whereas the other involves letters.
+However, there are subjects who can perform both tasks but whose performance is not different for hands and letters \citep{Fiori:2012fk}.
+These are subjects suffering Amyotrophic Lateral Sclerosis (ALS), which impairs motor representation \citep{parsons:1998_cerebrally}.
+Since ALS and ordinary subjects encounter the same stimuli and perform the same tasks, there seems to be no reason (other than our hypothesis about a difference in format) to suppose that the two groups' performance involves representations with different contents.
+So if the hand-letter difference in performance were entirely explained by a difference in content, we would expect ALS and ordinary subjects to exhibit the same difference in performance.
+But this is not the case.
+This is an obstacle to supposing that the hand-letter difference in performance in ordinary subjects could be explained by appeal to content.
+
+The hypothesis that visual and motor representations differ in format is consistent with evidence that imagining acting and imagining seeing involve different processes \citep{kosslyn:2001_imagining}.
 For instance, each can be selectively impaired \citep{sirigu:2011_motor}; and factors such as limb amputation or hand posture can interfere with imagining acting without interfering with  imagining seeing \citep{nico:2004_left, vargas:2004s_influence,  fourkas:2006_influence}.
@@
 The contrast we require is roughly between the way a former player might imagine this and the way that someone who has only ever read about basketball might imagine it.
-As we have seen, the way the phenomenologically action-like kind of imagination unfolds in time and the amount of effort it involves will depend on bio-mechanical, dynamical and postural constraints, among others.
+As we have seen, the way phenomenologically action-like imagination unfolds in time and the amount of effort it involves will depend on bio-mechanical, dynamical and postural constraints, among others.
 These constraints are closely related to those which govern actually performing such actions \citep{Jeannerod:2001yb}, and some can be altered by acquiring or losing motor expertise.
@@
 If the object is manifestly far out of reach it will not normally be possible to do this using the phenomenologically action-like kind of imagination, whereas no such difficulty need occur where a cognitive kind of imagination is involved.
-After all, where a cognitive kind of imagination is involved one can imagine having much longer limbs (or an entirely different body) whereas this cannot be achieved at will where the phenomenologically action-like kind of imagination is involved.
+After all, where a cognitive kind of imagination is involved one might imagine having much longer limbs (or an entirely different body) whereas this cannot be achieved at will where the phenomenologically action-like kind of imagination is involved.
 Finally, where there are constraints on a cognitive kind of imagination these are generally mediated by beliefs or suppositions about one's own abilities; it seems unlikely that this is true of the phenomenologically action-like kind of imagination.
+
+\rm
 
@@
 
-The interface problem is the problem of explaining how, some of the time, there could be reliable matches between outcomes variously specified by intentions and motor representations.
+The interface problem is the problem of explaining how, some of the time, there could be non-accidental matches between outcomes variously specified by intentions and motor representations.
 As the previous section explained, the problem arises because intentions and motor representations have different representational formats.
@@
 Of course this analogy doesn't show that demonstrative concepts can defer to motor representations (at least not unless pantomimes are motor representations of actions).
-But now consider purely mental pantomime---that is, the phenomenologically-action like kind of imagination.
+But now consider purely mental pantomime---that is, phenomenologically action-like imagination.
 One might use this kind of imagination to explore different ways of completing a task and then, having hit on a good solution, think to oneself `Do that!' It seems possible that in some such cases the demonstrative refers by deferring to a motor representation of action involved in imagining acting.
@@
 How can we be sure that it is ever really a motor representation rather than an intention that one defers to in thinking `Do that!'?
-Contrast two cases of imagining acting, both involving a tool.
+Contrast two cases of phenomenologically action-like imagination, both involving a tool.
 In the first case, imagine grasping an object with the tool; and in the second case, imagine releasing the same object with the tool.
@@
 Could it be a difference in intention that explains how they differ?
-Since it is possible to grasp without intentionally grasping, it is surely also possible to imagine grasping without imagining intentionally grasping.
-And note that to imagine grasping without imagining intentionally grasping does not necessarily require one to imagine non-intentionally grasping.
+Since it is possible to grasp without intending to grasp, it is surely also possible to imagine grasping without imagining intending to grasping.
+And note that to imagine grasping without imagining intending to grasping does not necessarily require one to imagine non-intentionally grasping.
 (It is easy to miss this point by confusing it with the issue of whether the imagining must itself be intentional.
-Whether or not it is possible to imagine grasping or releasing without \emph{intending to imagine} so acting, one can certainly imagine grasping or releasing without \emph{imagining intentionally} so acting.) So to rule out the possibility that the two imaginings differ with respect to intention, let both be neutral with respect to intention.
+Whether or not it is possible to imagine grasping or releasing without intending to imagine so acting, one can certainly imagine grasping or releasing without  imagining intending to so act.) So to rule out the possibility that the two imaginings differ with respect to intention, let both be neutral with respect to intention.
 Is the difference between the two cases then due to differing movements or patterns of muscle contraction?
@@
 We are not committed to claiming that demonstrative concepts can only refer to actions by deferring to motor representations involved in imagination.
-Given the parallels between the phenomenologically action-like kind of imagination and motor preparation \citep{Jeannerod:2001yb}, it is plausible that motor representations feature in experiences of action in such a way that motor actions are available for demonstrative reference.
-If so, the possibility of demonstrative concepts which refer to actions by deferring to motor representations is supported not just by experiences associated with imagining acting but also by experiences associated with actually acting.
+Given the parallels between the phenomenologically action-like kind of imagination and actually acting already discussed, it is plausible that demonstrative reference to action by deference to motor representation is made possible not just by experiences associated with imagining acting but also by experiences associated with actually acting.
 
-But how do demonstrative concepts feature in practical reasoning, which does not appear always to involve imagining or experiencing acting?
-Our proposal is similar to what one might say about demonstrative route concepts.
+\bf
+
+Having an intention with such a demonstrative concept does not generally require imagining or actually acting at the time the intention is formed.
+To see why, first note that many of the things we do are either things we have already done or else novel combinations of actions we have already performed, such as reaching, grasping and throwing.
+Relatedly, motor representations of actions which, taken as a whole, are novel can often be built up from motor representations of familiar actions \citep[pp.\ 45-49]{rizzolatti_mirrors_2008}.
+The second step in our proposal is similar to what one might say about demonstrative route concepts.
 Someone encounters a map with a route marked on it.
@@
 
+\rm
+
 We have been arguing for the existence of demonstrative concepts which refer to actions by deferring to motor representations in order to solve the interface problem.
-As already stated, the interface problem is the problem of explaining how outcomes variously specified by intentions and motor representations, despite their different formats, sometimes non-accidentally match.
+As already stated, the interface problem is the problem of explaining how it is sometimes no accident that an intention and a motor representation specify matching outcomes despite differing in format.
 As long as we think of intentions and motor representations as having logically independent contents, it seems that fully solving the problem would require appeal to processes of translation linking intention with motor representation.
@@
 Reference by deference to motor representation could therefore explain how action outcomes match with respect not only to ways of acting but also with respect to what is to be acted on.%
-\footnote{ In fact, there may be another way of explaining non-accidental matching with respect to what is to be acted on due to \citet[pp.\ 36-8, 44-5, 48-57]{Campbell:2002ge}.
+\footnote{
+\label{fn:campbell_location}
+In fact, there may be another way of explaining non-accidental matching with respect to what is to be acted on due to \citet[pp.\ 36-8, 44-5, 48-57]{Campbell:2002ge}.
 In outline, Campbell's idea is to note that intentions can refer to objects by means of perceptual demonstrative elements, and that perceptual and motor representations of objects may be sufficiently commensurate for a perceptually represented object to be reliably selected as a target for motor processes.
@@
 
+\bf
+
 At this point we should acknowledge a complication.
-Reference by deference to motor representation depends on motor experience.
-Above we suggested that motor experience is needed for the acquisition of demonstrative concepts of ways of acting but not for the continued use of such concepts in practical reasoning.
-However, things are more complicated where constituents of intentions refer to particular things to be acted on.
-Agents will not always have experienced these things in the past.
-Accordingly fully resolving the interface problem by appeal to demonstratives requires that motor representations of particular things are generally associated with experiences of those things.
-When we spontaneously intend to act on an object we have no previous experience of, our experience of that object would need to have motor as well as perhaps visual, tactile or acoustic aspects.
+Above we argued that experience is needed for the acquisition of demonstrative concepts of ways of acting but not for their continued use in practical reasoning.
+This matters because forming an intention does not always involve newly experiencing a way of acting.
+However, things are more complicated concerning what to act on.
+Where an action involves particular things to be acted on, agents will not generally have past experiences of these.
+Accordingly, demonstrative reference by deference to motor representation requires that some motor representations of particular targets for action are associated with experiences of those targets.
+When we spontaneously intend to act on an object of which we have no previous experience (and perhaps in other cases too), our experience of that object would have to depend in part on motor representations of it.
 There is indeed some evidence for the ubiquity of motor aspects of experience (and also evidence that motor representation does not always modulate how objects are experienced, but that is not relevant here).
@@
 Given that such judgements are based on experience,\footnote{
-It may be objected that these judgements could reflect non-perceptual expectations; \citet[pp.\ 203-4]{witt:2011_action} review evidence against this possibility.} 
+It may be objected that these judgements could reflect non-experiential expectations; \citet[pp.\ 203-4]{witt:2011_action} reviews evidence against this possibility.
+} 
 this illustrates one way in which experience may be shaped in part by motor representation of objects.
-There is also evidence that perceived size can be influenced by ability \citep[e.g.][]{witt2009kicking}.
-In short, then, it is plausible that motor representation of objects modulates experiences of them.
-This indicates that it may be possible to refer not only to ways of acting by also to what to act on by deferring to motor representations.
+There is also evidence that experiential judgements of size can be influenced by ability \citep[e.g.][]{witt2009kicking}.
+It is plausible, then, that motor representation of objects modulates experiences of them.
+This indicates that it is possible to refer not only to ways of acting by also to what to act on by deferring to motor representations.
+
+\rm
 
@@
 
-The problem of explaining this we called \emph{the interface problem}.
+The problem of explaining this we called the interface problem.
 Theoretically uncomplicated approaches to solving it would involve appeal to common causes or to processes of translation.
@@
 
-A further conjecture is that whenever there is a non-accidental match between the contents of intentions and of motor representations, this is due to deferential action concepts.
-The plausibility of this conjecture rests in part on the absence of any other viable solution to the interface problem, and of course we do not claim to have shown that no alternatives exists.
-If the conjecture is right, if there is no way of solving the interface problem other than by means of deferential action concepts, then concepts which refer by deferring to motor representations are what tie explicit practical reasoning to motor processes.
+A further conjecture is that whenever there is a non-accidental match between the contents of intentions and of motor representations, this is due at least in part to deferential action concepts.
+The plausibility of this conjecture rests in part on the absence of any other viable full solution to the interface problem, and of course we do not claim to have shown that no alternatives exists.
+If the conjecture is right, if there is no way of fully solving the interface problem other than by means of deferential action concepts, then concepts which refer by deferring to motor representations are what tie explicit practical reasoning to motor processes.
 
@@
 
-On this view experience of action plays a novel role.
-Action experiences in which motor representations feature, such as those associated with motor imagery and those associated with really acting, are arguably necessary for there to be concepts which are constituents of intentions and refer to actions by deferring to motor representations.
-But if, as we conjecture, such deference is necessary for intentions to properly and reliably result in bodily movements, it may turn out that intentionally acting in the world depends on action experiences featuring motor representation.
+On this view experience of acting plays a novel role.
+Experiences made possible by motor representation, such as those associated with phenomenologically action-like imagination and those associated with actual action, are arguably necessary for there to be concepts which are constituents of intentions and refer to actions by deferring to motor representations.
+But if, as we conjecture, such deference is necessary for intentions to properly and reliably result in bodily movements, it may turn out that intending to act in the world depends on experiences which are made possible by motor representation.
 Much as on some views thought about objects depends on perceptual experience \citep[e.g.][]{Campbell:2002ge}, so also intending actions may depend on motor experience.
@@
 \end{document}
-
-\bibliography{$HOME/endnote/phd_biblio}
-
-\end{document}

### 2012-03-29 intention motor representation with corrado sinigaglia/intention and motor action.ppr.r1.final.tex
submitted version of intention and motor representation; new mindreading minimal theory of mind version

--- 2012-03-27 intention motor representation with corrado sinigaglia/intention and motor action.ppr.r1.tex
@@
 %\\ 
-%**Department of Philosophy, University of Warwick
+%**Department of Philosophy, University of Warwick, UK
 %\\ 
-%***Dipartimento di Filosofia, Università degli Studi di Milano}
-\def \iemail{s.butterfill@warwick.ac.uk}
+%***Dipartimento di Filosofia, Università degli Studi di Milano, Italia}
+%\def \iemail{s.butterfill@warwick.ac.uk}
 %\date{}
@@
 \input{$HOME/Documents/submissions/preamble_steve_paper2}
-\author{}
+%\author{}
+\author {S. Butterfill* \& C. Sinigaglia**
+\\ 
+**Department of Philosophy, University of Warwick, UK
+\\ 
+***Dipartimento di Filosofia, Università degli Studi di Milano, Italia}
 \date{}
@@
 Cognitive neuroscience frequently depends on inferences of just this type.
-\bf
 To illustrate, compare imagining seeing an object moving with actually seeing it move.
@@
 Finally, where there are constraints on a cognitive kind of imagination these are generally mediated by beliefs or suppositions about one's own abilities; it seems unlikely that this is true of the phenomenologically action-like kind of imagination.
-
-\rm
 
@@
 
-\bf
-
 Having an intention with such a demonstrative concept does not generally require imagining or actually acting at the time the intention is formed.
@@
 Similarly, on our view experience of action is necessary for the acquisition of demonstrative concepts of action such as concepts of grasping and reaching but, perhaps subject to requirements on memory, not for their continued use in practical thought.
-
-\rm
 
@@
 
-\bf
-
 At this point we should acknowledge a complication.
@@
 This indicates that it is possible to refer not only to ways of acting by also to what to act on by deferring to motor representations.
-
-\rm
 

### 2012-03-29 mindreading/minimal theory of mind (new version)/minimal theory of mind.nv.tex
submitted version of intention and motor representation; new mindreading minimal theory of mind version

%!TEX TS-program = xelatex
%!TEX encoding = UTF-8 Unicode

\def \papersize {a4paper}
\documentclass[12pt,\papersize]{extarticle}
% extarticle is like article but can handle 8pt, 9pt, 10pt, 11pt, 12pt, 14pt, 17pt, and 20pt text

\def \ititle {How to Construct a Minimal Theory of Mind}
\def \isubtitle {}
\def \iauthor {Stephen A. Butterfill \& Ian A. Apperly}
\def \iemail{s.butterfill@warwick.ac.uk}
%for anonymous submisison
%\def \iauthor {}
%\def \iemail{}
%\date{}

\input{$HOME/Documents/submissions/preamble_steve_paper2}

\begin{document}

\setlength\footnotesep{1em}

\bibliographystyle{newapa} %apalike

%these two lines are for anonymous submission --- they remove author and date
%but don't forget to remove defs above as well --- otherwise it will be in the metadata
%\author{}
%\date{}

\maketitle
%\tableofcontents

\begin{abstract}
\noindent
What could someone represent that would enable her to track, at least within limits, perceptions, beliefs including false beliefs, and other propositional attitudes? 
An obvious possibility is that she might represent these very propositional attitudes as such.
It is sometimes tacitly or explicitly assumed that this is the only possible answer.
However we argue that several recent discoveries in developmental, cognitive, and comparative psychology indicate the need for other, less obvious possibilities.
Our aim is to meet this need by describing the construction of a minimal theory of mind.  
Minimal theory of mind is rich enough to explain systematic success on tasks held to be acid tests for theory of mind cognition including many false belief tasks.  
It is also extensible in ways that can explain a wide range of findings from non-human animals and human infants that are sometimes presented as evidence for full-blown theory of mind cognition.  
Yet minimal theory of mind does not require representing propositional attitudes, or any other kind of representation, as such.
Minimal theory of mind may be what enables those with limited cognitive resources or little conceptual sophistication, such as infants, chimpanzees, scrub-jays and human adults under load, able to track, within limits, facts about perceptions and beliefs.

\

\noindent
\textbf{Keywords:}
Theory of Mind, False Belief, belief, perception, development, comparative
\end{abstract}

\section{Introduction}
What could someone represent that would enable her to track, at least within limits, perceptions, beliefs including false beliefs, and other propositional attitudes? 
One answer is obvious: she might track these things by virtue of representing them as such, that is, by representing perceptions, beliefs, and other propositional attitudes as such.
Our aim in what follows is to identify another, less obvious answer.
There is a form of cognition---minimal theory of mind---which does not involve representing propositional attitudes as such but is  rich enough to enable systematic success on tasks held to be acid tests for theory of mind cognition including many false belief tasks.
As we will explain, this has consequences 
for interpreting a range of findings concerning infants', adults' and nonhumans' performances on theory of mind tasks.
It may also help us to understand what enables those with limited cognitive resources or little conceptual sophistication, such as infants, chimpanzees, scrub-jays and human adults under load, to track, within limits, facts about perceptions and beliefs.

In this section we first defend the theoretical coherence of our question and then explain the findings which motivate facing it.

To many our question may initially appear incomprehensible.
Could abilities to track false beliefs (say) really involve anything other than representing false beliefs?
To see the possibility of a positive answer it may help to consider a non-mental analogy.
What could someone represent that would enable her to track, at least within limits, the toxicity of potential food items?
Here the most straightforward answer (she could represent their toxicity) is clearly not the only one.
After all, someone might track toxicity by representing odour or by representing visual features associated with putrefaction, say.
Suppose Sin\'ead has no conception of toxins but represents the odours given off by food items and 
treats those which she thinks smell foul as dangerous to eat,
so that she would not normally offer them to friends or family
nor conceal them from her competitors.
This brings nutritional and competitive benefits thanks to facts about the the toxicity of food.
If Sin\'ead tends to behave in this way because of these benefits, 
it follows that she has an ability to track, in a limited but useful range of cases,  toxicity; and this ability involves  representing  odour only.
Our question, put roughly, is whether false beliefs have something like an odour.

To make the question more precise it is useful to distinguish 
theory of mind abilities from theory of mind cognition.  
A  \textit{theory of mind ability} is an ability that exists in part because exercising it brings benefits obtaining which depends on exploiting or influencing facts about others’ mental states.
To illustrate,
suppose that Hannah is able to discern whether Isabel's eyes are in view,
that Hannah exercises this ability to escape detection while stealing from Isabel,
that Hannah's ability exists in part because it benefits her in this way,
and 
that Hannah's escaping detection depends on exploiting a fact about Isabel's mental states (namely that she cannot usually see Hannah's acts of theft when Hannah doesn't have Isabel's eyes in view).
Then Hannah has a theory of mind ability.
(This is not supposed to be a plausible, real-world example but only to illustrate what the definition requires.)
By an ability to \textit{track} perceptions or beliefs (say), we mean a theory of mind ability which involves exploiting or influencing facts about perceptions or beliefs, respectively.
By contrast, \textit{theory of mind cognition} paradigmatically involves ascribing propositional attitudes such as beliefs, desires and intentions to give rationalising causal explanations of action.  
This distinction is important because the facts about other minds which theory of mind abilities exploit are not necessarily the facts which are represented in theory of mind cognition.  
To return to the illustration, Hannah is able, within limits, to exploit facts about what others perceive without representing perceptions as such. 
She has a theory of mind ability while possibly lacking any theory of mind cognition.

\small
\bibliography{$HOME/endnote/phd_biblio_en_record_num_keys}

\end{document}

### 2012-04-02 mindreading/minimal theory of mind (new version)/minimal theory of mind.nv.tex
working on new version of mtom

@@
 
-In this section we first defend the theoretical coherence of our question and then explain the findings which motivate facing it.
+In this section we defend explain our question; in the next section we introduce the findings which motivate facing it before starting to answer it in the third section.
 
-To many our question may initially appear incomprehensible.
+Some may find our question initially incomprehensible.
 Could abilities to track false beliefs (say) really involve anything other than representing false beliefs?
@@
 Here the most straightforward answer (she could represent their toxicity) is clearly not the only one.
-After all, someone might track toxicity by representing odour or by representing visual features associated with putrefaction, say.
-Suppose Sin\'ead has no conception of toxins but represents the odours given off by food items and 
-treats those which she thinks smell foul as dangerous to eat,
+After all, someone might track toxicity by representing odours or by representing visual features associated with putrefaction, say.
+Suppose Sin\'ead has no conception of toxins but represents the odours of food items and 
+treats those with foul odours as dangerous to eat,
 so that she would not normally offer them to friends or family
-nor conceal them from her competitors.
-This brings nutritional and competitive benefits thanks to facts about the the toxicity of food.
+nor conceal them from competitors.
+This brings nutritional and competitive benefits obtaining which depends on facts about toxicity.
 If Sin\'ead tends to behave in this way because of these benefits, 
-it follows that she has an ability to track, in a limited but useful range of cases,  toxicity; and this ability involves  representing  odour only.
-Our question, put roughly, is whether false beliefs have something like an odour.
+representing odours enables her to track, in a limited but useful range of cases,  toxicity.
+Our question, put very roughly, is whether   belief has something like an odour.
 
@@
 To illustrate,
-suppose that Hannah is able to discern whether Isabel's eyes are in view,
-that Hannah exercises this ability to escape detection while stealing from Isabel,
+suppose that Hannah is able to discern whether another's eyes are in view,
+that Hannah exercises this ability to escape detection while stealing from others,
 that Hannah's ability exists in part because it benefits her in this way,
 and 
-that Hannah's escaping detection depends on exploiting a fact about Isabel's mental states (namely that she cannot usually see Hannah's acts of theft when Hannah doesn't have Isabel's eyes in view).
+that Hannah's escaping detection depends on exploiting a fact about other's mental states (namely that they usually cannot  see Hannah's acts of theft when Hannah doesn't have their eyes in view).
 Then Hannah has a theory of mind ability.
 (This is not supposed to be a plausible, real-world example but only to illustrate what the definition requires.)
-By an ability to \textit{track} perceptions or beliefs (say), we mean a theory of mind ability which involves exploiting or influencing facts about perceptions or beliefs, respectively.
-By contrast, \textit{theory of mind cognition} paradigmatically involves ascribing propositional attitudes such as beliefs, desires and intentions to give rationalising causal explanations of action.
-This distinction is important because the facts about other minds which theory of mind abilities exploit are not necessarily the facts which are represented in theory of mind cognition.  
+An ability to \textit{track} perceptions or beliefs (say) is a theory of mind ability which involves exploiting or influencing facts about these states.
+By contrast, \textit{theory of mind cognition} involves representing mental states or processes as such.
+And \textit{full-blown} theory of mind cognition involves  representing propositional attitudes such as beliefs, desires and intentions to construct reason-giving, causal explanations of action.  
+The distinction between theory of mind abilities and theory of mind cognition matters because the facts about other minds which theory of mind abilities exploit are not necessarily the facts which are represented in theory of mind cognition.
 To return to the illustration, Hannah is able, within limits, to exploit facts about what others perceive without representing perceptions as such. 
 She has a theory of mind ability while possibly lacking any theory of mind cognition.
+
+It should be uncontroversial that some theory of mind abilities do not necessarily involve any theory of mind cognition at all. 
+Our question concerns abilities to track what others perceive and believe, including their false beliefs; these have been central in psychological research.
+Can anything less than full-blown theory of mind cognition  explain systematic success on a range of false belief tasks?
+
+\section{Motivation}
+Our question is theoretical: it concerns 
+not what anyone does represent
+but what someone could represent that would enable her, at least within limits, to track perceptions, beliefs and other propositional attitudes.
+The motivation for facing up to this question is, of course, partly empirical.
+
+Start with ordinary adult humans.
+Since these subjects can represent beliefs and other propositional attitudes as such, 
+it is natural to assume that such representations underpin their abilities to track perceptions and beliefs.
+But is this natural assumption correct?
+
+To see that it might not be, consider a further question.
+Are processes in which representations of perceptions and beliefs feature automatic?
+Roughly speaking,
+a process is \emph{automatic} if whether it occurs is to a significant degree independent of its relevance to the particulars of the subject's motives and aims.
+(Note that a process may occur spontaneously without thereby being automatic.)  
+Some evidence suggests that, for ordinary adult humans, belief ascription is automatic.
+For example,
+\citet{kovacs_social_2010} asked adults to identify the location of a ball.
+They found that adults were significantly slower to identify the ball's location when an onlooker had a false belief about the location of the ball,
+even though the onlooker's belief was not relevant to the task at all.
+Relatedly, \citet{Samson:2010jm} provide evidence that identifying what another perceives is automatic;  this finding is indirectly supported by  evidence that identifying what another perceives is not disrupted by a secondary executive task \citep{qureshi:2010_executive}.
+Taken together, these findings suggest that, at least in adults, some processes involving representation of perception and belief are automatic.
+
+But there is also a body of evidence supporting a different conclusion.
+\citet{back:2010_apperly} found that subjects are significantly slower to answer an unexpected question about another's belief when that belief is false compared to when it is true \citep[see also][]{apperly:2006_belief}.
+This suggests that, at least in adults, belief ascription is not automatic.
+There is also evidence that, even in relatively simple situations, using representations of others' beliefs is not automatic \citep{Keysar:2003xu,apperly:2010_limits}.
+The case for nonautomaticity is indirectly supported by evidence that reasoning about perceptions and beliefs---and even merely holding in mind what another believes, where no inference is required---involves a measurable processing cost  \citep{apperly:2008_back,apperly:2010_limits}, consumes attention and working memory in fully competent adults (\citealp{en_1698, lin:2010_reflexively, en_1547} experiments 4-5), may require inhibition \citep{bull:2008_role} and makes demands on executive function \citep{apperly:2004_frontal,samson:2005_seeing}.
+These findings, taken together, suggest that some processes involving representation of perception and belief are not automatic.
+
+The question was whether, in adult humans, the processes in which representations of perception and belief feature are automatic.  
+We have just seen that there is a systematic incongruity in the evidence.
+The incongruity  cannot easily be explained away by appeal to simple methodological factors or extraneous task demands.
+For instance, it may be tempting to suppose that the incongruity can be explained by distinguishing between linguistic and nonlinguistic tasks.  
+But belief ascription fails to be automatic even on some nonlinguistic tasks \citep[e.g.][]{apperly:2004_frontal}, and we know of no reason to assume that belief ascription could not be automatic on verbal theory of mind tasks.
+
+If the incongruity is not a methodological artefact, how should we interpret the evidence?
+Perhaps it should be taken at face value.
+In other cases, such as number and causation, it is already quite widely accepted that, in adult humans, some abilities to track these things are automatic whereas others are not.\footnote{
+On number: \citet{trick:1994_small};
+on causation: \citet{Michotte:1946nz}, \citet{Scholl:2004dx}.
+}
+The evidence suggests that the same may be true for perception and belief.
+In adult humans, some theory of mind abilities involve automatic processes whereas others depend on nonautomatic processes.
+
+A closely related view has already been elaborated and defended in more detail by
+\citet[]{Apperly:2009ju},
+although their argument complements ours by drawing  primarily on developmental and comparative research. 
+According to them,
+adults may enjoy efficient but inflexible forms of theory of mind cognition in addition to the full-blown form which involves representing beliefs and other propositional attitudes as such.
+While aspects of this conjecture have already been tested \citep[]{Samson:2010jm, en_2397, surtees_direct_2011}, it raises two complementary questions (as Apperly \& Butterfill themselves note).
+
+First, why isn't all theory of mind cognition automatic?
+Even without knowing in any detail how theory of mind cognition is implemented, it is plausible that what makes full-blown theory of mind cognition demanding is some feature, or combination of features, of the propositional attitudes.%
+\footnote{
+Several hypotheses about which feature of the propositional attitudes explains why full-blown theory of mind cognition is cognitively and conceptually demanding have been defended \citep[e.g.][]{en_1263, en_634, en_1269, en_78, en_81, en_404, 	%en_687, %bibtex error authors and year too similar
+	en_643, en_1130}.  
+More than one feature may contribute. 
+We are agnostic about which feature or features are to blame.
+}  
+%
+On any standard view, propositional attitudes form complex causal structures, have arbitrarily nest-able contents, interact with each other in uncodifiably complex ways and are individuated by their causal and normative roles in explaining thoughts and actions \citep[]{en_809, en_249}.
+If anything should consume working memory and other scarce cognitive resources, it is surely representing states with this combination of properties.
+
+Second, how could theory of mind cognition be automatic?
+Automatic processes do not usually significantly depend on inhibition, working memory or attention.
+***HERE
+
+\section{The need for a constructive approach}
+Davidson ...
+
+This is reflected in the dichotomy between behaviour reading (P&Vonk) \& full-blown ToM (Baillargeon review).
+
+Davidson might be right about ordinary, commonsense psychology.
+
+So a constructive approach is needed.
+
+\section{***}
+
+There is a growing body of evidence that theory of mind abilities are widespread.
+Scrub-jays selectively re-cache their food in ways that deprive competitors of knowledge of its location \citep[]{Clayton:2007fh}.
+Chimpanzees select routes to approach food which conceal them from a competitor’s view \citep[]{en_1546}, and retrieve food using strategies that optimize their return given what a dominant competitor has seen \citep[]{en_1545}.
+And children in their second year use pointing to provide information to others \citep[]{en_1093} in ways that reflect a partner’s ignorance or knowledge \citep[]{en_1699}, as well as providing more information to ignorant than knowledgeable partners when making requests \citep[]{en_1140}.
+One-year-old children also predict actions of agents with false beliefs about the locations of objects \citep[]{en_1092, en_1208} and choose different ways of interacting with others depending on whether their beliefs are true or false \citep[]{en_1783,Knudsen:2011fk,southgate:2010fb}.
+And in much the way that irrelevant facts about the contents of others’ beliefs modulate adult subjects’ response times, such facts also affect how long 7-month-old infants look at some stimuli \citep[]{kovacs_social_2010}.
+
+Is it already known what young children, chimpanzees or scrub-jays represent that enables them, within limits, to track others’ perceptions or beliefs?   
+Some hypotheses have been offered, 
+but each faces significant objections, as we shall briefly outline here.
+We do not suggest that these hypotheses are known to be false, only that they are not known to be true.
+This is what motivates articulating and testing  alternative hypotheses, such as the minimal theory of mind hypothesis we develop in following sections.
+
+The most straightforward answer would be to suppose that they represent perceptions, beliefs and other propositional attitudes as such 
+(see, e.g.\ \citealp{en_1138} and \citealp{en_1691} on infants).  
+But this answer faces several objections.
+A body of evidence from humans suggests that representing beliefs requires conceptual sophistication, for it has a protracted developmental course stretching over several years \citep[]{en_87, en_89} and its acquisition is tied to the development of executive function \citep[]{en_410, en_1130} and language \citep[]{en_1209}.
+Infants, scrub-jays and chimpanzees are all deficient in these.
+Development of reasoning about beliefs in humans may also be facilitated by explicit training \citep[]{en_85} and environmental factors such as siblings \citep[]{en_507, en_1299}.
+This is evidence that representations of belief in humans typically emerge from extensive participation in social interactions (as \citealp{en_1300} suggest).  
+If any of this is right, we must reject the hypothesis that infants, scrub jays and chimpanzees are representing propositional attitudes like belief as such.
+
+***
+
+Some have hypothesised that infants', chimpanzees' or scrub-jays' theory of mind abilities 
+do not involve any theory of mind cognition at all
+but are instead based on 
+representations of nonintentional behaviour only.\footnote{ 
+	On scrub-jays: \citet{en_1417};
+	 on chimpanzees: \citet{en_1413}; 
+	 on young children: \citet{perner:1988_developing,en_1168, en_1169}.
+}  
+Here we must be careful to distinguish two types of  hypothesis.
+One says of a particular group of subjects, nonhumans for instance, that their theory of mind abilities could in principle be fully explained by hypothetical behaviour reading capacities.
+
+A second, much stronger hypothesis is that a particular group of subjects' theory of mind abilities are explained by their actual behaviour reading capacities.
+One difficulty for this second type of hypothesis is that relatively little is known about actual behaviour reading capacities.\footnote{
+Key studies include
+	\citet{Newtson:1976ni}, 
+	\citet{Byrne:1999jk},
+	\citet{Baldwin:2001rs},
+	\citet{Saylor:2007pj} and
+	\citet{Baldwin:2008mw}.
+}
+Applied to infants, scrub-jays or chimpanzees several objections have been raised \citep[]{Apperly:2009ju,en_1553, Clayton:2007fh, en_1691, en_1393}.
+
+The first hypothesis is that, for some subjects, some actual theory of mind abilities can be explained by those subjects actual behaviour reading capacities.
+This is plausible, although relatively little is known about actual behaviour reading capacities.\footnote{
+Key studies include
+	\citet{Newtson:1976ni}, 
+	\citet{Byrne:1999jk},
+	\citet{Baldwin:2001rs},
+	\citet{Saylor:2007pj} and
+	\citet{Baldwin:2008mw}.
+}
+This hypothesis faces several objections \citep[]{en_1553, Clayton:2007fh, en_1691, en_1393}.
+
+\section{***}
+
+But what could other, non full-blown forms of theory of mind cognition involve representing?
+Several related proposals exist.  
+To mention only those we draw most directly on, Gomez (\citeyear[][p.\ 730]{en_1259}) has emphasized primitive intentional relations to objects established by gaze, O’Neil and Doherty have separately discussed a notion of engagement with objects (\citealp{en_1159, en_1140}), Call and Tomasello (\citeyear[][p.\ 58]{en_1669}) have suggested that chimpanzees track the `likely target' of others’ visual access and understand something about its effects on behaviour, and Whiten (\citeyear[]{en_1415, en_1416}) uses the notion of an `intervening variable' to explain primitive theory of mind notions.
+These are illuminating ideas and what follows can be seen as an elaboration and partial synthesis of them.
+The result---minimal theory of mind---is unlike its precursors in that it is rich enough to explain systematic success on a range of false belief tasks.
+Our approach is novel in this and others respects to which we return (in the Conclusion) after presenting the substance of our account.
 

### 2012-04-03 mindreading/minimal theory of mind (new version)/minimal theory of mind.nv.tex
mtom in progress

@@
 Our aim in what follows is to identify another, less obvious answer.
-There is a form of cognition---minimal theory of mind---which does not involve representing propositional attitudes as such but is  rich enough to enable systematic success on tasks held to be acid tests for theory of mind cognition including many false belief tasks.
+There is a form of cognition---minimal theory of mind---which does not involve representing propositional attitudes as such but does involve representing simpler, relational mental states which could, within limits, enable one to track propositional attitudes such as beliefs.
+Minimal theory of mind 
+is  rich enough to enable systematic success on tasks held to be acid tests for theory of mind cognition including many false belief tasks.
 As we will explain, this has consequences 
 for interpreting a range of findings concerning infants', adults' and nonhumans' performances on theory of mind tasks.
-It may also help us to understand what enables those with limited cognitive resources or little conceptual sophistication, such as infants, chimpanzees, scrub-jays and human adults under load, to track, within limits, facts about perceptions and beliefs.
+It may help us to understand what enables those with limited cognitive resources or little conceptual sophistication, such as infants, chimpanzees, scrub-jays and human adults under load, to track, within limits, facts about perceptions and beliefs.
 
@@
 Can anything less than full-blown theory of mind cognition  explain systematic success on a range of false belief tasks?
+We do not aim to argue that someone could track beliefs, true and false, without any theory of mind cognition at all.
+Our concern is rather with the construction of a minimal form of theory of mind cognition.
+As we shall explain, minimal theory of mind does involve representing  belief-like  states, but it does not involve representing beliefs or other propositional attitudes as such.
+
+%The notion that theory of mind cognition could involve representing non-propositional counterparts of belief 
+%According to Davidson, 
+%\begin{quote}
+%`We are stuck with our two main ways of describing and explaining things, one which treats objects and events as mindless, and the other which treats objects and events as having propositional attitudes.
+I see no way of bridging the gap by introducing an intermediate vocabulary.' \citep[p.\ 697]{Davidson:2003bw}
+%\end{quote}
+%It may be tempting to dismiss this assertion on the grounds that we can readily describe an infant as excited by a clapping game, or as preferring one toy to another. 
+%On the surface at least, these descriptions seem neither to involve propositional attitudes nor to involve treating the infant as mindless; it seems there are non-propositional forms of excitement and preference which are nevertheless mental states.
+%
+%Even so, Davidson is right that there is a genuine difficulty when it comes to understanding non-propositional counterparts of attitudes like belief and perception.
+We cannot rely entirely on commonsense here because our commonsense concepts of perception, belief, intention and action exhibit a form of holism: we grasp them only if we understand their interdependent roles in reason explanations (Davidson 1995b, 1995a).
+%
 
@@
 
-Start with ordinary adult humans.
-Since these subjects can represent beliefs and other propositional attitudes as such, 
+Consider ordinary adult humans.
+Since they can represent beliefs and other propositional attitudes as such, 
 it is natural to assume that such representations underpin their abilities to track perceptions and beliefs.
@@
 To see that it might not be, consider a further question.
-Are processes in which representations of perceptions and beliefs feature automatic?
+Is tracking others' perceptions and beliefs automatic?
 Roughly speaking,
@@
 (Note that a process may occur spontaneously without thereby being automatic.)  
-Some evidence suggests that, for ordinary adult humans, belief ascription is automatic.
+Some evidence suggests that, for ordinary adult humans, belief tracking is automatic.
 For example,
@@
 even though the onlooker's belief was not relevant to the task at all.
-Relatedly, \citet{Samson:2010jm} provide evidence that identifying what another perceives is automatic;  this finding is indirectly supported by  evidence that identifying what another perceives is not disrupted by a secondary executive task \citep{qureshi:2010_executive}.
-Taken together, these findings suggest that, at least in adults, some processes involving representation of perception and belief are automatic.
+Relatedly, \citet{Samson:2010jm} provide evidence that identifying what another perceives is automatic;  this finding is indirectly supported by  evidence that tracking others' perceptions is not disrupted by a secondary executive task \citep{qureshi:2010_executive}.
+Taken together, these findings suggest that, at least in adults, tracking others' perceptions and beliefs is sometimes automatic.
 
@@
 \citet{back:2010_apperly} found that subjects are significantly slower to answer an unexpected question about another's belief when that belief is false compared to when it is true \citep[see also][]{apperly:2006_belief}.
-This suggests that, at least in adults, belief ascription is not automatic.
-There is also evidence that, even in relatively simple situations, using representations of others' beliefs is not automatic \citep{Keysar:2003xu,apperly:2010_limits}.
-The case for nonautomaticity is indirectly supported by evidence that reasoning about perceptions and beliefs---and even merely holding in mind what another believes, where no inference is required---involves a measurable processing cost  \citep{apperly:2008_back,apperly:2010_limits}, consumes attention and working memory in fully competent adults (\citealp{en_1698, lin:2010_reflexively, en_1547} experiments 4-5), may require inhibition \citep{bull:2008_role} and makes demands on executive function \citep{apperly:2004_frontal,samson:2005_seeing}.
-These findings, taken together, suggest that some processes involving representation of perception and belief are not automatic.
+This suggests that, at least in adults, belief tracking is not automatic.
+There is also evidence that, even in relatively simple situations, 
+using facts about others' beliefs is not automatic \citep{Keysar:2003xu,apperly:2010_limits}.
+The case for nonautomaticity is indirectly supported by evidence that tracking perceptions and beliefs---and even merely holding in mind what another believes, where no inference is required---involves a measurable processing cost  \citep{apperly:2008_back,apperly:2010_limits}, consumes attention and working memory in fully competent adults (\citealp{en_1698, lin:2010_reflexively, en_1547} experiments 4-5), may require inhibition \citep{bull:2008_role} and makes demands on executive function \citep{apperly:2004_frontal,samson:2005_seeing}.
+These findings, taken together, suggest that tracking others' perceptions and beliefs is sometimes not automatic.
 
-The question was whether, in adult humans, the processes in which representations of perception and belief feature are automatic.  
-We have just seen that there is a systematic incongruity in the evidence.
-The incongruity  cannot easily be explained away by appeal to simple methodological factors or extraneous task demands.
-For instance, it may be tempting to suppose that the incongruity can be explained by distinguishing between linguistic and nonlinguistic tasks.  
-But belief ascription fails to be automatic even on some nonlinguistic tasks \citep[e.g.][]{apperly:2004_frontal}, and we know of no reason to assume that belief ascription could not be automatic on verbal theory of mind tasks.
+The question was whether, in adult humans,  tracking perception and belief is automatic.  
+If we assume, further, that either all such processes are automatic or else none are, then the evidence creates a conflict.
+This conflict  cannot easily be explained away by appeal to simple methodological factors or extraneous task demands.
+For instance, it may be tempting to suppose that the conflict can be explained by distinguishing between linguistic and nonlinguistic tasks.  
+But belief ascription may fail to be automatic even on some nonlinguistic tasks \citep[e.g.][]{apperly:2004_frontal}, and we know of no reason to assume that belief ascription could not be automatic on some linguistic theory of mind tasks (such as those where spontaneous tracking is already established, e.g.\ \citet{ferguson_listeners_2012}).
 
-If the incongruity is not a methodological artefact, how should we interpret the evidence?
+If the conflict is not a methodological artefact, how should we interpret the evidence?
 Perhaps it should be taken at face value.
+This means we must reject the assumption that tracking others' perceptions and beliefs is either always automatic or else always nonautomatic.
 In other cases, such as number and causation, it is already quite widely accepted that, in adult humans, some abilities to track these things are automatic whereas others are not.\footnote{
@@
 
-First, why isn't all theory of mind cognition automatic?
-Even without knowing in any detail how theory of mind cognition is implemented, it is plausible that what makes full-blown theory of mind cognition demanding is some feature, or combination of features, of the propositional attitudes.%
+First, why isn't tracking belief and perception always automatic?
+Consider what is involved in representing beliefs and other propositional attitudes.
+On any standard view, propositional attitudes form complex causal structures, have arbitrarily nest-able contents, interact with each other in uncodifiably complex ways and are individuated by their causal and normative roles in explaining thoughts and actions \citep[]{en_809, en_249}.
+If anything should consume working memory and other scarce cognitive resources, it is surely representing states with this combination of properties.
+So even without knowing in any detail how theory of mind cognition is implemented, it is plausible that some feature, or combination of features, of the propositional attitudes  makes full-blown theory of mind cognition demanding.%
 \footnote{
@@
 %
-On any standard view, propositional attitudes form complex causal structures, have arbitrarily nest-able contents, interact with each other in uncodifiably complex ways and are individuated by their causal and normative roles in explaining thoughts and actions \citep[]{en_809, en_249}.
-If anything should consume working memory and other scarce cognitive resources, it is surely representing states with this combination of properties.
+A possible explanation, then, is this.
+Tracking perception or belief is not always automatic because 
+it sometimes involves representing propositional attitudes as such,
+which typically or always places demands on working memory, attention and executive function  that are incompatible with automaticity.
 
-Second, how could theory of mind cognition be automatic?
-Automatic processes do not usually significantly depend on inhibition, working memory or attention.
-***HERE
+Second, how could  tracking perceptions or beliefs ever be automatic?
+If we assumed that such tracking always involved propositional attitudes as such, this question would present a puzzle.
+For, as we saw, representing propositional attitudes as such generally places demands on working memory, attention and executive function that are incompatible with automaticity.
+In some cases these demands might be overcome through automatization in something like the way that initially effortful numerical operations can through practice become automatic.\footnote{
+On the automatization of simple sums, see \citet{lefevre:1988_cognitive}.
+For the suggestion that something similar might happen concerning mental states, see \citet{Suddendorf:2003co}.
+%which is also discussed by \citet[p.\ 961]{Apperly:2009ju}.
+}
+However, almost nothing is known about to what extent, if any, automatization occurs in theory of mind. 
+And in any case automatization can only explain the automaticity of routine inferences.
+So it is possible that automatization, although perhaps important, does not fully explain the automaticity of some of adult humans' perception- and belief-tracking abilities.
+A full explanation may  depend on showing that tracking perceptions and beliefs can be done without  representing beliefs or other propositional attitudes as such.
+
+This is a source motivation for our question about what someone could represent that would enable her to track perceptions and beliefs.
+The existence of both automatic and nonautomatic tracking of perceptions and beliefs in human adults 
+suggests (without decisively showing, of course), 
+contrary to a natural assumption mentioned above,
+that not all of their abilities to track perceptions and beliefs involve representing propositional attitudes as such.
+
+Additional motivation for our question comes from evidence for theory of mind abilities in young children and infants.
+Children in their second year use pointing to provide information to others \citep[]{en_1093} in ways that reflect a partner’s ignorance or knowledge \citep[]{en_1699}, as well as providing more information to ignorant than knowledgeable partners when making requests \citep[]{en_1140}.
+One-year-old children also predict actions of agents with false beliefs about the locations of objects \citep[]{en_1092, en_1208} and choose different ways of interacting with others depending on whether their beliefs are true or false \citep[]{en_1783,Knudsen:2011fk,southgate:2010fb}.
+And in much the way that irrelevant facts about the contents of others’ beliefs modulate adult subjects’ response times, such facts also affect how long 7-month-old infants look at some stimuli \citep[]{kovacs_social_2010}.
+
+What do these infants and young children represent that enables them, within limits, to track others’ beliefs and other propositional attitudes?   
+The most straightforward answer would be to suppose that they represent perceptions, beliefs and other propositional attitudes as such \citep[e.g.][]{en_1138, en_1691}.  
+But this answer faces several objections.
+A body of evidence  suggests that representing beliefs requires conceptual sophistication, for it has a protracted developmental course stretching over several years \citep[]{en_87, en_89} and its acquisition is tied to the development of executive function \citep[]{en_410, en_1130} and language \citep[]{en_1209}.
+Infants and young children are deficient in these.
+Development of reasoning about beliefs in humans may also be facilitated by explicit training \citep[]{en_85} and environmental factors such as siblings \citep[]{en_507, en_1299}.  
+This is evidence that representations of belief in humans typically emerge from extensive participation in social interactions (as \citealp{en_1300} suggest).  
+If any of this is right, we must reject the hypothesis that infants are representing beliefs or other propositional attitudes as such.
+
+In principle one might suppose that infants and young childrens' abilities to track perception and belief 
+do not involve any theory of mind cognition at all
+but are instead based on 
+representations of nonintentional behaviour only.
+While it is possible in principle to explain some belief-tracking abilities by appeal to hypothetical behaviour reading capacities,\footnote{ 
+ See \citet{perner:1988_developing,en_1168, en_1169}.
+}  
+there is ***objections ... too little is known
+
+involve reasoning about behaviour only and so do not involve any theory of mind cognition at all.
+While it is possible that
+
+do not involve any theory of mind cognition at all
+but are instead based on 
+representations of nonintentional behaviour only.\footnote{ 
+	On scrub-jays: \citet{en_1417};
+	 on chimpanzees: \citet{en_1413}; 
+	 on young children: \citet{perner:1988_developing,en_1168, en_1169}.
+}  
+Here we must be careful to distinguish two types of  hypothesis.
+One says of a particular group of subjects, nonhumans for instance, that their theory of mind abilities could in principle be fully explained by hypothetical behaviour reading capacities.
+
+%Some readers may doubt that tracking perceptions or beliefs is ever automatic.
+%Others may endorse automaticity while doubting our suggestion that automatic tracking requires cognitive efficiency and therefore cannot generally involve representing propositional attitudes as such.
 
@@
 
-This is reflected in the dichotomy between behaviour reading (P&Vonk) \& full-blown ToM (Baillargeon review).
+%This is reflected in the dichotomy between behaviour reading (P&Vonk) \& full-blown ToM (Baillargeon review).
 
@@
 
-There is a growing body of evidence that theory of mind abilities are widespread.
-Scrub-jays selectively re-cache their food in ways that deprive competitors of knowledge of its location \citep[]{Clayton:2007fh}.
-Chimpanzees select routes to approach food which conceal them from a competitor’s view \citep[]{en_1546}, and retrieve food using strategies that optimize their return given what a dominant competitor has seen \citep[]{en_1545}.
-And children in their second year use pointing to provide information to others \citep[]{en_1093} in ways that reflect a partner’s ignorance or knowledge \citep[]{en_1699}, as well as providing more information to ignorant than knowledgeable partners when making requests \citep[]{en_1140}.
-One-year-old children also predict actions of agents with false beliefs about the locations of objects \citep[]{en_1092, en_1208} and choose different ways of interacting with others depending on whether their beliefs are true or false \citep[]{en_1783,Knudsen:2011fk,southgate:2010fb}.
-And in much the way that irrelevant facts about the contents of others’ beliefs modulate adult subjects’ response times, such facts also affect how long 7-month-old infants look at some stimuli \citep[]{kovacs_social_2010}.
-
 Is it already known what young children, chimpanzees or scrub-jays represent that enables them, within limits, to track others’ perceptions or beliefs?   
@@
 This is what motivates articulating and testing  alternative hypotheses, such as the minimal theory of mind hypothesis we develop in following sections.
-
-The most straightforward answer would be to suppose that they represent perceptions, beliefs and other propositional attitudes as such 
-(see, e.g.\ \citealp{en_1138} and \citealp{en_1691} on infants).  
-But this answer faces several objections.
-A body of evidence from humans suggests that representing beliefs requires conceptual sophistication, for it has a protracted developmental course stretching over several years \citep[]{en_87, en_89} and its acquisition is tied to the development of executive function \citep[]{en_410, en_1130} and language \citep[]{en_1209}.
-Infants, scrub-jays and chimpanzees are all deficient in these.
-Development of reasoning about beliefs in humans may also be facilitated by explicit training \citep[]{en_85} and environmental factors such as siblings \citep[]{en_507, en_1299}.
-This is evidence that representations of belief in humans typically emerge from extensive participation in social interactions (as \citealp{en_1300} suggest).  
-If any of this is right, we must reject the hypothesis that infants, scrub jays and chimpanzees are representing propositional attitudes like belief as such.
 
@@
 }
-Applied to infants, scrub-jays or chimpanzees several objections have been raised \citep[]{Apperly:2009ju,en_1553, Clayton:2007fh, en_1691, en_1393}.
+Applied to infants, scrub-jays or chimpanzees several objections have been raised \citep[]{Apperly:2009ju,en_1553, Clayton:2007fh, en_1691, Suddendorf:2003co}.
 
@@
 }
-This hypothesis faces several objections \citep[]{en_1553, Clayton:2007fh, en_1691, en_1393}.
+This hypothesis faces several objections \citep[]{en_1553, Clayton:2007fh, en_1691, Suddendorf:2003co}.
 

### 2012-04-03 mindreading/minimal theory of mind (new version)/minimal theory of mind.nv.tex
mtom in progress

@@
 
-In this section we defend explain our question; in the next section we introduce the findings which motivate facing it before starting to answer it in the third section.
+In this section we defend explain our question; in the next sections we introduce the findings which motivate facing it before starting to answer it in the fourth section.
 
@@
 As we shall explain, minimal theory of mind does involve representing  belief-like  states, but it does not involve representing beliefs or other propositional attitudes as such.
+
+The notion that some abilities to track perceptions or beliefs involve only on theory of mind cognition which does not involve representing perceptions or beliefs as such is not entirely novel.
+To mention only those we draw most directly on, Gomez (\citeyear[][p.\ 730]{en_1259}) has emphasized primitive intentional relations to objects established by gaze, O’Neil and Doherty have separately discussed a notion of engagement with objects (\citealp{en_1159, en_1140}), Call and Tomasello (\citeyear[][p.\ 58]{en_1669}) have suggested that chimpanzees track the `likely target' of others’ visual access and understand something about its effects on behaviour, and Whiten (\citeyear[]{en_1415, en_1416}) uses the notion of an `intervening variable' to explain primitive theory of mind notions.
+These are illuminating ideas and what follows can be seen as an elaboration and partial synthesis of them.
+The result---minimal theory of mind---is unlike its precursors in that it is rich enough to explain systematic success on a range of false belief tasks.
+Our approach is novel in this and others respects to which we return (in the Conclusion) after presenting the substance of our account.
 
@@
 Consider what is involved in representing beliefs and other propositional attitudes.
-On any standard view, propositional attitudes form complex causal structures, have arbitrarily nest-able contents, interact with each other in uncodifiably complex ways and are individuated by their causal and normative roles in explaining thoughts and actions \citep[]{en_809, en_249}.
+On any standard view, propositional attitudes form complex causal structures, have arbitrarily nestable contents, interact with each other in uncodifiably complex ways and are individuated by their causal and normative roles in explaining thoughts and actions \citep[]{en_809, en_249}.
 If anything should consume working memory and other scarce cognitive resources, it is surely representing states with this combination of properties.
@@
 
-Additional motivation for our question comes from evidence for theory of mind abilities in young children and infants.
+\section{More motivation}
+Further motivation for our question comes from evidence for theory of mind abilities in young children and infants.
 Children in their second year use pointing to provide information to others \citep[]{en_1093} in ways that reflect a partner’s ignorance or knowledge \citep[]{en_1699}, as well as providing more information to ignorant than knowledgeable partners when making requests \citep[]{en_1140}.
@@
 
-In principle one might suppose that infants and young childrens' abilities to track perception and belief 
+In principle an alternative would be to suppose that infants' and young children's abilities to track perceptions and beliefs 
 do not involve any theory of mind cognition at all
@@
 representations of nonintentional behaviour only.
-While it is possible in principle to explain some belief-tracking abilities by appeal to hypothetical behaviour reading capacities,\footnote{ 
- See \citet{perner:1988_developing,en_1168, en_1169}.
+It is arguably possible in principle to explain some belief-tracking abilities by appeal to hypothetical behaviour reading capacities  \citep{perner:1988_developing,en_1168, en_1169}.
+But there are several objections to the claim that the full range of even infants' abilities to track perceptions and beliefs could be explained in this way \citep{en_1691,Apperly:2009ju}.  
+And what is is currently known about humans' actual behaviour reading capacities suggests that they are unlikely to explain systematic success on false belief tasks.%
+\footnote{
+Key studies include
+	\citet{Newtson:1976ni}, 
+	\citet{Byrne:1999jk},
+	\citet{Baldwin:2001rs},
+	\citet{Saylor:2007pj} and
+	\citet{Baldwin:2008mw}.
 }
-there is ***objections ... too little is known
 
-involve reasoning about behaviour only and so do not involve any theory of mind cognition at all.
-While it is possible that
+Here, then, is a second source of motivation for our question about what someone could represent that would enable her, within limits, to track perceptions and beliefs.
+As we have seen, there are significant if not decisive objections to the two best developed conjectures about infant theory of mind abilities, the conjecture that these involve representing beliefs and other propositional attitudes as such and the conjecture that these involve representing nonintentional behaviour only.
+These objections, while not decisive, justify exploring alternatives.
 
-do not involve any theory of mind cognition at all
-but are instead based on 
-representations of nonintentional behaviour only.\footnote{ 
-	On scrub-jays: \citet{en_1417};
-	 on chimpanzees: \citet{en_1413}; 
-	 on young children: \citet{perner:1988_developing,en_1168, en_1169}.
-}  
-Here we must be careful to distinguish two types of  hypothesis.
-One says of a particular group of subjects, nonhumans for instance, that their theory of mind abilities could in principle be fully explained by hypothetical behaviour reading capacities.
+Theory of mind abilities are not only found in humans.
+For instance,
+scrub-jays can selectively re-cache their food in ways that deprive competitors of knowledge of its location \citep{Clayton:2007fh}, and  chimpanzees can both select routes to approach food which conceal them from a competitor’s view \citep[]{en_1546} and also retrieve food using strategies that optimize their return given what a dominant competitor has seen \citep[]{en_1545}.
+There is debate about the cognitive underpinnings of these abilities (e.g.\ 
+	\citealp{povinelli:2004vonk, 
+			Penn:2007ey,
+			Tomasello:2005ce,
+			Call:2008di}). 
+If it is not yet known precisely what explains these abilities 
+and if the available evidence does not already tightly constrain the space of admissible conjectures,
+then the construction of a minimal theory of mind may be relevant to these debates.
+For the conjecture that minimal theory of mind explains chimpanzees' or scrub-jays' abilities to track perceptions or beliefs can be empirically distinguished from conjectures about representations of nonintentional behaviour only and from conjectures about representations of perceptions, beliefs and other propositional attitudes as such (as we explain in section *** below).
 
-%Some readers may doubt that tracking perceptions or beliefs is ever automatic.
-%Others may endorse automaticity while doubting our suggestion that automatic tracking requires cognitive efficiency and therefore cannot generally involve representing propositional attitudes as such.
+%The construction of minimal theory of mind shows that there is a form of theory of mind cognition which does not involve representing beliefs or other propositional attitudes as such but is capable of explaining, within limits, systematic success on a range of false belief tasks.
+%As we shall show,
+%the conjecture that minimal theory of mind cognition underpins some abilities to track perceptions and belief in human adults and infants generates testable predictions that distinguish it from the alternatives.
 
@@
 
-So a constructive approach is needed.
+So a constructive approach is needed. ***
 
-\section{***}
+\section{Minimal theory of mind}
 
-Is it already known what young children, chimpanzees or scrub-jays represent that enables them, within limits, to track others’ perceptions or beliefs?   
-Some hypotheses have been offered, 
-but each faces significant objections, as we shall briefly outline here.
-We do not suggest that these hypotheses are known to be false, only that they are not known to be true.
-This is what motivates articulating and testing  alternative hypotheses, such as the minimal theory of mind hypothesis we develop in following sections.
+In this section we identify concepts and principles necessary for constructing a form of theory of mind cognition.
+We begin with someone, call her Lucky, capable  of representing nonintentional behaviour only and ask what more is needed for minimal theory of mind cognition.
+We describe Lucky’s progress with a series of five principles.
+The principles are constructed in such a way that it would be coherent to suppose that Lucky has the abilities codified by the first \textit{n} principles only.
+They are not intended to represent a developmental or evolutionary progression.
+The principles can also be extended to explain more sophisticated theory of mind abilities than those considered here.
+We restrict ourselves to these five principles because they are sufficient to explain success on some well-known false belief tasks.%
+\footnote{
+In standard false belief tasks, “[t]he subject is aware that he/she and another person [call him Maxi] witness a certain state of affairs x.
+Then, in the absence of the other person the subject witnesses an unexpected change in the state of affairs from x to y” \citep[][p.\ 106]{en_89}.
+The test concerns whether the subject realises that Maxi will falsely believe x to obtain.
+In many cases the states of affairs, x and y, differ only with respect to the location of an object \citep[e.g.][]{en_1092, en_1208, en_1824}.
+As we go on to discuss, our proposal for a minimal theory of mind could easily be extended to cover a range of other cases;
+but importantly there are also false belief tasks success on which cannot be explained by minimal theory of mind cognition (see section ***).
+}
 
-***
+	We aim to provide the core elements of a computational theory in Marr’s sense (\citeyear[][pp.\ 15-29]{en_917}) where our computational theory, unlike the standard full-blown theory of mind which hinges on beliefs, desires and other propositional attitudes, is one that could be realised in a cognitively efficient manner without requiring conceptual sophistication.\footnote{ 	To provide a computation theory in Marr’s sense is not to imply that its implementation will involve computational processes.}  There are multiple ways in which this computational theory might be implemented.
+We shall not discuss how the theory might be implemented here other than to note that it seems unlikely that the principles formulated below are represented explicitly.
+It is valuable to articulate the computational theory in some detail before formulating and testing conjectures about implementation.
 
-Some have hypothesised that infants', chimpanzees' or scrub-jays' theory of mind abilities 
-do not involve any theory of mind cognition at all
-but are instead based on 
-representations of nonintentional behaviour only.\footnote{ 
-	On scrub-jays: \citet{en_1417};
-	 on chimpanzees: \citet{en_1413}; 
-	 on young children: \citet{perner:1988_developing,en_1168, en_1169}.
+\subsection{First principle}
+
+***HERE
+
+Until now Lucky has been aware only of object-directed actions.
+A further basic unit of action in theory of mind cognition is goal-directed action.
+To illustrate, a paradigm goal-directed action would be drinking from a bottle, which typically involves multiple object-directed actions such as first grasping a bottle and then lifting a bottle.
+Many predictions do not directly concern object-directed actions but rather goal-directed actions.
+Awareness of goal-directed actions is therefore an important further step in acquiring minimal theory of mind cognition.
+
+The term ‘goal-directed action’ can be used to mean several things.
+One is intentional action.
+This notion is no use for constructing a minimal theory of mind.
+To represent intentional actions as such you also have to represent intentions and related propositional attitudes such as beliefs \citep[]{en_18}.
+Constructing a minimal theory of mind requires a basic notion of goal-directed action.
+We take the units of goal-directed action to be collections of one or more object-directed behaviours.
+We stipulate that for an outcome, \textit{g}, to be the goal of a collection of object-directed behaviours is for it to occur in order to bring about \textit{g}; that is, \textit{g} is the function of this collection.
+Here ‘function’ should be understood teleologically.
+On the simplest teleological construal of function, for an action to have the function of bringing about \textit{g} would be for actions of this type to have brought about \textit{g} in the past and for this action to occur in part because of this fact \citep[see further][]{en_144, en_141, en_162, en_139, en_161}.
+Lucky needs some ability to track the functions of things (in this sense of ‘function’) so that she can distinguish from among all the outcomes of an action those which are its goals.%
+\footnote{
+Note that the requirement is not that Lucky understands the theoretical account of functions, only that she can distinguish between things which have different functions in this theoretical sense of ‘function’.
+A wide variety of research supports the claim that young children, non-human primates and corvids track the functions of things (including
+	\citealp{en_1086},
+	\citealp{en_1318},
+	\citealp{en_1431},
+	\citealp{en_1447},
+	\citealp{en_1325} and
+	\citealp{en_1708}%
+).
 }
-Here we must be careful to distinguish two types of  hypothesis.
-One says of a particular group of subjects, nonhumans for instance, that their theory of mind abilities could in principle be fully explained by hypothetical behaviour reading capacities.
 
-A second, much stronger hypothesis is that a particular group of subjects' theory of mind abilities are explained by their actual behaviour reading capacities.
-One difficulty for this second type of hypothesis is that relatively little is known about actual behaviour reading capacities.\footnote{
-Key studies include
-	\citet{Newtson:1976ni}, 
-	\citet{Byrne:1999jk},
-	\citet{Baldwin:2001rs},
-	\citet{Saylor:2007pj} and
-	\citet{Baldwin:2008mw}.
-}
-Applied to infants, scrub-jays or chimpanzees several objections have been raised \citep[]{Apperly:2009ju,en_1553, Clayton:2007fh, en_1691, Suddendorf:2003co}.
+This is not an account of full-blown goal-directed action.
+It is what someone who has only a minimal grasp of goal-directed action might understand.
+Representing goals in this sense does not require representing representations because such goals are non-representational.
+Representing goals only involves representing outcomes as functions of behaviours.
 
-The first hypothesis is that, for some subjects, some actual theory of mind abilities can be explained by those subjects actual behaviour reading capacities.
-This is plausible, although relatively little is known about actual behaviour reading capacities.\footnote{
-Key studies include
-	\citet{Newtson:1976ni}, 
-	\citet{Byrne:1999jk},
-	\citet{Baldwin:2001rs},
-	\citet{Saylor:2007pj} and
-	\citet{Baldwin:2008mw}.
-}
-This hypothesis faces several objections \citep[]{en_1553, Clayton:2007fh, en_1691, Suddendorf:2003co}.
+To illustrate the difference between object-directed and goal-directed actions, consider a scrub-jay repeatedly caching and re-caching a worm (we take our inspiration from studies such as Emery \& Clayton, 2001).
+The outcome of this collection of object-directed caching behaviours is (let’s say) that an observer is deprived of information about the location of the worm.
+This is not an outcome of any of the constituent object-directed behaviours taken individually.
+It is only an outcome of the collection considered as a whole.
+The significance of the whole sequence can only by appreciated by Lucky once she is equipped to represent goals.
 
-\section{***}
-
-But what could other, non full-blown forms of theory of mind cognition involve representing?
-Several related proposals exist.  
-To mention only those we draw most directly on, Gomez (\citeyear[][p.\ 730]{en_1259}) has emphasized primitive intentional relations to objects established by gaze, O’Neil and Doherty have separately discussed a notion of engagement with objects (\citealp{en_1159, en_1140}), Call and Tomasello (\citeyear[][p.\ 58]{en_1669}) have suggested that chimpanzees track the `likely target' of others’ visual access and understand something about its effects on behaviour, and Whiten (\citeyear[]{en_1415, en_1416}) uses the notion of an `intervening variable' to explain primitive theory of mind notions.
-These are illuminating ideas and what follows can be seen as an elaboration and partial synthesis of them.
-The result---minimal theory of mind---is unlike its precursors in that it is rich enough to explain systematic success on a range of false belief tasks.
-Our approach is novel in this and others respects to which we return (in the Conclusion) after presenting the substance of our account.
+The first principle, then, is that object-directed actions form larger units of action which are directed to goals.
+This second principle is sufficient to explain some cases of imitative learning, which can be defined as attempting to reproduce the actions necessary to achieve a goal \citep[]{en_1317}.
 

### 2012-04-04 mindreading/minimal theory of mind (new version)/minimal theory of mind.nv.tex
before move away from give principles

@@
 
-In this section we identify concepts and principles necessary for constructing a form of theory of mind cognition.
-We begin with someone, call her Lucky, capable  of representing nonintentional behaviour only and ask what more is needed for minimal theory of mind cognition.
+In this section we begin with someone, call her Lucky, capable  of representing nonintentional behaviour only and ask what more is needed for minimal theory of mind cognition.
 We describe Lucky’s progress with a series of five principles.
@@
 The principles can also be extended to explain more sophisticated theory of mind abilities than those considered here.
-We restrict ourselves to these five principles because they are sufficient to explain success on some well-known false belief tasks.%
+We restrict ourselves to these five principles because they are sufficient to explain success on some false belief tasks.%
 \footnote{
@@
 In many cases the states of affairs, x and y, differ only with respect to the location of an object \citep[e.g.][]{en_1092, en_1208, en_1824}.
-As we go on to discuss, our proposal for a minimal theory of mind could easily be extended to cover a range of other cases;
+As we go on to discuss, our proposal for a minimal theory of mind could be extended to cover a range of other cases;
 but importantly there are also false belief tasks success on which cannot be explained by minimal theory of mind cognition (see section ***).
@@
 
-	We aim to provide the core elements of a computational theory in Marr’s sense (\citeyear[][pp.\ 15-29]{en_917}) where our computational theory, unlike the standard full-blown theory of mind which hinges on beliefs, desires and other propositional attitudes, is one that could be realised in a cognitively efficient manner without requiring conceptual sophistication.\footnote{ 	To provide a computation theory in Marr’s sense is not to imply that its implementation will involve computational processes.}  There are multiple ways in which this computational theory might be implemented.
+	We aim to provide the core elements of a computational theory in Marr’s sense (\citeyear[][pp.\ 15-29]{en_917}) where our computational theory, unlike the standard full-blown theory of mind which hinges on beliefs, desires and other propositional attitudes, is one that could be realised in a cognitively efficient manner without requiring conceptual sophistication.\footnote{ 	Computational theories in Marr’s sense are not necessarily  implemented by computational processes.}  There are multiple ways in which this computational theory might be implemented.
 We shall not discuss how the theory might be implemented here other than to note that it seems unlikely that the principles formulated below are represented explicitly.
@@
 
-***HERE
-
-Until now Lucky has been aware only of object-directed actions.
-A further basic unit of action in theory of mind cognition is goal-directed action.
-To illustrate, a paradigm goal-directed action would be drinking from a bottle, which typically involves multiple object-directed actions such as first grasping a bottle and then lifting a bottle.
-Many predictions do not directly concern object-directed actions but rather goal-directed actions.
-Awareness of goal-directed actions is therefore an important further step in acquiring minimal theory of mind cognition.
-
+The first principle concerns goal-directed action.
 The term ‘goal-directed action’ can be used to mean several things.
 One is intentional action.
-This notion is no use for constructing a minimal theory of mind.
-To represent intentional actions as such you also have to represent intentions and related propositional attitudes such as beliefs \citep[]{en_18}.
-Constructing a minimal theory of mind requires a basic notion of goal-directed action.
-We take the units of goal-directed action to be collections of one or more object-directed behaviours.
-We stipulate that for an outcome, \textit{g}, to be the goal of a collection of object-directed behaviours is for it to occur in order to bring about \textit{g}; that is, \textit{g} is the function of this collection.
+To represent intentional actions as such you also have to represent intentions or  propositional attitudes such as beliefs and desires \citep[]{en_18}.
+This notion is therefore no use for constructing a minimal theory of mind---our aim is to explain how Lucky could  track perceptions and beliefs without representing beliefs or other propositional attitudes as such.
+Instead we need a more basic notion of goal-directed action.
+
+A suitable notion of goal-directed action can be chracterised in terms of functions \citep{Taylor:1964tr,Dretske:1988sq}.
+The units of goal-directed action are events comprising mere bodily movements.  
+We stipulate that for an outcome, \textit{g}, to be the goal of some bodily movements is for these bodily movements to occur in order to bring about \textit{g}; that is, \textit{g} is the function of this collection.
 Here ‘function’ should be understood teleologically.
 On the simplest teleological construal of function, for an action to have the function of bringing about \textit{g} would be for actions of this type to have brought about \textit{g} in the past and for this action to occur in part because of this fact \citep[see further][]{en_144, en_141, en_162, en_139, en_161}.
-Lucky needs some ability to track the functions of things (in this sense of ‘function’) so that she can distinguish from among all the outcomes of an action those which are its goals.%
+Lucky needs some ability to track the functions of things (in this special sense of ‘function’) so that she can link some bodily movements to the goals to which they are directed.%
 \footnote{
@@
 
-This is not an account of full-blown goal-directed action.
-It is what someone who has only a minimal grasp of goal-directed action might understand.
-Representing goals in this sense does not require representing representations because such goals are non-representational.
-Representing goals only involves representing outcomes as functions of behaviours.
+This is not supposed to be a fully adequate account of goal-directed action.
+For our purposes what matters  is not whether the account correctly identifies what goal-directed action is, but rather that it characterises what someone who has only a minimal grasp of goal-directed action might understand.
+For comparison, consider an individual whose understanding of kinship could be characterised by an incorrect theory of social relations.
+In practice it may not matter that the individual fails to fully understand kinship providing that she can reliably identify who is whose kin in her everyday life.  
+Similarly, Lucky does not need to fully understand goal-directed action in order to be able to pick out, in a limited but useful range of cases, which bodily movements are directed to which outcomes.
 
-To illustrate the difference between object-directed and goal-directed actions, consider a scrub-jay repeatedly caching and re-caching a worm (we take our inspiration from studies such as Emery \& Clayton, 2001).
-The outcome of this collection of object-directed caching behaviours is (let’s say) that an observer is deprived of information about the location of the worm.
-This is not an outcome of any of the constituent object-directed behaviours taken individually.
-It is only an outcome of the collection considered as a whole.
-The significance of the whole sequence can only by appreciated by Lucky once she is equipped to represent goals.
+Note that representing goal-directed action as we have characterised it does not require representing representations.
+It only requires representing outcomes as functions of bodily movements. 
+(The term `goal' is sometimes used, perhaps improperly, to refer to intentions or other representations;
+but as we use the term, a goal is simply an outcome to which an action is directed.)
 
-The first principle, then, is that object-directed actions form larger units of action which are directed to goals.
-This second principle is sufficient to explain some cases of imitative learning, which can be defined as attempting to reproduce the actions necessary to achieve a goal \citep[]{en_1317}.
+The first principle, then, is that bodily movements form units which are directed to goals.
+This first principle is sufficient to explain some cases of imitative learning, which can be defined as attempting to reproduce the actions necessary to achieve a goal \citep[]{en_1317}.
+
+\subsection{Second principle}
+Before describing the second principle we need to introduce two concepts.
+An agent’s \textit{field} at any given time is a set of objects.
+Whether an object falls within the agent’s field is determined by spatial and physical constraints such as proximity and lighting.
+The agent’s orientation and posture will also play a role in determining which objects fall into an agent’s field, as will eye direction in some species.
+To fall within an agent’s field, there must be no opaque barriers between the agent and the object, unless the object was recently in motion and not behind a barrier.
+These constraints ensure that objects which fall into an agent’s field are approximately those the agent can perceive.\footnote{ 	A variety of research in spatial and motor cognition suggests that adult humans (and perhaps others) not only compute other agents’ fields but also spontaneously locate objects within the spatial perspectives of other agents \citep[e.g.][]{en_1700, en_1701}.
+}
+
+Let us say that an agent is \textit{encountering} an object if it is in her field.
+The notion of encountering defines a relation between an agent and an object.
+Within limits, this notion of encountering can do some of the work that the concept of perception does.
+Encountering an object is like perceiving one to the extent that both notions involve a relation between agents and objects, both notions have approximately the same extension (someone perceives an object just if she encounters it), and both notions are bound up with action.
+But to track who encounters what does not require understanding perception or related psychological concepts.
+
+With these concepts in place, we can state the second principle: one can’t initiate goal-directed actions on an object unless one is encountering it.
+So for an object-directed action to occur it is not sufficient merely that the object is present; the agent must also be encountering the object.
+This principle is not true but it is a useful heuristic.
+
+What applications does the first principle have?
+Sometimes it is possible to exercise control over an agent’s behaviour by manipulating whether she encounters an object.
+A simple application of this would be keeping some food concealed from competitors close enough to grab it.
+Concealing the food prevents the competitor from encountering it and therefore from grabbing it.
+
+The hypothesis that an individual tracks what others encounter is also sufficient to explain success on some tests for level 1 perspective taking.
+\citet[]{en_1550} asked children to hide objects from an experimenter.
+The setup meant that the only hiding place left the objects in full view of the child herself.
+Even the youngest children tested, two-and-a-half-year-olds, could do this reliably.
+They were also able to say who could see an object in different conditions.
+The researchers concluded that children can think about seeing:
+%
+\begin{quote}
+“children could conceptually distinguish what they saw from what another person might see, [and] could think about what the other person saw rather than what they saw when asked to do so.” \citep[][p.\ 1210]{en_1550}
+\end{quote}
+%
+An alternative hypothesis consistent with these results is that children relied on knowledge of encountering rather than of seeing.
+On this hypothesis, children interpreted the experimenter as requesting them to place the object where she would not be encountering it.
+And they interpreted questions about who can see what in terms of who is encountering what.
+
+We are not claiming that the best explanation of these findings is the hypothesis that two-year-olds represent encountering rather than seeing; any such claim would require new experiments.
+But this hypothesis is not known to be less likely than any of the competing hypotheses, and it can explain children’s performance without ascribing them sophisticated psychological concepts such as seeing.
+
+The principle about encountering being required for object-directed action could also explain success on another test for level 1 perspective taking by Moll and Tomasello (\citeyear[]{en_1202}).
+An experimenter was positioned so that she could see only one of two objects (see Figure \vref{figure:level_1_perspective_taking}).
+She then asked the child, ‘Where is the other toy?
+Where is it?
+I cannot find it!’ and then ‘Can you give it to me?’ (2006: 607).
+Two-year-olds were significantly more likely to give the experimenter the toy she could not see.
+
+\begin{figure}
+\begin{center}
+\includegraphics{figure_level_1_perspective_taking.png}
+\caption{
+\label{figure:level_1_perspective_taking}
+Experimental setup for Moll \& Tomasello’s 2006 experiment on level 1 perspective taking.
+The experimenter was unable to see one of two objects clearly visible to the child.
+When the experimenter asked “I cannot find it!
+… Can you give it to me?”, two-year-olds reliably handed the hidden toy to the experimenter \citep[source: ][figure 1]{en_1202}.
+}
+\end{center}
+\end{figure}
+
+The authors suggest that children of this age “knew what another person could and could not see when that differed from what they saw” \citep[][p.\ 609]{en_1202}.
+An alternative possibility is that children were representing encountering rather than seeing.
+Children may interpret the experimenter’s saying “Where is it?
+I cannot find it!” as expressing an inability to act.
+Since encountering an object is a condition for acting on it, this would explain why they reliably handed the experimenter the object she was not encountering.\footnote{ 	Children in Moll and Tomasello’s experiment reliably work out which object the experimenter is requesting.
+This suggests some understanding of pragmatics which could be taken as evidence of further theory of mind cognition.
+The hypothesis about encountering is neutral on this issue.
+Its purpose is only to show that success on these tasks may involve representing encountering rather than seeing.}
+
+Perhaps it is a mistake to think of our hypothesis about encountering as an alternative to claims about seeing and perceiving.
+It is striking that philosophers have quite different views about what seeing involves—there are debates about whether it is possible to see things without seeing that something is the case \citep[e.g.][]{en_1671}, about whether and in what sense seeing is a representational notion \citep[]{en_1619, en_1705}, and about whether vision is intrinsically a source of knowledge \citep[]{en_1706}, for example.
+By contrast, developmental researchers say little about what they take understanding seeing or perceiving to consist in.
+Which notion of perception is the one children understand?
+Without an answer to this question it is unclear what it means to claim that children understand seeing or perceiving and unclear what predictions follow from this claim.
+The notion of encountering might be taken as a first step towards specifying what subjects understand when they understand perceiving.
+It is only a first step because it may need elaborating in order to characterise what subjects understand about perception (see footnote 9 above); but it is a vital step because it involves minimising reliance on commonsense psychological concepts in favour of a constructive account.
 

### 2012-04-04 mindreading/minimal theory of mind (new version)/minimal theory of mind.nv.tex
mTOM in progress

@@
 In this section we begin with someone, call her Lucky, capable  of representing nonintentional behaviour only and ask what more is needed for minimal theory of mind cognition.
-We describe Lucky’s progress with a series of five principles.
+We describe Lucky’s progress with a series of principles.
 The principles are constructed in such a way that it would be coherent to suppose that Lucky has the abilities codified by the first \textit{n} principles only.
@@
 \footnote{
-In standard false belief tasks, “[t]he subject is aware that he/she and another person [call him Maxi] witness a certain state of affairs x.
-Then, in the absence of the other person the subject witnesses an unexpected change in the state of affairs from x to y” \citep[][p.\ 106]{en_89}.
+In standard false belief tasks, `[t]he subject is aware that he/she and another person [call him Maxi] witness a certain state of affairs x.
+Then, in the absence of the other person the subject witnesses an unexpected change in the state of affairs from x to y' \citep[][p.\ 106]{en_89}.
 The test concerns whether the subject realises that Maxi will falsely believe x to obtain.
@@
 Within limits, this notion of encountering can do some of the work that the concept of perception does.
-Encountering an object is like perceiving one to the extent that both notions involve a relation between agents and objects, both notions have approximately the same extension (someone perceives an object just if she encounters it), and both notions are bound up with action.
-But to track who encounters what does not require understanding perception or related psychological concepts.
+Encountering an object is like perceiving one to the extent that both notions involve a relation between agents and objects, both notions have approximately the same extension (someone perceives an object just if she encounters it), and both notions are bound up with action, as we shall explain.
+%This is why representing encounterings and exploiting principles linking encountering to goal-directed action could enable Lucky to track perceptions.
 
-With these concepts in place, we can state the second principle: one can’t initiate goal-directed actions on an object unless one is encountering it.
-So for an object-directed action to occur it is not sufficient merely that the object is present; the agent must also be encountering the object.
+With these concepts in place, we can state the second principle: 
+one cannot goal-directedly act on an object unless one has encountered it.  
+More carefully,
+if an outcome involves a particular object
+and the agent has not encountered that object,
+then that outcome cannot be a goal of her actions.
+As with the other principles, this is plainly not a fact.
+What matters is just that, in a limited but useful range of cases, the principles collectively enable lucky to track perceptions and goal-directed actions.
+
+The second principle has many applications.
+Someone who is aware of this principle can be motivated to prevent others from encountering her food even when they are not in a position to steal it immediately.
+Take scrub-jays.
+When choosing where to cache food in the presence of a competitor they prefer far to near, darker to lighter, and occluded to in-view locations \citep[]{en_1451, en_1452}.
+These scrub-jays may be trying to hinder future thefts, for these behaviours are not found when caching non-food items \citep[]{en_1419} or when caching in the presence of a partner \citep[][p.\ 514]{en_1418, en_1408}.
+Clayton and Emery note that `[s]uch skills suggest visual perspective taking—computing what another can or cannot see' \citep[]{en_1451}.
+That is to say, they ascribed to scrub jays the concept of seeing.
+Another possibility is that scrub-jays compute encountering rather than seeing.
+Perhaps scrub-jays take having encountered food to be a condition for performing goal-directed actions targeting that food.
+If so they may be trying to minimize the chance that others will encounter their food in order to prevent future theft.
+Of course we  are not claiming that this is the actual explanation of these findings.  
+Our question is not about what scrub-jays actually represent but about what someone could represent that would enable her to track perceptions.
+Our claim is that the ability to track perceptions in the ways scrub-jays do could involve representing encounterings only.
+
+%For another application of the second principle, consider two-year-old children's 
+% Level 1 perspective abilities. 
+%In an experiment designed by by Moll and Tomasello (\citeyear[]{en_1202}), an experimenter was positioned so that she could see only one of two objects (see Figure \vref{figure:level_1_perspective_taking}).
+She then asked the child, ‘Where is the other toy?
+Where is it?
+I cannot find it!’ and then ‘Can you give it to me?’ (2006: 607).
+Two-year-olds were significantly more likely to give the experimenter the toy she could not see.
+%%
+%\begin{figure}
+%\begin{center}
+%\includegraphics{figure_level_1_perspective_taking.png}
+%\caption{
+%\label{figure:level_1_perspective_taking}
+%Experimental setup for Moll \& Tomasello’s 2006 experiment on level 1 perspective taking.
+The experimenter was unable to see one of two objects clearly visible to the child.
+When the experimenter asked `I cannot find it!
+… Can you give it to me?', two-year-olds reliably handed the hidden toy to the experimenter \citep[source: ][figure 1]{en_1202}.
+%}
+%\end{center}
+%\end{figure}
+%%
+%The authors suggest that children of this age `knew what another person could and could not see when that differed from what they saw'.\footnote{
+%	\citet[][p.\ 609]{en_1202}.  
+%	Compare \citet[][p.\ 1210]{en_1550} who take subjects' success on a level 1 perspective taking task to show that they
+%`could conceptually distinguish what they saw from what another person might see, [and] could think about what the other person saw rather than what they saw when asked to do so.'
+%}
+%An alternative possibility is that children were representing encountering rather than seeing.
+Children may interpret the experimenter’s saying `Where is it?
+I cannot find it!'\ as expressing an inability to act.
+If they also treat having encountered an object as a condition for acting on it, this would explain why they reliably handed the experimenter the object she was not encountering.\footnote{ 	Children in Moll and Tomasello’s experiment reliably work out that the experimenter is requesting something.
+This suggests some understanding of pragmatics which could be taken as evidence of much richer theory of mind cognition.
+The hypothesis about encountering is neutral on this issue.
+Its purpose is only to show that success on these tasks may involve representing encounterings rather than seeings or perceptions.
+%}
+
+For another application of the second principle, consider Hare, Call and Tomasello’s finding that chimpanzees reliably adopt strategies which are appropriate given what dominant competitors know about the locations of food \citep[]{en_1545}.
+In their `uninformed' condition, a subordinate chimpanzee observed a food item being hidden while a dominant competitor’s view was blocked (see Figure \vref{figure:hare_food_perception}).
+In this condition subordinates chose to approach the food significantly more often than in a control condition where the dominant competitor saw the food being hidden.
+This indicates that the subordinate chimpanzees were at least indirectly sensitive to facts about what the dominants had perceived.
+Several explanations of this finding have already been suggested \citep[]{Call:2008di, en_1551, Suddendorf:2003co}.
+A further possibility is that subordinate chimpanzees are aware that the dominant chimpanzee has not encountered the food and take encountering the food to be a condition for the dominant to act with the goal of recovering it.
+That would enable them to predict that the subordinate will not be able to retrieve this food in the misinformed condition.
+
+\begin{figure}
+\begin{center}
+\includegraphics{figure_hare_food_perception.png}
+\caption{
+	\label{figure:hare_food_perception}
+	A subordinate observes as food is placed.
+The subordinate can also see the dominant.
+There are three conditions: control—the dominant sees food being placed; `uninformed'—the dominant’s view is blocked while the food is placed; and `misinformed'—the dominant sees the food being placed then  has their view blocked while it is moved.
+	 \citep[Source:][pp.\ 142, fig. 1]{en_1545}
+}
+\end{center}
+\end{figure}
+
+In short, 
+one way to explain abilities to track others' perceptions  is to represent perceptions as such.
+But another way to track perceptions is to represent encounterings and to suppose (as the second principle states) that goal-directed actions involving an object are only possible when the agent has encountered that object.
+
+So far we have been taking for granted that representing encounterings is different from representing perceptions as such.
+Why accept this claim? 
+It is striking that philosophers have quite different views about what seeing involves—there are debates about whether it is possible to see things without seeing that something is the case \citep[e.g.][]{en_1671}, about whether and in what sense seeing is a representational notion \citep[]{en_1619, en_1705}, and about whether vision is intrinsically a source of knowledge \citep[]{en_1706}, for example.
+There is at least as much uncertainty concerning what might be involved in representing perceptions as such. 
+So it is not possible for us to say with certainty what it is to represent perceptions as such.  
+Even so, we can be sure that there are some differences between encountering and perception.
+If perceptions are representations, then representing perceptions as such plausibly involves representing representations.
+Since encounterings are relations not representations (by definition), representing encounterings will  differ from representing perceptions in that only the latter involves representing representations.
+But, as mentioned, not everyone agrees that perceptions are representations.
+We should therefore consider a bundle of features.
+Perception constitutively involves appearances, modalities or the possibility of illusion or is constitutively linked to reasons, knowledge or informational states.
+It is plausible that representing perceptions as such involves understanding something about some of these things.
+Since encountering does not constitutively involve any of these (by definition), 
+representing encounterings may differ from representing perceptions in that it does not require any kind of sensitivity to constitutive links with reasons, knowledge or informational states and nor does it require understanding  anything about appearances, modalities or the possibility of illusion.
+
+\subsection{Third principle}
+
+At this point we switch our attention from conditions on the \textit{occurrence} of goal-directed actions to conditions on their \textit{success}.
+Such conditions are more stringent.
+Where a goal specifies a particular object, it’s often not enough to have encountered it.
+To succeed you must often also have encountered it \textit{in its current location} and to have done so \textit{on your most recent encounter}.
+(There are exceptions, of course.)
+
+We can capture this condition and more by introducing a new notion, \textit{registration}.
+Registration is a relation between an individual, an object and a location which will be implicitly defined by principles linking it to encountering and action.
+(Where these principles conflict, principles mentioned later trump principles mentioned earlier.)  The first of these principles is that an individual registers an object at a location if she most recently encountered it at that location.
+As is already clear from this principle, registration is like belief in that it has a correctness condition which may not obtain: a registration is correct when the object is in the location.
+Since representing registration brings no benefit for Lucky independently of its connection to action, this principle cannot stand alone in our sequence; it is only the first half of the third principle.
+
+The second half of the third principle states that correct registration is a condition of successful action.
+More precisely, in order to successfully perform a goal-directed action with a goal that specifies a particular object, the agent must correctly register that object.%
+\footnote{ 	
+Some commentators suggested that this principle requires an ability to remember particular encounterings and their temporal order.
+There is some evidence that some nonhumans possess this ability \citep[e.g.][]{en_1714, en_1715}, but our proposal does not depend on this.
+To track another agent’s most recent encounter, it is not necessary to remember all encounters and their temporal order; forgetting all but the most recent will do.
+}  
+This principle can be applied in two directions.
+In one direction, it licenses Lucky to predict that a competitor who does not have a correct registration of an object will not be successful in performing actions whose goals specify that object.
+In the other direction, it allows Lucky, on the basis of observing a successful goal-directed action, to infer that the agent has correctly registered the location of an object.%
+\footnote{ 	
+The ‘other direction’ is required (in conjunction with the fifth principle, below) for explaining infants’ success on false belief tasks where information about another’s beliefs is provided not by what she sees but by what she does \citep[as in][]{en_1824}.
+}
+So the principle not only extends Lucky’s ability to predict actions but also her ability to detect what someone registers.
+
+The correctness of someone’s registrations can be manipulated in their absence by moving or destroying objects they have registered.
+So with theory of mind cognition partially characterized by the third principle, Lucky can intentionally prevent others from stealing a food item they have already encountered simply by moving it in their absence.
+
+For an application of this principle, consider Hare, Call and Tomasello’s (\citeyear[]{en_1545}) experiment again.
+In a further condition, the `misinformed' condition, a subordinate observer watched as a dominant competitor saw food being hidden.
+The subordinate continued to watch as the competitor’s view was blocked and the food moved.
+In this case the competitor has encountered the food but does not correctly register it.
+The subordinate observers went for the food more often in this condition than in a control condition where the dominant saw the food being moved.
+This cannot be explained in terms of the third principle.
+That principle involved taking encountering an object to be a condition on acting on it.
+This condition is met: the competitor \textit{has} encountered the food.
+To explain why the observer goes for the food that has been moved, we need to appeal to the third principle—to correct registration as a condition on success.
+It is possible that the subordinate observer realized that the dominant competitor last encountered the food in a location other than its current location.
+Suppose the observer also understood that correct registration is a condition on successful goal-directed action.
+Then the observer could predict that the competitor would not succeed in retrieving the food.
+This could explain why the observer approached the food in the `misinformed' condition.
+
+Another application of the third principle is to scrub-jays who strategically re-cache food depending on who saw what.
+In one experiment, scrub-jays cached some food in the presence of Competitor A and then cached more food in the presence of Competitor B.
+Later they had an opportunity to recover food in the presence of Competitor A.  
+The scrub-jays preferred to recover and re-cache food cached in the presence of Competitor A, leaving untouched food cached in the presence of the absent competitor \citep[][pp.\ 517-9]{en_1418}.
+This strategy reduces the chances of Competitor A knowing where any of the scrub-jays’ food is.
+What might explain this sensitivity to who saw what in choosing which food to recover?
+Clayton and colleagues postulate a capacity to attribute knowledge or `informational states' \citep[]{en_1418}.
+They are opposed by Penn and colleagues who postulate that scrub-jays are not ascribing knowledge but using rules such as `Try to re-cache food in a site different from the one where it was cached when the competitor was present' \citep[]{en_1417, Penn:2007ey}.
+As an alternative to both proposals, the scrub-jays’ sensitivity could be explained in terms of registration.
+Suppose scrub-jays understand that correct registration is a condition on successful goal-directed action.
+Then they may be trying to prevent competitors from stealing their cached food by means of preventing them from correctly registering its location.
+As before, the motivation for our minimal theory of mind hypothesis is that it avoids the onerous information processing burden entailed by supposing that scrub-jays understand knowledge as such.
+
+A third application of this principle is to infant pointing.
+Liszkowski and colleagues showed that 12- and 18-month-olds point in order to provide relevant information to adults about the locations of objects \citep[]{en_1093}.
+In this experiment, infants watch as an adult uses an object in some task.
+Then the adult appears to accidentally misplace it.
+Later, when the adult visibly needs that object to complete the task again, infants reliably point to it.
+Apparently infants point not in order to get the object or to share interest in it, but to enable the adult to complete a task.
+The authors suggest that this could be explained either by supposing that infants understand what the adult does not know, or that they understand what the adult is not attending to \citep[][p.\ 185]{en_1093}.
+As knowledge and attention are both complex psychological notions, these suggestions raise the hard question of what children might understand of them.
+Registration suggests a third possible explanation.
+Maybe these infants understand that correct registration is necessary for successful goal-directed action and that pointing is a way of generating correct registration.
+(None of these explanations bear on what is perhaps most interesting about these findings, that infants care to inform others and do so spontaneously.
+Minimal theory of mind as described here is at most a small part of a larger story.)
+
+\section{***BIN***}
+
+With these concepts in place, we can state the second principle: one can’t complete a goal-directed action on an object unless one is encountering it.
+So for a goal-directed action to succeed it is not sufficient merely that the object is present; the agent must also be encountering the object.
 This principle is not true but it is a useful heuristic.
 
-What applications does the first principle have?
-Sometimes it is possible to exercise control over an agent’s behaviour by manipulating whether she encounters an object.
-A simple application of this would be keeping some food concealed from competitors close enough to grab it.
-Concealing the food prevents the competitor from encountering it and therefore from grabbing it.
-
-The hypothesis that an individual tracks what others encounter is also sufficient to explain success on some tests for level 1 perspective taking.
+The hypothesis that an individual tracks what others encounter is sufficient to explain success on some tests for level 1 perspective taking.
 \citet[]{en_1550} asked children to hide objects from an experimenter.
@@
 \begin{quote}
-“children could conceptually distinguish what they saw from what another person might see, [and] could think about what the other person saw rather than what they saw when asked to do so.” \citep[][p.\ 1210]{en_1550}
+`children could conceptually distinguish what they saw from what another person might see, [and] could think about what the other person saw rather than what they saw when asked to do so.'\footnote{
+\citet[][p.\ 1210]{en_1550}.
+Compare \citet[][p.\ 609]{en_1202} who take subjects' success on a level 1 perspective taking task to show that they
+`knew what another person could and could not see when that differed from what they saw'.
+}
 \end{quote}
@@
 And they interpreted questions about who can see what in terms of who is encountering what.
-
-We are not claiming that the best explanation of these findings is the hypothesis that two-year-olds represent encountering rather than seeing; any such claim would require new experiments.
-But this hypothesis is not known to be less likely than any of the competing hypotheses, and it can explain children’s performance without ascribing them sophisticated psychological concepts such as seeing.
-
-The principle about encountering being required for object-directed action could also explain success on another test for level 1 perspective taking by Moll and Tomasello (\citeyear[]{en_1202}).
-An experimenter was positioned so that she could see only one of two objects (see Figure \vref{figure:level_1_perspective_taking}).
-She then asked the child, ‘Where is the other toy?
-Where is it?
-I cannot find it!’ and then ‘Can you give it to me?’ (2006: 607).
-Two-year-olds were significantly more likely to give the experimenter the toy she could not see.
-
-\begin{figure}
-\begin{center}
-\includegraphics{figure_level_1_perspective_taking.png}
-\caption{
-\label{figure:level_1_perspective_taking}
-Experimental setup for Moll \& Tomasello’s 2006 experiment on level 1 perspective taking.
-The experimenter was unable to see one of two objects clearly visible to the child.
-When the experimenter asked “I cannot find it!
-… Can you give it to me?”, two-year-olds reliably handed the hidden toy to the experimenter \citep[source: ][figure 1]{en_1202}.
-}
-\end{center}
-\end{figure}
-
-The authors suggest that children of this age “knew what another person could and could not see when that differed from what they saw” \citep[][p.\ 609]{en_1202}.
-An alternative possibility is that children were representing encountering rather than seeing.
-Children may interpret the experimenter’s saying “Where is it?
-I cannot find it!” as expressing an inability to act.
-Since encountering an object is a condition for acting on it, this would explain why they reliably handed the experimenter the object she was not encountering.\footnote{ 	Children in Moll and Tomasello’s experiment reliably work out which object the experimenter is requesting.
-This suggests some understanding of pragmatics which could be taken as evidence of further theory of mind cognition.
-The hypothesis about encountering is neutral on this issue.
-Its purpose is only to show that success on these tasks may involve representing encountering rather than seeing.}
-
-Perhaps it is a mistake to think of our hypothesis about encountering as an alternative to claims about seeing and perceiving.
-It is striking that philosophers have quite different views about what seeing involves—there are debates about whether it is possible to see things without seeing that something is the case \citep[e.g.][]{en_1671}, about whether and in what sense seeing is a representational notion \citep[]{en_1619, en_1705}, and about whether vision is intrinsically a source of knowledge \citep[]{en_1706}, for example.
-By contrast, developmental researchers say little about what they take understanding seeing or perceiving to consist in.
-Which notion of perception is the one children understand?
-Without an answer to this question it is unclear what it means to claim that children understand seeing or perceiving and unclear what predictions follow from this claim.
-The notion of encountering might be taken as a first step towards specifying what subjects understand when they understand perceiving.
-It is only a first step because it may need elaborating in order to characterise what subjects understand about perception (see footnote 9 above); but it is a vital step because it involves minimising reliance on commonsense psychological concepts in favour of a constructive account.
+Of course we  are not claiming that this is the best explanation of these findings.  
+Our point is just that the ability to track perceptions measured by level 1 perspective taking tasks could involve representing encounterings only and need not involve representing perceptual states as such.
 

### 2012-04-04 intention motor representation with corrado sinigaglia/intention and motor action.ppr.r1.final.tex
working on new version of mtom
mTOM in progress

@@
 \def \iauthor {}
-%\def \iauthor {S. Butterfill* \& C. Sinigaglia**
+\def \iauthor {S. Butterfill* \& C. Sinigaglia**
+\\ 
+**Department of Philosophy, University of Warwick, UK
+\\ 
+***Dipartimento di Filosofia, Università degli Studi di Milano, Italia}
+\def \iemail{s.butterfill@warwick.ac.uk}
+%\date{}
+
+\input{$HOME/Documents/submissions/preamble_steve_paper2}
+%\author{}
+%\author {S. Butterfill* \& C. Sinigaglia**
 %\\ 
@@
 %***Dipartimento di Filosofia, Università degli Studi di Milano, Italia}
-%\def \iemail{s.butterfill@warwick.ac.uk}
-%\date{}
-
-\input{$HOME/Documents/submissions/preamble_steve_paper2}
-%\author{}
-\author {S. Butterfill* \& C. Sinigaglia**
-\\ 
-**Department of Philosophy, University of Warwick, UK
-\\ 
-***Dipartimento di Filosofia, Università degli Studi di Milano, Italia}
 \date{}
@@
 \footnote{
-\label{fn:campbell_location}
+\label{fn:campbell_location}%
 In fact, there may be another way of explaining non-accidental matching with respect to what is to be acted on due to \citet[pp.\ 36-8, 44-5, 48-57]{Campbell:2002ge}.
@@
 But if, as we conjecture, such deference is necessary for intentions to properly and reliably result in bodily movements, it may turn out that intending to act in the world depends on experiences which are made possible by motor representation.
-Much as on some views thought about objects depends on perceptual experience \citep[e.g.][]{Campbell:2002ge}, so also intending actions may depend on motor experience.
+Much as on some views thought about objects depends on perceptual experience \citep[e.g.][]{Campbell:2002ge}, so also intending actions may depend on motor experience.%
+\footnote{
+Heartfelt thanks to Chiara Brozzo, Silvano Zipoli Caiani, Naomi Eilan, Vittorio Gallese, Francesco Guala, \'Elisabeth Pacherie and Natalie Sebanz.
+Corado Sinigaglia was supported by EU grant TESIS, a Fellowship from the Institute of Philosophy (University of London) and a Fellowship from the Center for the Epistemology of Cognitive Sciences, Ecole Normale Superiéure de Lyon.
+}
 

### 2012-04-04 mindreading/minimal theory of mind (new version)/minimal theory of mind.nv.tex
mTOM in progress

@@
 \noindent
-What could someone represent that would enable her to track, at least within limits, perceptions, beliefs including false beliefs, and other propositional attitudes? 
+What could someone represent that would enable her to track, at least within limits, perceptions, knowledge states and beliefs including false beliefs? 
 An obvious possibility is that she might represent these very propositional attitudes as such.
@@
 \section{Introduction}
-What could someone represent that would enable her to track, at least within limits, perceptions, beliefs including false beliefs, and other propositional attitudes? 
+What could someone represent that would enable her to track, at least within limits, perceptions, knowledge states and beliefs including false beliefs? 
 One answer is obvious: she might track these things by virtue of representing them as such, that is, by representing perceptions, beliefs, and other propositional attitudes as such.
@@
 not what anyone does represent
-but what someone could represent that would enable her, at least within limits, to track perceptions, beliefs and other propositional attitudes.
+but what someone could represent that would enable her, at least within limits, to track perceptions, knowledge states and beliefs.
 The motivation for facing up to this question is, of course, partly empirical.
@@
 
-What do these infants and young children represent that enables them, within limits, to track others’ beliefs and other propositional attitudes?   
-The most straightforward answer would be to suppose that they represent perceptions, beliefs and other propositional attitudes as such \citep[e.g.][]{en_1138, en_1691}.  
+What do these infants and young children represent that enables them, within limits, to track others’ perceptions, knowledge states and beliefs?   
+The most straightforward answer would be to suppose that they represent perceptions, knowledge states and beliefs as such \citep[e.g.][]{en_1138, en_1691}.  
 But this answer faces several objections.
@@
 Here, then, is a second source of motivation for our question about what someone could represent that would enable her, within limits, to track perceptions and beliefs.
-As we have seen, there are significant if not decisive objections to the two best developed conjectures about infant theory of mind abilities, the conjecture that these involve representing beliefs and other propositional attitudes as such and the conjecture that these involve representing nonintentional behaviour only.
+As we have seen, there are significant if not decisive objections to the two best developed conjectures about infant theory of mind abilities, the conjecture that these involve representing perceptions, knowledge states and beliefs as such and the conjecture that these involve representing nonintentional behaviour only.
 These objections, while not decisive, justify exploring alternatives.
@@
 then the construction of a minimal theory of mind may be relevant to these debates.
-For the conjecture that minimal theory of mind explains chimpanzees' or scrub-jays' abilities to track perceptions or beliefs can be empirically distinguished from conjectures about representations of nonintentional behaviour only and from conjectures about representations of perceptions, beliefs and other propositional attitudes as such (as we explain in section *** below).
+For the conjecture that minimal theory of mind explains chimpanzees' or scrub-jays' abilities to track perceptions or beliefs can be empirically distinguished from conjectures about representations of nonintentional behaviour only and from conjectures about representations of perceptions, knowledge states and beliefs (as we explain in section *** below).
 
@@
 In short, 
-one way to explain abilities to track others' perceptions  is to represent perceptions as such.
-But another way to track perceptions is to represent encounterings and to suppose (as the second principle states) that goal-directed actions involving an object are only possible when the agent has encountered that object.
+abilities to track others' perceptions  may depend on representing perceptions as such.
+But another way to track perceptions would be to represent encounterings and to suppose (as the second principle states) that goal-directed actions involving an object are only possible when the agent has encountered that object.
 
@@
 Registration is a relation between an individual, an object and a location which will be implicitly defined by principles linking it to encountering and action.
-(Where these principles conflict, principles mentioned later trump principles mentioned earlier.)  The first of these principles is that an individual registers an object at a location if she most recently encountered it at that location.
+By design, applying all of these principles would sometimes lead to inconsistencies.  
+Where it would be impossible to apply all the principles consistently, consistency is to be achieved by having principles mentioned later trump principles mentioned earlier.
+The first  principle defining registration is that an individual registers an object at a location if and only if she most recently encountered it at that location.
 As is already clear from this principle, registration is like belief in that it has a correctness condition which may not obtain: a registration is correct when the object is in the location.
@@
 \footnote{ 	
-The ‘other direction’ is required (in conjunction with the fifth principle, below) for explaining infants’ success on false belief tasks where information about another’s beliefs is provided not by what she sees but by what she does \citep[as in][]{en_1824}.
+The ‘other direction’ is required (in conjunction with the fourth principle, below) for explaining infants’ success on false belief tasks where information about another’s beliefs is provided not by what she sees but by what she does \citep[as in][]{en_1824}.
 }
@@
 In this case the competitor has encountered the food but does not correctly register it.
-The subordinate observers went for the food more often in this condition than in a control condition where the dominant saw the food being moved.
-This cannot be explained in terms of the third principle.
+Subordinate observers went for the food more often in this condition than in a control condition where the dominant saw the food being moved.  
+This cannot be explained in terms of the second principle.
 That principle involved taking encountering an object to be a condition on acting on it.
 This condition is met: the competitor \textit{has} encountered the food.
-To explain why the observer goes for the food that has been moved, we need to appeal to the third principle—to correct registration as a condition on success.
+To explain why the subordinate observer goes for the food that has been moved, we need to appeal to the third principle—to correct registration as a condition on success.
 It is possible that the subordinate observer realized that the dominant competitor last encountered the food in a location other than its current location.
@@
 Then the observer could predict that the competitor would not succeed in retrieving the food.
-This could explain why the observer approached the food in the `misinformed' condition.
+This could explain why  subordinate observers  more often approach the food in the `misinformed' condition than in the control condition.
 
-Another application of the third principle is to scrub-jays who strategically re-cache food depending on who saw what.
+For another application of the third principle, consider scrub-jays who strategically re-cache food depending on who saw what.
 In one experiment, scrub-jays cached some food in the presence of Competitor A and then cached more food in the presence of Competitor B.
@@
 Then they may be trying to prevent competitors from stealing their cached food by means of preventing them from correctly registering its location.
-As before, the motivation for our minimal theory of mind hypothesis is that it avoids the onerous information processing burden entailed by supposing that scrub-jays understand knowledge as such.
 
-A third application of this principle is to infant pointing.
-Liszkowski and colleagues showed that 12- and 18-month-olds point in order to provide relevant information to adults about the locations of objects \citep[]{en_1093}.
-In this experiment, infants watch as an adult uses an object in some task.
+A third application of the third principle is to infant pointing.
+\citet{en_1093} show that 12- and 18-month-olds point in order to provide relevant information to adults about the locations of objects.
+In their experiment, infants watch as an adult uses an object in some task.
 Then the adult appears to accidentally misplace it.
@@
 As knowledge and attention are both complex psychological notions, these suggestions raise the hard question of what children might understand of them.
-Registration suggests a third possible explanation.
-Maybe these infants understand that correct registration is necessary for successful goal-directed action and that pointing is a way of generating correct registration.
+Another possible explanation involves registration.
+Maybe the infants understand that correct registration is necessary for successful goal-directed action and that pointing is a way of generating correct registration.
 (None of these explanations bear on what is perhaps most interesting about these findings, that infants care to inform others and do so spontaneously.
 Minimal theory of mind as described here is at most a small part of a larger story.)
+
+Note that have not argued for the truth of any hypothesis about what chimpanzees, scrub-jays or infants represent.
+As explained above, what motivates our enquiry is not that there is already evidence for hypotheses about particular subjects' minimal theory of mind cognition: it is that, given the limited evidence for competing hypotheses and the objections they face, new hypotheses are worth formulating and testing.
+And in later sections we will explain how hypotheses about minimal theory of mind generate testable predictions, enabling them to be empirically distinguished from competing hypotheses.
+For now our focus is the question of what a subject could represent that would enable her to track track perceptions, knowledge states and beliefs in ways measured by various experimental paradigms.
 

### 2012-04-06 mindreading/minimal theory of mind (new version)/minimal theory of mind.nv.tex
mtom redraft got to end

@@
 \input{$HOME/Documents/submissions/preamble_steve_paper2}
+
+%avoid overhang
+\tolerance=5000
 
@@
 theory of mind abilities from theory of mind cognition.  
-A  \textit{theory of mind ability} is an ability that exists in part because exercising it brings benefits obtaining which depends on exploiting or influencing facts about others’ mental states.
+A  \textit{theory of mind ability} 
+	\label{df:tom_ability}
+is an ability that exists in part because exercising it brings benefits obtaining which depends on exploiting or influencing facts about others’ mental states.
 To illustrate,
@@
 (This is not supposed to be a plausible, real-world example but only to illustrate what the definition requires.)
-An ability to \textit{track} perceptions or beliefs (say) is a theory of mind ability which involves exploiting or influencing facts about these states.
-By contrast, \textit{theory of mind cognition} involves representing mental states or processes as such.
+An ability to \textit{track} 
+	\label{df:track} 
+perceptions or beliefs (say) is a theory of mind ability which involves exploiting or influencing facts about these states.
+By contrast, \textit{theory of mind cognition} 
+	\label{df:tom_cognition}
+involves representing mental states or processes as such.
 And \textit{full-blown} theory of mind cognition involves  representing propositional attitudes such as beliefs, desires and intentions to construct reason-giving, causal explanations of action.  
@@
 
-The notion that some abilities to track perceptions or beliefs involve only on theory of mind cognition which does not involve representing perceptions or beliefs as such is not entirely novel.
+The notion that some abilities to track perceptions, knowledge states or beliefs involve only theory of mind cognition which does not involve representing perceptions or beliefs as such is not entirely novel.
 To mention only those we draw most directly on, Gomez (\citeyear[][p.\ 730]{en_1259}) has emphasized primitive intentional relations to objects established by gaze, O’Neil and Doherty have separately discussed a notion of engagement with objects (\citealp{en_1159, en_1140}), Call and Tomasello (\citeyear[][p.\ 58]{en_1669}) have suggested that chimpanzees track the `likely target' of others’ visual access and understand something about its effects on behaviour, and Whiten (\citeyear[]{en_1415, en_1416}) uses the notion of an `intervening variable' to explain primitive theory of mind notions.
@@
 \section{Motivation}
+\label{sec:motivation}
 Our question is theoretical: it concerns 
@@
 \section{More motivation}
+\label{sec:more_motivation} 
 Further motivation for our question comes from evidence for theory of mind abilities in young children and infants.
@@
 scrub-jays can selectively re-cache their food in ways that deprive competitors of knowledge of its location \citep{Clayton:2007fh}, and  chimpanzees can both select routes to approach food which conceal them from a competitor’s view \citep[]{en_1546} and also retrieve food using strategies that optimize their return given what a dominant competitor has seen \citep[]{en_1545}.
-There is debate about the cognitive underpinnings of these abilities (e.g.\ 
-	\citealp{povinelli:2004vonk, 
-			Penn:2007ey,
-			Tomasello:2005ce,
-			Call:2008di}). 
-If it is not yet known precisely what explains these abilities 
+There is debate about the cognitive underpinnings of these abilities. 
+Some researchers argue that they may involve representations of nonintentional behaviour only,\footnote{
+For example, on chimpanzees: 
+	\citet{povinelli:2004vonk}; 
+	\citet[pp.\ 364-5]{Vonk:2006cq}; and on scrub-jays and chimpanzees \citet{Penn:2007ey}.
+} 
+others that they involve representations of perceptions, knowledge states and other propositional attitudes.\footnote{
+For example, on chimpanzees: \citet{Tomasello:2005ce,
+			Call:2008di}; and on scrub-jays: 
+			\citet[p.\ 73]{Emery:2007ze}.
+}
+If these arguments are not yet decisive
 and if the available evidence does not already tightly constrain the space of admissible conjectures,
-then the construction of a minimal theory of mind may be relevant to these debates.
+then the construction of a minimal theory of mind may also be relevant to these debates.
 For the conjecture that minimal theory of mind explains chimpanzees' or scrub-jays' abilities to track perceptions or beliefs can be empirically distinguished from conjectures about representations of nonintentional behaviour only and from conjectures about representations of perceptions, knowledge states and beliefs (as we explain in section *** below).
@@
 %the conjecture that minimal theory of mind cognition underpins some abilities to track perceptions and belief in human adults and infants generates testable predictions that distinguish it from the alternatives.
-
-\section{The need for a constructive approach}
-Davidson ...
-
-%This is reflected in the dichotomy between behaviour reading (P&Vonk) \& full-blown ToM (Baillargeon review).
-
-Davidson might be right about ordinary, commonsense psychology.
-
-So a constructive approach is needed. ***
 
@@
 When choosing where to cache food in the presence of a competitor they prefer far to near, darker to lighter, and occluded to in-view locations \citep[]{en_1451, en_1452}.
-These scrub-jays may be trying to hinder future thefts, for these behaviours are not found when caching non-food items \citep[]{en_1419} or when caching in the presence of a partner \citep[][p.\ 514]{en_1418, en_1408}.
+These scrub-jays may be trying to hinder future thefts, for these behaviours are not found when caching non-food items \citep[]{en_1419} or when caching in the presence of a partner \citep[][p.\ 514]{Clayton:2007fh, Emery:2007ze}.
 Clayton and Emery note that `[s]uch skills suggest visual perspective taking—computing what another can or cannot see' \citep[]{en_1451}.
@@
 Later they had an opportunity to recover food in the presence of Competitor A.  
-The scrub-jays preferred to recover and re-cache food cached in the presence of Competitor A, leaving untouched food cached in the presence of the absent competitor \citep[][pp.\ 517-9]{en_1418}.
+The scrub-jays preferred to recover and re-cache food cached in the presence of Competitor A, leaving untouched food cached in the presence of the absent competitor \citep[][pp.\ 517-9]{Clayton:2007fh}.
 This strategy reduces the chances of Competitor A knowing where any of the scrub-jays’ food is.
 What might explain this sensitivity to who saw what in choosing which food to recover?
-Clayton and colleagues postulate a capacity to attribute knowledge or `informational states' \citep[]{en_1418}.
+Clayton and colleagues postulate a capacity to attribute knowledge or `informational states' \citep[]{Clayton:2007fh}.
 They are opposed by Penn and colleagues who postulate that scrub-jays are not ascribing knowledge but using rules such as `Try to re-cache food in a site different from the one where it was cached when the competitor was present' \citep[]{en_1417, Penn:2007ey}.
@@
 
-\section{***BIN***}
+\subsection{Fourth principle}
 
-With these concepts in place, we can state the second principle: one can’t complete a goal-directed action on an object unless one is encountering it.
-So for a goal-directed action to succeed it is not sufficient merely that the object is present; the agent must also be encountering the object.
-This principle is not true but it is a useful heuristic.
+So far Lucky thinks of correct registration as a condition for the success of goal-directed action.
+This does not tell her anything about what happens if the condition is not met.
+In particular it tells her nothing about how an agent will act when she registers an object incorrectly.
+The fourth principle involves a switch from thinking of registration as a success condition to thinking of it as a causal factor.
+This principle states that when an agent performs a goal-directed action with a goal that specifies a particular object, the agent will act as if the object were in the location she registers it in.
 
-The hypothesis that an individual tracks what others encounter is sufficient to explain success on some tests for level 1 perspective taking.
-\citet[]{en_1550} asked children to hide objects from an experimenter.
-The setup meant that the only hiding place left the objects in full view of the child herself.
-Even the youngest children tested, two-and-a-half-year-olds, could do this reliably.
-They were also able to say who could see an object in different conditions.
-The researchers concluded that children can think about seeing:
+Now that Lucky understands registration as a factor influencing action it can serve her as a proxy for false belief.
+Just as, in a limited but useful range of cases, you can track food sources' toxicities by representing their odours
+%, objects’ masses by measuring their weights (within limits) 
+and prospective sexual partners’ virtues by representing their plumage, so also you can track beliefs by representing registrations.
+
+Applications of the fourth principle therefore include Onishi and Baillargeon’s (\citeyear[]{en_1092}) false belief task.
+Subjects are shown an adult observer who is present while a piece of melon is placed in one box.
+In the critical condition, the adult observer is then absent while the melon moves to another box.
+Comparative looking times indicate that the subjects, who are 14-month-old infants, expect that the adult will reach into the box not containing the melon.\footnote{ 	This finding is supported by a growing body of related research \citep[including][]{en_1666, en_1690, en_1691, en_1208, en_1261}.}  The authors explain this finding by hypothesizing that the infants are ascribing beliefs about the melon’s location to the adults \citep[][p.\ 257]{en_1092}.
+Alternatively, the findings could be explained on the hypothesis that they are tracking registration as a cause of action.
+
+\subsection{Extensions and variations}
+With the fourth principle we completed the construction of a minimal theory of mind capable of underwriting success on some false belief tasks.  
+We stop here because false belief tasks are often taken to be an acid test for theory of mind.
+But of course additonal principles can be added accommodate further theory of mind abilities.  
+For instance, further principles might extend the definition of registration.  
+Registration was been defined as a relation between agents, objects and locations. 
+This definition could be extended to include  other types of property as relata in addition to locations.  
+Further notions, such as a relational proxy for tracking desires, could also be added.  
+These and other modifications would enable hypotheses about minimal theory of mind cognition  to explain a wider range of theory of mind abilities.%
+\footnote{
+Infants and two-year-olds' theory of mind abilities involve more than tracking false beliefs about location.
+To illustrate, \citet{en_1691} show that infants can use communicated information in predicting actions based on false belief, 
+\citet{he:2011_false}  show that 2.5-year-olds can succeed on false belief tasks involving unexpected contents rather than location properties, 
+and  \citet{en_1795}  show that 18-month olds are sensitive to inferences that others are likely to make. 
+}
+
+Variations on the principles are also possible.
+To illustrate, scrub-jays’ caching strategies do not obviously call for an ability to track relations to particular food items: their ultimate aim probably isn’t to prevent others pilfering particular worms but to prevent them from pilfering any worms at all.
+Accordingly scrub-jays’ protective caching strategies may involve tracking relations between agents, locations and food types rather than particular objects.
+In general it is possible that there is variation across species in what individuals represent that enables them to track perceptions, knowledge states and beliefs. 
+If we attempted to characterise theory of mind cognition using only adult human commonsense psychological notions of perception, knowledge and belief, it would be hard to make systematic  sense of the possibility of such variation.
+These commonsense psychological notions are not easy to take apart because they resemble the decision-theoretic notion of expected utility in this respect: they are characterised in part by their roles in a system of explanation \citep{Davidson:1985qg,Davidson:1999ju}.
+One virtue of the minimal theory of mind construction is that it enables us to make systematic sense of the possibility of variations. 
+%Further enhancements will be needed should their general sensitivity to time and episodes also feature in their theory of mind cognition \citep[see][]{en_1409}.
+
+\subsection{But is it theory of mind cognition?}
+The term ‘theory of mind’ has been used in several apparently distinct ways \citep[]{en_1415}.  
+On some definitions, all theory of mind cognition involves metarepresentation.\footnote{
+See \citet{en_81}.  This view can seem obviously correct if it is assumed that  mental states are all representations.
+However, there reasons to doubt this assumption \citep[see, for instance, ][]{Campbell:2002ge}.
+}  
+Minimal theory of mind involves representing goals to which actions are directed, encounterings and registrations, none of which are are representations. 
+So if we accepted this definition, what we have constructed would not count as a form of theory of mind cognition just because no metarepresentation is involved.
+
+On another influential definition, theory of mind cognition begins when subjects ascribe states which function as variables intervening between environmental or behavioural inputs and behavioural outputs, and which play some roles characteristic of mental states (\citealp{en_1416}; \citealp[][p.\ 732]{Penn:2007ey}).
+On this definition, the endpoint in our construction (but no earlier point) does count as theory of mind cognition because registrations are intermediate variables and play a subset of the causal roles characteristic of belief.
+For registrations, like beliefs, are characterised by principles generalising across all goal-directed actions,
+can be assigned correctness conditions
+and causally influence actions.
+(Of course they also differ from beliefs in many ways:
+they are not propositional attitudes,  they cannot involve reference to never-existent objects and they are not subject to the norms that arguably characterise belief.)
+We use the label ‘minimal theory of mind’ because, relative to this definition, the construction describes a minimally elaborate form of theory of mind cognition,
+one that does not the entail cognitive and conceptual demands associated with representing perceptions, knowledge states and beliefs as such  
+but is capable of grounding theory of mind abilities that generalise across goal-directed actions and are sufficient for systematic success on some false belief tasks.
+
+In what follows we show that minimal theory of mind cognition has clear signature limits.  These limits generate predictions capable of distinguishing hypotheses about minimal theory of mind both 
+	from hypotheses  about  representations of nonintentional behaviour only 
+	and also 
+	from  hypotheses about full-blown theory of mind cognition.
+
+\section{Limits: how to distinguish minimal from full-blown theory of mind cognition}
+\label{sec:limits1}
+How could we distinguish minimal from full-blown theory of mind cognition experimentally?
+The point of minimal theory of mind is to enable agents to fake it—that is, to act as if they were reasoning about propositional attitudes, within limits.
+Where a task goes beyond these limits, we can be sure an agent is not using minimal theory of mind only.
+
+Some limits on minimal theory of mind cognition arise from the fact that the theory makes use of objects and their relations to agents, rather than representations of objects, to predict others’ behaviours.
+This means that false beliefs involving quantification or identity cannot be tracked by representing registrations.
+To see why not, consider the following inference:
 %
 \begin{quote}
-`children could conceptually distinguish what they saw from what another person might see, [and] could think about what the other person saw rather than what they saw when asked to do so.'\footnote{
-\citet[][p.\ 1210]{en_1550}.
-Compare \citet[][p.\ 609]{en_1202} who take subjects' success on a level 1 perspective taking task to show that they
-`knew what another person could and could not see when that differed from what they saw'.
-}
+(1) Mitch believes that Charly is in Baltimore.
+
+(2) Charly is Samantha.
+
+Therefore:
+
+(3) Mitch believes that Samantha is in Baltimore.
 \end{quote}
 %
-An alternative hypothesis consistent with these results is that children relied on knowledge of encountering rather than of seeing.
-On this hypothesis, children interpreted the experimenter as requesting them to place the object where she would not be encountering it.
-And they interpreted questions about who can see what in terms of who is encountering what.
-Of course we  are not claiming that this is the best explanation of these findings.  
-Our point is just that the ability to track perceptions measured by level 1 perspective taking tasks could involve representing encounterings only and need not involve representing perceptual states as such.
+On almost any account of belief, this inference is not valid
+\citep[][pp.\ 214-5]{frege:1948_sense}.
+Its central role in a popular film \citep[]{en_1793} indicates that human adults typically appreciate that this inference is not valid.
+Contrast the above inference with the corresponding inference in the case of registration:
+%
+\begin{quote}
+(1$'$) Mitch registers \(<\)Charly, Baltimore\(>\)
+
+(2) Charly is Samantha.
+
+Therefore:
+
+(3$'$) Mitch registers \(<\)Samantha, Baltimore\(>\)
+\end{quote}
+%
+This inference from (1$'$) and (2) to (3$'$) is logically valid.
+It is valid because registration is a relation to objects.
+We can compare registration with other relations like being left of something.
+If Charly is Samantha (whether you know it or not), then anyone who is left of Charly is left of Samantha; similarly for registering Charly’s location.
+
+This formal difference between belief and registration entails a limit on minimal theory of mind cognition.
+Consider Lucky who tracks beliefs by means of representing registrations only and is unable to represent beliefs.
+Lucky should have no problem predicting actions based on false beliefs about the locations of objects but she should encounter difficulties in predicting actions based on beliefs essentially involving mistakes about identity.
+In particular, Lucky should not be able to understand why, when Mitch registers \(<\)Charly, Baltimore\(>\), he continues searching for Samantha.\footnote{ 	This assumes that Lucky herself knows that Charly is Samantha.
+To ease exposition we assume throughout that Lucky has no false beliefs involving identity.}  For to register \(<\)Charly, Baltimore\(>\) is the same thing as registering \(<\)Samantha, Baltimore\(>\).
+And Lucky should be equally at a loss when those she observes mistakenly believe that two distinct people are identical.
+By contrast, subjects who can represent beliefs as such should have no special problem with false beliefs essentially involving identity.
+This is how mistakes about the identities of objects can be used to distinguish minimal from full-blown theory of mind cognition.%
+\footnote{ 	
+Related points about quantification entail further testable distinctions.
+For instance, minimal theory of mind should make it impossible to track beliefs whose contents essentially involve \textit{most} objects having a certain property.
+It is beyond the scope of this paper to elaborate on these predictions.
+}
+
+How could this be exploited experimentally?
+One paradigm suitable for infants would involve a puppet which has a concave face and body so that,  unusually, it looks very different when seen from different angles.
+Viewed from one angle, the puppet looks like one person (Charly); viewed from another angle, the puppet looks like another person (Samantha); and viewed head-on both aspects of its appearance become visible simultaneously.
+Crucially, it is possible to see either aspect without having reason to suppose that the puppet would look so different from another angle.
+Subjects sit at a table opposite the protagonist.
+There is a screen in the middle of the table which blocks each person’s view of the other side of the table.
+In the critical condition, the puppet is initially on the subjects’ side of the screen (see Figure \vref{figure:charlie_samantha_experiment}).
+In the first scene, the puppet emerges from the screen so that both observers can see it.
+Only one aspect of the puppet is revealed to each; the protagonist sees the Charly-aspect, the subject sees the other aspect.
+In the second scene the puppet returns to the subjects’ side of the screen; she then rotates.
+The subject but not the protagonist is thereby shown that the face and body have two aspects.
+Finally, in scene three, the puppet emerges from the screen.
+The puppet again reveals only one aspect of its face and body to each observer but this time the protagonist sees the Samantha-aspect and the subject sees the other aspect.
+At this point the puppet leaves the stage altogether.
+Subjects can see that there is nothing on their side of screen, but the protagonist cannot see this.
+Critically, this means that there is reason for the  protagonist to expect there still to be a puppet (Charly) behind the screen.  
+So when, finally, the protagonist reaches around the screen, this action makes sense:
+	though there is actually nothing behind the screen, the protagonist has reason to expect there is a puppet.
+But for subjects to appreciate that the protagonist has this expectation, and to make sense of her reaching action, they would have to ascribe a false belief about identity.%
+\footnote{
+	To ensure this, the design must be such that subjects cannot readily categorise the puppet’s two appearances and such that subjects do not use verbal labels for the two aspects.
+Otherwise it would be possible for subjects to infer the protagonist’s expectation by reasoning about types of object.
+}  
+Subjects' lack of surprise at the protagonist's reaching around the screen (as compared to  a control condition which differs from this one only in that the protagonist manifestly knows that the two aspects of the puppet’s face and body are aspects of a single puppet) would therefore be evidence that they can track false beliefs about identity.
+This would be evidence for full-blown, not minimal, theory of mind.
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=\textwidth]{figure_charlie_samantha_experiment.png}
+\caption{
+\label{figure:charlie_samantha_experiment}
+	How to test whether infants can ascribe false beliefs about identity.
+The subject is in the foreground with only the back of  her head visible; the protagonist sits opposite, facing the subject.
+A screen conceals some of the action from the protagonist.
+}
+\end{center}
+\end{figure}
+
+So far we have discussed mistaken beliefs about identity where the truth of the belief would require there to be more individuals in the world (Charly and Samantha; a teacher and an assassin) than there actually are.
+The converse is also possible \citep[see, e.g.,][]{en_1794}.
+Here the mistake is to have beliefs whose truth would require there to be fewer individuals in the world than there actually are.
+The main condition of a possible experiment is outlined in Figure \vref{figure:identity_balls_experiment}.
+In this experiment there are two perceptually indistinguishable balls.
+Whereas subjects can see both balls simultaneously, a screen ensures that the protagonist never sees both at once.
+The movements of the two balls are timed in such a way that what the protagonist sees indicates the existence of a single ball.%
+\footnote{ 	
+Song and Baillargeon (\citeyear[]{en_1823}) provide evidence that 15-month-old infants are able to predict actions on the basis of how things appear to observers who are ignorant of their true nature.
+\citet{en_1822} show that chimpanzees also make inferences about the appearances of things in predicting behaviours.
+These and other findings \citep[e.g.][]{en_1820} suggest that infants and perhaps others are generally sensitive to facts about the ways things look in tracking others’ beliefs.
+}
+When the protagonist sees a ball appear from behind the screen and leave the scene (panels 5-7 in Figure 5), this provides a reason for the protagonist to believe that there is nothing behind the screen.
+Infants could only be sensitive to this belief if they were able to track beliefs involving identity.
+Sensitivity to this belief might be manifested by comparing this condition with a modified condition in which the protagonist  manifestly knows that there are two balls.
+If subjects show more surprise when the protagonist reaches around the screen (say) in the original condition than in the modified condition, this too would be evidence for full-blown over minimal theory of mind cognition.
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=\textwidth]{figure_identity_balls_experiment.png}
+\caption{
+\label{figure:identity_balls_experiment}
+	How to test whether infants can ascribe false beliefs about identity where the mistake involves belief whose truth would entail that there were fewer objects than there in fact are.
+As in Figure \ref{figure:charlie_samantha_experiment}, the subject is in the foreground with only the back of  her head visible; the protagonist sits opposite, facing the subject.
+A screen conceals some of the action from the protagonist.
+}
+\end{center}
+\end{figure}
+
+Given that this analysis, it is notable that one study already claims to have found positive evidence that infants can ascribe beliefs about identity. 
+Scott and Baillargeon (\citeyear{en_1690}) have already conducted experiments which aim to show that 18-month-old infants are able to represent beliefs involving mistakes about `which particular object \textit{token}' another is facing rather than only about `which \textit{type} of object' is present (p.\ 1179).
+However, while their aim is clearly relevant, we shall show that these experiments do not succeed in testing whether infants ascribe beliefs about identity.
+In essence, their paradigm involves two types of object, a divisible penguin which affords hiding a key and an indivisible penguin which lacks this affordance.
+The divisible penguin has two states, divided and whole.
+When whole it is visually indistinguishable from the indivisible penguin which does not afford hiding a key.
+Infants sit opposite a protagonist.
+Familiarisation trials provide information that the protagonist can expect there to be two penguins in the scene at all times, one of each type; these trials also provide information that the protagonist can expect the divisible penguin to appear in its divided state.
+In the critical condition, the scene contains an uncovered and a covered penguin.
+The uncovered penguin is divisible but, because it is in the whole state, it might easily be taken to be indivisible.
+The covered penguin is indivisible.
+The findings suggest that infants know that the uncovered penguin is divisible but expect the protagonist to believe that the uncovered penguin is indivisible and to infer that the covered penguin is divisible.
+Does this pattern of knowledge and expectation require that the infants have attributed a false belief about identity?
+This is not necessary.
+Suppose that infants attributed the following reasoning to the protagonist: this penguin visible on my right is \textit{an} indivisible penguin; \textit{a} divisible penguin is always present; therefore there is \textit{a} divisible penguin under the cover.
+This reasoning concerns only the types of objects present and does not involve identity.
+Yet it is sufficient (together with some further inferences about the protagonists’ goal) for success.
+So while Scott and Baillargeon’s (2010) findings are interesting in many respects, their implementation does not provide evidence that infants ascribe beliefs about identity.
+
+In this section we have explained ways of distinguishing minimal theory of mind cognition from its full-blown counterpart.
+Where subjects are able to track beliefs about identity we know they are not relying only on minimal theory of mind; and, conversely, where subjects are able to track beliefs about objects’ properties but not about identity, we can infer that they are not using full-blown theory of mind.
+We next consider how  minimal theory of mind might be distinguished experimentally from strategies which involve representing nonintentional behaviour only.
+
+\section{Limits: how to distinguishing minimal theory of mind from behavioural strategies}
+\label{sec:limits2}
+What could show that subjects are not solving a task by representing nonintentional behaviour only but are using minimal theory of mind or something stronger?  
+We shall use the term ‘behavioural strategy’ to refer to processes underpinning theory of mind abilities which involve representing nonintentional behaviours only.
+
+An obstacle to distinguishing minimal theory of mind from behavioural strategies is that there is no in-principle limit to the complexity of the predictions which behavioural strategies might support.
+The behaviour-representing counterpart of Laplace’s demon knows our entire behavioural histories, has perfect knowledge of the regularities governing object-directed actions and is not limited by scarce cognitive resources.
+This demon can readily predict our future behaviours without any theory of mind.
+
+One way to avoid the problem of the behaviour-representing demon is to set tasks involving explanation rather than prediction \citep[compare][p.\ Chapter 8]{en_167}.
+Subjects observe a standard false belief scenario in which a character, Maxi, first searches for an object in the wrong location.
+The test question is why Maxi searched as he did \citep[see e.g.][]{en_80, en_352}.
+Where subjects’ answers include reference to incorrect registration or a related notion (which might, of course, be verbalised in terms of what an agent ‘thinks’), they have at least minimal theory of mind cognition.
+Of course, this only works where subjects can be asked ‘why’ questions.
+
+For other subjects we need an indirect measure.
+If subjects sometimes represent encounters or perceptions as such, then representations of others’ encounters may interfere with expectations or judgements about reality or with judgements about one’s own perceptions and beliefs \citep[so-called `altercentric interference',][]{Samson:2010jm}.
+For example, representing a protagonists’ encounter with an object at a particular location may interfere with expectations about the object’s actual location.
+Accordingly, evidence that there is such interference when in a situation where others have different perceptions or beliefs would provide indirect evidence for theory of mind cognition (minimal or full-blown) over behavioural strategies, for such interference would be difficult to explain if subjects were representing behaviours only.
+Evidence showing this kind of altercentric interference has recently been found in human adults \citep[]{en_1797, Samson:2010jm}.
+These studies show that adults’ judgements about the number of objects they could see in a visual scene were slower and more error-prone when the scene contained an irrelevant other whose perspective was different, suggesting that another’s perspective can cause altercentric interfere.
+
+Of course it was never in question that adults represent encounters or perceptions and not only behaviours; the relevance of these studies is that they demonstrate that altercentric interference can be used as an indirect measure for distinguishing behavioural strategies from minimal theory of mind cognition.
+\citet[]{kovacs_social_2010} report evidence of altercentric interference in infants’ expectations about the presence of an object.
+Their study suggests that infants’ surprise when their expectations about an object’s presence or absence are violated is modulated by facts about another’s expectations about that object’s location.
+Infants looked for less time at the unexpected presence of an object behind a screen when another observer incorrectly expected it to be there than when the observed did not expect it to be there.
+These findings seem to be evidence of altercentric interference, which is consistent with either minimal or full-blown theory of mind but difficult to explain if participants are exclusively using behavioural strategies.
+
+A related way to distinguish behavioural strategies from minimal theory of mind or something stronger is by means of self-other inference.
+Emery and Clayton (\citeyear[]{en_1411}) divided scrub-jays into those who had previously pilfered food and those who had never pilfered.
+All subjects then cached some food in private and cached some more food while being observed.
+When later given privacy, the pilferers re-cached the food they had earlier cached in the presence of the observer but did not re-cache the food they had cached in private.
+The non-pilferers did not show this pattern of behaviour and did little re-caching.\footnote{ 	Similar results might be found if this paradigm were adapted for human infants \citep[see][]{en_1707}.}  This is difficult to reconcile with the hypothesis that scrub-jays rely entirely on behavioural strategies.
+By contrast, the hypothesis that scrub-jays have at least minimal theory of mind and so are able to represent goal-directed actions as such allows us to suppose that engaging in pilfering alerts them to the possibility that others may do the same.\footnote{ 	In later discussions these authors claim that the pilferers may be projecting their own experiences onto others in the sense that they are using `their own experiences to infer experiences in others' \citep[][p.\ 81]{Emery:2007ze}.
+This claim goes beyond our suggestion, which involves goals rather than experiences and salience rather than projection.
+}
+
+Another route to distinguishing minimal theory of mind cognition from behavioural strategies derives from an analogy with speech perception.
+It is well established that phoneme recognition is facilitated by lexical information \citep[]{en_1704}.
+Similarly, for subjects with (at least) minimal theory of mind cognition, there is evidence that identification of significant segments of behaviour is facilitated by information about goals.
+For instance, Zacks (\citeyear[]{en_1442}) varied the availability of information about goals while holding motion features of behaviour constant.
+He found that subjects given information about goals segmented behaviour differently from subjects not given such information.
+Of course for Zacks it was not in question that his subjects could represent goals; the issue was how their representation of goals interacted with their representations of behaviour.
+But with modification the  paradigm could be used to test participants who are not already known to be capable of representing goals.
+If making information about goals, encounterings or registrations available alters how subjects segment behaviours, this would be evidence that they are not relying on behavioural strategies only.
+
+In short, a variety of routes to distinguishing behavioural strategies from minimal theory of mind are available.
+Some of these involve established paradigms, others adapt ideas from related research.
+A decisive argument either way would require converging evidence from multiple paradigms \citep[p.\ 622]{premack_does_1978}.
+
+\section{Conclusion}
+
+Abilities to track\footnote{
+The notion of \emph{tracking} was defined  \vpageref{df:track}; see also the definitions of \emph{theory of mind ability} (\vpageref*{df:tom_ability}) and \emph{theory of mind cognition} (\vpageref*{df:tom_cognition}).
+}
+perceptions and beliefs are  sometimes but not always  automatic in human adults (or so we argued in Section \ref{sec:motivation}).   
+But, as a variety of theoretical and empirical considerations indicate,  representing beliefs and other propositional attitudes as such is associated with demands on working memory, inhibition or attention that are incompatible with automaticity.
+This motivated asking what  someone could represent that would enable her to track perceptions, knowledge states and beliefs without meeting these cognitive demands.
+
+Further motivation for asking this question comes from  evidence that theory of mind abilities are widespread, occurring not only in human adults but also in infants, chimpanzees and scrub-jays at least (see Section \ref{sec:more_motivation}).
+By contrast full-blown theory of mind cognition may be comparatively rare 
+thanks to the conceptual and cognitive demands associated with representing beliefs and other propositional attitudes as such.  
+If so it is useful to identify what else someone could represent that would enable her to track, in a limited but useful range of situations, perceptions, knowledge states and beliefs, including false beliefs.
+
+To answer this question we constructed a minimal theory of mind.
+The construction is sufficiently rich to explain systematic success on tasks held to be acid tests for theory of mind cognition including many false belief tasks.
+It is also extensible in ways that can explain a range of findings from non-human animals and human infants that are often presented as evidence for full-blown theory of mind cognition.
+Where minimal theory of mind must break down is in cases involving quantification or mistakes about identity.
+Because such cases require full-blown theory of mind, it is possible to distinguish whether an individual’s performance on a particular task involves minimal or full-blown theory of mind cognition.
+
+The central notions for minimal theory of mind are encountering and registration.
+We characterised these by their structures, acquisition conditions and roles in producing goal-directed actions.
+Encountering and registration serve as non-representational proxies for perception and belief in this sense: in a limited but useful range of everyday circumstances, agents perceive an object just when they encounter it and they believe that an object has a given property just when they register it having that property.
+But encountering differs from perception in not constitutively involving appearances, modalities or the possibility of illusion and in not being constitutively linked to reasons, knowledge or informational states.
+Similarly, registration differs from belief in lacking representational content and the possibility of embedding, in not interacting in arbitrarily complex ways with other states, and in the simplicity of its parameter-setting role in causing actions.
+These differences ensure that representing encountering or registration need not involve meeting conceptual and cognitive demands associated with representing perceptions, beliefs and other propositional attitudes as such.
+
+The novelty of our constructive approach lies in several features.  It does not rely directly on everyday psychological concepts, whose exact nature is a source of controversy.
+Nor does it rely on infants or non-human animals holding theoretical commitment to simplified versions of these concepts (contrast 
+	\citealp{en_75};
+	\citealp{en_81};
+	\citealp{en_2372} and
+	\citealp{en_1614}%
+).
+Importantly, we do not assume that minimal theory of mind develops into full-blown theory of mind in humans.  
+It may instead remain distinct, supporting cognitively efficient theory of mind across the lifespan \citep[see][]{Samson:2010jm, en_2397}.
+The construction of minimal theory of mind is systematic enough to generate testable predictions distinguishing it from both behavioural stratetgies and full-blown theory of mind cognition (see Sections \ref{sec:limits1} and \ref{sec:limits2}).
+Our construction makes detailed sense of the notion that there are degrees of theory of mind cognition.  
+(This is a virtue because while is widely recognised that degrees of theory of mind cognition are needed \citep[e.g.][]{en_610, en_1415}, there have been few detailed attempts to make systematic sense of this possibility.)
+It also pushes much further than earlier work the boundaries of what can be achieved without full-blown theory of mind cognition; in particular, it explains how systematic success on a range of false belief tasks (but not those which essentially involve identity or quantification) is possible without representing beliefs or other propositional attitudes as such.
+Minimal theory of mind may be what enables those with limited cognitive resources or little conceptual sophistication, such as infants, chimpanzees, scrub-jays and human adults under load,  to track  perceptions, knowledge states and beliefs.
 

### 2012-04-06 mindreading/minimal theory of mind (new version)/minimal theory of mind.nv.tex
mtom revision sent to Ian

@@
 \noindent
-What could someone represent that would enable her to track, at least within limits, perceptions, knowledge states and beliefs including false beliefs? 
-An obvious possibility is that she might represent these very propositional attitudes as such.
+What could someone represent that would enable her to track, at least within limits, others' perceptions, knowledge states and beliefs including false beliefs? 
+An obvious possibility is that she might represent these very attitudes as such.
 It is sometimes tacitly or explicitly assumed that this is the only possible answer.
@@
 Minimal theory of mind is rich enough to explain systematic success on tasks held to be acid tests for theory of mind cognition including many false belief tasks.  
-It is also extensible in ways that can explain a wide range of findings from non-human animals and human infants that are sometimes presented as evidence for full-blown theory of mind cognition.  
 Yet minimal theory of mind does not require representing propositional attitudes, or any other kind of representation, as such.
-Minimal theory of mind may be what enables those with limited cognitive resources or little conceptual sophistication, such as infants, chimpanzees, scrub-jays and human adults under load, able to track, within limits, facts about perceptions and beliefs.
+Minimal theory of mind may be what enables those with limited cognitive resources or little conceptual sophistication, such as infants, chimpanzees, scrub-jays and human adults under load,  to track others' perceptions, knowledge states and beliefs.
 
@@
 \section{Introduction}
-What could someone represent that would enable her to track, at least within limits, perceptions, knowledge states and beliefs including false beliefs? 
+What could someone represent that would enable her to track, at least within limits, others' perceptions, knowledge states and beliefs including false beliefs? 
 One answer is obvious: she might track these things by virtue of representing them as such, that is, by representing perceptions, beliefs, and other propositional attitudes as such.
@@
 If these arguments are not yet decisive
-and if the available evidence does not already tightly constrain the space of admissible conjectures,
+and if the available evidence does not already tightly constrain the space of admissible hypotheses,
 then the construction of a minimal theory of mind may also be relevant to these debates.
-For the conjecture that minimal theory of mind explains chimpanzees' or scrub-jays' abilities to track perceptions or beliefs can be empirically distinguished from conjectures about representations of nonintentional behaviour only and from conjectures about representations of perceptions, knowledge states and beliefs (as we explain in section *** below).
+For the hypothesis that minimal theory of mind explains particular subjects' abilities to track perceptions or beliefs can be empirically distinguished from hypotheses about representations of nonintentional behaviour only and also from hypotheses about representations of perceptions, knowledge states and beliefs (as we explain in Sections \ref{sec:limits1} and \ref{sec:limits2} below).
 
@@
 \section{Minimal theory of mind}
-
+\label{sec:minimal}
 In this section we begin with someone, call her Lucky, capable  of representing nonintentional behaviour only and ask what more is needed for minimal theory of mind cognition.
@@
 As we go on to discuss, our proposal for a minimal theory of mind could be extended to cover a range of other cases;
-but importantly there are also false belief tasks success on which cannot be explained by minimal theory of mind cognition (see section ***).
+but importantly there are also false belief tasks success on which cannot be explained by minimal theory of mind cognition (see Section \vref{sec:limits1}).
 }
@@
 Of course it was never in question that adults represent encounters or perceptions and not only behaviours; the relevance of these studies is that they demonstrate that altercentric interference can be used as an indirect measure for distinguishing behavioural strategies from minimal theory of mind cognition.
-\citet[]{kovacs_social_2010} report evidence of altercentric interference in infants’ expectations about the presence of an object.
+\citet[]{kovacs_social_2010} report evidence of possible altercentric interference in infants’ expectations about the presence of an object.
 Their study suggests that infants’ surprise when their expectations about an object’s presence or absence are violated is modulated by facts about another’s expectations about that object’s location.
 Infants looked for less time at the unexpected presence of an object behind a screen when another observer incorrectly expected it to be there than when the observed did not expect it to be there.
-These findings seem to be evidence of altercentric interference, which is consistent with either minimal or full-blown theory of mind but difficult to explain if participants are exclusively using behavioural strategies.
+These findings may be evidence of altercentric interference,%
+\footnote{
+An anonymous referee pointed out that  this is not the only possible interpretation.
+% raised the possibility that the pattern of infants'  looking times in this experiment might be explained not by their representing others' expectations but rather by their representing others' presence or absence at critical moments.
+We do not mean to suggest that a single study provides decisive evidence for altercentric interference, only that in the wider context of research on infant theory of mind  it provides an indication and constitutes progress on the question of how altercentric interference could be measured.
+}
+which would be consistent with either minimal or full-blown theory of mind but difficult to explain in a principled way if subjects are  using behavioural strategies only.
 
@@
 To answer this question we constructed a minimal theory of mind.
-The construction is sufficiently rich to explain systematic success on tasks held to be acid tests for theory of mind cognition including many false belief tasks.
-It is also extensible in ways that can explain a range of findings from non-human animals and human infants that are often presented as evidence for full-blown theory of mind cognition.
-Where minimal theory of mind must break down is in cases involving quantification or mistakes about identity.
+The construction is  rich enough to explain systematic success on tasks held to be acid tests for theory of mind cognition including many false belief tasks.
+%It is also extensible in ways that can explain a range of findings from non-human animals and human infants that are often presented as evidence for full-blown theory of mind cognition.  
+Where minimal theory of mind must break down is in cases involving quantification or mistakes about identity (see Section \ref{sec:limits1}).  
 Because such cases require full-blown theory of mind, it is possible to distinguish whether an individual’s performance on a particular task involves minimal or full-blown theory of mind cognition.
@@
 The central notions for minimal theory of mind are encountering and registration.
-We characterised these by their structures, acquisition conditions and roles in producing goal-directed actions.
+We characterised these by their structures, acquisition conditions and roles in producing goal-directed actions (see Section \ref{sec:minimal}).
 Encountering and registration serve as non-representational proxies for perception and belief in this sense: in a limited but useful range of everyday circumstances, agents perceive an object just when they encounter it and they believe that an object has a given property just when they register it having that property.
@@
 It also pushes much further than earlier work the boundaries of what can be achieved without full-blown theory of mind cognition; in particular, it explains how systematic success on a range of false belief tasks (but not those which essentially involve identity or quantification) is possible without representing beliefs or other propositional attitudes as such.
-Minimal theory of mind may be what enables those with limited cognitive resources or little conceptual sophistication, such as infants, chimpanzees, scrub-jays and human adults under load,  to track  perceptions, knowledge states and beliefs.
+Minimal theory of mind may be what enables those with limited cognitive resources or little conceptual sophistication, such as infants, chimpanzees, scrub-jays and human adults under load,  to track others'  perceptions, knowledge states and beliefs.
 

### 2012-04-15 intention motor representation with corrado sinigaglia/conf lyon glasgow was spp submit rejected/intention and motor representation GLASGOW.tex
talk

--- 2012-03-20 intention motor representation with corrado sinigaglia/conf spp submit/intention and motor representation SPP submit.tex
@@
 
-\def \ititle {Intention and motor representation:}
-\def \isubtitle {how to resist temptation}
+\def \ititle {Intention and motor representation in purposive action}
+\def \isubtitle {}
 \def \iauthor {
-%S. Butterfill* \& C. Sinigaglia**
-%\\ 
-%**Department of Philosophy, University of Warwick
-%\\ 
-%***Dipartimento di Filosofia, Università degli Studi di Milano
+S. Butterfill* \& C. Sinigaglia**
+\\ 
+**Department of Philosophy, University of Warwick
+\\ 
+***Dipartimento di Filosofia, Università degli Studi di Milano
 }
 \def \iemail{
-%	s.butterfill@warwick.ac.uk
+	s.butterfill@warwick.ac.uk
 }
-%\date{}
 
 \input{$HOME/Documents/submissions/preamble_steve_paper2}
-\author{}
-\date{}
 
@@
 \noindent
-Are there distinctive roles for intention and motor representation in explaining the purposiveness of action?
-Standard philosophical accounts of action assign a role to intention but are silent on motor representation.
+Are there distinct roles for intention and motor representation in explaining the purposiveness of action?
+Standard accounts of action assign a role to intention but are silent on motor representation.
 The temptation is to suppose that nothing need be said here because motor representation is either only an enabling condition for purposive action or else merely a variety of intention.
-Drawing on cognitive and neuroscientific research, this talk identifies reasons for resisting temptation.
+This paper provides reasons for resisting that temptation.
 Some motor representations, like intentions, coordinate actions in virtue of representing outcomes; but, unlike intentions, motor representations cannot feature as premises or conclusions in practical reasoning.
@@
 It also gives rise to a problem: were the roles of intention and motor representation entirely independent, this would impair effective action.
-Resisting temptation therefore requires explaining how intentions interlock with motor representations.
-The solution, we argue, is to recognise that the contents of intention can be partially determined by the contents of motor representations.
-Understanding this content-determining relation enables better understanding how intention is related to action.
-
-\
-
-\noindent
-Word count: 3731 including notes, excluding references
+It is therefore necessary to explain how intentions interlock with motor representations.
+The solution, we argue, is to recognise that the contents of intentions can be partially determined by the contents of motor representations.
+Understanding this content-determining relation enables better understanding how intentions relate to actions.
 
@@
 
+\section{Pre-Introduction}
+
+Which events are actions?
+
+In philosophy this question is sometimes answered in terms of intention.
+For an event to be an action is for it to stand in a special relation to an intention.
+This view says nothing at all about how intentions could result in bodily movements; that is considered to be merely a matter of implementation.
+
+In psychology the question is sometimes answered in terms of motor representation.
+For an event to be an action is for it to stand in a special relation to a motor representation.
+This view says little about actions whose purposes involve things the motor system doesn’t care about---for example, your motor system doesn’t care whether you’re ripping up a real Armani shirt or a fake Armani one, but this sort of difference can affect whether your actions succeed or fail.
+
+You might just say that the two disciplines are talking past each other.
+But we want to argue that these two pictures are aspects of a larger theory of action.
+
 \section{Introduction}
 
+A basic question about action is,
 What is the relation between a purposive action and the outcome or outcomes to which it is directed?

### 2012-04-16 intention motor representation with corrado sinigaglia/conf lyon glasgow was spp submit rejected/intention and motor representation HANDOUT.tex
talk

--- 2012-03-20 intention motor representation with corrado sinigaglia/conf spp submit/intention and motor representation HANDOUT.tex
@@
 
-\def \ititle {Intention and Motor Representation in Action Explanation}
+\def \ititle {Intention and Motor Representation in Purposive Action}
 \def \isubtitle {}
@@
 {\Large
-\textbf{Intention and Motor Representation in Action Explanation}
+\textbf{Intention and Motor Representation in Purposive Action}
 }
@@
 
-\item My intention that I visit Paris on Friday is a propositional attitude.
+\item My intention that I visit Glasgow on Monday is a propositional attitude.
 

### 2012-04-24 intention motor representation with corrado sinigaglia/intention and motor action.ppr.corrected.tex
a

--- 2012-04-04 intention motor representation with corrado sinigaglia/intention and motor action.ppr.r1.final.tex
@@
 For example, suppose you are faced with an array of objects one of which---the target---you will shortly be required to grasp.
-Subjects who imagine grasping an object other than the target object tend to be slower in subsequently grasping the target object than subjects who do not imagine acting or subjects who imagine grasping the target object \citep{ramsey:2010_incongruent_}.
+Subjects who imagine grasping an object other than the target object tend to be slo wer in subsequently grasping the target object than subjects who do not imagine acting or subjects who imagine grasping the target object \citep{ramsey:2010_incongruent_}.
 Just as the similarities between imagining seeing and actually seeing are evidence for the hypothesis that the representations involved in imagining seeing and actually seeing have a common format, so also the similarities in characteristic performance profile between imagining acting and actually acting, together with the particular patterns of interference between the two, suggest that imagining acting and actually acting involve a common representational format.
@@
 Similarly, we mentioned that factors making actually acting more effortful also make imagining acting more effortful.
-For instance, in some cases rotating a ball clockwise is easier than rotating it anti-clockwise, and so is imagining a ball rotate.
+For instance, in some cases rotating a ball clockwise is easier than rotating it anti-clockwise,
+%*correction!
+and so is imagining rotating a ball.
 By contrast, the effort involved in actually seeing or imagining seeing a ball rotate does not similarly differ depending on direction.

### 2012-04-25 mindreading/minimal theory of mind (new version)/minimal theory of mind.nv.tex
mToM resubmitted to mind and language

@@
 %for anonymous submisison
-%\def \iauthor {}
-%\def \iemail{}
+\def \iauthor {}
+\def \iemail{}
 %\date{}
@@
 %but don't forget to remove defs above as well --- otherwise it will be in the metadata
-%\author{}
-%\date{}
+\author{}
+\date{}
 
@@
 
-In this section we defend explain our question; in the next sections we introduce the findings which motivate facing it before starting to answer it in the fourth section.
+In this section we defend and explain our question; in the next sections we introduce the findings which motivate facing it before starting to answer it in the fourth section.
 
@@
 but what someone could represent that would enable her, at least within limits, to track perceptions, knowledge states and beliefs.
-The motivation for facing up to this question is, of course, partly empirical.
+However, the motivation for facing up to this question is, of course, partly empirical.
 
@@
 But there is also a body of evidence supporting a different conclusion.
-\citet{back:2010_apperly} found that subjects are significantly slower to answer an unexpected question about another's belief when that belief is false compared to when it is true \citep[see also][]{apperly:2006_belief}.
+\citet{back:2010_apperly} found that subjects are significantly slower to answer an unexpected question about another's true or false belief compared with matched questions about reality \citep[see also][]{apperly:2006_belief}.
 This suggests that, at least in adults, belief tracking is not automatic.
@@
 
-A closely related view has already been elaborated and defended in more detail by
+A closely related view has already been elaborated and defended in some detail by
 \citet[]{Apperly:2009ju},
-although their argument complements ours by drawing  primarily on developmental and comparative research. 
+whose argument complements ours by drawing  primarily on developmental and comparative research, and less on theoretical motivations. 
 According to them,
@@
 
-What do these infants and young children represent that enables them, within limits, to track others’ perceptions, knowledge states and beliefs?   
+What do these infants and young children represent that enables them to track others’ perceptions, knowledge states and beliefs in such cases?   
 The most straightforward answer would be to suppose that they represent perceptions, knowledge states and beliefs as such \citep[e.g.][]{en_1138, en_1691}.  
@@
 But there are several objections to the claim that the full range of even infants' abilities to track perceptions and beliefs could be explained in this way \citep{en_1691,Apperly:2009ju}.  
-And what is is currently known about humans' actual behaviour reading capacities suggests that they are unlikely to explain systematic success on false belief tasks.%
+And what is currently known about humans' actual behaviour reading capacities suggests that they are unlikely to explain systematic success on false belief tasks.%
 \footnote{
@@
 
-Theory of mind abilities are not only found in humans.
+Finally, theory of mind abilities are not only found in humans.
 For instance,
@@
 }
-If these arguments are not yet decisive
-and if the available evidence does not already tightly constrain the space of admissible hypotheses,
-then the construction of a minimal theory of mind may also be relevant to these debates.
+There is certainly no current resolution to these arguments,
+and since the available evidence does not already tightly constrain the space of admissible hypotheses,
+the construction of a minimal theory of mind may also be relevant to these debates.
 For the hypothesis that minimal theory of mind explains particular subjects' abilities to track perceptions or beliefs can be empirically distinguished from hypotheses about representations of nonintentional behaviour only and also from hypotheses about representations of perceptions, knowledge states and beliefs (as we explain in Sections \ref{sec:limits1} and \ref{sec:limits2} below).
@@
 As we go on to discuss, our proposal for a minimal theory of mind could be extended to cover a range of other cases;
-but importantly there are also false belief tasks success on which cannot be explained by minimal theory of mind cognition (see Section \vref{sec:limits1}).
+but importantly there are also false belief tasks success on which success could not be explained by minimal theory of mind cognition (see Section \vref{sec:limits1}).
 }
@@
 Even so, we can be sure that there are some differences between encountering and perception.
-If perceptions are representations, then representing perceptions as such plausibly involves representing representations.
+Perhaps most obviously, if perceptions are representations, then representing perceptions as such plausibly involves representing representations.
 Since encounterings are relations not representations (by definition), representing encounterings will  differ from representing perceptions in that only the latter involves representing representations.
+
 But, as mentioned, not everyone agrees that perceptions are representations.
 We should therefore consider a bundle of features.
-Perception constitutively involves appearances, modalities or the possibility of illusion or is constitutively linked to reasons, knowledge or informational states.
+According to one or other of the theories just described, perception constitutively involves appearances, modalities or the possibility of illusion or is constitutively linked to reasons, knowledge or informational states.
 It is plausible that representing perceptions as such involves understanding something about some of these things.
@@
 As is already clear from this principle, registration is like belief in that it has a correctness condition which may not obtain: a registration is correct when the object is in the location.
+
 Since representing registration brings no benefit for Lucky independently of its connection to action, this principle cannot stand alone in our sequence; it is only the first half of the third principle.
@@
 
-Note that have not argued for the truth of any hypothesis about what chimpanzees, scrub-jays or infants represent.
+Note that we have not argued for the truth of any hypothesis about what chimpanzees, scrub-jays or infants represent.
 As explained above, what motivates our enquiry is not that there is already evidence for hypotheses about particular subjects' minimal theory of mind cognition: it is that, given the limited evidence for competing hypotheses and the objections they face, new hypotheses are worth formulating and testing.
-And in later sections we will explain how hypotheses about minimal theory of mind generate testable predictions, enabling them to be empirically distinguished from competing hypotheses.
+In later sections we will explain how hypotheses about minimal theory of mind generate testable predictions, enabling them to be empirically distinguished from competing hypotheses.
 For now our focus is the question of what a subject could represent that would enable her to track track perceptions, knowledge states and beliefs in ways measured by various experimental paradigms.
@@
 Applications of the fourth principle therefore include Onishi and Baillargeon’s (\citeyear[]{en_1092}) false belief task.
-Subjects are shown an adult observer who is present while a piece of melon is placed in one box.
+Infant subjects are shown an adult observer who is present while a piece of melon is placed in one box.
 In the critical condition, the adult observer is then absent while the melon moves to another box.
@@
 \subsection{Extensions and variations}
-With the fourth principle we completed the construction of a minimal theory of mind capable of underwriting success on some false belief tasks.  
+With the fourth principle we have completed the construction of a minimal theory of mind capable of underwriting success on some false belief tasks.  
 We stop here because false belief tasks are often taken to be an acid test for theory of mind.
@@
 To ease exposition we assume throughout that Lucky has no false beliefs involving identity.}  For to register \(<\)Charly, Baltimore\(>\) is the same thing as registering \(<\)Samantha, Baltimore\(>\).
-And Lucky should be equally at a loss when those she observes mistakenly believe that two distinct people are identical.
+And Lucky should be equally at a loss when those she observes someone mistakenly believe that two distinct people are identical.
 By contrast, subjects who can represent beliefs as such should have no special problem with false beliefs essentially involving identity.
@@
 
-Given that this analysis, it is notable that one study already claims to have found positive evidence that infants can ascribe beliefs about identity. 
+Given this analysis, it is notable that one study already claims to have found positive evidence that infants can ascribe beliefs about identity. 
 Scott and Baillargeon (\citeyear{en_1690}) have already conducted experiments which aim to show that 18-month-old infants are able to represent beliefs involving mistakes about `which particular object \textit{token}' another is facing rather than only about `which \textit{type} of object' is present (p.\ 1179).

### 2012-06-15 joint action and knowing others minds/submit philosophical studies  2012-01/second submit/reply to reviewer.txt
submitted revision to Phil Studies

The reviewer's comments and two suggestions were extremely useful.  The paper has been  revised to incorporate both suggestions.

--- First suggestion:
"The authors might address what a passive observer C might think of what B is doing.  Suppose the passive observer knows what A is doing (trying to get the pram on the bus).  Then C, observing A's cues for engaging in joint activity with B, could infer as easily as B that A is trying to lift the stroller to help get it on the bus.  I don't think that this undermines the author's argument, because C relies on the very capacities to engage in interaction that the authors emphasize.  But it might be nice to address this sort of case, just to clarify their thesis:  after all, C does not on this occasion engage in any reciprocity, interaction, or willingness to interact.   (Nor does C rely on any deception or cognitive sophistication; cf. p14)."

--- Reply to first suggestion:
This has been done in a new section, section 8 on pp. 20-1 titled 'Observing Interaction'.  There is also a new paragraph on p. 15 (the third full paragraph) and corresponding minor changes to statements of the argument throughout.  As the reviewer points out, the suggestion doesn't undermine the main line of argument.  Even so, it seemed appropriate to devote a whole section to it both because it clarifies our thesis and because it suggests a further way in which interaction may matter for mindreading.

--- Second suggestion:
"Concerning the your-goal-is-my-goal inference: the authors rely on the idea that the mind-reader A has a fixed goal G so that when the other person B expresses willingness to engage in joint activity, A can infer that B will have the same goal.  But is this always the case?  Often when someone makes an overture to engage in shared activity, it's not always clear that you are supposed to keep on doing exactly what you were doing.  There are, of course, cases where someone joins in and adopts a goal that the other or others have.  But often shared activity gets individuals doing things they weren't doing before.  So when A signals that they are willing to act jointly, can the mindreader (B) assume that A has the same goal?  Couldn't A be trying to get him to have some different goal with him? So, clarifying the inference to rule out this case might be helpful."

--- Reply to first suggestion:
This has been done in with a slight change to premise (1) on page 13 and a new paragraph on page 14 (the second full paragraph) .

I've mentioned a couple of suggestions, but the paper is pretty much fit to be published as is.  I think it meets the high standards for Philosophical Studies.

### 2012-06-15 mindreading/conf espp 2012 symposium false belief counterfactuals perspective schurz rafetseder/false beliefs, perspectives and counterfactuals espp 2012 symposium submission.tex
minimal theory of mind sent to mind and language final (i hope)

--- 2012-02-27 conf espp 2012 symposium false belief counterfactuals perspective schurz rafetseder/false beliefs, perspectives and counterfactuals espp 2012 symposium submission.tex
@@
 this proposal faces several objections \citep[e.g.][]{Song:2008qo}.
-The core issue is not whether hypothetical behavioural strategies might in principle explain theory of mind abilities; it is whether all of these subjects’ actual behaviour reading capacities are flexible enough to explain the full range of their theory of mind abilities \citep[]{Apperly:2009ju}
+The core issue is not whether hypothetical behavioural strategies might in principle explain theory of mind abilities; it is whether all of these subjects’ actual behaviour reading capacities are flexible enough to explain the full range of their theory of mind abilities \citep[]{Apperly:2009ju}.
 In my view this is unlikely.
@@
 (\citealp{Call:1999co}; \citealp{low:2010_preschoolers} Study 2).
-We should also note that children who fail Category A tasks can answer questions about perception or pretence that are word-for-word identical with the questions about false belief that they cannot answer correctly (\citealp{Gopnik:1994cb}; see also\citealp{Cluster:1996ht}).
+We should also note that children who fail Category A tasks can answer questions about perception or pretence that are word-for-word identical with the questions about false belief that they cannot answer correctly (\citealp{Gopnik:1994cb}; see also \citealp{Cluster:1996ht}).
 So the Extraneous Requirement cannot be straightforwardly linked to language or communication.
@@
 and some involve prediction actions whereas others involve predicting desires \citep{Astington:1991kk} or selecting an argument appropriate for someone with a false belief \citep{Bartsch:2000es}.
-Despite all this variation and more, the Category A tasks all appear to measure a developmental transition \citep{Wellman:2001lz}.
+Despite all this variation and more, the Category A tasks all appear to measure a single developmental transition \citep{Wellman:2001lz}.
 This is a significant obstacle to identifying the Extraneous Requirement.

### 2012-06-20 joint action and knowing others minds/submit philosophical studies  2012-01/second submit/interacting_mindreaders_source/interacting mindreaders.3.ps.corrections.tex
final version submitted to phil studies accepted

--- 2012-01-07 joint action and knowing others minds/interacting mindreaders.2.tex
@@
 
-\documentclass[12pt,a4paper]{extarticle}
+%\def \papersize {a4paper}
+\def \papersize {letterpaper}
+\documentclass[12pt,\papersize]{extarticle}
 % extarticle is like article but can handle 8pt, 9pt, 10pt, 11pt, 12pt, 14pt, 17pt, and 20pt text
@@
 \def \isubtitle {}
-%\def \iauthor {Stephen A. Butterfill}
-%\def \iemail{s.butterfill@warwick.ac.uk}
-\def \iauthor {}
-\def \iemail{}
+\def \iauthor {Stephen A. Butterfill}
+\def \iemail{s.butterfill@warwick.ac.uk}
+%\def \iauthor {}
+%\def \iemail{}
 \date{}
 
-\input{$HOME/Documents/submissions/preamble_steve_paper2}
-\author{}
+\input{preamble_steve_paper2}
+%\author{}
 
@@
 \bibliographystyle{newapa} %apalike
+
+\tolerance=5000
 
@@
 \begin{abstract}
+\noindent
 Could interacting mindreaders be in a position to know things which they would be unable to know if they were manifestly passive observers?
 This paper argues that they could.
-Mindreading is sometimes reciprocal: 
-the mindreader's target reciprocates by taking the mindreader as a target for mindreading.
-The paper explains how such reciprocity can significantly narrow the range of possible interpretations of behaviour 
-where mindreaders are, or appear to be, in a position to interact.
-It follows that
-revisions and extensions 
-are needed to 
-standard theories of mindreading
-on which
-all the evidence 
-is in principle available to manifestly passive mindreaders.
-The view also has consequences for 
-understanding 
-how abilities to interact combined with comparatively simple forms of mindreading
-may explain the emergence,
-in evolution or development,
-of sophisticated forms of social cognition.
+Mindreading is sometimes reciprocal: the mindreader's target reciprocates by taking the mindreader as a target for mindreading.
+The paper explains how such reciprocity can significantly narrow the range of possible interpretations of behaviour where mindreaders are, or appear to be, in a position to interact.
+A consequence is that revisions and extensions are needed to standard theories of the evidential basis of mindreading.
+The view also has consequences for understanding how abilities to interact combined with comparatively simple forms of mindreading may explain the emergence, in evolution or development, of sophisticated forms of social cognition.
 
@@
 The two are on a par in this sense:
-in principle the same evidence could be available to each, and each can exploit the same principles in moving from evidence to ascriptions of thought and action.
+in principle the same evidence could be available to each, and each can exploit the routes to knowledge in moving from evidence to ascriptions of thought and action.
 Of course these theories are all compatible with the idea that interaction might be useful for mindreading in practice.  
@@
 that mindreaders actually or apparently capable of interacting with their targets are at an advantage not only in practice but also in theory.
-Their ascriptions could rely on evidence and principles which would be unavailable if they were entirely passive observers.
+Their ascriptions could exploit routes to knowledge which would be unavailable if they were entirely passive observers.
 
@@
 an individual's access to evidence for mindreading  depends in part on her access to evidence for goal ascription.
-Our plan is to show that interacting mindreaders could make use, in ascribing goals, of evidence and principles which are not available to mere observers (who do not even appear to be in a position to interact with their targets).
+Our plan is to show that interacting mindreaders could exploit routes to knowledge of the goals of others' actions which are not available to mere observers.
 In doing this we will be showing that
-mindreaders who are (or appear) poised to interact with others 
+mindreaders poised to interact with others 
 could know things about their minds 
@@
 According to 
-what Csibra and Gergely call `the principle of rational action':
+what Csibra and Gergely call `the principle of rational action',
 %
@@
 \footnote{
-Knowledge of the premises may not require explicit knowledge of 
+Knowledge of the conclusion may not require explicit knowledge of 
 	what is salient to others,
@@
 the next step is to identify an obstacle to acquiring knowledge of the goals of actions,
-one which we will eventually show can be avoided where mindreaders can (or appear to be able to) interact with their targets.
+one which we will eventually show can be avoided where mindreaders can (or appear able to) interact with their targets.
 
@@
  takes the possibility of interaction into account.
-In what follows we shall explain how abilities to interact with others provide a route to knowledge of the goals of other agents' actions  which avoids the problem of opaque means.
-This is one way in which moving away from a purely observational model of mindreading yields a richer evidential base for ascriptions.
+In what follows we shall explain how being able to interact with another sometimes makes available a route to knowledge of the goals of her actions  which avoids the problem of opaque means.
+This will support our claim that an interacting mindreader might be in a position to know things which she would be unable to know if she were only observing.
 
@@
 \label{sec:joint_action}
-Our overall aim is to defend this claim:
+We aim eventually is to defend this claim:
 there are routes to knowledge of the goals of others' actions
@@
 suppose that, while doing some gardening,
-we find a turnip too big for any of us easily pull out of the ground alone.
+we find a turnip too big for any of us to easily pull out of the ground alone.
 We each individually intend that we all pull up the enormous turnip
@@
 \footnote{
-For an argument that it is possible for each of us individually to intend that we do something, see \citet{Bratman:1999fr} and \citet[Chapter 2]{Bratman:2012fk}.
+For an argument that it is possible for each of us individually to intend that we do something, see \citet{Bratman:1999fr} and \citet{Bratman:2012fk}.
 }
@@
 In fact our actions might have a distributive goal even though none of us is aware of this, or even of the others' existence. 
-(We shall refrain from illustrating this point by contriving a truly gigantic turnip and a completely dark, very stormy night; readers can probably guess how this would go.)
+(We resist the temptation to contrive a truly gigantic turnip and a pitch dark, very stormy night; readers can probably guess how this would go.)
 One consequence of this is that 
@@
 if she is not limited to merely observing them,
-how might this provide her with a route to knowledge of the goals of their actions?
+how might this enable her to exploit a route to knowledge of the goals of their actions?
 The answer hinges on interactions involving distributive goals
@@
 Fortunately there is a way around this.  
-For there are various cues which signal that one agent is prepared to engage in joint action with another,
+For there are various cues which signal that one agent is prepared to engage in some joint action or other with another,
 and joint actions involve distributive goals.  
 \label{twin_pram}
-Seeing a stranger with the haggard look of a new parent struggling to get his twin pram on to the bus, you grab the front wheels and make eye contact, raising your eyebrows and smiling.
+Seeing you struggling to get your twin pram onto a bus and noticing you have the haggard look of a new parent, a passing stranger  grabs the front wheels and makes eye contact with you, raising her eyebrows and smiling.
 (The noise of the street rules out talking.)   
-In this way you signal that you are about to act jointly with the stranger, to lift the pram with him.   
-Since the stranger is fully committed to getting his pram onto the bus,
-he knows what the sole goal of his own actions will be.
-But he also knows that you will engage in joint action with him,
-which means that your actions and his actions will have a distributive goal.
-This may enable him to infer the goal of your imminent actions: 
-your goal is his goal, to get the pram onto the bus.
+In this way she signals that she is about to act jointly with you.   
+Since you are fully committed to getting your pram onto the bus,
+you know what the sole goal of your own actions will be.
+But you also know that the stranger will engage in joint action with you,
+which means that, taken together, her actions and your actions will have a distributive goal.
+This may enable you to infer the goal of the stranger's imminent actions: 
+her goal is your goal, to get the pram onto the bus.
 
@@
 \begin{enumerate}
-\item You are willing to engage in some joint action\footnote{
+\label{your_goal_is_my_goal}
+\item You are 
+%willing to 
+about to attempt to 
+engage in some joint action\footnote{
 We leave open the issue of how joint action is to be characterised subject only to the 
@@
 }
-or other with me
+or other with me.
 %(for example, because you have made eye contact with me while I was in the middle of attempting to do something).
@@
 in the right contexts
-an individual can recognize in another's facial expressions, engaging gestures or synchronized bodily movements that she is about to engage in joint action with her.
+an individual can recognize in another's facial expressions, engaging gestures or synchronized bodily movements that she is about to attempt to engage in joint action with her.
 Exploiting these indicators does not typically depend on knowing the particular contents of any beliefs, desires or goals.  
@@
 }  
-Of course these indicators provide no guarantee that others are genuinely willing to engage in joint action.  
+Of course these indicators do not guarantee that others are  about to attempt to engage in joint action.  
 But they are sufficiently reliable to ground knowledge in some cases.  
@@
 Not everything needs to rest on indicators, however.
-It is sometimes possible to know that others are willing to engage in joint action with you even without relying on such indicators.  
-Thanks to widespread dispositions to act jointly, in some situations it is reasonable to take for granted that others are willing to act jointly.  
+It is sometimes possible to know that others are about to attempt to engage in joint action with you even without relying on such indicators.  
+Thanks to widespread dispositions to act jointly, in some situations it is reasonable to take for granted that others will act jointly.  
 For example, this is often so for children struggling with a coat while surrounded by family or familiar adults.  
@@
 Of course dispositions to engage in joint action may vary between cultural groups and situations.
-This may be fatal for the peripatetic cosmopolitan,
+This may be fatal for the frequent traveller,
 but for others what matters is not whether the dispositions are universal but only that they are sufficiently widespread to be predictable.
@@
 These background conditions include the other having largely true beliefs concerning which goal your actions are or will be directed to.
-After all, the other agent may be willing to act jointly with you while being entirely mistaken about the goals to which your actions will be directed.
+After all, another agent may attempt to act jointly with you while being entirely mistaken about the goals of your actions.
 Where this happens, the premises of the inference might be true but the conclusion false.
-(We return to this point in section \vref{sec:false_belief}.)  
-But in some situations there is no requirement  to consider  this possibility, as for instance when stereotypes, conventions or simplicity should and do make the goal of your actions obvious to the other agent.
-Given the right conditions,
-that others are willing to engage in joint action with you 
+We shall return to this point later (in section \vref{sec:false_belief}). 
+For now, note that in some situations there is no requirement  to consider  this possibility, as for instance when stereotypes, conventions or simplicity should and do make the goal of your actions obvious to the other agent.
+
+As well as having true beliefs concerning the goal of your actions, the other must be willing to pursue this goal. 
+This is not guaranteed by the truth of the premises of the inference. 
+For all these require, the other might be aware of the goal you envisage but attempting to initiate a joint action directed to another, quite different goal%
+%and would be unwilling to perform actions directed to the goal you envisage, so that the conclusion of the inference is false
+.  
+Perhaps, for example, the other disapproves of your project and is offering to act with you as part of an attempt to divert you.\footnote{
+We are grateful to anonymous referee for emphasising this point.
+}
+Note, however, that it is only reasonable for the other to attempt a joint action with you where you are in a position to know which goal the joint action would be directed to. 
+If no alternative goal ought to be salient to you and the other has done nothing to make an alternative salient, then it is not unreasonable for you to ignore the possibility that the other is unwilling to pursue the goal to which your actions will be directed.
+This might easily be the case where, for instance, you are in a tidy kitchen patiently chopping your way through a large pile of carrots; it is perhaps less likely to be the case in a more cluttered kitchen where you are surrounded by all kinds of potential ingredient in different states of preparation.
+In general, how strongly the truth of the premises of the your-goal-is-my-goal inference supports the truth of the conclusion will depend (among other conditions) on how salient alternative goals to which your actions might be directed ought to be to you.
+
+Given these and perhaps other background conditions,
+that others are about to attempt to engage in joint action with you 
 is sometimes sufficient reason to hold that
@@
 in which a goal ascriber is manifestly in a position to interact with her target.
+
+We should also acknowledge that 
+mindreaders who only observe can make use of the premise about joint action that your-goal-is-my-goal depends on.  
+Knowing that the target of their mindreading  is about to attempt to engage in joint action with another agent might enable them to identify the goal of their target's actions. 
+We return to this issue in section \vref{sec:observing_joint_action}. 
+For now what matters is just that only interacting mindreaders can exploit your-goal-is-my-goal.
 
@@
 %As we saw, this problem is particularly likely to arise in cases involving novel tools and communicative gestures.
-Showing how your-goal-is-my-goal makes  avoiding this problem possible is a way of demonstrating the potential value   of interaction for mindreading.
+Showing how your-goal-is-my-goal makes it possible to avoid this problem is a way of demonstrating the potential value   of interaction for mindreading.
 
 In our earlier example 
-a novice parent is struggling to lift his heavy twin pram onto a bus when a stranger joins in and they lift the pram together (\vpageref*{twin_pram}).
-Suppose the stranger starts tipping the pram in a way that the inexperienced parent fails to recognise as a means, indeed the only means, of getting it onto the bus.
+a novice parent is struggling to lift his heavy twin pram onto a bus when a stranger joins in and they lift the pram together (see page \pageref*{twin_pram}).
+Suppose the stranger starts tipping the pram in a way that the novice parent fails to recognise as a means, indeed the only means, of getting it onto the bus.
 Outside the context of joint action, 
@@
 But it will do sometimes, and this is all we need.
-We have identified evidence for goal ascription that is available independently of a goal ascriber's knowing which ends actions are means to, and we have illustrated how this will sometimes enable mindreaders to escape the problem of opaque means
+We have identified evidence for goal ascription that is available independently of a goal ascriber's knowing which ends actions are means to, and we have illustrated how this will sometimes enable interacting mindreaders to avoid the problem of opaque means
 
@@
 knowing that she will perform actions directed to achieving this outcome.  
-We then signal to Ayesha that we are willing to act jointly with her.  
+We then signal to Ayesha that we will act jointly with her.  
 Now she is in a position to know what the goal of our action will be when we deploy the tool.  
@@
 because she knows that this is her goal
-and that we were willing to engage in joint action with her.
+and that we were attempting to engage in joint action with her.
 This is one illustration of how 
@@
 and 
-that we had exploited the my-goal-is-your-goal inference.
+that we had exploited the your-goal-is-my-goal inference.
 Then we would believe,
@@
 merely associating a gesture with its referent falls short of being able to respond appropriately.
-But if a mindreader supposes that her target is willing to engage in joint action with her,
+But if a mindreader supposes that her target is attempting to engage in joint action with her,
 then she may infer that the goal of her target's action is her goal
@@
 But we have shown that there is another possibility:
-the referential stance may 
+ referential understanding may 
 emerge from
-the teleological stance
+the teleological understanding 
  together with abilities to engage in simple forms of joint action.
@@
 anything's actual evolution or development.
-%Our point is just to illustrate one of several ways in which interaction may matter for mindreading.
-But widening the evidential basis for goal ascription
-to include evidence available only to mindreaders actually or apparently capable of interacting with their targets
+But widening the evidential basis for goal ascription to include facts about interactions between mindreaders and their targets 
 does reveal the coherence of  
@@
 
+\section{Observing interaction}
+\label{sec:observing_joint_action}
+We have been assuming that the your-goal-is-my-goal inference is  available to interacting mindreaders only. 
+This is required for our argument that 
+%some routes to knowledge of the goals of actions are available only to interacting mindreaders. 
+an interacting mindreader could be in a position to know things  she would be unable to know were she a mere observer.
+We note, however, that a mere observer can use a related inference.%
+\footnote{
+Here we are  indebted to an anonymous referee.
+} 
+To illustrate, return once more to our example of a novice parent who, struggling to get his twin pram onto a bus, is about to act jointly with a stranger.
+Now consider a third person observing the novice parent and the stranger. 
+Suppose this third person can manifestly only observe what  happens. 
+Even so, she might know that the stranger is about to attempt to engage in joint action with the parent, and she might come to know this in roughly the way that the parent does. 
+%So there is a sense in which the third person, who merely observes, is in  the same position as the novice parent.  
+%For each, knowing that the parent's actions are directed to lifting the pram onto the bus could, given some background conditions, put him or her in a position to know that this will be the goal to which the stranger's actions are directed.
+
+To generalise, the following inference characterises a route to knowledge of others' goals which could be exploited by a mere observer:
+\begin{enumerate}
+\item She is about to attempt to engage in some joint action or other with him.
+
+\item This is the single goal to which his actions will be directed.
+
+\end{enumerate}
+%
+Therefore:
+%
+\begin{enumerate}[resume]
+%
+\item A goal of her actions will be his goal, the goal to which his actions will be directed.
+\end{enumerate}
+%
+Call this inference \emph{her-goal-is-his-goal}.  
+This inference shows that mere observers who understand interaction also have a route to knowledge of other's goals that avoids the problem of opaque means. 
+The significance of interaction for mindreading extends beyond cases in which the mindreader herself interacts.
+
+Is this an objection to our claim that an interacting mindreader could be in a position to know things  she would be unable to know were she a mere observer?
+No, for the above inference differs from the earlier your-goal-is-my-goal inference (\vpageref*{your_goal_is_my_goal}) in requiring knowledge of another's goal. 
+The second premise makes this explicit.
+Where your-goal-is-my-goal depends on a mindreader's knowing to which goal her own actions will be directed, this inference depends on the mindreader's knowing to which goal another's actions will be directed.
+Now there could be---and surely are---cases where a mindreader knows to which goals her own actions are directed even though if another were to act as she is acting, mimicking her movements, she would be unable to recognise the goal to which these actions are directed.
+So there could be situations where an interacting mindreader could exploit your-goal-is-my-goal even though if she were only observing she would be unable to exploit her-goal-is-his-goal.  
+This is sufficient to support our claim that an interacting mindreader could be in a position to know things  she would be unable to know were she only observing.
+
 \section{The problem of false belief}
 \label{sec:false_belief}
-So far we have been arguing for the value of interaction for goal ascription
-on the grounds that 
-interacting mindreaders, unlike mere observers, have 
-a route to knowledge of the goals of others' actions
-which avoids the problem of opaque means.
+So far we have been arguing that interaction can matter for goal ascription
+because it  makes available routes to knowledge of the goals of others' actions
+which avoid the problem of opaque means.
 There is another reason for supposing that interaction matters for goal ascription and (at least indirectly) for mindreading more generally.
@@
 But of course if you had believed instead that the cat was in the south box,
-then, in acting on the same goal, you would have reached for the south box.
+then, in acting with the same goal, you would have reached for the south box.
 Now consider Ayesha who is observing your actions.
@@
 So despite the difference in belief,
-the possibility of interaction means that
+the possibility of interaction may mean that
 Ayesha can knowledgeably identify the goal of your action.
@@
 
-In short, then, abilities to engage in joint action open a route to knowledge of the goals of other agents' actions which does not depend on the goal ascriber knowing what her target believes even when their beliefs relevantly differ.
+In short, then, abilities to engage in joint action make available a route to knowledge of the goals of other agents' actions which does not depend on the goal ascriber knowing what her target believes even when their beliefs relevantly differ.\footnote{
+The her-goal-is-his-goal inference (see section \vref{sec:observing_joint_action}) 
+shows that mindreaders who only observe may also be able to exploit a route to knowledge of the goal of another agent's actions which does not depend on knowing what that agent believes but rather on knowing what a further agent interacting with the first believes.
+}
 This is not because abilities to engage in joint action provide a way to avoid the problem of false beliefs altogether.
@@
 Our suggestion was this.
-Where a mindreader recognizes that her target is willing to engage in joint action
+Where a mindreader recognizes that her target is about to attempt to engage in joint action with her 
 and where the mindreader is unwilling to change which goals her own actions will be directed to,
@@
 that the goals of her target's actions will be the goals of her own actions.
-This in outline is the `your-goal-is-my-goal' route to  knowledge (see  section \vref{sec:your_goal_is_my_goal}).
+This is the your-goal-is-my-goal route to  knowledge (see  section \vref{sec:your_goal_is_my_goal}).
 
@@
 Our point in mentioning it is merely to illustrate how
- better understanding the role of interaction in goal ascription and  mindreading more generally
+ better understanding the roles of interaction in goal ascription and  mindreading more generally
 might eventually lead to discoveries
@@
 %	encounter anomalies
-which they might not be able to know 
+which they might not have been able to know 
 %or encounter
- as mere observers.
+if they were only observing.
 % matters because it enables us to better understand the evidential basis of mindreading---%
@@
 
-\bibliography{$HOME/endnote/phd_biblio}
+\bibliography{phd_biblio}
 

### 2012-06-20 developing mind/chapter_01.tex
various

--- 2011-12-30 joint action and knowing others minds/interacting mindreaders.2.tex
@@
 
-\documentclass[14pt,a4paper]{extarticle}
+\def \papersize {a4paper}
+\documentclass[12pt,\papersize]{extarticle}
 % extarticle is like article but can handle 8pt, 9pt, 10pt, 11pt, 12pt, 14pt, 17pt, and 20pt text
 
-\def \ititle {Interacting Mindreaders}
+\def \ititle {A}
 \def \isubtitle {}
@@
 \def \iemail{s.butterfill@warwick.ac.uk}
-%\date{}
+%\def \iauthor {}
+%\def \iemail{}
+\date{}
 
 \input{$HOME/Documents/submissions/preamble_steve_paper2}
+%\author{}
 
@@
 
+\tolerance=5000
+
 \maketitle
@@
 \begin{abstract}
-Could interacting mindreaders be in a position to know things which they would be unable to know if they were manifestly passive observers?
-This paper defends a positive answer.
-Mindreading is sometimes reciprocal: 
-the mindreader's target reciprocates by taking the mindreader as a target for mindreading.
-The paper explains how such reciprocity can significantly narrow the range of possible interpretations of behaviour 
-where mindreaders are, or appear to be, in a position to interact.
-It follows that
-revisions and extensions 
-are needed for 
-standard theories of mindreading
-on which
-all the evidence relevant to mindreading 
-is in principle available to manifestly passive mindreaders.
-The view also has consequences for 
-understanding 
-how abilities to interact combined with comparatively simple forms of mindreading
-may explain the emergence,
-in evolution or development,
-of sophisticated forms of social cognition.
-
-%
-%What evidence grounds ascriptions of thoughts and actions,
-%and how does the evidence support the ascriptions?
-%In answering this question, 
-%philosophers sometimes focus on mere observation, ignoring interaction.
-%The only evidence considered is 
-% evidence that would be available to mindreaders 
-%observing their targets but in no position to interact with them.
-%The present paper,
-%which focuses on goal ascription,
-%argues that this is a mistake
-%and
-%identifies evidence 
-%available only to mindreaders capable of interacting with their targets.
-%Being poised to interact with others may make it possible to know things about their thoughts and actions which one might not otherwise be in a position to know.
-%This has consequences for the possible roles of interaction in explaining the evolution and development of mindreading. 
-%
-%What evidence grounds ascriptions of goals to other agents' actions,
-%and how does the evidence support the ascriptions?
-%It is usually assumed that, in answering this question, 
-%we can focus on observation and ignore interaction.
-%We can restrict ourselves to evidence that would be available to mindreaders who observe their targets but are not necessarily in a position to interact with them.
-%The present paper identifies evidence 
-%which supports goal ascriptions
-%and 
-%is available only to mindreaders who are capable of interacting with their targets.
-
+\noindent
+***
 \end{abstract}
 
-\section{Mindreading: the normative project}
-\label{sec:intro}
+\section{The puzzle}
+\label{sec:puzzle}
 
-Mindreading is
-the process of identifying thoughts and actions on the basis of bodily movements,
-somewhat as reading is the process of identifying propositions on the basis of inscriptions \citep[p.\ 4]{Apperly:2010kx}.
-Contrast
-a mindreader who is, or appears to be, capable of interacting with her targets 
-and
-a mindreader who can manifestly only observe.
-Is it possible that the interacting mindreader is in a position to know things which she would be unable to know if she were unable to interact with her targets?
-Our aim in this paper is to argue that the answer is a qualified `yes'.
+Plan:
 
-The question is about the evidential basis of mindreading, 
-not about what sorts of mechanisms are involved.
-While philosophers have engaged with questions about mechanisms (such as whether mindreading involves a process of simulation or of theorizing or some combination of the two),
-comparatively little effort has recently been devoted to issues about what evidence could ground mindreading.
+Simple theory of the mind
 
-Our question is part of a broader question,
-What is the evidential basis for ascriptions of thought and action and how does the evidence support the ascriptions?
-%\footnote{
-%See e.g.\ \citet[pp.\ 126-7]{Davidson:1973jx}.
-%}
-%To stress that this question is not directly about about \emph{how} anyone actually ascribes thoughts or identifies actions% 
-%% nor directly about the mechanisms, processes or representations involved in mindreading
-%, Davidson 
-%sometimes formulates the question by asking what someone \emph{could} know that \emph{would} put them in a position to identify another's thoughts and actions.%
-%\footnote{
-%See, e.g., \citet[p.\ 126]{Davidson:1973jx}.
-%While what follows draws  on Davidson's questions and insights,
-%our aims here are more modest than his.
-%For here we are concerned only with a  fraction of the problem of ascribing mental states and meanings,
-%and we are not concerned with larger claims about mind, meaning and truth 
-%(on which see 
-%\citet{Davidson:1990du} and 
-%\citet{lepore_donald_2005}).
-%}
-The most sustained attempts to answer this question,
-Davidson's (\citeyear{Davidson:1984wh}; \citeyear{Davidson:1990du}), Lewis' (\citeyear{lewis:1974ri}) and Dennett's (\citeyear{Dennett:1987sf}),
-do not exploit the possibility of interaction.
-The evidence and principles they consider are available to manifestly passive observers.
-So on their theories,
-a purely passive mindreader observing from behind a one-way mirror
-is on a par with
-a mindreader who, actually or apparently, could interact with those she seeks to interpret.
-The two are on a par in this sense:
-in principle the same evidence could be available to each, and each can exploit the same principles in moving from evidence to ascriptions of thought.
-Of course none of these theories is incompatible with the idea that interaction might be useful for mindreading in practice.  
-The point is just that, on these theories, interaction makes no difference to the theory.
-In this paper we aim to show
-that mindreaders actually or apparently capable of interacting with their targets are at an advantage not only in practice but also in theory.
-Their ascriptions could rely on evidence and principles which would be unavailable if they were entirely passive observers.
+When can infants first represent objects they can't see
 
-Why suppose, in advance of considering the details, that interacting mindreaders might know more?
-A mindreader's target is often also a mindreader and may sometimes reciprocate by taking the mindreader as a target for mindreading.
-%Where this happens we have mindreaders mindreading and being mindread by mindreaders.
-%The mindreader is her target's target for mindreading.
-It ought to be possible, in mindreading, to make use of this reciprocity.
-%of the fact that the mindreader's target reciprocates by taking the mindreader to be a target for mindreading.
-But how could such reciprocity facilitate mindreading?
-If we assume the mindreader is merely observing her target,
-that there is no potential for interaction,
-then it seems that any way of exploiting reciprocity would involve higher-order ascriptions. 
-The mindreader would ascribe to her target beliefs (say) about the mindreader's own beliefs and other mental states.
-And if her target reciprocates, she might escalate by ascribing to the target beliefs about her own beliefs about the target's beliefs about her beliefs.
-While this might be useful in some situations, 
-the nesting this approach requires quickly becomes dauntingly complex.
-And the basic intuition about reciprocal mindreading goes unsatisfied.
-Reciprocal mindreading should sometimes result in 
-something like a meeting of minds
-rather than
-an escalation of higher-order ascriptions.
-Perhaps, then, fully exploiting reciprocity in mindreading
-requires the mindreaders to be in a position to interact with each other.
-Perhaps in some cases 
-being or appearing poised to interact
-can somehow enable mindreaders to exploit reciprocity without first having to ascribe of higher-order mental states.
-This is the hunch we develop in what follows.
+Generalisation
 
-Even if the hunch turns out to be right, why investigate it?
-One reason is that the investigation will enable us to revise and extend existing accounts of the evidential basis of mindreading  in ways that make them more accurate and comprehensive.
-Another motive concerns the emergence, in evolution and in development, of mindreading.
-Several researchers have offered quite general conjectures about how interaction might explain the emergence of sophisticated forms of cognition.
-One view is that needs to interact with others have driven and shaped some aspects of cognition.%
-\footnote{
-E.g.\ \citet[p.\ 103]{Knoblich:2006bn} suggest that 
-`functions traditionally considered hallmarks of individual cognition originated through the need to interact with others'
-and that
-`perception, action, and cognition are grounded in social interaction.'
-}
-We shall not speak to this view here.
-Another view is that abilities to interact with others
-may have fostered the emergence in evolution---or, on a related view, in development---%
-of sophisticated forms of cognition including social cognition and, in particular, mindreading.%
-\footnote{
-\citet[p.\ 1]{Moll:2007gu}
-argue for the `Vygotskian Intelligence Hypothesis' according to which `the unique aspects of human cognition ... were driven by, or even constituted by, social co-operation.'
-See also
-	\citet{Hughes:2004zj},
-	\citet{Hughes:2006fu},
-	\citet{Tomasello:2007gl} and
-	\citet{tomasello:2008origins}.
-}
-Studying how interaction broadens the evidential basis of mindreading will eventually point to one way of filling in some details on just how interaction might facilitate the emergence of sophisticated forms of mindreading.
+Preview of cases
 
-Our claim needs to be qualified in several ways.
-First, we should highlight something already explicit but not yet emphasised:
-the claim applies not just to 
- mindreaders who actually could interact with their targets
-but also to
- mindreaders who appear to their targets to be potential interaction partners (even if they are not).
-Trading the risk of a minor misunderstanding for concision we use the term `interacting mindreader' to refer to both groups;
-similarly, references to mindreaders who `only observe' must be understood as excluding those who appear to their targets to be interaction partners.
-Only cases involving the first group---mindreaders who actually could interact---are likely to be of wider interest.
-Cases involving misleading appearances might be regarded as artefacts of the model.
-Any application of our claim to understanding evolution or development is bound to focus on cases where mindreaders actually could interact with their targets.
+Preview of remedies
 
-A second qualification concerns the scope of mindreading.
-In most discussions of mindreading, the focus is on  ascription of beliefs and other mental states to individuals.
-But in what follows we shall focus on the ascription of goals to actions.
-Some might claim that goal ascription is not mindreading, perhaps because 
-identifying relations between actions and the outcomes to which they are directed
-does not necessarily involve ascribing mental states.
-Our view is consistent with this claim (and with its negation).
-\label{goal_ascription_verify}
-Evidence for ascriptions of belief and of other mental states often includes the occurrence (and non-occurrence) of goal-directed actions.
-Identifying these actions typically requires goal ascription.
-In some cases, then, 
-evidence for the ascription of a goal
-will indirectly support the ascription of a belief or other mental state---%
-the evidence will support the ascription of a goal to an action,
-and this ascription will in turn support the ascription of a belief (say) to the agent of the action.
-Accordingly, even those who deny that goal ascription is mindreading should agree that sometimes evidence for a goal ascription is indirectly evidence for the ascription of a belief or other mental state.
-So whether or not goal ascription is deemed to be mindreading,
-a mindreader's access to evidence for mindreading  depends in part on her access to evidence for goal ascription.
-Our plan is to show that interacting mindreaders could make use, in ascribing goals, of evidence and principles which are not available to mere observers (who do not even appear to be in a position to interact with their targets).
-In doing this we will be showing that
-mindreaders who are (or appear) poised to interact with others 
-could know things about their minds 
-which they might not otherwise be in a position to know.
+\section{Why}
+Why study philosophy of mind and developmental psychology together?
 
-%Because the leading theories have supposed that being able to interact makes no difference,
-%they have treated the evidential basis for mindreading as more narrow than it truly is.
-%This matters
-%because broadening the evidential basis will enable us to give a more accurate and comprehensive theory of mindreading.
-%And, as we shall argue, this in turn matters because it may bear on the possibility of explaining the emergence, in evolution or in development, of mindreading.
-
-Developing these ideas requires us to fill in some background on goal ascription and its limits, as well as on notions of goal-directed interactions.
-Readers impatient to get to the central idea should skip to section \vref{sec:your_goal_is_my_goal}.
-
-\section{Goal ascription}
-Purposive action is action directed to the realisation of one or more outcomes.
-Goal ascription is the process of identifying to which outcomes others' purposive actions are directed.
-To illustrate, suppose that
-Hannah kicks a ball thereby both preventing her sisters from scoring and also breaking a window.
-Asked about episode,
-Hannah might protest that the goal of her action was not to break the window but only to reverse the others' advance.
-As this illustrates,
-among the actual and possible outcomes of an action,
-only some are outcomes to which the action was directed.
-Goal ascription is the process of identifying from among an action's actual and possible outcomes
-those to which it is directed.
-%Our immediate question is therefore, 
-%What evidence could support hypotheses about the outcomes to which actions are directed, 
-%and how would the evidence support the hypotheses?
-
-We focus on goal ascription 
-partly because this simplifies our argument,
-but mainly because goal ascription is widely thought to be among the very earliest components of mindreading to emerge (or, if goal ascription is not mindreading, then it is a late precursor).%
-\footnote{
-See, for example,
-\citet{Gergely:1995sq} and 
-\citet{Woodward:1998dm}.
-See also
-\citet[p.\ 111, Box 1]{Baillargeon:gx}
-on two subsystems
-and 
-\citet{Povinelli:2001jf} on `behavioural regularities' whose specification sometimes appears to involve goal-directed action.
-}
-By showing that
-  interacting mindreaders may have access to 
-  evidence for goal ascriptions 
-  which is unavailable to those who merely observe,
-we may be able to indicate ways in which interaction could facilitate the emergence, in evolution or development, of more sophisticated forms of mindreading.%
-\footnote{
-There are various ways of understanding the idea that mindreading may come in several forms, some less conceptually or cognitively demanding than others.
-For a range of views, see
- \citet{Apperly:2009ju},
- \citet{Call:2005qe},
- \citet{Doherty:2006wz},
- \citet{ONeill:2005ff}
- and
- \citet{Wellman:2001if}.
-}
-
-Because goal ascription has received little attention,
-in this section
-we shall 
-briefly review some potential benefits of being able to identify goals
-and then 
-consider what sort of evidence might support goal ascription.
-
-It is a familiar idea that goal ascription enables one to learn from others' successes.
-For example,
-if you know or can guess that another agent's actions are directed to opening a nut,
-you may then be in a position to infer that the unfamiliar pattern of actions she is performing constitute a means to open nuts.
-%\citep{Horner:2005pj}
-In a different case, knowing the goal of another agent's actions may enable you to discover a new use for a tool.
-A slightly less familiar idea is that goal ascription enables one to learn from others' failures as well as their successes.
-For example, suppose that while you are searching for some peanuts 
-another agent attempts but fails to reach for a closed container.
-In some circumstances,
-if you know that the goal of the agent's action was to obtain the peanuts
-then you now have evidence as to where they might be.\footnote{
-\citet{hare_chimpanzees_2004} exploit this principle in testing chimpanzees' abilities to ascribe goals.
-}
-This is one illustration of how goal ascription could in principle enable us to learn from others' failures 
-(\citealp{Want:2001hp} offer another).
-
-Goal ascription enables one
-to predict and manipulate others' actions.
-If you know that an agent is engaged in a sequence of actions whose eventual goal involves retrieving some object,
-you may be able to predict that the agent will go to where the object is at some point.\footnote{
-For an application see \citet{Hare:2001ph}.
-}
-Equally, in some cases knowing this much about the goal of another's actions many enable you to assist them,
-either 
-by retrieving the object for them 
-or else
-by revealing the object's location to them.%
-\footnote{
-For a paradigm involving the former see \citet{warneken:2007sa};
-on the latter, see \citet{Liszkowski:2008al}.
-}
-
-Goal ascription is also instrumental for ascribing propositional attitudes.
-We have already mentioned that such ascriptions often involve 
-confirming predictions about actions 
-which in turn typically involves
-identify the goals of those actions (\vpageref*{goal_ascription_verify}).
-In addition,
-knowing which outcomes an action is directed to may constrain hypotheses about what an agent intends 
-as well as
-potentially providing information concerning what the agent knows, believes or desires.
-For example,
-if we know that the goal of an agent's action is to retrieve some peanuts,
-and if we also know where all the peanuts are,
-we may be able to infer that she does not know where the peanuts are,
-or that she falsely believes that some of the peanuts are over there.%
-\footnote{
-\citet{Wimmer:1998kx} exploit this possibility in testing children's abilities to ascribe false beliefs.
-}
-(Of course this can also work the other way:
-information about an agent's beliefs or other mental states may support conclusions about her goals.
-Belief- and goal-ascriptions are mutually constraining.)
-
-Now that we have reviewed some of the benefits goal ascription can bring,
-what evidence could support ascriptions of goals to actions?
-Consider the claim that 
-knowledgeably identifying the goals to which actions are directed
-depends on 
-knowledge of the agents' intentions
-which in turn depends 
-on knowledge of their beliefs, desires and other mental states.
-If this claim were true,
-there would be no such thing as a theory of the evidential basis of goal ascription,
- only a theory of the evidential basis of mindreading more generally.
-
-The falsity of this claim seems to be assumed in both developmental and comparative research on goal ascription.%
-\footnote{ 
-Compare \citet{Gergely:1995sq},
-	\citet{Woodward:1998dm} and
-	\citet{Penn:2007ey}
-among many others.
-}
-We know of no compelling reason for taking this claim to be true.
-After all,
-it is consistent with rejecting the claim to recognize that
- ascriptions of goals to actions
- constrain, and are constrained by,
- ascriptions of intentions and other mental states.
-The existence of such constraints does not show that one could 
-not know something about the goals of actions 
-while  knowing nothing about the agents' mental states.
-Further,
-there do seem to be situations where knowledge of agents' goals does not require any knowledge of their mental states.
-For instance,
-suppose that two people are sitting opposite each other at a low table
- which is 
-sparsely populated with objects.
-The objects are all out in the open; manifestly, both can clearly see them.
-If one person reaches to grasp one of these objects (the duck, say), 
-must the other ascribe beliefs or other mental states in order to knowledgeably identify the goal of her action as that of grasping the duck?
-On the face of it, she need not.  
-Even if she had no ability to ascribe mental states, it seems she might nevertheless be in a position to identify the goal of the other's action.
-
-While not decisive,
-these considerations are perhaps sufficient to motivate 
-exploring what evidence might ground goal ascription.
-According to 
-what Csibra and Gergely call `the principle of rational action':
-%
 \begin{quote}
-`an action can be explained by a goal state if, and only if, it is seen as the most justifiable action towards that goal state that is available within the constraints of reality.'%
-\footnote{
-\citet[p.\ 255]{Csibra:1998cx}; cf.\ \citet{Csibra:2003jv}.
-A related but different `principle of efficiency' has been formulated by \citet[p.\ 1061]{Southgate:2008el}:
-`goal attribution requires that agents expend the least possible amount of energy within their motor constraints to achieve a certain end.'
-}
+`Naturalism in philosophy …\ has the goal of articulating the application conditions of puzzling concepts (like knowledge and perception) so that empirical (scientific) methods can be used to answer questions  …\ Naturalism in epistemology is merely the attempt to get clear enough about what we mean when we talk about knowledge and perception to be able to tell—in ways a biologist or an experimental psychologist would recognise as scientifically respectable—whether what we are saying is true or not' \citep[p.\ x]{Dretske:2000ky}
 \end{quote}
-%
-Taking this idea as a rough starting point,
-we propose that
-%goal ascription could sometimes involve 
-%identifying which potentially desirable outcomes 
-%an observed action is the best available means to achieving.
-%More carefully, our proposal is that 
-these facts:
-%
-\begin{enumerate}
-%
-\item action $a$ is directed to some goal;
-%
-\item actions of $a$'s type are normally capable of being means of realising outcomes of $G$'s type in situations with the salient (to any concerned) features of this situation;
-% 
-\item no alternative type of action is both 
-typically available to agents of this type 
-and also 
-such that actions of this type would be normally be significantly better\footnotemark \ means of realising outcome $G$ in situations with the salient features of this situation;
-\footnotetext{
-An action of type $a'$ is a better means of realising outcome $G$ in a given situation than an action of type $a$ if, for instance, actions of type $a'$ normally involve less effort than actions of type $a$
-in situations with the salient features of this situation 
-and everything else is equal; 
-or if, for example, actions of type $a'$ are normally more likely to realise outcome $G$ than actions of type $a$
-in situations with the salient features of this situation 
-and everything else is equal.
-}
-%
-\item the occurrence of outcome $G$ is typically desirable for agents of this type;
-%
-\end{enumerate}
-%
-and
-%
-\begin{enumerate}[resume]
-\item there is no other outcome, $G'$, 
-the occurrence of which would be at least comparably desirable for agents of this type 
-and where (2) and (3) both hold of $G'$ and $a$
-%
-%where
-%	the occurrence of $G'$ would be at least comparably desirable for agents of this type 
-%%where
-%%	$G$'s occurrence is not a means to $G'$'s nor conversely,
-%and
-%where  
-%	(2) and (3) both hold of $G'$ and $a$
-\end{enumerate}
-%
-may jointly constitute defeasible evidence for the conclusion that:
-%
-\begin{enumerate}[resume]
-\item $G$ is a goal to which action $a$ is directed.
-\end{enumerate}
-%
-We suggest that the above inference, from (1)-(5) to (6), is a route to knowledge of the goals of actions in this sense:
-it would be possible to know the premises without already knowing the conclusion;
-and, 
-in some cases knowing the premises could put one in a position to know the conclusion.%
-\footnote{
-Knowledge of the premises may not require explicit knowledge of what others find obvious, what is salient to them or what makes actions better for them.
-It is arguably sufficient that
-the individual ascribing a goal 
-is entitled to 
-rely on being sufficiently similar to the target of her ascription 
-with respect to 
-	what she finds obvious, 
-	what makes one action a better or worse means of realising an outcome than another action for her,  
-	what is desirable to her,
-	and what is salient to her.
-}
-
-Why accept this?
-Suppose that 
-	(1)-(5) are facts 
-and that
-	$G$ is not a goal to which action $a$ is directed.
-Then from (1) we know that $a$ has a goal, call it $G'$.
-And from (5) we can infer that either 
-(i) 
-the action, $a$, was not of a type that is normally capable 
-of being a means to the outcome to which it was directed,
-or
-(ii) 
-a significantly better (for instance, a more reliable or less effortful) means of achieving $G'$ is typically available to agents of this type,
-or
-(iii)
-$G'$'s occurrence would be significantly less desirable than $G$'s occurrence for agents of this type.
-So the agent's action fell short of being appropriate or optimal in some way,
-or the agent has atypical preferences or capabilities,
-or the situation is not normal.
-While any of these is possible---%
-and perhaps in some cases even to be expected---%
-there are also circumstances in which it is reasonable to suppose that none obtains,
-that the situation will be normal in relevant respects,
-that agents will have typical preferences and capabilities 
-and that they will act in ways that are appropriate and (in the limited sense in play here) optimal.
-In such circumstances, the above inference can serve as a route to knowledge of the goals of actions.
-The existence of this route to knowledge shows that goal ascription doesn't invariably depend on mental state ascription.
-
-Note that we are not suggesting that the above inference captures the only route to knowledge of the goals of actions,
-or that it provides anything like a comprehensive theory of the evidential basis of goal ascription.
-Clearly it does not.
-The inference has only limited applications.
-For example, it cannot be used when an action is the best available means of achieving two or more comparably desirable outcomes. 
-Fortunately the argument that follows does not depend on having a comprehensive theory of evidence for goal ascription.
-The point of this brief discussion was only to 
-consider the sort of evidence that can bear on goal ascription
-and to 
-show that 
-even those incapable of ascribing mental states like belief 
-might nevertheless have evidence sufficient for goal ascription.
-
-Our narrow aim in what follows is to show that some routes to knowledge of the goals of actions 
-are available only to interacting mindreaders
-and not to those who merely observe.
-Having in this section introduced goal ascription as a topic,
-the next step is to identify an obstacle to acquiring knowledge of the goals of actions,
-one which we will eventually show can be avoided where mindreaders can (or appear to be able to) interact with their targets.
-
-\section{The problem of opaque means}
-\label{sec:opaque_means}
-
-While we lack a detailed theory of the evidential basis of goal ascription,
-it is certain that the evidence for goal ascription sometimes includes considerations about which ends actions are means to.
-Suppose an observer faces an action but cannot identify ends to which it could be a means.
-This may prevent her from recognizing the action's goal%
-\footnote{
-It is possible that some actions have more than one goal.
-To reduce parenthetical qualifications we shall write as if actions had only one goal.  
-All of our key claims and arguments are consistent with the possibility of actions with more than one goal.
-} 
- by depriving her of evidence.
-To illustrate, contrast two cases of tool use.
-In one case, someone uses a reamer to  juice  a lime; in the other, someone else scores shag with a lame to prevent a loaf from cracking.
-Without communication, repetition or convention,
-an observer familiar with reamers but not lames 
-may be able to identify the goal of the first action only.
-As this illustrates, ignorance about to which ends actions are means can be an obstacle to goal ascription.%
-Call this the problem of opaque means.
-
-We are not suggesting that no observer could ever identify the goal of any action she fails to recognise as a means to achieving that goal.
-Of course opaque means are not in every case an insurmountable obstacle to goal ascription, and
-they may only rarely be a problem for human adults with sophisticated social skills.
-Our point is neither novel nor surprising:
-opaque means sometimes  deprive mindreaders of evidence and so prevent goal ascription.
-This is more likely to happen where  goal ascribers lack sophistication in mindreading, communication and culture.
-
-Some of the most plausibly unique aspects of human cognition depend on our abilities to recognise the goals of novel behaviours involving tools, and of communicative gestures.
-The problem of opaque means is likely to arise in both cases
-if goal ascription is based  entirely on observation 
-(so that the possibility of interaction is ignored)
-and if the goal ascriber lacks sophisticated social skills.
-We have just seen an illustration of how the problem of opaque means arises where tools are used to unfamiliar ends.
-Relatedly, it is also likely to arise where actions involve multiple steps that do not form a familiar sequence, can occur in various orders and can be interspersed among other activities;
-as in preparing spirit from grain, for example.
-%\footnote{
-%*cut because 'plausibly unique aspects of human cognition'*
-%one example is food preparation in some groups of mountain gorillas
-%\citep[p.\ 531]{Byrne:2003wx}.
-%*This is slightly tricky because Byrne wants a pure behaviour reading account with no understanding of intention.  
-%Stress that this is consistent with supposing that goals are ascribed.
-%}
-
-The problem of opaque means also affects communicative actions 
- because these characteristically have  goals which the actions are means to realising only because others recognise them as means to realising those goals (a Gricean circle). 
-To illustrate, consider an experiment from 
-\citet[][experiment 3]{hare_chimpanzees_2004}
- whose two main conditions are depicted in figure \vref{fig:reach_point}.
-The pictures in the figure stand for what participants, who were chimpanzees, saw.
-The question was whether participants would be able to work out which of two containers concealed a reward.
-In the condition depicted in the left panel, participants saw a chimpanzee trying but failing to reach for the correct container. 
-Participants had no problem getting the reward in this case, suggesting that they understood the goal of the failed reach.
-In the condition depicted in the right panel, a human pointed at the correct container.
-Participants did not reliably  get the reward in this case, suggesting that they failed to understand the goal of the pointing action.%
-\footnote{
-The contrast between the two conditions is not due merely to the fact that one involves a human and the other a chimpanzee.
-Participants were also successful when the failed reach was executed by a human rather than another chimpanzee \citep[][experiment 1]{hare_chimpanzees_2004}. 
-}
-This may be because of the problem of opaque means.
-One theoretically possible explanation of these findings is that the participants could identify to which end a failed reach might be a means, but not to which end a communicative gesture might be a means.%
-\footnote{
-  \label{fn:communicative_intention}
-Hare and Tomasello 
-consider several explanations for their findings including `the hypothesis that chimpanzees do not understand the communicative intent of a cooperative-communicative experimenter' (\citeyear[p.\ 580]{hare_chimpanzees_2004}).
-\citet[pp.\ 5--7]{Moll:2007gu} argue for a hypothesis along these lines by appeal to a range of related findings.
-}
-Whatever the truth about the chimpanzees' performance,
-this possibility illustrates how the problem of opaque means can be an obstacle to exploiting communicative gestures.
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=12cm]{figure_hare_toma_2004_e3.png}
-\caption{
-\label{fig:reach_point}
-	A failed reach (left) and a helpful point (right).
-	Reproduced from \citet[p.\ 557, figure 4]{hare_chimpanzees_2004}.
-}
-\end{center}
-\end{figure}
-
-This, then, is the problem of opaque means:
-failures to identify to which ends actions are means can impair goal ascription.
-The problem is potentially a problem for mindreaders given the standard, purely observational models of mindreading.
-It is not our intention to suggest that the problem of opaque means is a problem \emph{for models of mindreading}.
-Rather, it is a potential problem \emph{for mindreaders}.
-
-That the problem of opaque means exists
-is the first premise in our argument
-that models of mindreading based on pure observation are less powerful than models taking into account the possibility of interaction.
-They are less powerful in this sense: 
-some routes to knowledge of the goals of actions are available only where a theory of the evidential basis of mindreading 
- takes the possibility of interaction into account.
-In what follows we shall explain how abilities to interact with others provide a route to knowledge of the goals of other agents' actions  which avoids the problem of opaque means.
-This is one way in which moving away from a purely observational model of mindreading yields a richer evidential base for ascriptions.
-
-\section{Interactions involving distributive goals}
-\label{sec:joint_action}
-Our overall aim is to defend this claim:
-there are routes to knowledge of the goals of others' actions
-which are closed to
-mindreaders who merely observe their targets
-but open to 
-mindreaders who are, or appear to be, capable of interacting with their targets.
-As a preliminary to defending this claim,
-we need to specify which types of interaction are relevant.
-That is the aim of this section.
-
-Let us stipulate that an outcome is a \emph{distributive goal} of two or more agents' actions just if two conditions are met.
-\label{df:distributive_goal}
-First, this outcome is a one to which each agent's actions are individually directed.
-Second, each agent's actions are related to the outcome in such a way that it is possible for all the agents (not just any agent, all of them together) to succeed in bringing about this outcome.
-
-To illustrate,
-suppose that, while doing some gardening,
-we find a turnip too big for any of us easily pull out of the ground alone.
-We each individually intend that we all pull up the enormous turnip
-and we act on this intention.%
-\footnote{
-For an argument that it is possible for each of us individually to intend that we do something, see \citet{Bratman:1999fr} and \citet[Chapter 2]{Bratman:2012fk}.
-}
-There is a single outcome,
-freeing the turnip,
-to which each of our actions is individually directed.
-And it is possible for all of us to succeed in bringing about this outcome.
-So this outcome is a distributive goal of our actions.
-
-%Contrast this episode with an alternative where no distributive goal is involved.
-%Instead of intending that we all pull up the enormous turnip,
-%each of us intends that she herself pull it up.
-%While all of our actions are directed to similar outcomes (they all involve pulling up the turnip),
-%these are clearly different outcomes.
-%They are clearly different because one could occur while another did not;
-%one of us might have broken an arm and been unable to pullwhile the others succeeded.
-%So in this case, freeing the turnip is not a distributive goal of our actions.
-%For some actions to have a distributive goal,
-%it is necessary that there be a single outcome (and not merely similar outcomes) to which they are directed.
-
-Note that our actions could have a distributive goal even without us each intending that we do something.
-As another variation on the turnip-pulling story,
-suppose that we each individually intend to 
-	contribute to pulling up the turnip or pull it up.
-Such intentions are agent neutral in this sense:
-relative to such an intention an agent 
-	could succeed by acting alone 
-and
-  she also could succeed by acting with unspecified other agents.
-Plausibly our having these intentions is sufficient for our actions to have a distributive goal.
-For in virtue of these intentions each of our actions is directed to the turnip's extraction,
-and this is an outcome relative to which it is possible that all of our actions succeed.
-
-In defining `distributive goal' we stipulated that it must be possible for all of the agents of actions with a distributive goal to succeed relative to that outcome.
-Consider a variation on our first turnip-pulling story.
-Instead of each intending that we pull up the turnip,
-we each simply intend  to pull up the turnip.
-In this case,
-some might claim that
-this amounts to us each intending that he or she pulls up the turnip.
-And this claim might be taken to support the further claim that
-there is no outcome to which each of our actions are directed
-in virtue of our having these intentions
-such that we could all succeed relative to this outcome.
-%, given our intentions in this alternative case, we could not all succeed relative to the outcome of pulling up the turnip---%
-%that if your actions were to succeed relative to the turnip's extraction,
-%then ours would fail.
-We take no stand on whether either claim is correct.
-For our purposes all that matters is that there are distributive goals.
-And the existence of distributive goals follows from the possibility of each of us 
-intending that we pull up the turnip
-(rather than simply
-intending to pull it up),
-and from the possibility of each of us
-intending to contribute to pulling up the turnip or pull it up.
-
-For two agents' actions to have a distributive goal it is sufficient that their actions constitute a joint action 
-(at least this is true on almost any account of joint action).%
-\footnote{
-\label{fn:joint_action_distributive_goal}
-We know of two definitions of joint action on which it is not straightforward that all joint actions involve distributive goals.
-One is provided 
- is provided by \citet[p.\ 366]{ludwig_collective_2007} and quoted in footnote \vref{fn:ludwig}; the other is due to 
-\citet[p.\ 70]{Sebanz:2006yq}.
-According to Sebanz and colleagues' `working definition', 
-`joint action can be regarded as any form of social interaction whereby two or more individuals coordinate their actions in space and time to bring about a change in the environment'
-(\citeyear[p.\ 70]{Sebanz:2006yq}).
-On one way of reading this, the latter part is equivalent to `a change in the environment is such that: to bring it about two or more individuals coordinate their actions';
- this does suggest that the actions have a distributive goal.
-On other standard accounts it is straightforward that joint action involves distributive goals.
-Compare 
-	\citet[pp.\ 329-31]{Bratman:1992mi},
-	\citet[pp.\ 96-7]{Searle:1990em} and
-	\citet[pp.\ 168-9]{gilbert:2009shared}.
-}
-However, the converse does not hold.
-As the first variation of the turnip-pulling story indicates,
-two or more agents' actions may have a distributive goal 
-even though the agents do not know about each others' intentions or actions.
-In fact our actions might have a distributive goal even though none of us is aware of this, or even of the others' existence. 
-(We shall refrain from illustrating this point by contriving a truly gigantic turnip and a completely dark, very stormy night; readers can probably guess how this would go.)
-One consequence of this is that 
-two or more agents' actions may have a distributive goal
-even though they are not engaged in joint action
-(at least not on any standard account of joint action).%
-\footnote{
-On many accounts of joint action, each agent involved in a joint action must believe, expect or know something about the jointness of her action.
-See 
-	\citet[p.\ 103]{Bratman:1993je}, %common knowledge
-	\citet[p.\ 40]{Butterfill:2011fk},
-	\citet[p.\ 10]{Kutz:2000si}, %Agents acting on a shared intention have `a conception of themselves as contributors to a collective end.'
-	\citet[p. 56]{miller_social_2001} % requires that each agent believes her actions are interdependent with the other agent's.
-	and
-	\citet[p.\ 361]{Roth:2004ki}.
-%: `each participant ...
-can answer the question of what he is doing or will be doing by saying for example ``We are walking together'' or ``We will/intend to walk together.'''
-Exceptions include  \citet{pacherie_framing_2011} and \citet{ludwig_collective_2007}.
-\label{fn:ludwig}
-According to Ludwig,
-`The concept of a joint action as such is just that of an event of which there are multiple agents' (\citeyear[p.\ 366]{ludwig_collective_2007}).
-Depending on what events are and what it is to be the agent of an event,
-it may turn out that,
-on Ludwig's definition,
-our actions' having a distributive goal 
-is sufficient for
-us to be engaged in joint action.
-}
-
-The notion of a distributive goal is a 
-narrowly technical one.
-In introducing this notion by stipulation we are not aiming to match anyone's intuitions about anything.
-What matters for now is just that the definition is theoretically coherent and that it is possible for actions to have distributive goals.
-Our aim in introducing distributive goals 
- is to capture a minimal requirement on the sort of interactions 
-relevant to characterising a route to knowledge of the goals of actions
-available only to interacting mindreaders.
-% 
-% relevant to explaining how 
-%interacting mindreaders 
-%might be able to exploit routes to knowledge of the goals of actions
-%which would be unavailable  were  they mere observers.
-
-\section{Your-goal-is-my-goal}
-\label{sec:your_goal_is_my_goal}
-If a mindreader is able to interact with her targets,
-if she is not limited to merely observing them,
-how might this provide her with a route to knowledge of the goals of their actions?
-The answer hinges on interactions involving distributive goals
-(as defined \vpageref{df:distributive_goal}).
-
-Here is an intuitive idea that doesn't quite work:
-if a mindreader is engaged in an interaction with her target that involves a distributive goal, 
-it may be easy for the mindreader to know what the goal of her target's actions is because this goal is the goal of her own actions.
-So if she knows the goal of her own actions and she knows that she is engaged with her target in an interaction involving a distributive goal,
-then she already knows what the goal of her target's actions are.
-%Roughly speaking, the mindreader can reason about her target thus: your goal is my goal.
-
-Of course this intuitive idea is no use it stands.  
-For the inference it captures relies on the premise that the mindreader and her target are engaged in actions with a distributive goal.
-But for the mindreader to know this premise it seems she must   already know which goal her target's actions are directed to.
-
-Fortunately there is a way around this.  
-For there are various cues which signal that one agent is prepared to engage in joint action with another,
-and joint actions involve distributive goals.  
-\label{twin_pram}
-Seeing a stranger with the haggard look of a new parent struggling to get his twin pram on to the bus, you grab the front wheels and make eye contact, raising your eyebrows and smiling.
-(The noise of the street rules out talking.)   
-In this way you signal that you are about to act jointly with the stranger, to lift the pram with him.   
-Since the stranger is fully committed to getting his pram onto the bus,
-he knows what the sole goal of his own actions will be.
-But he also knows that you will engage in joint action with him,
-which means that your actions and his actions will have a distributive goal.
-This may enable him to infer the goal of your imminent actions: 
-your goal is his goal, to get the pram onto the bus.
-
-Our suggestion, then, is that the following inference characterises a route to knowledge of others’ goals:
-%
-\begin{enumerate}
-\item You are willing to engage in some joint action\footnote{
-We leave open the issue of how joint action is to be characterised subject only to the 
-requirement that all joint actions must involve distributive goals (see further footnote \vref{fn:joint_action_distributive_goal}).
-Attempts to characterise joint action in ways relevant to explaining development include
-	\citet{Tollefsen:2005vh},
-	\citet{Carpenter:2009wq},
-	\citet{pacherie_framing_2011} and
-	\citet{Butterfill:2011fk}.
-The last of these 
-not only shows how a primitive form of joint action can be characterised without ascriptions of higher-order mental states
-but also shares our focus on relations between actions and goals rather than between agents and intentions.
-}
-or other with me
-%(for example, because you have made eye contact with me while I was in the middle of attempting to do something).
-
-\item I am not about to change the single goal to which my actions will be directed.
-
-\end{enumerate}
-%
-Therefore:
-%
-\begin{enumerate}[resume]
-%
-\item A goal of your actions will be my goal, the goal I now envisage that my actions will be directed to.
-\end{enumerate}
-%
-Call this inference \emph{your-goal-is-my-goal}.  
-To say that it characterises a route to knowledge implies two things.  
-First, in some cases it is possible to know the premises, 1–2, without already knowing the conclusion, 3.  
-Second, in some of those cases knowing the premises would put one in a position to know the conclusion.  
-We shall consider these points in turn
-
-Is it ever possible to know the premises without first knowing the conclusion?  
-Consider the first premise.  
-Sometimes 
-in the right contexts
-an individual can recognize in another's facial expressions, engaging gestures or synchronized bodily movements that she is about to engage in joint action with him.
-Exploiting these indicators does not typically depend on knowing the particular contents of any beliefs, desires or goals.  
-Expressions, gestures and movements can naturally indicate imminent jointness in much the way thay can also naturally indicate emotions.\footnote{
-Ideas along these lines are suggested by the discussion of {emergent coordination} in \citet{Knoblich:2010fk}.
-}  
-Of course these indicators provide no guarantee that others are genuinely willing to engage in joint action.  
-But they are sufficiently reliable to ground knowledge in some cases.  
-The existence of such indicators shows that knowing the first  premise of the above inference does not require already knowing which particular goals the other has.
-
-Not everything needs to rest on indicators, however.
-It is sometimes possible to know that others are willing to engage in joint action with you even without relying on such indicators.  
-Thanks to widespread dispositions to act jointly, in some situations it is reasonable to take for granted that others are willing to act jointly.  
-For example, this is often so for children struggling with a coat while surrounded by family or familiar adults.  
-And in at least some subcultures people using public transport can reasonably take for granted that, within limits, those around them will act jointly with them when the need is clear.
-Of course dispositions to engage in joint action may vary between cultural groups and situations.
-This may be fatal for the peripatetic cosmopolitan,
-but for others what matters is not whether the dispositions are universal but only that they are sufficiently widespread to be predictable.
-
-Turning to the second point,
-could knowing the premises of the your-goal-is-my-goal inference ever put one in a position to know the conclusion?  
-Of course the inference is not deductive and will only work when certain background conditions are met.  
-These background conditions include the other having largely true beliefs concerning which goal your actions are or will be directed to.
-After all, the other agent may be willing to act jointly with you while being entirely mistaken about the goals to which your actions will be directed.
-Where this happens, the premises of the inference might be true but the conclusion false.
-(We return to this point in section \vref{sec:false_belief}.)  
-But in some situations there is no requirement  to consider  this possibility, as for instance when stereotypes, conventions or simplicity should and do make the goal of your actions obvious to the other agent.
-Given the right conditions,
-that others are willing to engage in joint action with you 
-is sometimes sufficient reason to hold that
-you will end up acting jointly with them
- even where the goal to which your actions will be directed is already fixed.
-
-In short, then, the two requirements for the your-goal-is-my-goal inference to characterise a route to knowledge are met.  
-In some cases it is possible to know the premises without already knowing the conclusion thanks to natural expressions of willingness to engage in joint action.  
-And knowing the premises sometimes puts one in a position to know the conclusion thanks to the fact that, when things are going well, another's willingness to engage in joint action with you can be based on an accurate assessment of the goals of your actions.
-
-In principle,
-exploiting your-goal-is-my-goal does not require actually being in a position to interact with the target of goal ascription.
-It is sufficient (and would be necessary but for some special cases) to be taken by the target to be in a position to interact. 
-Our concern, however,
-is with cases that are likely to be important for understanding
-development or evolution.
-This motivates a focus on goal ascribers of limited sophistication
-who lack both deep insight into others' minds and fully-fledged communicative abilities.
-Such individuals are unlikely to be able to contrive or exploit situations in which they only appear to be in a position to interact with their targets.
-When we come to applications of your-goal-is-my-goal, the relevant cases will be those 
-in which a goal ascriber is manifestly in a position to interact with her target.
-
-The your-goal-is-my-goal route to knowledge is characterised by an inference.  
-However, exploiting this route to knowledge may not require actually making the inference or knowing the premises.  
-Depending on what knowing requires,
-it may be sufficient to believe the conclusion because one has reliably detected a situation in which the premises of the inference are true;
-it may not be necessary to think of this situation as a situation where the premises are true,
-nor even to be able to think of it in this way.
-Our aim in identifying the your-goal-is-my-goal inference is not to 
-defend a detailed hypothesis about  mechanisms of mindreading.
-Instead our present concerns are limited to a normative question about the evidential basis of mindreading.
-The your-goal-is-my-goal inference matters
-not because it describes exactly how anyone actually assigns goals (maybe it doesn't)
-but because 
-it characterises a route to knowledge
-that is closed to mere observers
-and open to interacting mindreaders.
-
-\section{Avoiding the problem of opaque means}
-
-The problem of opaque means was this:
- failures to identify to which ends actions are means can impair goal ascription (see section \vref{sec:opaque_means}).
-%As we saw, this problem is particularly likely to arise in cases involving novel tools and communicative gestures.
-Showing how your-goal-is-my-goal makes  avoiding this problem possible is a way of demonstrating the potential value   of interaction for mindreading.
-
-In our earlier example 
-a novice parent is struggling to lift his heavy twin pram onto a bus when a stranger joins in and they lift the pram together (\vpageref*{twin_pram}).
-Suppose the stranger starts tipping the pram in a way that the inexperienced parent fails to recognise as a means, indeed the only means, of getting it onto the bus.
-Outside the context of joint action, 
-failure to realise that the action is a means to getting the pram onto the bus
-might have the consequence that
-the parent's evidence on balance supports the conclusion that the goal of the stranger's actions is to take the pram off the bus.
-(Perhaps the stranger is impatient to get onto the bus herself.)
-%Or suppose the stranger starts by pulling the pram away from the bus in order to better position it for entry,
-%and the novice parent does not realise that this is necessary.
-%Outside the context of joint action, 
-%his evidence might on balance support the conclusion that the goal of the stranger's actions is to take the pram off the bus.
-%(Perhaps the stranger is impatient to get onto the bus herself.)
-But in the context of joint action,
-your-goal-is-my goal gives the parent  additional evidence for supposing that, 
-even though the stranger's actions do not seem to him to be a means to getting the pram onto the bus, 
-this really is the goal of her actions.
-Of course this additional evidence will not always outweigh   other evidence.
-But it will do sometimes, and this is all we need.
-We have identified evidence for goal ascription that is available independently of a goal ascriber's knowing which ends actions are means to, and we have illustrated how this will sometimes enable mindreaders to escape the problem of opaque means
-
-We saw earlier that the problem of opaque means  may impair goal ascription where actions involve novel uses for tools.
-How could your-goal-is-my-goal mitigate the problem in such cases?
-Imagine we are interacting with a young child,
-Ayesha,
-and want her to understand how a new tool is used.  
-It is difficult to convey this to her directly.  
-So we first get her interested in achieving an outcome that would require the new tool,
-knowing that she will perform actions directed to achieving this outcome.  
-We then signal to Ayesha that we are willing to act jointly with her.  
-Now she is in a position to know what the goal of our action will be when we deploy the tool.  
-She is able to identify this goal 
-despite being unable to recognize it as an end to which our tool-using action is a means.
-She is able to identify this goal 
-because she knows that this is her goal
-and that we were willing to engage in joint action with her.
-This is one illustration of how 
-interacting mindreaders have at their disposal ways of
-identifying the goals of actions involving novel uses of tools
-which are unavailable to mindreaders who can only observe.
-
-As this example indicates,
-exploiting your-goal-is-my-goal 
-can shift the burden of identifying goals from a mindreader to her target.
-In the example Ayesha is the focal mindreader and we are her target;
-but her success in identifying the goal of our actions
-depends on this,
-that our willingness to act jointly with her is based on \emph{our} knowledge of the goals of \emph{her} actions.
-In purely observational mindreading, the target's beliefs about the goals of the mindreader's actions are not normally relevant (except, of course, when the mindreader is ascribing such beliefs).
-But interacting mindreaders who rely on your-goal-is-my-goal
-thereby rely on their targets' having correctly identified the goals of their actions.
-Of course this is sometimes a reason not to rely on your-goal-is-my-goal.
-But where the target understands relevant means-ends relations,
-such as actions involving novel tools,
-the your-goal-is-my-goal route to knowledge of others' goals may sometimes be the only option.
-
-Here we have a first illustration of how
-interacting mindreaders can exploit reciprocity 
-without relying on ascriptions of higher-order mental states.
-The mindreader's knowledge of the goal of her target's actions depends on the target's knowledge of the goal of the mindreader's actions.
-What makes this dependence possible is not each knowing that the other knows something about the goal of her actions;
- it is  the actual or apparent  possibility of interaction.
-
-The distant promise of all this is that 
-understanding how interaction widens the evidential basis for goal ascription and mindreading more generally 
-may eventually enable us to explain
-the origins, in evolution or development (and ideally both), of abilities to learn novel and opaque uses for tools from others 
-without assuming that 
-rich communicative skills
-or 
-sophisticated forms of mindreading 
-%(such as the ability to ascribe propositional attitudes)
-must already be  present.
-
-\section{Communicative gestures}
-\label{sec:communicative_gestures}
-When introducing the problem of opaque means 
-(in section \vref{sec:opaque_means})
-we saw that it could  affect
-communicative actions.
-To illustrate this suggestion we drew on 
-an experiment by \citet{hare_chimpanzees_2004}
-in which chimpanzees had to find a reward 
-and were helped by being shown either a failed reach or  a helpful point to the target location.
-Strikingly, for chimpanzees the helpful point is no help at all---even though it superficially resembles the failed reach, which  did help.
-Taking this paradigm as a case study, 
-we want to suggest that your-goal-is-my-goal might 
-enable us to understand how abilities to engage in joint action 
-could be part of what enables mindreaders to make the transition from 
-a simple understanding of goals 
-to 
-an early understanding communicative actions.%
-\footnote{
-Note that although our discussion borrows an experimental paradigm,
-our aim is not to argue for empirical hypotheses about chimpanzee social cognition.
-Our aim is only to argue for the theoretical significance of interaction for mindreading by showing 
-that your-goal-is-my-goal
-could {in principle} enable
-individuals to make the transition 
-from
- a simple understanding of goals
-to
- an early understanding of communicative actions.
-Of course it would powerfully demonstrate the relevance of our theoretical argument if we could provide evidence to show that this actually happens.
-But for now we are concerned with more narrowly conceptual issues.
-}
-
-Let us imagine ourselves as the chimpanzee  for a moment.
-We witness the pointing action.
-With our eyes we follow the point to a container \citep[see][p.\ 6]{Moll:2007gu}.
-So we do associate the pointing action with its target.
-But we are no more likely to choose this container than the other in seeking the reward.
-So we probably do not think of the pointing action as having any goal which would clue us in to the relevance of the container it indicates.
-(In principle we might perfectly understand the pointing action while failing to react to it in any systematic way because we are uncertain about the agent's integrity; but let us discount this possibility for the sake of illustration.)
-Now suppose 
-that, 
-before pointing, the agent had used facial gestures to signal willingness to engage in joint action with us
-%; and that we were able to think of retrieving the food as a possible distributive goal%
-%\footnote{
-%    On \emph{distributive goal} see section \vref{sec:joint_action}.
-%  }
-%of our actions;
-and 
-that we had exploited the my-goal-is-your-goal inference.
-Then we would believe,
-perhaps mistakenly,
-that a goal of the pointing action was to retrieve the food.
-In which case the pointing action would have been no less helpful in enabling us to succeed than the failed reach---which, as you may recall, was very helpful.
-So the your-goal-is-my-goal inference can 
-enable a goal ascriber
-to  
-misunderstand pointing actions as something like  failed reaches.
-This means that,
-even without any deeper understanding of communication,
-goal ascribers can respond appropriately to helpful pointing actions in the context of joint action.
-
-It is natural to suppose that the difficulty chimpanzees have in Hare and Tomasello's experiments with responding appropriately to helpful pointing but not to failed reaching
-is due to a failure to understand communicative intention.%
-\footnote{
-See  footnote \vref{fn:communicative_intention}.
-Relatedly,
-in their discussion of these findings
-Moll and Tomasello suggest that
-`to understand pointing, the subject needs to understand more than the individual goal-directed behaviour. 
-She needs to understand that by pointing towards a location, the other attempts to communicate to her where a desired object is located; that the other tries to inform her about something that is relevant for her'
-\citep[p.\ 6]{Moll:2007gu}.
-Assuming this is right, our suggestion is that individuals could reliably  respond  appropriately to pointing actions in the context of joint action without understanding pointing.
-}
-What we are suggesting is that participants must also have been unable or unwilling to 
-exploit the your-goal-is-my-goal inference.
-
-Consider a related experiment by \citet{leekam_adults_2010}.
-Again participants had to retrieve a reward from one of several closed containers, but this time they were two- and three-year-old children.
-In one condition participants were shown an adult holding up a replica of the target container.
-Leekam and colleagues found that 
-when this action was accompanied by an engaging facial expression,
-three-year-old children were significantly better at identifying the correct container compared to
-when the the action was accompanied by a neutral facial expression (p.\ 116).
-Why did the engaging facial expression enhance performance?
-The authors consider the idea that 
-engaging facial gestures somehow help children to understand communicative intentions.%
-\footnote{
-\citet[p.\ 118]{leekam_adults_2010}: `the adult’s social cues conveyed her communicative intent, which in turn encouraged the child to ‘see through the sign’ ...
-helping them  to take a dual stance to it.'
-}
-An alternative possibility is that children succeeded without understand the replica as a sign at all.
-Instead they may have 
-associated the replica with the container it resembled
-(which by itself is not enough to motivate selecting this container, of course),
-regarded the engaging facial gestures as expressing willingness to engage in joint action,
-and 
-exploited your-goal-is-my-goal 
-	to infer that a goal of the action of holding up the replica was to find the reward.
-In this way they might have understood (or misunderstood) the action of holding up the replica as 
-like a failed reach
-in being an attempt to retrieve the reward.
-
-So far we have illustrated how your-goal-is-my-goal enables responding appropriately to communicative gestures with two examples, pointing and holding up a replica.
-The pattern of reasoning generalises to a wider range of communicative gestures including single-word utterances.
-The basic requirement is this: in a particular context, the goal ascriber must associate a communicative gesture with its referent.
-For instance, she must associate the pointing gesture with the object indicated; or, if (say) she is looking to see who has an object she must associate an utterance of `daddy' with the daddy.%
-\footnote{
-Such associations appear to be in place by six months of age or earlier \citep{tincoff:1999_beginnings_,tincoff:2011}.
-There is debate on when infants are first able to rapidly associate words with novel objects \citep{werker:1998_acquisition,friedrich:2011_word_}.
-Interestingly,
-a possible developmental decalage separating 
-two measures of whether infants have formed a word-object association,
-preferential looking and comprehension as measured by reaching in response to a request  \citep{gurteen:2011_rapid},
-may suggest that
-infants can form a word-object association before they can comprehend communicative actions involving the word.
-}
-As we saw, 
-outside the context of joint action,
-merely associating a gesture with its referent falls short of being able to respond appropriately.
-But if a mindreader supposes that her target is willing to engage in joint action with her,
-then she may infer that the goal of her target's action is her goal
-and so be motivated to treat the thing associated with a communicative gesture as relevant to the goal of her own actions.
-This will reliably (but not always) enable her to respond appropriately to the communicative gesture even without understanding it as a communicative gesture.
-And once she has experienced how that communicative gesture works as a tool for guiding others' actions in the context of joint action,
-she may be in a position to realise, further, that the same tool can be used in other contexts.
-
-This, in barest outline, is 
-how
-possessing abilities to engage in joint action
-means that 
-an individual with an ability to ascribe simple goals only and no understanding of communicative intent
-might 
-nevertheless reliably respond appropriately to some communicative gestures,
-and so come be in a position to understand how such gestures can be used to guide others' actions.%
-\footnote{
-Contrast Csibra's claim that,
-early in human development,
-goal ascription (`teleological understanding' in his terms)
-and 
-identifying the referents of communicative gestures (`referential understanding')
-`rely on different kinds of action understanding' %\citep[p.\ 456]{Csibra:2003kp}
-and
-are initially two distinct `action interpretation systems'   \citep[p.\ 456]{Csibra:2003kp}.
-%(although of course they come together later in development)
-We have not shown that this claim wrong.
-But we have shown that there is another possibility:
-the referential stance may 
-emerge from
-the teleological stance
- together with abilities to engage in simple forms of joint action.
-}
-Of course we have only argued that this transition is theoretically possible;
-we have not attempted to defend any hypothesis about 
-anything's actual evolution or development.
-%Our point is just to illustrate one of several ways in which interaction may matter for mindreading.
-But widening the evidential basis for goal ascription
-to include evidence available only to mindreaders actually or apparently capable of interacting with their targets
-does reveal the coherence of  
-novel hypotheses about a role for interaction in the emergence  of communicative gestures.
-
-%This is one illustration of how capacities for joint action, even very simple forms of joint action, might be relevant to explaining the development or evolution of richer forms of cultural cognition.
-
-\section{The problem of false belief}
-\label{sec:false_belief}
-So far we have been arguing for the value of interaction for goal ascription
-on the grounds that 
-interacting mindreaders, unlike mere observers, have 
-a route to knowledge of the goals of others' actions
-which avoids the problem of opaque means.
-There is another reason for supposing that interaction matters for goal ascription and (at least indirectly) for mindreading more generally.
-To introduce this reason we must first describe another problem affecting goal ascription, the problem of false belief.
-
-Recall that the problem of opaque means occurs when mindreaders  must rely entirely on observation and cannot identify to which ends actions are means 
-(see section \vref{sec:opaque_means}).
-A yet more familiar problem affecting goal ascription 
-arises from the interdependence of beliefs and goals.
-To illustrate, 
-imagine sitting at a table.
-On the table are two closed opaque boxes.
-One box contains an owl, the other a cat.
-If the goal of your action is to retrieve the cat, 
-and you believe that the cat is in the north box,
-then 
-(unless things are going very badly) 
-you will reach for the north box.
-But of course if you had believed instead that the cat was in the south box,
-then, in acting on the same goal, you would have reached for the south box.
-Now consider Ayesha who is observing your actions.
-Suppose Ayesha has sufficient reason to believe, falsely, that you know the cat is in the south box.
-Then she may be justified in supposing, incorrectly, that 
-the goal of your action, in reaching for the north box,
-is to retrieve the owl.
-As this illustrates,
-differences in belief between observers and protagonists can 
-impair goal ascription
-when the goal ascriber is unaware of those differences.
-Call this the problem of false belief.
-
-The your-goal-is-my-goal
-route to knowledge
-sometimes enables mindreaders to avoid the problem of false belief.
-To illustrate consider a counterfactual alternative to the above example.
-Ayesha dislikes the owl and is concerned with retrieving the cat.
-As you reach for the north box,
-your facial gestures signal willingness to engage in joint action with Ayesha.
-She then concludes, correctly, that your goal is her goal, to retrieve the cat.
-So despite the difference in belief,
-the possibility of interaction means that
-Ayesha can knowledgably identify the goal of your action.
-
-Exploiting your-goal-is-my-goal does not make it possible to  avoid the problem of false belief entirely,
-it only shifts  the problem.
-To illustrate,
-in the above example Ayesha's ability to correctly identify the goal of your action
-depends on your correctly anticipating the goals of her actions---on your knowing that she is concerned with retrieving the cat.
-If you incorrectly anticipated that her actions would be directed to retrieving the owl,
-Ayesha would have been mistaken in taking your actions to be directed to retrieving the cat.
-As this indicates,
-exploiting your-goal-is-my-goal 
-can shift 
-both 
-\emph{which} differences in belief have the potential to impair  goal ascription
-and also 
-\emph{who} needs to be aware of those differences.
-This is a second illustration of how the possibility of interaction 
-may enable 
-	even individuals with only limited insights into others' minds and actions
- 	to exploit reciprocity in mindreading---to exploit the fact that their targets are sometimes themselves mindreading the mindreaders.
-
-In short, then, abilities to engage in joint action open a route to knowledge of the goals of other agents' actions which does not depend on the goal ascriber knowing what her target believes even when their beliefs relevantly differ.
-This is not because abilities to engage in joint action provide a way to avoid the problem of false beliefs altogether.
-Rather they shift the burden of resolving the problem of false belief from a goal ascriber to her target.
-This is potentially valuable for those who have limited insight into their targets' beliefs,
-or who lack abilities to track differences in belief altogether.
-
-%On a purely observational model of mindreading,
-%it seems inevitable,
-%given that observed behaviours will be consistent with many different ascriptions of beliefs and goals,
-%that a goal ascriber who was unable to track others' beliefs 
-%would often have to proceed as if there were no relevant differences in beliefs \citep[compare][p.\ 49]{Bennett:1976rg}. 
-%%\footnote{
-%%Compare \citet[p.\ 49]{Bennett:1976rg}.  For a related argument concerning beliefs and desires rather than beliefs and goals, see 
-%%	\citet[p.\ 17]{Davidson:1984pr}
-%%%what can fairly be treated as evidence for any one of the propositional attitudes can count as evidence only if untested assumptions are made about others.”  [1984d, p. 17]
-%%and
-%%	(\citeyear[pp.\ 159--167]{Davidson:1984wh}).
-%%%`This creates a problem, for it means that behaviour, which is the main evidential basis for attributions of belief and desire, is reckoned the result of two forces less open to public observation.
-Thus where one constellation of beliefs and desires will rationalize an action, it is always possible to find a quite different constellation that will do as well.
-Even a generous sample of actions threatens to leave open an unacceptably large number of alternative explanations.'
-%%%`of course it cannot be assumed that speakers never have false beliefs.
-Error is what gives belief its point.
-We can, however, take it as given that most beliefs are correct.
-The reason for this is that a belief is identified by its location in a pattern of beliefs; it is this pattern that determines the subject matter of the belief, what the belief is about.
-Before some object in, or aspect of, the world can become part of the subject matter of a belief (true of false) there must be endless true beliefs about the subject matter.
-False beliefs tend to undermine the identification of the subject matter; to undermine, therefore, the validity of a description of the belief as being about that subject.'
-%%}
-% So on a purely observational model of mindreading
-%there may fail to be any straightforward way in which
-%the need to track differences in belief 
-%could become apparent to a mindreader who did not already understand that beliefs can be false.
-%Things are different
-%if we extend our model of mindreading
-%to take into account the possibility of interaction.
-%We can then see how
-%a goal ascriber who is unable to track others' beliefs
-%will encounter anomalies.
-%For some differences in belief 
-%(the mere fact of such differences, not anyone's awareness of them)
-%will mean that
-% ordinary, purely observational routes to knowledge will result in ascribing one goal
-% whereas
-% exploiting your-goal-is-my-goal would result in a ascribing a different goal.
-%That is, the evidence available to a goal ascriber may inexplicably support two incompatible ascriptions. 
-%In this way,
-%differences in belief have the potential to cause puzzling anomalies
-%even before a mindreader is able to track such differences.
-%This 
-%raises the possibility
-%that abilities to engage in joint action
-%may play a role in explaining how  awareness of belief emerges.%
-%\footnote{
-%Some researchers have suggested that awareness of belief is innate in humans
-%\citep{Leslie:1998nq,Leslie:2005ef,Baillargeon:gx},
-%although this view is currently controversial
-%\citep{Apperly:2009ju,Ruffman:2005yl}.
-%But whether or not the emergence in development of awareness of belief stands in need of explanation,
-%few would doubt that its emergence in evolution does need explaining.
-%}
-
-\section{Conclusion}
-Our aim was to show that interaction can facilitate mindreading in this sense:
-some routes to knowledge are closed to mindreaders who rely exclusively on observation
-but
-open to interacting mindreaders.
-In pursuing this aim we focused on interactions which are joint actions.
-Our suggestion was this.
-Where a mindreader recognizes that her target is willing to engage in joint action
-and where the mindreader is unwilling to change which goals her own actions will be directed to,
-she may be in a position to know
-that the goals of her target's actions will be the goals of her own actions.
-This in outline is the `your-goal-is-my-goal' route to  knowledge (see  section \vref{sec:your_goal_is_my_goal}).
-
-%***Return to the leading theories?
-%We do not take this to show that the leading theories must be entirely discarded.
-%If we are right about interaction making evidence available, 
-%either leading theory can be extended accordingly.
-%***
-
-To show that this route to knowledge is potentially valuable,
-we argued that it enables mindreaders to overcome two problems,
-the problem of opaque means
-and the problem of false belief.
-These problems may rarely arise for human adults thanks to sophisticated linguistic communication and extensive knowledge of how things can be achieved.  
-But interest in the evolution or development of mindreading
-motivates focusing on 
-those with
-  absent or fledgling communicative skills,
-  limited insight into others' minds
-  and narrow knowledge of how things work.
-For such individuals
-the problems of opaque means and false beliefs may easily arise,
-particularly in cases involving novel tools 
-and 
-communicative gestures.
-Both problems can be overcome 
-by exploiting the your-goal-is-my-goal route to knowledge.
-It follows that 
-expertise with 
-  tools 
-  and
-  communicative gestures
-does not presuppose knowledge of others' intentions and beliefs.
-This shows that it is coherent to suppose that 
- expertise with tools or communicative gestures (or both),
-far from  presupposing sophisticated forms of mindreading,
-might instead
-play a role in explaining their emergence in evolution or development.
-
-How might this work?
-Here is a wild conjecture about
-a path from simple forms of interaction to sophisticated forms of mindreading.
-We have seen that 
-	abilities to ascribe goals to actions 
- together with
-	abilities to engage in simple forms of joint action involving distributive goals
- may be sufficient to explain how individuals come to productively misunderstand communicative gestures in the context of joint action (see section \vref{sec:communicative_gestures}).
-This might  be a step towards responding appropriately to, and using, communicative gestures in other contexts,
-which might then play a role in explaining how early forms of communication emerge.
-But we know that early, non-linguistic forms of communication are essential for the later appearance of linguistic communication,
-and abilities to communicate by language in turn appear to play an important role in the development (at least) of sophisticated forms of mindreading.%
-\footnote{
-A range of evidence is reviewed in contributions to \citet{Astington2005ot}.
-}
-Now clearly this wild conjecture is not one that we have provided any reason for accepting.
-Our point in mentioning it is merely to illustrate how
- better understanding the role of interaction in goal ascription and  mindreading more generally
-might eventually lead to discoveries
-on how
- sophisticated forms of cognition
- emerge from
- conceptually and cognitively undemanding forms of interaction.
-Take cultures of tool use,
-	the flexible use of communicative gestures,
-	or insight into interpersonal differences in belief and other mental states:
-if any or all of these 
-originate in abilities to interact,
-then we need to know how interaction could lead to their emergence.
-And one small step towards acquiring this knowledge might just be 
-understanding how interacting mindreaders 
-could 
-	come to know things 
-%	and
-%	encounter anomalies
-which they might not be able to know 
-%or encounter
- as mere observers.
-% matters because it enables us to better understand the evidential basis of mindreading---%
-%and it might just also matter because it plays a role in demonstrating the theoretical coherence of hypotheses about mindreading's emergence.
 

### 2012-06-20 intention motor representation with corrado sinigaglia/intention and motor action.ppr.corrected.tex
proof reading nearly done

--- 2012-04-04 intention motor representation with corrado sinigaglia/intention and motor action.ppr.r1.final.tex
@@
 For instance, how a grasping action is represented may depend on a relatively distal outcome: on whether (say) it results in the object grasped being eaten or placed.
-Also the joint displacements which realise grasping in one context might in another context realise a different action, such as scratching; or they might fail to realise any action at all (because there is no target object, say).
+In addition, the joint displacements which realise grasping in one context might in another context realise a different action, such as scratching; or they might fail to realise any action at all (because there is no target object, say).
 Third, a motor representation of an action outcome locates the action within a hierarchical structure.
-To illustrate, a left-handed precision grip and a right-handed whole hand grip are different action outcomes and both instances of grasping.\footnote{On the notion of action outcome in motor representation, see \citet{gallese:2000_inner}, \citet{jeannerod_motor_2006} and \citet{rizzolatti_mirrors_2008}.}
+To illustrate, a left-handed precision grip and a right-handed whole hand grip are different action outcomes and both instances of grasping.%
+\footnote{On the notion of action outcome in motor representation, see \citet{gallese:2000_inner}, \citet{jeannerod_motor_2006} and \citet{rizzolatti_mirrors_2008}.}
 
@@
 A related alternative is to contrast the same grasping movements in the presence of objects which could, or manifestly could not, be grasped by means of such movements \citep{koch:2010_resonance}.
-In each of these cases there is evidence that some markers of motor processing are correlated with action outcomes rather than narrowly kinematic or dynamic features of action.\footnote{ Of course some researchers have raised doubts concerning some of the evidence that there are markers of motor processing which carry information about action outcomes (e.g.\ \citealp{cavallo_grasping_2011,borroni_mirroring_2011}).
+In each of these cases there is evidence that some markers of motor processing are correlated with action outcomes rather than narrowly kinematic or dynamic features of action.%
+\footnote{ Of course some researchers have raised doubts concerning some of the evidence for the claim that there are markers of motor processing which carry information about action outcomes (e.g.\ \citealp{cavallo_grasping_2011,borroni_mirroring_2011}).
 On balance, however, the evidence supports this conclusion.}
@@
 Moreover, there is a clear resemblance between the natural way of understanding intentions as grounding outcome-directedness and the way in which motor representations ground outcome-directedness (as \citealp{pacherie:2008_action}, pp.\ 189-90 has also argued).
-So given the functional roles and representational objects of motor representations, there is little room for doubt that motor representations can ground the outcome-directedness of purposive actions.
+So given the functional roles and contents of motor representations, there is little room for doubt that motor representations can ground the outcome-directedness of purposive actions.
 
@@
 
-As background we first need a generic distinction between content and format.
-Imagine you are in an unfamiliar city and are trying to get to the central station.
+As background we first need a generic distinction between content and format. Imagine you are in an unfamiliar city and are trying to get to the central station. 
 A stranger offers you two routes. 
@@
 \footnote{Note that the distinction between content and format is orthogonal to issues about representational medium. 
-The maps in our illustration may be paper map or electronic maps, and the instructions may be spoken, signed or written.
-This difference is one of medium.
+The maps in our illustration may be paper map or electronic maps, and the instructions may be spoken, signed or written. This difference is one of medium.
 } 
@@
 But now consider purely mental pantomime---that is, phenomenologically action-like imagination.
-One might use this kind of imagination to explore different ways of completing a task and then, having hit on a good solution, think to oneself `Do that!' It seems possible that in some such cases the demonstrative refers by deferring to a motor representation of action involved in imagining acting.
+One might use this kind of imagination to explore different ways of completing a task and then, having hit on a good solution, think to oneself `Do that!'. 
+It seems possible that in some such cases the demonstrative refers by deferring to a motor representation of action involved in imagining acting.
 
@@
 Could it be a difference in intention that explains how they differ?  
-Since it is possible to grasp without intending to grasp, it is surely also possible to imagine grasping without imagining intending to grasping.
-And note that to imagine grasping without imagining intending to grasping does not necessarily require one to imagine non-intentionally grasping.
+Since it is possible to grasp without intending to grasp, it is surely also possible to imagine grasping without imagining intending to grasp. 
+And note that to imagine grasping without imagining intending to grasp does not necessarily require one to imagine non-intentionally grasping. 
 (It is easy to miss this point by confusing it with the issue of whether the imagining must itself be intentional.
@@
 There is also evidence that experiential judgements of size can be influenced by ability \citep[e.g.][]{witt2009kicking}.  
-It is plausible, then, that motor representation of objects modulates experiences of them.
+It is plausible, then, that some motor representations of objects  modulate experiences of them.  
 This indicates that it is possible to refer not only to ways of acting by also to what to act on by deferring to motor representations.

### 2012-06-20 intention motor representation with corrado sinigaglia/intention and motor action.ppr.corrected.tex
proofing done

@@
 A further conjecture is that whenever there is a non-accidental match between the contents of intentions and of motor representations, this is due at least in part to deferential action concepts. 
-The plausibility of this conjecture rests in part on the absence of any other viable full solution to the interface problem, and of course we do not claim to have shown that no alternatives exists.
-If the conjecture is right, if there is no way of fully solving the interface problem other than by means of deferential action concepts, then concepts which refer by deferring to motor representations are what tie explicit practical reasoning to motor processes.
+The plausibility of this conjecture rests in part on the absence of any other viable full solution to the interface problem, and of course we do not claim to have shown that no alternatives exist. 
+If the conjecture is right, then concepts which refer by deferring to motor representations are what tie explicit practical reasoning to motor processes.
 
@@
 \footnote{
-Heartfelt thanks to Chiara Brozzo, Silvano Zipoli Caiani, Naomi Eilan, Vittorio Gallese, Francesco Guala, \'Elisabeth Pacherie and Natalie Sebanz.
+Heartfelt thanks to Gabriella Bottini, Chiara Brozzo, Silvano Zipoli Caiani, Naomi Eilan, Vittorio Gallese, Francesco Guala, \'Elisabeth Pacherie, Natalie Sebanz and an anonymous referee.
 Corado Sinigaglia was supported by EU grant TESIS, a Fellowship from the Institute of Philosophy (University of London) and a Fellowship from the Center for the Epistemology of Cognitive Sciences, Ecole Normale Superiéure de Lyon.

### 2012-07-06 developing mind/chapter_01.tex
wokring on intor

@@
 
-\def \ititle {A}
+\def \ititle {The Developing Mind}
 \def \isubtitle {}
@@
 
-\begin{abstract}
-\noindent
-***
-\end{abstract}
+%\begin{abstract}
+%\noindent
+%***
+%\end{abstract}
+
+\section{From Myths to Mechanisms}
+
+How do humans come to know  about---and to knowingly manipulate---%
+objects,
+causes,
+words,
+numbers,
+colours,
+actions
+and
+thoughts?
+In a beautiful myth, Plato suggests that the answer is recollection. 
+Before birth, in another world, we become acquainted with the truth. 
+Then, in falling to earth, we forget everything. 
+But as we grow we are sometimes able to recall part of what we once knew.
+So it is by recollection that humans come to know about objects, causes, numbers and everything else.
+
+How else could this happen?
+Since Plato philosophers and psychologists have offered other stories.  
+Some hold, like Plato, that knowledge is in some sense present at birth.
+Or else that the concepts which make knowledge possible are already present at birth. 
+Others suggest that concepts and knowledge are acquired through sensory experience, through learning to act, through training in language or through social interaction.
+These are seductive ideas but none has survived rigorous scientific testing; some have failed, others have yet to be  thoroughly tested or are perhaps not  currently precise enough test. 
+And when we look at particular cases in detail---for instance, when we look at how humans come to know about colours---we will discover complexities that seem to be incompatible with any of the stories.
+That's why this book doesn't tour nativism, empiricism and other myths about the developmental origins of human knowledge, valuable though these are. 
+Development is like climate change in one respect.
+Lots of different mechanisms are simultaneously at work and many interact with each other. 
+%While it might initially seem more satisfying to have a unifying myth, 
+To make progress we need to identify various mechanisms and attempt to model their interactions.
+
+In broad outline, then, understanding the developmental origins of knowledge calls detailed investigation of many interacting but separate phenomena rather than large-scale theory building. 
+This attitude is nicely captured by some psychologists reflecting on difficulties in understanding knowledge of objects:
+%
+\begin{quote}
+`there are many separable systems of mental representations ...\ and thus many different kinds of knowledge.
+...\ the task ...\ is to contribute to the enterprise of finding the distinct systems of mental representation and to understand their development and integration%
+%.
+...\ the challenge ...\ is to characterize what it is about the mental representations of the presence or absence of the barriers that constrains representations of where the object should be in looking-time studies and how these representations differ from those that guide search or pointing in the present studies%
+'
+\citep[p.\ 1522]{Hood:2000bf}.
+\end{quote}
+%
+%obstacle --- half-formed minds
+Performing this task  requires philosophical as well as more narrowly psychological investigation. 
+The question is how humans come to know about objects, words, thoughts and other things. 
+In pursuing this question we have to consider minds where the knowledge is neither clearly present nor obviously absent. 
+This is challenging because both commonsense and theoretical tools for describing minds are generally designed for characterising fully developed adults. 
+Davidson writes:
+%
+\begin{quote}
+`%
+%if you want to describe what is going on in the head of the child when it has a few words which it utters in appropriate situations, you will fail for lack of the right sort of words of your own. 
+We have many vocabularies for describing nature when we regard it as mindless, and we have a mentalistic vocabulary for describing thought and intentional action; what we lack is a way of describing what is in between' \citep[p.\ 11]{Davidson:1999ju}.
+\end{quote}
+%
+%Of course you might not agree with Davidson that children just starting to use words are incapable of thought or of intentional action. 
+%In fact research on children's pre-verbal communicative behaviours such as pointing seems not only to manifest purposiveness but also a sensitivity to others' thoughts.
+%(Davidson seems to have paid little attention to development, and he held that being able to think, to act intentionally and to communicate with language all require understanding the possibility that one's own beliefs may be false.)
+%Even so, Davidson 
+%
+This is why philosophy is needed. 
+To understand the emergence of knowledge we need to find ways of describing what is in between: individuals whose movements are neither mindless nor guided by intention and knowledge.  
+%in between ways of describing individuals, ways that neither treat them as mindless nor presuppose capacities for knowledge.
+
+This difficulty emerges in a concrete way in developmental research. 
+When can infants first know things about objects they aren't perceiving?   
+For instance, when a ball falling behind a chair disappears from view, when do infants first realise that the ball is somewhere behind the chair?
+The ability to realise this is known as `object permanence'.  
+One way to test for object permanence is to ask when infants first reach for objects they can’t see or when they first remove barriers to retrieve objects concealed behind them.  
+Infants don’t do this until around eight months \citep[p.\ 202]{Meltzoff:1998wp} or maybe later \citep{moore:2008_factors}.  
+Since four-month-olds already have the planning skills they would need to execute the reach \citep{Shinskey:2001fk}, 
+their failure to reach is evidence that infants can first think about objects they aren’t perceiving at around eight months or later.  
+But another way to test for object permanence is to ask how infants respond to apparently impossible events. 
+Suppose, for example, that infants watch as a solid object is placed immediately behind a screen and then the screen falls backwards, ending up flat as though the object were not there, which is apparently impossible \citep{baillargeon:1985_object,baillargeon:1987_object}.
+If infants show heightened interest in this and similar cases, perhaps by looking at the display for longer than might otherwise be expected, this would be evidence that they can know things about objects they aren't perceiving.\footnote{
+These particular studies have been attacked on methodological grounds \citep[e.g.][]{sirois:2007_social_}, but there are other, related studies \citep[some are mentioned by][]{Aguiar:2002ob}.
+}
+As it turns out, infants show such heightened interest from around four months or earlier. 
+Put together, the two sorts of findings give rise to the `paradox of early permanence' (as \citealp{Meltzoff:1998wp} call it). 
+The best explanation of the first sort of findings seems to be that infants cannot think about objects they aren't perceiving until eight months or later; 
+but the best explanation of the second sort of findings seems to be that infants can do this from around four months or earlier. 
+Clearly these explanations cannot both be correct. 
+But neither seems to be wrong. 
+Resolving this apparent conflict requires responding to an instance of Davidson's challenge and find ways of describing phenomena in between mindless ignorance of unseen objects and adult-like knowledge of objects.
+%key review paper: \citep{moore:2010_numerical}
+
+In what follows we will repeatedly encounter instances of this challenge  in considering the developmental origins of different domains of knowledge, 
+and we will examine tools philosophers and psychologists have invented in responding to different instances of the challenge.  
+Puzzling, often apparently contradictory patterns of findings concerning how humans come to know about things such as objects, colours, causes, numbers and the rest matter for two reasons.
+One is that they can advance our understanding of  the origins of knowledge.
+The other is that they show how little is actually understood and how much is yet to be discovered.
+In what follows there is no much by way of answers; the primary aim is just to identify the puzzles.
+
+%Focus is on development and origins of knowledge ... but a side-effect will be that we understand representation, perception and knowledge better.
+
+%source!
+%research/Development as Rediscovery--core knowledge and speech perception.doc
 
@@
 
+\section{Quotes}
+
+\citep[p.\ 198]{Fodor:1975pb}: “the fundamental explicandum, is the organism and its propositional attitudes: what it believes, what it learns, what it wants and fears, what it perceives to be the case.
+Cognitive psychologists accept ...
+the facticity of ascriptions of propositional attitudes to organisms and the consequent necessity of explaining how organisms come to have the attitudes to propositions that they do.”
+
+\citep[p.\ 104]{Locke:1975qo}: “...
+’tis past doubt, that Men have in their Minds several Ideas, such as are those expressed by the words, Whiteness, Hardness, Sweetness, Thinking, Motion, Man, Elephant, Army, Drunkenness, ...
+and others: It is in the first place to be enquired, How he comes by them?”
+
+\citep[p.\ 12]{Dewey:1938yp}: “How does it come about that the development of organic behavior into controlled inquiry brings about the differentiation and cooperation of observational and conceptual operations?”
+
 \section{Why}

### 2012-07-12 developing mind/chapter_01.tex
working
working on proposal

@@
 
-\def \ititle {The Developing Mind}
-\def \isubtitle {}
+\def \ititle {The Developing Mind: }
+\def \isubtitle {A philosophical introduction to issues in cognitive development}
 \def \iauthor {Stephen A. Butterfill}
@@
 and
-thoughts?
+minds?
 In a beautiful myth, Plato suggests that the answer is recollection. 
@@
 
-In broad outline, then, understanding the developmental origins of knowledge calls detailed investigation of many interacting but separate phenomena rather than large-scale theory building. 
-This attitude is nicely captured by some psychologists reflecting on difficulties in understanding knowledge of objects:
-%
-\begin{quote}
-`there are many separable systems of mental representations ...\ and thus many different kinds of knowledge.
-...\ the task ...\ is to contribute to the enterprise of finding the distinct systems of mental representation and to understand their development and integration%
-%.
-...\ the challenge ...\ is to characterize what it is about the mental representations of the presence or absence of the barriers that constrains representations of where the object should be in looking-time studies and how these representations differ from those that guide search or pointing in the present studies%
-'
-\citep[p.\ 1522]{Hood:2000bf}.
-\end{quote}
-%
 %obstacle --- half-formed minds
@@
 
+\section{Core Knowledge and Perception}
 %source!
 %research/Development as Rediscovery--core knowledge and speech perception.doc
+
+In broad outline, then, understanding the developmental origins of knowledge calls detailed investigation of many interacting but separate phenomena rather than large-scale theory building. 
+This attitude is nicely captured by some psychologists reflecting on difficulties in understanding knowledge of objects:
+%
+\begin{quote}
+`there are many separable systems of mental representations ...\ and thus many different kinds of knowledge.
+...\ the task ...\ is to contribute to the enterprise of finding the distinct systems of mental representation and to understand their development and integration%
+%.
+...\ the challenge ...\ is to characterize what it is about the mental representations of the presence or absence of the barriers that constrains representations of where the object should be in looking-time studies and how these representations differ from those that guide search or pointing in the present studies%
+'
+\citep[p.\ 1522]{Hood:2000bf}.
+\end{quote}
+%
+
+\section{Development as rediscovery}
+
+\section{Social Origins of Knowledge}
 

### 2012-07-12 developing mind/proposal developing mind.tex
working on proposal

%!TEX TS-program = xelatex
%!TEX encoding = UTF-8 Unicode

\def \papersize {a4paper}
\documentclass[12pt,\papersize]{extarticle}
% extarticle is like article but can handle 8pt, 9pt, 10pt, 11pt, 12pt, 14pt, 17pt, and 20pt text

\def \ititle {The Developing Mind: A philosophical introduction to issues in cognitive development}
\def \isubtitle {-- Book Proposal --}
\def \iauthor {Stephen A. Butterfill}
\def \iemail{s.butterfill@warwick.ac.uk}
%\def \iauthor {}
%\def \iemail{}
\date{}

\input{$HOME/Documents/submissions/preamble_steve_paper2}
%\author{}

%\setromanfont[Mapping=tex-text]{Sabon LT Std}

\begin{document}

\setlength\footnotesep{1em}

\bibliographystyle{newapa} %apalike

\tolerance=5000

\maketitle
%\tableofcontents

%\begin{abstract}
%\noindent
%***
%\end{abstract}

\section{Brief Description}
How do humans come to know about %
objects,
causes,
words,
numbers,
colours,
actions
and
minds? 
The question goes back to Plato or earlier and remains unanswered.
%Philosophers have been pursuing this question since Plato or before. 
%More recently it has become the focus of a branch of psychology, developmental psychology.
%So far no one has convincingly answered the question, but two major breakthroughs have been made in the last three decades or so.
Two recent scientific breakthroughs appear to bring us closer to an answer, and to show that the question is even less straightforward than philosophers have  assumed.
The first breakthrough is the discovery that preverbal infants enjoy surprisingly rich social skills, skills which
% begin to be built from the first months of life and 
may well be foundational for later linguistic abilities and enable the emergence of knowledge \citep[e.g.][]{Csibra:2009xr,Meltzoff:2007pj,Tomasello:2005wx}. 
A second breakthrough concerns the use of increasingly sensitive---and sometimes controversial---methods to detect  sophisticated expectations concerning  causal interactions, numerosity, mental states and more besides in preverbal infants \citep[e.g.][]{Spelke:1990jn,Baillargeon:gx}.
These expectations or the representations and processes underpinning them arguably also enable the emergence of knowledge. 
Although each breakthrough has been discussed at length by leading psychologists, philosophers have yet to consider either systematically in print.
Further, the two breakthroughs are associated with different camps, one nativist and the other Vygotskian, and have rarely been considered together as twin factors enabling the emergence of knowledge.
The proposed book will familiarise readers with findings related to both breakthroughs and explain their relevance to philosophical accounts of knowledge, action and experience.  
Most importantly, the book aims to introduce readers to  philosophical issues raised by these findings in cognitive development.

Advances in neuroscience may be transforming parts of developmental psychology much as they have already transformed the study of perception and action \citep{Johnson:2005az}. 
The proposed book focusses on discoveries rather than methods but will introduce readers to some relevant findings in developmental cognitive neuroscience along the way.

The book is organised by domains of knowledge, so that one chapter concerns knowledge of objects, another focusses on  knowledge of number, and so on.  
The domains are chosen so that each set of developmental findings is linked to one or more philosophical issues. 
For instance, research on knowledge of objects gives bite to questions about modularity and the nature of tacit knowledge \citep{Davies:1989gg,Fodor:1983dg}; research on knowledge of number invites discussions of nativism \citep{Fodor:1981ep,Spelke:1998im}; and developmental findings on knowledge of colour may challenge some assumptions philosophers have made about relations between language, thought and perception \citep{Gilbert:2006yb,Regier:2009ve}.

\section{Audience}

The book is aimed at advanced undergraduate and graduate students in philosophy. 
It is suited to teaching a course on philosophical issues in cognitive development. 
The book could also be used in courses on the philosophy of mind and action, or the philosophy of psychology. 
To this end the book will not presuppose familiarity with the philosophy of mind and will include chapter summaries and suggestions for further reading.

Professional philosophers whose research connects with issues in cognitive development may also find the book useful.  
Judging by papers and manuscripts under review for journals and university presses, even those writing about cognitive development can have difficulty identifying the full range of findings relevant to their positions.

Students in psychology are often drawn to philosophy and should find in this book an accessible introduction to key distinctions and arguments. 
It might even be used for courses in developmental psychology. (I once co-taught such a course in psychology with Jim Russell in Cambridge).

\section{Length}
70,000--100,000 words

\section{Timetable}
I aim to submit a first draft of the manuscript by 30 September 2014.

I plan to try out material for the book while teaching courses at the University of Warwick (UK) and Central European University (Hungary).

\section{Competition}
At the time of writing there is no book or collection devoted to philosophical issues in cognitive development.

Philosophers have written monographs and edited collections on particular issues in cognitive development \citep[e.g.][]{Bermudez:2003dj,carruthers:1996_theories}.  
The proposed book complements these by providing a unified discussion of a wider range of topics. 
The advantage is not just convenience: understanding and properly evaluating theories about the developmental origins of knowledge requires bringing together research on different domains of knowledge.

There are also several collections bringing together research by philosophers and developmental psychologists or presenting developmental research in ways accessible to philosophers 
\citep[e.g.][]{hirschfeld:1994_mapping,carruthers:2005_innate_structure,carruthers:2006_innate_culture,mccormack:2011_tool}.
From the point of view of the proposed book, these provide useful sources of further reading for those who want more information on particular issues.

Finally, the proposed book will complement the many excellent books which expound particular theories about the origins of knowledge; these include 
	\citet{carey:2009_origin},
	\citet{Tomasello:1999xz},
	\citet{Gopnik:1997xq},
	and
	\citet{Elman:1996zd}.
The proposed book will include critical discussion of theories presented in these volumes.  
It's aims are clearly distinct from any of these volume's aims: the proposed book will review a range of existing theories and explore philosophical questions they raise.

\section{Contents}
(Details are likely to change.)

\subsection{Introduction}
How do humans come to know  about---and to knowingly manipulate---%
objects,
causes,
words,
numbers,
colours,
actions
and
minds?
In pursuing this question we have to consider minds where the knowledge is neither clearly present nor obviously absent. 
This is challenging because both commonsense and theoretical tools for describing minds are generally designed for characterising fully developed adults. 
Davidson writes:
%
\begin{quote}
`%
%if you want to describe what is going on in the head of the child when it has a few words which it utters in appropriate situations, you will fail for lack of the right sort of words of your own. 
We have many vocabularies for describing nature when we regard it as mindless, and we have a mentalistic vocabulary for describing thought and intentional action; what we lack is a way of describing what is in between' \citep[p.\ 11]{Davidson:1999ju}.
\end{quote}
%
%Of course you might not agree with Davidson that children just starting to use words are incapable of thought or of intentional action. 
%In fact research on children's pre-verbal communicative behaviours such as pointing seems not only to manifest purposiveness but also a sensitivity to others' thoughts.
%(Davidson seems to have paid little attention to development, and he held that being able to think, to act intentionally and to communicate with language all require understanding the possibility that one's own beliefs may be false.)
%Even so, Davidson 
%
To understand the emergence of knowledge we need to find ways of describing what is in between: individuals whose movements are neither mindless nor guided by intention and knowledge.

***

\subsection{Social Interaction before Knowledge}
Consider the hypothesis that social interaction is an enabler of cognitive development. 
This hypothesis immediately raises a question. 
***

\subsection{Objects and How They Interact}

\subsection{Number: From Perceptual Experience to Knowledge}

\subsection{Seeing and Talking about Colours}

\subsection{Words Are Tools for Communication}

\subsection{Actions: Teleology and Mirroring Motor Awareness}

\subsection{Mindreading}

\subsection{Conclusion}

\bibliography{$HOME/endnote/phd_biblio}

\end{document}

### 2012-07-13 developing mind/proposal developing mind.tex
working on proposal

@@
 
-\input{$HOME/Documents/submissions/preamble_steve_paper2}
+\input{$HOME/Documents/submissions/preamble_steve_report}
 %\author{}
+
+%use titlesec package to make subsections bold italics
+\titleformat*{\subsection}{\bfseries\itshape}
 
@@
 %\tableofcontents
+
+% disables chapter, section and subsection numbering
+\setcounter{secnumdepth}{-1}
 
@@
 Two recent scientific breakthroughs appear to bring us closer to an answer, and to show that the question is even less straightforward than philosophers have  assumed.
-The first breakthrough is the discovery that preverbal infants enjoy surprisingly rich social skills, skills which
+The first breakthrough is the discovery that preverbal infants enjoy surprisingly rich social abilities, abilities which
 % begin to be built from the first months of life and 
@@
 Although each breakthrough has been discussed at length by leading psychologists, philosophers have yet to consider either systematically in print.
-Further, the two breakthroughs are associated with different camps, one nativist and the other Vygotskian, and have rarely been considered together as twin factors enabling the emergence of knowledge.
+Further, the two breakthroughs are associated with different camps, one nativist and the other Vygotskian, and have rarely been considered together as identifying twin factors enabling the emergence of knowledge.
 The proposed book will familiarise readers with findings related to both breakthroughs and explain their relevance to philosophical accounts of knowledge, action and experience.  
@@
 \section{Contents}
-(Details are likely to change.)
+(This plan may change.)
 
@@
 To understand the emergence of knowledge we need to find ways of describing what is in between: individuals whose movements are neither mindless nor guided by intention and knowledge.  
+Some progress has already been made but many challenges remain. 
+Philosophers have much to contribute in crafting distinctions and conceptual tools useful for meeting the challenges involved in describing what is in between. 
+There is also much for philosophers to gain: studying what is in between mindless nature and the sorts of cognition captured by  adult humans' everyday, pre-theoretical mentalistic notions 
+ will reveal things about what minds are and how they work.
 
-***
+As well as explaining how research on developing minds bears on some existing philosophical issues and raises some new ones,
+this chapter introduces two scientific breakthroughs that have recently furthered understanding of how knowledge might emerge in development.
+As already mentioned (in the Brief Description), the first breakthrough is the discovery that preverbal infants enjoy surprisingly rich social abilities, abilities which
+% begin to be built from the first months of life and 
+may well be foundational for later linguistic abilities and enable the emergence of knowledge \citep[e.g.][]{Csibra:2009xr,Meltzoff:2007pj,Tomasello:2005wx}. 
+A second breakthrough concerns the use of increasingly sensitive---and sometimes controversial---methods to detect  sophisticated expectations concerning  causal interactions, numerosity, mental states and more besides in preverbal infants \citep[e.g.][]{Spelke:1990jn,Baillargeon:gx}.
+These breakthroughs are driven by researchers with conflicting theoretical positions and raise quite different issues.  
+But they both  identify ingredients necessary for understanding the developmental origins of human knowledge.
 
-\subsection{Social Interaction before Knowledge}
-Consider the hypothesis that social interaction is an enabler of cognitive development. 
-This hypothesis immediately raises a question. 
-***
+\subsection{Chapter 1. Social Interaction without Words}
+Could social interaction be an enabler of cognitive development? 
+This chapter examines arguments for the hypothesis that it could
+ (as offered in \citealp{Tomasello:2003tq} and \citealp{Moll:2007gu}); it also discusses objections to this hypothesis and introduces questions it raises.
 
-\subsection{Objects and How They Interact}
+An immediate problem is that much adult social interaction depends so heavily on communication by language (not to mention social networks) that it can be hard to imagine what social interaction with preverbal infants could involve.
+Relatedly, many theoretical accounts of social interaction presuppose linguistic abilities.
+To overcome this difficulty we shall review evidence for a package of social abilities manifested in preverbal infants.
+These include imitation, which can occur just days and even minutes after birth \citep{meltzoff:1977_imitation,field:1982_imitation,meltzoff:1983_newborn}, 
+%for a review on imitation, see \citep{jones:2009_development}
+imitative learning, 
+gaze following \citep{Csibra:2008be},
+goal ascription \citep{Gergely:1995sq,Woodward:2000jw},
+social referencing \citep{Baldwin:2000qq} 
+and 
+pointing. 
+Taken together, the evidence reveals that preverbal infants have surprisingly rich social abilities.
 
-\subsection{Number: From Perceptual Experience to Knowledge}
+The hypothesis that these social abilities enable cognitive development faces an objection. 
+For it seems that many of these abilities already presuppose  knowledge of objects, actions and minds. 
+For instance, twelve-month old infants  will helpfully point to inform ignorant but not knowledgable adults about the location of an object  \citep{Liszkowski:2008al}.
+Some researchers take this to show that these infants are already capable of knowing about others' knowledge and ignorance. 
+If that is right, pointing and other social abilities could play at most a limited role in explaining the developmental origins of knowledge of mind. 
+After all, we can hardly explain the emergence of something by appealing to abilities whose possession already presuppose it.
+We will encounter tools for replying to this objection in Chapters 2 (on Objects) and 6 (on Actions).
 
-\subsection{Seeing and Talking about Colours}
+The hypothesis that social abilities enable cognitive development naturally invites a question: How could that work? 
+Some have argued that early social abilities partially explain the emergence in humans of communication by language (\citealp{tomasello:2008origins}; see Chapter 5 on Words). 
+This is a valuable contribution.  
+But could social abilities also play a role in explaining the developmental origins of knowledge of things other than words---for example, in explaining the emergence of knowledge of objects, numbers, colours or minds?
+This question comes up in one way or another in each of the following chapters.
 
-\subsection{Words Are Tools for Communication}
+\subsection{Chapter 2. Objects and How They Interact}
+\label{ch:objects}
+When can infants first know things about objects they aren't perceiving?   
+For instance, when a ball falling behind a chair disappears from view, when do infants first realise that the ball is somewhere behind the chair?
+The ability to realise this is known as `object permanence'.  
+One way to test for object permanence is to ask when infants first reach for objects they can’t see or when they first remove barriers to retrieve objects concealed behind them.  
+Infants don’t do this until around eight months \citep[p.\ 202]{Meltzoff:1998wp} or maybe later \citep{moore:2008_factors}.  
+Since four-month-olds already have the planning skills they would need to execute the reach \citep{Shinskey:2001fk}, 
+their failure to reach is evidence that infants can first think about objects they aren’t perceiving at around eight months or later.  
+But another way to test for object permanence is to ask how infants respond to apparently impossible events. 
+Suppose, for example, that infants watch as a solid object is placed immediately behind a screen and then the screen falls backwards, ending up flat as though the object were not there, which is apparently impossible \citep{baillargeon:1985_object,baillargeon:1987_object}.
+If infants show heightened interest in this and similar cases, perhaps by looking at the display for longer than might otherwise be expected, this would be evidence that they can know things about objects they aren't perceiving.
+%\footnote{
+%These particular studies have been attacked on methodological grounds \citep[e.g.][]{sirois:2007_social_}, but there are other, related studies \citep[some are mentioned by][]{Aguiar:2002ob}.
+%}
+ As it turns out, infants show such heightened interest from around four months or earlier. 
+Put together, the two sorts of findings give rise to the `paradox of early permanence' (as \citealp{Meltzoff:1998wp} call it). 
+The best explanation of the first sort of findings seems to be that infants cannot think about objects they aren't perceiving until eight months or later; 
+but the best explanation of the second sort of findings seems to be that infants can do this from around four months or earlier. 
+Clearly these explanations cannot both be correct. 
+But neither seems to be wrong.
 
-\subsection{Actions: Teleology and Mirroring Motor Awareness}
+This conflict exemplifies a pattern that occurs again and again in investigating the developmental origins of knowledge. 
+Here's the pattern.
+We ask when humans can first know about X. 
+One set of findings provides converging evidence that the answer is: surprisingly early.
+Another set of findings, using a different set of techniques,  provides converging evidence that the answer is: much later. 
+Unless they arise from methodological failings, these conflicting answers force us to recognise that the question involves a mistake. 
+The mistake is to think that there is a single kind of knowledge. 
+%As some psychologists reflecting on difficulties in understanding knowledge of objects suggest, 
+The pattern of conflict shows that we must recognise:
+%
+\begin{quote}
+`there are many separable systems of mental representations ...\ and thus many different kinds of knowledge.
+...\ the task ...\ is to contribute to the enterprise of finding the distinct systems of mental representation and to understand their development and integration%
+%.
+...\ the challenge ...\ is to characterize what it is about the mental representations of the presence or absence of the barriers that constrains representations of where the object should be in looking-time studies and how these representations differ from those that guide search or pointing in the present studies%
+'
+\citep[p.\ 1522]{Hood:2000bf}.
+\end{quote}
+%
+It is one thing to propose that there are multiple kinds of knowledge and quite another to make systematic sense of this possibility.
+In this chapter we will consider attempts to do this by appeal to notions of modularity \citep{Fodor:1983dg}
+and core knowledge \citep{Spelke:2007hb}. 
+Both attempts raise further questions.  
+These questions about how to make sense of the possibility that there are multiple kinds of knowledge 
+run through the following chapters.
 
-\subsection{Mindreading}
+\subsection{Chapter 3. Number: From Perceptual Experience to Knowledge}
+
+\subsection{Chapter 4. Seeing and Talking about Colours}
+
+\subsection{Chapter 5. Words Are Tools for Communication}
+
+\subsection{Chapter 6. Actions: Teleology and Mirroring Motor Awareness}
+\label{ch:actions}
+
+\subsection{Chapter 7. Mindreading}
 

### 2012-07-13 joint action and knowing others minds/interacting mindreaders.3.ps.corrections.tex
changes to minimal theor yof mind requested after accept --- include papa on Jonathan Bennett acknolwedge

--- 2012-06-20 joint action and knowing others minds/submit philosophical studies  2012-01/second submit/interacting_mindreaders_source/interacting mindreaders.3.ps.corrections.tex
@@
 
-%\def \papersize {a4paper}
-\def \papersize {letterpaper}
+\def \papersize {a4paper}
+%\def \papersize {letterpaper}
 \documentclass[12pt,\papersize]{extarticle}
@@
 
-\input{preamble_steve_paper2}
+\input{$HOME/Documents/submissions/preamble_steve_paper2}
 %\author{}
@@
 	contribute to pulling up the turnip or pull it up.
-Such intentions are agent neutral in this sense:
+Such intentions are agent-neutral in this sense:
 relative to such an intention an agent 
@@
 If no alternative goal ought to be salient to you and the other has done nothing to make an alternative salient, then it is not unreasonable for you to ignore the possibility that the other is unwilling to pursue the goal to which your actions will be directed.
-This might easily be the case where, for instance, you are in a tidy kitchen patiently chopping your way through a large pile of carrots; it is perhaps less likely to be the case in a more cluttered kitchen where you are surrounded by all kinds of potential ingredient in different states of preparation.
+This might easily be the case where, for instance, you are in a tidy kitchen patiently chopping your way through a large pile of carrots; it is perhaps less likely to be the case in a more cluttered kitchen where you are surrounded by all kinds of potential ingredients in different states of preparation.
 In general, how strongly the truth of the premises of the your-goal-is-my-goal inference supports the truth of the conclusion will depend (among other conditions) on how salient alternative goals to which your actions might be directed ought to be to you.
@@
 But interacting mindreaders who rely on your-goal-is-my-goal
-thereby rely on their targets' having correctly identified the goals of their actions.
+thereby rely on their targets having correctly identified the goals of their actions.
 Of course this is sometimes a reason not to rely on your-goal-is-my-goal.
@@
 Call this inference \emph{her-goal-is-his-goal}.  
-This inference shows that mere observers who understand interaction also have a route to knowledge of other's goals that avoids the problem of opaque means. 
+This inference shows that mere observers who understand interaction also have a route to knowledge of others' goals that avoids the problem of opaque means. 
 The significance of interaction for mindreading extends beyond cases in which the mindreader herself interacts.
@@
 
-\bibliography{phd_biblio}
+\bibliography{$HOME/endnote/phd_biblio}
 

### 2012-07-13 mindreading/minimal theory of mind (new version)/minimal theory of mind.nv.tex
changes to minimal theor yof mind requested after accept --- include papa on Jonathan Bennett acknolwedge

@@
 %for anonymous submisison
-\def \iauthor {}
-\def \iemail{}
+%\def \iauthor {}
+%\def \iemail{}
 %\date{}
@@
 %but don't forget to remove defs above as well --- otherwise it will be in the metadata
-\author{}
-\date{}
+%\author{}
+%\date{}
 
@@
 It is valuable to articulate the computational theory in some detail before formulating and testing conjectures about implementation.
+
+Our approach is inspired by a tradition of creature construction in philosophy, and in particular by Bennett’s construction in \textit{Linguistic Behaviour} (\citeyear{Bennett:1976rg}).
+Of course, where Bennett and others aimed to understand something about what it is for a subject to have desires, beliefs and other mental states, our main aim is to understand what might be involved in tracking and thinking about these things.
 
@@
 This reasoning concerns only the types of objects present and does not involve identity.
-Yet it is sufficient (together with some further inferences about the protagonists’ goal) for success.
+Yet it is sufficient (together with some further inferences about the protagonist’s goal) for success.
 So while Scott and Baillargeon’s (2010) findings are interesting in many respects, their implementation does not provide evidence that infants ascribe beliefs about identity.
@@
 If subjects sometimes represent encounters or perceptions as such, then representations of others’ encounters may interfere with expectations or judgements about reality or with judgements about one’s own perceptions and beliefs \citep[so-called `altercentric interference',][]{Samson:2010jm}.
-For example, representing a protagonists’ encounter with an object at a particular location may interfere with expectations about the object’s actual location.
+For example, representing a protagonist's encounter with an object at a particular location may interfere with expectations about the object’s actual location.
 Accordingly, evidence that there is such interference when in a situation where others have different perceptions or beliefs would provide indirect evidence for theory of mind cognition (minimal or full-blown) over behavioural strategies, for such interference would be difficult to explain if subjects were representing behaviours only.
 Evidence showing this kind of altercentric interference has recently been found in human adults \citep[]{en_1797, Samson:2010jm}.
-These studies show that adults’ judgements about the number of objects they could see in a visual scene were slower and more error-prone when the scene contained an irrelevant other whose perspective was different, suggesting that another’s perspective can cause altercentric interfere.
+These studies show that adults’ judgements about the number of objects they could see in a visual scene were slower and more error-prone when the scene contained an irrelevant other whose perspective was different, suggesting that another’s perspective can cause altercentric interference.
 

### 2012-07-19 developing mind/proposal developing mind.tex
working on proposal

@@
 
-\def \ititle {The Developing Mind: A philosophical introduction to issues in cognitive development}
+\def \ititle {The Developing Mind: \\ {A philosophical introduction to %issues in 
+cognitive development}
+}
 \def \isubtitle {-- Book Proposal --}
@@
 minds? 
-The question goes back to Plato or earlier and remains unanswered.
+This question goes back to Plato or earlier and remains unanswered.
 %Philosophers have been pursuing this question since Plato or before. 
@@
 %So far no one has convincingly answered the question, but two major breakthroughs have been made in the last three decades or so.
-Two recent scientific breakthroughs appear to bring us closer to an answer, and to show that the question is even less straightforward than philosophers have  assumed.
+Two recent scientific breakthroughs appear to bring us closer to an answer, and to show that the question is even less straightforward than previously  assumed.
 The first breakthrough is the discovery that preverbal infants enjoy surprisingly rich social abilities, abilities which
 % begin to be built from the first months of life and 
-may well be foundational for later linguistic abilities and enable the emergence of knowledge \citep[e.g.][]{Csibra:2009xr,Meltzoff:2007pj,Tomasello:2005wx}. 
-A second breakthrough concerns the use of increasingly sensitive---and sometimes controversial---methods to detect  sophisticated expectations concerning  causal interactions, numerosity, mental states and more besides in preverbal infants \citep[e.g.][]{Spelke:1990jn,Baillargeon:gx}.
+may well facilitate the subsequent acquisition of linguistic abilities and enable the emergence of knowledge \citep[e.g.][]{Csibra:2009xr,Meltzoff:2007pj,Tomasello:2005wx}. 
+A second breakthrough involves the use of increasingly sensitive---and sometimes controversial---methods to detect  expectations without relying on subjects' abilities to talk or act.  
+These methods have revealed sophisticated expectations about  causal interactions, numerosity, mental states and more besides in preverbal infants \citep[e.g.][]{Spelke:1990jn,Baillargeon:gx}.
 These expectations or the representations and processes underpinning them arguably also enable the emergence of knowledge. 
-Although each breakthrough has been discussed at length by leading psychologists, philosophers have yet to consider either systematically in print.
-Further, the two breakthroughs are associated with different camps, one nativist and the other Vygotskian, and have rarely been considered together as identifying twin factors enabling the emergence of knowledge.
-The proposed book will familiarise readers with findings related to both breakthroughs and explain their relevance to philosophical accounts of knowledge, action and experience.  
-Most importantly, the book aims to introduce readers to  philosophical issues raised by these findings in cognitive development.
+Although each breakthrough has been discussed at length by leading psychologists, philosophers have yet to consider them systematically in print.
+Further, the two breakthroughs are associated with different camps, one nativist, the other Vygotskian.
+Perhaps for this reason the breakthrough findings have rarely been considered together as identifying twin factors enabling the emergence of knowledge.
+The proposed book aims to introduce readers to  
+%example of new---the question about simple joint action
+new philosophical issues raised by these findings  
+%example of existing question: modularity
+and to explain their relevance to longstanding philosophical questions about the mind.
 
+%if we're talking about breakthroughs we'd better mention neuroscience
 Advances in neuroscience may be transforming parts of developmental psychology much as they have already transformed the study of perception and action \citep{Johnson:2005az}. 
-The proposed book focusses on discoveries rather than methods but will introduce readers to some relevant findings in developmental cognitive neuroscience along the way.
+Although detailed discussion of neuroscientific methods is beyond its scope, the proposed book will introduce a selection of relevant findings in developmental cognitive neuroscience along the way.
 
-The book is organised by domains of knowledge, so that one chapter concerns knowledge of objects, another focusses on  knowledge of number, and so on.  
+The proposed book is organised by domains of knowledge, so that one chapter concerns knowledge of objects, another focuses on  knowledge of number, and so on.  
 The domains are chosen so that each set of developmental findings is linked to one or more philosophical issues. 
-For instance, research on knowledge of objects gives bite to questions about modularity and the nature of tacit knowledge \citep{Davies:1989gg,Fodor:1983dg}; research on knowledge of number invites discussions of nativism \citep{Fodor:1981ep,Spelke:1998im}; and developmental findings on knowledge of colour may challenge some assumptions philosophers have made about relations between language, thought and perception \citep{Gilbert:2006yb,Regier:2009ve}.
+For instance, research on knowledge of objects gives bite to questions about modularity and the nature of tacit knowledge \citep{Fodor:1983dg,Davies:1989gg}; research on knowledge of number invites discussions of nativism \citep{Fodor:1981ep,Spelke:1998im}; and developmental findings on knowledge of colour may challenge certain assumptions philosophers have made about relations between language, thought and perception \citep{Gilbert:2006yb,Regier:2009ve}.
 
@@
 
-Students in psychology are often drawn to philosophy and should find in this book an accessible introduction to key distinctions and arguments. 
-It might even be used for courses in developmental psychology. (I once co-taught such a course in psychology with Jim Russell in Cambridge).
+Psychology students, undergraduate and graduate, are often drawn to philosophy and should find in this book an accessible introduction to key distinctions and arguments. 
+It might  be used for courses in developmental psychology. (I once co-taught such a course in psychology with Jim Russell in Cambridge).
 
@@
 
-Finally, the proposed book will complement the many excellent books which expound particular theories about the origins of knowledge; these include 
+Finally, the proposed book will complement some excellent books which expound particular theories about the origins of knowledge; these include 
 	\citet{carey:2009_origin},
@@
 	\citet{Elman:1996zd}.
-The proposed book will include critical discussion of theories presented in these volumes.  
-It's aims are clearly distinct from any of these volume's aims: the proposed book will review a range of existing theories and explore philosophical questions they raise.
+The proposed book will include critical discussion of theories presented in some of these  books.  
+It's aims are clearly distinct from those of these books: the proposed book will review a range of existing findings and theories and explore philosophical questions they raise.
 
-\section{Contents}
-(This plan may change.)
+\section{Contents (provisional)}
 
@@
 
-As well as explaining how research on developing minds bears on some existing philosophical issues and raises some new ones,
+In explaining how research on developing minds bears on some existing philosophical issues and raises some new ones,
 this chapter introduces two scientific breakthroughs that have recently furthered understanding of how knowledge might emerge in development.
@@
 may well be foundational for later linguistic abilities and enable the emergence of knowledge \citep[e.g.][]{Csibra:2009xr,Meltzoff:2007pj,Tomasello:2005wx}. 
-A second breakthrough concerns the use of increasingly sensitive---and sometimes controversial---methods to detect  sophisticated expectations concerning  causal interactions, numerosity, mental states and more besides in preverbal infants \citep[e.g.][]{Spelke:1990jn,Baillargeon:gx}.
+A second breakthrough is the use of increasingly sensitive---and sometimes controversial---methods to detect  sophisticated expectations concerning  causal interactions, numerosity, mental states and more besides in preverbal infants \citep[e.g.][]{Spelke:1990jn,Baillargeon:gx}.
 These breakthroughs are driven by researchers with conflicting theoretical positions and raise quite different issues.  
@@
 %for a review on imitation, see \citep{jones:2009_development}
-imitative learning, 
+imitative learning \citep{carpenter:1998_fourteen}, 
 gaze following \citep{Csibra:2008be},
@@
 and 
-pointing. 
+pointing \citep{Liszkowski:2006ec}.  
 Taken together, the evidence reveals that preverbal infants have surprisingly rich social abilities.
@@
 After all, we can hardly explain the emergence of something by appealing to abilities whose possession already presuppose it.
-We will encounter tools for replying to this objection in Chapters 2 (on Objects) and 6 (on Actions).
+We will encounter tools for replying to this objection in Chapters 2 (on Objects), 6 (on Actions) and 7 (on Minds).
 
-The hypothesis that social abilities enable cognitive development naturally invites a question: How could that work? 
-Some have argued that early social abilities partially explain the emergence in humans of communication by language (\citealp{tomasello:2008origins}; see Chapter 5 on Words). 
+If this objection can be overcome, how could social abilities enable cognitive development? 
+Some have argued that early social abilities partially explain the emergence in humans of communication by language (\citealp{tomasello:2008origins}; see Chapter 5 on Words below). 
 This is a valuable contribution.  
@@
 %
-It is one thing to propose that there are multiple kinds of knowledge and quite another to make systematic sense of this possibility.
+It is one thing to propose that there are multiple kinds of knowledge, quite another to make systematic sense of this possibility.
 In this chapter we will consider attempts to do this by appeal to notions of modularity \citep{Fodor:1983dg}
@@
 
-\subsection{Chapter 3. Number: From Perceptual Experience to Knowledge}
+\subsection{Chapter 3. Numbers}
+
+How might abilities based on core knowledge enable the emergence in development of knowledge proper? 
+%(\textit{Knowledge proper} is knowledge not subject to the limits associated with  core knowledge.)
+In this chapter we consider answers to this question for the special case of knowledge of number.
+
+In Chapter 2 (on Objects) we saw that core knowledge of objects is manifested early in infancy, whereas more flexible abilities involving knowledge proper first appear months or even years later. 
+A similar pattern can be observed in the case of number. 
+From around five months of age or earlier, infants 
+are sensitive to the number of items in a repeatedly presented stimulus \citep{starkey:1980_perception}, 
+can sum over the number of items in a short sequence \citep{wynn:1996_infants}, 
+can discriminate between small and large sets of objects \citep{xu:2000_large}, 
+and can keep track of the precise number of items in a small set of objects following additions or subtractions \citep{wynn:1992_addition}.
+
+Should we conclude that infants already have adult-like number concepts?  
+Limits on infants' abilities show that we should not. 
+To illustrate, 
+at around fourteen months of age infants who have seen two objects put into a box will search in it until they have recovered two objects, and infants who have seen three objects put into the box will search three times.
+But after seeing four objects placed in the box, infants usually search just once \citep{feigenson_tracking_2003}. 
+In general, infants' abilities to deal with precise numerosity are strictly limited to situations involving at most three items \citep{feigenson_tracking_2003}.
+%debate about exact number?
+It seems clear, then, that infants do not have adult-like number concepts.
+
+The three-item limit on infants' abilities to deal with precise numerosity is the key to a revealing discovery.
+Many species besides humans have abilities to deal with precise numerosity which are also subject to limits like infants' three-item limit; this suggests that they are  closely related \citep{hauser:2003_spontaneous}.
+%In fact, infant-like, limited abilities can also be found in human adults who have number concepts. 
+This tells us that infants' core knowledge of number is unlikely to be a product of learning \citep{feigenson:2004_core} and so supports a modest form of nativism.
+
+A bolder form nativism would insist not only that core knowledge of number is innate but also that number concepts proper are innate \citep{Fodor:1981ep}. 
+Against this bolder view, some have attempted to explain how core knowledge together with mastery of a sequence number words could explain how number concepts can be acquired \citep{carey:2009_origin}.
+If successful, this explanation shows that not all primitive concepts are innate and provides us with one model of how core knowledge might enable the emergence of knowledge proper.
 
 \subsection{Chapter 4. Seeing and Talking about Colours}
+How do children acquire colour concepts and colour words---concepts and words for red, blue and green, say? 
+Exploring this question reveals unexpected complexity in the developmental origins of knowledge, as we will see in this chapter.
 
-\subsection{Chapter 5. Words Are Tools for Communication}
+%Frank Jackson is best known for the story of Mary, a brilliant scientist who spends the first part of her stuck in a monochrome environment learning all that can be known from books about everything including the colour blue \citep[p.\ 130]{jackson:1982_epiphenomenal}.
+%Then one day she is set free, goes out into the world and learns something she could not have learnt from a book, namely what it is like for others to sense blue \citep[p.\ 292]{jackson:1986_what}.
+%Although the story seems fantastic, it echoes a developmental pattern.
 
-\subsection{Chapter 6. Actions: Teleology and Mirroring Motor Awareness}
+Categorical perception of colour emerges early in infancy.  This has been demonstrated with four-month-olds using habituation \citep{Bornstein:1976of} and visual search \citep{Franklin:2005xk}.   
+%say something about ERP / oddball study \citep{Clifford:2009fo}
+Slightly older infants can make use of colour properties such as red and green to recognise objects.  
+For instance, nine-months-olds can determine whether an object they saw earlier is the same as a subsequently presented object on the basis of its colour \citep{Wilcox:2008jk}.  
+By the time they are two years old, toddlers who do not comprehend any colour terms can use colour categories implicitly in learning and using proper names; for instance, they are able to learn and use proper names for toy dinosaurs that differ only in colour \citep[][Experiment 3]{Soja:1994np}.
+So infants and toddlers enjoy categorical perception of colour and may benefit from it in recognising and learning about objects.  
+However children only acquire concepts of, and words for, colours some time later; and colour concepts, like colour words, are acquired gradually \citep{Pitchford:2005hm,Kowalski:2006hk,Sandhofer:1999if,Sandhofer:2006qo}.
+
+A natural hypothesis, then, is that the acquisition of colour words and concepts builds on categorical perception of colour: infants are first perceptually acquainted with colours and then learn names for them.
+And since infants enjoy categorical perception not only of colour but also of orientation \citep{franklin:2010_hemispheric}, speech \citep{Kuhl:1987la,Kuhl:2004nv,Jusczyk:1995it} and facial expressions of emotion \citep{Etcoff:1992zd,Kotsoni:2001ph,Campanella:2002aa},
+we might suppose that something similar holds in these cases too.
+The hypothesis, then, is that categorical perception provides `the building blocks—the elementary units—for higher-order categories' \citep[p.\ 3]{Harnad:1987ej}.
+
+This hypothesis turns out to be entirely mistaken. 
+The extensions of colour words may be subject to universal constraints but they certainly vary significantly between languages  \citep{kay:2003_resolving}.
+%can use figure from Kay & Regier 2007 Berinmo on p. 291
+The boundaries of adult humans' colour concepts and their perceptual colour categories  are influenced by the extensions of the colour words they use \citep{Kay:2006ly,Roberson:2007wg,Winawer:2007im}.
+It is even possible to alter the boundaries of adults' perceptual categories by teaching them new colour terms
+\citep{Ozgen:2002yk}. 
+But infants' perceptual colour categories are not influenced by colour words; after all, they don't use colour words. 
+Indeed, in toddlers who have recently acquired colour terms, the extensions of their colour concepts do not match their perceptual category boundaries \citep{Franklin:2005hp}. 
+So it is a mistake to think of the development of colour words and concepts as a matter of learning to label or demonstratively refer to categories infants are already perceptually acquainted with. 
+Quite the reverse: acquiring colour words and concepts is a prerequisite for being able to perceive the colour categories they refer to.
+
+%***Something about different processes:
+%Adult-mode categorical perception of colour disappears in the face of predictable verbal interference but not non-verbal interference (*Roberson, Davies and Davidoff 2000: 985; Pilling, Wiggett, et al.
+2003: 549–50; Wiggett and Davies 2008).
+%***Maybe also mention that language can play a role.
+
+Can the case of colour shed light on the question of how core knowledge enables the emergence of knowledge proper?
+There are some parallels between categorical perception and core knowledge: for instance, both phenomena are judgement-independent and have limited effects on action. 
+Indeed, there seems no less reason to postulate core knowledge of colour or speech than there is to postulate core knowledge of objects or agents. 
+So the case of colour suggests that the development of knowledge proper is not always a matter of making explicit, or building on, what was earlier implicit in core knowledge. 
+In fact core knowledge might sometimes be in competition with knowledge proper. 
+%It should not be taken for granted that the several kinds of knowledge involved in development are harmoniously integrated. 
+%Rather, core knowledge might sometimes be in competition with knowledge proper.
+
+\subsection{Chapter 5. Words and Other Communicative Tools}
+One problem in acquiring a language is to identify which things are words; among much else, this involves working out where, in a stream of speech or gesture, one word ends and another begins.  
+In solving this problem infants may rely on statistical learning \citep{Saffran:1996aj}.
+Another problem is learning which words pick out which objects and properties; this is often called the `mapping problem'. 
+Philosophers have sometimes held that in solving this problem children rely on association (or perhaps other forms of statistical learning).
+Compare Wittgenstein (\citeyear[p.\ 77]{Wittgenstein:1972lj}): 
+%
+\begin{quote}
+`The child learns this language from the grown-ups by being trained to its use.
+I am using the word ``trained'' in a way strictly analogous to that in which we talk of an animal being trained to do certain things.
+It is done by means of example, reward, punishment, and suchlike.'
+%See also Russell (1921: 71): “A child learning to speak is learning habits and associations which are just as much determined by the environment as the habit of expecting dogs to bark and cocks to crow.
+The community that speaks a language has learnt it, and modified it by processes almost all of which are not deliberate, but the results of causes operating according to more or less ascertainable laws.”
+\end{quote}
+%
+If this sort of view were right, abilities to communicate linguistically would not necessarily depend on knowledge proper and so it would be possible to appeal to linguistic communication in explaining the developmental origins of knowledge.
+Against Wittgenstein, experimental findings support the hypothesis that children solve the mapping problem by general reasoning and rely on social cognition
+\citep{Baldwin:1995xl,Bloom:2000qz,Sabbagh:2001sp}. 
+%Also talk about language creation here (see lexical innovation paper)
+
+This hypothesis conflicts with Davidson's claims that having thoughts involves grasping the concept of truth \citep[p.\ 189]{Davidson:2001sm}, 
+% `What gives my belief its content ... is my knowledge of what is required for the belief ... to be true' \citep[p.\ 189]{Davidson:2001sm}, 
+and that `we grasp the concept of truth only when we can communicate the contents—the propositional contents—of the shared experience, and this requires language' \citep[p.\ 27]{Davidson:1997wj}.
+The objection to Davidson's view is that that it has things backwards: acquiring your first words already involves thinking about what they pick out. 
+If this objection is correct, can anything be salvaged from Davidson's arguments? 
+While the objection rules out claiming that being able to think depends on being able to communicate propositional contents using language, it leaves open the possibility that thought is essentially intersubjective.
+One challenge, then, is to explain the intersubjectivity of thought without appeal to communication by language.
+
+So far our focus has been the mapping problem, the problem of determining which words pick what out. 
+A distinct and no less difficult problem is how infants come to understand words as communicative tools at all.
+Merely associating a word or gesture with what it picks out would not enable an infant to respond appropriately to uses of that word or gesture, nor to use it intelligently. 
+After all, chimpanzees associate pointing gestures with their referents \citep[][p.\ 6]{Moll:2007gu} but fail to respond appropriately to helpful pointing gestures \citep{hare_chimpanzees_2004}. 
+Unlike chimpanzees, infants not only map words to what they pick out but also use and understand them as tools for communication. 
+What does this understanding amount to---in particular, do infants have an adult-like understanding of communicative intention---and how do infants acquire it?
+%On one view, it involves a special `action interpretation system' distinct from that involved in understanding ordinary, non-communicative action \citep[p.\ 456]{Csibra:2003kp}. 
+%***This question is pressing because we want to appeal to language to explain the emergence of full-blown mindreading
+Answering this question may require revisions or extensions to existing philosophical accounts of communication which often presuppose conceptual sophistication \citep[e.g.][]{Grice:1957kg}.
+
+\subsection{Chapter 6. Actions: Teleology and Motor Awareness}
 \label{ch:actions}
+***
 
-\subsection{Chapter 7. Mindreading}
+\subsection{Chapter 7. Beliefs}
+What is involved in representing belief?
+The following three claims are separately defensible but appear to be jointly inconsistent: 
+%
+\begin{enumerate}
+\item Infants can represent false beliefs from around their first birthday or earlier. \label{infant_fb}
+%
+\item  \label{fb_is_perspectives}  Being able to represent false beliefs involves being able to  (i) process perspective differences or (ii) reason counterfactually (or both).
+%
+\item Infants cannot (i) process perspective differences nor (ii) engage in counterfactual reasoning until they are at least one year old.  \label{infant_perspectives}
+%
+\end{enumerate}
+%
+This chapter will first consider evidence and theoretical arguments in support of each claim in turn.
+
+There is a growing body of evidence for (\ref{infant_fb}).
+From around their first birthday infants  predict actions of agents with false beliefs about the locations of objects \citep[]{Onishi:2005hm, Southgate:2007js}. 
+and choose different ways of interacting with others depending on whether their beliefs are true or false \citep[]{Buttelmann:2009gy,Knudsen:2011fk,southgate:2010fb}.  
+And in much the way that irrelevant facts about the contents of others’ beliefs modulate adult subjects’ response times, such facts also affect how long 7-month-old infants look at some stimuli \citep[]{kovacs_social_2010}.
+The variety of  measures---looking time, anticipatory looking, pointing and helping---makes it hard to dismiss these findings on methodological grounds.
+So we cannot straightforwardly reject (\ref{infant_fb}).
+
+Why accept (\ref{fb_is_perspectives})?  
+Part of the answer is theoretical. 
+On any standard view, beliefs are propositional attitudes and are individuated (at least in part) by their causal and normative roles in explaining thoughts and actions \citep[]{Davidson:1980xp, Davidson:1990du}.
+Representing states with these features arguably involves being able to process perspective differences and a core component of genuine counterfactual reasoning \citep{perner:2007_objects}. 
+Another part of the argument for accepting (\ref{fb_is_perspectives}) is empirical.
+Until they are around four years of age,
+children systematically fail a wide range of false belief tasks.
+These tasks are very varied: most are verbal but some are nonverbal (\citealp{Call:1999co}; \citealp{low:2010_preschoolers} Study 2),
+some involve prediction whereas others involve retrodiction or justification \citep[e.g.][]{Wimmer:1998kx},
+some concern the first-person perspective, whereas others involve a second- or third-person perspective \citep[e.g.][]{Gopnik:1991db},
+some involve interaction whereas in others the subject is a mere observer \citep[e.g.][]{Chandler:1989qa},
+and some involve prediction actions whereas others involve predicting desires \citep{Astington:1991kk} or selecting an argument appropriate for someone with a false belief \citep{Bartsch:2000es}. 
+Despite all this variation and more, these false belief tasks all appear to measure a single developmental transition \citep{Wellman:2001lz},
+and that developmental transition is linked to developments in abilities to process perspective difficulties \citep{Perner:2002jj} and to reason counterfactually \citep{rafetseder:2012_submitted}.
+While some researchers have claimed that the developmental transition these false belief tasks measure is not a transition in understanding belief 
+	(e.g.\ 
+	\citealp[][p.\ 417]{Carpenter:2002gc},
+	\citealp{Bloom:2000bt}, and
+	\citealp{Leslie:1998nq}),
+the wide variety of tasks used is evidence against such claims.
+So we cannot straightforwardly reject  (\ref{fb_is_perspectives}).
+
+Can we reject (\ref{infant_perspectives}), the claim that infants neither process perspective differences nor engage in counterfactual reasoning until they are at least one year old?
+One difficulty is that current developmental evidence almost uniformly supports this claim  \citep{rafetseder:2010_counterfactual,beck:2011_almost}.
+A second problem is that the development of abilities to reason counterfactually,
+%\footnote{Here and throughout I follow Eva Rafetseder in using the term `counterfactual reasoning'  to refer to reasoning which essentially involves comparing two possible situations; the term is sometimes used more broadly to include the sorts of representations involved in pretence.}
+like the development of abilities to represent false belief,
+appears to involve working memory and inhibitory control 
+(on counterfactuals: 
+	\citealp{drayton:2011_counterfactual,beck:2011_supporting};
+on false beliefs: \citealp{apperly:2008_back, Apperly:2009cc}; \citealp{lin:2010_reflexively, McKinnon:2007rr} experiments 4-5; \citealp{saxe_reading_2006}).
+As even committed nativists are likely to agree, capacities for inhibitory control and working memory develop over several years and are limited in infants \citep[e.g.][]{carlson:2005_developmentally}.
+So we cannot straightforwardly reject (\ref{infant_perspectives}).
+
+If each of the three claims (1)-(3) is true, they cannot be inconsistent after all. 
+Can we avoid the apparent inconsistency by appealing to the distinction between core knowledge (or modular representation) and knowledge proper introduced in Chapter 2 (on Objects)?  
+One objection to such an appeal concerns the flexibility of false belief understanding in infants.
+One-year-old infants succeed on false belief tasks which involve actively helping others, interpreting their utterances and pointing to provide information \citep{Buttelmann:2009gy,Knudsen:2011fk,southgate:2010fb}.
+By contrast, core knowledge is supposed to have only limited consequences for purposive action.
+
+To the apparent inconsistency of (1)-(3) there are three main responses.
+One is to reject (1) \citep[e.g.][]{Perner:2005hq}, 
+another is to reject (2),
+ and a third is to argue that the inconsistency is only apparent by appeal to a distinction such as that between core knowledge and knowledge proper \citep[e.g.][]{Clements:1994cw,low:2010_preschoolers}.
+We have seen objections to each of these responses.
+This indicates that we do not adequately understand what is involved in representing beliefs and, more generally, what it is to have a theory of mind.
+
+%This is intended as leading in to measurement and minimal theory of mind
 
 \subsection{Conclusion}
+Are we closer to understanding 
+how humans come to know about %
+objects,
+causes,
+words,
+numbers,
+colours,
+actions
+and
+minds? 
+***
 

### 2012-07-19 mindreading/conf espp 2012 symposium false belief counterfactuals perspective schurz rafetseder/false beliefs, perspectives and counterfactuals espp 2012 symposium submission.tex
typos

@@
 %
-\item Infants cannot (i) process perspective differences nor (ii) engage in counterfactual reasoning until they are at least one year old.  \label{infant_perspectives}
+\item Infants cannot (i) process perspective differences nor (ii) engage in counterfactual reasoning until they are well over one year old.  \label{infant_perspectives}
 %
@@
 
-\item Infants cannot (i) process perspective differences nor (ii) engage in counterfactual reasoning until they are at least one year old.  \label{infant_perspectives}
+\item Infants cannot (i) process perspective differences nor (ii) engage in counterfactual reasoning until they are well over one year old.  \label{infant_perspectives}
 
@@
 This is a significant obstacle to identifying the Extraneous Requirement.
-In fact, my view the currently available evidence supports the conclusion that there is no Extraneous Requirement.
+In fact, in my view the currently available evidence supports the conclusion that there is no Extraneous Requirement.
 And as far as I know the only plausible candidate for the transition  Category A tasks measure is a transition in the ability to represent beliefs.

### 2012-07-20 developing mind/proposal developing mind.tex
working on proposal

@@
 Two recent scientific breakthroughs appear to bring us closer to an answer, and to show that the question is even less straightforward than previously  assumed.
-The first breakthrough is the discovery that preverbal infants enjoy surprisingly rich social abilities, abilities which
-% begin to be built from the first months of life and 
-may well facilitate the subsequent acquisition of linguistic abilities and enable the emergence of knowledge \citep[e.g.][]{Csibra:2009xr,Meltzoff:2007pj,Tomasello:2005wx}. 
+The first breakthrough concerns social interaction;
+is the discovery that preverbal infants enjoy surprisingly rich social abilities.
+These abilities enable infants to engage in some forms of social interaction. 
+This may well facilitate the subsequent acquisition of linguistic abilities and enable the emergence of knowledge \citep[e.g.][]{Csibra:2009xr,Meltzoff:2007pj,Tomasello:2005wx}. 
 A second breakthrough involves the use of increasingly sensitive---and sometimes controversial---methods to detect  expectations without relying on subjects' abilities to talk or act.  
-These methods have revealed sophisticated expectations about  causal interactions, numerosity, mental states and more besides in preverbal infants \citep[e.g.][]{Spelke:1990jn,Baillargeon:gx}.
+These methods have revealed sophisticated expectations about  causal interactions, numerosity, actions, mental states and more besides in preverbal infants \citep[e.g.][]{Spelke:1990jn,Baillargeon:gx}.
 These expectations or the representations and processes underpinning them arguably also enable the emergence of knowledge. 
@@
 The domains are chosen so that each set of developmental findings is linked to one or more philosophical issues. 
-For instance, research on knowledge of objects gives bite to questions about modularity and the nature of tacit knowledge \citep{Fodor:1983dg,Davies:1989gg}; research on knowledge of number invites discussions of nativism \citep{Fodor:1981ep,Spelke:1998im}; and developmental findings on knowledge of colour may challenge certain assumptions philosophers have made about relations between language, thought and perception \citep{Gilbert:2006yb,Regier:2009ve}.
+For instance, research on knowledge of objects gives bite to questions about modularity and the nature of tacit knowledge%\citep{Fodor:1983dg,Davies:1989gg}
+; research on knowledge of number invites discussions of nativism% \citep{Fodor:1981ep,Spelke:1998im}
+; and developmental findings on knowledge of colour may challenge certain assumptions philosophers have made about relations between language, thought and perception.
 
@@
 Professional philosophers whose research connects with issues in cognitive development may also find the book useful.  
-Judging by papers and manuscripts under review for journals and university presses, even those writing about cognitive development can have difficulty identifying the full range of findings relevant to their positions.
+Judging by papers and manuscripts under review, philosophers sometimes have difficulty identifying the full range of findings in cognitive development relevant to their positions.
 
@@
 It might  be used for courses in developmental psychology. (I once co-taught such a course in psychology with Jim Russell in Cambridge).
+
+For readers captivated by Gopnik's popular \textit{The Philosophical Baby} \citep{gopnik:2009_philosophical} who can cope with academic texts, the proposed book might provide bridge to studying philosophy or psychology.
 
@@
 \section{Timetable}
-I aim to submit a first draft of the manuscript by 30 September 2014.
+I plan to submit a first draft of the manuscript by 30 September 2014.
 
-I plan to try out material for the book while teaching courses at the University of Warwick (UK) and Central European University (Hungary).
+I aim to try out material for the book while teaching courses at  Central European University (Hungary) and the University of Warwick (UK).
 
 \section{Competition}
-At the time of writing there is no book or collection devoted to philosophical issues in cognitive development.
+At the time of writing there is no book or collection devoted to philosophical issues in cognitive development other than \textit{The Philosophical Baby} \citep{gopnik:2009_philosophical} which is primarily aimed at general readers.
 
 Philosophers have written monographs and edited collections on particular issues in cognitive development \citep[e.g.][]{Bermudez:2003dj,carruthers:1996_theories}.  
-The proposed book complements these by providing a unified discussion of a wider range of topics. 
+The proposed book complements these by providing integrated discussion of a wider range of topics. 
 The advantage is not just convenience: understanding and properly evaluating theories about the developmental origins of knowledge requires bringing together research on different domains of knowledge.
@@
 There are also several collections bringing together research by philosophers and developmental psychologists or presenting developmental research in ways accessible to philosophers 
-\citep[e.g.][]{hirschfeld:1994_mapping,carruthers:2005_innate_structure,carruthers:2006_innate_culture,mccormack:2011_tool}.
-From the point of view of the proposed book, these provide useful sources of further reading for those who want more information on particular issues.
+\citep[e.g.][]{Whiten:1991qn,hirschfeld:1994_mapping,carruthers:2005_innate_structure,carruthers:2006_innate_culture,hoerl:2011_understanding}.
+In relation to the proposed book, these provide useful sources of further reading for those who want more information on particular issues.
 
-Finally, the proposed book will complement some excellent books which expound particular theories about the origins of knowledge; these include 
-	\citet{carey:2009_origin},
+Finally, the proposed book will complement some excellent books which introduce particular theories of the origins of knowledge; these include 
+	\citet{Elman:1996zd},
+	\citet{Gopnik:1997xq},
 	\citet{Tomasello:1999xz},
-	\citet{Gopnik:1997xq},
 	and
-	\citet{Elman:1996zd}.
+	\citet{carey:2009_origin}.
 The proposed book will include critical discussion of theories presented in some of these  books.  
@@
 In pursuing this question we have to consider minds where the knowledge is neither clearly present nor obviously absent. 
-This is challenging because both commonsense and theoretical tools for describing minds are generally designed for characterising fully developed adults. 
+This is challenging because both commonsense and theoretical tools for describing minds are generally suitable only for fully developed adult minds. 
 Davidson writes:
@@
 Philosophers have much to contribute in crafting distinctions and conceptual tools useful for meeting the challenges involved in describing what is in between. 
-There is also much for philosophers to gain: studying what is in between mindless nature and the sorts of cognition captured by  adult humans' everyday, pre-theoretical mentalistic notions 
+There is also much for philosophers to gain: studying what is in between mindless nature and the sorts of cognition captured by   commonsense psychological notions 
  will reveal things about what minds are and how they work.
@@
 These breakthroughs are driven by researchers with conflicting theoretical positions and raise quite different issues.  
-But they both  identify ingredients necessary for understanding the developmental origins of human knowledge.
+But both breakthroughs are needed to identify ingredients necessary for understanding the developmental origins of human knowledge.
 
 \subsection{Chapter 1. Social Interaction without Words}
-Could social interaction be an enabler of cognitive development? 
-This chapter examines arguments for the hypothesis that it could
- (as offered in \citealp{Tomasello:2003tq} and \citealp{Moll:2007gu}); it also discusses objections to this hypothesis and introduces questions it raises.
+Could social interaction enable cognitive development? 
+This chapter examines arguments for the hypothesis that it could (as offered in \citealp{Tomasello:2003tq} and \citealp{Moll:2007gu}). 
+It also discusses objections to this hypothesis and introduces questions it raises.
 
-An immediate problem is that much adult social interaction depends so heavily on communication by language (not to mention social networks) that it can be hard to imagine what social interaction with preverbal infants could involve.
-Relatedly, many theoretical accounts of social interaction presuppose linguistic abilities.
-To overcome this difficulty we shall review evidence for a package of social abilities manifested in preverbal infants.
+One obstacle to even understanding the hypothesis is that much adult social interaction depends so heavily on communication by language (not to mention social networks) that it can be hard to imagine what social interaction with preverbal infants could involve.
+To overcome this obstacle we shall review evidence for a package of social abilities manifested in preverbal infants.
 These include imitation, which can occur just days and even minutes after birth \citep{meltzoff:1977_imitation,field:1982_imitation,meltzoff:1983_newborn}, 
@@
 For instance, twelve-month old infants  will helpfully point to inform ignorant but not knowledgable adults about the location of an object  \citep{Liszkowski:2008al}.
-Some researchers take this to show that these infants are already capable of knowing about others' knowledge and ignorance. 
-If that is right, pointing and other social abilities could play at most a limited role in explaining the developmental origins of knowledge of mind. 
+Doesn't this show that these infants are already capable of knowing about others' knowledge and ignorance? 
+%could also ask whether imitation involves understand action --- after all, need to understand what to imitate!
+If so, pointing and other social abilities could play at most a limited role in explaining the developmental origins of knowledge of mind. 
 After all, we can hardly explain the emergence of something by appealing to abilities whose possession already presuppose it.
@@
 One way to test for object permanence is to ask when infants first reach for objects they can’t see or when they first remove barriers to retrieve objects concealed behind them.  
-Infants don’t do this until around eight months \citep[p.\ 202]{Meltzoff:1998wp} or maybe later \citep{moore:2008_factors}.  
+Infants don’t do this until around eight months of age \citep[p.\ 202]{Meltzoff:1998wp} or  later \citep{moore:2008_factors}.  
 Since four-month-olds already have the planning skills they would need to execute the reach \citep{Shinskey:2001fk}, 
@@
 %}
- As it turns out, infants show such heightened interest from around four months or earlier. 
+ As it turns out, infants show such heightened interest from around four months of age or earlier. 
 Put together, the two sorts of findings give rise to the `paradox of early permanence' (as \citealp{Meltzoff:1998wp} call it). 
@@
 %As some psychologists reflecting on difficulties in understanding knowledge of objects suggest, 
-The pattern of conflict shows that we must recognise:
+In fact:
 %
 \begin{quote}
-`there are many separable systems of mental representations ...\ and thus many different kinds of knowledge.
-...\ the task ...\ is to contribute to the enterprise of finding the distinct systems of mental representation and to understand their development and integration%
+`there are many separable systems of mental representations ...\ and thus many different kinds of knowledge. ...\ 
+[T]he task ...\ is to contribute to the enterprise of finding the distinct systems of mental representation and to understand their development and integration% 
 %.
@@
 The three-item limit on infants' abilities to deal with precise numerosity is the key to a revealing discovery.
-Many species besides humans have abilities to deal with precise numerosity which are also subject to limits like infants' three-item limit; this suggests that they are  closely related \citep{hauser:2003_spontaneous}.
+Many species besides humans have abilities to deal with precise numerosity.
+Because these abilities are also subject to limits like infants' three-item limit, it is reasonable to hold that they are  closely related \citep{hauser:2003_spontaneous}.  
 %In fact, infant-like, limited abilities can also be found in human adults who have number concepts. 
-This tells us that infants' core knowledge of number is unlikely to be a product of learning \citep{feigenson:2004_core} and so supports a modest form of nativism.
+This makes it unlikely that infants' core knowledge of number is  a product of learning \citep{feigenson:2004_core}, and so supports a modest form of nativism.
 
@@
 Against this bolder view, some have attempted to explain how core knowledge together with mastery of a sequence number words could explain how number concepts can be acquired \citep{carey:2009_origin}.
-If successful, this explanation shows that not all primitive concepts are innate and provides us with one model of how core knowledge might enable the emergence of knowledge proper.
+If successful, this explanation shows that not all primitive concepts are innate and provides us with one model of how core knowledge together with linguistic abilities might enable the emergence of knowledge proper.
 
@@
 For instance, nine-months-olds can determine whether an object they saw earlier is the same as a subsequently presented object on the basis of its colour \citep{Wilcox:2008jk}.  
-By the time they are two years old, toddlers who do not comprehend any colour terms can use colour categories implicitly in learning and using proper names; for instance, they are able to learn and use proper names for toy dinosaurs that differ only in colour \citep[][Experiment 3]{Soja:1994np}.
+By the time they are two years old, toddlers who do not comprehend any colour words can use colour categories implicitly in learning and using proper names; for instance, they are able to learn and use proper names for toy dinosaurs that differ only in colour \citep[][Experiment 3]{Soja:1994np}.
 So infants and toddlers enjoy categorical perception of colour and may benefit from it in recognising and learning about objects.  
@@
 The boundaries of adult humans' colour concepts and their perceptual colour categories  are influenced by the extensions of the colour words they use \citep{Kay:2006ly,Roberson:2007wg,Winawer:2007im}.
-It is even possible to alter the boundaries of adults' perceptual categories by teaching them new colour terms
+It is even possible to alter the boundaries of adults' perceptual categories by teaching them new colour words
 \citep{Ozgen:2002yk}. 
-But infants' perceptual colour categories are not influenced by colour words; after all, they don't use colour words. 
-Indeed, in toddlers who have recently acquired colour terms, the extensions of their colour concepts do not match their perceptual category boundaries \citep{Franklin:2005hp}. 
+But preverbal infants' perceptual colour categories are not influenced by colour words. 
+Indeed, in toddlers who have recently acquired colour words, the extensions of their colour concepts do not match their perceptual category boundaries \citep{Franklin:2005hp}. 
 So it is a mistake to think of the development of colour words and concepts as a matter of learning to label or demonstratively refer to categories infants are already perceptually acquainted with. 
@@
 It is done by means of example, reward, punishment, and suchlike.'
+%careful with this quote --- from the context Wittgenstein is imagining how things might be, not how they are.
+%but this quote shows that he is serious about how children are:
+%`A child uses such primitive forms of language when it learns to talk. Here the teaching of language is not explanation, but training.' \citep[§5, p.\ 4]{Wittgenstein:1953mm}
+%And:
+%`The primitive language-game which children are taught needs no justification; attempts at justification need to be rejected.' \citep[p.\ 200]{Wittgenstein:1953mm}
+%And:
+%`I myself wrote in my book that children learn to understand a word in such and such a way.
+Do I know that, or do I believe it?
+Why in such a case do I write not "I believe etc." but simply the indicative sentence?' \citep[§290, p.\ 38]{Wittgenstein:1974dk}
+%
 %See also Russell (1921: 71): “A child learning to speak is learning habits and associations which are just as much determined by the environment as the habit of expecting dogs to bark and cocks to crow.
@@
 If this sort of view were right, abilities to communicate linguistically would not necessarily depend on knowledge proper and so it would be possible to appeal to linguistic communication in explaining the developmental origins of knowledge.
-Against Wittgenstein, experimental findings support the hypothesis that children solve the mapping problem by general reasoning and rely on social cognition
+%Compare Wittgenstein again:
+%`535. The child knows what something is called if he can reply correctly to the question "what is that called?"
+%536. Naturally, the child who is just learning to speak has not yet got the concept is called at all.
+%537. Can one say of someone who hasn't this concept that he knows what such-and-such is called?
+%538. The child, I should like to say, learns to react in such-and-such a way; and in so reacting it doesn't so far know anything. Knowing only begins at a later level.'
+%\citep[p.\ 71]{Wittgenstein:1974dk}
+This does indeed seem plausible for the special case of colour words, as we saw in Chapter 4 (on Colours); 
+and something like it may hold for number words too (see Chapter 3 on Numbers).  
+%ref to Wittgenstein on child learning about numbers somewhere?
+But putting these special cases aside, the experimental findings do not support Wittgenstein's view.
+Instead they support the hypothesis that children solve the mapping problem by general reasoning and rely on social cognition
 \citep{Baldwin:1995xl,Bloom:2000qz,Sabbagh:2001sp}. 
@@
 and that `we grasp the concept of truth only when we can communicate the contents—the propositional contents—of the shared experience, and this requires language' \citep[p.\ 27]{Davidson:1997wj}.
-The objection to Davidson's view is that that it has things backwards: acquiring your first words already involves thinking about what they pick out. 
+% also:
+%`language ...
+is, or has become, a mode of perception.
+However, speech is not just one more organ; it is essential to the other senses if they are to yield propositional knowledge' \citep[p.\ 22]{Davidson:1997wj}
+% also:
+%`Mastery of the word or concept [`table'] requires in addition, I believe, that the child understand that error is possible, that "That's a table' expresses a judgment that has a truth value independent of its merely being uttered.' \citep[p.\ 9]{Davidson:2001np}
+%
+The objection to Davidson's view is that  it has things backwards: acquiring your first words already involves thinking about what they pick out. 
 If this objection is correct, can anything be salvaged from Davidson's arguments? 
@@
 A distinct and no less difficult problem is how infants come to understand words as communicative tools at all.
-Merely associating a word or gesture with what it picks out would not enable an infant to respond appropriately to uses of that word or gesture, nor to use it intelligently. 
+Merely associating a word or gesture with what it picks out would not enable a child to respond appropriately to uses of that word or gesture, nor to use it intelligently (contra \citealp[p.\ 75]{Russell:1921ww}).
+%`four ways of understanding words:
+%[...]
+%(4) When the word is being first learnt, you may associate it with an object, which is what it "means," or a representative of various objects that it "means."
+%In the fourth case, the word acquires, through association, some of the same causal efficacy as the object.
+The word "motor" can make you leap aside, just as the motor can, but it cannot break your bones.
+The effects which a word can share with its object are those which proceed according to laws other than the general laws of physics, i.e.
+those which, according to our terminology, involve vital movements as opposed to merely mechanical movements.'
+%\citep[p.\ 75]{Russell:1921ww} 
 After all, chimpanzees associate pointing gestures with their referents \citep[][p.\ 6]{Moll:2007gu} but fail to respond appropriately to helpful pointing gestures \citep{hare_chimpanzees_2004}. 
-Unlike chimpanzees, infants not only map words to what they pick out but also use and understand them as tools for communication. 
-What does this understanding amount to---in particular, do infants have an adult-like understanding of communicative intention---and how do infants acquire it?
+Unlike chimpanzees, children not only map words to what they pick out but also use and understand them as tools for communication. 
+What does this understanding amount to---in particular, do children learning their first words already have an adult-like understanding of communicative intention?
+And how do children acquire this understanding?
 %On one view, it involves a special `action interpretation system' distinct from that involved in understanding ordinary, non-communicative action \citep[p.\ 456]{Csibra:2003kp}. 
 %***This question is pressing because we want to appeal to language to explain the emergence of full-blown mindreading
-Answering this question may require revisions or extensions to existing philosophical accounts of communication which often presuppose conceptual sophistication \citep[e.g.][]{Grice:1957kg}.
+Answering these questions may require revisions or extensions to existing philosophical accounts of communication which often presuppose conceptual sophistication \citep[e.g.][]{Grice:1957kg}.
 
 \subsection{Chapter 6. Actions: Teleology and Motor Awareness}
-\label{ch:actions}
+%Astington's paradox:
+%\begin{quote}
+%`There is something paradoxical about intention attribution.
+%Though it is true that motivational states are more obvious and more frequently inferred than beliefs, the attribution of intention---in a precise and complete sense---is not simpler, and may indeed be harder, than the attribution of belief'
+%\citep[p.\ 85]{astington:2001_paradox}
+
+Understanding goal-directed action involves being able to identify which events are actions and to link actions to the goals to which they are directed.
+Infants' expectations indicate that they can do both of these things from some time around the end of their first year of life \citep{Csibra:2003kp}. 
+Suppose we take these expectations to show that infants understand action.
+In what could this understanding consist?
+
+On a standard view, for an event to be a goal-directed action requires that it stands in some relation to an intention. 
+Intentions are propositional attitudes which play characteristic roles in planning and coordinating action, are linked to practical reasoning and are subject to characteristic norms \citep{Bratman:1987xw}.
+On this view, an action is directed to a goal in virtue of being an event related  an intention whose content specifies this goal. 
+Accordingly, understanding action involves understanding intentions and so having mastered propositional attitude psychology.
+This view of what understanding action consists in coheres with the views of those who hold that one-year-old infants understand psychological notions like belief and desire \citep{Baillargeon:gx}.
+But this view is incompatible with  views on which early understanding of action plays a role in explaining how children later acquire knowledge of belief, desire and other mental states.
+
+An alternative view of action and action understanding does not demand knowledge of intentions.
+Instead of thinking of the relation between actions and the goals to which they are directed as determined by intentions representing those goals,
+we might instead attempt to explain the relation in terms of justification:
+%
+\begin{quote}
+`the principle of rational action ...
+states that an action can be explained by a goal state if, and only if, it is seen as the most justifiable action towards that goal state that is available within the constraints of reality'
+(Csibra and Gergely 1998: 255; cf. Csibra, Bíró, et al. 2003)
+\end{quote}
+%
+***Don't we have to separate out the question of which events are actions?
+***Want to argue that the view is still demanding---and incompatible with modularity---because requires reasoning about justification.  
+Indeed:
+%
+\begin{quote}
+`when taking the teleological stance one-year-olds apply the same inferential principle of rational action that drives everyday mentalistic reasoning about intentional actions in adults' \citep[p.\ 259]{Csibra:1998cx}\citep[compare][p.\ 290]{Gergely:2003gb}
+\end{quote}
+%
 ***
+
+%Some philosophers oppose the standard view of goal-directed action.
+%Instead of explaining the directedness of actions to goals by appeal to intentions which represent those goals, 
+%it is at least theoretically coherent to suppose that an action's being directed to an outcome may consist in its having the function of bringing about that outcome. 
+%Here function should be construed teleologically. 
+%One perhaps inadequate but illustrative construal says that an action has the function of bringing about an outcome just if (i) actions of this type have brought about outcomes of this type in the past, and (ii) this action occurs now in part because of that fact.
+%Teleological accounts of function, and of the application of this notion to understanding goal-directed action, have been extensively developed \citep[see further][]{Godfrey-Smith:1996ln,Millikan:1984ib,Neander:1991to,Price:2001hs,Wright:1976ls}.
+
+There is a radical alternative ... understanding consist in motor awareness of action ...
+\begin{quote}
+\citep[p.\ 154]{bekkering:2000_imitation}: `behaviours are not simply replicated as unified, non-decomposed motor patterns.
+Rather, imitation involves first a decomposition of the motor patterns into their constituent components and later a reconstruction of the action pattern from these components.
+Second, the decomposition--reconstruction process is guided by an interpretation of the motor pattern as a goal- directed behaviour.
+Thus, the constituent elements in the mediating representation involve goals rather than motor segments.
+Third, we assume that these goals are represented in a hierarchical pattern with some of the encoded goals being dominant over others.'
+\end{quote}
+
+***parallels between how visual representations and processes enable experiences of objects and how motor representations and processes enable experiences of actions.
+Of course, just as not every visual representation leaves its mark on experience, there are surely many motor representations that are only distantly if at all related to experience.
+But, equally, there are experiences which depend in part on motor representation in much the way that some visual experiences depend in part on visual representations.
 
@@
 There is a growing body of evidence for (\ref{infant_fb}).
-From around their first birthday infants  predict actions of agents with false beliefs about the locations of objects \citep[]{Onishi:2005hm, Southgate:2007js}. 
+From around their first birthday infants  predict actions of agents with false beliefs about the locations of objects \citep[]{Onishi:2005hm, Southgate:2007js}
 and choose different ways of interacting with others depending on whether their beliefs are true or false \citep[]{Buttelmann:2009gy,Knudsen:2011fk,southgate:2010fb}.  
@@
 some involve interaction whereas in others the subject is a mere observer \citep[e.g.][]{Chandler:1989qa},
-and some involve prediction actions whereas others involve predicting desires \citep{Astington:1991kk} or selecting an argument appropriate for someone with a false belief \citep{Bartsch:2000es}. 
+and some involve predicting actions whereas others involve predicting desires \citep{Astington:1991kk} or selecting an argument appropriate for someone with a false belief \citep{Bartsch:2000es}. 
 Despite all this variation and more, these false belief tasks all appear to measure a single developmental transition \citep{Wellman:2001lz},
@@
 We have seen objections to each of these responses.
-This indicates that we do not adequately understand what is involved in representing beliefs and, more generally, what it is to have a theory of mind.
+This indicates that we do not adequately understand what is involved in representing beliefs and, more generally, what it is to have %a theory of mind 
+knowledge of minds.
 

### 2012-07-21 developing mind/proposal developing mind.tex
working on proposal

@@
 
-Understanding goal-directed action involves being able to identify which events are actions and to link actions to the goals to which they are directed.
-Infants' expectations indicate that they can do both of these things from some time around the end of their first year of life \citep{Csibra:2003kp}. 
-Suppose we take these expectations to show that infants understand action.
-In what could this understanding consist?
+Tracking a goal-directed action involves at least two things: distinguishing the action from other events 
+and
+identifying to which goal or goals it is directed.
+Infants' expectations indicate that they can do both of these things from some time around the end of their first year of life \citep[e.g.][]{Csibra:2003kp}. 
+But tracking is not understanding.
+To say that infants can track goal-directed action is not yet to say anything about what they understand of action.
+Consider two questions about the nature of action.
+First, which events are actions?
+Second, what is the relation between an action and the goal or goals to which the action is directed?
+For each of these we can ask corresponding questions about infants' understanding of action.
+Which events do infants take to be actions?
+And how do they understand the relation between actions and the goals to which they are directed?
+We need to answer these questions in order to understand how early abilities to track action are developmentally related to knowledge of action, and also how they relate to developments in other domains including language.
+In this chapter we consider what infants understand of action.
 
-On a standard view, for an event to be a goal-directed action requires that it stands in some relation to an intention. 
-Intentions are propositional attitudes which play characteristic roles in planning and coordinating action, are linked to practical reasoning and are subject to characteristic norms \citep{Bratman:1987xw}.
-On this view, an action is directed to a goal in virtue of being an event related  an intention whose content specifies this goal. 
-Accordingly, understanding action involves understanding intentions and so having mastered propositional attitude psychology.
-This view of what understanding action consists in coheres with the views of those who hold that one-year-old infants understand psychological notions like belief and desire \citep{Baillargeon:gx}.
+The question of what actions are is usually answered by appeal to intention, a propositional attitude which plays a characteristic role in planning and coordinating action, is linked to practical reasoning and is subject to characteristic norms (\citealp{Bratman:1987xw}).
+On the standard view, an action is directed to an goal in virtue of the action's being appropriately related to an intention which represents this goal or some related goal.
+The corresponding view about understanding has it that 
+understanding action involves understanding intentions and so having mastered propositional attitude psychology.
+Could this be how infants understand actions?
+This view would be consistent with the views of those who hold that one-year-old infants understand psychological notions like belief and desire \citep{Baillargeon:gx}.
 But this view is incompatible with  views on which early understanding of action plays a role in explaining how children later acquire knowledge of belief, desire and other mental states.
+We should therefore consider alternative, less conceptually demanding accounts of action understanding.
 
-An alternative view of action and action understanding does not demand knowledge of intentions.
+Are there alternative views of action understanding which do not demand knowledge of intentions?
+As before, it is helpful to start with a view about what actions are and then consider corresponding views about understanding actions.
 Instead of thinking of the relation between actions and the goals to which they are directed as determined by intentions representing those goals,
-we might instead attempt to explain the relation in terms of justification:
+we might instead attempt to explain the relation in terms of justification.
+This is the idea behind Gergely and Csibra's `principle of rational action' which states that:
 %
 \begin{quote}
-`the principle of rational action ...
-states that an action can be explained by a goal state if, and only if, it is seen as the most justifiable action towards that goal state that is available within the constraints of reality'
+`%the principle of rational action ... states that 
+an action can be explained by a goal state if, and only if, it is seen as the most justifiable action towards that goal state that is available within the constraints of reality' 
 (Csibra and Gergely 1998: 255; cf. Csibra, Bíró, et al. 2003)
@@
 %
-***Don't we have to separate out the question of which events are actions?
-***Want to argue that the view is still demanding---and incompatible with modularity---because requires reasoning about justification.  
-Indeed:
+%*Don't we have to separate out the question of which events are actions?
+This view of action understanding requires less conceptual sophistication thanks to not requiring knowledge of intentions as such. 
+But remains quite demanding, for it entails that one-year-olds  are able to justify actions and engage in means end reasoning.
+Gergely and Csibra stress this aspect of their view:
 %
 \begin{quote}
-`when taking the teleological stance one-year-olds apply the same inferential principle of rational action that drives everyday mentalistic reasoning about intentional actions in adults' \citep[p.\ 259]{Csibra:1998cx}\citep[compare][p.\ 290]{Gergely:2003gb}
+`%when taking the teleological stance 
+one-year-olds apply the same inferential principle of rational action that drives everyday mentalistic reasoning about intentional actions in adults' (\citealp[p.\ 259]{Csibra:1998cx}; \citealp[compare][p.\ 290]{Gergely:2003gb})
 \end{quote}
 %
-***
+This is incompatible with the notion that infants' abilities to track action depend on core knowledge or modular representations.
+For, as we saw in Chapter 2 (on Objects),  processes involving core knowledge  are supposed to be informationally encapsulated;
+and informationally encapsulated processes cannot represent relations involving rationality or justification as such \citep{Fodor:2000cj}.
+To see whether it is even theoretically coherent to suppose that understanding action could involve core knowledge only, we must find alternative accounts of what it is to understand actions, accounts that involve neither intention nor rationality.
 
@@
 
-There is a radical alternative ... understanding consist in motor awareness of action ...
-\begin{quote}
-\citep[p.\ 154]{bekkering:2000_imitation}: `behaviours are not simply replicated as unified, non-decomposed motor patterns.
+So far we have ignored the possibility that motor representations and processes may play a role in understanding action.
+It is now well established that motor processes and representations are involved in observing actions---not only in producing them \citep{rizzolatti_functional_2010}.
+Furthermore, some motor representations represent outcomes to which actions are directed \citep{jeannerod_motor_2006}. 
+For instance, observing another grasping a cup can involve a motor representation of the outcome, of the cup's being grasped, and not only motor representations of merely kinematic or dynamic features.
+Apparently, then, motor planning can be used in reverse: it can be used for inferring a goal from an observed action and not only for  determining what to do in order to achieve a particular goal.
+We also know that, in adults and infants alike, motor representation can facilitate judgements about the goals of actions 
+(on adults: \citealp{serino:2009_lesions_,pazzaglia:2008_sound_}; on infants: \citealp{bekkering:2000_imitation}).
+It is possible, therefore, that infants can track actions not because they know about agent's intentions or consider the rationality of actions,
+but because their motor expertise enables  them to  determine to which goals others' actions could be directed. 
+(This is surely not the whole story, but it can be extended to involve statistical learning as well as motor representation; see \citet{paulus:2011_role}, \citet{Baldwin:2001rn}, and \citet{Saylor:2007pj}.)
+
+%\begin{quote}
+%\citep[p.\ 154]{bekkering:2000_imitation}: `behaviours are not simply replicated as unified, non-decomposed motor patterns.
 Rather, imitation involves first a decomposition of the motor patterns into their constituent components and later a reconstruction of the action pattern from these components.
@@
 Third, we assume that these goals are represented in a hierarchical pattern with some of the encoded goals being dominant over others.'
-\end{quote}
-
-***parallels between how visual representations and processes enable experiences of objects and how motor representations and processes enable experiences of actions.
+%\end{quote}
+%
+%*parallels between how visual representations and processes enable experiences of objects and how motor representations and processes enable experiences of actions.
 Of course, just as not every visual representation leaves its mark on experience, there are surely many motor representations that are only distantly if at all related to experience.
 But, equally, there are experiences which depend in part on motor representation in much the way that some visual experiences depend in part on visual representations.
+%
+%
 

### 2012-07-22 developing mind/proposal developing mind.tex
proposal finished draft check conclusion

@@
 
-\input{$HOME/Documents/submissions/preamble_steve_report}
+%\input{$HOME/Documents/submissions/preamble_steve_report}
+\input{$HOME/Documents/submissions/preamble_steve_paper2}
+
+%adjust title size & font
+\pretitle{
+	\begin{center}
+	%\sffamily
+	\LARGE
+}
+
 %\author{}
+
+%change fonts (not anonymous)
+%\setromanfont[Mapping=tex-text]{Linux Libertine O} 
+%\setsansfont[Mapping=tex-text]{Linux Biolinum O} 
+%\setmonofont[Scale=MatchLowercase]{Andale Mono}
 
@@
 \section{Competition}
-At the time of writing there is no book or collection devoted to philosophical issues in cognitive development other than \textit{The Philosophical Baby} \citep{gopnik:2009_philosophical} which is primarily aimed at general readers.
+At the time of writing there is no book or collection devoted to philosophical issues in cognitive development other than \textit{The Philosophical Baby} \citep{gopnik:2009_philosophical}  is primarily aimed at general readers.
 
@@
 To understand the emergence of knowledge we need to find ways of describing what is in between: individuals whose movements are neither mindless nor guided by intention and knowledge.  
+And the ways we find to describe these individuals 
+%that are not only theoretically coherent and empirically motivated---in addition,
+%we need ways of describing these individuals 
+must enable us to understand what it is like to be them, to
+see things from their points of view.
 Some progress has already been made but many challenges remain. 
@@
 %
-If this sort of view were right, abilities to communicate linguistically would not necessarily depend on knowledge proper and so it would be possible to appeal to linguistic communication in explaining the developmental origins of knowledge.
+If this sort of view were right, abilities to communicate linguistically would not necessarily depend on knowledge proper. 
+So it would be possible to conjecture that the emergence in development of knowledge proper is a consequence of 
+training with language.
 %Compare Wittgenstein again:
@@
 %ref to Wittgenstein on child learning about numbers somewhere?
-But putting these special cases aside, the experimental findings do not support Wittgenstein's view.
+But putting these special cases aside, the experimental findings do not support Wittgenstein's claim that language is entirely acquired through training.
 Instead they support the hypothesis that children solve the mapping problem by general reasoning and rely on social cognition
 \citep{Baldwin:1995xl,Bloom:2000qz,Sabbagh:2001sp}. 
-%Also talk about language creation here (see lexical innovation paper)
+This hypothesis is further supported by findings that children who have learnt few or no words sometimes invent their own words \citep{Clark:1993bv}, 
+and by
+research on profoundly deaf children brought up in purely oral environments and therefore without experience of language who, individually or in groups, create their own signed languages \citep{Kegl:1999es,Senghas:2001zm,Goldin-Meadow:2003pj}.
 
-This hypothesis conflicts with Davidson's claims that having thoughts involves grasping the concept of truth \citep[p.\ 189]{Davidson:2001sm}, 
+This hypothesis---children solve the mapping problem by general reasoning---conflicts with Davidson's claims that having thoughts involves grasping the concept of truth \citep[p.\ 189]{Davidson:2001sm},
 % `What gives my belief its content ... is my knowledge of what is required for the belief ... to be true' \citep[p.\ 189]{Davidson:2001sm}, 
@@
 We need to answer these questions in order to understand how early abilities to track action are developmentally related to knowledge of action, and also how they relate to developments in other domains including language.
-In this chapter we consider what infants understand of action.
+In this chapter we explore three hypotheses about the understanding which underpins infants' abilities to track action.
 
@@
 
-Are there alternative views of action understanding which do not demand knowledge of intentions?
+Are there alternative views of what understanding actions involves which do not demand knowledge of intentions?
 As before, it is helpful to start with a view about what actions are and then consider corresponding views about understanding actions.
@@
 %*Don't we have to separate out the question of which events are actions?
-This view of action understanding requires less conceptual sophistication thanks to not requiring knowledge of intentions as such. 
-But remains quite demanding, for it entails that one-year-olds  are able to justify actions and engage in means end reasoning.
+As part of an account of what actions are, this is unlikely to be correct.
+After all, agents sometimes fail to select actions that could be justified (or seen as justifiable) given the goal to which they are directed.
+\citet[p.\ 123]{Csibra:2003jv} themselves note this point.
+But the principle of rationality might serve as a useful approximation.
+Perhaps someone who applied the principle of rationality to work out which goals actions are directed to would be right much of the time. 
+So perhaps infants' abilities to track actions involve understanding action in accordance with the principle of rationality.
+This view of action understanding requires less conceptual sophistication thanks to not requiring knowledge of intentions or other propositional attitudes as such. 
+But it remains quite demanding, for it entails that one-year-olds  are able to justify actions and engage in means end reasoning.
 Gergely and Csibra stress this aspect of their view:
@@
 
-So far we have ignored the possibility that motor representations and processes may play a role in understanding action.
+So far we have ignored the possibility that motor representations and processes play a role in understanding action.
 It is now well established that motor processes and representations are involved in observing actions---not only in producing them \citep{rizzolatti_functional_2010}.
@@
 but because their motor expertise enables  them to  determine to which goals others' actions could be directed. 
-(This is surely not the whole story, but it can be extended to involve statistical learning as well as motor representation; see \citet{paulus:2011_role}, \citet{Baldwin:2001rn}, and \citet{Saylor:2007pj}.)
+(This is surely not the whole story, but it can be extended to involve statistical learning as well as motor representation; see \citet{paulus:2011_role}.)
+%don't cite here: these are to do with segmentation
+%, \citet{Baldwin:2001rn}, and \citet{Saylor:2007pj}.)
+On this hypothesis, infants' abilities to track action do not involve understanding actions in the way that human adults do (or are standardly held to). 
+For this reason the hypothesis might  appear  to make infants' abilities concerning action philosophically uninteresting.
+We can see that this appearance is mistaken by reflecting on how intentions (and other propositional attitudes) might refer to actions involving bodily movements.
+The considerations for the view that 
+thought about objects depends on perceptual experience of them  \citep[see][]{Campbell:2002ge} 
+also support the view that 
+all thought about actions ultimately depends on motor experience of them.
 
@@
 minds? 
-***
+Maybe.
+%In some special cases---number, colour, words and minds---there are theories which make testable predictions.
+%These theories conflict with some philosophers' views about development.
+%Most obviously, these theories provide objections to any form of nativism on which number, colour or psychological concepts are innate.
+%The findings that inform these theories also lead to objections to ***Wittgenstein ***Davidson
+We have seen that apparent inconsistencies in developmental findings concerning knowledge of objects (Chapter 2) and knowledge of minds (Chapter 7) can be resolved by distinguishing kinds of knowledge.
+One challenge was to make such distinctions in ways that are both  theoretically coherent and empirically motivated.
+A first step was to appeal to the notions of modularity and core knowledge; a further step was to consider how, in particular cases (including objects, numbers, colours and actions), perceptual and motor representations and processes can constitute core knowledge.
+
+A second, related challenge arises from hypotheses that social interaction enables the emergence in development of knowledge.  
+For this hypothesis to be coherent, there must be forms of social interaction which do not already presuppose the knowledge whose emergence is to be explained.
+Existing accounts of social interaction apply only to those who already have knowledge of several domains, including objects, actions and minds.
+The challenge, then, is to characterise these simple forms of social interaction.
+
+Distinguishing core knowledge from knowledge proper 
+and
+characterising simple forms of social interaction
+are 
+two key theoretical ingredients of
+developmental theories with testable predictions for some special cases, including numbers, colours, words and minds.
+These theories will probably turn out to be wrong, of course.
+What seems secure, though, is the idea that explaining development requires 
+%1
+distinguishing multiple kinds of knowledge, 
+%2
+identifying simple forms of social interaction 
+and 
+%3
+understanding how 
+	these jointly enable development.
+	%social interaction can transform knowledge.
 

### 2012-07-23 developing mind/proposal developing mind.tex
proposal about to send it

@@
 %The findings that inform these theories also lead to objections to ***Wittgenstein ***Davidson
-We have seen that apparent inconsistencies in developmental findings concerning knowledge of objects (Chapter 2) and knowledge of minds (Chapter 7) can be resolved by distinguishing kinds of knowledge.
-One challenge was to make such distinctions in ways that are both  theoretically coherent and empirically motivated.
-A first step was to appeal to the notions of modularity and core knowledge; a further step was to consider how, in particular cases (including objects, numbers, colours and actions), perceptual and motor representations and processes can constitute core knowledge.
+We have seen that apparent inconsistencies in developmental findings concerning knowledge of objects (Chapter 2) and minds (Chapter 7) can be resolved by distinguishing kinds of knowledge.
+One challenge is to make such distinctions in ways that are both  theoretically coherent and empirically motivated.
+In addressing this challenge we appealed to the notions of modularity and core knowledge and considered how, in particular cases (including objects, numbers, colours and actions), perceptual and motor representations and processes can realise core knowledge.
 
-A second, related challenge arises from hypotheses that social interaction enables the emergence in development of knowledge.  
+A second, related challenge arises from the hypothesis that social interaction enables the emergence in development of knowledge.  
 For this hypothesis to be coherent, there must be forms of social interaction which do not already presuppose the knowledge whose emergence is to be explained.
@@
 characterising simple forms of social interaction
-are 
-two key theoretical ingredients of
-developmental theories with testable predictions for some special cases, including numbers, colours, words and minds.
-These theories will probably turn out to be wrong, of course.
+yielded 
+two key ingredients of
+developmental theories for some special cases, including numbers, colours, words and minds.
+We know that these  theories advance our understanding of development because they lead to new objections to various philosophical views about nativism, language acquisition and the dependence of thought on language.
+Of course these theories will probably turn out to be wrong.
 What seems secure, though, is the idea that explaining development requires 

### 2012-07-31 intention motor representation goal ascription with corrado sinigaglia (paper 2)/Motor Representation in Goal Ascription.tex
second paper tex formatting

%!TEX TS-program = xelatex
%!TEX encoding = UTF-8 Unicode

%\def \papersize {a4paper}
\def \papersize {letterpaper}

\documentclass[12pt,\papersize]{extarticle}
% extarticle is like article but can handle 8pt, 9pt, 10pt, 11pt, 12pt, 14pt, 17pt, and 20pt text

\def \ititle {Intention and motor representation in purposive action}
\def \isubtitle {}
\def \iauthor {}
\def \iauthor {C. Sinigaglia* \& S. Butterfill**
\\ 
**Dipartimento di Filosofia, Università degli Studi di Milano, Italia
\\ 
***Department of Philosophy, University of Warwick, UK}
\def \iemail{corrado.sinigaglia@unimi.it}
%\date{}

\input{$HOME/Documents/submissions/preamble_steve_paper2}
%\author{}
%\author {S. Butterfill* \& C. Sinigaglia**
%\\ 
%**Department of Philosophy, University of Warwick, UK
%\\ 
%***Dipartimento di Filosofia, Università degli Studi di Milano, Italia}
\date{}

\begin{document}

\setlength\footnotesep{1em}

\bibliographystyle{newapa} %apalike

\maketitle
%\tableofcontents
\title{}

\begin{abstract}
\noindent
%
How could judgements about the goals of actions depend on motor representations? Many findings show that they do, but several obstacles ... Overcome obstacles by showing that motor representations support experiences of action in something like the ways in which visual representations support experiences of objects ... Implications for mindreading.
%
\end{abstract}

\tolerance=5000

\section{Introduction}

Motor representation sometimes occurs in action observation and enables judgements about the goals to which actions are directed (see \citealp{rizzolatti_functional_2010} for a review). For instance, suppose you observe someone's hand moving towards a pair of scissors. You may judge that she will shortly perform an action directed to cutting something up. Making this judgement may depend on representing the other's actions in motor terms \citep{boria:2009_intention, ortigue:2010_understanding}.  This is not to say, of course, that you could not make a judgement of this type without representing the action in motor terms, only that making this particular judgement may depend on motor representation. How is this possible? How could any judgements about the goals of actions ever depend on motor representations?

Two obstacles stand in the way of answering this question. The first is an apparent mismatch between what goal ascription requires and what motor representations can capture. Goal ascription requires capturing not only the outcome to which an action is directed but also the directedness of the action to that outcome. Where goal ascription involves representing intentions, it is clear enough how both can be captured. But it is not obvious how both can be captured with motor representation alone. To understand how goal ascription judgements could depend on motor representations we need to understand how the contents of the motor representations are related to the contents of the judgements.

The second obstacle is the apparent isolation of motor representation from thought. It is often assumed that motor processes and representations are entirely cut off from experience and so could not have any rational effect on judgement. But ascribing goals does not seem from the inside to be like gazing into a crystal ball and simply being struck with an idea. From a thinker's point of view, experience reveals not only movements of objects but also the goals of actions. To understand how goal ascription judgements could depend on motor representations we need to understand how the role of motor representations in enabling judgements is compatible with it appearing (perhaps veridically) to the thinker that she has reasons for those judgements.

In this paper we investigate ways of overcoming both obstacles. The first obstacle concerns how the contents of the motor representations are related to the contents of the judgements. On this we shall identify evidence for two claims. First, some motor representations represent not merely kinematic or dynamic features of actions but outcomes to which actions are directed. Second, although the directedness of an action to an outcome cannot be represented motorically, it can be captured by motor processes which determine how an action should unfold in order for the outcome to be realised. So the status of a particular motor representation as a representation of a goal to which an action is directed depends both on its content---on its being a representation of an outcome---and on its role in motor processes which determine how that action should unfold. This removes one obstacle to understanding how motor representations could enable judgements about the goals of actions.  But in removing this first obstacle we run straight into a second.  This second obstacle concerns how it could appear to the thinker that she has reasons for judgements when the judgements are enabled by motor representations. To remove this obstacle we shall argue that motor representations are related to experiences.  There is a parallel between visual and motor representations. Visual representations of the particular colours of objects are related to experiences of those objects and their colours, and these experiences arguably provide reasons for judgements about the colours of objects.%
\footnote{
*Some controversy.  Campbell: yes; Davidson: no (experience plays a merely causal role).  We simply assume yes.  (Or we can duck: we're talking about appearance to the thinker of having reasons, not the fact of having reasons.)}
So the fact that visual representations enable judgements about particular colours in no way conflicts with the claim that thinkers have reasons for those judgements.  We shall argue that motor representations of outcomes are related to experiences of actions and their directedness to outcomes.  This is why thinkers can have reasons for judgements about the directedness of an action to a goal even where the judgements are enabled by motor representations---and also why it appears (veridically) to thinkers that they have such reasons.  Motor representation stands to experience of action as visual representation stands to experience of particular colours. This anyway is the view we shall defend.

This view has implications for understanding how goal ascription relates to mindreading. A familiar view in philosophy has it that goal ascription depends on the ascription of intentions, which in turn is interdependent with ascriptions of belief and desire (*refs). On this view goal ascription is possible only as part of a larger mindreading project. By contrast we shall show that in some cases the only representations involved in goal ascription are motor representations. So goal ascription can occur independently of any knowledge of mental states. It is therefore coherent to conjecture that facts about the goals of actions form part of the evidential basis for mental state ascription. Knowledge of others' minds ultimately depends on knowledge of their actions, which in turn depends on experiences made possible by motor representations.  Mindreading is grounded in motor representation.

\section{How to capture the directedness of actions to outcomes}

Our project is to understand how judgements about the goals of actions  could depend on motor representations.  One obstacle to understanding this is an apparent mismatch between what goal ascription requires and what motor representations can capture. To illustrate, let us return to our opening example. Observing someone's hand moving towards a pair of scissors you judge that she will shortly perform an action directed to cutting something.  In making this judgement you are representing an outcome and linking it to an action. How could motor representation enable you to make such a judgement?

If all motor representations were about merely kinematic or dynamic features of action only, their role in facilitating judgements about outcomes would be mysterious. However some motor representations represent outcomes to which actions are directed, as we shall now argue.  For instance, observing someone's hand moving towards a pair of scissors can involve a motor representation of an outcome such as the cutting of something and not only motor representations of individual joint displacements that might bring about this outcome.  Motor representations can facilitate goal ascriptions in part because some motor representations represent the outcomes which are the goals to be ascribed.

But how can we distinguish a representation of an outcome from a representation of merely kinematic features of actions?  This is possible because a single outcome can be realised by actions with quite different kinematic features and, conversely, actions with arbitrarily similar kinematic features can realise different outcomes in different contexts.  To illustrate, consider this outcome: the grasping of a particular pen. This could be realised by any of three types of action which vary kinematically: a hand action, a foot action or a tool-using action.  If we had marks that we could use to identify motor representations without already knowing their contents, and if we had evidence that these marks were constant across all three types of action, then we could infer that some motor representations capture something more abstract than the kinematic features particular to the hand, foot or tool-using action. But we do have well established marks of motor representation, for motor representation can be inferred from patterns of neuronal discharge, from motor-evoked potentials, from where blood flows in the brain, from behavioural performance profiles and in other ways.  And we do have evidence that such marks are constant across all three actions which realise the grasping outcome \citep{rizzolatti:1988_functional, Rizzolatti:2001ug, cattaneo:2010_state-dependent}. So we can infer that some motor representations do not represent merely kinematic features of actions. To infer, further, that some motor representations represent outcomes we need to consider what happens where, conversely, we hold kinematic features constant while varying to which outcome an action is directed. Take an action which realises the grasping of a particular pen and compare it with a second action which is as similar as possible to the first with respect to its kinematic features but differs with respect to which outcome it is directed to because the pen is manifestly absent (so the action is not plausibly directed to grasping that pen) or manifestly too large or too small to grasp.  Some marks of motor representations distinguish these actions \citep{Umilta:2001zr, villiger:2010_activity, koch:2010_resonance}.  This together with the constancy of some motor representations across variations in kinematic features is good (if not conclusive) evidence that some motor representations represent outcomes to which actions are directed.%
\footnote{
\citet{pacherie:2008_action} and \citet{butterfill:2012_intention} also argue for this conclusion; the latter further supported this claim by appeal to the functions of motor processes in planning and monitoring action.
}

Actions are typically directed to many outcomes.  For instance, an action might be directed to grasping the scissors, to cutting a newspaper and to removing a particularly useful article to file later.  In reaching for the scissors the agent is already performing actions directed to all of these outcomes.  Which outcomes can be represented motorically?  Outcomes can be partially ordered by the means-end relation. It might be assumed that motor representations capture only least outcomes relative to the means-end ordering. It is certainly true that not all outcomes can be represented motorically.  For instance, the difference between cutting out an article and cutting out a picture from newspapers is not as such a difference that need be captured by any motor representations.  But we should not infer from this that motor representations are limited to representing least (in the means-end ordering) outcomes. In fact, differences such as that between reaching for a pair of scissors to cut something and reaching for a pair of scissors to place them somewhere (perhaps to put them away) are captured by motor representations (\citealp{Fogassi:2005nf, cattaneo:2007_impairment, bonini:2010_ventral, }). This suggests that motor representations are not limited to representing least outcomes.

Our overall question is how motor representation could enable goal ascription judgements. The fact that motor representations can represent outcomes is a step towards answering this question. However we are not yet in a position to answer the question. To see why, consider that someone might (at least in principle) observe an action and represent an outcome without yet relating the outcome to the action. Kate is observing Ludwig kick a ball. Kate has represented this outcome, the ball's being kicked, but she does not represent the outcome as in any way related to Ludwig's action. Rather, she simply finds herself somehow disposed to kick the ball. This is not because she intends to pass the ball back to Ludwig or to compete with him or because of any other intention: it is just because representing the outcome has suggested to her the possibility of performing this action. So Kate represents the outcome to which Ludwig's action is directed but she does not identify that outcome as the goal of his action. More would be required for goal ascription. To ascribe a goal to an action is to represent the action and the goal and also to relate the goal to the action.

But how might the directedness of actions to outcomes be represented or otherwise captured? In philosophy the notion of goal-directedness is standardly explained in terms of intention: for an action to be directed to a goal is for the action to be appropriately related to an intention whose content specifies that goal (or, on some views, an intention specifying an appropriately related goal). Given this, it is perhaps natural to suppose that ascribing a particular goal to a given action might involve representing that action, representing an intention whose content specifies the goal, and representing the action as appropriately related to the intention.  If so, motor representation could have at most a limited role in goal ascription.  After all, no motor representation represents an intention.

How can you get from motor representations to the directedness of actions to outcomes?

If motor representations were representations of kinematic or dynamic features only, the claim that they enable goal ascription would be mysterious.  In fact, however, motor representations represent outcomes: or so we shall argue.  This is a first step towards

How could motor representation represent the outcome to which an action is directed?

To understand how motor representation could enable goal ascription we need to understand the relation between motor representations and outcomes and the relation between

it is perhaps natural to suppose that ascribing a particular goal to a given action might involve representing the action, representing an intention whose content specifies the goal, and representing the action as appropriately related to the intention.

\section{* PLAN *}
[Corrado's Challenge: Why is this not just about implementation details?]
\begin{enumerate}
\item What roles, if any, might motor representation play in goal ascription?
\item Some motor representations occur in action observation and sometimes facilitate goal ascription (Serino; basketball; \& others?) (otherwise the question would be bizarre).  This is what makes the question pressing (timely).
\item The Question: How is this possible?  How could motor representation facilitate goal ascription?
\item Philosophical motivation: A familiar view in philosophy has it that goal ascription depends on the ascription of intentions, which in turn is interdependent with ascriptions of belief and desire (*refs). On this view goal ascription is possible only as part of a larger mindreading project. By contrast, developmental psychologists have assumed that goal ascription is possible even without knowing anything of an agent's mental states (*refs).  Our concern is with this pure form of goal ascription, goal ascription that occurs independently of knowledge of mental states.
\item There are two problems.  (a) adequacy of motor representation; (b) link between motor representation and judgement about the action and its goal
\item Problem (a).  Motor representations may represent outcomes but not the directedness of actions to outcomes?
\item Solution to problem (a): Capturing vs. representing the directedness of actions.
\item Problem (b): If we thought of motor representation as cut off from experience, it would seem impossible that differences in motor representation could make a difference to the reasons available from the observer's point of view.  
\item However, there is evidence that motor representation influences experience (e.g. acoustic experience of pitch direction).
\item One possibility: motor representation influences visual and acoustic experiences of movements and sounds, and these somehow facilitate goal ascription
\item This does not seem to be a fully adequate explanation because ?
\item Furthermore, motor representation enables experience which are (i) revelatory of action and (ii) distinct from visual or auditory experiences (Berti \&c).
\item So maybe: motor representation enables experiences of action as such
\item Conclusion: motor representation enables goal ascription by making possible experiences of action as such.  The role of motor representation in goal ascription is like the role of visual representation in judgements of motion.
\end{enumerate}

MR----> | acquisition| Action Judgment (action concept) <----> B \& D Judgment (b\&d concept)
tracking goal -----> || ascribing goal <----------------- > ascribing beliefs and desires

MR<----> Action Judgment (action concept) <----> B \& D Judgment (b\&d concept)

\bibliography{$HOME/endnote/phd_biblio}

\end{document}

### 2012-07-31 intention motor representation joint action (paper 3)/intention and motor representation in joint action.tex
various
second paper tex formatting

@@
 \noindent
-On the assumption that social motor representation plays a role in explaining how effective joint action is possible, is there also a role for motor representation in explaining what joint action is?
+On the assumption that motor representation plays a role in explaining how effective joint action is possible, is there also a role for motor representation in explaining what joint action is?  
 Philosophers tend to assume that motor representation is only an enabling condition for joint action and of no direct interest to narrowly philosophical theories of joint action and shared intention.
-In this talk I shall argue that social motor representation and shared intention  have distinctive roles in explaining the purposiveness of joint action.
+In this paper we identify a certain structure of motor representations---call it \textit{social motor representation}---which can enable joint action,
+and we argue that social motor representation and shared intention  have distinctive roles in explaining the purposiveness of joint action.
 This gives rise to a challenge.
@@
 You might first write down a plan of action without specifying who will act; the plan simply describes what is to be done.
-The plan  will eventually be implemented by you and your friends but the plan itself does not specify this and so it is agent-neutral.
+The plan  will eventually be implemented by you and your friends
+ but this is not written in plan itself  and so it is agent-neutral.
 Of course the fact that this plan is your collective plan may be represented elsewhere; this fact may also be implicit in  the plan's being stapled to the door of your communal kitchen.
@@
 
-Why accept the premise that social motor representation enables joint action?
+Why even consider the premise that social motor representation enables joint action?
 A direct way to test the premise would be to 
@@
 is this.
-If you are engaged in a joint action with someone, one which involves moving an object by passing it between you, then each of you has motor representations of actions whose (perhaps improper) components include actions which will be performed by the other  and these motor representations are functionally equivalent to motor representations of your own actions in the sense that they are just the sorts of representation that might have caused you to do what the other is doing (if you were in her position).
+If you are engaged in a joint action with someone, one which involves moving an object by passing it between you, then each of you has motor representations of actions whose (perhaps improper) components include actions which will be performed by the other  and these motor representations are functionally equivalent to motor representations of your own actions in the sense that they are just the sorts of representation that might have caused you to do what the other is doing if you were in her position.
 
@@
 The experiment by \citet{kourtis:2010_favoritism} was not designed to distinguish
-a case in which each agent represents the outcome to which the  whole joint action is directed (in this case, the entire movement of the object)
+cases in which each agent represents the outcome to which the  whole joint action is directed (in this case, the entire movement of the object)
 from
-a case in which 
+ cases in which 
 each agent represents separately outcomes to which each of the individual actions comprising the joint action (in this case, the several movements of the object each of which just one agent is responsible for).
-However we are making a leap and assuming (without argument) that in cases like that studied by Kourtis and colleagues, there is sometimes social motor representation.
+We are making a leap and assuming (without argument) that in cases like that studied by Kourtis and colleagues, there is sometimes social motor representation.
 
@@
 let us take a step back and consider an individual action.
-Suppose an agent moves a mug from one place to another, passing in from her left hand to her right hand half way.
+Suppose an agent moves a mug from one place to another, passing it from her left hand to her right hand half way.
 It is a familiar idea that motor planning, like planning generally, involves starting with relatively abstract representations of outcomes and gradually filling in details.
@@
 They also face a similar coordination problem, requiring a precisely timed swap from one hand to another.
-Now suppose (inspired by Koutis et al’s and others' findings) that the same planning is involved in the individual case (where one agent performs the whole action) and in the joint action case (where the action is distributed between two agents).
+Now suppose, 
+inspired by Koutis et al’s and others' findings,
+that the same planning is involved in the individual case (where one agent performs the whole action) and in the joint action case (where the action is distributed between two agents).
 The planning is the same almost up to the actual muscle contractions.
@@
 So far we have only been considering a possible role for social motor representation in enabling joint action.  
-But I think the details already give us grounds for holding that motor representation has a role to play in explaining what joint action is.
+Our primary concern, though, is with what joint action is.
+And of course not all enabling features are constitutive. 
+But the details of how 
+social motor representation enables joint action
+ already give us grounds for holding that motor representation has a role to play in explaining what joint action is.
 To see why,

### 2012-08-03 intention motor representation joint action (paper 3)/intention and motor representation in joint action.tex
working on motor joint paper

@@
 
-\bibliographystyle{newapa} %apalike
+\bibliographystyle{$HOME/Documents/submissions/mynewapa} %apalike
 
@@
 This needs unpacking.
-A representation (motor or not) is \emph{agent-neutral} if its content does not specify any agent or agents.
+A representation (motor or not) is \emph{agent-neutral} if its content does not specify any agent or agents.%
+\footnote{
+Our use of the term `agent-netural' to describe motor representations bears no relation to the use of the same term  to describe reasons \citep[on the latter, see][]{Parfit:1984fk}.
+}
 To illustrate, agent neutral representations are sometimes found at the early stages of planning.
@@
 Our primary concern, though, is with what joint action is.
-And of course not all enabling features are constitutive. 
-But the details of how 
+The details of how 
 social motor representation enables joint action
@@
 
-\section{Slide}
 A basic question about ordinary, individual action is:
 What is the relation between a purposive action and the outcome or outcomes to which it is directed?
+Purposive actions typically have many actual outcomes.
+Grabbing little Isabel by the hands I swing her around, causing her to laugh and, simultaneously, breaking a vase.
+Either or both of these might be outcomes to which my action is directed.
+Note also that some or all of the outcomes to which my action was directed might not be among its actual outcomes; after all, actions can fail.
+So among all the actual and possible outcomes of my action, one or some are singled out as specially related to this action.
+The question is what singles out the outcome or outcomes, actual or merely possible, to which a particular purposive is directed.
 
-Many ordinary purposive actions have many different outcomes.
-Grabbing little Isabel by the hands I swing her around, causing her to laugh and, simultaneously, breaking a vase.
-In fact the outcome to which this purposive action was directed might not be among its actual outcomes; after all, actions can fail.
+This question is closely related to a second.
+Ordinary purposive actions are sometimes composed of more than one motor action.  My swinging Isabel around includes my reaching for her wrists, grasping them and then spinning us around.  
+But  my action doesn’t include other things which I might be doing simultaneously, like refusing a cup of tea with my eyes or  trying to determine whether that smell is coming from Isabel’s sister Hannah’s nappy.
+The second question, then, is this: for a particular action directed to a specified goal, what determines which activities%
+\footnote{
+As we use the term `activities',
+all actions are activities 
+and so are  things like reaching and grasping.
+This should make it clear that our position does not depend on whether or not reachings and the like are  actions.
+} comprise the purposive action and which do not.
 
-\section{Slide}
-So among all the actual and possible outcomes of my action, one or some are singled out as specially related to this action.
-One aspect of the question concerns what singles out the outcome or outcomes, actual or merely possible, to which a particular purposive is directed.
-But there is also a second aspect ...
-
-\section{Slide}
-Ordinary purposive actions are sometimes composed of more than one motor action.
-My swinging Isabel around starts with my reaching for her wrists, grasping them and then spinning us around ...
-and my action doesn’t include other things which I might be doing simultaneously, like refusing a cup of tea with my eyes or  trying to determine whether that smell is coming from Isabel’s sister Hannah’s nappy.
-
-\section{Slide}
-So another aspect of our question is what determines which activities comprise the purposive action and which do not.
-
-\section{Slide}
-The standard answer to this question involves intention.
-An intention (1) represents an outcome, (2) coordinates the one or several activities which comprise the action; and (3) coordinate these activities in a way that would normally facilitate the outcome’s occurrence.
-
-What binds particular component actions together into larger purposive actions?  It is the fact that these actions are all parts of plans involving a single intention.
-What singles out an actual or possible outcome as one to which the component actions are collectively directed?  It is the fact that this outcome is represented by the intention.
-
+The standard answer to both questions involves intention.
+An intention represents an outcome, coordinates the one or several activities which comprise the action, and coordinates these activities in a way that would normally facilitate the outcome's occurrence.
+What binds component activities together into larger purposive actions?  
+It is the fact that these actions are all consequences of plans involving a single intention (and are all appropriately related to those plans).
+What singles out an actual or possible outcome as one to which the component activities are collectively directed?  
+It is the fact that this outcome is represented by the intention.
 So the intention is what binds component actions together into purposive actions and links the action taken as a whole to the outcomes to which they are directed.
 
-\section{Slide}
-Now as Elisabeth Pacherie has argued \citep[pp.\ 189-90]{pacherie:2008_action} (and I’ve had a go at arguing this in joint work with Corrado Sinigaglia recently too),
-motor representations are relevantly similar to intentions.
-
-Of course motor representations differ from intentions in some important ways (as Pacherie also notes).
-
+Motor representations are relevantly similar to intentions,
+as some have recently argued (\citealp[pp.\ 189-90]{pacherie:2008_action}; \citealp{butterfill:2012_intention}).
+Of course motor representations differ from intentions in some important ways (as these authors note).
 But they are similar in the respects that matter for explaining the purposiveness of action.
-(1) Like intentions, some motor representations represent outcomes (and not merely patterns of joint displacements, say).
-(2) Like intentions, some motor representations play a role in coordinating multiple  component activities by virtue of their role as elements in hierarchically structured plans.
-(3) And, like intentions, some motor representations coordinate these activities in a way that would normally facilitate the outcome’s occurrence.
-
-So in the individual case, it seems to me quite straightforward that there is a role for motor representation to play in explaining the purposiveness of action [*explaining the possibility of purposive action?].
-
-The claim is not that \emph{all} purposive actions are linked to outcomes by motor representations, just that some are.
-In some cases, the purposiveness of an action is grounded in a motor representation of an outcome; in other cases it is grounded in an intention.
+For, like intentions, some motor representations represent outcomes and not merely kinematic features of action.
+Like intentions, some motor representations play a role in coordinating multiple  component activities by virtue of their role as elements in hierarchically structured plans.
+And, like intentions, some motor representations coordinate these activities in a way that would normally facilitate the outcome’s occurrence.
+So anyone who accepts the standard story about purposive action and intention
+should also accept that a similar story about purposive action and motor representation.
+Given that the two basic questions about ordinary, individual purposive action can also be answered by appeal to intention,
+they can also be answering by appeal to motor representation.
+This motivates the following view.
+Not all purposive actions are bound together and linked to outcomes by intentions: in some cases what binds together purposive actions and links them to outcomes are motor representations.
+%In some cases, the purposiveness of an action is grounded in a motor representation of an outcome; in other cases it is grounded in an intention.
 And of course in many cases it may be that both intention and motor representation are involved.
 
-\section{Slide}
-Now let’s turn to joint action.
+This is one reason for supposing that motor representation is not merely an enabling condition for ordinary, individual action but also plays a role in explaining what action is (just as intention does).
+Now let’s return to joint action.
 
-\section{Slide}
-The same question we asked about ordinary, individual action also arises for joint action.
-What is the relation between a purposive joint action and the outcome or outcomes to which it is directed?
+The same two questions we asked about ordinary, individual action also arise for joint action.
+%What is the relation between a purposive joint action and the outcome or outcomes to which it is directed?
+What singles out the outcome or outcomes to which a purposive joint action is directed?
+And what binds together the various activities (of several agents) that make up the joint action?
+%A difference in the case of joint action is, of course, that the component activities are not activities of a single agent.
 
-\section{Slide}
-And again the question has two aspects.
-What singles out the outcome or outcomes to which a purposive joint action is directed?
-
-\section{Slide}
-And what binds together the various activities that make up the joint action?
-The difference in the case of joint action is, of course, that these activities are not necessarily activities of a single agent.
-
-\section{Slide}
-The answer to this question for the case of joint action is also superficially similar in the answer we gave in the case of ordinary, individual action.
-
+If we appeal to a notion of shared intention,
+we can answer these questions about joint action in a way that is superficially similar to way we answered the parallel questions about ordinary, individual action.
 A shared intention is what relates purposive joint actions to the outcomes to which they are directed.
 For the shared intention 
-(1) involves a representation, on the part of each agent, of an outcome
-(2) coordinates the several agents’ activities
+involves a representation, on the part of each agent, of an outcome,
+coordinates the several agents’ activities
 and 
-(3) coordinates the several agents’ activities in such a that would normally facilitate the occurrence of the represented outcome.
+coordinates the several agents’ activities in such a that would normally facilitate the occurrence of the represented outcome.
+This is how a shared intention can bind together the activities comprising a joint action and link them to an outcome.
+%Shared intention does for joint action roughly what ordinary intention does for ordinary, individual action.
 
-It is in this sense that a shared intention can ground the purposiveness of a joint action.
+% [individual]	intention : motor representation
+% 					::
+% [joint]			shared intention : social motor representation
 
-\section{Slide}
-But what we saw earlier, what the research by Kourtis et al and others indicates, is that social motor representation can play a  role similar to that of shared intention.
+Our earlier discussion of how social motor representation might enable joint action already shows that social motor representation resembles shared intention in this respect.
+Return to the example of two agents moving an object in a way that involves passing it between them.
+Suppose that their passing involves a social motor representation of the outcome,
+%reciprocal agent-neutral motor representations of the outcome, 
+which is the movement of the object. 
+Then there are motor representations, one for each agent, 
+of an outcome to which the joint action is directed.
+And these representations coordinate the several agents' activities,
+and 
+do so in ways that would normally facilitate the occurrence of the  outcome represented.%
+\footnote{
+This implies that social motor representation and the associated processes underwrite what \citet{Butterfill:2011_wija} calls \textit{collective goals}.
+}
+So social motor representation can bind together the activities comprising a joint action and link them to an outcome in much the way that shared intention can.
 
-Return to the example of two agents moving an object in a way that involves passing it between them.
-Suppose that their passing involves reciprocal agent-neutral motor representations of the outcome, which is the movement of the object. 
-These motor representations
-(1) represent an outcome to which the joint action is directed,
-(2) coordinate the several agents’ activities
+%[*use above?] Because each agent represents the whole movement and plans all of its implementation irrespective of which parts she will actually perform, each agent plans the action in a way that should coordinate with the other agent's plans providing they use similar planning procedures
+
+%[*What I’m saying here, in effect, is that both shared intention and social motor representation can yield a COLLECTIVE GOAL]
+
+What we are suggesting is very simple.
+Given the correctness of a standard view about shared intention in joint action, 
 and 
-(3) coordinate the several agents’ activities in a way that would normally facilitate the occurrence of the represented outcome.
-Because each agent represents the whole movement and plans all of its implementation irrespective of which parts she will actually perform, each agent plans the action in a way that should coordinate with the other agent’s plans providing they use similar planning procedures
+given that in ordinary, individual action, motor representations and bind together activities and link them to outcomes,
+it is plausible that 
+in joint action, several agents' activities can be bound together and linked to an outcome by social motor representation.
+That is,
+the purposiveness of a joint action can be grounded not only in shared intention, but also in social motor representation.
 
-[*What I’m saying here, in effect, is that both shared intention and social motor representation can yield a COLLECTIVE GOAL]
+\section{Social motor representation: like shared intention}
+It may be helpful to compare and contrast the notion of social motor representation with a notion of shared intention. 
+We shall use Bratman's account of shared intention as it is the best developed. 
+Here are Bratman’s collectively sufficient\footnotemark \ conditions for you and I to have a shared intention that we J:
+%
+\footnotetext{
+In \citet{Bratman:1992mi}, the following were offered as jointly sufficient \textit{and individually necessary} conditions; the retreat to sufficient conditions occurs in \citet[][pp.\ 143-4]{Bratman:1999fr} where he notes that `for all that I have said, shared intention might be multiply realizable.'
+} 
+%
+\begin{quote}
+\label{quote:bratman_account}
+`1. (a) I intend that we J and (b) you intend that we J
 
-\section{Slide}
-What I’m suggesting is very simple.
+`2. I intend that we J in accordance with and because of la, lb, and meshing subplans of la and lb; you intend that we J in accordance with and because of la, lb, and meshing subplans of la and lb
 
-If you think that in ordinary, individual action, the purposiveness of actions can be grounded by motor representations
-(and you should think this because it’s true),
-then you should also think the same about actions involving two or more agents---the purposiveness of a joint action can be grounded in motor representations as well as in shared intentions.
+`3. 1 and 2 are common knowledge between us' \citep[][p.\ View 4]{Bratman:1993je}
+\end{quote}
+%
+Let us take each of these three conditions in turn.
 
-\section{Slide}
-Let me try another way of presenting the same idea.
-Here are Michael Bratman’s sufficient conditions for shared intention.
-I want to suggest that social motor representation provides a parallel
+To see a parallel with the first condition, (1), recall two (empirical) claims on which the notion of social motor representation is based.
+First, some motor representations represent outcomes.
+Second, some motor representations represent the outcomes of actions not all of whose components will be executed by the agent whose motor representation it is.
+Given these claims, there is a direct parallel with Bratman's first condition, (1).
+In the case of social motor representation, each agent represents the outcome: but it is represented motorically rather than by having an intention.
+% (in our earlier example, this was the movement of the object).
 
-\section{Slide}
-There is a direct parallel with the first condition: in the case of motor representation, each agent represents the outcome (e.g. the movement of the object).
-The key claim here is that some motor representations (i) represent outcomes, and (ii) represent the outcomes of actions not all of whose components will be executed by the agent whose motor representation it is.
+Concerning the second condition, (2), there is clearly no {direct} parallel. 
+Whereas one intention can be about another intention, 
+we assume that one motor representation cannot be about another motor representation.
+But we do think there is a parallel of sorts. 
+A function of the second condition, (2), is to ensure meshing of subplans. 
+Each agent's having a motor representation of the outcome to which their joint action is directed does ensure meshing of subplans.
+What ensures this meshing is not the fact that each agent represents the other's plans {as the other's plans}.
+Rather in the case where joint action is grounded in social motor representation, what ensures meshing of subplans is this:
+each agent plans all of the agents' actions, and the agents rely on planning strategies that are sufficiently similar to ensure meshing subplans.
 
-\section{Slide}
-Here there is clearly no \emph{direct} parallel. 
-I don’t think motor representations can represent motor representations in the way that intentions can represent intentions.
-But I do think there is a parallel of sorts.
-Each agent’s having a motor representation of the distributed goal of their action does ensure meshing of subplans.
-What ensures this meshing is not the fact that each agent represents the other's plans \emph{as the other's plans}.
-Rather in the case where joint action is grounded in social motor representation, what ensures meshing of subplans is two facts (i) each agent plans all of the agents’ actions, and (ii) the agents rely on similar planning strategies (planning strategies that are sufficiently similar to ensure meshing subplans).
+The third condition, (3), concerns common knowledge.
+Why is this condition needed?
+Bratman himself says little.%
+\footnote{
+See \citet[p.\ 117]{Bratman:1993je}:
+`it seems reasonable to suppose that in shared intention the fact that each has the relevant attitudes is itself out in the open, is public.' 
+In other words, common knowledge is needed because it is.
+}
+One possible justification for supposing that shared intention involves common knowledge concerns a normative link between intention and reasons.
+In acting on an intention, there should be reasons for which the  agent acts.
+And, arguably, a consideration can only be among the reasons for which an agent acts if she knows that consideration (or at least is in a position to know it).
+So the need for common knowledge may arise from the need to explain how reasons for which an agent acts could include facts about others' intentions.
+This need does not arise in the case of social motor representation (at least not in the same way).
+For, arguably, where actions involve motor representations, it is not true that there should be reasons for which the agent acts.
+(Of course there are reasons which explain why motor actions happen; but these need not be reasons for which agents act.)
+So motor joint action does not require that one agent's motor representations provide reasons for which another agent acts.
+Instead, what is required is this.
+There should be a good chance---good relative to the potential costs and benefits of attempting this particular joint action now---that social motor representation will provide the necessary coordination.
+Of course this could be guaranteed by common knowledge. 
+But common knowledge is not required.
+Alternatively it can be ensured by common planning processes and a common {background} of dispositions, habits and expectations. 
+%common background is for reciprocity of motor representation and similar willingness to engage in joint action.
 
-\section{Slide}
-I’m less sure about a parallel to the common knowledge condition.
-As I see things, the justification for supposing that shared intention involves common knowledge concerns a normative link between intention and reasons.
-In acting on intentions, one should be acting for reasons.
-And a consideration can only be among your reasons if you know that consideration.
-So I think the need for common knowledge arises from the need to explain how another person’s intentions could be among your reasons for acting.
-I don’t think this need arises in the case of motor representation because it seems to me that the sort of planning of which motor representation is an element does not involve acting for reasons in the same sense.
-(In motor action, there are reasons why we do things (of course!) but these are not  reasons for which we act.)
-What motor joint action requires is not that your motor plans provide reasons for mine.
-There just has to be a good chance that this is true relative to the costs and benefits of joint action and the alternatives to joint action.
-So \textbf{I think that instead of common \emph{knowledge}, in the case of social motor representation there is a common \emph{background} of dispositions, habits and expectations.}%
-\footnote{
-[***CUT but one thing that might do the work of common knowledge is a custom or habit 
-that would allow the agents, in their particular social context, to rely on each other’s cooperation.
-In some countries this sort of thing works on public transport; 
-it is reasonable to take for granted that, if you are obviously struggling with a pram or suitcase, then someone nearby will help.]
-}
+If, as we have just argued, social motor representations play a role analogous to the structure of intentions and knowledge which Bratman identifies as sufficient for shared intention, then this is a  (non-decisive)  reason to think that motor representations can ground the purposiveness of joint action.
 
-\section{Slide}
-If this is right, if social motor representations play a role analogous to the structure of intentions and knowledge which Bratman identifies as sufficient for shared intention, then this is another reason to think that motor representations can ground the purposiveness of joint action.
+\section{Social motor representation: unlike shared intention}
+Our opening question was, Which events are joint actions?
+The standard view is that for an event to be a joint action it must be appropriately related to a shared intention.
+We are in the process of arguing that this is not the whole truth about joint action,
+and that some events are joint actions by virtue of being appropriately related to a structure of motor representations---social motor representations.
+We don’t mean to suggest that all joint actions involve social motor representation.
+The picture we are aiming to establish is rather this: some but not all joint actions involve shared intention; some but not all joint actions involve social motor representation; and there could be joint actions which involve social motor representation but not shared intention.
+This is why in our view 
+%Recap: the question was: Does social motor representation  play a role in explaining what joint is?
+%I have just been arguing for a positive answer.
+%My thesis is this:
+%\textbf{Reciprocal agent-neutral motor representations coordinate multiple agents’ actions around an outcome in part by virtue of representing that outcome.}
+%That is, reciprocal social motor representations can ground the purposiveness of joint action.
+%This is why I think that 
+fully understanding what joint action is requires understanding the coordinating role of social motor representation and not only understanding shared intention.
 
-\section{Slide}
-Recap: the question was: Does social motor representation  play a role in explaining what joint is?
-I have just been arguing for a positive answer.
-My thesis is this:
-\textbf{Reciprocal agent-neutral motor representations coordinate multiple agents’ actions around an outcome in part by virtue of representing that outcome.}
-That is, reciprocal social motor representations can ground the purposiveness of joint action.
-This is why I think that fully understanding what joint action is requires understanding the coordinating role of social motor representation and not only understanding shared intention.
-
-I don’t mean to suggest that all joint actions involve social motor representation.  Surely some joint actions do not.
+***HERE
+***Two lines of objection:
+(1) all joint actions involve share intention
+(2) social motor representation is just a species of shared intention (have just been filling in some details of how shared intentions are realised).
 

### 2012-08-05 intention motor representation joint action (paper 3)/intention and motor representation in joint action.tex
wokring

@@
 On this widely held view, for an event to be a joint action is for it to be appropriately related to a shared intention.
-For comparison, in the case of ordinary, individual action 
-one might hold that for an event to be an action is for it to be appropriately related to an intention.
-This assumption immediately raises the question of what shared intention is.
-On this question several philosophers offer competing theories.
-It is striking, however, that few have challenged the assumption that for an event to be a joint action it must be appropriately related to a shared intention.
-This is striking because the parallel assumption about ordinary, individual action is not as widely held among leading theorists.
-On several leading views, there could be agents whose actions are purposive---and, in some cases, even intentional---even though the agents have no intentions.
+This view is analogous to the view that, in the case of ordinary, individual action, 
+ for an event to be an action is for it to be appropriately related to an intention.
+
+Few have argued for, or objected to, 
+the view that for an event to be a joint action it must be appropriately related to a shared intention.
+This is striking because the parallel assumption about ordinary, individual action is not as widely held among leading theorists.%
+\footnote{
+For instance, \citet[p.\ ]{Velleman:2000fq}
+}
+In fact, some allow that there could be agents whose actions are purposive---and, in some cases, even intentional---although  the agents have no intentions at all.%
+\footnote{
+For instance, \citet[p.\ 251]{bratman:2000_valuing} % = \citep[p.\ 51]{Bratman:2007bd}
+describes an agent called `Creature 2' who `acts on the basis of its beliefs and considered desires' only (not intentions) as doing things `intentionally'.
+
+}
 If any such view is correct, it is impossible to fully explain what it is for an event to be an action by appeal to intention alone.
-Our aim is to defend a parallel claim about joint action.
-More specifically, we shall defend these claims: 
-first, 
-that fully explaining which events are joint actions requires appeal to structures of motor representation and not only to shared intention;
-and, 
-second, 
-that recognising a constitutive role for motor representation in joint action makes possible an account of shared intention which is simpler and more elegant than any of the leading alternatives.
+Our aim is to provide material for a parallel view about joint action.
+We shall offer considerations in favour of two claims: 
+First, fully explaining which events are joint actions requires appeal to structures of motor representation and not only to shared intention.
+And, second, recognising a constitutive role for motor representation in joint action makes possible an account of shared intention which is simpler and more elegant than any of the leading alternatives.
 
@@
 But  my action doesn’t include other things which I might be doing simultaneously, like refusing a cup of tea with my eyes or  trying to determine whether that smell is coming from Isabel’s sister Hannah’s nappy.
-The second question, then, is this: for a particular action directed to a specified goal, what determines which activities%
+The second question, then, is this: For a particular action directed to a specified goal, what determines which activities%
 \footnote{
@@
 This should make it clear that our position does not depend on whether or not reachings and the like are  actions.
-} comprise the purposive action and which do not.
+} comprise the purposive action and which do not?
 
@@
 This motivates the following view.
-Not all purposive actions are bound together and linked to outcomes by intentions: in some cases what binds together purposive actions and links them to outcomes are motor representations.
+Not all purposive actions are bound together and linked to outcomes by intentions. 
+In some cases what binds together purposive actions and links them to outcomes are motor representations.
 %In some cases, the purposiveness of an action is grounded in a motor representation of an outcome; in other cases it is grounded in an intention.
@@
 Given these claims, there is a direct parallel with Bratman's first condition, (1).
-In the case of social motor representation, each agent represents the outcome: but it is represented motorically rather than by having an intention.
-% (in our earlier example, this was the movement of the object).
+Where some agents have either a shared intention or a social motor representation, there is an outcome to which their actions are directed and each agent represents this outcome.
+Of course there is also a difference: In the case of social motor representation, the outcome is represented motorically and need not feature in the content of any intention.%
+\footnote{
+Here and below were are assuming that no motor representations are intentions. 
+If this assumption is wrong (as \citealp{pacherie:2008_action} suggests), social motor representation may be even more closely related to shared intention that we suggest here.
+}
 
@@
 we assume that one motor representation cannot be about another motor representation.
-But we do think there is a parallel of sorts. 
+But there is a parallel of sorts. 
 A function of the second condition, (2), is to ensure meshing of subplans. 
-Each agent's having a motor representation of the outcome to which their joint action is directed does ensure meshing of subplans.
+Each agent's having a motor representation of the outcome to which all their actions are together directed does ensure meshing of subplans.
 What ensures this meshing is not the fact that each agent represents the other's plans {as the other's plans}.
-Rather in the case where joint action is grounded in social motor representation, what ensures meshing of subplans is this:
-each agent plans all of the agents' actions, and the agents rely on planning strategies that are sufficiently similar to ensure meshing subplans.
+Rather what ensures meshing of subplans is this:
+Each agent plans all of the agents' actions, and the agents rely on planning strategies that are sufficiently similar to ensure meshing subplans.
 
@@
 But common knowledge is not required.
-Alternatively it can be ensured by common planning processes and a common {background} of dispositions, habits and expectations. 
+Alternatively it can be ensured by common planning processes and a common {background} of dispositions, habits and expectations.%
+\footnote{
+Another possible line of justification for a common knowledge component of shared intention might start from Davidson's claim that
+`[a]ction does require that what the agent does is intentional under some description, and this in turn requires ...\ that what the agent does is known to him under some description' \citep[p.\ 50]{Davidson:1971fz}.
+If we accepted this claim, then it seems we might have to conclude that not all joint actions are actions. 
+While this may appear to be an obstacle to accepting our view, there is an independent reason for thinking that not all joint actions are actions.
+For if we also follow Davidson in accepting that all actions are bodily movements (or Hornsby in accepting that all actions are tryings), it turns out that few paradigm cases of joint action are actions.
+For instance, we might paint a house together or make a hollandaise sauce together without there being any bodily movements of which we are both agents.
+} 
 %common background is for reciprocity of motor representation and similar willingness to engage in joint action.
 
-If, as we have just argued, social motor representations play a role analogous to the structure of intentions and knowledge which Bratman identifies as sufficient for shared intention, then this is a  (non-decisive)  reason to think that motor representations can ground the purposiveness of joint action.
+If, as we have just argued, social motor representations play a role analogous to the structure of intentions and knowledge which Bratman identifies as sufficient for shared intention, then this is a  (non-decisive)  reason to think that motor representations can ground the purposiveness of a joint action.
 
-\section{Social motor representation: unlike shared intention}
+\section{Joint action without shared intention}
 Our opening question was, Which events are joint actions?
@@
 We are in the process of arguing that this is not the whole truth about joint action,
-and that some events are joint actions by virtue of being appropriately related to a structure of motor representations---social motor representations.
+and that some events are joint actions by virtue of being appropriately related to a structure of motor representations we call social motor representation.
 We don’t mean to suggest that all joint actions involve social motor representation.
-The picture we are aiming to establish is rather this: some but not all joint actions involve shared intention; some but not all joint actions involve social motor representation; and there could be joint actions which involve social motor representation but not shared intention.
-This is why in our view 
+The view we are aiming to establish is rather this: Some but not all joint actions involve social motor representation; and there could be joint actions which involve social motor representation but not shared intention.
+This is why 
 %Recap: the question was: Does social motor representation  play a role in explaining what joint is?
@@
 
-***HERE
-***Two lines of objection:
-(1) all joint actions involve share intention
-(2) social motor representation is just a species of shared intention (have just been filling in some details of how shared intentions are realised).
+%*Two lines of objection:
+%(1) all joint actions involve share intention [here]
+%(2) social motor representation is just a species of shared intention (have just been filling in some details of how shared intentions are realised). [later]
+We are some way off having argued for this view.
+One consequence of our view is that there could be joint actions without shared intention. 
+But this does not straightforwardly follow from the claim we have argued for, namely that social motor representation is involved in some joint actions.
+And it also goes against a widely held view.
+To focus on just one case, Alonso says `the key property of joint action lies in its internal component [...] in the participants' having a ``collective'' or ``shared'' intention' \citep[pp.\ 444-5]{alonso_shared_2009}.%
+\footnote{
+***others
+}
+%Now for all we have said so far, 
+%someone might allow that, as we have argued, social motor representation is involved in some joint actions 
+%but also insist that all joint actions involve shared intention.
+If this were right, it would be false or at least unclear that social motor representation is needed to answer the question about which events are joint actions.
+We must therefore defend the view that joint action need not involve shared intention.
 
-But, equally, \emph{there could be purposive joint actions which do not involve shared intentions},
-just as there can be purposive actions which do not involve intentions at all.
-Let me go slowly in explaining why I think there could be joint action without shared intention and start by returning to the case of ordinary, individual action.
-
-\section{Slide}
-So what are intentions for?
-
-I’m going to assume that intentions are something over and above basic beliefs and desires; that an intention is not, for instance, merely a strongest desire or, as Donald Davidson held at an early stage of his thinking, merely a belief-desire pair.
-Intention involves more than this.
-
-There is a temptation to assume that intention is involved in every case of purposive action.
-But it’s hard to see what the argument for this assumption could be.
-In many cases it seems that beliefs, desires and motor representations are all that is needed to explain purposive action. 
-You offer me a biscuit.
-I want one, and I believe I can get one by reaching out for it.
-So I do reach for it.
-As far as I can see, there’s no need to suppose that, in addition to the belief and desire, it must be the case that I also intend to take a biscuit.
-(At least not unless we take ‘intention’ to mean ‘strongest desire’, which it does not.)  
-Maybe I do intend this.  
-But it’s possible for an agent to take and eat a biscuit, and to do so purposively, without having any intentions at all.  
-Beliefs, desires and motor representations are sufficient.
-
-So if we don't need intentions merely to perform a purposive action, what are intentions for?
-
-\section{Slide}
-This question becomes more pressing if you consider that motor representations enable quite sophisticated planning over short periods of time and sequences of action; for example, how you grasp a pointer will depend on what you are about to do with it \citep{zhang:2007_planning}.
-
-This sort of planning does not need intentions at all.  So (again) what are intentions for?
-
-\section{Slide}
-Michael Bratman suggests that 
-\emph{Intentions are for planning multiple separate actions over longer periods of time; and for planning multiple separate actions whose execution is mutually constraining where the outcomes cannot be represented motorically.}
-
-This is a case where intentions are really needed [see figure in slide] --- here one can’t act on strongest desire (for the big reward) if want to maximise rewards by collecting the small and the large reward.
-And one can’t rely on motor representation because the motor system doesn’t care about things that cannot be represented in motor terms.
-
-I don't think, of course, that intentions are only involved in actions which require planning of this sort.
-But I do think it's only in such actions that absolutely require intentions.
-
-Not all purposive actions involve any planning of this sort.
-
-Now you could imagine a two-person version of this task where we are rewarded for what we collectively achieve.
-In this case it’s optimal if one of us goes for the small reward and the other goes for the large reward.
-I think it’s this kind of planning that shared intention is really for.
-
-\section{Slide}
-By contrast, in many ordinary cases of joint action there is no need for planning of this sort and so no need for shared intention.
-Actions such as these \emph{might} involve shared intention but they do not \emph{necessarily} involve shared intention.
-
-I’m suggesting that some joint actions---like the one two people move an object in a way that involves passing it between them---don’t require this kind of planning and so don’t necessarily involve shared intentions.
-In some cases, social motor representation alone is sufficient for purposive joint action.
-
-\section{Slide}
-This goes against a widely shared view in the literature
-
-\section{Slide}
-To focus on just one case, Alonso says ‘the key property of joint action lies in its internal component [...] in the participants’ having a “collective” or “shared” intention.’ \citep[pp.\ 444-5]{alonso_shared_2009}
-
-I want to spend a bit of time on this claim because it matters for my second theme, which is a problem concerning how motor representation and shared intention interface in joint action.
-
-As far as I know, this claim is not explicitly argued for.  From conversation with philosophers I think the central argument hinges on  contrast cases.
-Contrast cases are pairs of actions which are similar in terms of behaviour and coordination but where one is joint but the other isn’t....
-
-\section{Slide}
-[nothing]
-
-\section{Slide}
-[Gilbert's Jack and Sue walking together case]
-
-\section{Slide}
-[Searle's park vistor contrast case]
-
+As far as we know, the claim that all joint action involves shared intention is not explicitly argued for.  
+The best argument we know of hinges on so-called contrast cases.
+Contrast cases are pairs of actions which are similar in terms of behaviour and coordination but where one is joint but the other isn't.  
+\citet{gilbert_walking_1990} contrasts two people walking together with two people individually walking side by side.  
+The two pairs' movements may be the same and similarly coordinated (to avoid collision), but walking together is a joint action whereas merely walking side by side is not. 
+Relatedly,  \citet{Searle:1990em}  contrasts a case in which several park visitors simultaneously run to a central shelter in order to perform a dance with another case in which the park visitors run to the central shelter in order to escape a storm.
+The first is a case of joint action, the second is not; but the same movements occur in both.
 These sorts of contrast case invite the question, 
-How do joint actions differ from individual actions which may occur in parallel? 
-What is the difference between Jack and Sue walking together and their walking side-by-side?  
-Gilbert’s example shows that the difference can’t just be a matter of coordination, because people merely walking alongside each other also need to coordinate their actions in order to avoid colliding with each other.
+How do joint actions differ from individual but parallel actions? 
+Gilbert’s example shows that the difference can’t just be a matter of coordination, because people who are merely walking alongside each other also need to coordinate their actions in order to avoid colliding.
 And Searle’s example shows that the difference between joint action and parallel individual action can’t just be that the actions have a common effect because merely parallel actions can have common effects too.
 
-How might this lead someone to think that all joint action involves shared intention?
-I think the idea is supposed to be this.
-One difference between the genuinely joint actions and their merely parallel counterparts is that the genuinely joint actions involve shared intentions.
-And there is no further difference that enables one systematically to distinguish the two cases.
+How could the contrast cases be used to argue that all joint action involves shared intention?
+The argument rests on this premise:
+It is possible to distinguish systematically between the contrast cases by appeal to shared intention and only by appeal to shared intention.
+To defeat this argument we shall show that there is another, more general way to distinguish contrast cases.
+
+Take Searle's example of several people running to the shelter.  
+In the joint case, there is a single outcome to which each person's actions is individually directed, namely their collective arrival at the shelter.  
+In the contrasting individual case, where park visitors individually run to the shelter to escape a storm, there is no single outcome to which each of their actions is individually directed.  
+Instead each visitor's actions are directed to that visitor's own arrival at the shelter.  
+Similarly, turning to Gilbert's example, when two people walk together, there is a single outcome to which both of their actions are individually directed (their collective arrival at a corner, say); whereas when two people merely walk side by side there is no single outcome to which each agent's actions are directed.
+In general, where two or more agent's actions constitute a joint action, there is an outcome to which each agent's actions are individually directed such that it is possible for them all to succeed relative to this outcome.%
+\footnote{
+Some readers may be sceptical of this claim.
+But note that it is a consequence of the view we are opposing, the view that all joint actions involve shared intention.
+Or, rather, it is a consequence of this view given a further premise which is a consequence of any standard account of shared intention: Where several agents act on a shared intention that they J, 
+each will perform an action directed to J and it is possible for them all to succeed in J-ing.
+} 
+Since this is not true in the contrasting cases of non-joint action, it allows us to distinguish systematically between joint actions and their non-joint but behaviourally and coordinatively indistinguishable counterparts.
+
+For concision let us stipulate that an outcome is a \emph{distributive goal} of two or more agents' actions just if two conditions are met.
+First, this outcome is a goal to which each agent's actions are individually directed.
+Second, each agent's actions are related to the goal in such a way that it is possible for all the agents (not just any agent, all of them together) to succeed relative to this goal.
+Let us also say that a structure is \emph{something like shared intention} if it stands to joint action roughly as intention stands to ordinary, individual action.  
+In these terms, our claim is that joint actions involve distributive goals and that this distinguishes joint actions from non-joint actions in the standard contrast cases.  
+Given that the existence of a distributive goal need involve nothing like a shared intention, we can conclude that systematically distinguishing between the contrast cases does not require anything like shared intention.
 
@@
 
+\section{***CUT}
+So what are intentions for?
+
+Let us to assume that intentions are something over and above basic beliefs and desires; that an intention is not, for instance, merely a strongest desire or, as Donald Davidson held at an early stage of his thinking, merely a belief-desire pair.
+Intention involves more than this.
+
+There is a temptation to assume that intention is involved in every case of purposive action.
+But it’s hard to see what the argument for this assumption could be.
+In many cases it seems that beliefs, desires and motor representations are all that is needed to explain purposive action. 
+You offer me a biscuit.
+I want one, and I believe I can get one by reaching out for it.
+So I do reach for it.
+As far as I can see, there’s no need to suppose that, in addition to the belief and desire, it must be the case that I also intend to take a biscuit.
+(At least not unless we take ‘intention’ to mean ‘strongest desire’, which it does not.)  
+Maybe I do intend this.  
+But it’s possible for an agent to take and eat a biscuit, and to do so purposively, without having any intentions at all.  
+Beliefs, desires and motor representations are sufficient.
+
+So if we don't need intentions merely to perform a purposive action, what are intentions for?
+
+This question becomes more pressing if you consider that motor representations enable quite sophisticated planning over short periods of time and sequences of action; for example, how you grasp a pointer will depend on what you are about to do with it \citep{zhang:2007_planning}.
+
+This sort of planning does not need intentions at all.  So (again) what are intentions for?
+
+ Bratman suggests that 
+\emph{Intentions are for planning multiple separate actions over longer periods of time; and for planning multiple separate actions whose execution is mutually constraining where the outcomes cannot be represented motorically.}
+
+This is a case where intentions are really needed [see figure in slide] --- here one can’t act on strongest desire (for the big reward) if want to maximise rewards by collecting the small and the large reward.
+And one can’t rely on motor representation because the motor system doesn’t care about things that cannot be represented in motor terms.
+
+I don't think, of course, that intentions are only involved in actions which require planning of this sort.
+But I do think it's only in such actions that absolutely require intentions.
+
+Not all purposive actions involve any planning of this sort.
+
+Now you could imagine a two-person version of this task where we are rewarded for what we collectively achieve.
+In this case it’s optimal if one of us goes for the small reward and the other goes for the large reward.
+I think it’s this kind of planning that shared intention is really for.
+
+By contrast, in many ordinary cases of joint action there is no need for planning of this sort and so no need for shared intention.
+Actions such as these \emph{might} involve shared intention but they do not \emph{necessarily} involve shared intention.
+
+We are suggesting that some joint actions---like the one two people move an object in a way that involves passing it between them---don’t require this kind of planning and so don’t necessarily involve shared intentions.
+In some cases, social motor representation alone is sufficient for purposive joint action.
+
 \bibliography{$HOME/endnote/phd_biblio}

### 2012-08-06 intention motor representation joint action (paper 3)/intention and motor representation in joint action.tex
just about to change the contrast cases section

@@
 By contrast, 
-most philosophers and many (but not all) psychologists appear to assume that this question can be fully answered just in terms of a special kind of intention or structure of intention and knowledge often called a shared (or `collective') intention.
+most philosophers and many (but not all) psychologists appear to assume that this question can be fully answered just in terms of a special kind of intention or structure of intention, knowledge and committment often called a shared (or `collective') intention.
 On this widely held view, for an event to be a joint action is for it to be appropriately related to a shared intention.
@@
 Few have argued for, or objected to, 
-the view that for an event to be a joint action it must be appropriately related to a shared intention.
-This is striking because the parallel assumption about ordinary, individual action is not as widely held among leading theorists.%
-\footnote{
-For instance, \citet[p.\ ]{Velleman:2000fq}
-}
+the view that for an event to be a joint action is for it to be appropriately related to a shared intention.
+This is striking because the parallel assumption about ordinary, individual action is not universally accepted. 
 In fact, some allow that there could be agents whose actions are purposive---and, in some cases, even intentional---although  the agents have no intentions at all.%
@@
 describes an agent called `Creature 2' who `acts on the basis of its beliefs and considered desires' only (not intentions) as doing things `intentionally'.
-
 }
+And others have argued that motor representation and intention are both needed in explaining which events are ordinary, individual actions \citep{butterfill:2012_intention}.
 If any such view is correct, it is impossible to fully explain what it is for an event to be an action by appeal to intention alone.
@@
 First, fully explaining which events are joint actions requires appeal to structures of motor representation and not only to shared intention.
+%***what's the argument for this second claim!?
 And, second, recognising a constitutive role for motor representation in joint action makes possible an account of shared intention which is simpler and more elegant than any of the leading alternatives.
+
+Note that the main claims of this paper are arguably consistent  with the claim that for an event to be an action is for it to be appropriately related to an intention.
+This is because it is possible to reject the claim that joint actions are actions. 
+Indeed, \citet[p.\ 376]{ludwig_collective_2007} has offered an independent reason to reject this claim.
+Suppose we accept, following \citet[pp.\ 49, 59]{Davidson:1971fz}, that all actions are bodily movements.
+Or suppose we follow \citet{hornsby_actions_1980} in holding that actions are tryings. 
+Then it turns out that few or no paradigm cases of joint action are actions.
+In making the sauce, one agent stirs while the other pours;
+in painting the house, one covers the outside while the other does the inside; 
+and in walking together each walker moves her own legs.
+In each of these supposedly paradigm cases of joint action, as in many others, no bodily movements which are actions (and no tryings) need have two or more agents.
+Apparently, then, joint actions are not actions. 
+This raises the possibility that the ingredients involved in explaining which events are actions differ from those involved in explaining which events are joint actions.
+Ludwig's observation on joint actions not being actions 
+also suggests, more generally, that caution is needed in generalising claims about ordinary, individual action to include joint action. 
+Although we shall draw parallels between individual and joint action, such parallels will serve merely exegetical ends.
+
+In what follows we aim to extend and generalise existing theories, not to replace them.
+We take for granted that some notion of shared intention is needed for explaining which events are joint actions. 
+Our aim is only to show that additional ingredients are needed by characterising one such ingredient.
 
@@
 social motor representation enables joint action
- already give us grounds for holding that motor representation has a role to play in explaining what joint action is.
+ already give us grounds for holding that motor representation has a role to play in explaining which events are joint actions.
 To see why,
@@
 and 
-given that in ordinary, individual action, motor representations and bind together activities and link them to outcomes,
+given that in ordinary, individual action, motor representations  bind together activities and link them to outcomes,
 it is plausible that 
@@
 \footnote{
-Another possible line of justification for a common knowledge component of shared intention might start from Davidson's claim that
-`[a]ction does require that what the agent does is intentional under some description, and this in turn requires ...\ that what the agent does is known to him under some description' \citep[p.\ 50]{Davidson:1971fz}.
-If we accepted this claim, then it seems we might have to conclude that not all joint actions are actions. 
-While this may appear to be an obstacle to accepting our view, there is an independent reason for thinking that not all joint actions are actions.
-For if we also follow Davidson in accepting that all actions are bodily movements (or Hornsby in accepting that all actions are tryings), it turns out that few paradigm cases of joint action are actions.
-For instance, we might paint a house together or make a hollandaise sauce together without there being any bodily movements of which we are both agents.
+Another possible line of justification the claim that common knowledge is involved in shared intention might start from a generalisation of Davidson's claim that
+`[a]ction does require %that what the agent does is intentional under some description, and this in turn requires 
+...\ that what the agent does is known to him under some description' \citep[p.\ 50]{Davidson:1971fz}.
+%If we accepted this claim, then it seems we might have to conclude that not all joint actions are actions. 
+%While this may appear to be an obstacle to accepting our view, there is an independent reason for thinking that not all joint actions are actions.
+%For if we also follow Davidson in accepting that all actions are bodily movements (or Hornsby in accepting that all actions are tryings), it turns out that few paradigm cases of joint action are actions.
+%For instance, we might paint a house together or make a hollandaise sauce together without there being any bodily movements of which we are both agents.
 } 
@@
 
-\section{Joint action without shared intention}
+\section{The Contrast Cases}
+
+\section{Social motor representation: unlike shared intention}
+
+*function (not to coordinate planning in Bratman's sense)
+
+*format (based on earlier paper ...)
+
+\section{***Joint action without shared intention}
 Our opening question was, Which events are joint actions?
@@
 We don’t mean to suggest that all joint actions involve social motor representation.
-The view we are aiming to establish is rather this: Some but not all joint actions involve social motor representation; and there could be joint actions which involve social motor representation but not shared intention.
+The view we are aiming to establish is rather this: Some joint actions involve social motor representation; and there could be joint actions which involve social motor representation but not shared intention.
 This is why 

### 2012-08-23 intention motor representation joint action (paper 3)/conf collective intentionality manchester 2012/motor representation and shared intention.tex
manchester talk

--- 2012-08-06 intention motor representation joint action (paper 3)/intention and motor representation in joint action.tex
@@
 
-\def \ititle {Intention and Motor Representation in Joint Action}
+\def \ititle {Motor Representation and Shared Intention}
 \def \isubtitle {}
-\def \iauthor {S. Butterfill* \& C. Sinigaglia**
-\\ 
-**Department of Philosophy, University of Warwick
-\\ 
-***Dipartimento di Filosofia, Università degli Studi di Milano}
+\def \iauthor {
+%S. Butterfill* \& C. Sinigaglia**
+%\\ 
+%**Department of Philosophy, University of Warwick
+%\\ 
+%***Dipartimento di Filosofia, Università degli Studi di Milano
+}
 \def \iemail{s.butterfill@warwick.ac.uk}
@@
 \noindent
-On the assumption that motor representation plays a role in explaining how effective joint action is possible, is there also a role for motor representation in explaining what joint action is?  
+On the assumption that motor representation plays a role in explaining how effective joint action is possible, do we also need motor representation to explain what joint action is?
 Philosophers tend to assume that motor representation is only an enabling condition for joint action and of no direct interest to narrowly philosophical theories of joint action and shared intention.
-In this paper we identify a certain structure of motor representations---call it \textit{social motor representation}---which can enable joint action,
-and we argue that social motor representation and shared intention  have distinctive roles in explaining the purposiveness of joint action.
+In this talk I shall argue that social motor representation and shared intention have distinctive roles in explaining the purposiveness of joint action.
 This gives rise to a challenge.
-On the one hand, effective joint action---imagine two people erecting a tent in a gale together---sometimes requires both shared intentions and social motor representations plus a certain kind of harmony between the two.
+On the one hand, effective joint action—imagine two people erecting a tent in a gale together—sometimes requires both shared intentions and social motor representations plus a certain kind of harmony between the two.
 On the other hand, recognizing their distinctive roles precludes the existence of direct inferential links between shared intentions and social motor representations.
@@
 Which events are joint actions?
-We shall argue that fully answering this question involves appeal to structures of motor representation.
-By contrast, 
-most philosophers and many (but not all) psychologists appear to assume that this question can be fully answered just in terms of a special kind of intention or structure of intention, knowledge and committment often called a shared (or `collective') intention.
+Most philosophers and many (but not all) psychologists appear to assume that this question can be fully answered just in terms of a special kind of intention or structure of intention, knowledge and commitment often called a shared (or `collective') intention.
 On this widely held view, for an event to be a joint action is for it to be appropriately related to a shared intention.
-This view is analogous to the view that, in the case of ordinary, individual action, 
- for an event to be an action is for it to be appropriately related to an intention.
 
-Few have argued for, or objected to, 
-the view that for an event to be a joint action is for it to be appropriately related to a shared intention.
-This is striking because the parallel assumption about ordinary, individual action is not universally accepted. 
-In fact, some allow that there could be agents whose actions are purposive---and, in some cases, even intentional---although  the agents have no intentions at all.%
+[\textit{Short version: One consideration in favour of this view is ...
+and jump to the second consideration towards the end of  section \ref{text:second_consideration} on page \vpageref{text:second_consideration}.}]
+
+\subsection{What is shared intention?}
+Of course this answer to the question about which events are joint actions assumes that we know what shared or collective intention is.
+But there are many accounts of it.
+Still, I think most people would agree that shared intention stands to joint action approximately as plain vanilla intention stands to ordinary, individual action
+Further,
+on all or most leading accounts of shared intention, each of the following is a necessary condition:
+
+\begin{description}
+
+\item[awareness of joint-ness] Agents acting on a shared intention know that they are not acting individually; they have `a conception of themselves as contributors to a collective end.'\footnote{
+	\citet[p.\ 10]{Kutz:2000si}.
+Compare \citet[p.\ 361]{Roth:2004ki}: `each participant ...
+can answer the question of what he is doing or will be doing by saying for example ``We are walking together'' or ``We will/intend to walk together.'''
+Relatedly, \citet[p. 56]{miller_social_2001} requires that each agent believes her actions are interdependent with the other agent's.
+}
+
+\item[awareness of others' agency]  When agents act on a shared intention, each is aware of at least one of the others as an intentional agent.\footnote{
+	Compare \citet[p.\ 333]{Bratman:1992mi}: `Cooperation ...
+is cooperation between intentional agents each of whom sees and treats the other as such'.
+See also \citet[p.\ 105]{Searle:1990em}: `The biologically primitive sense of the other person as a candidate for shared intentionality is a necessary condition of all collective behavior'
+}
+
+\item[awareness of others' states or commitments] When two agents share an intention that they F, each is aware of, or has individuating beliefs about, some of the other's intentions, beliefs or commitments concerning F.\footnote{
+This condition is necessary for shared intention even on what \citet[p.\ 40]{tuomela_collective_2000} calls `the weakest kind of collective intention'.
+But it may not be necessary if, as \citet{Gold:2007zd} suggest, shared intentions are constitutively intentions formed by a certain kind of reasoning.
+% "if the distinctive feature of collective intentions is to be found in the reasoning by which they were formed, then an analysis that focuses on the intentions themselves will miss the feature that makes collective intentions collective.
+"
+}
+
+\end{description}
+%
+I think we can go even further and say that, 
+on most accounts,
+where joint action involves shared intention, the agents act in part \emph{because} of their awareness of joint-ness, of others' agency and of others' states or commitments.
+
+What follows assumes that where one or more of these three conditions is not met, there is no shared intention.
+
+\subsection{Why shared intention?}
+Why might anyone hold that for an event to be a joint action is for it to be appropriately related to a shared intention?
+Two considerations favour this view.
+
+First, the view might be motivated by appeal to some well-known contrast cases.
+Contrast cases are pairs of events which are similar in terms of the behaviour and coordination they involve but where one is a joint action while the other is not.  
+Thus \citet{gilbert_walking_1990} contrasts two people walking together with two people individually walking side by side.  
+The two pairs' movements may be the same and similarly coordinated (to avoid collision), but walking together is a joint action whereas merely walking side by side is not. 
+Relatedly,  \citet{Searle:1990em}  contrasts a case in which several park visitors simultaneously run to a central shelter in order to perform a dance with another case in which the park visitors run to the central shelter in order to escape a storm.
+The first is a case of joint action, the second is not; but the same movements occur in both.
+These sorts of contrast case invite the question, 
+How do joint actions differ from individual but parallel actions? 
+Gilbert’s example shows that the difference can’t just be a matter of coordination, because people who are merely walking alongside each other also need to coordinate their actions in order to avoid colliding.
+And Searle’s example shows that the difference between joint action and parallel individual action can’t just be that the actions have a common effect because merely parallel actions can have common effects too.
+If someone thought that it is possible to distinguish systematically between the contrast cases by appeal to shared intention (and perhaps only by appeal to shared intention), this could motivate holding that for an event to be a joint action is for it to be appropriately related to a shared intention.
+
+{\label{text:second_consideration} A second consideration} in favour of this view is a possible parallel between shared intention and ordinary, individual action.
+If you think that events are actions in virtue of being appropriately related to intentions, 
+and if you think that shared intention stands to joint action as plain vanilla intention stands to ordinary, individual action,
+then it seems reasonable to suppose that for an event to be a joint action is for it to be appropriately related to a shared intention (*cite Pacherie Synthese forthcoming).
+
+\subsection{My view}
+But the parallel assumption about ordinary, individual action is not universally accepted. 
+In fact, some allow that there could be agents whose actions are purposive---and, in some cases, even intentional---although  the agents have no intentions at all.
+For instance, Michael Bratman describes a creature who `acts on the basis of its beliefs and considered desires' only (not intentions) as doing things `intentionally'.%
 \footnote{
@@
 }
-And others have argued that motor representation and intention are both needed in explaining which events are ordinary, individual actions \citep{butterfill:2012_intention}.
+Others have argued that motor representation and intention are both needed in explaining which events are ordinary, individual actions \citep{pacherie:2000_content,butterfill:2012_intention}.
 If any such view is correct, it is impossible to fully explain what it is for an event to be an action by appeal to intention alone.
-Our aim is to provide material for a parallel view about joint action.
-We shall offer considerations in favour of two claims: 
-First, fully explaining which events are joint actions requires appeal to structures of motor representation and not only to shared intention.
+
+Our aim in this talk is to defend a parallel view about joint action. 
+%(Parallel to the latter view, that is: not saying that creatures without any shared intentions are capable of joint actions.)
+We shall consider the possibility that fully explaining which events are joint actions requires appeal to structures of motor representation and not only to shared intention.
 %***what's the argument for this second claim!?
-And, second, recognising a constitutive role for motor representation in joint action makes possible an account of shared intention which is simpler and more elegant than any of the leading alternatives.
 
-Note that the main claims of this paper are arguably consistent  with the claim that for an event to be an action is for it to be appropriately related to an intention.
-This is because it is possible to reject the claim that joint actions are actions. 
-Indeed, \citet[p.\ 376]{ludwig_collective_2007} has offered an independent reason to reject this claim.
-Suppose we accept, following \citet[pp.\ 49, 59]{Davidson:1971fz}, that all actions are bodily movements.
-Or suppose we follow \citet{hornsby_actions_1980} in holding that actions are tryings. 
-Then it turns out that few or no paradigm cases of joint action are actions.
-In making the sauce, one agent stirs while the other pours;
-in painting the house, one covers the outside while the other does the inside; 
-and in walking together each walker moves her own legs.
-In each of these supposedly paradigm cases of joint action, as in many others, no bodily movements which are actions (and no tryings) need have two or more agents.
-Apparently, then, joint actions are not actions. 
-This raises the possibility that the ingredients involved in explaining which events are actions differ from those involved in explaining which events are joint actions.
-Ludwig's observation on joint actions not being actions 
-also suggests, more generally, that caution is needed in generalising claims about ordinary, individual action to include joint action. 
-Although we shall draw parallels between individual and joint action, such parallels will serve merely exegetical ends.
-
-In what follows we aim to extend and generalise existing theories, not to replace them.
-We take for granted that some notion of shared intention is needed for explaining which events are joint actions. 
-Our aim is only to show that additional ingredients are needed by characterising one such ingredient.
+Please bear in mind that we are not trying to get rid of, or replace, the notion of shared intention.
+%I take for granted that a notion of shared intention is needed for explaining which events are joint actions. 
+%My concern is only with the possibility that additional ingredients are needed as well.
+In fact recognising that motor representation is needed to say which events are joint actions may make it easier to understand what shared intention is.
 
 \section{Social motor representation}
-We start from the premise that reciprocal agent-neutral motor representation enables some joint actions.
+We start from the premise that some joint actions are facilitated by reciprocal, agent-neutral motor representations of outcomes whose obtaining would normally involve action on the part of each agent.
+
 This needs unpacking. 
+A \textit{motor} representation is the sort of representation that enables us to reach for, grasp and transfer objects in a coordinated and fluid way.
+We follow several psychologists and philosophers in supposing that motor representations feature in planning and monitoring action \citep[e.g.]{wolpert:1995internal, miall:1996_forward}.
+One consequence is that motor representations are not concerned with merely kinematic or dynamic features of actions only.
+Rather, some motor representations represent outcomes, such as the movement of a target object from one place to another.
+
 A representation (motor or not) is \emph{agent-neutral} if its content does not specify any agent or agents.%
@@
 Of course the fact that this plan is your collective plan may be represented elsewhere; this fact may also be implicit in  the plan's being stapled to the door of your communal kitchen.
-The agent-neutrality of a representation does not require that the agents are nowhere specified, only that they are not specified in the content of the representation.\footnote{
+The agent-neutrality of a representation does not require that the agents are nowhere specified, only that they are not specified in the content of the representation.%
+\footnote{
 Strictly the following argument does not hinge on the agent-neutrality of representations.
-It is sufficient for our purposes that there are reciprocal motor representations which concern actions which are partly but not entirely the agent's own.
+It is sufficient for our purposes that there are reciprocal motor representations  of outcomes whose obtaining would normally involve action on the part of each of the reciprocating agents.
 In principle such representations could have contents which specify other agents or multiple agents. 
 We focus on agent-neutral representations to simplify exposition.
+It may be important that our view is consistent with the possibility that reciprocal motor representations are agent-neutral. 
 }
+
 Two or more agents have \emph{reciprocal} motor representations  just if there is a single outcome and each agent has a motor representation of that outcome. 
-We shall use the term \emph{social} motor representation as an abbreviation for reciprocal agent-neutral motor representation.
-(Of course we cannot assume in advance of argument that reciprocal agent-neutral motor representations are social in any interesting sense;
+It is hardly controversial that reciprocal motor representations exist, for their existence is suggested by by a large body of research on motor cognition in action observation.
+What is much more controversial is that reciprocal motor representation occurs in joint action.
+Some of the research that Gunether Knoblich presented supports the claim that there are reciprocal motor representations in joint action, and he suggests that these enable joint action.%
+\footnote{
+See \citet{kourtis:2012_predictive}: `the partner’s expected action is simulated at the motor level, which probably facilitates effective performance of the joint action.'
+\citet{kourtis:2010_favoritism} show that 
+reciprocal motor representation is more likely to occur in joint action than is mere observation. 
+See also \citet{Knoblich:2003nf}.
+ }
+
+I want to go a tiny step further and suggest that in joint action there are sometimes reciprocal, agent-neutral representations of outcomes whose obtaining would normally involve action on the part of all of the agents.
+So when we act together, some of my motor representations may concern outcomes that are partly but not entirely to be realised by my actions.
+For example, suppose our task is to move an object from A to B, where you pick it up and pass it to me so that I can then place it.
+In this case I may represent the movement of the object from A to B and not only the component movements.
+I represent a collective outcome of our actions and not just outcomes to which each of our actions are individually directed.
+%
+%So it is not just that, when we act together, I represent the outcomes to which your actions are directed and the outcomes to which my actions are directed; I may also represent an outcome to which all of our actions are directed.
+
+We shall use the term \emph{social} motor representation as an abbreviation for the reciprocal, agent-neutral motor representation of outcomes whose obtaining would normally require action on the part of each reciprocating agent.
+(Of course we cannot assume in advance of argument that such reciprocal, agent-neutral motor representations are social in any interesting sense;
 but we are using the term `social' in a non-standard way as an abbreviation, one that reflects our aim.)
-So for two or more agents to have social motor representation is just for there to be a single outcome of which each agent has an agent-neutral motor representation.
 
-Our premise is just that social motor representation is among the factors enabling some joint actions.
-It does not follow that social motor representation is involved in every joint action,
-nor that the existence of social motor representation would be sufficient for there to be joint action.
+What follows is speculative philosophy: we take for granted that sometimes social motor representation facilitates joint action and ask whether this conjecture bears on our question about which events are joint actions.
 
-This premise is doubly controversial.
-While it is now well established both that there are motor representations of action outcomes (and not merely patterns of joint displacement, say), and also that there are motor representations of the outcomes of others' actions rather than only of one's own actions,
-less research has so far considered the conjecture that motor representations concern outcomes of actions which are partly but not entirely one's own.
-There is also relatively little evidence for the further conjecture that such representations enable some joint actions.
-We shall not argue for either conjecture; they are empirical premises.  
-What follows is speculative philosophy: an attempt to explain the consequences and significance of the conjecture that social motor representation enables some joint actions.
-
-Why even consider the premise that social motor representation enables joint action?
-A direct way to test the premise would be to 
-measure how (if at all) selective interventions on social motor representation affected agents’ performance of joint actions.
-As far as we know no one has done this and it may be impractical.
-But there are some indirect findings.
-We shall focus on just one, which will serve to illustrate the premise (but does not provide a decisive reason to accept it).
-Kourtis and colleagues report that `[s]imulation of another person’s action, as reflected in the activation of motor cortices, gets stronger the more the other is perceived as an interaction partner’  \citep[p.\ 4]{kourtis:2010_favoritism}.
-To establish this they had two subjects sit opposite each other at a table; there was also a confederate at the table, the `loner'.  
-Sometimes the subjects acted alone, picking up and replacing an object.  And sometimes they had to act together, passing an object between them.  
-The loner always acted alone.
-EEG measurements of motor activation were recorded.
-The researchers compared patterns of motor activation in four conditions: when the first subject acted, when the second subject acted, when the loner acted and when there was a `no go' signal so that no one acted.
-They found that when one of the subjects acted, patterns of activation in each of the two subjects were similar regardless of which was acting and which was observing.
-By contrast, when the loner acted the subjects' patterns of activation resembled those found in the `no go' condition when there was no action.
-Since the researchers did not include a condition where there was a competitive, non-joint interaction, we cannot conclude that the effect they found is specific to joint action as opposed to other kinds of interaction.
-What we do take this paper to indicate, 
-together with a body of other research on motor representation of observed action,
-is this.
-If you are engaged in a joint action with someone, one which involves moving an object by passing it between you, then each of you has motor representations of actions whose (perhaps improper) components include actions which will be performed by the other  and these motor representations are functionally equivalent to motor representations of your own actions in the sense that they are just the sorts of representation that might have caused you to do what the other is doing if you were in her position.
-
-For each of the agents involved in a joint action to represent motorically actions whose (perhaps improper) components include the others' action is not yet for there to be social motor representation.
-For there to be social motor representation there must be an outcome (such as the movement of an object from one place to another) which each agent has an agent-neutral motor representation of.
-The experiment by \citet{kourtis:2010_favoritism} was not designed to distinguish
-cases in which each agent represents the outcome to which the  whole joint action is directed (in this case, the entire movement of the object)
-from
- cases in which 
-each agent represents separately outcomes to which each of the individual actions comprising the joint action (in this case, the several movements of the object each of which just one agent is responsible for).
-We are making a leap and assuming (without argument) that in cases like that studied by Kourtis and colleagues, there is sometimes social motor representation.
-
-\section{How could social motor representation enable joint action?}
-Suppose that social motor representation is present in some joint action contexts, and, inspired by \citet{kourtis:2010_favoritism}, that 
-social motor representations are more likely to occur in joint action than when one is merely observing.
-It doesn't follow, of course, that social motor representation enables some joint actions.
+\section{How could social motor representation facilitate joint action?}
+Suppose that social motor representation is present in some joint action contexts.
+It doesn't follow, of course, that social motor representation facilitates some joint actions.
 But let us consider just the possibility that it might.
-How could social motor representation enable joint action even in principle?
+How could social motor representation facilitate joint action even in principle?
 
@@
 The planning is the same almost up to the actual muscle contractions.
+
 How could this be helpful?
@@
 (iii) each  agent's plans for their own action are constrained by their plans for the other agent's action.
+
 So what enables the two agents' plans to mesh is not that they represent each other's plans but more simply that they plan each other's actions as well as their own actions as if they were each about to do the whole thing themselves.
+
 Each agent is planning (and monitoring) both their actions almost as if a single agent were going to execute the whole action.
 And of course this is exactly what we want for small-scale joint action---we want two or more agents to act as one.
-This may be why the performance of dyads in joint actions often resembles the performance of individuals tasked with performing the whole action alone (*refs: Knoblich circles following task).
+This may be why the performance of dyads in joint actions often resembles the performance of individuals tasked with performing the whole action alone \citep{Knoblich:2003nf}.
 
@@
 \section{Grounding the purposiveness of joint action}
-So far we have only been considering a possible role for social motor representation in enabling joint action.  
+So far we have only been considering a possible role for social motor representation in facilitating joint action.  
 Our primary concern, though, is with what joint action is.
@@
 This should make it clear that our position does not depend on whether or not reachings and the like are  actions.
-} comprise the purposive action and which do not?
+} 
+comprise the purposive action and which do not?
 
@@
 
-\section{The Contrast Cases}
+%
+%\section{The Contrast Cases}
+%We have been arguing that social motor representation can bind multiple agents' activities together and link them to an outcome, and this in much the way that shared intention does  so.
+%This is a reason, 
+%not decisive but perhaps sufficient in the absence of strong contrary reasons, 
+%for supposing that social motor representation plays a role in explaining which events are joint actions.
+%Perhaps some events are joint actions in virtue of being appropriately related not to shared intention but to social motor representation.
+%
+%As mentioned at the start, our aim is to extend and generalise existing theories, not to replace them. 
+%But at this point our answer to the question of which events are joint actions seems to involve a disjunction.
+%This is both puzzling and inelegant. 
+%(Unless, of course, social motor representation can be regarded as a special case of shared intention.  
+%We explain why we reject this possibility below.)
+%In this section we explain how an account of joint action in terms of shared intention can be generalised to accommodate the possibility that social motor representation plays a role in explaining which events are joint actions.
+%
+%What guides and constrains theorising about shared intention? 
+%While there is little agreement on what shared intention is or what it is for, intuitions about it are sometimes grounded in contrast cases.
+%Contrast cases are pairs of events which are similar in terms of the behaviour and coordination they involve but where one is a joint action while the other is not.  
+%Thus \citet{gilbert_walking_1990} contrasts two people walking together with two people individually walking side by side.  
+%The two pairs' movements may be the same and similarly coordinated (to avoid collision), but walking together is a joint action whereas merely walking side by side is not. 
+%Relatedly,  \citet{Searle:1990em}  contrasts a case in which several park visitors simultaneously run to a central shelter in order to perform a dance with another case in which the park visitors run to the central shelter in order to escape a storm.
+The first is a case of joint action, the second is not; but the same movements occur in both.
+%These sorts of contrast case invite the question, 
+%How do joint actions differ from individual but parallel actions? 
+%Gilbert’s example shows that the difference can’t just be a matter of coordination, because people who are merely walking alongside each other also need to coordinate their actions in order to avoid colliding.
+%And Searle’s example shows that the difference between joint action and parallel individual action can’t just be that the actions have a common effect because merely parallel actions can have common effects too.
+%
+%How might the contrast cases be used to guide and constrain theorising about shared intention?
+%The idea that they serve this function seems to rest on a premise:
+%It is possible to distinguish systematically between the contrast cases by appeal to shared intention and only by appeal to shared intention.
+%We shall show that there is another, more general way to distinguish contrast cases.
+%
+%Take Searle's example of several people running to the shelter.  
+%In the joint case, there is a single outcome to which each person's actions is individually directed, namely their collective arrival at the shelter.  
+%In the contrasting individual case, where park visitors individually run to the shelter to escape a storm, there is no single outcome to which each of their actions is individually directed.  
+%Instead each visitor's actions are directed to that visitor's own arrival at the shelter.  
+%Similarly, turning to Gilbert's example, when two people walk together, there is a single outcome to which both of their actions are individually directed (their collective arrival at a corner, say); whereas when two people merely walk side by side there is no single outcome to which each agent's actions are directed.
+%In general, where two or more agent's actions constitute a joint action, there is an outcome to which each agent's actions are individually directed such that it is possible for them all to succeed relative to this outcome.%
+%\footnote{
+%Some readers may be sceptical of this claim.
+%But note that it is a consequence of the view we are opposing, the view that all joint actions involve shared intention.
+%Or, rather, it is a consequence of this view given a further premise which is a consequence of any standard account of shared intention: Where several agents act on a shared intention that they J, 
+%each will perform an action directed to J and it is possible for them all to succeed in J-ing.
+%} 
+%Since this is not true in the contrasting cases of non-joint action, it allows us to distinguish systematically between joint actions and their non-joint but behaviourally and coordinatively indistinguishable counterparts.
+%
+%For concision let us stipulate that an outcome is a \emph{distributive goal} of two or more agents' actions just if two conditions are met.
+%First, this outcome is a goal to which each agent's actions are individually directed.
+%Second, each agent's actions are related to the goal in such a way that it is possible for all the agents (not just any agent, all of them together) to succeed relative to this goal.
+%In these terms, our claim is that joint actions involve distributive goals and that this distinguishes joint actions from non-joint actions in the standard contrast cases.  
+%Since social motor representation and shared intention are each sufficient for the existence of a distributive goal, we can conclude that systematically distinguishing between standardly considered contrast cases does not require shared intention.
+% 
+%A natural response to this argument would be to suggest that the standard contrast cases are insufficient. 
+%For in some cases two agents' actions can have a distributive goal although arguably there is no joint action. 
+%Nora and Olive killed Fred.  
+%Each fired a shot.
+%Each intended that her shooting ground or partially ground Fred's death.%
+%\footnote{
+%Events $D_1$, ...\ $D_n$ \emph{ground} $E$, if: $D_1$, ...\ $D_n$ and $E$ occur; 
+%$D_1$, ...\ $D_n$ are each part of $E$; and 
+%every event that is 
+%	a part of $E$
+%	but does not overlap $D_1$, ...\ $D_n$ 
+%is caused by some or all of $D_1$, ...\ $D_n$.
+%(This is a refinement and generalisation of a notion due to \citet{pietroski_actions_1998}.)
+%Event $D$ \emph{partially grounds} event $E$ if $D$ alone does not ground $E$ but there are events including $D$ which do ground $E$.
+%} 
+%%
+%As it turned out, both intentions were fulfilled.
+%Neither shot was individually fatal but together they were deadly.
+%An ambulance arrived on the scene almost at once but Fred didn't make it to the hospital.
+%Now Nora and Olive's actions have a distributive goal.
+%After all, 
+%	each agent's actions are individually directed to Fred's death
+%	and  
+%	it is consistent with the stipulations made about this scenario that these goal relations are compatible in the sense that both agents could succeed together. 
+%But is their killing of Fred a joint action?
+%Imagine that Nora and Olive had no knowledge of each other, nor of each other's actions, and that their efforts were entirely uncoordinated.
+%We might even suppose that Nora and Olive are so antagonistic to each other that they would, if either knew the other's location, turn their guns on each other.
+%Their actions nevertheless have a distributive goal.
+%But given these further suppositions it is likely to seem counterintuitive to suppose that their killing Fred was a joint action. 
+%So although we can distinguish the standardly considered contrast cases just by appeal to the notion of a distributive goal, we should not assume that this notion captures the intended contrast between joint and merely parallel actions.
+%
+%***HERE refine: collective goals
+%
+%
+%
+%\section{Social motor representation: unlike shared intention}
+%
+%*function (not to coordinate planning in Bratman's sense)
+%
+%*format (based on earlier paper ...)
+%
 
-\section{Social motor representation: unlike shared intention}
-
-*function (not to coordinate planning in Bratman's sense)
-
-*format (based on earlier paper ...)
-
-\section{***Joint action without shared intention}
+\section{Conclusion}
 Our opening question was, Which events are joint actions?
@@
 
-%*Two lines of objection:
-%(1) all joint actions involve share intention [here]
-%(2) social motor representation is just a species of shared intention (have just been filling in some details of how shared intentions are realised). [later]
-We are some way off having argued for this view.
-One consequence of our view is that there could be joint actions without shared intention. 
-But this does not straightforwardly follow from the claim we have argued for, namely that social motor representation is involved in some joint actions.
-And it also goes against a widely held view.
-To focus on just one case, Alonso says `the key property of joint action lies in its internal component [...] in the participants' having a ``collective'' or ``shared'' intention' \citep[pp.\ 444-5]{alonso_shared_2009}.%
-\footnote{
-***others
-}
-%Now for all we have said so far, 
-%someone might allow that, as we have argued, social motor representation is involved in some joint actions 
-%but also insist that all joint actions involve shared intention.
-If this were right, it would be false or at least unclear that social motor representation is needed to answer the question about which events are joint actions.
-We must therefore defend the view that joint action need not involve shared intention.
-
-As far as we know, the claim that all joint action involves shared intention is not explicitly argued for.  
-The best argument we know of hinges on so-called contrast cases.
-Contrast cases are pairs of actions which are similar in terms of behaviour and coordination but where one is joint but the other isn't.  
-\citet{gilbert_walking_1990} contrasts two people walking together with two people individually walking side by side.  
-The two pairs' movements may be the same and similarly coordinated (to avoid collision), but walking together is a joint action whereas merely walking side by side is not. 
-Relatedly,  \citet{Searle:1990em}  contrasts a case in which several park visitors simultaneously run to a central shelter in order to perform a dance with another case in which the park visitors run to the central shelter in order to escape a storm.
-The first is a case of joint action, the second is not; but the same movements occur in both.
-These sorts of contrast case invite the question, 
-How do joint actions differ from individual but parallel actions? 
-Gilbert’s example shows that the difference can’t just be a matter of coordination, because people who are merely walking alongside each other also need to coordinate their actions in order to avoid colliding.
-And Searle’s example shows that the difference between joint action and parallel individual action can’t just be that the actions have a common effect because merely parallel actions can have common effects too.
-
-How could the contrast cases be used to argue that all joint action involves shared intention?
-The argument rests on this premise:
-It is possible to distinguish systematically between the contrast cases by appeal to shared intention and only by appeal to shared intention.
-To defeat this argument we shall show that there is another, more general way to distinguish contrast cases.
-
-Take Searle's example of several people running to the shelter.  
-In the joint case, there is a single outcome to which each person's actions is individually directed, namely their collective arrival at the shelter.  
-In the contrasting individual case, where park visitors individually run to the shelter to escape a storm, there is no single outcome to which each of their actions is individually directed.  
-Instead each visitor's actions are directed to that visitor's own arrival at the shelter.  
-Similarly, turning to Gilbert's example, when two people walk together, there is a single outcome to which both of their actions are individually directed (their collective arrival at a corner, say); whereas when two people merely walk side by side there is no single outcome to which each agent's actions are directed.
-In general, where two or more agent's actions constitute a joint action, there is an outcome to which each agent's actions are individually directed such that it is possible for them all to succeed relative to this outcome.%
-\footnote{
-Some readers may be sceptical of this claim.
-But note that it is a consequence of the view we are opposing, the view that all joint actions involve shared intention.
-Or, rather, it is a consequence of this view given a further premise which is a consequence of any standard account of shared intention: Where several agents act on a shared intention that they J, 
-each will perform an action directed to J and it is possible for them all to succeed in J-ing.
-} 
-Since this is not true in the contrasting cases of non-joint action, it allows us to distinguish systematically between joint actions and their non-joint but behaviourally and coordinatively indistinguishable counterparts.
-
-For concision let us stipulate that an outcome is a \emph{distributive goal} of two or more agents' actions just if two conditions are met.
-First, this outcome is a goal to which each agent's actions are individually directed.
-Second, each agent's actions are related to the goal in such a way that it is possible for all the agents (not just any agent, all of them together) to succeed relative to this goal.
-Let us also say that a structure is \emph{something like shared intention} if it stands to joint action roughly as intention stands to ordinary, individual action.  
-In these terms, our claim is that joint actions involve distributive goals and that this distinguishes joint actions from non-joint actions in the standard contrast cases.  
-Given that the existence of a distributive goal need involve nothing like a shared intention, we can conclude that systematically distinguishing between the contrast cases does not require anything like shared intention.
-
-I think this argument is mistaken for two reasons.
-
-\section{Slide}
-First because there are contrast cases in which the joint action does not involve shared intention.
-For instance, contrast our lifting a sofa together with us each individually raising either end of the sofa, coincidentally at the same time.
-The former is a joint action whereas the latter is arguably not.
-But I don't think it's right to assume that lifting a sofa together necessarily involves shared intention (although our \emph{buying} a sofa together might).
-So it seems to me that the contrast cases give us no reason at all to suppose that all joint actions involve shared intention.
-
-In fact the existence of cases like the sofa indicate that appeal to shared intention does not provide a sufficiently general way of explaining the difference illustrated by the contrast cases.
-
-This brings me to a second consideration.
-There is a more general notion appeal to which enables us to distinguish the contrast cases.
-Genuine joint action differs from merely parallel action because the former involves each agent’s representing an outcome to which all of their actions are directed where these outcome representations coordinate their actions in a way that would normally facilitate their collective success in bring about this outcome.
-
-\section{Slide}
-(Actually I don’t think this is quite general enough because I think some joint actions involve non-representational coordinative structures only; I am also doubtful that there is a sharp distinction between merely parallel and genuinely joint action and I think it is possible to see the difference as a matter of degree.
-But I don’t want to get into that here.
-It’s sufficient that we have moved away from the bare shared intention account.
-[*WHAT WE REALLY NEED IS A COLLECTIVE OR SHARED GOAL, and the possibility of gradual construction shows that there’s no magic moment separating joint from parallel action.)
-
-To sum up so far, I reject the claim that reflection on the contrast cases provides any support for the idea that all joint action involves shared intention.
-
-\section{Slide}
-So far I have been assuming that there is an inconsistency between two claims:
-(1) my claim that reciprocal  agent-neutral motor representation can ground purposive joint action
-(2) many philosophers’ claim that all joint actions involve shared intention.
-
-But maybe it is a mistake to think that these are inconsistent.
-Why not suppose that some social motor representations are  shared intentions?
-
-\section{Slide}
-If some reciprocal agent-neutral motor representations are shared intentions, then there is no inconsistency.
-And what I’ve just been arguing is that agent-neutral motor representations resemble shared intentions in that both play a role in coordinating agents’ actions by virtue of representing outcomes.
-Isn’t that enough to justify identifying them as shared intentions?
-
-This issue might easily seem narrowly conceptual or terminological.  
-At the end of the day it doesn’t much matter if we want to call some motor representations ‘shared intentions’.  
-After all, on most accounts shared intentions are neither shared nor intentions so we would hardly be doing more violence to the term than is already being done.  
-So insofar as labeling some social motor representations shared intentions might help to avoid unnecessarily philosophical discussion, I’m all in favour.
-However, there is an important difference between the states normally regarded as  shared intentions and any motor representations.
-And this difference matters for understanding the interface between shared intention and motor representation.
-
-[*PLAN for what follows:*]
-
-Difference in format.
-Leads to the interface problem.  
-Shared intentions can be inferentially integrated with other shared intentions; but not they cannot be inferentially integrated with social motor intentions (two disjoint planning processes).  
-Elisabeth Pacherie’s proposal: shared intentions set outcomes to be achieved by social motor representations.  
-I borrow this idea from her, but it raises a further problem.
-The problem is how the one sets outcomes for the other given the difference in representational format.
-
-\section{Slide}
-In what respects do reciprocal agent-neutral motor representations differ from shared intentions?
-
-\section{Slide}
-As background we first need a generic distinction between content and format.
-Imagine you are in an unfamiliar city and are trying to get to the central station.
-A stranger offers you two routes.
-Each route could be represented by a distinct line on a paper map.
-The difference between the two lines is a difference in content.
-
-\section{Slide}
-Each of the routes could alternatively have been represented by a distinct series of instructions written on the same piece of paper; these cartographic and propositional representations differ in format.
-
-Format matters because only where two representations have the same format can they be straightforwardly inferentially integrated.
-
-\section{Slide}
-To illustrate, let’s stay with representations of routes.  
-Suppose you are given some verbal instructions describing a route.
-You are then shown a representation of a route on a map and asked whether this is the same route that was verbally described.
-You are not allowed to find out by following the routes or by imagining following them.
-Special cases aside, answering the question will involve a process of translation because two distinct representational formats are involved, propositional and cartographic.
-It is not be enough that you could follow either representation of the route.
-You will also need to be able to translate from at least one representational format into at least one other format.
-
-\section{Slide}
-This brings me to the argument ...
-
-\begin{enumerate}
-\item Only representations with a common format can be inferentially integrated.
-
-\item Any two intentions can be inferentially integrated in practical reasoning.
-
-\item My intention that I visit Paris on Friday is a propositional attitude.
-\end{enumerate}
-
-\section{Slide}
-It follows that 
-All intentions are propositional attitudes
-
-\section{Slide}
-But it is also widely agreed that:
-No motor representations are propositional attitudes.
-
-\section{Slide}
-So we can conclude that:
- No motor representations are intentions
-
-\section{Slide}
-This step---Any two intentions can be inferentially integrated in practical reasoning---is questionable.  I don’t have an argument for this and I’m not sure it isn’t terminological.  
-What I care about is that we distinguish attitudes according to the processes in which they feature.
-So if you like we could distinguish two kinds of intention, one propositional the other motor.
-As long as we distinguish representations of different formats I don’t see that it matters too much whether we call them all intentions or whether we use that term for only some of them.
-
-\section{Slide}
-So where does this leave us?
-The question was whether reciprocal agent-neutral motor intentions could count as shared intentions.
-
-\section{Slide}
-In answer to that question, I think this:
-(a)
-IF you agree motor representations are not intentions,
-THEN reciprocal agent-neutral motor representations are not shared intentions.
-
-And
-(b)
-IF motor representations are a non-propositional variety of intention,
-THEN reciprocal agent-neutral motor representations are a non-standard variety of shared intentions.
-
-The key thing is that, either way, reciprocal agent-neutral motor representations cannot be inferentially integrated with shared intentions in practical reasoning.
-This leads to what I’ll call ‘The Interface Problem’
-
-\section{Slide}
-It will take me a moment to explain what the problem is.
-
-\section{Slide}
-The first step is to note that:
-Some joint actions involve both shared  intention and reciprocal agent-neutral motor representation.
-Imagine two people setting up a tent in a gale together, for example.
-It seems success will often require many ingredients including shared intention for the large-scale planning and reciprocal agent-neutral motor representation for passing objects and bending the poles together.
-
-\section{Slide}
-Earlier I argued that reciprocal agent-neutral motor representations:
-
-\begin{itemize}
-\item represent outcomes;
-
-\item ground the purposiveness of some 
-joint actions; and
-\end{itemize}
-
-\section{Slide}
-This means, I think, that in at least some cases of effective joint action there will be a certain kind of harmony between the contents of 
- reciprocal agent-neutral motor representations
- and the contents of 
- shared intentions.
-Further, I think it's probable that, in some cases, the harmony is non-accidental.
-
-To be clearer about what `harmony' means, we need the notion of matching:
-\begin{quote}
-Two  outcomes, A and B, \emph{match} in a particular context just if, in that context, either the occurrence of A would normally constitute or cause, at least partially, the occurrence of B or vice versa.
-\end{quote}
-In some cases of joint action, the outcomes each kind of representation specifies non-accidentally match.
-
-\section{Slide}
-This leads to a question: how are non-accidental matches between the outcomes specified by shared intentions and by reciprocal agent-neutral motor representations possible?
-
-A natural suggestion is that matching is achieved through practical reasoning.
-It is in this way that the contents of shared intentions could partially determine the contents of reciprocal agent-neutral motor representations.
-But there is \textbf{an obstacle} to this idea.
-
-\section{Slide}
-For as we saw earlier, reciprocal agent-neutral motor representations
-\begin{itemize}
-\item differ in format from (the constituent attitudes of) shared intentions.
-\end{itemize}
-
-Given that only representations with a common format can be inferentially integrated in practical reasoning,
-we cannot appeal to reasoning to explain how the outcomes specified by shared intentions match the outcomes specified by reciprocal agent-neutral motor representations.
-This is why the interface problem is a problem.
-
-\section{Slide}
-In conclusion,
-I have suggested that 
-reciprocal agent-neutral motor representations
-can 
-ground purposive joint action
-for they
-can coordinate two or more agents' actions in virtue of representing a single outcome to which each agent's actions are directed.
-
-I also suggested, further, that
-reciprocal agent-neutral motor representation
-and
-shared intention  
-have distinctive roles in explaining the purposiveness of joint action.
-
-And this suggestion has two consequences.
-One is that we cannot properly understand what joint action is if we focus only on shared intention.
-The other is that we face a challenge.
-The challenge is to explain how these two kinds of representation could sometimes harmoniously contribute to effective joint action despite not being inferentially integrated in practical reasoning.
-
-\section{***CUT}
-So what are intentions for?
-
-Let us to assume that intentions are something over and above basic beliefs and desires; that an intention is not, for instance, merely a strongest desire or, as Donald Davidson held at an early stage of his thinking, merely a belief-desire pair.
-Intention involves more than this.
-
-There is a temptation to assume that intention is involved in every case of purposive action.
-But it’s hard to see what the argument for this assumption could be.
-In many cases it seems that beliefs, desires and motor representations are all that is needed to explain purposive action. 
-You offer me a biscuit.
-I want one, and I believe I can get one by reaching out for it.
-So I do reach for it.
-As far as I can see, there’s no need to suppose that, in addition to the belief and desire, it must be the case that I also intend to take a biscuit.
-(At least not unless we take ‘intention’ to mean ‘strongest desire’, which it does not.)  
-Maybe I do intend this.  
-But it’s possible for an agent to take and eat a biscuit, and to do so purposively, without having any intentions at all.  
-Beliefs, desires and motor representations are sufficient.
-
-So if we don't need intentions merely to perform a purposive action, what are intentions for?
-
-This question becomes more pressing if you consider that motor representations enable quite sophisticated planning over short periods of time and sequences of action; for example, how you grasp a pointer will depend on what you are about to do with it \citep{zhang:2007_planning}.
-
-This sort of planning does not need intentions at all.  So (again) what are intentions for?
-
- Bratman suggests that 
-\emph{Intentions are for planning multiple separate actions over longer periods of time; and for planning multiple separate actions whose execution is mutually constraining where the outcomes cannot be represented motorically.}
-
-This is a case where intentions are really needed [see figure in slide] --- here one can’t act on strongest desire (for the big reward) if want to maximise rewards by collecting the small and the large reward.
-And one can’t rely on motor representation because the motor system doesn’t care about things that cannot be represented in motor terms.
-
-I don't think, of course, that intentions are only involved in actions which require planning of this sort.
-But I do think it's only in such actions that absolutely require intentions.
-
-Not all purposive actions involve any planning of this sort.
-
-Now you could imagine a two-person version of this task where we are rewarded for what we collectively achieve.
-In this case it’s optimal if one of us goes for the small reward and the other goes for the large reward.
-I think it’s this kind of planning that shared intention is really for.
-
-By contrast, in many ordinary cases of joint action there is no need for planning of this sort and so no need for shared intention.
-Actions such as these \emph{might} involve shared intention but they do not \emph{necessarily} involve shared intention.
-
-We are suggesting that some joint actions---like the one two people move an object in a way that involves passing it between them---don’t require this kind of planning and so don’t necessarily involve shared intentions.
-In some cases, social motor representation alone is sufficient for purposive joint action.
-
 \bibliography{$HOME/endnote/phd_biblio}

### 2012-08-26 mindreading/conf espp 2012 symposium false belief counterfactuals perspective schurz rafetseder/false beliefs, perspectives and counterfactuals espp 2012 symposium submission.tex
espp talk done

@@
 
-Can we reject (\ref{infant_perspectives}), the claim that infants neither process perspective differences nor engage in counterfactual reasoning until they are at least one year old?
+Can we reject (\ref{infant_perspectives}), the claim that infants neither process perspective differences nor engage in counterfactual reasoning until they well over one year old?
 One difficulty is that current developmental evidence almost uniformly supports this claim  \citep{rafetseder:2010_counterfactual,beck:2011_almost}.
@@
 some involve interaction whereas in others the subject is a mere observer \citep[e.g.][]{Chandler:1989qa},
-and some involve prediction actions whereas others involve predicting desires \citep{Astington:1991kk} or selecting an argument appropriate for someone with a false belief \citep{Bartsch:2000es}.
+and some involve predicting actions whereas others involve predicting desires \citep{Astington:1991kk} or selecting an argument appropriate for someone with a false belief \citep{Bartsch:2000es}.
 Despite all this variation and more, the Category A tasks all appear to measure a single developmental transition \citep{Wellman:2001lz}.
@@
 
-\item[3$'$.] Infants cannot (i) process perspective differences nor (ii) engage in counterfactual reasoning until they are at least one year old.  \label{infant_perspectives}
+\item[3$'$.] Infants cannot (i) process perspective differences nor (ii) engage in counterfactual reasoning until they are well over one year old.  \label{infant_perspectives}
 \end{description}
@@
 %
-\item[3$''$.] Infants cannot (i) process perspective differences nor (ii) engage in counterfactual reasoning until they are at least one year old.  
+\item[3$''$.] Infants cannot (i) process perspective differences nor (ii) engage in counterfactual reasoning until they are well over one year old.  
 \end{description}
@@
 %
-\item[3$'''$.] Infants cannot (i) process perspective differences nor (ii) engage in counterfactual reasoning until they are at least one year old.  
+\item[3$'''$.] Infants cannot (i) process perspective differences nor (ii) engage in counterfactual reasoning until they are well over one year old.  
 \end{description}

### 2012-08-26 intention motor representation goal ascription with corrado sinigaglia (paper 2)/Motor Representation in Goal Ascription.tex
working on motor joint paper
updated paper, megered version 2 and 3

@@
 
-\def \ititle {Intention and motor representation in purposive action}
+\def \ititle {Motor representation in goal ascription}
 \def \isubtitle {}
@@
 %***Dipartimento di Filosofia, Università degli Studi di Milano, Italia}
-\date{}
+%\date{}
 
@@
 %
+
+\ % blank line
+
+\noindent
+Key words: ***
+
+\ % blank line
+
+\noindent
+Word count: 7 500
 \end{abstract}
@@
 
-Motor representation sometimes occurs in action observation and enables judgements about the goals to which actions are directed (see \citealp{rizzolatti_functional_2010} for a review).
-For instance, suppose you observe someone's hand moving towards a pair of scissors.
+Motor representation sometimes occurs in action observation and enables judgements about the goals to which actions are directed (see section \vref{sec:evidence} and \citealp{rizzolatti_functional_2010} for a review).
+Suppose you observe someone's hand moving towards a pair of scissors. 
 You may judge that she will shortly perform an action directed to cutting something up. 
@@
 This is not to say, of course, that you could not make a judgement of this type without representing the action in motor terms, only that making this particular judgement may depend on motor representation.
-How is this possible?
-How could any judgements about the goals of actions ever depend on motor representations?
+How could this work? 
+How could some judgements about the goals of actions ever depend on motor representations?
 
-Two obstacles stand in the way of answering this question.
-The first is an apparent mismatch between what goal ascription requires and what motor representations can capture.
-Goal ascription requires capturing not only the outcome to which an action is directed but also the directedness of the action to that outcome.
-Where goal ascription involves representing intentions, it is clear enough how both can be captured.
-But it is not obvious how both can be captured with motor representation alone.
-To understand how goal ascription judgements could depend on motor representations we need to understand how the contents of the motor representations are related to the contents of the judgements.
-
-The second obstacle is the apparent isolation of motor representation from thought.
-It is often assumed that motor processes and representations are entirely cut off from experience and so could not have any rational effect on judgement.
-But ascribing goals does not seem from the inside to be like gazing into a crystal ball and simply being struck with an idea.
-From a thinker's point of view, experience reveals not only movements of objects but also the goals of actions.
-To understand how goal ascription judgements could depend on motor representations we need to understand how the role of motor representations in enabling judgements is compatible with it appearing (perhaps veridically) to the thinker that she has reasons for those judgements.
-
-In this paper we investigate ways of overcoming both obstacles.
-The first obstacle concerns how the contents of the motor representations are related to the contents of the judgements.
-On this we shall identify evidence for two claims.
+Our aim in this paper is to answer this question. 
+The first step is to ask about the contents of motor representations.
+Where judgements about the goals of observed actions depend on the observers' motor representations, what do those motor representations represent?
+To understand how goal ascription judgements could depend on motor representations we need to understand how the contents of the motor representations compare with the contents of the judgements.
+In section \vref{sec:content} we aim to remedy this need by identifying evidence for two claims.
 First, some motor representations represent not merely kinematic or dynamic features of actions but outcomes to which actions are directed.
@@
 So the status of a particular motor representation as a representation of a goal to which an action is directed depends both on its content---on its being a representation of an outcome---and on its role in motor processes which determine how that action should unfold.
-This removes one obstacle to understanding how motor representations could enable judgements about the goals of actions.
-But in removing this first obstacle we run straight into a second.
-This second obstacle concerns how it could appear to the thinker that she has reasons for judgements when the judgements are enabled by motor representations.
-To remove this obstacle we shall argue that motor representations are related to experiences.
+We propose that where a judgement about the goal to which an action is directed depends on motor representations, there is a motor representation of an outcome which captures the directedness of the action to the goal.
+
+The second step in our attempt to understand how some judgements depend on motor representations concerns the apparent isolation of motor representation from thought.
+When judgements about the goals of observed actions depend on the observers' motor representations, which processes connect motor representations to judgements?
+We shall argue (in section \vref{sec:processes} that motor representations are related to experiences.
 There is a parallel between visual and motor representations.
@@
 Knowledge of others' minds ultimately depends on knowledge of their actions, which in turn depends on experiences made possible by motor representations.
-Mindreading is grounded in motor representation.
+Mindreading starts from motor experiences of actions.
 
-\section{How to capture the directedness of actions to outcomes}
+***NOTE: There is not enough about the importance of the question and its wider philosophical significance.
+
+\section{Motor representation enables judgements about the goals to which actions are directed}
+\label{sec:evidence}
+Before launching into the main task for this paper we shall briefly review evidence for our opening premise.
+Why think that motor representation enables judgements about the goals to which actions are directed?
+
+Some evidence for this claim arises from studies of deficits.
+Serino and colleagues (\citeyear{serino:2009_lesions_}) compared the performances of control and hemiplegic subjects who were asked to identify actions from point-light displays.
+They found that ability to perform an observed action was correlated with ability to recognise that action.
+In particular, hemiplegic subjects were less accurate in identifying actions performed with limbs on the hemiplegic side of their bodies than actions performed with limbs on the unaffected side.
+Other evidence for a role for motor representation in observation-based judgements of action comes from research on apraxia.
+In one study subjects were asked to identify actions such as cutting paper and drinking with a straw on the basis of the sounds these actions produced.
+Subjects with limb apraxia showed an impairment in recognising hand-related actions (such as cutting paper) whereas subjects with buccofacial apraxia were impaired in recognising mouth-related actions (such as drinking); but no subjects showed a general impairment in recognising sounds and their significance \citep{pazzaglia:2008_sound_}.
+This link between motor deficits and action recognition indicates that observation-based action judgements sometimes depend on motor representation.
+
+***mention other sources of evidence
+
+see `How your hands drive my eyes' (but note that our focus is effects \emph{on judgement}):
+\begin{quote}
+ `PMv is critically involved in outcome prediction (for a review see Schubotz, 2007).
+For instance, a fMRI study (Schubotz \& von Cramon, 2004) and lesion (Schubotz, Sakreida, Tittgemeyer, \& von Cramon, 2004) data showed that predictions of outcomes even more abstract than bodily movements recruit the premotor cortex"
+There is evidence that action observation and action execution selectively interact with each other.
+Brass et al.
+(Brass, Bekkering, \& Prinz, 2001) instructed participants to execute finger movements in response to observed compatible and incompatible finger movements, and found a reaction time advantage for compatible trials.
+Similarly, Kilner et al.
+(Kilner, Paulignan, \& Blakemore, 2003) showed an increased variance in arm movement execution when observing a human model executing a qualitatively different, rather than the same arm movement.
+This interference effect has been argued to be due to the simultaneous recruitment of the motor representation of both the observed and the executed movements
+There is evidence that action observation may impact on action execution.
+In particular, people have been shown to be faster in performing fingers movements while observing compatible rather than incompatible finger movements (Brass, et al., 2001).
+In the same vein, grasping movements are faster when primed by a picture of a compatible rather than incompatible grip postures (Craighero, Bello, Fadiga, \& Rizzolatti, 2002; Craighero, Fadiga, Rizzolatti, \& Umilta, 1999).
+
+`Our findings show that this effect may be related also to the motor representation of action outcome as measured by the gaze behaviour of the observer.
+Target-specific proactive eye movements have been demonstrated to be driven by the motor representation of an action outcome during both the execution and the observation of a hand action such as a grasping action (Ambrosini, Costantini, et al., 2011; Ambrosini, Sinigaglia, \& Costantini, 2011; Falck-Ytter, 2012; Falck-Ytter, Gredeback, \& von Hofsten, 2006; Flanagan \& Johansson, 2003; Rotman, et al., 2006).
+The present study indicates that the proactivity of the observer's eye movements significantly decreases when the motor representation of the observed action outcome turns out to be incompatible with the motor representation of the action outcome to be executed.'
+\end{quote}
+
+\section{Content}
+\label{sec:content}
 
@@
 
+[***TODO: write above in terms of Question One (rather than obstacle).]
+
+\subsection{Some motor representations represent outcomes}
 If all motor representations were about merely kinematic or dynamic features of action only, their role in facilitating judgements about outcomes would be mysterious.
@@
 This could be realised by any of three types of action which vary kinematically: a hand action, a foot action or a tool-using action.
-If we had marks that we could use to identify motor representations without already knowing their contents, and if we had evidence that these marks were constant across all three types of action, then we could infer that some motor representations capture something more abstract than the kinematic features particular to the hand, foot or tool-using action.
+If we had marks we could use to identify motor representations without already knowing their contents, and if we had evidence that these marks were constant across all three types of action, then we could infer that some motor representations capture something more abstract than the kinematic features particular to the hand, foot or tool-using action.
 But we do have well established marks of motor representation, for motor representation can be inferred from patterns of neuronal discharge, from motor-evoked potentials, from where blood flows in the brain, from behavioural performance profiles and in other ways.
@@
 To infer, further, that some motor representations represent outcomes we need to consider what happens where, conversely, we hold kinematic features constant while varying to which outcome an action is directed.
-Take an action which realises the grasping of a particular pen and compare it with a second action which is as similar as possible to the first with respect to its kinematic features but differs with respect to which outcome it is directed to because the pen is manifestly absent (so the action is not plausibly directed to grasping that pen) or manifestly too large or too small to grasp.
+Take an action which realises the grasping of a particular pen and compare it with a second action which is as similar as possible to the first with respect to its kinematic features but differs with respect to which outcome it is directed to because the pen is manifestly absent (so the action is not plausibly directed to grasping that pen) or because the pen is manifestly too large or too small to grasp.
 Some marks of motor representations distinguish these actions \citep{Umilta:2001zr, villiger:2010_activity, koch:2010_resonance}.
@@
 \footnote{
-\citet{pacherie:2008_action} and \citet{butterfill:2012_intention} also argue for this conclusion; the latter further supported this claim by appeal to the functions of motor processes in planning and monitoring action.
+\citet{pacherie:2008_action} and \citet{butterfill:2012_intention} also argue for this conclusion; the latter further support this claim by appeal to the functions of motor processes in planning and monitoring action.
 }
@@
 Which outcomes can be represented motorically?
-Outcomes can be partially ordered by the means-end relation.
+Not all outcomes can be represented motorically, of course.
+For instance, the difference between cutting out an article and cutting out a picture from newspapers is not as such a difference that need be captured by any motor representations.
+One idea about which outcomes can be represented motorically starts from the observation that outcomes can be partially ordered by the means-end relation.
 It might be assumed that motor representations capture only least outcomes relative to the means-end ordering.
-It is certainly true that not all outcomes can be represented motorically.
-For instance, the difference between cutting out an article and cutting out a picture from newspapers is not as such a difference that need be captured by any motor representations.
-But we should not infer from this that motor representations are limited to representing least (in the means-end ordering) outcomes.
-In fact, differences such as that between reaching for a pair of scissors to cut something and reaching for a pair of scissors to place them somewhere (perhaps to put them away) are captured by motor representations (\citealp{Fogassi:2005nf, cattaneo:2007_impairment, bonini:2010_ventral, }).
+This is wrong, however.
+Differences such as that between reaching for a pair of scissors to cut something and reaching for a pair of scissors to place them somewhere (perhaps to put them away) are captured by motor representations (\citealp{Fogassi:2005nf, cattaneo:2007_impairment, bonini:2010_ventral}).
 This suggests that motor representations are not limited to representing least outcomes.
 
+\subsection{How to capture the directedness of actions to outcomes}
 Our overall question is how motor representation could enable goal ascription judgements.
@@
 Rather, she simply finds herself somehow disposed to kick the ball.
-This is not because she intends to pass the ball back to Ludwig or to compete with him or because of any other intention: it is just because representing the outcome has suggested to her the possibility of performing this action.
-So Kate represents the outcome to which Ludwig's action is directed but she does not identify that outcome as the goal of his action.
-More would be required for goal ascription.
-To ascribe a goal to an action is to represent the action and the goal and also to relate the goal to the action.
+This is not because she intends to pass the ball back to Ludwig or to compete with him or because of any intention concerning Ludwig: it is just because representing the outcome has suggested to her the possibility of performing this action.
+So Kate represents the outcome to which Ludwig's action is directed but she does not identify that outcome as a goal to which his action is directed.
+(She does not relate the outcome to his action at all.)  More would be required for goal ascription.
+To ascribe a goal to an action involves not only representing the action and the outcome which is the goal but also relating the outcome to the action.
 
-But how might the directedness of actions to outcomes be represented or otherwise captured?
-In philosophy the notion of goal-directedness is standardly explained in terms of intention: for an action to be directed to a goal is for the action to be appropriately related to an intention whose content specifies that goal (or, on some views, an intention specifying an appropriately related goal).
+It is important in what follows that we sometimes represent intentions without linking them to actions.
+For instance, you might know that a friend intends to break an egg and see the friend handling an egg which breaks without yet knowing whether this breaking was appropriately related to that intention.
+To represent an intention and an action does not necessarily involve representing any relation between the action and the outcome.
+
+How might the directedness of actions to outcomes be represented or otherwise captured?
+In philosophy the notion of goal-directedness is standardly explained in terms of intention: for an action to be directed to a outcome is for the action to be appropriately related to an intention whose content specifies that outcome (or, on some views, an intention specifying an appropriately related outcome).
 Given this, it is perhaps natural to suppose that ascribing a particular goal to a given action might involve representing that action, representing an intention whose content specifies the goal, and representing the action as appropriately related to the intention.
@@
 
-How can you get from motor representations to the directedness of actions to outcomes?
+The problem is not significantly different if we suppose that goal-directedness can be explained in terms of mental representations other than intentions.
+For instance, some have argued that the directedness of an action to an outcome can be explained in terms of motor representation and not only in terms of intention \citep{butterfill:2012_intention}.
+Suppose for the sake of argument that this is correct.
+We still face substantially the same puzzle about how motor representation in goal ascription could capture the directedness of an action to an outcome.
+For motor representations do not represent any mental representations at all.
+So the directedness of an action to an outcome cannot be represented motorically.
 
-If motor representations were representations of kinematic or dynamic features only, the claim that they enable goal ascription would be mysterious.
-In fact, however, motor representations represent outcomes: or so we shall argue.
-This is a first step towards
+One way to avoid our argument for this conclusion would be to claim that the directedness of an action to a goal can be understood in ways that do not involve intention or indeed any representation.
+For instance, an action's being directed to an outcome may consist in its having the function of bringing about that outcome.
+Here function could be construed teleologically.
+Teleological accounts of function, and of the application of this notion to understanding goal-directed action, have been extensively developed (*see further Godfrey-Smith 1996; Millikan 1984; Neander 1991; Price 2001; Wright 1976).
+Suppose we accept for the sake of argument  that goal-directedness can be explained in terms of function rather than intention or any other kind of mental representation.
+Substantially the same puzzle remains.
+After all, motor representations no more represent functions than they represent intentions.
 
-How could motor representation represent the outcome to which an action is directed?
+At this point it seems to us plausible that motor representations cannot represent the directedness of an action to an outcome.
+If we supposed that all goal ascription involved representing directedness, it would be hard to understand the role of motor representation in enabling goal ascription.
+But it would be a mistake to suppose that directedness must be represented in goal ascription.
+There are ways of capturing directedness which do not involve representing it.
+In the rest of this section we shall argue that in goal ascription the link between actions and outcomes to which they are directed can be captured not only by representing it but alternatively by a planning-like process which involves determining how an outcome is to be achieved and whether given movements are appropriate to achieving that outcome.
 
-To understand how motor representation could enable goal ascription we need to understand the relation between motor representations and outcomes and the relation between
+To see how the directedness of an action to an outcome can be captured without being represented, let us first step back to consider cases of goal ascription which involve representing intentions only (so motor representation is not yet in the picture).
+Observing Ayesha gathering some papers during a long meeting that shows no sign of ending, you conjecture that she is intending to leave discretely.
+You use this conjecture about Ayesha's intention to determine one or more likely sequences of behaviours, almost as if you were planning from her point of view how to act on that intention.
+This leaves you with expectations concerning how she will act.
+You might be waiting for Ayesha, who has no other friends present, to give you a nod before quietly pushing away from the table.
+If what you observe matches these expectations, or if there are only deviations that can be explained in ways consistent with your conjecture about Ayesha's intention, then you may come to regard the conjecture as confirmed.
+But if there is a mismatch---if, for example, Ayesha signals that she wants attention and looks around to see whether people are looking at her---then you might reject the conjecture about her intention to leave discretely.
+As this illustrates, goal ascription often (perhaps not always) involves determining how an action will or would unfold, and this process of determination may ensure that there is a reliable connection between the action's being directed to a certain goal and your ascribing that goal to the action.
+So the directedness of an action to a goal can be captured by a planning-like process in addition to being represented.
 
-it is perhaps natural to suppose that ascribing a particular goal to a given action might involve representing the action, representing an intention whose content specifies the goal, and representing the action as appropriately related to the intention.
+You might suggest that where an observer represents another agent's intention, she would only ever determine how the agent will act if she also supposes the actions will be directed to fulfilling this intention.
+This may be right.
+The illustration is only supposed to show that goal-directedness could in principle be captured without being represented.
+
+Motor representations and processes may likewise capture the directedness of an action to an outcome.
+In action observation it is not just that there may be motor representations of outcomes.
+There is evidence that a motor representation of an outcome sometimes causes a determination of which movements are likely to be performed to achieve that outcome  \citep{kilner:2004_motor, urgesi:2010_simulating}.
+The processes involved in determining how observed actions are likely to unfold given their outcomes are closely related, or identical, to processes involved in performing actions.
+This is suggested both by the involvement of regions of the brain associated with action (*refs), and by patterns of facilitation and interference which occur when simultaneously observing and performing actions \citep{brass:2000_compatibility, kilner:2003_interference, craighero:2002_hand}.
+It is almost as if the observer were planning from the agent's point of view how to act.
+These planning-like processes sometimes leave observers with expectations about how an action will unfold.
+For instance, if you were to observe just the early phases of a grasping movement your eyes may jump to its likely target, ignoring nearby objects \citep{ambrosini:2011_grasping}.
+These proactive eye movements resemble those you would typically make if you were acting yourself \citep{Flanagan:2003lm}: in reaching for something, it is natural to look at the target rather than your hand.
+Importantly, the occurrence of such proactive eye movements in action observation depends both on representing the outcome of an action (even temporary interference in your ability to represent the outcome will interfere with the eye movements,  \citealp{Costantini:2012fk}) and also on planning-like processes (having to perform actions incongruent with those you are observing also interferes with proactive eye movements, \citealp{Costantini:2012uq}).
+Finally, going slightly beyond what there is currently direct evidence for, it is also plausible that whether the expectations about how the action will unfold are met or not will influence whether the motor representation of that outcome persists.
+This ensures that, where someone is observing an action, there is a reliable connection between the observer representing an outcome and the action being directed to that outcome.
+There is a sense, then, in which the directedness of an action to a goal can be captured by motor representations and processes: As might also happen in the case of intention, the directedness of the action to a goal is captured by planning-like processes.
+
+Let us stipulate that to capture the directedness of an action to an outcome is to (a) represent the action; and to (b) determine how the action is likely to unfold given that it would normally realise that outcome; where (c) this determination leaves the observer with expectations about how the action will unfold; and (d) whether the representation of the outcome persists depends on the expectations being met.
+We have been arguing that there is evidence that motor representations and processes can capture the directedness of actions.
+
+It is perhaps worth pausing here to stress a parallel between performing and observing actions.
+Start with performance.
+In virtue of what is an action related to the outcome or outcomes to which it is directed?
+Suppose that in performing an action there is a motor representation of an outcome.
+Now an action is not directed to an outcome just in virtue of there being a representation of that outcome, of course.
+Rather the outcome has to play some guiding role in determining how the action will unfold.
+So what makes it the case that the action is directed to the outcome is not just the content of the representation but also its role in determining how the action is to unfold.
+Now switch to action observation.
+How might an observer capture the directedness of an action to an outcome?
+As in performing an action, there is sometimes a motor representation of an outcome and a process of working out how the action will unfold, as we have seen.
+But unlike performing action, in observation the process of working out how the action will unfold does not normally directly cause of the action (of course).
+So if expectations about how the action will unfold are not met, it is not normally possible to alter the action; instead it may be necessary to link the action to an outcome different from the one whose representation generated these unmet expectations.
+This is how the representations and processes which ground the directedness of actions to outcomes in performing actions may also be involved in capturing the directedness in observing actions.
+%
+\begin{quote}
+[***\textit{A previous, rough attempt at the above paragraph}:  \\
+grounding /capturing: In performing an action there is a representation of an outcome.
+Now an action is not directed to an outcome just in virtue of there being a representation of that outcome, of course.
+Rather the outcome has to play some guiding role in working out how the action will unfold.
+So what determines that the action is directed to the outcome is both the representation and the process of working out how the action is to unfold.
+Now the case of action observation is similar.
+There is a representation of an outcome and a process of working out how the action will unfold.
+A difference is that the process of working out how the action will unfold is not (of course) normally a cause of the action.
+So the representations and processes which ground the directedness of actions to outcomes in performing actions: these representations and processes also capture the directedness in observing actions.]
+\end{quote}
+
+In this section we have considered a first obstacle to understanding how judgements about the goals of actions sometimes depend on motor representations.
+The obstacle was that motor representations might appear not to concern the goals of actions at all, making their influence on judgements about goals difficult to understand.
+In removing this obstacle we made two points.
+The first was that motor representations do not all represent merely kinematic features of action; rather, some motor representations represent outcomes to which actions are directed.
+The second point was that although motor representations probably do not represent the directedness of an action to an outcome, reflection on the role of motor representations and processes in action observation shows that this directedness can be captured motorically.
+Potentially, then, there is a respect in which the content of a motor representation may resemble the content of judgement about the goal to which an action is directed: they may represent a single outcome and capture the directedness of an action to that outcome.
+We therefore propose that where a judgement about the goal to which an action is directed depends on motor representations, there is a motor representation of an outcome which captures the directedness of the action to the goal.%
+\footnote{
+***Do we need to allow that the outcomes represented motorically and in judgement may not be identical?
+}
+
+\section{Action and Experience}
+\label{sec:processes}
+Our overall question is how judgements about the goals of actions could ever depend on motor representations.
+The previous section concerned the possibility of a match between the contents of motor representations and judgements.
+Now to say that two varieties of representations---motor representations and judgements---sometimes have nonaccidentally matching contents is not yet to explain how that match might come about.
+Which processes could be responsible for the match?
+When judgements about the goals of observed actions depend on the observers' motor representations, which processes connect motor representations to judgements?
+
+In broad terms, our will be proposal is that the dependence of observational judgements on motor representations goes via experience.
+It is because motor representations can affect our experiences when we observe actions that those motor representations can facilitate judgements about these actions.
+However this is consistent with two quite different hypotheses concerning the experiences associated with observing action.
+
+One hypothesis is that in observing action we experience not only movements and sounds but also goal-directed actions as such: it is this experience which enables judgements of action, and motor impairments affect judgements of action by preventing one from experiencing actions as such.
+On the other hypothesis, in observing actions presented as point-light stimuli or acoustically we experience movements and sounds only.
+Motor representations somehow make these particular experiences of movements and sounds possible, and these experiences in turn ground judgements about actions.
+On both hypotheses motor representations enable experiences which are revelatory of action in the sense that they can provide reasons for correct judgements about which action someone is performing.
+The difference concerns what these revelatory experiences are experiences of: whether some are experiences of goal-directed actions as such (the first hypothesis) or whether all are experiences of movements and sounds only (the second hypothesis).
+
+In what follows we first consider evidence that the second hypothesis is plausible and then, in a further section, we consider arguments for the first hypothesis.
+Although we shall conclude that the first hypothesis is correct---there are experience of action as such---it is helpful to approach this by way of evidence relevant to the second hypothesis.
+
+\subsection{Motor representation can enable visual and acoustic experiences}
+For convenience, in what follows we shall use the term `perceptual experiences' to mean experiences of the visual, acoustic, tactile and other things commonly associated with perception.
+As a terminological stipulation (for which there are reasons, but they come later), no experiences of action as such are perceptual experiences.
+Later, when we have considered arguments that some experiences are experiences of actions as such, we shall contrast `perceptual experience' with `motor experience'.
+
+The possibility that some motor representations enable visual experiences is suggested by research showing that motor plans, such as might be involved in producing a gesture with the hand, can influence visual processes \citep{bortoletto:2011_action}.
+The influence of motor representations on visual processes varies depending on both the gesture to be performed and the visual stimulus presented.
+Importantly, these influences can be detected before any action occurs, showing that it is the motor representations rather than any consequent movements which are influencing visual processes.
+[*Some other evidence; stress that the effect can be found using different measures.]  As it may be possible to influence visual processes in ways that make no difference to experience, we do not take these findings alone to establish the claim that motor representations could enable visual experiences.
+But these findings do at least show that the claim is consistent with the neuropsychological facts.%
+\footnote{
+For a recent review of some of the evidence concerning motor influences on perceptual processing, see \citet{halasz:2012_unconscious}. (*refs)
+}
+
+What would establish that some motor representations enable perceptual experiences?
+For some very special pairs of tones, the first is sometimes perceived as lower in pitch than the second and sometimes as higher in pitch.
+Repp and Knoblich (\citeyear{repp:2007_action}) had a group of expert pianists and a control group of non-experts press two keys in sequence, where the first key was sometimes left and sometimes right of the second key.
+They found that, for the expert pianists, the direction of the key presses modulated the experience: it influenced the perceived direction of the change in pitch.
+Could what modulates experience in this case be not a motor representation but the occurrence of a movement, or perhaps even the perception of a movement of the subject's own fingers?
+Against this, note that the effect was not observed in non-expert pianists: for them the direction of action did not measurably influence the experienced pitch direction.
+Since both groups moved in the same way, if the modulation were due to movement only we would expect it to occur irrespective of expertise.
+Instead it seems likely that differences in expertise between the two groups of subjects affected how the movements they performed were represented motorically, and these differences in motor representation are in turn what explains their experience of the change in pitch.
+This study, and others like it,%
+\footnote{
+See also \citet{zwickel:2010_interference} who investigate effects of action on visual experience of motion.  *more
+} 
+provide relatively direct evidence that motor representations can enable acoustic experiences.
+
+How do these findings bear on the second hypothesis above, according to which motor representation in action observation enables experiences of sounds and movements that somehow facilitate the identification of actions?
+Perhaps the motor representations enable certain experiences of kinematic features which somehow enable subjects to better identify actions.
+We do not claim that this is entirely plausible, only that the second hypothesis cannot be rejected out of hand.
+This is enough to show that we cannot infer that there are experiences of action just from the fact that motor representation enables experiences which are revelatory of action.
+Stronger evidence for the first hypothesis is needed.
+
+\subsection{Experiences of action}
+Our focus is experiences involved in observing actions, but let us first consider experiences involved in performing actions.
+We shall argue that acting sometimes involves experiences which can ground judgements about what one has done even though these are not experiences of movements or sounds.
+In effect, the idea will be that you can strip away perceptual experiences and still be left with experiences revelatory of action.
+These can only be experiences of action, or so we claim.
+Two phenomena are involved in making this argument.
+The first is anosognosia for hemiplegia, which shows that there can be experiences revelatory of action even in the absence of any actual movements.
+The second is deafference, which confirms that these experiences are unlikely to be illusory experiences of movement.
+Together these phenomena support the view that, in performing actions, we have experiences of actions and not only of movements, sounds and the like.
+
+The argument sketched so far concerns experiences associated with performing actions, not observing them.
+Note, however, that the basis of these experiences of action is motor representation and planning.
+Since these are the very things which, in action observation, underlie the experiences enabling one to identify actions, it is reasonable to conclude that there is a phenomenal element common to performing an action and observing an action performed, namely the action itself.
+Roughly, what you experience in performing an action is what you would experience if you were observing it.
+But this is all to come; let us start by introducing some background.
+
+Anosognosia for hemiplegia is a clinical condition in which patients deny, and appear largely unaware of, a severe paralysis of one or more limbs on one side of their body.
+It typically follows a stroke involving damage to motor areas in the right hemisphere of the brain (*refs).
+To illustrate the effects of this condition, consider a subject with anosognosia for hemiplegia who was asked to brush her hair holding a brush in her contralesional hand.
+Although she was unable to move the hand, she proceeded to move her head as if her hair were being brushed, and then reported having successfully brushed her hair \citep{berti:2008_motor}.
+This form of anosognosia cannot be explained as entirely due to neglect or complete ignorance of part of one's body \citep[p.\ 165]{berti:2008_motor}.
+
+Could anosognosia for hemiplegia be explained as mere  confabulation or a top-down effect of judgement on experience?
+This is incompatible with the performance of anosognosics on bimanual tasks \citep{berti:2008_motor, garbarini:2012_moving}.
+To illustrate, consider an anosognosic who has been asked to wash with both hands.
+Although she can only actually move one hand, she may move this hand in much the way that she would do if she were actually washing with both hands (*ref: personal communication?).
+This is hard for ordinary subjects to reproduce (try it).
+For a more carefully controlled illustration, consider anosognosic subjects who were asked to draw simultaneously with both hands, one hand was supposed to continuously draw a vertical line and the other hand to continuously draws a circle.
+Anosognosic subjects can of course only actually move one hand, but they show interference as if both hands were actually being moved.
+That is, if you were to perform this task, your vertical line would become somewhat oval; this interference is an effect of bimanual coordination.
+The striking finding is that anosognosic subjects show the same pattern of interference although not actually drawing a circle.
+By contrast, subjects with motor neglect and subjects with hemiplegia but no anosognosia do not show this pattern of interference \citep{garbarini:2012_moving}.
+This suggests that anosognosia involves relatively detailed action-related experiences and cannot be entirely explained as confabulation.
+
+On the leading, best supported explanation, anosognosia for hemiplegia arises from defects in monitoring action.
+When a subject with anosognosia for hemiplegia is asked to perform an action involving a hemiparetic limb, motor planning occurs as it might do in ordinary subjects.
+Now in ordinary subjects, and in hemiplegic subjects without anosognosia, the predicted effects of a planned action are compared with motor representations of action outcomes.
+Consequently if such subjects tried and failed to initiate an action, they would normally be aware of their failure even without seeing or touching their limbs.
+However, in anosognosic subjects the comparator is damaged and no signal of failure is generated.
+This is why it seems to them that they have acted despite failing even to initiate the action.
+Of course, anosognosic subjects can perceive that their objectives have not been met: after apparently opening a bottle, for instance, they can see that the bottle remains shut; they would not normally proceed to drink from it.
+And these subjects do not necessarily have visual, proprioceptive or tactile sensory impairments.
+But the experience associated with motor planning dominates, so that it seems to them as if they have acted.
+From their point of view, it seems that the bottle is peculiarly difficult to open rather than that they are unable to move a limb \citep[pp.\ 173-4]{berti:2008_motor}.
+
+How is anosognosia for hemiplegia relevant to the issue of whether there are experiences of action as such?
+It seems that anosognosic subjects have some experiences which are revelatory of action---revelatory in the sense that they provide reasons for judgements about actions---despite not actually moving a limb.
+What could these experiences be experiences of?
+Consider the claim that these are experiences of movement only.
+While it is arguably not generally correct to infer an absence of experience of movement from the absence of actual movement, in this particular case it does seems unlikely that there could be experience of movement.
+After all, some anosognosic subjects have intact perceptual systems: if they experience movement at all, they should experience an absence of movement.
+So any experiences of movement will not be the sort of experiences that would provide reasons for judgements about successful action.
+Given that (as argued above) the experiences on which anosognosic subjects' judgements of action are based are also not entirely a product of confabulation, it is plausible to accept that they are experiences of action as such.
+
+This view is further supported by research on deafferented subjects.
+These subjects lack tactile and proprioceptive sensory feedback from their actions but, unlike subjects with anosognosia for hemiplegia, can actually move their limbs.
+Such subjects may continue to have some experiences characteristic of action; for instance, when asked to flex an index finger they can report having done so even in the absence of visual information \citep{kristeva:2006}.
+These subjects' sensory deficits make it unlikely that their experiences could be of movements or the like: it is more plausible that, concerning acting, their experiences are like those of subjects with anosognosia for hemiplegia.
+None experience movement: what they experience is action.
+
+Anosognosia and deafference comprise a useful pair of cases: in anosognosia we have preserved perceptual senses and an absence of actual bodily movement, whereas in deafference we have damaged senses and actual bodily movement.
+But in both cases there are experiences revelatory of action.
+The existence of these disorders indicates that motor representation is sufficient for experiences capable of grounding judgements about actions even in the absence of any perceptual experiences of movements, sounds or the like.
+It is plausible to infer that these experiences are experiences of action as such.
+
+***HERE: todo.
+Two things.
+First, explain the timing experiment---it's not all about subjects with brain damage, there is converging evidence from ordinary subjects.
+Second, explain the comparitor model.
+
+***Anosognosia for hemiplegia : experience of action without acting (cannot be an experience of movement, \&c).
+Or could this be an illusory experience of one's body moving?
+\citep{garbarini:2012_moving}%
+\footnote{
+See also \citet{gallese:2010_bodily}: "Brugger, Kollias, Muri, Crelier, and Hepp-Reymond (2000) described one of these cases.
+fMRI imaging of their patient during phantom limb sensations of hand movements showed activation of premotor and posterior parietal cortex.
+Aplasic phantoms can therefore be explained as the phenomenal correlate of planning/monitoring action of an absent limb."}
+and hand-washing: \citep{berti:2008_motor}
+
+***parallel with speech (short)
+
+\section{Conclusion}
+Motor representation enables judgements about the goals of actions by making possible experiences of action as such.
+
+\subsection{Reasons}
+In making judgements about the goals of actions, it appears (perhaps veridically) to the thinker that she has reasons for those judgements, much as it appears to her that she has when making visual judgements about the particular colours of objects.
+Or, if we are sceptical about thinkers having reasons for visual judgements, at least we can say that it does not seem to the thinker that she is simply landed with the judgement.
+Contrast someone who `knew at a glance' that a vase is genuine Ding ware, or someone on a blind date who, on first catching sight of her partner for the night, `just knows' that this is the one she wants to spend the rest of her life with.
+Action judgements often seem different from these: rather than simply being landed with them, it may seem to a thinker that they bear a straightforward relation to her experience.
+
+Our view is consistent with the claim that it appears to a thinker that she has reasons for making judgements about the goals of actions even when  those judgements depend on motor representations.
+  Motor representation stands to experience of action as visual representation stands to experience of particular colours.
+
+\subsection{Goal ascription without metarepresentation}
+Having noted above that goal ascription often or always involves representing the directedness of an action to an outcome, we argued that this is something no motor representation can represent.
+Equally, however, we noted that the directedness of actions to outcomes can be captured by planning-like processes, and we suggested that motor representations can feature in such processes.
+One possibility, then, is that there might be a kind of goal ascription---or something resembling goal ascription---where the only representations involved are motor representations.
+This enables us to see how there could be goal ascription, or something very like it (we are not insisting that the term 'goal ascription' be used) could occur without metarepresentation.
+This gives us a first role for motor representation in goal ascription which distinguishes motor representation from representations of intentions or other mental states.
+Where goal ascription involves capturing the directedness of actions to goals by representations of intentions it thereby involves metarepresentation; but where the only representations involved in goal ascription are motor representations, there is no metarepresentation at all.
 

### 2012-08-26 intention motor representation joint action (paper 3)/intention and motor representation in joint action.tex
bits and pieces

@@
 By contrast, 
-most philosophers and many (but not all) psychologists appear to assume that this question can be fully answered just in terms of a special kind of intention or structure of intention, knowledge and committment often called a shared (or `collective') intention.
+most philosophers and many (but not all) psychologists appear to assume that this question can be fully answered just in terms of a special kind of intention or structure of intention, knowledge and commitment often called a shared (or `collective') intention.
 On this widely held view, for an event to be a joint action is for it to be appropriately related to a shared intention.
@@
 \section{The Contrast Cases}
+We have been arguing that social motor representation can bind multiple agents' activities together and link them to an outcome, and this in much the way that shared intention does  so.
+This is a reason, 
+not decisive but perhaps sufficient in the absence of strong contrary reasons, 
+for supposing that social motor representation plays a role in explaining which events are joint actions.
+Perhaps some events are joint actions in virtue of being appropriately related not to shared intention but to social motor representation.
+
+As mentioned at the start, our aim is to extend and generalise existing theories, not to replace them. 
+But at this point our answer to the question of which events are joint actions seems to involve a disjunction.
+This is both puzzling and inelegant. 
+(Unless, of course, social motor representation can be regarded as a special case of shared intention.  
+We explain why we reject this possibility below.)
+In this section we explain how an account of joint action in terms of shared intention can be generalised to accommodate the possibility that social motor representation plays a role in explaining which events are joint actions.
+
+What guides and constrains theorising about shared intention? 
+While there is little agreement on what shared intention is or what it is for, intuitions about it are sometimes grounded in contrast cases.
+Contrast cases are pairs of events which are similar in terms of the behaviour and coordination they involve but where one is a joint action while the other is not.  
+Thus \citet{gilbert_walking_1990} contrasts two people walking together with two people individually walking side by side.  
+The two pairs' movements may be the same and similarly coordinated (to avoid collision), but walking together is a joint action whereas merely walking side by side is not. 
+Relatedly,  \citet{Searle:1990em}  contrasts a case in which several park visitors simultaneously run to a central shelter in order to perform a dance with another case in which the park visitors run to the central shelter in order to escape a storm.
+The first is a case of joint action, the second is not; but the same movements occur in both.
+These sorts of contrast case invite the question, 
+How do joint actions differ from individual but parallel actions? 
+Gilbert’s example shows that the difference can’t just be a matter of coordination, because people who are merely walking alongside each other also need to coordinate their actions in order to avoid colliding.
+And Searle’s example shows that the difference between joint action and parallel individual action can’t just be that the actions have a common effect because merely parallel actions can have common effects too.
+
+How might the contrast cases be used to guide and constrain theorising about shared intention?
+The idea that they serve this function seems to rest on a premise:
+It is possible to distinguish systematically between the contrast cases by appeal to shared intention and only by appeal to shared intention.
+We shall show that there is another, more general way to distinguish contrast cases.
+
+Take Searle's example of several people running to the shelter.  
+In the joint case, there is a single outcome to which each person's actions is individually directed, namely their collective arrival at the shelter.  
+In the contrasting individual case, where park visitors individually run to the shelter to escape a storm, there is no single outcome to which each of their actions is individually directed.  
+Instead each visitor's actions are directed to that visitor's own arrival at the shelter.  
+Similarly, turning to Gilbert's example, when two people walk together, there is a single outcome to which both of their actions are individually directed (their collective arrival at a corner, say); whereas when two people merely walk side by side there is no single outcome to which each agent's actions are directed.
+In general, where two or more agent's actions constitute a joint action, there is an outcome to which each agent's actions are individually directed such that it is possible for them all to succeed relative to this outcome.%
+\footnote{
+Some readers may be sceptical of this claim.
+But note that it is a consequence of the view we are opposing, the view that all joint actions involve shared intention.
+Or, rather, it is a consequence of this view given a further premise which is a consequence of any standard account of shared intention: Where several agents act on a shared intention that they J, 
+each will perform an action directed to J and it is possible for them all to succeed in J-ing.
+} 
+Since this is not true in the contrasting cases of non-joint action, it allows us to distinguish systematically between joint actions and their non-joint but behaviourally and coordinatively indistinguishable counterparts.
+
+For concision let us stipulate that an outcome is a \emph{distributive goal} of two or more agents' actions just if two conditions are met.
+First, this outcome is a goal to which each agent's actions are individually directed.
+Second, each agent's actions are related to the goal in such a way that it is possible for all the agents (not just any agent, all of them together) to succeed relative to this goal.
+In these terms, our claim is that joint actions involve distributive goals and that this distinguishes joint actions from non-joint actions in the standard contrast cases.  
+Since social motor representation and shared intention are each sufficient for the existence of a distributive goal, we can conclude that systematically distinguishing between standardly considered contrast cases does not require shared intention.
+
+A natural response to this argument would be to suggest that the standard contrast cases are insufficient. 
+For in some cases two agents' actions can have a distributive goal although arguably there is no joint action. 
+Nora and Olive killed Fred.  
+Each fired a shot.
+Each intended that her shooting ground or partially ground Fred's death.%
+\footnote{
+Events $D_1$, ...\ $D_n$ \emph{ground} $E$, if: $D_1$, ...\ $D_n$ and $E$ occur; 
+$D_1$, ...\ $D_n$ are each part of $E$; and 
+every event that is 
+	a part of $E$
+	but does not overlap $D_1$, ...\ $D_n$ 
+is caused by some or all of $D_1$, ...\ $D_n$.
+(This is a refinement and generalisation of a notion due to \citet{pietroski_actions_1998}.)
+Event $D$ \emph{partially grounds} event $E$ if $D$ alone does not ground $E$ but there are events including $D$ which do ground $E$.
+} 
+%
+As it turned out, both intentions were fulfilled.
+Neither shot was individually fatal but together they were deadly.
+An ambulance arrived on the scene almost at once but Fred didn't make it to the hospital.
+Now Nora and Olive's actions have a distributive goal.
+After all, 
+	each agent's actions are individually directed to Fred's death
+	and  
+	it is consistent with the stipulations made about this scenario that these goal relations are compatible in the sense that both agents could succeed together. 
+But is their killing of Fred a joint action?
+Imagine that Nora and Olive had no knowledge of each other, nor of each other's actions, and that their efforts were entirely uncoordinated.
+We might even suppose that Nora and Olive are so antagonistic to each other that they would, if either knew the other's location, turn their guns on each other.
+Their actions nevertheless have a distributive goal.
+But given these further suppositions it is likely to seem counterintuitive to suppose that their killing Fred was a joint action. 
+So although we can distinguish the standardly considered contrast cases just by appeal to the notion of a distributive goal, we should not assume that this notion captures the intended contrast between joint and merely parallel actions.
+
+***HERE refine: collective goals
 
@@
 The best argument we know of hinges on so-called contrast cases.
-Contrast cases are pairs of actions which are similar in terms of behaviour and coordination but where one is joint but the other isn't.  
-\citet{gilbert_walking_1990} contrasts two people walking together with two people individually walking side by side.  
-The two pairs' movements may be the same and similarly coordinated (to avoid collision), but walking together is a joint action whereas merely walking side by side is not. 
-Relatedly,  \citet{Searle:1990em}  contrasts a case in which several park visitors simultaneously run to a central shelter in order to perform a dance with another case in which the park visitors run to the central shelter in order to escape a storm.
-The first is a case of joint action, the second is not; but the same movements occur in both.
-These sorts of contrast case invite the question, 
-How do joint actions differ from individual but parallel actions? 
-Gilbert’s example shows that the difference can’t just be a matter of coordination, because people who are merely walking alongside each other also need to coordinate their actions in order to avoid colliding.
-And Searle’s example shows that the difference between joint action and parallel individual action can’t just be that the actions have a common effect because merely parallel actions can have common effects too.
-
-How could the contrast cases be used to argue that all joint action involves shared intention?
-The argument rests on this premise:
-It is possible to distinguish systematically between the contrast cases by appeal to shared intention and only by appeal to shared intention.
-To defeat this argument we shall show that there is another, more general way to distinguish contrast cases.
-
-Take Searle's example of several people running to the shelter.  
-In the joint case, there is a single outcome to which each person's actions is individually directed, namely their collective arrival at the shelter.  
-In the contrasting individual case, where park visitors individually run to the shelter to escape a storm, there is no single outcome to which each of their actions is individually directed.  
-Instead each visitor's actions are directed to that visitor's own arrival at the shelter.  
-Similarly, turning to Gilbert's example, when two people walk together, there is a single outcome to which both of their actions are individually directed (their collective arrival at a corner, say); whereas when two people merely walk side by side there is no single outcome to which each agent's actions are directed.
-In general, where two or more agent's actions constitute a joint action, there is an outcome to which each agent's actions are individually directed such that it is possible for them all to succeed relative to this outcome.%
-\footnote{
-Some readers may be sceptical of this claim.
-But note that it is a consequence of the view we are opposing, the view that all joint actions involve shared intention.
-Or, rather, it is a consequence of this view given a further premise which is a consequence of any standard account of shared intention: Where several agents act on a shared intention that they J, 
-each will perform an action directed to J and it is possible for them all to succeed in J-ing.
-} 
-Since this is not true in the contrasting cases of non-joint action, it allows us to distinguish systematically between joint actions and their non-joint but behaviourally and coordinatively indistinguishable counterparts.
-
-For concision let us stipulate that an outcome is a \emph{distributive goal} of two or more agents' actions just if two conditions are met.
-First, this outcome is a goal to which each agent's actions are individually directed.
-Second, each agent's actions are related to the goal in such a way that it is possible for all the agents (not just any agent, all of them together) to succeed relative to this goal.
-Let us also say that a structure is \emph{something like shared intention} if it stands to joint action roughly as intention stands to ordinary, individual action.  
-In these terms, our claim is that joint actions involve distributive goals and that this distinguishes joint actions from non-joint actions in the standard contrast cases.  
-Given that the existence of a distributive goal need involve nothing like a shared intention, we can conclude that systematically distinguishing between the contrast cases does not require anything like shared intention.
+***[moved]
 

### 2012-08-26 joint action and the emergence of mindreading/joint_action_and_the_emergence_of_mindreading.tex
bits and pieces

%!TEX TS-program = xelatex
%!TEX encoding = UTF-8 Unicode

%\def \papersize {a4paper}
\def \papersize {letterpaper}

\documentclass[12pt,\papersize]{extarticle}
% extarticle is like article but can handle 8pt, 9pt, 10pt, 11pt, 12pt, 14pt, 17pt, and 20pt text

\def \ititle {Joint Action and the Emergence of Mindreading}
\def \isubtitle {}
\def \iauthor {}
\def \iauthor {S. Butterfill*
\\ 
**Department of Philosophy, University of Warwick, UK
}
\def \iemail{s.butterfill@warwick.ac.uk}
%\date{}

\input{$HOME/Documents/submissions/preamble_steve_paper2}
%\author{}
%\author {S. Butterfill* \& C. Sinigaglia**
%\\ 
%**Department of Philosophy, University of Warwick, UK
%\\ 
%***Dipartimento di Filosofia, Università degli Studi di Milano, Italia}
\date{}

\begin{document}

\setlength\footnotesep{1em}

\bibliographystyle{newapa} %apalike

\maketitle
%\tableofcontents
\title{}

\begin{abstract}
\noindent
***
\end{abstract}

\tolerance=5000

\section{A Challange and a Conjecture}

I want to start with a challenge.  
The challenge is to explain the emergence, in evolution or development (or both), of sophisticated forms of mindreading. 
By `mindreading' I mean the ascription to others of mental states and the use of such ascriptions in explaining and predicting their thoughts or actions. 
%%compare:
%`In saying that an individual has a theory of mind, we mean that the individual imputes mental states to himself and to others ... [T]he system can be used to make predictions, specifically about ... behavior'
%(Premack & Woodruff 1978: 515)
And by  `sophisticated forms of mindreading', I mean those which involve representing perceptions, knowledge states, intentions, beliefs and other propositional attitudes. 
%In other words, the sort of mindreading  measured by tasks such as the famous false belief task \citep{Wimmer:1983dz}.

Several researchers have conjectured that social interaction partially explains how sophisticated forms of cognition, including mindreading, emerge in development or evolution.
According to Moll and Tomasello' Vygotskian Intelligence Hypothesis, 
%
\begin{quote} 
`the unique aspects of human cognition ... were driven by, or even constituted by, social co-operation'
\citep[p.\ 1]{Moll:2007gu}.
\end{quote}
%
Independently, Knoblich and Sebanz hypothesise that: 
\begin{quote} 
`perception, action, and cognition are grounded in social interaction
 … functions traditionally considered hallmarks of individual cognition originated through the need to interact with others' \citep[p.\ 103]{Knoblich:2006bn}.
\end{quote}
%
Although these hypotheses concern social interaction or social co-operation generally, both sets of researchers focus on a particular kind of social interaction, 
I also want to focus on joint action.
The conjecture I am interested in links mindreading and joint action:
%
\begin{quote}
The existence of abilities to engage in joint action partially explain how sophisticated forms of mindreading emerge in evolution or development (or both).%
\end{quote}
%
I am not certain that this conjecture would be endorsed by Moll and Tomasello, or by Knoblich and Sebanz.
But it is inspired by their ideas and findings.
In what follows I present a series of objections to the conjecture and explain how they can be overcome.
I do not hope to show that the conjecture is correct. 
My aim is only to show that, so far, there are no good reasons to reject it.
Doing this will involve making clearer what the conjecture amounts to and relating it to some testable predictions.  
This may not sound like much of an achievement. 
Shouldn't I roll my sleeves up and actually test the conjecture before writing about it? 
Perhaps.
But note that there is no good alternative to this conjecture, and that we currently lack suitable conceptual tools even to properly understand, let alone test, it. 
In particular the notions of mindreading and joint action are  in need of some empirically motivated philosophical attention.
Anyway, we have to start somewhere if we are going to explain the emergence of mindreading: I proposal we start with the conjecture.

I have been using the term `joint action' without explanation. 
It is not easy to say what joint action is; this (along with how to distinguish kinds of mindreading) is a central issue in what follows. 
Researchers normally introduce it just by giving examples.  Paradigm cases in philosophy include two people 
	painting a house together \citep{Bratman:1992mi},
	lifting a heavy sofa together  \citep{Velleman:1997oo},
	preparing a hollandaise sauce together \citep{Searle:1990em},
	going to Chicago together \citep{Kutz:2000si},
	and walking together \citep{gilbert_walking_1990}.
In developmental psychology paradigm cases of joint action include  two people 
	tidying up the toys together \citep{Behne:2005qh},
	cooperatively pulling handles in sequence to make a dog-puppet sing \citep{Brownell:2006gu},
	and bouncing a block on a large trampoline together \citep{Tomasello:2007gl}.
Other paradigm cases from research in cognitive psychology include two people
	lifting a two-handled basket  \citep{Knoblich:2008hy},
	putting a stick through a ring \citep{ramenzoni_joint_2011},
	and swinging their legs in phase \citep%[p. 284]
	{schmidt_richardons:_2008}.
We need to be careful because it is not obvious that there is a single phenomenon of which all these are paradigm cases.

A better way to introduce joint action is by using contrast cases.
Contrast cases are pairs of events which are similar in terms of the behaviour and coordination they involve but where one is a joint action while the other is not.  
Thus \citet{gilbert_walking_1990} contrasts two people walking together with two people individually walking side by side.  
The two pairs' movements may be the same and similarly coordinated (to avoid collision), but walking together is a joint action whereas merely walking side by side is not. 
Relatedly,  \citet{Searle:1990em}  contrasts a case in which several park visitors simultaneously run to a central shelter in order to perform a dance with another case in which the park visitors run to the central shelter in order to escape a storm.  The first is a case of joint action, the second is not; but the same movements occur in both.  
These contrast cases invite the question, 
How do joint actions differ from individual but parallel actions? 
Gilbert's example shows that the difference can’t just be a matter of coordination, because people who are merely walking alongside each other also need to coordinate their actions in order to avoid colliding.  
And Searle's example shows that the difference between joint action and parallel individual action can’t just be that the actions have a common effect because merely parallel actions can have common effects too. 
Contrast cases at least enable us agree on what joint action isn't.

My plan is simple. 
The first objection is that sophisticated forms of mindreading emerge (in evolution or development) before joint action. 
In that case, the emergence of mindreading could hardly involve joint action. 
The reply to this objection focuses on kinds of mindreading.  The second (of two) objections is that engaging in joint action requires sophisticated mindreading.
If this were right, any appeal to joint action in an account of how sophisticated forms of mindreading emerge  would presuppose what was supposed to be explained.   
In considering this objection, we will see that some kinds of joint action do require sophisticated mindreading. 
A good reply therefore depends on showing that there is more than one kind of joint action, and that not all joint action requires sophisticated kinds of mindreading.
At this point we will have removed some objections to the conjecture.
It is then natural to ask how it could to work.
How could joint action be involved in explaining the emergence of sophisticated forms of mindreading?
We will see how materials from the replies to the two objections can be assembled to answer this question.

\section{First Objection}

\bibliography{$HOME/endnote/phd_biblio}

\end{document}

### 2012-08-27 intention motor representation joint action (paper 3)/conf collective intentionality manchester 2012/motor representation and shared intention.tex
before change question

@@
 
-\def \ititle {Motor Representation and Shared Intention}
+\def \ititle {Motor Representation \\ and Shared Intention}
 \def \isubtitle {}
 \def \iauthor {
-%S. Butterfill* \& C. Sinigaglia**
-%\\ 
-%**Department of Philosophy, University of Warwick
-%\\ 
-%***Dipartimento di Filosofia, Università degli Studi di Milano
-}
+\textit{Collective Intentionality}, Manchester, August 2012}
 \def \iemail{s.butterfill@warwick.ac.uk}
@@
 
-{\label{text:second_consideration} A second consideration} in favour of this view is a possible parallel between shared intention and ordinary, individual action.
+{\label{text:second_consideration} A second consideration} in favour of this view is a possible parallel between joint action and ordinary, individual action.
 If you think that events are actions in virtue of being appropriately related to intentions, 
 and if you think that shared intention stands to joint action as plain vanilla intention stands to ordinary, individual action,
-then it seems reasonable to suppose that for an event to be a joint action is for it to be appropriately related to a shared intention (*cite Pacherie Synthese forthcoming).
+then it seems reasonable to suppose that for an event to be a joint action is for it to be appropriately related to a shared intention (as \citealp{Pacherie:2012fk} suggests).
 
 \subsection{My view}
-But the parallel assumption about ordinary, individual action is not universally accepted. 
-In fact, some allow that there could be agents whose actions are purposive---and, in some cases, even intentional---although  the agents have no intentions at all.
-For instance, Michael Bratman describes a creature who `acts on the basis of its beliefs and considered desires' only (not intentions) as doing things `intentionally'.%
-\footnote{
-For instance, \citet[p.\ 251]{bratman:2000_valuing} % = \citep[p.\ 51]{Bratman:2007bd}
-describes an agent called `Creature 2' who `acts on the basis of its beliefs and considered desires' only (not intentions) as doing things `intentionally'.
-}
-Others have argued that motor representation and intention are both needed in explaining which events are ordinary, individual actions \citep{pacherie:2000_content,butterfill:2012_intention}.
+On the second consideration, 
+we take for granted that there is a parallel between joint action and ordinary, individual action.
+But is the standard view about the relation between action and intention correct?
+
+Not everyone accepts the standard view.
+Some philosophers allow that there could be agents whose actions are purposive---and, in some cases, even intentional---although  the agents have no intentions at all.
+For instance, Michael Bratman describes a creature who `acts on the basis of its beliefs and considered desires' only (not intentions) as doing things `intentionally' \citet[p.\ 251]{bratman:2000_valuing}.%
+%\footnote{
+%For instance, \citet[p.\ 251]{bratman:2000_valuing} % = \citep[p.\ 51]{Bratman:2007bd}
+%describes an agent called `Creature 2' who `acts on the basis of its beliefs and considered desires' only (not intentions) as doing things `intentionally'.
+%}
+
+Others reject the standard view without asserting there can be action without intention.
+Instead they claim that ingredients other than intention are needed to say which events are actions.
+In particular, some claim that motor representation and intention are both needed in explaining which events are ordinary, individual actions \citep{pacherie:2000_content,butterfill:2012_intention}.
 If any such view is correct, it is impossible to fully explain what it is for an event to be an action by appeal to intention alone.
@@
 In fact recognising that motor representation is needed to say which events are joint actions may make it easier to understand what shared intention is.
+
+\section{Intention and Motor Representation}
+So that you can see where this is going, let me briefly explain what stands behind the view that motor representation is needed in addition to intention in saying which events are actions.
+
+A basic question about ordinary, individual action is:
+What is the relation between a purposive action and the outcome or outcomes to which it is directed?
+Purposive actions typically have many actual outcomes.
+Grabbing little Isabel by the hands I swing her around, causing her to laugh and, simultaneously, breaking a vase.
+Either or both of these might be outcomes to which my action is directed.
+Note also that some or all of the outcomes to which my action was directed might not be among its actual outcomes; after all, actions can fail.
+So among all the actual and possible outcomes of my action, one or some are singled out as specially related to this action.
+The question is what singles out the outcome or outcomes, actual or merely possible, to which a particular purposive is directed.
+
+This question is closely related to a second.
+Ordinary purposive actions are sometimes composed of more than one motor action.  My swinging Isabel around includes my reaching for her wrists, grasping them and then spinning us around.  
+But  my action doesn’t include other things which I might be doing simultaneously, like refusing a cup of tea with my eyes or  trying to determine whether that smell is coming from Isabel’s sister Hannah’s nappy.
+The second question, then, is this: For a particular action directed to a specified goal, what determines which activities%
+\footnote{
+As we use the term `activities',
+all actions are activities 
+and so are  things like reaching and grasping.
+This should make it clear that our position does not depend on whether or not reachings and the like are  actions.
+} 
+comprise the purposive action and which do not?
+
+The standard answer to both questions involves intention.
+An intention represents an outcome, coordinates the one or several activities which comprise the action, and coordinates these activities in a way that would normally facilitate the outcome's occurrence.
+What binds component activities together into larger purposive actions?  
+It is the fact that these actions are all consequences of plans involving a single intention (and are all appropriately related to those plans).
+What singles out an actual or possible outcome as one to which the component activities are collectively directed?  
+It is the fact that this outcome is represented by the intention.
+So the intention is what binds component actions together into purposive actions and links the action taken as a whole to the outcomes to which they are directed.
+
+Motor representations are relevantly similar to intentions,
+as some have recently argued (\citealp[pp.\ 189-90]{pacherie:2008_action}; \citealp{butterfill:2012_intention}).
+Of course motor representations differ from intentions in some important ways (as these authors note).
+But they are similar in the respects that matter for explaining the purposiveness of action.
+For, like intentions, some motor representations represent outcomes and not merely kinematic features of action.
+Like intentions, some motor representations play a role in coordinating multiple  component activities by virtue of their role as elements in hierarchically structured plans.
+And, like intentions, some motor representations coordinate these activities in a way that would normally facilitate the outcome’s occurrence.
+So anyone who accepts the standard story about purposive action and intention
+should also accept that a similar story about purposive action and motor representation.
+Given that the two basic questions about ordinary, individual purposive action can also be answered by appeal to intention,
+they can also be answering by appeal to motor representation.
+This motivates the following view.
+Not all purposive actions are bound together and linked to outcomes by intentions. 
+In some cases what binds together purposive actions and links them to outcomes are motor representations.
+%In some cases, the purposiveness of an action is grounded in a motor representation of an outcome; in other cases it is grounded in an intention.
+And of course in many cases it may be that both intention and motor representation are involved.
+
+This is one reason for supposing that motor representation is not merely an enabling condition for ordinary, individual action but also plays a role in explaining what action is (just as intention does).
+Now let’s return to joint action.
 
@@
 
-A basic question about ordinary, individual action is:
-What is the relation between a purposive action and the outcome or outcomes to which it is directed?
-Purposive actions typically have many actual outcomes.
-Grabbing little Isabel by the hands I swing her around, causing her to laugh and, simultaneously, breaking a vase.
-Either or both of these might be outcomes to which my action is directed.
-Note also that some or all of the outcomes to which my action was directed might not be among its actual outcomes; after all, actions can fail.
-So among all the actual and possible outcomes of my action, one or some are singled out as specially related to this action.
-The question is what singles out the outcome or outcomes, actual or merely possible, to which a particular purposive is directed.
-
-This question is closely related to a second.
-Ordinary purposive actions are sometimes composed of more than one motor action.  My swinging Isabel around includes my reaching for her wrists, grasping them and then spinning us around.  
-But  my action doesn’t include other things which I might be doing simultaneously, like refusing a cup of tea with my eyes or  trying to determine whether that smell is coming from Isabel’s sister Hannah’s nappy.
-The second question, then, is this: For a particular action directed to a specified goal, what determines which activities%
-\footnote{
-As we use the term `activities',
-all actions are activities 
-and so are  things like reaching and grasping.
-This should make it clear that our position does not depend on whether or not reachings and the like are  actions.
-} 
-comprise the purposive action and which do not?
-
-The standard answer to both questions involves intention.
-An intention represents an outcome, coordinates the one or several activities which comprise the action, and coordinates these activities in a way that would normally facilitate the outcome's occurrence.
-What binds component activities together into larger purposive actions?  
-It is the fact that these actions are all consequences of plans involving a single intention (and are all appropriately related to those plans).
-What singles out an actual or possible outcome as one to which the component activities are collectively directed?  
-It is the fact that this outcome is represented by the intention.
-So the intention is what binds component actions together into purposive actions and links the action taken as a whole to the outcomes to which they are directed.
-
-Motor representations are relevantly similar to intentions,
-as some have recently argued (\citealp[pp.\ 189-90]{pacherie:2008_action}; \citealp{butterfill:2012_intention}).
-Of course motor representations differ from intentions in some important ways (as these authors note).
-But they are similar in the respects that matter for explaining the purposiveness of action.
-For, like intentions, some motor representations represent outcomes and not merely kinematic features of action.
-Like intentions, some motor representations play a role in coordinating multiple  component activities by virtue of their role as elements in hierarchically structured plans.
-And, like intentions, some motor representations coordinate these activities in a way that would normally facilitate the outcome’s occurrence.
-So anyone who accepts the standard story about purposive action and intention
-should also accept that a similar story about purposive action and motor representation.
-Given that the two basic questions about ordinary, individual purposive action can also be answered by appeal to intention,
-they can also be answering by appeal to motor representation.
-This motivates the following view.
-Not all purposive actions are bound together and linked to outcomes by intentions. 
-In some cases what binds together purposive actions and links them to outcomes are motor representations.
-%In some cases, the purposiveness of an action is grounded in a motor representation of an outcome; in other cases it is grounded in an intention.
-And of course in many cases it may be that both intention and motor representation are involved.
-
-This is one reason for supposing that motor representation is not merely an enabling condition for ordinary, individual action but also plays a role in explaining what action is (just as intention does).
-Now let’s return to joint action.
+[***moved to earlier]
 

### 2012-08-27 intention motor representation joint action (paper 3)/conf collective intentionality manchester 2012/motor representation and shared intention.tex
final version? Monday's effort

@@
 \section{Intention and Motor Representation}
+[\emph{*Probably skip all of this section in the talk}]
+
 So that you can see where this is going, let me briefly explain what stands behind the view that motor representation is needed in addition to intention in saying which events are actions.
@@
 and 
-coordinates the several agents’ activities in such a that would normally facilitate the occurrence of the represented outcome.
+coordinates the several agents’ activities in such a way that would normally facilitate the occurrence of the represented outcome.
 This is how a shared intention can bind together the activities comprising a joint action and link them to an outcome.

### 2012-08-28 intention motor representation joint action (paper 3)/conf collective intentionality manchester 2012/motor representation and shared intention.tex
manchester tuesday morning

@@
 We don’t mean to suggest that all joint actions involve social motor representation.
-The view we are aiming to establish is rather this: Some joint actions involve social motor representation; and there could be joint actions which involve social motor representation but not shared intention.
+The view we are aiming to establish is rather this: Some events are joint actions in virtue of being appropriately related to social motor representations which bind their components together and ensure that there is a single outcome to which these components are collectively directed.
 This is why 
@@
 %This is why I think that 
-fully understanding what joint action is requires understanding the coordinating role of social motor representation and not only understanding shared intention.
+understanding which events are joint actions requires understanding not only shared intention but also
+ the coordinating role of social motor representation.
+
+This gives rise to a challenge. 
+Effective joint action—imagine two people erecting a tent in a gale together—sometimes requires both shared intentions and social motor representation plus a certain kind of harmony between the two. 
+To characterise the harmony required, let us define a notion of matching:
+%
+\begin{quote}
+Two  outcomes, A and B, \emph{match} in a particular context just if, in that context, either the occurrence of A would normally constitute or cause, at least partially, the occurrence of B or vice versa.
+\end{quote}
+%
+When joint action involves both shared intention and social motor representation, the outcomes each of these representations specify should match.
+This is the required harmony.
+The challenge is to explain how matching could ever non-accidentally occur.
+%This will be particularly difficult if you suppose, as we do, that motor representation and intention are not inferentially integrated.
+
+I think we can see a possible solution to this challenge by considering another problem.
+In building shared intention from ordinary individual intention, we need intentions \emph{that we J}.
+As has been much discussed, the contents of these intentions cannot all refer to actions involving shared intentions.
+For this reason Michael Bratman suggests that things we intend are  cooperatively neutral activities.
+It is then necessary to add further intentions in order to transform cooperatively neutral activities into joint actions.
+But it also seems possible that in some cases, what we intend when we intend that we J is not a cooperatively neutral activity but instead a joint action of the sort which involves social motor representation.
+
+So perhaps harmony between shared intention and social motor representation is sometimes achieved in this way: what we intend when we share an intention is the sort of joint action that involves social motor representation.
 

### 2012-09-05 intention motor representation joint action (paper 3)/conf collective intentionality manchester 2012/motor representation and shared intention.tex
tidying up after various talks

@@
 \begin{quote}
-Two  outcomes, A and B, \emph{match} in a particular context just if, in that context, either the occurrence of A would normally constitute or cause, at least partially, the occurrence of B or vice versa.
+Two  outcomes, A and B, \emph{match} in a particular context just if, in that context, either the occurrence of A would normally cause or constitute, at least partially, the occurrence of B or vice versa.
 \end{quote}
@@
 I think we can see a possible solution to this challenge by considering another problem.
-In building shared intention from ordinary individual intention, we need intentions \emph{that we J}.
+In constructing realisers of shared intention from ordinary individual intention, we need intentions \emph{that we J}.
 As has been much discussed, the contents of these intentions cannot all refer to actions involving shared intentions.
@@
 
+In conclusion, two things.
+First, some events are joint actions in virtue of being appropriately related to social motor representations which bind their components together and ensure that there is a single outcome to which these components are collectively directed.
+Second, and much more tentatively, in some cases social motor representation may be among the ingredients that realise a shared intention.
+
 \bibliography{$HOME/endnote/phd_biblio}

### 2012-09-05 joint action and knowing others minds/submit philosophical studies  2012-01/second submit/conf budapest ceu talk cognitive 2012-09/interacting mindreaders HANDOUT.tex
tidying up after various talks

%!TEX TS-program = xelatex
%!TEX encoding = UTF-8 Unicode

\documentclass[11pt]{extarticle}
% extarticle is like article but can handle 8pt, 9pt, 10pt, 11pt, 12pt, 14pt, 17pt, and 20pt text

\def \ititle {Joint Action and the Emergence of Mindreading}
\def \isubtitle {}
\def \iauthor {Stephen A. Butterfill}
\def \iemail{s.butterfill@warwick.ac.uk}
\date{}

%remove all section numbers
 \setcounter{secnumdepth}{0}

%for strikethrough
\usepackage[normalem]{ulem}

\input{$HOME/Documents/submissions/preamble_steve_handout}

\bibpunct{}{}{,}{s}{}{,}  %use superscript TICS style bib
%remove hanging indent for TICS style bib
%TODO doesnt work
\setlength{\bibhang}{0em}
%\setlength{\bibsep}{0.5em}

%itemize bullet should be dash
\renewcommand{\labelitemi}{$-$}

\begin{document}

\begin{multicols}{3}

\setlength\footnotesep{1em}

\bibliographystyle{newapa} %apalike

%\maketitle
%\tableofcontents

\

\begin{center}
{
\textbf{Joint Action and the Emergence of Mindreading}
\Large{Interacting Mindreaders}
}

%Stephen A. Butterfill

<s.butterfill@warwick.ac.uk>

\end{center}

Could an interacting mindreader be in a position to know things which she would be unable to know were she unable to interact with her targets?

\section{Ordinary 3$^{\textrm{\tiny rd}}$ person interpretation}

Csibra \& Gergely's principle of rational action: `an action can be explained by a goal state if, and only if, it is seen as the most justifiable action towards that goal state that is available within the constraints of reality.'\citep{Csibra:1998cx,Csibra:2003jv}

%(Cf.\ Southgate et al's principle of efficiency:
%`goal attribution requires that agents expend the least possible amount of energy within their motor constraints to achieve a certain end.' \citep%[p.\ 1061]
%{Southgate:2008el})

%

These facts:
%
\begin{enumerate}
%
\item action $a$ is directed to some goal;
%
\item actions of $a$'s type are normally capable of being means of realising outcomes of $G$'s type in situations with the salient (to any concerned) features of this situation;
% 
\item no alternative type of action is both 
typically available to agents of this type 
and also 
such that actions of this type would be normally be significantly better* means of realising outcome $G$ in situations with the salient features of this situation;
%
\item the occurrence of outcome $G$ is typically desirable for agents of this type;
%
\end{enumerate}
%
\begin{enumerate}[resume]
\item there is no other outcome, $G'$, 
the occurrence of which would be at least comparably desirable for agents of this type 
and where (2) and (3) both hold of $G'$ and $a$
%
%where
%	the occurrence of $G'$ would be at least comparably desirable for agents of this type 
%%where
%%	$G$'s occurrence is not a means to $G'$'s nor conversely,
%and
%where  
%	(2) and (3) both hold of $G'$ and $a$
\end{enumerate}
%
may jointly constitute defeasible evidence for the conclusion that:
%
\begin{enumerate}[resume]
\item $G$ is a goal to which action $a$ is directed.
\end{enumerate}
%

{
\footnotesize
*An action of type $a'$ is a better means of realising outcome $G$ in a given situation than an action of type $a$ if, for instance, actions of type $a'$ normally involve less effort than actions of type $a$ 
in situations with the salient features of this situation 
and everything else is equal; 
or if, for example, actions of type $a'$ are normally more likely to realise outcome $G$ than actions of type $a$
in situations with the salient features of this situation 
and everything else is equal.
}

%\section{The problem of opaque means}
%Ordinary 3$^{\textrm{\tiny rd}}$ person interpretation may fail when it is not known which outcomes a behaviour is a means of bringing about, especially where a novel tool or communicative device is used.

\section{Your-goal-is-my-goal}
\begin{enumerate}
\item You are willing to engage in some joint action or other with me.
%(for example, because you have made eye contact with me while I was in the middle of attempting to do something).

\item I am not about to change the single goal to which my actions will be directed.

\end{enumerate}
%
Therefore:
%
\begin{enumerate}[resume]
%
\item A goal of your actions will be my goal, the goal I now envisage that my actions will be directed to.
\end{enumerate}
%

\section{Application}

`to understand pointing, the subject needs to understand more than the individual goal-directed behaviour. She needs to understand that ... the other attempts to communicate to her ...  and ... the communicative intention behind the gesture'\citep{Moll:2007gu}

\begin{center}
\includegraphics[width=6cm]{figure_hare_toma_2004_e3.png}
\label{fig:reach_point}

A failed reach (left) and a helpful point (right).\citep%[p.\ 557, figure 4]
	{hare_chimpanzees_2004}
\end{center}

`the adult’s social cues conveyed her communicative intent, which in turn encouraged the child to `see through the sign'.'
\citep%[p.\ 118]
{leekam_adults_2010}

\footnotesize 
\bibliography{$HOME/endnote/phd_biblio}

\end{multicols}

\end{document}

### 2012-09-07 joint action and knowing others minds/conf budapest ceu talk cognitive 2012-09/interacting mindreaders HANDOUT.tex
moved folders

--- 2012-09-05 joint action and knowing others minds/submit philosophical studies  2012-01/second submit/conf budapest ceu talk cognitive 2012-09/interacting mindreaders HANDOUT.tex
@@
 {
-\textbf{Joint Action and the Emergence of Mindreading}
 \Large{Interacting Mindreaders}
@@
 
-<s.butterfill@warwick.ac.uk>
+<butterfillS@ceu.hu>
 
@@
 
-Could an interacting mindreader be in a position to know things which she would be unable to know were she unable to interact with her targets?
+Could interacting mindreaders be in a position to know things which they would be unable to know if they were manifestly passive observers?
+
+\section{Terminology}
+\emph{Mindreading}: `In saying that an individual has a theory of mind, we mean that the individual imputes mental states to himself and to others ...
+the system can be used to make predictions, specifically about ...
+behavior.'\citep{premack_does_1978} %(Premack & Woodruff 1978: 515)
+
+\emph{Sophisticated forms of mindreading} involve imputing propositional attitudes such as belief, desire and intention in the course of constructing or evaluating  reason-giving, causal explanations of action.
+
+%\section{Joint action}
+%Paradigm cases in philosophy include two people 
+%	painting a house together, \citep{Bratman:1992mi}
+%	lifting a heavy sofa together, \citep{Velleman:1997oo} 
+%	preparing a hollandaise sauce together, \citep{Searle:1990em}
+%	going to Chicago together, \citep{Kutz:2000si} 
+%	and walking together. \citep{gilbert_walking_1990}
+%
+%In developmental psychology paradigm cases of joint action include  two people 
+%	tidying up the toys together, \citep{Behne:2005qh}
+%	cooperatively pulling handles in sequence to make a dog-puppet sing, \citep{Brownell:2006gu} 
+%	and bouncing a block on a large trampoline together.\citep{Tomasello:2007gl}
+%
+%Other paradigm cases from research in cognitive psychology include two people
+%	lifting a two-handled basket,  \citep{Knoblich:2008hy}
+%	putting a stick through a ring, \citep{ramenzoni_joint_2011}
+%	and swinging their legs in phase.\citep%[p. 284]
+%	{schmidt_richardons:_2008}
+
+\section{The Problem of Opaque Means}
+Ignorance about to which ends actions are means can be an obstacle to goal ascription.
+
+\section{Distributive Goals}
+A \emph{goal} is an outcome to which actions are, or might be, directed.  A \emph{goal-state} is an intention or other state of an agent linking an action to a goal to which it is directed.
+
+%A \emph{goal-directed joint action} is a joint action which, taken as a whole, is directed to a goal.
+
+The \emph{distributive goal} of two or more actions is G: (a) each action is individually directed to G; and (b) it is possible that: all actions succeed relative to this outcome.
+
+%\textbf{Collective goal}.  The \emph{collective goal} of two or more actions is G:
+%(a) G is a distributive goal of the outcomes;
+%(b) the actions are coordinated; and 
+%(c) coordination of this type would normally  facilitate occurrences of outcomes of G's type
 
@@
 
-%(Cf.\ Southgate et al's principle of efficiency:
-%`goal attribution requires that agents expend the least possible amount of energy within their motor constraints to achieve a certain end.' \citep%[p.\ 1061]
-%{Southgate:2008el})
+(Contrast a principle of efficiency:
+`goal attribution requires that agents expend the least possible amount of energy within their motor constraints to achieve a certain end.' \citep%[p.\ 1061]
+{Southgate:2008el})
 
@@
 
-`the adult’s social cues conveyed her communicative intent, which in turn encouraged the child to `see through the sign'.'
+Leekam et al.: `the adult’s social cues conveyed her communicative intent, which in turn encouraged the child to `see through the sign'.'
 \citep%[p.\ 118]
 {leekam_adults_2010}
+
+Csibra: Early in human development,
+teleological and referential action interpretation 
+`rely on different kinds of action understanding' %\citep[p.\ 456]{Csibra:2003kp}
+and
+involve two distinct `action interpretation systems'   %\citep[p.\ 456]{Csibra:2003kp} 
+which come together later in development.\citep{Csibra:2003kp}
 

### 2012-09-07 joint action and knowing others minds/conf budapest ceu talk cognitive 2012-09/interacting mindreaders ceu talk.tex
moved folders

--- 2012-07-13 joint action and knowing others minds/interacting mindreaders.3.ps.corrections.tex
@@
 
-\def \ititle {Interacting Mindreaders}
-\def \isubtitle {}
+\def \ititle {Interacting Mindreaders:}
+\def \isubtitle {CEU Talk}
 \def \iauthor {Stephen A. Butterfill}
@@
 
-Mindreading is
-the process of identifying thoughts and actions on the basis of bodily movements,
-somewhat as reading is the process of identifying propositions on the basis of inscriptions \citep[p.\ 4]{Apperly:2010kx}.
+Thank you for inviting me.  I’m visiting until Christmas to immerse myself in your work here, with the aim continuing one collaboration and perhaps starting new collaborative work.
+
+My work is mostly on mindreading and joint action.
+Here I'm aiming to bring the two together by discussing mindreading by the agents of a joint action.
+The question I aim to answer is  whether interacting mindreaders could  be in a position to know things which they would be unable to know if they were manifestly passive observers.
+
+But before I get to the question,
+ let me step back to explain the wider issues that motivate it.
+
+\section{slide: Vygotskian Intelligence Hypothesis}
+Moll and Tomasello say ...
+
+Coming at things from a very different starting point, 
+Sebanz and Knoblich make a different but related claim.  
+They say ...
+
+Faced with these bold conjectures, it is natural to ask,
+How could that work?
+How could social interaction drive or ground sophisticated forms of cognition such as those involved in communication by language or sophisticated forms of mindreading?
+
+I think we might be able to make a small step towards understanding these conjectures by considering the role of  social interaction in simple forms of mindreading.
+
+\section{slide: Question}
+So this talk is based on a narrow question.
 Contrast
@@
 Is it possible that the interacting mindreader is in a position to know things which she would be unable to know if she were unable to interact with her targets?
-Our aim in this paper is to argue that the answer is a qualified `yes'.
-
-The question is about the evidential basis of mindreading, 
-not about what sorts of mechanisms are involved.
-While philosophers have engaged with questions about mechanisms (such as whether mindreading involves a process of simulation or of theorizing or some combination of the two),
-comparatively little effort has recently been devoted to issues about what evidence could ground mindreading.
-
-Our question is part of a broader question,
-What is the evidential basis for ascriptions of thought and action and how does the evidence support the ascriptions?
-%\footnote{
-%See e.g.\ \citet[pp.\ 126-7]{Davidson:1973jx}.
-%}
-%To stress that this question is not directly about about \emph{how} anyone actually ascribes thoughts or identifies actions% 
-%% nor directly about the mechanisms, processes or representations involved in mindreading
-%, Davidson 
-%sometimes formulates the question by asking what someone \emph{could} know that \emph{would} put them in a position to identify another's thoughts and actions.%
-%\footnote{
-%See, e.g., \citet[p.\ 126]{Davidson:1973jx}.
-%While what follows draws  on Davidson's questions and insights,
-%our aims here are more modest than his.
-%For here we are concerned only with a  fraction of the problem of ascribing mental states and meanings,
-%and we are not concerned with larger claims about mind, meaning and truth 
-%(on which see 
-%\citet{Davidson:1990du} and 
-%\citet{lepore_donald_2005}).
-%}
-The most sustained attempts to answer this question,
-Davidson's (\citeyear{Davidson:1984wh}; \citeyear{Davidson:1990du}), Lewis' (\citeyear{lewis:1974ri}) and Dennett's (\citeyear{Dennett:1987sf}),
-do not exploit the possibility of interaction.
-The evidence and principles they consider are available to manifestly passive observers.
+
+Most philosophers who have written about the evidential basis of mindreading have assumed that the answer is `no'.
+Of course they would allow that interaction can facilitate mindreading in practice.
+But their theories on the evidential basis of mindreading assume that interacting mindreaders have no advantage \emph{in principle} over mere observers.
 So on their theories,
@@
 in principle the same evidence could be available to each, and each can exploit the routes to knowledge in moving from evidence to ascriptions of thought and action.
-Of course these theories are all compatible with the idea that interaction might be useful for mindreading in practice.  
-But on these theories interaction makes no difference to what can in principle be known.
-In this paper we aim to show
-that mindreaders actually or apparently capable of interacting with their targets are at an advantage not only in practice but also in theory.
+
+In this talk I aim to show that mindreaders actually or apparently capable of interacting with their targets are at an advantage not only in practice but also in theory.
 Their ascriptions could exploit routes to knowledge which would be unavailable if they were entirely passive observers.
 
+\section{slide: Reciprocity}
 Why suppose, in advance of considering the details, that interacting mindreaders might know more?
@@
 
-Even if the hunch turns out to be right, why investigate it?
-One reason is that the investigation will enable us to revise and extend existing accounts of the evidential basis of mindreading  in ways that make them more accurate and comprehensive.
-Another motive concerns the emergence, in evolution or in development, of mindreading.
-Several researchers have offered quite general conjectures about how interaction might explain the emergence of sophisticated forms of cognition.
-One view is that needs to interact with others have driven and shaped some aspects of cognition.%
-\footnote{
-E.g.\ \citet[p.\ 103]{Knoblich:2006bn} suggest that 
-`functions traditionally considered hallmarks of individual cognition originated through the need to interact with others'
-and that
-`perception, action, and cognition are grounded in social interaction.'
-}
-We shall not speak to this view here.
-Another view is that abilities to interact with others
-may have fostered the emergence in evolution---or, on a different view, in development---%
-of sophisticated forms of cognition including social cognition and, in particular, mindreading.%
-\footnote{
-\citet[p.\ 1]{Moll:2007gu}
-argue for the `Vygotskian Intelligence Hypothesis' according to which `the unique aspects of human cognition ... were driven by, or even constituted by, social co-operation.'
-See also
-	\citet{Hughes:2004zj},
-	\citet{Hughes:2006fu},
-	\citet{Tomasello:2007gl} and
-	\citet{tomasello:2008origins}.
-}
-Studying how interaction broadens the evidential basis of mindreading will eventually point to one way of filling in some details on just how interaction might facilitate the emergence of sophisticated forms of mindreading.
-
-Our claim needs to be qualified in several ways.
-First, we should highlight something already explicit but not yet emphasised:
-the claim applies not just to 
- mindreaders who actually could interact with their targets
-but also to
- mindreaders who appear to their targets to be potential interaction partners (even if they are not).
-Trading the risk of a minor misunderstanding for concision we use the term `interacting mindreader' to refer to both groups;
-similarly, references to mindreaders who `only observe' must be understood as excluding those who appear to their targets to be interaction partners.
-Only cases involving the first group---mindreaders who actually could interact---are likely to be of wider interest.
-%Cases involving misleading appearances might be regarded as artefacts of the model.
-Any application of our claim to understanding evolution or development is bound to focus on cases where mindreaders actually could interact with their targets.
-
-A second qualification concerns the scope of mindreading.
+\section{slide: Goal Ascription}
+
 In most discussions of mindreading, the focus is on  ascription of beliefs and other mental states to individuals.
-But in what follows we shall focus on the ascription of goals to actions.
-Some might claim that goal ascription is not mindreading, perhaps because 
-identifying relations between actions and the outcomes to which they are directed
-does not necessarily involve ascribing mental states.
-Our view is consistent with this claim (and with its negation).
-\label{goal_ascription_verify}
-Evidence for ascriptions of belief and of other mental states often includes the occurrence (and non-occurrence) of goal-directed actions.
-Identifying these actions typically requires goal ascription.
-In some cases, then, 
-evidence for the ascription of a goal
-will indirectly support the ascription of a belief or other mental state---%
-the evidence will support  a goal ascription which will in turn support the ascription of a belief (say) to the agent of the action.
-Accordingly, even those who deny that goal ascription is mindreading should agree that sometimes evidence for goal ascription is indirectly evidence for ascription of belief and other mental states.
-So whether or not goal ascription is deemed to be mindreading,
-an individual's access to evidence for mindreading  depends in part on her access to evidence for goal ascription.
-Our plan is to show that interacting mindreaders could exploit routes to knowledge of the goals of others' actions which are not available to mere observers.
-In doing this we will be showing that
-mindreaders poised to interact with others 
-could know things about their minds 
-which they might not otherwise be in a position to know.
-
-%Because the leading theories have supposed that being able to interact makes no difference,
-%they have treated the evidential basis for mindreading as more narrow than it truly is.
-%This matters
-%because broadening the evidential basis will enable us to give a more accurate and comprehensive theory of mindreading.
-%And, as we shall argue, this in turn matters because it may bear on the possibility of explaining the emergence, in evolution or in development, of mindreading.
-
-Developing these ideas requires us to fill in some background on goal ascription and its limits, as well as on notions of goal-directed interaction.
-Readers impatient to get to the central idea might skip to section \vref{sec:your_goal_is_my_goal}.
-
-\section{Goal ascription}
-Purposive action is action directed to the realisation of one or more outcomes.
+But in what follows I want to focus on goal ascription.
+
 Goal ascription is the process of identifying to which outcomes others' purposive actions are directed.
@@
  those outcomes.
-%, among an action's actual and possible outcomes,
-%those to which it is directed.
-%Our immediate question is therefore, 
-%What evidence could support hypotheses about the outcomes to which actions are directed, 
-%and how would the evidence support the hypotheses?
-
-We focus on goal ascription 
-partly because this simplifies our argument,
-but mainly because goal ascription is widely thought to be among the very earliest components of mindreading to emerge (or, if goal ascription is not mindreading, then it is a late precursor).%
-\footnote{
-See, for example,
-\citet{Gergely:1995sq} and 
-\citet{Woodward:1998dm}.
-See also
-\citet[p.\ 111, Box 1]{Baillargeon:gx}
-on two subsystems
-and 
-\citet{Povinelli:2001jf} on `behavioural regularities' whose specification sometimes appears to involve goal-directed action.
-}
-By showing that
-  interacting mindreaders may have access to 
-  evidence for goal ascriptions 
-  which is unavailable to those who merely observe,
-we will eventually be able to indicate ways in which interaction could facilitate the emergence, in evolution or development, of more sophisticated forms of mindreading.%
-\footnote{
-There are various ways of understanding the idea that mindreading may come in several forms, some less conceptually or cognitively demanding than others.
-For a range of views, see
- \citet{Apperly:2009ju},
- \citet{Call:2005qe},
- \citet{Doherty:2006wz},
- \citet{ONeill:2005ff}
- and
- \citet{Wellman:2001if}.
-}
-
-Because goal ascription has received little attention,
-in this section
-we shall 
-briefly review some potential benefits of being able to identify goals
-and then 
-consider what sort of evidence might support goal ascription.
-
-It is a familiar idea that goal ascription enables one to learn from others' successes.
-For example,
-if you know or can guess that another agent's actions are directed to opening a nut,
-you may then be in a position to infer that the unfamiliar pattern of actions she is performing constitute a means to open nuts.
-%\citep{Horner:2005pj}
-%In a different case, knowing the goal of another agent's actions may enable you to discover a new use for a tool.
-A slightly less familiar idea is that goal ascription enables one to learn from others' failures as well as their successes.
-For example, suppose that while you are searching for some peanuts 
-another agent attempts but fails to reach for a closed container.
-In some circumstances,
-if you know that the goal of the agent's action was to obtain the peanuts
-then you now have evidence as to where they might be.\footnote{
-\citet{hare_chimpanzees_2004} exploit this fact in testing chimpanzees' abilities to ascribe goals.
-}
-This is one illustration of how goal ascription could in principle enable us to learn from others' failures 
-(\citealp{Want:2001hp} offer another).
-
-Goal ascription enables one
-to predict and manipulate others' actions.
-If you know that an agent is engaged in a sequence of actions whose eventual goal involves retrieving some object,
-you may be able to predict that the agent will go to where the object is at some point.\footnote{
-For an application see \citet{Hare:2001ph}.
-}
-Equally, in some cases knowing this much about the goal of another's actions many enable you to assist them,
-either 
-by retrieving the object for them 
-or else
-by revealing the object's location to them.%
-\footnote{
-For a paradigm involving the former see \citet{warneken:2007sa};
-on the latter, see \citet{Liszkowski:2008al}.
-}
-
-Goal ascription is also instrumental for ascribing propositional attitudes.
-We have already mentioned 
-   (\vpageref*{goal_ascription_verify})
-that such ascriptions often involve 
-confirming predictions about actions 
-which in turn typically requires
-identifying the goals of those actions.
-In addition,
-knowing which outcomes an action is directed to may constrain hypotheses about what an agent intends 
-as well as
-potentially providing information concerning what the agent knows, believes or desires.
-For example,
-if we know that the goal of an agent's action is to retrieve some peanuts,
-and if we also know where all the peanuts are,
-we may be able to infer that she does not know where the peanuts are,
-or that she falsely believes that some of the peanuts are over there.%
-\footnote{
-\citet{Wimmer:1998kx} exploit this possibility in testing children's abilities to ascribe false beliefs.
-}
-(Of course this can also work the other way:
-information about an agent's beliefs or other mental states may support conclusions about her goals.
-Belief- and goal-ascriptions are mutually constraining.)
-
-Now that we have reviewed some of the benefits goal ascription can bring,
-what evidence could support ascriptions of goals to actions?
-Consider the claim that 
-knowledgeably identifying the goals to which actions are directed
-depends on 
-knowledge of the agents' intentions
+
+Some might claim that goal ascription is not mindreading, perhaps because 
+identifying relations between actions and the outcomes to which they are directed
+does not necessarily involve ascribing mental states.
+I'm happy if you want to define mindreading so as to exclude mere goal ascription.
+
+Still, everyone should agree that mindreading very often depends on goal ascriptions. 
+After all,
+evidence for ascriptions of belief and of other mental states often includes the occurrence (and non-occurrence) of goal-directed actions.
+Identifying these actions typically requires goal ascription.
+In some cases, then, 
+evidence for the ascription of a goal
+will indirectly support the ascription of a belief or other mental state---%
+the evidence will support  a goal ascription which will in turn support the ascription of a belief (say) to the agent of the action.
+Accordingly, even those who deny that goal ascription is mindreading should agree that sometimes evidence for goal ascription is indirectly evidence for ascription of belief and other mental states.
+
+\section{slide: Evidence for goal ascription}
+What evidence could support ascriptions of goals to actions?
+
+Some philosophers seem to assume that goal ascription depends on knowledge of the agents' intentions
 which in turn depends 
@@
 it would make no sense to discuss the evidential basis of goal ascription except in discussing the evidential basis of mindreading more generally.
+
 However,
@@
 }
-And we know of no compelling reason for taking this claim to be true.
-After all,
-it is consistent with rejecting the claim to recognize that
- ascriptions of goals to actions
- constrain, and are constrained by,
- ascriptions of intentions and other mental states.
-The existence of such constraints does not show that one could 
-not know something about the goals of actions 
-while  knowing nothing about the agents' mental states.
-Further,
-there do seem to be situations where knowledge of agents' goals does not require any knowledge of their mental states.
-For instance,
-suppose that two people are sitting opposite each other at a low table
- which is 
-sparsely populated with objects.
-The objects are all out in the open; manifestly, both can clearly see them.
-If one person reaches to grasp one of these objects (the duck, say), 
-must the other ascribe beliefs or other mental states in order to knowledgeably identify the goal of her action as that of grasping the duck?
-On the face of it, she need not.  
-Even if she had no ability to ascribe mental states, it seems she might nevertheless be in a position to identify the goal of the other's action.
-
-While not decisive,
-these considerations are perhaps sufficient to motivate 
-exploring what evidence might support goal ascription.
+
+To show that goal ascription is possible without ascribing mental states, we need to say what evidence someone could have that would enable her to know to which goals an action was directed.
+
 According to 
@@
 %
-Taking this idea as a rough starting point,
-we propose that
-%goal ascription could sometimes involve 
-%identifying which potentially desirable outcomes 
-%an observed action is the best available means to achieving.
-%More carefully, our proposal is that 
-these facts:
+This principle plays two distinct roles.
+One role is mechanistic: this principle forms part of an account of how infants (and others) actually ascribe goals.
+Another role is normative: this principle also identifies grounds on which it would be rational to ascribe a goal.
+I'm concerned only with its normative role.
+
+As Csibra and Gergely formulate it, the principle might seem simple.
+But actually their eloquence is hiding some complexity.
+It is perhaps worth spelling out what might be involved in applying this principle.
+Let me try to spell it out as an inference with premises and a conclusion ...
 %
@@
 
+[\textit{skip paragraph}]
 Why accept this?
@@
 
-Note that we are not suggesting that the above inference captures the only route to knowledge of the goals of actions,
-or that it provides anything like a comprehensive theory of the evidential basis of goal ascription.
-Clearly it does not.
-The inference has only limited applications.
+Of course the inference is not deductive.  It is possible for the premises to be true and the conclusion false.  But this is typically of any inference describing a route to knowledge.
+
+I'm not suggesting, of course, that this inference describes how anyone actually ascribes a goal.
+My concern is with a related, but more narrowly theoretical question: what evidence could support goal ascription.
+
+I also want to concede that this can't be the whole story about goal ascription  
+because this inference has only limited applications.
 For example, it cannot be used when an action is the best available means of achieving two or more comparably desirable outcomes.
-Fortunately the argument that follows does not depend on having a comprehensive theory of evidence for goal ascription.
-The point of this brief discussion was only to 
-consider the sort of evidence that can bear on goal ascription
-and to 
-show that 
-even those incapable of ascribing mental states like belief 
-might nevertheless have evidence sufficient for goal ascription.
-
-Our narrow aim in what follows is to show that some routes to knowledge of the goals of actions 
-are available only to interacting mindreaders
+
+Anyway, here is one route to knowledge of the goals of actions.
+My aim in what follows is to show that there is another route 
+to knowledge of the goals of actions 
+that is available only to interacting mindreaders
 and not to those who merely observe.
-Having in this section introduced goal ascription as a topic,
-the next step is to identify an obstacle to acquiring knowledge of the goals of actions,
-one which we will eventually show can be avoided where mindreaders can (or appear able to) interact with their targets.
+But first let me explain why we might want another route to knowledge of the goals of actions.
+
+To apply Csibra and Gergely's principle in a particular case where you first observe an action and then conclude that its goal is $G$,
+you need to know that the action is a `justifiable action towards' $G$.
+In other words, you need to know that actions of this type are a means to achieving $G$.
+
+Whatever you think of the details of this principle of rational action, 
+it seems plausible that the evidence for goal ascription sometimes includes considerations about which ends actions are means to.
 
@@
 \label{sec:opaque_means}
-
-While we lack a detailed theory of the evidential basis of goal ascription,
-it is certain that the evidence for goal ascription sometimes includes considerations about which ends actions are means to.
-Suppose an observer faces an action but cannot identify ends to which it could be a means.
-This may prevent her from recognizing the action's goal%
-\footnote{
-It is possible that some actions have more than one goal.
-To reduce parenthetical qualifications we shall write as if actions had only one goal.  
-All of our key claims and arguments are consistent with the possibility of actions with more than one goal.
-} 
- by depriving her of evidence.
+This the problem of opaque means:
+When an observer faces an action but cannot identify ends to which it could be a means, she might be unable to recognise the action's goal.
+
+Ignorance about to which ends actions are means can be an obstacle to goal ascription.
+
+Of course the opaque means are not always  an insurmountable obstacle to goal ascription, and
+they may only rarely be a problem for human adults who can rely on culture, convention, or communication.
+Where goal ascribers lack such sophistication, the problem of opaque means is likely to be more pressing.
+
+Some of the most plausibly unique aspects of human cognition depend on our abilities to recognise the goals of novel behaviours involving tools, and of communicative gestures.
+The problem of opaque means is likely to arise in both cases
+if goal ascription is based  entirely on observation 
+(so that the possibility of interaction is ignored)
+and if the goal ascriber lacks sophisticated social skills.
+
 To illustrate, contrast two cases of tool use.
@@
 may be able to identify the goal of the first action only.
-As this illustrates, ignorance about to which ends actions are means can be an obstacle to goal ascription.
-Call this the problem of opaque means.
-
-We are not suggesting that no observer could ever identify the goal of any action she fails to recognise as a means to achieving that goal.
-Of course opaque means are not in every case an insurmountable obstacle to goal ascription, and
-they may only rarely be a problem for human adults with sophisticated social skills.
-Our point is neither novel nor surprising:
-opaque means sometimes  deprive mindreaders of evidence and so prevent goal ascription.
-This is more likely to happen where  goal ascribers lack sophistication in mindreading, communication and culture.
-
-Some of the most plausibly unique aspects of human cognition depend on our abilities to recognise the goals of novel behaviours involving tools, and of communicative gestures.
-The problem of opaque means is likely to arise in both cases
-if goal ascription is based  entirely on observation 
-(so that the possibility of interaction is ignored)
-and if the goal ascriber lacks sophisticated social skills.
-We have just seen an illustration of how the problem of opaque means arises where tools are used to unfamiliar ends.
-Relatedly, it is also likely to arise where actions involve multiple steps that do not form a familiar sequence, can occur in various orders and can be interspersed among other activities;
+
+The problem of opaque means is also likely to arise where actions involve multiple steps that do not form a familiar sequence, can occur in various orders and can be interspersed among other activities;
 as in preparing spirit from grain, for example.
@@
 One theoretically possible explanation of these findings is that the participants could identify to which end a failed reach might be a means, but not to which end a communicative gesture might be a means.%
-\footnote{
-  \label{fn:communicative_intention}
+\footnote{\label{fn:communicative_intention}%
 Hare and Tomasello 
@@
 The problem is potentially a problem for mindreaders given the standard, purely observational theories of mindreading.
-Note that it is not our intention to suggest that the problem of opaque means is a problem \emph{for theories of mindreading};
-what matters for our purposes is only that it is a potential problem \emph{for mindreaders}.
-
-That the problem of opaque means exists
-is the first step in our argument
-that 
-theories of the evidential basis of mindreading based on pure observation 
-are less powerful than theories taking into account the possibility of interaction.
-They are less powerful in this sense: 
-some routes to knowledge of the goals of actions are available only where a theory of the evidential basis of mindreading 
- takes the possibility of interaction into account.
-In what follows we shall explain how being able to interact with another sometimes makes available a route to knowledge of the goals of her actions  which avoids the problem of opaque means.
+
+\textbf{In what follows we shall explain how being able to interact with another sometimes makes available a route to knowledge of the goals of her actions  which avoids the problem of opaque means.}
 This will support our claim that an interacting mindreader might be in a position to know things which she would be unable to know if she were only observing.
@@
 \label{sec:joint_action}
-We aim eventually is to defend this claim:
-there are routes to knowledge of the goals of others' actions
-which are closed to
-mindreaders who merely observe their targets
-but open to 
-mindreaders who are, or appear to be, capable of interacting with their targets.
 As a preliminary to defending this claim,
 we need to specify which types of interaction are relevant.
-That is the aim of this section.
 
@@
 %it is necessary that there be a single outcome (and not merely similar outcomes) to which they are directed.
-
-Note that our actions could have a distributive goal even without us each intending that we do something.
-As another variation on the turnip-pulling story,
-suppose that we each individually intend to 
-	contribute to pulling up the turnip or pull it up.
-Such intentions are agent-neutral in this sense:
-relative to such an intention an agent 
-	could succeed by acting alone 
-and
-  she also could succeed by acting with unspecified other agents.
-Plausibly our having these intentions is sufficient for our actions to have a distributive goal.
-For in virtue of these intentions each of our actions is directed to the turnip's extraction,
-and this is an outcome relative to which it is possible that all of our actions succeed.
 
@@
 if she is not limited to merely observing them,
-how might this enable her to exploit a route to knowledge of the goals of their actions?
-The answer hinges on interactions involving distributive goals
-(as defined \vpageref{df:distributive_goal}).
+how might this enable know things about the goals of their actions?
 
@@
 %This is one illustration of how capacities for joint action, even very simple forms of joint action, might be relevant to explaining the development or evolution of richer forms of cultural cognition.
-
-\section{Observing interaction}
-\label{sec:observing_joint_action}
-We have been assuming that the your-goal-is-my-goal inference is  available to interacting mindreaders only. 
-This is required for our argument that 
-%some routes to knowledge of the goals of actions are available only to interacting mindreaders. 
-an interacting mindreader could be in a position to know things  she would be unable to know were she a mere observer.
-We note, however, that a mere observer can use a related inference.%
-\footnote{
-Here we are  indebted to an anonymous referee.
-} 
-To illustrate, return once more to our example of a novice parent who, struggling to get his twin pram onto a bus, is about to act jointly with a stranger.
-Now consider a third person observing the novice parent and the stranger. 
-Suppose this third person can manifestly only observe what  happens. 
-Even so, she might know that the stranger is about to attempt to engage in joint action with the parent, and she might come to know this in roughly the way that the parent does. 
-%So there is a sense in which the third person, who merely observes, is in  the same position as the novice parent.  
-%For each, knowing that the parent's actions are directed to lifting the pram onto the bus could, given some background conditions, put him or her in a position to know that this will be the goal to which the stranger's actions are directed.
-
-To generalise, the following inference characterises a route to knowledge of others' goals which could be exploited by a mere observer:
-\begin{enumerate}
-\item She is about to attempt to engage in some joint action or other with him.
-
-\item This is the single goal to which his actions will be directed.
-
-\end{enumerate}
-%
-Therefore:
-%
-\begin{enumerate}[resume]
-%
-\item A goal of her actions will be his goal, the goal to which his actions will be directed.
-\end{enumerate}
-%
-Call this inference \emph{her-goal-is-his-goal}.  
-This inference shows that mere observers who understand interaction also have a route to knowledge of others' goals that avoids the problem of opaque means. 
-The significance of interaction for mindreading extends beyond cases in which the mindreader herself interacts.
-
-Is this an objection to our claim that an interacting mindreader could be in a position to know things  she would be unable to know were she a mere observer?
-No, for the above inference differs from the earlier your-goal-is-my-goal inference (\vpageref*{your_goal_is_my_goal}) in requiring knowledge of another's goal. 
-The second premise makes this explicit.
-Where your-goal-is-my-goal depends on a mindreader's knowing to which goal her own actions will be directed, this inference depends on the mindreader's knowing to which goal another's actions will be directed.
-Now there could be---and surely are---cases where a mindreader knows to which goals her own actions are directed even though if another were to act as she is acting, mimicking her movements, she would be unable to recognise the goal to which these actions are directed.
-So there could be situations where an interacting mindreader could exploit your-goal-is-my-goal even though if she were only observing she would be unable to exploit her-goal-is-his-goal.  
-This is sufficient to support our claim that an interacting mindreader could be in a position to know things  she would be unable to know were she only observing.
 
@@
 Our point in mentioning it is merely to illustrate how
- better understanding the roles of interaction in goal ascription and  mindreading more generally
+ \textbf{better understanding the roles of interaction in goal ascription and  mindreading more generally
 might eventually lead to discoveries
@@
 %	encounter anomalies
-which they might not have been able to know 
+which they would not have been able to know 
 %or encounter
-if they were only observing.
+if they were only observing.}
 % matters because it enables us to better understand the evidential basis of mindreading---%

### 2012-09-08 joint action and knowing others minds/conf budapest ceu talk cognitive 2012-09/interacting mindreaders ceu talk.tex
ceu talk

@@
 The two are on a par in this sense:
-in principle the same evidence could be available to each, and each can exploit the routes to knowledge in moving from evidence to ascriptions of thought and action.
+in principle the same evidence could be available to each, and each can exploit the same routes to knowledge in moving from evidence to ascriptions of thought and action.
 
@@
 Second, in some of those cases knowing the premises would put one in a position to know the conclusion.
-We shall consider these points in turn
-
-Is it ever possible to know the premises without first knowing the conclusion?  
-Consider the first premise.  
-Sometimes 
-in the right contexts
-an individual can recognize in another's facial expressions, engaging gestures or synchronized bodily movements that she is about to attempt to engage in joint action with her.
-Exploiting these indicators does not typically depend on knowing the particular contents of any beliefs, desires or goals.  
-Expressions, gestures and movements can naturally indicate imminent jointness in much the way they can also naturally indicate emotions.\footnote{
-Ideas along these lines are suggested by the discussion of {emergent coordination} in \citet{Knoblich:2010fk}.
-}  
-Of course these indicators do not guarantee that others are  about to attempt to engage in joint action.  
-But they are sufficiently reliable to ground knowledge in some cases.  
-The existence of such indicators shows that knowing the first  premise of the above inference does not require already knowing which particular goals the other's actions will be directed to.
-
-Not everything needs to rest on indicators, however.
-It is sometimes possible to know that others are about to attempt to engage in joint action with you even without relying on such indicators.  
-Thanks to widespread dispositions to act jointly, in some situations it is reasonable to take for granted that others will act jointly.  
-For example, this is often so for children struggling with a coat while surrounded by family or familiar adults.  
-And in at least some subcultures people using public transport can reasonably take for granted that, within limits, those around them will act jointly with them when the need is clear.
-Of course dispositions to engage in joint action may vary between cultural groups and situations.
-This may be fatal for the frequent traveller,
-but for others what matters is not whether the dispositions are universal but only that they are sufficiently widespread to be predictable.
-
-Turning to the second point,
-could knowing the premises of the your-goal-is-my-goal inference ever put one in a position to know the conclusion?  
-Of course the inference is not deductive and will only work when certain background conditions are met.  
-These background conditions include the other having largely true beliefs concerning which goal your actions are or will be directed to.
-After all, another agent may attempt to act jointly with you while being entirely mistaken about the goals of your actions.
-Where this happens, the premises of the inference might be true but the conclusion false.
-We shall return to this point later (in section \vref{sec:false_belief}). 
-For now, note that in some situations there is no requirement  to consider  this possibility, as for instance when stereotypes, conventions or simplicity should and do make the goal of your actions obvious to the other agent.
-
-As well as having true beliefs concerning the goal of your actions, the other must be willing to pursue this goal. 
-This is not guaranteed by the truth of the premises of the inference. 
-For all these require, the other might be aware of the goal you envisage but attempting to initiate a joint action directed to another, quite different goal%
-%and would be unwilling to perform actions directed to the goal you envisage, so that the conclusion of the inference is false
-.  
-Perhaps, for example, the other disapproves of your project and is offering to act with you as part of an attempt to divert you.\footnote{
-We are grateful to anonymous referee for emphasising this point.
-}
-Note, however, that it is only reasonable for the other to attempt a joint action with you where you are in a position to know which goal the joint action would be directed to. 
-If no alternative goal ought to be salient to you and the other has done nothing to make an alternative salient, then it is not unreasonable for you to ignore the possibility that the other is unwilling to pursue the goal to which your actions will be directed.
-This might easily be the case where, for instance, you are in a tidy kitchen patiently chopping your way through a large pile of carrots; it is perhaps less likely to be the case in a more cluttered kitchen where you are surrounded by all kinds of potential ingredients in different states of preparation.
-In general, how strongly the truth of the premises of the your-goal-is-my-goal inference supports the truth of the conclusion will depend (among other conditions) on how salient alternative goals to which your actions might be directed ought to be to you.
-
-Given these and perhaps other background conditions,
-that others are about to attempt to engage in joint action with you 
-is sometimes sufficient reason to hold that
-you will end up acting jointly with them
- even where the goal to which your actions will be directed is already fixed.
-
-In short, then, the two requirements for the your-goal-is-my-goal inference to characterise a route to knowledge are met.  
-In some cases it is possible to know the premises without already knowing the conclusion thanks to natural expressions of willingness to engage in joint action.  
-And knowing the premises sometimes puts one in a position to know the conclusion thanks to the fact that, when things are going well, another's willingness to engage in joint action with you can be based on an accurate assessment of the goals of your actions.
-
-In principle,
-exploiting your-goal-is-my-goal does not require actually being in a position to interact with the target of goal ascription.
-It is sufficient (and would be necessary but for some special cases) to be taken by the target to be in a position to interact. 
-Our concern, however,
-is with cases that are likely to be important for understanding
-development or evolution.
-This motivates a focus on goal ascribers of limited sophistication
-who lack both deep insight into others' minds and fully-fledged communicative abilities.
-Such individuals are unlikely to be able to contrive or exploit situations in which they only appear to be in a position to interact with their targets.
-When we come to applications of your-goal-is-my-goal, the relevant cases will be those 
-in which a goal ascriber is manifestly in a position to interact with her target.
-
-We should also acknowledge that 
-mindreaders who only observe can make use of the premise about joint action that your-goal-is-my-goal depends on.  
-Knowing that the target of their mindreading  is about to attempt to engage in joint action with another agent might enable them to identify the goal of their target's actions. 
-We return to this issue in section \vref{sec:observing_joint_action}. 
-For now what matters is just that only interacting mindreaders can exploit your-goal-is-my-goal.
-
-The your-goal-is-my-goal route to knowledge is characterised by an inference.  
-However, exploiting this route to knowledge may not require actually making the inference or knowing the premises.  
-Depending on what knowing requires,
-it may be sufficient to believe the conclusion because one has reliably detected a situation in which the premises of the inference are true;
-it may not be necessary to think of this situation as a situation where the premises are true,
-nor even to be able to think of it in this way.
-Our aim in identifying the your-goal-is-my-goal inference is not to 
-defend a detailed hypothesis about  mechanisms of mindreading.
-Instead our present concerns are limited to a normative question about the evidential basis of mindreading.
-The your-goal-is-my-goal inference matters
-not because it describes exactly how anyone actually assigns goals (maybe it doesn't)
-but because 
-it characterises a route to knowledge
-that is closed to mere observers
-and open to interacting mindreaders.
 

### 2012-10-09 intention motor representation joint action (paper 3)/conf budapest philosophy 2012-11/shared agency and motor representation.tex
shared agency version for budapest

--- 2012-08-27 intention motor representation joint action (paper 3)/conf collective intentionality manchester 2012/motor representation and shared intention.tex
@@
 
-\def \ititle {Motor Representation \\ and Shared Intention}
+\def \ititle {Shared Agency and Motor Representation}
 \def \isubtitle {}
-\def \iauthor {
-\textit{Collective Intentionality}, Manchester, August 2012}
+\def \iauthor {CEU Philosophy, Budapest, 20 November 2012}
 \def \iemail{s.butterfill@warwick.ac.uk}
@@
 
-\input{$HOME/Documents/submissions/preamble_steve_paper2}
+\input{$HOME/Documents/submissions/preamble_steve_paper3}
 %\author{}
@@
 \noindent
-On the assumption that motor representation plays a role in explaining how effective joint action is possible, do we also need motor representation to explain what joint action is?
-Philosophers tend to assume that motor representation is only an enabling condition for joint action and of no direct interest to narrowly philosophical theories of joint action and shared intention.
-In this talk I shall argue that social motor representation and shared intention have distinctive roles in explaining the purposiveness of joint action.
-This gives rise to a challenge.
-On the one hand, effective joint action—imagine two people erecting a tent in a gale together—sometimes requires both shared intentions and social motor representations plus a certain kind of harmony between the two.
-On the other hand, recognizing their distinctive roles precludes the existence of direct inferential links between shared intentions and social motor representations.
-The challenge is to explain how these two kinds of representation could sometimes harmoniously contribute to effective joint action despite the lack of inferential integration.
+Shared agency is paradigmatically involved when two or more people paint a house together, tidy the toys away together, or lift a two-handled basket together.
+To characterise shared agency, some philosophers have appealed to a special kind of intention or structure of intention, knowledge or commitment often called `shared intention'.
+In this paper we argue that fully characterising shared agency may additionally require appeal to  motor representation.
+Shared agency is not only a matter of what we intend: sometimes it also depends on interlocking structures of motor representation.
+This may have consequences for some metaphysical, normative and phenomenological questions about shared agency.
+
 \end{abstract}
 
+%\section{***EVIDENCE***}
+%I now think three bits of evidence are relevant:
+%\begin{enumerate}
+%\item The GROOP effect shows that there are representations which specify each agent's task in relation to the other (so they are not simple representing the outcomes to which each of their actions are directed; they are representing an outcome to which their actions taken together are directed.)
+%\item \citet{kourtis:2012_predictive} shows that motor planning can occur for others' actions when we are engaged in joint action with them.
+%\item Vesper's ESPP paper (on jumping together and imagining jumping together --- jumping is published \citep{vesper:2012_jumping}, imagining jumping is about to be submitted (as of August 2012).
+This shows that individuals are capable of running motor simulations of multiple roughly simultaneous actions.
+%\end{enumerate}
+%
+
 \section{Introduction}
-Which events are joint actions?
-Most philosophers and many (but not all) psychologists appear to assume that this question can be fully answered just in terms of a special kind of intention or structure of intention, knowledge and commitment often called a shared (or `collective') intention.
-On this widely held view, for an event to be a joint action is for it to be appropriately related to a shared intention.
+Shared agency is paradigmatically involved when two (or more) people 
+	paint a house together \citep{Bratman:1992mi}, 
+	lift a heavy sofa together \citep{Velleman:1997oo}, 
+	prepare a hollandaise sauce together \citep{Searle:1990em}, 
+	go to Chicago together \citep{Kutz:2000si}, 
+	and walk together \citep{gilbert_walking_1990}.
+In developmental psychology paradigm cases of shared agency include  two people 
+	tidying up the toys together \citep{Behne:2005qh},
+	cooperatively pulling handles in sequence to make a dog-puppet sing \citep{Brownell:2006gu},
+	bouncing a ball on a large trampoline together \citep{Tomasello:2007gl}.
 
-[\textit{Short version: One consideration in favour of this view is ...
-and jump to the second consideration towards the end of  section \ref{text:second_consideration} on page \vpageref{text:second_consideration}.}]
+A \emph{joint action} is an action or event involving shared agency.
 
-\subsection{What is shared intention?}
-Of course this answer to the question about which events are joint actions assumes that we know what shared or collective intention is.
-But there are many accounts of it.
-Still, I think most people would agree that shared intention stands to joint action approximately as plain vanilla intention stands to ordinary, individual action
-Further,
-on all or most leading accounts of shared intention, each of the following is a necessary condition:
+Shared agency raises a tangle of scientific and philosophical questions.  Psychologically we want to know which mechanisms make it possible to engage in and understand different sorts of shared agency
+\citep{vesper_minimal_2010}.  
+Developmentally we want to know when shared agency emerges, what it presupposes and whether abilities to engage in it somehow facilitate socio-cognitive, pragmatic or symbolic development \citep{Moll:2007gu,Hughes:2004zj,Brownell:2006gu}.
+Phenomenologically we want to characterise what (if anything) is special about experiences of action and agency when shared agency is involved \citep{Pacherie:2010fk}.  
+Metaphysically we want to know what kinds of entities and structures are implied by the existence of shared agency \citep{Gilbert:1992rs,Searle:1994lb}.  
+And normatively we want to know what kinds of commitments (if any) are entailed by shared agency and how these commitments arise \citep{Roth:2004ki}.
+%, plus a formal account of how practical reasoning for joint action differs (if at all) from individual practical reasoning \citep{Sugden:2000mw,Gold:2007zd}
 
-\begin{description}
+To support investigation of these questions it may be useful to have an account of shared agency that gives us a principled way of distinguishing shared from individual agency \citep[compare][]{Bratman:2009lv}.
 
-\item[awareness of joint-ness] Agents acting on a shared intention know that they are not acting individually; they have `a conception of themselves as contributors to a collective end.'\footnote{
-	\citet[p.\ 10]{Kutz:2000si}.
-Compare \citet[p.\ 361]{Roth:2004ki}: `each participant ...
-can answer the question of what he is doing or will be doing by saying for example ``We are walking together'' or ``We will/intend to walk together.'''
-Relatedly, \citet[p. 56]{miller_social_2001} requires that each agent believes her actions are interdependent with the other agent's.
-}
+Where could we start in constructing or evaluating an account of shared agency?
+One possibility is to start with some well-known contrast cases.
+Contrast cases are pairs of events which are similar in terms of the behaviour and coordination they involve but where one exemplifies shared agency  while the other does not.  
+Thus \citet{gilbert_walking_1990} contrasts two people walking together with two people individually walking side by side.  
+The two pairs' movements may be the same and similarly coordinated (to avoid collision), but walking together involves shared agency whereas merely walking side by side is not. 
+Relatedly,  \citet{Searle:1990em}  contrasts a case in which several park visitors simultaneously run to a central shelter in order to perform a dance with another case in which the park visitors run to the central shelter in order to escape a storm.
+The first is a case involves shared agency, the second does not; but the same movements occur in both.
+These sorts of contrast case invite the question, 
+How does shared agency differ from individual but parallel agency? 
+Gilbert’s example shows that the difference can’t just be a matter of coordination, because people who are merely walking alongside each other also need to coordinate their actions in order to avoid colliding.
+And Searle’s example shows that the difference between shared agency and parallel individual agency can’t just be that the resulting actions have a common effect because merely parallel actions can have common effects too.
+So one constraint on any account of shared agency is that it should enable us to make systematic sense of these and other contrast cases.
 
-\item[awareness of others' agency]  When agents act on a shared intention, each is aware of at least one of the others as an intentional agent.\footnote{
-	Compare \citet[p.\ 333]{Bratman:1992mi}: `Cooperation ...
-is cooperation between intentional agents each of whom sees and treats the other as such'.
-See also \citet[p.\ 105]{Searle:1990em}: `The biologically primitive sense of the other person as a candidate for shared intentionality is a necessary condition of all collective behavior'
-}
+\section{Shared Intention}
+On the leading, best developed accounts of shared agency in philosophy and psychology, 
+shared agency is explained  in terms of a special kind of intention or structure of intention, knowledge and commitment often called a shared intention.
+On this widely held view, the difference between our walking together and two strangers merely walking side-by-side is that one event is appropriately related to a shared intention whereas the other is not.
 
-\item[awareness of others' states or commitments] When two agents share an intention that they F, each is aware of, or has individuating beliefs about, some of the other's intentions, beliefs or commitments concerning F.\footnote{
-This condition is necessary for shared intention even on what \citet[p.\ 40]{tuomela_collective_2000} calls `the weakest kind of collective intention'.
-But it may not be necessary if, as \citet{Gold:2007zd} suggest, shared intentions are constitutively intentions formed by a certain kind of reasoning.
-% "if the distinctive feature of collective intentions is to be found in the reasoning by which they were formed, then an analysis that focuses on the intentions themselves will miss the feature that makes collective intentions collective.
-"
-}
+Of course, explaining shared agency in terms of shared intention pushes us back to the question, What is shared intention?
+The term `shared intention' is a term of art.
+On many accounts, shared intentions are neither shared nor intentions.
+Minimally, a shared intention stands to a joint action in roughly the way that an ordinary, individual intention stands to an ordinary, individual action.
+Now an ordinary, individual intention represents an outcome, coordinates an agent's actions, and coordinates the agent's actions in such a way that, normally, this coordination would facilitate the occurrence of the represented outcome.
+Similarly, a shared intention involves one or more representations of an outcome, coordinates the several agents' actions, and coordinates them in such a way that, normally, the coordination would facilitate the occurrence of the represented outcome.
 
-\end{description}
+Beyond this parallel there is much divergence on what shared intention is.
+Some hold that it differs from
+ordinary intention with respect to the attitude involved (\citealp{Searle:1990em}). 
+Others have explored the notion that it differs from ordinary intention with respect to its subject, which is plural \citep{Gilbert:1992rs,helm_plural_2008}, 
+or that it differs from ordinary intention in the way it arises, namely through team reasoning \citep{Gold:2007zd}, 
+or that it involves distinctive obligations or commitments to others (\citealp{Gilbert:1992rs}; \citealp{Roth:2004ki}).
+Opposing all such views, \citet{Bratman:1992mi,Bratman:2009lv} argues that shared intention can be realised by multiple ordinary individual intentions and other attitudes whose contents interlock in a distinctive way.
+
+We can distinguish strong and weak claims linking shared agency to shared intention.
+The weak claim is that shared intention will enable us to understand \emph{some forms} of shared agency.
+This claim is neutral on whether there might be joint actions without shared intention, and on whether ingredients other than shared intention might be needed in giving a full account of shared agency.
+The strong claim is that an account of shared intention gives us the whole story about shared agency, so that no other ingredients are needed.
+Given this claim, for an event to be a joint action is for it to be appropriately related to a shared intention.
+
+Some philosophers and psychologists have endorses the strong claim.  For instance:  
 %
-I think we can go even further and say that, 
-on most accounts,
-where joint action involves shared intention, the agents act in part \emph{because} of their awareness of joint-ness, of others' agency and of others' states or commitments.
+\begin{quote} 
+`I take a collective action to involve a collective intention.'  \citep[p.\ 5]{Gilbert:2006wr}
+\end{quote}
+%
+Further examples could be given  (%
+	\citealp[p.\ 381]{Carpenter:2009wq}; 
+	\citealp[p.\ 369]{Call:2009fk};
+	\citealp{Kutz:2000si}; 
+	\citealp[p.\ 117]{rakoczy_pretend_2006}; 
+	\citealp{Tollefsen:2005vh}%
+	).
+In what follows we remain neutral on the weak claim but argue against the strong claim.
+We argue against the strong claim on the grounds that shared agency is sometimes not only a matter of shared intention but may also involves structures of interlocking motor representations.
+This may have consequences for some of the metaphysical, normative and phenomenological questions about shared agency mentioned above.
 
-What follows assumes that where one or more of these three conditions is not met, there is no shared intention.
+\section{The parallel with ordinary intentions}
+Why might anyone endorse the strong claim that for an event to be a joint action is for it to be appropriately related to a shared intention? 
+Those who endorse it rarely or never provide arguments.
+However, two considerations might appear favour this view.
 
-\subsection{Why shared intention?}
-Why might anyone hold that for an event to be a joint action is for it to be appropriately related to a shared intention?
-Two considerations favour this view.
-
-First, the view might be motivated by appeal to some well-known contrast cases.
-Contrast cases are pairs of events which are similar in terms of the behaviour and coordination they involve but where one is a joint action while the other is not.  
-Thus \citet{gilbert_walking_1990} contrasts two people walking together with two people individually walking side by side.  
-The two pairs' movements may be the same and similarly coordinated (to avoid collision), but walking together is a joint action whereas merely walking side by side is not. 
-Relatedly,  \citet{Searle:1990em}  contrasts a case in which several park visitors simultaneously run to a central shelter in order to perform a dance with another case in which the park visitors run to the central shelter in order to escape a storm.
-The first is a case of joint action, the second is not; but the same movements occur in both.
-These sorts of contrast case invite the question, 
-How do joint actions differ from individual but parallel actions? 
-Gilbert’s example shows that the difference can’t just be a matter of coordination, because people who are merely walking alongside each other also need to coordinate their actions in order to avoid colliding.
-And Searle’s example shows that the difference between joint action and parallel individual action can’t just be that the actions have a common effect because merely parallel actions can have common effects too.
-If someone thought that it is possible to distinguish systematically between the contrast cases by appeal to shared intention (and perhaps only by appeal to shared intention), this could motivate holding that for an event to be a joint action is for it to be appropriately related to a shared intention.
+First, if someone thought that it is possible to distinguish systematically between the contrast cases by appeal to shared intention (and perhaps only by appeal to shared intention), this could motivate holding that for an event to be a joint action is for it to be appropriately related to a shared intention.
+We shall argue, however, that distinguishing systematically between the contrast cases cannot be done entirely in terms of shared intention.
+[*HY: this argument is missing from this draft.]
 
@@
 
-\subsection{My view}
-On the second consideration, 
-we take for granted that there is a parallel between joint action and ordinary, individual action.
-But is the standard view about the relation between action and intention correct?
-
-Not everyone accepts the standard view.
+Concerning this second consideration, 
+we take for granted that the parallel obtains but we do not accept the view about ordinary, individual action.
+Not everyone accepts that for an event to be an action is for it to be appropriately related to an intention.
 Some philosophers allow that there could be agents whose actions are purposive---and, in some cases, even intentional---although  the agents have no intentions at all.
@@
 %}
-
 Others reject the standard view without asserting there can be action without intention.
@@
 If any such view is correct, it is impossible to fully explain what it is for an event to be an action by appeal to intention alone.
-
-Our aim in this talk is to defend a parallel view about joint action. 
-%(Parallel to the latter view, that is: not saying that creatures without any shared intentions are capable of joint actions.)
-We shall consider the possibility that fully explaining which events are joint actions requires appeal to structures of motor representation and not only to shared intention.
+So depending on your views about ordinary, individual intention,
+the second consideration---the idea that the relation of shared intention to joint action is in some ways parallel to the relation between ordinary, individual intention and plain vanilla action---might actually support the view we aim to establish, namely that fully explaining shared agency requires appeal to structures of motor representation and not only to shared intention.
 %***what's the argument for this second claim!?
 
-Please bear in mind that we are not trying to get rid of, or replace, the notion of shared intention.
-%I take for granted that a notion of shared intention is needed for explaining which events are joint actions. 
-%My concern is only with the possibility that additional ingredients are needed as well.
-In fact recognising that motor representation is needed to say which events are joint actions may make it easier to understand what shared intention is.
+\section{Motor Representation}
+Since we aim to argue that a fully adequate account of shared agency must appeal to motor representation, it may be helpful first to provide some background about motor representation.
 
-\section{Intention and Motor Representation}
-[\emph{*Probably skip all of this section in the talk}]
+A \textit{motor} representation is the sort of representation that enables us to reach for, grasp and transfer objects in a coordinated and fluid way.
+We follow several psychologists and philosophers in supposing that motor representations feature in planning and monitoring action \citep[e.g.]{wolpert:1995internal, miall:1996_forward}.
 
-So that you can see where this is going, let me briefly explain what stands behind the view that motor representation is needed in addition to intention in saying which events are actions.
-
-A basic question about ordinary, individual action is:
-What is the relation between a purposive action and the outcome or outcomes to which it is directed?
-Purposive actions typically have many actual outcomes.
-Grabbing little Isabel by the hands I swing her around, causing her to laugh and, simultaneously, breaking a vase.
-Either or both of these might be outcomes to which my action is directed.
-Note also that some or all of the outcomes to which my action was directed might not be among its actual outcomes; after all, actions can fail.
-So among all the actual and possible outcomes of my action, one or some are singled out as specially related to this action.
-The question is what singles out the outcome or outcomes, actual or merely possible, to which a particular purposive is directed.
-
-This question is closely related to a second.
-Ordinary purposive actions are sometimes composed of more than one motor action.  My swinging Isabel around includes my reaching for her wrists, grasping them and then spinning us around.  
-But  my action doesn’t include other things which I might be doing simultaneously, like refusing a cup of tea with my eyes or  trying to determine whether that smell is coming from Isabel’s sister Hannah’s nappy.
-The second question, then, is this: For a particular action directed to a specified goal, what determines which activities%
-\footnote{
-As we use the term `activities',
-all actions are activities 
-and so are  things like reaching and grasping.
-This should make it clear that our position does not depend on whether or not reachings and the like are  actions.
-} 
-comprise the purposive action and which do not?
-
-The standard answer to both questions involves intention.
-An intention represents an outcome, coordinates the one or several activities which comprise the action, and coordinates these activities in a way that would normally facilitate the outcome's occurrence.
-What binds component activities together into larger purposive actions?  
-It is the fact that these actions are all consequences of plans involving a single intention (and are all appropriately related to those plans).
-What singles out an actual or possible outcome as one to which the component activities are collectively directed?  
-It is the fact that this outcome is represented by the intention.
-So the intention is what binds component actions together into purposive actions and links the action taken as a whole to the outcomes to which they are directed.
-
-Motor representations are relevantly similar to intentions,
+Motor representations are in some ways similar to intentions,
 as some have recently argued (\citealp[pp.\ 189-90]{pacherie:2008_action}; \citealp{butterfill:2012_intention}).
-Of course motor representations differ from intentions in some important ways (as these authors note).
-But they are similar in the respects that matter for explaining the purposiveness of action.
-For, like intentions, some motor representations represent outcomes and not merely kinematic features of action.
+First, like intentions, some motor representations represent outcomes and not merely kinematic features of action.
 Like intentions, some motor representations play a role in coordinating multiple  component activities by virtue of their role as elements in hierarchically structured plans.
 And, like intentions, some motor representations coordinate these activities in a way that would normally facilitate the outcome’s occurrence.
-So anyone who accepts the standard story about purposive action and intention
-should also accept that a similar story about purposive action and motor representation.
-Given that the two basic questions about ordinary, individual purposive action can also be answered by appeal to intention,
-they can also be answering by appeal to motor representation.
-This motivates the following view.
-Not all purposive actions are bound together and linked to outcomes by intentions. 
-In some cases what binds together purposive actions and links them to outcomes are motor representations.
-%In some cases, the purposiveness of an action is grounded in a motor representation of an outcome; in other cases it is grounded in an intention.
-And of course in many cases it may be that both intention and motor representation are involved.
 
-This is one reason for supposing that motor representation is not merely an enabling condition for ordinary, individual action but also plays a role in explaining what action is (just as intention does).
-Now let’s return to joint action.
+One major dissimilarity between intentions and motor representations is that motor representations lead a kind of double life.
+For motor representations are involved not only in producing actions but also in observing actions.
+Indeed there seem to be some striking similarities between the sorts of processes and representations usually involved in performing a particular action and those which typically occur when observing someone else perform that action \citep{rizzolatti_functional_2010,rizzolatti_mirrors_2008}.%
+\footnote{ 
+If motor representations occur in action observation, then observing actions might sometimes facilitate performing compatible actions and interfere with performing incompatible actions.
+Both effects do indeed occur, as several studies have shown \citep{brass:2000_compatibility, craighero:2002_hand, kilner:2003_interference, costantini:2012_does}.
+}
+
+[*HY: I'm still thinking about how best to introduce motor representation: any pointers?]
+
+So much for the bare notion of motor representation.
+Why suppose that motor representation might be needed to explain shared agency?
 
 \section{Social motor representation}
-We start from the premise that some joint actions are facilitated by reciprocal, agent-neutral motor representations of outcomes whose obtaining would normally involve action on the part of each agent.
+We start from an empirical premise which is about enabling conditions for shared agency only: some joint actions are facilitated by reciprocal, agent-neutral motor representations of outcomes whose obtaining would normally involve action on the part of each agent.
 
 This needs unpacking. 
-A \textit{motor} representation is the sort of representation that enables us to reach for, grasp and transfer objects in a coordinated and fluid way.
-We follow several psychologists and philosophers in supposing that motor representations feature in planning and monitoring action \citep[e.g.]{wolpert:1995internal, miall:1996_forward}.
-One consequence is that motor representations are not concerned with merely kinematic or dynamic features of actions only.
+As already mentioned, a \textit{motor} representation is the sort of representation that enables us to reach for, grasp and transfer objects in a coordinated and fluid way.
+To repeat, motor representations are not concerned with merely kinematic or dynamic features of actions only.
 Rather, some motor representations represent outcomes, such as the movement of a target object from one place to another.
@@
 It is hardly controversial that reciprocal motor representations exist, for their existence is suggested by by a large body of research on motor cognition in action observation.
-What is much more controversial is that reciprocal motor representation occurs in joint action.
-Some of the research that Gunether Knoblich presented supports the claim that there are reciprocal motor representations in joint action, and he suggests that these enable joint action.%
+It is more controversial that reciprocal motor representation occurs in joint action, but there is some evidence for this claim too.%
 \footnote{
@@
 
-I want to go a tiny step further and suggest that in joint action there are sometimes reciprocal, agent-neutral representations of outcomes whose obtaining would normally involve action on the part of all of the agents.
-So when we act together, some of my motor representations may concern outcomes that are partly but not entirely to be realised by my actions.
+We want to go a tiny step further and suggest that in joint action there are sometimes reciprocal, agent-neutral representations of outcomes whose obtaining would normally involve action on the part of each of the agents.
+So when we act together, some of my motor representations may concern outcomes that are partly but not entirely to be realised by my actions (and likewise for you).
 For example, suppose our task is to move an object from A to B, where you pick it up and pass it to me so that I can then place it.
@@
 We shall use the term \emph{social} motor representation as an abbreviation for the reciprocal, agent-neutral motor representation of outcomes whose obtaining would normally require action on the part of each reciprocating agent.
-(Of course we cannot assume in advance of argument that such reciprocal, agent-neutral motor representations are social in any interesting sense;
-but we are using the term `social' in a non-standard way as an abbreviation, one that reflects our aim.)
+(This is probably a bad term to pick; we are still trying to come up with a better one. 
+Certainly we cannot assume in advance of argument that such reciprocal, agent-neutral motor representations are social in any interesting sense;
+for now we are using the term `social' in a non-standard way as an abbreviation, one that reflects our aim.)
 
-What follows is speculative philosophy: we take for granted that sometimes social motor representation facilitates joint action and ask whether this conjecture bears on our question about which events are joint actions.
+What follows is speculative philosophy: we take for granted that  social motor representation sometimes facilitates joint action and ask whether this conjecture bears on our question about shared agency.
+As a first step, we shall ask how social motor representation might facilitate joint action.
+Note that answering this question does not directly commit us to any view about shared agency.
+After all, social motor representation might be an enabling condition for some instances of shared agency but completely unnecessary for explaining what shared agency is.
+
+[*HY: I'm planning a short section on evidence for social motor representation.  But as I'm going beyond the evidence anyway, I don't plan to include it in presenting this paper to philosophers.]
 
@@
 Action-relevant details are progressively filled in by representations at lower stages of the hierarchy. 
-Now in the action we are considering there is a need, even for the single agent, to coordinate the exchange between the two hands.
+Now in the action we are considering there is a need, even for the single agent, to coordinate the exchange between her two hands.
 How is this achieved? 
@@
 They also face a similar coordination problem, requiring a precisely timed swap from one hand to another.
-Now suppose, 
-inspired by Koutis et al’s and others' findings,
-that the same planning is involved in the individual case (where one agent performs the whole action) and in the joint action case (where the action is distributed between two agents).
+Now suppose, that the same planning is involved in the individual case (where one agent performs the whole action) and in the joint action case (where the action is distributed between two agents).
 The planning is the same almost up to the actual muscle contractions.
@@
 and 
-(iii) each  agent's plans for their own action are constrained by their plans for the other agent's action.
+(iii) each  agent's plans for their own action are constrained by the plans for the other agent's action.
 
@@
 \section{Grounding the purposiveness of joint action}
-So far we have only been considering a possible role for social motor representation in facilitating joint action.  
-Our primary concern, though, is with what joint action is.
+So far we have only been considering a possible role for social motor representation in enabling joint action.  
+How does any of this bear on our main question about shared agency?
 The details of how 
 social motor representation enables joint action
- already give us grounds for holding that motor representation has a role to play in explaining which events are joint actions.
-To see why,
-let’s go back to individual action for a moment again.
+ already give us grounds for holding that motor representation has a role to play in explaining shared agency.
 
-[***moved to earlier]
-
-The same two questions we asked about ordinary, individual action also arise for joint action.
+Here are two basic question about  joint action are.
 %What is the relation between a purposive joint action and the outcome or outcomes to which it is directed?
@@
 If we appeal to a notion of shared intention,
-we can answer these questions about joint action in a way that is superficially similar to way we answered the parallel questions about ordinary, individual action.
+we can answer these questions about joint action.
 A shared intention is what relates purposive joint actions to the outcomes to which they are directed.
@@
 the purposiveness of a joint action can be grounded not only in shared intention, but also in social motor representation.
+This is why we suppose that an account of shared agency must appeal not only to shared intention but also to social motor representation.
 
-\section{Social motor representation: like shared intention}
+\section{How social motor representation resembles shared intention}
 It may be helpful to compare and contrast the notion of social motor representation with a notion of shared intention. 
@@
 
-If, as we have just argued, social motor representations play a role analogous to the structure of intentions and knowledge which Bratman identifies as sufficient for shared intention, then this is a  (non-decisive)  reason to think that motor representations can ground the purposiveness of a joint action.
+If, as we have just argued, social motor representations play a role analogous to the structure of intentions and knowledge which Bratman identifies as sufficient for shared intention, then this is a  (non-decisive) reason to think that motor representation is also needed in characterising shared agency.
 
@@
 
+\section{Are social motor representations shared intentions?}
+
+We have been arguing that an account of shared agency cannot appeal to shared intention only  but must also appeal to  social motor representation (and perhaps to other ingredients besides).
+Our argument rests on the premise that social motor representations are not shared intentions.
+But since we have just bee pointing to broad similarities between shared intention and social motor representation in that both play a role in coordinating agents’ actions by virtue of representing outcomes,
+it may be tempting to suppose that some social motor representations are  shared intentions.
+
+This issue might easily seem narrowly conceptual or terminological.  
+At the end of the day it doesn’t much matter if we want to call some motor representations ‘shared intentions’.  
+After all, as already noted, on some accounts shared intentions are neither shared nor intentions so we would hardly be doing more violence to the term than is already being done.
+
+However exactly one decides to use the term `shared intention',  at least three substantive issues remain. 
+The first concerns conceptual demands.
+Whereas having a shared intention arguably demands an ability to represent others' intentions \citep[pp.\ *][]{Butterfill:2011fk},
+social motor representation imposes no such demands.
+The second concerns planning.
+Whereas shared intentions are elements in long-term plans and function in part to enable agents to coordinate their plans, social motor representation is incapable of playing this role.
+A third, and related substantive issue is that social motor representations are structures of representations with a non-propositional format and so cannot be inferentially integrated with ordinary intentions and knowledge \citep[pp.\ *][]{butterfill:2012_intention}, whereas shared intentions can.
+
+While prefer (for narrowly terminological reasons) to state our claim by saying that explaining shared agency requires ingredients other than shared intention,
+the claim could alternatively be formulated by saying that  explaining shared agency requires importantly different (in the ways described) varieties of shared intention.
+
 \section{Conclusion}
-Our opening question was, Which events are joint actions?
-The standard view is that for an event to be a joint action it must be appropriately related to a shared intention.
-We are in the process of arguing that this is not the whole truth about joint action,
-and that some events are joint actions by virtue of being appropriately related to a structure of motor representations we call social motor representation.
+We have been considering how to provide an account of shared agency that might contribute to investigating a tangle philosophical and scientific questions.
+
+Whereas some have claimed that shared agency can be fully explained in terms of a notion of shared intention, we have argued that some events are joint actions by virtue of being appropriately related to a structure of motor representations we call social motor representation.
 We don’t mean to suggest that all joint actions involve social motor representation.
-The view we are aiming to establish is rather this: Some joint actions involve social motor representation; and there could be joint actions which involve social motor representation but not shared intention.
+The view we are aiming to establish is rather this: Some events are joint actions in virtue of being appropriately related to social motor representations which bind their components together and ensure that there is a single outcome to which these components are collectively directed.
 This is why 
@@
 %This is why I think that 
-fully understanding what joint action is requires understanding the coordinating role of social motor representation and not only understanding shared intention.
+understanding shared agency requires understanding not only shared intention but also
+ the coordinating role of social motor representation.
+
+None of this is to deny that shared intention is among the ingredients needed to characterise shared agency.
+Indeed, it may be that the notion of social motor representation has a role to play in explaining what shared intention is.
+In constructing realisers of shared intention from ordinary individual intention, we need intentions \emph{that we $J$}.
+As has been much discussed \citep[e.g.][]{petersson_collectivity_2007}, the contents of these intentions cannot all refer to actions involving shared intentions.
+For this reason Michael Bratman suggests that things we intend are  cooperatively neutral activities. %*ref
+It is then necessary to add further intentions in order to transform cooperatively neutral activities into joint actions.
+But it also seems possible that in some cases, what we intend when we intend that we $J$ is not a cooperatively neutral activity but instead a joint action of the sort which involves social motor representation.
+
+So perhaps harmony between shared intention and social motor representation is sometimes achieved in this way: what we intend when we share an intention is the sort of joint action that involves social motor representation.
+
+In conclusion, two things.
+First, some events are joint actions in virtue of being appropriately related to social motor representations which bind their components together and ensure that there is a single outcome to which these components are collectively directed.
+Second, and much more tentatively, in some cases social motor representation may be among the ingredients that realise a shared intention.
 

### 2012-10-09 intention motor representation goal ascription with corrado sinigaglia (paper 2)/Motor Representation in Goal Ascription.tex
tidying up

@@
 
-Motor representation sometimes occurs in action observation and enables judgements about the goals to which actions are directed (see section \vref{sec:evidence} and \citealp{rizzolatti_functional_2010} for a review).
+Motor representation sometimes occurs in action observation and facilitates judgements about the goals to which actions are directed (see section \vref{sec:evidence} and \citealp{rizzolatti_functional_2010} for a review).
 Suppose you observe someone's hand moving towards a pair of scissors. 
@@
 
-\section{Motor representation enables judgements about the goals to which actions are directed}
+[*rephrase (also introduction): there is evidence that is naturally interpreted as showing that motor representation facilitates judgements about the goals to which actions are directed.
+But there is also an obstacle to accepting this conclusion: it is unclear how this could occur.
+In this paper we aim to remove the obstacle by explaining how motor representation could facilitate judgements about the goals of actions.]
+
+\section{Does motor representation sometimes facilitate action judgements?}
 \label{sec:evidence}
+
 Before launching into the main task for this paper we shall briefly review evidence for our opening premise.
-Why think that motor representation enables judgements about the goals to which actions are directed?
+Why suppose that motor representation sometimes facilitates judgements about the goals to which actions are directed?
 
-Some evidence for this claim arises from studies of deficits.
-Serino and colleagues (\citeyear{serino:2009_lesions_}) compared the performances of control and hemiplegic subjects who were asked to identify actions from point-light displays.
+As a preliminary, note that it is now well established that motor processes and representations are involved not only in producing actions but also in observing actions.
+Indeed there seem to be some striking similarities between the sorts of processes and representations usually involved in performing a particular action and those which typically occur when observing someone else perform that action \citep{rizzolatti_functional_2010,rizzolatti_mirrors_2008}.
+
+If motor representations occur in action observation, then observing actions might sometimes facilitate performing compatible actions and interfere with performing incompatible actions.
+Both effects do indeed occur, as several studies have shown \citep{brass:2000_compatibility, craighero:2002_hand, kilner:2003_interference, costantini:2012_does}.
+To give just one example, \citet{kilner:2003_interference} asked subjects to move their arms horizontally while they observed another person moving their arms either horizontally (that is, congruently) or vertically (that is, incongruently).
+They found that subjects' movements deviated significantly more when observing incongruent actions, and much as they would deviate if the subjects had been required to move their other arm incongruently.
+The effect cannot be explained as a consequence of observing mere movements because the same effect was not found when subjects observed a robot moving its arm.
+
+Of course the fact that motor representation occurs in action observation does not show, further, that motor representation actually facilitates judgements about the goals to which actions are directed.
+
+What would count as evidence for this further claim that motor representation can facilitate goal ascription?
+One possible source of evidence involves motor expertise.
+If manipulating subjects' motor expertise can selectively affect their abilities to identify the goals of actions, this would be evidence that motor representation sometimes facilitates goal ascription.
+Accordingly, \citet{casile:2006_nonvisual} asked subjects to make judgements about observed actions, including were some that are typically difficult to perform without training.
+These actions were presented to subjects as point-light stimuli to ensure that only visual information about the actors' joint displacements was available.%
+\footnote{
+Readers who unfamiliar with point-light stimuli may view examples at http://www.biomotionlab.ca/Demos/BMLwalker.html.
+The technique was introduced by \citet{johansson:1973_visual}.
+}
+They compared performance before and after subjects were trained to perform those actions themselves.
+They found that subjects' ability to accurately judge which action was being performed improved with the training.
+This is plausibly a consequence of increased motor expertise rather than any form of perceptual learning because the subjects were blindfolded during training (and the point-light stimuli on which their judgements were based are, of course, entirely visual).
+A related way to measure the effects of motor representation on judgements is to temporarily lesion part of the motor cortex using repetitive transcranial magnetic stimulation  \citealp{urgesi:2007_representation, moro:2008_neural}.
+\citet{urgesi:2007_representation} measured how long subjects took to make judgements about the type of action they were observing.
+They found that a temporary lesion to the premotor cortex, but not a temporary lesion to another brain area, increased the time taken to make action judgements.
+And the effect of lesioning the premotor cortex was specific to action judgements: subjects' abilities to make observational judgements identifying body parts was unaffected.
+Putting these studies together, we have evidence that enhancing motor representation (through training) sometimes facilitates judgements about the goals of actions, while destroying motor representation even temporarily can impair such judgements.
+
+Further evidence comes from  studies of neurological deficits.
+If deficits in abilities to perform certain actions specifically impair abilities to make observational judgements about actions of that type, this would also indicate that motor representation can facilitate action judgements.
+\citet{serino:2009_lesions_} compared the performances of control and hemiplegic subjects who were asked to identify actions.
 They found that ability to perform an observed action was correlated with ability to recognise that action.
 In particular, hemiplegic subjects were less accurate in identifying actions performed with limbs on the hemiplegic side of their bodies than actions performed with limbs on the unaffected side.
-Other evidence for a role for motor representation in observation-based judgements of action comes from research on apraxia.
+By contrast, no such pattern was found for control subjects, which included both a group of healthy subjects and also a group of brain-damaged patients with no motor deficit.
+(Including the latter group makes it possible to rule out the possibility that the findings were due to a general deficit in higher-order visual processing).
+Other evidence for a role for motor representation in judgements of action comes from research on apraxia.
 In one study subjects were asked to identify actions such as cutting paper and drinking with a straw on the basis of the sounds these actions produced.
 Subjects with limb apraxia showed an impairment in recognising hand-related actions (such as cutting paper) whereas subjects with buccofacial apraxia were impaired in recognising mouth-related actions (such as drinking); but no subjects showed a general impairment in recognising sounds and their significance \citep{pazzaglia:2008_sound_}.
-This link between motor deficits and action recognition indicates that observation-based action judgements sometimes depend on motor representation.
+These links between motor deficits and action judgements provide further evidence that motor representation sometimes facilitates action judgements.%
+\footnote{
+In addition to facilitating judgements about the goals of observed actions, motor representations and processes may also sometimes facilitate judgements about the possibility of performing an action \citep{grosjean:2007_fitts_law, eskenazi:2009_role}.
+}
 
-***mention other sources of evidence
+Motor representation may not always be necessary for observational judgements about the goals of actions.
+Indeed, as \citet{mahon:2008_action} notes, some studies suggest that apraxic subjects can recognise actions they cannot perform \citep[see also][]{hickok:2009_eight}.
+This is consistent with our premise that motor representation sometimes facilitates judgements about actions.
+Even if this facilitation were an extremely rare phenomenon, the question of how it is possible would still arise.
 
-see `How your hands drive my eyes' (but note that our focus is effects \emph{on judgement}):
-\begin{quote}
- `PMv is critically involved in outcome prediction (for a review see Schubotz, 2007).
-For instance, a fMRI study (Schubotz \& von Cramon, 2004) and lesion (Schubotz, Sakreida, Tittgemeyer, \& von Cramon, 2004) data showed that predictions of outcomes even more abstract than bodily movements recruit the premotor cortex"
-There is evidence that action observation and action execution selectively interact with each other.
-Brass et al.
-(Brass, Bekkering, \& Prinz, 2001) instructed participants to execute finger movements in response to observed compatible and incompatible finger movements, and found a reaction time advantage for compatible trials.
-Similarly, Kilner et al.
-(Kilner, Paulignan, \& Blakemore, 2003) showed an increased variance in arm movement execution when observing a human model executing a qualitatively different, rather than the same arm movement.
-This interference effect has been argued to be due to the simultaneous recruitment of the motor representation of both the observed and the executed movements
-There is evidence that action observation may impact on action execution.
-In particular, people have been shown to be faster in performing fingers movements while observing compatible rather than incompatible finger movements (Brass, et al., 2001).
-In the same vein, grasping movements are faster when primed by a picture of a compatible rather than incompatible grip postures (Craighero, Bello, Fadiga, \& Rizzolatti, 2002; Craighero, Fadiga, Rizzolatti, \& Umilta, 1999).
-
-`Our findings show that this effect may be related also to the motor representation of action outcome as measured by the gaze behaviour of the observer.
-Target-specific proactive eye movements have been demonstrated to be driven by the motor representation of an action outcome during both the execution and the observation of a hand action such as a grasping action (Ambrosini, Costantini, et al., 2011; Ambrosini, Sinigaglia, \& Costantini, 2011; Falck-Ytter, 2012; Falck-Ytter, Gredeback, \& von Hofsten, 2006; Flanagan \& Johansson, 2003; Rotman, et al., 2006).
-The present study indicates that the proactivity of the observer's eye movements significantly decreases when the motor representation of the observed action outcome turns out to be incompatible with the motor representation of the action outcome to be executed.'
-\end{quote}
-
-\section{Content}
+\section{What can motor representations represent?}
 \label{sec:content}
 
-Our project is to understand how judgements about the goals of actions  could depend on motor representations.
-One obstacle to understanding this is an apparent mismatch between what goal ascription requires and what motor representations can capture.
-To illustrate, let us return to our opening example.
-Observing someone's hand moving towards a pair of scissors you judge that she will shortly perform an action directed to cutting something.
-In making this judgement you are representing an outcome and linking it to an action.
-How could motor representation enable you to make such a judgement?
+We have just reviewed evidence that judgements about the goals of actions  are sometimes facilitated by motor representation.
+Our project is to understand how this could happen.
+A first step is to understand how the contents of the judgements compare with the contents of motor representations.
+Where judgements about the goals of actions depend on motor representations, could those motor representations represent outcomes to which the actions are directed?
+In this section we argue for a positive answer on the grounds that some motor representations do represent outcomes and that there are processes which ensure that, in observing an action, the outcomes represented motorically are normally outcomes to which the observed action is directed.
+This will allow us to conclude that motor representations can facilitate judgements about goals in part because some motor representations represent outcomes which are among the goals of the actions.
 
-[***TODO: write above in terms of Question One (rather than obstacle).]
+We take for granted that action observation sometimes involves motor representations of merely kinematic features of actions.
+For example, when observing someone about to press a button, it may be useful to be able to identify with some degree of precision the trajectory her finger will follow and how long it will take to reach the button.
+In some cases observers represent both features of such actions motorically (*ref).
+But our concern is with the further claim that some motor representations represent outcomes.
 
-\subsection{Some motor representations represent outcomes}
-If all motor representations were about merely kinematic or dynamic features of action only, their role in facilitating judgements about outcomes would be mysterious.
-However some motor representations represent outcomes to which actions are directed, as we shall now argue.
-For instance, observing someone's hand moving towards a pair of scissors can involve a motor representation of an outcome such as the cutting of something and not only motor representations of individual joint displacements that might bring about this outcome.
-Motor representations can facilitate goal ascriptions in part because some motor representations represent the outcomes which are the goals to be ascribed.
-
-But how can we distinguish a representation of an outcome from a representation of merely kinematic features of actions?
+How can we distinguish a representation of an outcome from a representation of merely kinematic features of actions?
 This is possible because a single outcome can be realised by actions with quite different kinematic features and, conversely, actions with arbitrarily similar kinematic features can realise different outcomes in different contexts.
 To illustrate, consider this outcome: the grasping of a particular pen.
-This could be realised by any of three types of action which vary kinematically: a hand action, a foot action or a tool-using action.
-If we had marks we could use to identify motor representations without already knowing their contents, and if we had evidence that these marks were constant across all three types of action, then we could infer that some motor representations capture something more abstract than the kinematic features particular to the hand, foot or tool-using action.
-But we do have well established marks of motor representation, for motor representation can be inferred from patterns of neuronal discharge, from motor-evoked potentials, from where blood flows in the brain, from behavioural performance profiles and in other ways.
-And we do have evidence that such marks are constant across all three actions which realise the grasping outcome \citep{rizzolatti:1988_functional, Rizzolatti:2001ug, cattaneo:2010_state-dependent}.
+This could be realised by any of at least three types of action which vary kinematically: a hand action, a foot action or a tool-using action.
+If we had marks we could use to identify motor representations independently of knowing their contents, and if we had evidence that these marks were constant across instances of all three types of action, then we could infer that some motor representations capture something more abstract than the kinematic features particular to the hand, foot or tool-using action.
+But we do have well established marks of motor representation, for motor representation can be inferred from patterns of neuronal discharge, from motor-evoked potentials, from where blood flows in motor areas of the brain, from behavioural performance profiles and in other ways besides.
+And we do have evidence that such marks are constant across instances of all three types of action which realise the grasping outcome \citep{rizzolatti:1988_functional, Rizzolatti:2001ug, cattaneo:2010_state-dependent}.
 So we can infer that some motor representations do not represent merely kinematic features of actions.
 To infer, further, that some motor representations represent outcomes we need to consider what happens where, conversely, we hold kinematic features constant while varying to which outcome an action is directed.
-Take an action which realises the grasping of a particular pen and compare it with a second action which is as similar as possible to the first with respect to its kinematic features but differs with respect to which outcome it is directed to because the pen is manifestly absent (so the action is not plausibly directed to grasping that pen) or because the pen is manifestly too large or too small to grasp.
+Take an action which realises the grasping of a particular pen and compare it with a second action which is as similar as possible to the first with respect to its kinematic features but differs with respect to which outcome it is directed to because the pen is manifestly absent (so the action is not plausibly directed to grasping anything) or because the pen is manifestly too large or too small to grasp.
 Some marks of motor representations distinguish these actions \citep{Umilta:2001zr, villiger:2010_activity, koch:2010_resonance}.
@@
 \footnote{
-\citet{pacherie:2008_action} and \citet{butterfill:2012_intention} also argue for this conclusion; the latter further support this claim by appeal to the functions of motor processes in planning and monitoring action.
+We offer only a brief argument for this conclusion here because \citet{pacherie:2008_action} and \citet{butterfill:2012_intention} also argue for it; the latter further support this conclusion by appeal to the functions of motor processes in planning and monitoring action.
 }
 
+***HERE
+
+So far we have been arguing that some motor representations are about outcomes by contrasting this claim with the hypothesis that motor all representations concern merely kinematic features of action.
+[*make more vivid!]  This is not quite enough for a convincing argument because some people have claimed that some motor representations in some way concern sensory consequences of actions.
+To illustrate, grasping a small ball between your finger and thumb will have predictable tactile and visual consequences, and it is possible (at least in principle) that some motor representations involved in grasping the ball might somehow concern these sensory consequences.
+Note that the outcomes to which actions are directed are often distinct from any of the action's sensory consequences.
+Take grasping a ball again: although the outcome---the grasping of the ball---reliably has sensory consequences, these are neither necessary nor sufficient for the outcome to occur.  
+They are not necessary because the grasping could also be performed with a foot or a tool, which could substantially alter the sensory consequences of the action; and because grasping could also in principle occur in complete darkness with numb (or even deafferented) hands.
+The sensory consequences are also not sufficient because in principle the pattern of movements of finger and thumb in relation to the ball could occur without there being any event directed to the grasping of an object---such movements might in principle occur as part of an exercise in moving one's fingers in which the ball is an unwanted distraction rather than a target.
+So our claim that motor representations represent outcomes is conceptually quite distinct from the claim that some motor representations in some way concern sensory consequences of actions: in principle, both claims might be true, or one might be true while the other is false.
+
+Suppose, at least for the sake of argument, that some motor representations do indeed somehow concern sensory consequences of actions. 
+Does we still have grounds for holding that there are motor representations of outcomes?
+Yes.
+We noted just above that it is possible to vary the sensory consequences of an action while holding one of its outcomes constant.
+To return to the example, grasping a ball can be done with a hand, a foot or a tool.
+Using different effectors will alter how the action looks (its visual consequences) but not the goal to which it is directed. 
+So the fact that observing instances of all three types of action can involve a common motor representation shows that not all motor representations can concern sensory consequences and merely kinematic features only.
+
+***The claim that some motor representations represent outcomes raises two further questions.
+First, which outcomes can be represented motorically?
+And, second, what (if anything) ensures that an outcome represented motorically in action observation is an outcome to which the observed action is directed?
+
 Actions are typically directed to many outcomes.
-For instance, an action might be directed to grasping the scissors, to cutting a newspaper and to removing a particularly useful article to file later.
+For instance, a single action might be directed to grasping the scissors, to cutting a newspaper and to removing a particularly useful article to file later.
 In reaching for the scissors the agent is already performing actions directed to all of these outcomes.
@@
 Not all outcomes can be represented motorically, of course.
-For instance, the difference between cutting out an article and cutting out a picture from newspapers is not as such a difference that need be captured by any motor representations.
+For instance, the difference between cutting out an article and cutting out a picture from a newspaper is not a difference that need be captured by any motor representations.
 One idea about which outcomes can be represented motorically starts from the observation that outcomes can be partially ordered by the means-end relation.
@@
 
-\subsection{How to capture the directedness of actions to outcomes}
+[*So the plan is to end this section after arguing that mr don't represent directedness, and then have a new section break on 'Capturing without representing'.
+The Q.
+for that section might be: What sort of process links the outcome represented to the observed action and ensures that the outcome represented is one to which the action is directed?]
+
+\subsection{Directedness}
 Our overall question is how motor representation could enable goal ascription judgements.
 The fact that motor representations can represent outcomes is a step towards answering this question.
-However we are not yet in a position to answer the question.
+However we are not yet in a position to answer it.
 To see why, consider that someone might (at least in principle) observe an action and represent an outcome without yet relating the outcome to the action.
@@
 For instance, you might know that a friend intends to break an egg and see the friend handling an egg which breaks without yet knowing whether this breaking was appropriately related to that intention.
-To represent an intention and an action does not necessarily involve representing any relation between the action and the outcome.
+To represent an intention and an action does not necessarily involve representing any relation between the action and the intention.
 
@@
 
-Anosognosia and deafference comprise a useful pair of cases: in anosognosia we have preserved perceptual senses and an absence of actual bodily movement, whereas in deafference we have damaged senses and actual bodily movement.
+Anosognosia and deafference comprise a useful pair of cases: some anosognosics we have preserved perceptual senses and an absence of actual bodily movement, whereas individuals suffering deafference have damaged senses and actual bodily movement.
 But in both cases there are experiences revelatory of action.

### 2012-10-09 intention motor representation joint action (paper 3)/conf collective intentionality manchester 2012/motor representation and shared intention.tex
tidying up

@@
 \end{abstract}
+
+\section{***REFRAME***}
+Shouldn't start with the question, Which events are joint actions?
+Instead: either it's about making sense of purposive joint action, or else it's the question about the relation between purposive joint actions and the goals to which they are directed.
+(Was just trying to make it sound more important by asking Which events are joint actions.  But I don't think this question bears any weight.)
+
+\section{***EVIDENCE***}
+I now think three bits of evidence are relevant:
+\begin{enumerate}
+\item The GROOP effect shows that there are representations which specify each agent's task in relation to the other (so they are not simple representing the outcomes to which each of their actions are directed; they are representing an outcome to which their actions taken together are directed.)
+\item \citet{kourtis:2012_predictive} shows that motor planning can occur for others' actions when we are engaged in joint action with them.
+\item Vesper's ESPP paper (on jumping together and imagining jumping together --- jumping is published \citep{vesper:2012_jumping}, imagining jumping is about to be submitted (as of August 2012).
+This shows that individuals are capable of running motor simulations of multiple roughly simultaneous actions.
+\end{enumerate}
 
@@
 \subsection{My view}
-On the second consideration, 
-we take for granted that there is a parallel between joint action and ordinary, individual action.
-But is the standard view about the relation between action and intention correct?
+Let us focus on the second consideration, 
+which hinges on  a parallel between joint action and ordinary, individual action.
+We take for granted that the parallel obtains but we do not accept the view about ordinary, individual action.
 
-Not everyone accepts the standard view.
+Not everyone accepts that for an event to be an action is for it to be appropriately related to an intention.
 Some philosophers allow that there could be agents whose actions are purposive---and, in some cases, even intentional---although  the agents have no intentions at all.
@@
 \section{Intention and Motor Representation}
+\label{sec:individual_action_motor_representation}
 [\emph{*Probably skip all of this section in the talk}]
@@
 Motor representations are relevantly similar to intentions,
-as some have recently argued (\citealp[pp.\ 189-90]{pacherie:2008_action}; \citealp{butterfill:2012_intention}).
+as some have recently argued (\citealp[pp.\ 189-90]{pacherie:2008_action}; \citealp{butterfill:2012_intention}).%
+\footnote{
 Of course motor representations differ from intentions in some important ways (as these authors note).
 But they are similar in the respects that matter for explaining the purposiveness of action.
-For, like intentions, some motor representations represent outcomes and not merely kinematic features of action.
+}
+A \textit{motor} representation is the sort of representation that enables us to reach for, grasp and transfer objects in a coordinated and fluid way.
+We follow several psychologists and philosophers in supposing that motor representations feature in planning and monitoring action \citep[e.g.]{wolpert:1995internal, miall:1996_forward}.
+
+How do motor representations resemble intentions?
+First, like intentions, some motor representations represent outcomes and not merely kinematic features of action.
 Like intentions, some motor representations play a role in coordinating multiple  component activities by virtue of their role as elements in hierarchically structured plans.
@@
 
-This is one reason for supposing that motor representation is not merely an enabling condition for ordinary, individual action but also plays a role in explaining what action is (just as intention does).
-Now let’s return to joint action.
+This is one reason for supposing that motor representation is not merely an enabling condition for ordinary, individual action but also plays a role in explaining which events are actions (and plays a role resembling that of intention).
+Now let's return to joint action.
+%Suppose that the purposiveness of some actions really is grounded in motor representation in the way just described.
+Could there be a parallel role for motor representation in explaining which events are joint actions?
 
@@
 This needs unpacking. 
-A \textit{motor} representation is the sort of representation that enables us to reach for, grasp and transfer objects in a coordinated and fluid way.
-We follow several psychologists and philosophers in supposing that motor representations feature in planning and monitoring action \citep[e.g.]{wolpert:1995internal, miall:1996_forward}.
-One consequence is that motor representations are not concerned with merely kinematic or dynamic features of actions only.
+As already mentioned, a \textit{motor} representation is the sort of representation that enables us to reach for, grasp and transfer objects in a coordinated and fluid way.
+To repeat, motor representations are not concerned with merely kinematic or dynamic features of actions only.
 Rather, some motor representations represent outcomes, such as the movement of a target object from one place to another.
@@
 It is hardly controversial that reciprocal motor representations exist, for their existence is suggested by by a large body of research on motor cognition in action observation.
-What is much more controversial is that reciprocal motor representation occurs in joint action.
+It is more controversial that reciprocal motor representation occurs in joint action.
 Some of the research that Gunether Knoblich presented supports the claim that there are reciprocal motor representations in joint action, and he suggests that these enable joint action.%
@@
 
-I want to go a tiny step further and suggest that in joint action there are sometimes reciprocal, agent-neutral representations of outcomes whose obtaining would normally involve action on the part of all of the agents.
+I want to go a tiny step further and suggest that in joint action there are sometimes reciprocal, agent-neutral representations of outcomes whose obtaining would normally involve action on the part of each of the agents.
 So when we act together, some of my motor representations may concern outcomes that are partly but not entirely to be realised by my actions.
@@
 
-What follows is speculative philosophy: we take for granted that sometimes social motor representation facilitates joint action and ask whether this conjecture bears on our question about which events are joint actions.
+What follows is speculative philosophy: we take for granted that  social motor representation sometimes facilitates joint action and ask whether this conjecture bears on our question about which events are joint actions.
 
@@
 Now suppose, 
-inspired by Koutis et al’s and others' findings,
+inspired by \citet{kourtis:2010_favoritism,kourtis:2012_predictive} and other findings mentioned above,
 that the same planning is involved in the individual case (where one agent performs the whole action) and in the joint action case (where the action is distributed between two agents).
@@
 So far we have only been considering a possible role for social motor representation in facilitating joint action.  
-Our primary concern, though, is with what joint action is.
+How does any of this bear on our main question about which events are joint actions?
 The details of how 
@@
 
-[***moved to earlier]
-
-The same two questions we asked about ordinary, individual action also arise for joint action.
+Earlier, in section \vref{sec:individual_action_motor_representation}, we asked two questions we asked about ordinary, individual action; the same two questions also arise for joint action.
 %What is the relation between a purposive joint action and the outcome or outcomes to which it is directed?

### 2012-10-09 joint action and knowing others minds/interacting mindreaders.3.ps.corrections.tex
tidying up

@@
 The two are on a par in this sense:
-in principle the same evidence could be available to each, and each can exploit the routes to knowledge in moving from evidence to ascriptions of thought and action.
+in principle the same evidence could be available to each, and each can exploit the 
+%correction! ***
+same 
+routes to knowledge in moving from evidence to ascriptions of thought and action.
 Of course these theories are all compatible with the idea that interaction might be useful for mindreading in practice.  
@@
 Second, in some of those cases knowing the premises would put one in a position to know the conclusion.  
-We shall consider these points in turn
+We shall consider these points in turn. %correction! (the period) ***
 

### 2012-10-09 mindreading/minimal theory of mind (new version)/minimal theory of mind.nv.tex
tidying up

@@
 using facts about others' beliefs is not automatic \citep{Keysar:2003xu,apperly:2010_limits}.
-The case for nonautomaticity is indirectly supported by evidence that tracking perceptions and beliefs---and even merely holding in mind what another believes, where no inference is required---involves a measurable processing cost  \citep{apperly:2008_back,apperly:2010_limits}, consumes attention and working memory in fully competent adults (\citealp{en_1698, lin:2010_reflexively, en_1547} experiments 4-5), may require inhibition \citep{bull:2008_role} and makes demands on executive function \citep{apperly:2004_frontal,samson:2005_seeing}.
+The case for nonautomaticity is indirectly supported by evidence that tracking perceptions and beliefs---and even merely holding in mind what another believes, where no inference is required---involves a measurable processing cost  \citep{apperly:2008_back,apperly:2010_limits}, consumes attention and working memory in fully competent adults (\citealp{Apperly:2009cc, lin:2010_reflexively, McKinnon:2007rr} experiments 4-5), may require inhibition \citep{bull:2008_role} and makes demands on executive function \citep{apperly:2004_frontal,samson:2005_seeing}.
 These findings, taken together, suggest that tracking others' perceptions and beliefs is sometimes not automatic.
@@
 We stop here because false belief tasks are often taken to be an acid test for theory of mind.
-But of course additonal principles can be added accommodate further theory of mind abilities.  
+But of course additional principles can be added accommodate further theory of mind abilities.  
 For instance, further principles might extend the definition of registration.  

### 2012-10-11 intention motor representation joint action (paper 3)/conf budapest philosophy 2012-11/motor representation and shared intention HANDOUT.tex
working

--- 2012-03-08 conf joint action --- sharing goals vs sharing intentions cambridge 2012 talk/intention and motor representation in joint action HANDOUT.tex
@@
 
-\def \ititle {Intention and Motor Representation in Joint Action}
+\def \ititle {Shared Agency and Motor Representation}
 \def \isubtitle {}
@@
 {\Large
-\textbf{Intention and Motor Representation in Joint Action}
+\textbf{Shared Agency and Motor Representation}
 }
 
-%Stephen A. Butterfill
+Stephen A. Butterfill \& Corrado Sinigaglia
 
-<s.butterfill@warwick.ac.uk>
+%<s.butterfill@warwick.ac.uk>
+<butterfillS@ceu.hu>
 
@@
 \section{Introduction}
-A motor representation is \emph{agent-neutral} if it concerns an action which is not one’s own or, in the case of joint action, not entirely one’s own.
+A representation (motor or not) is \emph{agent-neutral} if its content does not specify any particular agent or agents.
+
+%***CHECK DEFINITIONS***
 

### 2012-10-11 intention motor representation joint action (paper 3)/conf budapest philosophy 2012-11/shared agency and motor representation.tex
working

@@
 To characterise shared agency, some philosophers have appealed to a special kind of intention or structure of intention, knowledge or commitment often called `shared intention'.  
-In this paper we argue that fully characterising shared agency may additionally require appeal to  motor representation.
+In this paper we argue that there are forms of shared agency characterising which requires appeal to  motor representation.  
 Shared agency is not only a matter of what we intend: sometimes it also depends on interlocking structures of motor representation.
@@
 
-%\section{***EVIDENCE***}
-%I now think three bits of evidence are relevant:
-%\begin{enumerate}
-%\item The GROOP effect shows that there are representations which specify each agent's task in relation to the other (so they are not simple representing the outcomes to which each of their actions are directed; they are representing an outcome to which their actions taken together are directed.)
-%\item \citet{kourtis:2012_predictive} shows that motor planning can occur for others' actions when we are engaged in joint action with them.
-%\item Vesper's ESPP paper (on jumping together and imagining jumping together --- jumping is published \citep{vesper:2012_jumping}, imagining jumping is about to be submitted (as of August 2012).
-This shows that individuals are capable of running motor simulations of multiple roughly simultaneous actions.
-%\end{enumerate}
-%
+\section{***Discussion with Hong Yu}
+`Once we move beyond the single agent there's an a priori reason to suppose that you have a single kind of agentive structure' (Hong Yu).
+This is slightly tricky for me because I don't think there's a single kind of agentive structure in individual action either.
+
+Dissimilarity between this paper and the paper on intention and motor representation: whereas it was maybe plausible to conjecture that all intentions bottom out in motor representation, it is not similarly plausible to claim that shared intentions bottom out in social motor representations (there are surely cases of shared agency where the sharing is entirely a matter of intention and does not involve motor representations at all).
+
+Hong Yu suggested that we need to include the argument about format in order to argue that motor representations have a role distinct from that of shared intention.
+
+In discussion: I need to shift this paper from saying:
+\begin{quote}
+There is a form of shared agency and that form cannot be characterised without reference to motor representation.
+\end{quote}
+to saying:
+\begin{quote}
+There are at least two forms of shared agency (at least two agentive structures) and some of these are such that characterising them involves appeal to motor representation.
+\end{quote}
+This shift reveals a new problem: we become more open to the objection that philosophy should concern forms of shared agency involving intention only.
+
+I need to stress the \emph{parity argument}.
+The argument is this:
+If 
+\begin{quote}
+
+	two or more agents' actions being appropriately related to a shared intention (as Bratman characterises it) is sufficient for shared agency, 
+\end{quote}
+then
+\begin{quote}
+	two or more agents' actions being appropriately related to a structure of social motor representation is also sufficient for shared agency.
+\end{quote}
 
@@
 
-Shared agency raises a tangle of scientific and philosophical questions.  Psychologically we want to know which mechanisms make it possible to engage in and understand different sorts of shared agency
-\citep{vesper_minimal_2010}.  
-Developmentally we want to know when shared agency emerges, what it presupposes and whether abilities to engage in it somehow facilitate socio-cognitive, pragmatic or symbolic development \citep{Moll:2007gu,Hughes:2004zj,Brownell:2006gu}.
+Shared agency raises a tangle of scientific and philosophical questions.  Psychologically we want to know which mechanisms make it possible \citep{vesper_minimal_2010}.  
+Developmentally we want to know when shared agency emerges, what it presupposes and whether it might somehow facilitate socio-cognitive, pragmatic or symbolic development \citep{Moll:2007gu,Hughes:2004zj,Brownell:2006gu}.
 Phenomenologically we want to characterise what (if anything) is special about experiences of action and agency when shared agency is involved \citep{Pacherie:2010fk}.  
@@
 Thus \citet{gilbert_walking_1990} contrasts two people walking together with two people individually walking side by side.  
-The two pairs' movements may be the same and similarly coordinated (to avoid collision), but walking together involves shared agency whereas merely walking side by side is not. 
+The two pairs' movements may be the same and similarly coordinated (to avoid collision), but walking together involves shared agency whereas merely walking side by side does not. 
 Relatedly,  \citet{Searle:1990em}  contrasts a case in which several park visitors simultaneously run to a central shelter in order to perform a dance with another case in which the park visitors run to the central shelter in order to escape a storm.
-The first is a case involves shared agency, the second does not; but the same movements occur in both.
+The first  case involves shared agency, the second does not; but the same movements occur in both.%
+\footnote{
+Compare Pears on individual action, using contrast cases to argue that whether something is an action depends on its antecedents ***
+}  
 These sorts of contrast case invite the question, 
@@
 
-\section{Shared Intention}
+\section{Shared intention}
 On the leading, best developed accounts of shared agency in philosophy and psychology, 
@@
 `I take a collective action to involve a collective intention.'  \citep[p.\ 5]{Gilbert:2006wr}
+
+\emph{
+***This quote is not clearly endorsing the strong claim (and we don't want to deny that all shared agency will involve shared intention (cf not arguing that there are individual actions which involve intention in paper-1).)
+}
 \end{quote}
@@
 and if you think that shared intention stands to joint action as plain vanilla intention stands to ordinary, individual action,
-then it seems reasonable to suppose that for an event to be a joint action is for it to be appropriately related to a shared intention (as \citealp{Pacherie:2012fk} suggests).
+then it seems reasonable to suppose that for an event to be a joint action is for it to be appropriately related to a shared intention (\citealp{Pacherie:2012fk} offers this argument).
 
@@
 Some philosophers allow that there could be agents whose actions are purposive---and, in some cases, even intentional---although  the agents have no intentions at all.
-For instance, Michael Bratman describes a creature who `acts on the basis of its beliefs and considered desires' only (not intentions) as doing things `intentionally' \citet[p.\ 251]{bratman:2000_valuing}.%
+For instance, Michael Bratman describes a creature who `acts on the basis of its beliefs and considered desires' only (not intentions) as doing things `intentionally' \citet[p.\ 251]{bratman:2000_valuing}.
 %\footnote{
@@
 
-\section{Motor Representation}
+\section{Motor representation}
 Since we aim to argue that a fully adequate account of shared agency must appeal to motor representation, it may be helpful first to provide some background about motor representation.
@@
 A \textit{motor} representation is the sort of representation that enables us to reach for, grasp and transfer objects in a coordinated and fluid way.  
-We follow several psychologists and philosophers in supposing that motor representations feature in planning and monitoring action \citep[e.g.]{wolpert:1995internal, miall:1996_forward}.
+%Imagine someone who, while remaining seated, moves an object from her right to her left, passing it between her hands as she does so. 
+%The transition between hands is not something she needs to think or reason about. 
+%But if the transition occurs without any pause---if she does not have to wait before releasing the object from her right hand into her left in order to check that the object is securely in her left hand---this is a sign that her action involves predicting the position of her left hand.
+%***continue this example?
+We follow several psychologists and philosophers in supposing that motor representations feature in planning and monitoring action \citep[e.g.][]{wolpert:1995internal, miall:1996_forward}.
+%*HY asks why focus here on forward models
 
@@
 
+\section{Interpersonal coordination and motor representation}
+
+These bits of evidence are relevant (***):
+%
+\begin{enumerate}
+
+\item \citet{kourtis:2012_predictive} shows that motor planning can occur for others' actions when we are engaged in joint action with them.
+
+\item Vesper's ESPP paper (on jumping together and imagining jumping together --- jumping is published \citep{vesper:2012_jumping}, imagining jumping is about to be submitted (as of August 2012).
+This shows that individuals are capable of running motor simulations of multiple roughly simultaneous actions.
+(The important point for me is that one can simulate roughly simultaneous actions, not that the simulations are simultaneous.)
+
+\item The GROOP effect shows that there are representations which specify each agent's task in relation to the other (so they are not simple representing the outcomes to which each of their actions are directed; they are representing an outcome to which their actions taken together are directed.)
+
+\item Vesper says forthcoming EEG paper using piano playing paradigm on agent-neutral identification of error: one brain wave signals whether there is an error, and a different brain wave signals whose error it is (also tells you whether the overall harmonics are affected)
+\end{enumerate}
+
 \section{Social motor representation}
@@
 
-A representation (motor or not) is \emph{agent-neutral} if its content does not specify any agent or agents.%
+A representation (motor or not) is \emph{agent-neutral} if its content does not specify any particular agent or agents.%
 \footnote{
@@
 In principle such representations could have contents which specify other agents or multiple agents. 
-We focus on agent-neutral representations to simplify exposition.
-It may be important that our view is consistent with the possibility that reciprocal motor representations are agent-neutral. 
+We focus on agent-neutral representations to show that our view is consistent with the possibility that reciprocal motor representations are agent-neutral. 
 }
@@
 It is hardly controversial that reciprocal motor representations exist, for their existence is suggested by by a large body of research on motor cognition in action observation.
+[***Rest of this paragraph belongs in the evidence section?
+Should be clearer about transition:
+(a) I am observing and we reciprocally represent an outcome of your action (mirroring).
+(b) We are interacting and reciprocally represent an outcome of your action (mirroring in joint action).
+(c) We are interacting and reciprocally represent an outcome to which our actions are distributively directed.
+]
 It is more controversial that reciprocal motor representation occurs in joint action, but there is some evidence for this claim too.%
@@
 
-\section{How could social motor representation facilitate joint action?}
-Suppose that social motor representation is present in some joint action contexts.
-It doesn't follow, of course, that social motor representation facilitates some joint actions.
+\section{How could social motor representation coordinate actions?}
+Suppose that social motor representation is sometimes present when two or more agents act together.
+It doesn't follow, of course, that social motor representation will play any role in coordinating the agents' actions.
 But let us consider just the possibility that it might.
-How could social motor representation facilitate joint action even in principle?
+How could this happen?
+How could social motor representation play a role in coordinating action when two or more agents act together?
+
+To answer this question let us step back from motor representation to illustrate a general principle about planning.
+Suppose that Aravinda runs the trains and Gerhard runs the busses.
+Twice a year, on the equinoxes, Aravinda updates the train timetables and sends Gerhard the changes.
+For his part Gerhard also updates the bus timetables twice a year, on the solstices, and sends Aravinda the changes.
+Each wants their services to be optimally coordinated with the other's services in the sense of minimising passengers' journey times (perhaps only in order to maximise profits).
+So how successful Gerhard's plans are depends on Aravinda's plans, and conversely. 
+This is why each is responsive to the other's plans; indeed, each may even try to predict changes in the other's plans.  
+But from each individual's point of view, the other's plans are  merely a constraint.
+This approach to planning suffers from two defects.
+First, it is unlikely to be optimal in the sense of resulting, eventually, in a combination of plans such that no other combination of plans would have been better for at least one service and no worse for either service.
+Second, it is unlikely to be efficient in the sense of allowing Gerhard and Aravinda to arrive at an optimal combination of plans for the two services, trains and buses, with the fewest iterations.
+How could they do better?
+One possibility may be to have a single plan covering busses and trains---perhaps for example Aravinda could buy Gerhard's franchise. 
+But suppose that this is not possible, and that there are limits on how much information Aravinda and Gerhard can share. 
+Now there is a single goal to which Aravinda and Gerhard's activities are both directed.
+
+%Let us go very slowly here.
+%Suppose that, in moving the mug, you try to locate two separate plans:
+%a plan for your left hand and a plan for your right hand.
+%In the Left Plan (as we might call it), the movements of the right hand feature as a constraint.
+%And conversely for the Right Plan.
+%It is almost as if you two hands are under the control of two separate agents who periodically share information about plans.
+%So if you update the Right Plan, you will need to make compensating changes to the Left Plan.
+%And of course those compensating changes to the left plan may in turn trigger further changes to the Right Plan.
+%And so on.
+%This approach might work in some cases. 
+%But it imposes two costs.
+%First, it is unlikely to be efficient in the sense of minimising the range of possible combinations that are considered. 
+%Second, it is unlikely to be optimal in the sense of finding a combination of plans such that no other combination of plans would have been better for at least one hand and no worse for any hand.
+%How could you have done beter?
+%The mistake was to have two separate plans, Left and Right, in each of which the movements of the other hand are a constraint.
+%The better
 
@@
 (iii) each  agent's plans for their own action are constrained by the plans for the other agent's action.
+
+[***TODO: (a) contrast this case with a team of experts, each with different motor expertise (e.g.\ musicians playing together).  
+They can't plan each others' actions.
+(b) Discuss in how much detail each others' actions should be planned so as to enable coordination.
+]
 

### 2012-10-25 intention motor representation joint action (paper 3)/conf budapest philosophy 2012-11/shared agency and motor representation.tex
working on it

@@
 In this paper we argue that there are forms of shared agency characterising which requires appeal to  motor representation.  
-Shared agency is not only a matter of what we intend: sometimes it also depends on interlocking structures of motor representation.
+Shared agency is not only a matter of what we intend: sometimes it  constitutively involves interlocking structures of motor representation.
 This may have consequences for some metaphysical, normative and phenomenological questions about shared agency.
@@
 \section{***Discussion with Hong Yu}
-`Once we move beyond the single agent there's an a priori reason to suppose that you have a single kind of agentive structure' (Hong Yu).
+`Once we move beyond the single agent there's no a priori reason to suppose that you have a single kind of agentive structure' (Hong Yu).
 This is slightly tricky for me because I don't think there's a single kind of agentive structure in individual action either.
@@
 	and walk together \citep{gilbert_walking_1990}.
-In developmental psychology paradigm cases of shared agency include  two people 
+In developmental psychology paradigm cases involving shared agency include  two people 
 	tidying up the toys together \citep{Behne:2005qh},
@@
 
-A \emph{joint action} is an action or event involving shared agency.
+Which representations %and processes 
+are needed in giving a philosophical account of shared agency?
+Philosophers have so far considered intentions and other propositional attitudes but ignored the possibility that appeal to motor representation may also be needed. 
+%To considerations jointly motivate considering whether fully characterising shared agency may require appeal to motor representation.
+By contrast, concerning individual rather than shared agency, some have argued that motor representation may play a role in explaining agency \citep[e.g.][]{pacherie:2000_content,pacherie:2008_action,butterfill:2012_intention}.
+Scientific research provides evidence that motor representations sometimes play a role in enabling agents to exercise shared agency, as we shall explain in detail below.
+So it is possible to ask whether appeal to motor representation  %might not only sometimes enable agents to exercise shared agency but also
+could also play a role in explaining what shared agency is.
+In this paper we argue that it could.
+Shared agency is not only a matter of what we intend or know, or of our commitments: in some cases it also constitutively involves certain structures of motor representation, or so we aim to show.
 
+%A \emph{joint action} is an action or event involving shared agency.
+
+Before starting on the argument, let us step back to ask why anyone might want a philosophical account of shared agency.
 Shared agency raises a tangle of scientific and philosophical questions.  Psychologically we want to know which mechanisms make it possible \citep{vesper_minimal_2010}.  
@@
 %, plus a formal account of how practical reasoning for joint action differs (if at all) from individual practical reasoning \citep{Sugden:2000mw,Gold:2007zd}
+One use for a philosophical account of shared agency is to support investigation of these questions by providing a principled way to distinguish shared from individual agency (as \citealp{Bratman:2009lv} suggests).
+Or, if there is more than one form of shared agency, the account should provide a principled way of distinguishing among forms of individual and shared agency.   
+%
+%To make this aim vivid, consider some well-known contrast cases.
+%Contrast cases are pairs of episodes which are similar in terms of the behaviour and coordination they involve but where one episode exemplifies shared agency  while the other does not.  
+%Thus \citet{gilbert_walking_1990} contrasts two people walking together with two people individually walking side by side.  
+%The two pairs' movements may be the same and similarly coordinated (to avoid collision), but walking together involves shared agency whereas merely walking side by side does not. 
+%Relatedly,  \citet{Searle:1990em}  contrasts a case in which several park visitors simultaneously run to a central shelter in order to perform a dance with another case in which the park visitors run to the central shelter in order to escape a storm.
+The first  case involves shared agency, the second does not; but the same movements occur in both.%
+%\footnote{
+%Compare \citet{Pears:1971fk} who uses contrast cases to argue that whether something is an ordinary, individual action depends on its antecedents.
+%}  
+%
+%Contrast cases invite the question, 
+%How do actions involving shared agency differ from merely parallel actions involving  only individual agency? 
+%Gilbert’s example shows that the difference can’t just be a matter of coordination, because people who are merely walking alongside each other also need to coordinate their actions in order to avoid colliding.
+%And Searle’s example shows that the difference between shared agency and parallel individual agency can’t just be that the resulting actions have a common effect because merely parallel actions can have common effects too.
+%Minimally, a philosophical account of shared agency should identify a difference in these and other contrast cases.
+%
 
-To support investigation of these questions it may be useful to have an account of shared agency that gives us a principled way of distinguishing shared from individual agency \citep[compare][]{Bratman:2009lv}.
-
-Where could we start in constructing or evaluating an account of shared agency?
-One possibility is to start with some well-known contrast cases.
-Contrast cases are pairs of events which are similar in terms of the behaviour and coordination they involve but where one exemplifies shared agency  while the other does not.  
-Thus \citet{gilbert_walking_1990} contrasts two people walking together with two people individually walking side by side.  
-The two pairs' movements may be the same and similarly coordinated (to avoid collision), but walking together involves shared agency whereas merely walking side by side does not. 
-Relatedly,  \citet{Searle:1990em}  contrasts a case in which several park visitors simultaneously run to a central shelter in order to perform a dance with another case in which the park visitors run to the central shelter in order to escape a storm.
-The first  case involves shared agency, the second does not; but the same movements occur in both.%
-\footnote{
-Compare Pears on individual action, using contrast cases to argue that whether something is an action depends on its antecedents ***
-}  
-These sorts of contrast case invite the question, 
-How does shared agency differ from individual but parallel agency? 
-Gilbert’s example shows that the difference can’t just be a matter of coordination, because people who are merely walking alongside each other also need to coordinate their actions in order to avoid colliding.
-And Searle’s example shows that the difference between shared agency and parallel individual agency can’t just be that the resulting actions have a common effect because merely parallel actions can have common effects too.
-So one constraint on any account of shared agency is that it should enable us to make systematic sense of these and other contrast cases.
+As already mentioned, our aim is to show that distinguishing individual from shared agency sometimes involves appeal to certain structures of motor representation.
+First we outline a standard approach to shared agency.
 
@@
 shared agency is explained  in terms of a special kind of intention or structure of intention, knowledge and commitment often called a shared intention.
-On this widely held view, the difference between our walking together and two strangers merely walking side-by-side is that one event is appropriately related to a shared intention whereas the other is not.
+For two or more agents to exercise shared agency is for them to have a shared intention and for this shared intention to be appropriately related to their actions.
+Such actions are called a `shared intentional activity' (\citealp[p.\ 142]{Bratman:1999fr}).%
+%On this widely held view, the difference between our walking together and two strangers merely walking side-by-side is that one event is appropriately related to a shared intention whereas the other is not.
 
-Of course, explaining shared agency in terms of shared intention pushes us back to the question, What is shared intention?
-The term `shared intention' is a term of art.
+Explaining shared agency in terms of shared intention pushes us back to the question, What is shared intention?
+The term `shared intention' is potentially misleading.
 On many accounts, shared intentions are neither shared nor intentions.
-Minimally, a shared intention stands to a joint action in roughly the way that an ordinary, individual intention stands to an ordinary, individual action.
-Now an ordinary, individual intention represents an outcome, coordinates an agent's actions, and coordinates the agent's actions in such a way that, normally, this coordination would facilitate the occurrence of the represented outcome.
-Similarly, a shared intention involves one or more representations of an outcome, coordinates the several agents' actions, and coordinates them in such a way that, normally, the coordination would facilitate the occurrence of the represented outcome.
-
-Beyond this parallel there is much divergence on what shared intention is.
+And there is much divergence on what shared intention is.
 Some hold that it differs from
@@
 
-We can distinguish strong and weak claims linking shared agency to shared intention.
-The weak claim is that shared intention will enable us to understand \emph{some forms} of shared agency.
-This claim is neutral on whether there might be joint actions without shared intention, and on whether ingredients other than shared intention might be needed in giving a full account of shared agency.
-The strong claim is that an account of shared intention gives us the whole story about shared agency, so that no other ingredients are needed.
-Given this claim, for an event to be a joint action is for it to be appropriately related to a shared intention.
+To avoid the need to take sides on what precisely shared intention is, 
+let us abstract from some details. 
+There are features of shared intention on which there should be widespread agreement.
+Minimally, a shared intention stands to the actions comprising a shared intentional activity in roughly the way that an ordinary, individual intention stands to an ordinary, individual action.
+Now an ordinary, individual intention represents an outcome, coordinates an agent's actions, and coordinates the agent's actions in such a way that, normally, this coordination would facilitate the occurrence of the represented outcome.
+Similarly, 
+on almost any account,
+a shared intention involves there being a single outcome which each agent represents, where these representations coordinate the several agents' actions and coordinate them in such a way that, normally, the coordination would facilitate the occurrence of the represented outcome.
 
-Some philosophers and psychologists have endorses the strong claim.  For instance:  
-%
-\begin{quote} 
-`I take a collective action to involve a collective intention.'  \citep[p.\ 5]{Gilbert:2006wr}
+Our view involves rejecting the claim that a full account of shared agency can be given in terms of shared intention alone.
+Instead, we shall argue, it is also necessary to appeal to certain structures of motor representation. 
+Note, however, that our view is compatible with various weaker claims linking shared intention to shared agency.
+In particular, our view is consistent with the claims that appeal shared intention is also needed in giving an account of shared agency, 
+and that \emph{some forms} of shared agency can be fully characterised by appeal to shared intention alone.
+Our view is also consistent with the claim that all exercises of shared agency involve shared intentions in some way.%
+\footnote{
+See \citet[p.\ 5]{Gilbert:2006wr}: `I take a collective action to involve a collective intention.'  
+Related claims are made by 
+	\citet[p.\ 381]{Carpenter:2009wq}, 
+	\citet[p.\ 369]{Call:2009fk}, 
+	\citet{Kutz:2000si},  
+	\citet[p.\ 117]{rakoczy_pretend_2006},  and
+	\citet{Tollefsen:2005vh}.
+}
 
-\emph{
-***This quote is not clearly endorsing the strong claim (and we don't want to deny that all shared agency will involve shared intention (cf not arguing that there are individual actions which involve intention in paper-1).)
-}
-\end{quote}
-%
-Further examples could be given  (%
-	\citealp[p.\ 381]{Carpenter:2009wq}; 
-	\citealp[p.\ 369]{Call:2009fk};
-	\citealp{Kutz:2000si}; 
-	\citealp[p.\ 117]{rakoczy_pretend_2006}; 
-	\citealp{Tollefsen:2005vh}%
-	).
-In what follows we remain neutral on the weak claim but argue against the strong claim.
-We argue against the strong claim on the grounds that shared agency is sometimes not only a matter of shared intention but may also involves structures of interlocking motor representations.
-This may have consequences for some of the metaphysical, normative and phenomenological questions about shared agency mentioned above.
-
-\section{The parallel with ordinary intentions}
-Why might anyone endorse the strong claim that for an event to be a joint action is for it to be appropriately related to a shared intention? 
-Those who endorse it rarely or never provide arguments.
-However, two considerations might appear favour this view.
+Why might anyone hold, contrary to what we shall argue, that a full account of shared intention without appeal to ingredients other than shared intention? 
+Two considerations might initially appear favour this view.
 
@@
 We shall argue, however, that distinguishing systematically between the contrast cases cannot be done entirely in terms of shared intention. 
-[*HY: this argument is missing from this draft.]
+[*CV: this argument is missing from this draft.]
 
-{\label{text:second_consideration} A second consideration} in favour of this view is a possible parallel between joint action and ordinary, individual action.
-If you think that events are actions in virtue of being appropriately related to intentions, 
-and if you think that shared intention stands to joint action as plain vanilla intention stands to ordinary, individual action,
-then it seems reasonable to suppose that for an event to be a joint action is for it to be appropriately related to a shared intention (\citealp{Pacherie:2012fk} offers this argument).
-
-Concerning this second consideration, 
-we take for granted that the parallel obtains but we do not accept the view about ordinary, individual action.
-Not everyone accepts that for an event to be an action is for it to be appropriately related to an intention.
-Some philosophers allow that there could be agents whose actions are purposive---and, in some cases, even intentional---although  the agents have no intentions at all.
-For instance, Michael Bratman describes a creature who `acts on the basis of its beliefs and considered desires' only (not intentions) as doing things `intentionally' \citet[p.\ 251]{bratman:2000_valuing}.
-%\footnote{
-%For instance, \citet[p.\ 251]{bratman:2000_valuing} % = \citep[p.\ 51]{Bratman:2007bd}
-%describes an agent called `Creature 2' who `acts on the basis of its beliefs and considered desires' only (not intentions) as doing things `intentionally'.
-%}
-Others reject the standard view without asserting there can be action without intention.
-Instead they claim that ingredients other than intention are needed to say which events are actions.
-In particular, some claim that motor representation and intention are both needed in explaining which events are ordinary, individual actions \citep{pacherie:2000_content,butterfill:2012_intention}.
-If any such view is correct, it is impossible to fully explain what it is for an event to be an action by appeal to intention alone.
-So depending on your views about ordinary, individual intention,
-the second consideration---the idea that the relation of shared intention to joint action is in some ways parallel to the relation between ordinary, individual intention and plain vanilla action---might actually support the view we aim to establish, namely that fully explaining shared agency requires appeal to structures of motor representation and not only to shared intention.
-%***what's the argument for this second claim!?
+{\label{text:second_consideration} A second, related consideration} which might appear to favour this view is a parallel between shared agency and ordinary, individual agency.
+If a full account of individual agency could be given by appeal to intention and other propositional attitudes with no appeal to motor representation,
+then it would seem unlikely that giving a full account of shared agency must somehow involve appeal to motor representation.
+However, as mentioned above, some have argued that fully characterising individual agency will require appeal to motor representation as well as intention \citep{pacherie:2000_content,butterfill:2012_intention}.
+Given these arguments, the parallel between individual and shared agency motivates considering the possibility that  there is a constitutive role for motor representation in shared agency. 
+We also note that some philosophers allow that there could be agents whose actions are purposive---and, in some cases, even intentional---although  the agents have no intentions at all.
+For instance, Bratman describes a creature who `acts on the basis of its beliefs and considered desires' only (not intentions) as doing things `intentionally' (\citeyear[p.\ 251]{bratman:2000_valuing}; see \citealp[pp.\ 137--8]{Bratman:1987xw} for further justification).
+Certainly, then, it is not unreasonable to question the claim that a full account of ordinary, individual agency can be given just in terms of intention.
 
@@
 We follow several psychologists and philosophers in supposing that motor representations feature in planning and monitoring action \citep[e.g.][]{wolpert:1995internal, miall:1996_forward}.
-%*HY asks why focus here on forward models
+[*HY asks why focus here on forward models]
 
@@
 (b) We are interacting and reciprocally represent an outcome of your action (mirroring in joint action).
-(c) We are interacting and reciprocally represent an outcome to which our actions are distributively directed.
+(c) We are interacting and reciprocally represent an outcome to which our actions are distributively directed.%
 ] 
@@
 
-[*HY: I'm planning a short section on evidence for social motor representation.  But as I'm going beyond the evidence anyway, I don't plan to include it in presenting this paper to philosophers.]
-
 \section{How could social motor representation coordinate actions?}
@@
 
-To answer this question let us step back from motor representation to illustrate a general principle about planning.
-Suppose that Aravinda runs the trains and Gerhard runs the busses.
-Twice a year, on the equinoxes, Aravinda updates the train timetables and sends Gerhard the changes.
-For his part Gerhard also updates the bus timetables twice a year, on the solstices, and sends Aravinda the changes.
-Each wants their services to be optimally coordinated with the other's services in the sense of minimising passengers' journey times (perhaps only in order to maximise profits).
-So how successful Gerhard's plans are depends on Aravinda's plans, and conversely. 
-This is why each is responsive to the other's plans; indeed, each may even try to predict changes in the other's plans.  
+As a first step towards answering  this question let us illustrate a general principle about planning by scaling up from motor action all the way to corporate action.
+
+Suppose that Aravinda runs the trains and Gerhard  the busses.
+Let us stipulate that we measure the extent to which the two services, train and bus, are coordinated is to be measured by the sum of all passengers' total journey times.
+So new timetables would result in better coordination than the current timetables just if the new timetables would decrease the sum of passengers' total journey times.
+To illustrate, improving coordination might involve having trains arrive at a station shortly before, rather than shortly after, busses depart from there.
+Now Aravinda and Gerhard each intend they, Aravinda and Gerhard, coordinate the trains with the busses  to the greatest extent possible given other constraints; and they intend to do this by way of these intentions and meshing subplans of them, and this is common knowledge between them.
+Aravinda and Gerhard thus meet Bratman's sufficient conditions for them to have a shared intention.
+Acting on this intention, 
+every January Aravinda updates the train timetables and sends Gerhard the changes.
+Likewise, Gerhard  updates the bus timetables every June and sends Aravinda the changes.
+In this way  each is responsive to the other's intentions and subplans of these. 
+Indeed, each may even try to predict changes in the other's subplans and modify their own accordingly.  
 But from each individual's point of view, the other's plans are  merely a constraint.
+
 This approach to planning suffers from two defects.
@@
 Second, it is unlikely to be efficient in the sense of allowing Gerhard and Aravinda to arrive at an optimal combination of plans for the two services, trains and buses, with the fewest iterations.
+
 How could they do better?
-One possibility may be to have a single plan covering busses and trains---perhaps for example Aravinda could buy Gerhard's franchise. 
+One possibility may be to have a single person planning the busses and trains---perhaps, for example, Aravinda could buy Gerhard's franchise. 
 But suppose that this is not possible, and that there are limits on how much information Aravinda and Gerhard can share. 
-Now there is a single goal to which Aravinda and Gerhard's activities are both directed.
+Then another way that might improve things would be to have Aravinda and Gerhard each plan everything.
+So Aravinda would plan the best timetable for all the trains and busses, and Gerhard would do the same.
+Since Aravinda only controls the trains, she can only implement the train-related part of this master timetable.
+Likewise, Gerhard can only implement the bus part of his master timetable.
+But if Aravinda's and Gerhard's separate planning processes are sufficiently similar, this process may result in coordinated services.
+In this new case, 
+Aravinda and Gerhard coordinate their actions 
+not by knowing or predicting each others' plans 
+but by thinking about the best overall plan.
+Instead of each viewing the others' plans and actions as constraint on their own,
+in planning actions they absorb the others' actions.
+
+The question was how social motor representation might play a role in coordinating action.
+We propose that it might do so in the way illustrated by Aravinda and Gerhard's final attempt at coordination,
+where each plans everything despite only being in a position to execute half of the plan.
 
@@
 
-To answer this question
-let us take a step back and consider an individual action.
+To see how this might work,
+let us step down to motor action but first consider only  an individual action.
 Suppose an agent moves a mug from one place to another, passing it from her left hand to her right hand half way.
@@
 \footnote{
-Another possible line of justification the claim that common knowledge is involved in shared intention might start from a generalisation of Davidson's claim that
+Another possible line of justification for the claim that common knowledge is involved in shared intention might start from a natural generalisation of Davidson's claim that
 `[a]ction does require %that what the agent does is intentional under some description, and this in turn requires 

### 2012-10-27 intention motor representation joint action (paper 3)/conf budapest philosophy 2012-11/tmp.tex
before change introduction messing up contrast cases

--- 2012-10-25 intention motor representation joint action (paper 3)/conf budapest philosophy 2012-11/shared agency and motor representation.tex
@@
 For two or more agents to exercise shared agency is for them to have a shared intention and for this shared intention to be appropriately related to their actions.
-Such actions are called a `shared intentional activity' (\citealp[p.\ 142]{Bratman:1999fr}).%
+Such actions comprise a \emph{shared intentional activity} (\citealp[p.\ 142]{Bratman:1999fr}).%
 %On this widely held view, the difference between our walking together and two strangers merely walking side-by-side is that one event is appropriately related to a shared intention whereas the other is not.
@@
 
-To avoid the need to take sides on what precisely shared intention is, 
-let us abstract from some details. 
-There are features of shared intention on which there should be widespread agreement.
+To avoid having to take sides on what shared intention is, 
+let us abstract the details
+and focus on features that everyone should agree on.
 Minimally, a shared intention stands to the actions comprising a shared intentional activity in roughly the way that an ordinary, individual intention stands to an ordinary, individual action.
@@
 
-Our view involves rejecting the claim that a full account of shared agency can be given in terms of shared intention alone.
-Instead, we shall argue, it is also necessary to appeal to certain structures of motor representation. 
-Note, however, that our view is compatible with various weaker claims linking shared intention to shared agency.
-In particular, our view is consistent with the claims that appeal shared intention is also needed in giving an account of shared agency, 
-and that \emph{some forms} of shared agency can be fully characterised by appeal to shared intention alone.
+There is another feature of shared intention that should be uncontroversial.  
+Shared intentions or their components feature in practical reasoning alongside ordinary, individual intentions,
+ and there are normative requirements which apply to combinations of individual and shared intentions.
+To illustrate, tonight there is a party and a ceremony.
+It is impossible for anyone to attend both, and this is common knowledge among us.
+We have a shared intention that we attend the ceremony together.
+While having this shared intention, I also intend to go to the party. 
+Give our common knowledge,
+this combination of shared and individual intentions is irrational. 
+Its irrationality is related to that which would be involved in my individually intending to attend the ceremony while also intending to go to the party.
+In short, shared intentions or their components are inferentially and normatively integrated with ordinary, individual intentions.
+
+So much for shared intention.
+Can we give a full account of shared agency by saying (roughly) that exercises of shared agency involve agents acting on a shared intention?
+Or are further ingredients necessary to fully explicate shared agency?
+Our view is that further ingredients are necessary.
+In giving a full account of shared agency, it is also necessary to appeal to certain structures of motor representation. 
+Note, however, that our view is consistent with various weaker claims linking shared intention to shared agency.
+Our view is consistent with the claim that appeal shared intention is also needed in giving an account of shared agency, 
+and with the claim that \emph{some forms} of shared agency can be fully characterised by appeal to shared intention alone.
 Our view is also consistent with the claim that all exercises of shared agency involve shared intentions in some way.%
@@
 
-First, if someone thought that it is possible to distinguish systematically between the contrast cases by appeal to shared intention (and perhaps only by appeal to shared intention), this could motivate holding that for an event to be a joint action is for it to be appropriately related to a shared intention.
+First, if someone thought that it is possible to distinguish systematically between the contrast cases by appeal to shared intention (and perhaps only by appeal to shared intention), this could motivate holding that further ingredients are unnecessary in explicating shared agency.
 We shall argue, however, that distinguishing systematically between the contrast cases cannot be done entirely in terms of shared intention. 
@@
 then it would seem unlikely that giving a full account of shared agency must somehow involve appeal to motor representation.
-However, as mentioned above, some have argued that fully characterising individual agency will require appeal to motor representation as well as intention \citep{pacherie:2000_content,butterfill:2012_intention}.
+However, as mentioned above, some have argued that fully characterising individual agency requires appeal to motor representation as well as intention \citep{pacherie:2000_content,butterfill:2012_intention}.
 Given these arguments, the parallel between individual and shared agency motivates considering the possibility that  there is a constitutive role for motor representation in shared agency. 
@@
 For instance, Bratman describes a creature who `acts on the basis of its beliefs and considered desires' only (not intentions) as doing things `intentionally' (\citeyear[p.\ 251]{bratman:2000_valuing}; see \citealp[pp.\ 137--8]{Bratman:1987xw} for further justification).
-Certainly, then, it is not unreasonable to question the claim that a full account of ordinary, individual agency can be given just in terms of intention.
+So there are several grounds for doubting that a full account of ordinary, individual agency could be given just in terms of intention. 
+If, as we believe, these doubts are well founded, then it is reasonable to ask whether ingredients other than shared intention are need to give a full account of shared agency.
 
 \section{Motor representation}
-Since we aim to argue that a fully adequate account of shared agency must appeal to motor representation, it may be helpful first to provide some background about motor representation.
+Since we aim to argue that a full account of shared agency must appeal to motor representation, it may be helpful first to provide some background about motor representation.
 
@@
 So much for the bare notion of motor representation.
-Why suppose that motor representation might be needed to explain shared agency?
-
-\section{Interpersonal coordination and motor representation}
-
-These bits of evidence are relevant (***):
-%
-\begin{enumerate}
-
-\item \citet{kourtis:2012_predictive} shows that motor planning can occur for others' actions when we are engaged in joint action with them.
-
-\item Vesper's ESPP paper (on jumping together and imagining jumping together --- jumping is published \citep{vesper:2012_jumping}, imagining jumping is about to be submitted (as of August 2012).
-This shows that individuals are capable of running motor simulations of multiple roughly simultaneous actions.
-(The important point for me is that one can simulate roughly simultaneous actions, not that the simulations are simultaneous.)
-
-\item The GROOP effect shows that there are representations which specify each agent's task in relation to the other (so they are not simple representing the outcomes to which each of their actions are directed; they are representing an outcome to which their actions taken together are directed.)
-
-\item Vesper says forthcoming EEG paper using piano playing paradigm on agent-neutral identification of error: one brain wave signals whether there is an error, and a different brain wave signals whose error it is (also tells you whether the overall harmonics are affected)
-\end{enumerate}
+Why suppose that motor representation might be needed in giving an account of shared agency? 
+Our answer involves several steps. 
+We first describe a structure of motor representation,
+and then present evidence that such structures sometimes occur when two or more agents coordinate their actions. 
+This will allow us to conclude only
 
 \section{Social motor representation}
-We start from an empirical premise which is about enabling conditions for shared agency only: some joint actions are facilitated by reciprocal, agent-neutral motor representations of outcomes whose obtaining would normally involve action on the part of each agent.
+We start from an empirical premise about enabling conditions for shared agency: some joint actions are facilitated by reciprocal, agent-neutral motor representations of outcomes whose obtaining would normally involve action on the part of each agent.
 
@@
 It is sufficient for our purposes that there are reciprocal motor representations  of outcomes whose obtaining would normally involve action on the part of each of the reciprocating agents.
-In principle such representations could have contents which specify other agents or multiple agents. 
-We focus on agent-neutral representations to show that our view is consistent with the possibility that reciprocal motor representations are agent-neutral. 
+In principle such representations could have contents which specify particular agents. 
+However, we focus on agent-neutral representations to show that our view is consistent with the possibility that reciprocal motor representations are agent-neutral. 
 }
@@
 Two or more agents have \emph{reciprocal} motor representations  just if there is a single outcome and each agent has a motor representation of that outcome. 
+Reciprocal motor representations can occur both when one agent observes another acting alone and also when two or more agents act together.
+Where on agent observes another acting alone,
+it is possible that both have motor representations of an outcome to which the agent's action is directed.
+For the agent, the representation plays a role in planning and monitoring; for the observer, the representation facilitates recognition.
+This case of reciprocal motor representation---often called mirroring---has been extensively studied.
+But our main interest is in a different case,
+one where two or more agents are acting together,
+ there is a single outcome to which their actions are directed,
+ and the agents each have a motor representation of this outcome.
+So in this case of reciprocal motor representation, the agents each represent an outcome whose obtaining would normally involve not only their own actions but also those of other agents.
+(Note that our having reciprocal motor representations does not imply that we know this.)
+When we act together, some of my motor representations may concern outcomes that are partly but not entirely to be realised by my actions; and likewise for you.
+To illustrate, suppose our task is to move an object from A to B, where you pick it up and pass it to me so that I can then place it.
+In this case I may represent the movement of the object from A to B and not only the component movements.
+I represent a collective outcome of our actions and not just outcomes to which each of our actions are individually directed.
+%
+%So it is not just that, when we act together, I represent the outcomes to which your actions are directed and the outcomes to which my actions are directed; I may also represent an outcome to which all of our actions are directed.
+
+We shall use the term \emph{shared} motor representation as an abbreviation for the reciprocal, agent-neutral motor representation of outcomes whose obtaining would normally require action on the part of each reciprocating agent.
+Note that shared motor representations are shared only in the sense in which two people can share a name.  
+(We do not mean to suggest that they are shared in the sense in which two people can share a parent: there need be no representation with two or more subjects.)
+Note also that a shared motor representation is not a special kind of motor representation: it is merely a structure of ordinary motor representations. 
+At this stage we are not strictly entitled use the term `shared motor representation' as an abbreviation.
+After all we have yet to show that these structures of motor representation play any constitutive role (or even any enabling role) in explaining shared agency.
+The choice of term merely reflects our aim.
+
+\section{Interpersonal coordination and motor representation}
+
+These bits of evidence are relevant (***):
+%
+\begin{enumerate}
+
+\item \citet{kourtis:2012_predictive} shows that motor planning can occur for others' actions when we are engaged in joint action with them.
+
+\item Vesper's ESPP paper (on jumping together and imagining jumping together --- jumping is published \citep{vesper:2012_jumping}, imagining jumping is about to be submitted (as of August 2012).
+This shows that individuals are capable of running motor simulations of multiple roughly simultaneous actions.
+(The important point for me is that one can simulate roughly simultaneous actions, not that the simulations are simultaneous.)
+
+\item The GROOP effect shows that there are representations which specify each agent's task in relation to the other (so they are not simple representing the outcomes to which each of their actions are directed; they are representing an outcome to which their actions taken together are directed.)
+
+\item Vesper says forthcoming EEG paper using piano playing paradigm on agent-neutral identification of error: one brain wave signals whether there is an error, and a different brain wave signals whose error it is (also tells you whether the overall harmonics are affected)
+\end{enumerate}
+
 It is hardly controversial that reciprocal motor representations exist, for their existence is suggested by by a large body of research on motor cognition in action observation.
@@
 
-We want to go a tiny step further and suggest that in joint action there are sometimes reciprocal, agent-neutral representations of outcomes whose obtaining would normally involve action on the part of each of the agents.
-So when we act together, some of my motor representations may concern outcomes that are partly but not entirely to be realised by my actions (and likewise for you).
-For example, suppose our task is to move an object from A to B, where you pick it up and pass it to me so that I can then place it.
-In this case I may represent the movement of the object from A to B and not only the component movements.
-I represent a collective outcome of our actions and not just outcomes to which each of our actions are individually directed.
-%
-%So it is not just that, when we act together, I represent the outcomes to which your actions are directed and the outcomes to which my actions are directed; I may also represent an outcome to which all of our actions are directed.
-
-We shall use the term \emph{social} motor representation as an abbreviation for the reciprocal, agent-neutral motor representation of outcomes whose obtaining would normally require action on the part of each reciprocating agent.
-(This is probably a bad term to pick; we are still trying to come up with a better one. 
-Certainly we cannot assume in advance of argument that such reciprocal, agent-neutral motor representations are social in any interesting sense;
-for now we are using the term `social' in a non-standard way as an abbreviation, one that reflects our aim.)
-
-What follows is speculative philosophy: we take for granted that  social motor representation sometimes facilitates joint action and ask whether this conjecture bears on our question about shared agency.
-As a first step, we shall ask how social motor representation might facilitate joint action.
+What follows is speculative philosophy: we now take for granted that  social motor representation sometimes facilitates the exercise of shared agency and ask whether this matters for a philosophical account of shared agency.
+As a first step, we shall ask how social motor representation might facilitate agents in coordinating their actions.
 Note that answering this question does not directly commit us to any view about shared agency.
-After all, social motor representation might be an enabling condition for some instances of shared agency but completely unnecessary for explaining what shared agency is.
+After all, social motor representation might sometimes facilitate coordination while being extraneous to a philosophical account of shared agency.
 

### 2012-10-28 intention motor representation joint action (paper 3)/conf budapest philosophy 2012-11/shared agency and motor representation.tex
working on it

@@
 
+\section{***Discussion with Cordula Vesper}
+\begin{enumerate}
+\item Say more about what shared agency at the start: use the contrast cases instead of examples, and say that it involves some kind of fusing of agency.
+\item Note that in the jumping experiments, only the subject with the shorter jump changes what she does depending on the difference in jump-length between the two subjects.
+Vesper et al interpret this as showing that the subject with the shorter jump plans the whole activity.
+But it isn't obviously reciprocal: there is no evidence that the subject with the longer jump is similarly planning.
+And this makes sense: here they achieve coordination by first distributing roles (a strategy) and then having at least one subject plan both actions (so there is slightly less reliance on the two plans coinciding).
+Of course the appearance that they are distributing roles might just be an artefact of reciprocal planning where the upshot of the planning is that the person with the longer jump just does what she would normally do.
+
+\item {[done]}  Don't say `neither shared nor an intention' unless you explain this.
+\item {[done]} Can cite \citet{konvalinka:2010_follow} as example of what happens when two people each try to predict the other.
+\end{enumerate}
+
 \section{***Discussion with Hong Yu}
@@
 \begin{quote}
-	two or more agents' actions being appropriately related to a structure of social motor representation is also sufficient for shared agency.
+	two or more agents' actions being appropriately related to a structure of social motor representation is also sufficient for shared 
+	agency.
 \end{quote}
@@
 \section{Introduction}
-Shared agency is paradigmatically involved when two (or more) people 
-	paint a house together \citep{Bratman:1992mi}, 
-	lift a heavy sofa together \citep{Velleman:1997oo}, 
-	prepare a hollandaise sauce together \citep{Searle:1990em}, 
-	go to Chicago together \citep{Kutz:2000si}, 
-	and walk together \citep{gilbert_walking_1990}.
-In developmental psychology paradigm cases involving shared agency include  two people 
-	tidying up the toys together \citep{Behne:2005qh},
-	cooperatively pulling handles in sequence to make a dog-puppet sing \citep{Brownell:2006gu},
-	bouncing a ball on a large trampoline together \citep{Tomasello:2007gl}.
+When members of a flash mob in the Central Cafe respond to a pre-arranged cue by noisily opening their newspapers, they exercise shared agency. 
+But when someone not part of the mob just happens to noisily open her newspaper in response to the same cue, her action does not involve shared agency.%
+\footnote{
+See \citet{Searle:1990em}; in his example park visitors simultaneously run to a shelter, in once case as part of dancing together and in another case because of a storm. 
+Compare \citet{Pears:1971fk} who uses contrast cases to argue that whether something is an ordinary, individual action depends on its antecedents. 
+}
+To give another example, 
+ two former members of the mob exercise shared agency when they 
+   later walk to the metro station together. 
+But two people who merely happened to be walking to the metro station side by side would not be exercising shared agency \citep{gilbert_walking_1990}. 
+These sorts of case invite the question, 
+How do actions involving shared agency differ from merely parallel actions involving  only individual agency?
+
+The first example shows that the difference between shared agency and parallel individual agency can’t just be that the resulting actions have a common effect because merely parallel actions can have common effects too.
+And the second example shows that the difference can’t just be a matter of coordination, because people who are merely happen to be walking side by side each other also need to coordinate their actions in order to avoid colliding.
+Apparently, then, systematically distinguishing cases of shared agency from cases of parallel but merely individual agency  requires reflection on causes of actions.
+
+%
+%Shared agency is paradigmatically involved when two (or more) people 
+%	paint a house together \citep{Bratman:1992mi}, 
+%	lift a heavy sofa together \citep{Velleman:1997oo}, 
+%	prepare a hollandaise sauce together \citep{Searle:1990em}, 
+%	go to Chicago together \citep{Kutz:2000si}, 
+%	and walk together \citep{gilbert_walking_1990}.
+%In developmental psychology paradigm cases involving shared agency include  two people 
+%	tidying up the toys together \citep{Behne:2005qh},
+%	cooperatively pulling handles in sequence to make a dog-puppet sing \citep{Brownell:2006gu},
+%	bouncing a ball on a large trampoline together \citep{Tomasello:2007gl}.
 
 Which representations %and processes 
-are needed in giving a philosophical account of shared agency?
-Philosophers have so far considered intentions and other propositional attitudes but ignored the possibility that appeal to motor representation may also be needed. 
+are needed to distinguish shared from individual agency?
+Philosophers have so far considered intentions and commitments but ignored the possibility that appeal to motor representations may also be needed. 
 %To considerations jointly motivate considering whether fully characterising shared agency may require appeal to motor representation.
 By contrast, concerning individual rather than shared agency, some have argued that motor representation may play a role in explaining agency \citep[e.g.][]{pacherie:2000_content,pacherie:2008_action,butterfill:2012_intention}.
-Scientific research provides evidence that motor representations sometimes play a role in enabling agents to exercise shared agency, as we shall explain in detail below.
+In addition, scientific research provides evidence that motor representations sometimes play a role in enabling agents to exercise shared agency, as we shall explain in detail below.
 So it is possible to ask whether appeal to motor representation  %might not only sometimes enable agents to exercise shared agency but also
-could also play a role in explaining what shared agency is.
+could also play a role in distinguishing shared from individual agency.
 In this paper we argue that it could.
-Shared agency is not only a matter of what we intend or know, or of our commitments: in some cases it also constitutively involves certain structures of motor representation, or so we aim to show.
+Shared agency is not always only a matter of our intentions or commitments: in some cases it also constitutively involves certain structures of motor representation.
+Or so we aim to show.
 
-%A \emph{joint action} is an action or event involving shared agency.
-
-Before starting on the argument, let us step back to ask why anyone might want a philosophical account of shared agency.
-Shared agency raises a tangle of scientific and philosophical questions.  Psychologically we want to know which mechanisms make it possible \citep{vesper_minimal_2010}.  
+Before going further,
+let us step back to ask 
+	why it might be useful 
+	to  distinguish systematically between shared and individual agency?
+Shared agency raises a tangle of scientific and philosophical questions.  Psychologically we want to know which mechanisms make it possible \citep{Sebanz:2006yq,vesper_minimal_2010}.  
 Developmentally we want to know when shared agency emerges, what it presupposes and whether it might somehow facilitate socio-cognitive, pragmatic or symbolic development \citep{Moll:2007gu,Hughes:2004zj,Brownell:2006gu}.
@@
 %, plus a formal account of how practical reasoning for joint action differs (if at all) from individual practical reasoning \citep{Sugden:2000mw,Gold:2007zd}
-One use for a philosophical account of shared agency is to support investigation of these questions by providing a principled way to distinguish shared from individual agency (as \citealp{Bratman:2009lv} suggests).
-Or, if there is more than one form of shared agency, the account should provide a principled way of distinguishing among forms of individual and shared agency.   
-%
-%To make this aim vivid, consider some well-known contrast cases.
-%Contrast cases are pairs of episodes which are similar in terms of the behaviour and coordination they involve but where one episode exemplifies shared agency  while the other does not.  
-%Thus \citet{gilbert_walking_1990} contrasts two people walking together with two people individually walking side by side.  
-%The two pairs' movements may be the same and similarly coordinated (to avoid collision), but walking together involves shared agency whereas merely walking side by side does not. 
-%Relatedly,  \citet{Searle:1990em}  contrasts a case in which several park visitors simultaneously run to a central shelter in order to perform a dance with another case in which the park visitors run to the central shelter in order to escape a storm.
-The first  case involves shared agency, the second does not; but the same movements occur in both.%
-%\footnote{
-%Compare \citet{Pears:1971fk} who uses contrast cases to argue that whether something is an ordinary, individual action depends on its antecedents.
-%}  
-%
-%Contrast cases invite the question, 
-%How do actions involving shared agency differ from merely parallel actions involving  only individual agency? 
-%Gilbert’s example shows that the difference can’t just be a matter of coordination, because people who are merely walking alongside each other also need to coordinate their actions in order to avoid colliding.
-%And Searle’s example shows that the difference between shared agency and parallel individual agency can’t just be that the resulting actions have a common effect because merely parallel actions can have common effects too.
-%Minimally, a philosophical account of shared agency should identify a difference in these and other contrast cases.
-%
+A philosophical account of shared agency 
+that made it possible to systematically distinguish shared from individual agency  
+may support investigation of these questions (as \citealp{Bratman:2009lv} suggests).  
+%Or, if there is more than one form of shared agency, the account should provide a principled way of distinguishing among forms of individual and shared agency.
 
 As already mentioned, our aim is to show that distinguishing individual from shared agency sometimes involves appeal to certain structures of motor representation.
-First we outline a standard approach to shared agency.
+Our first step is to review standard approaches to shared agency.
 
@@
 Explaining shared agency in terms of shared intention pushes us back to the question, What is shared intention?
-The term `shared intention' is potentially misleading.
-On many accounts, shared intentions are neither shared nor intentions.
-And there is much divergence on what shared intention is.
+Here is much disagreement.
 Some hold that it differs from
@@
 
-Our view involves rejecting the claim that a full account of shared agency can be given in terms of shared intention alone.
-Instead, we shall argue, it is also necessary to appeal to certain structures of motor representation. 
-Note, however, that our view is compatible with various weaker claims linking shared intention to shared agency.
-In particular, our view is consistent with the claims that appeal shared intention is also needed in giving an account of shared agency, 
-and that \emph{some forms} of shared agency can be fully characterised by appeal to shared intention alone.
-Our view is also consistent with the claim that all exercises of shared agency involve shared intentions in some way.%
+There is another feature of shared intention that should be uncontroversial.  
+Shared intentions or their components feature in practical reasoning alongside ordinary, individual intentions,
+ and there are normative requirements which apply to combinations of individual and shared intentions.
+To illustrate, tonight there is a party and a ceremony.
+It is impossible for anyone to attend both, and this is common knowledge among us.
+We have a shared intention that we attend the ceremony together.
+While having this shared intention, I also intend to go to the party. 
+Give our common knowledge,
+this combination of shared and individual intentions is irrational. 
+Its irrationality is related to that which would be involved in my individually intending to attend the ceremony while also intending to go to the party.
+In short, shared intentions or their components are inferentially and normatively integrated with ordinary, individual intentions.
+
+So much for shared intention.
+Can we give a full account of shared agency by saying (roughly) that exercises of shared agency involve agents acting on a shared intention?
+Or are further ingredients necessary in systematically distinguishing shared from individual  agency?
+Our view is that further ingredients are necessary.
+In giving a full account of shared agency, it is also necessary to appeal to certain structures of motor representation. 
+Note, however, that our view is consistent with various weaker claims linking shared intention to shared agency.
+Our view is consistent with the claim that appeal shared intention is also needed in giving an account of shared agency,  
+%and with the claim that \emph{some forms} of shared agency can be fully characterised by appeal to shared intention alone.
+and with the claim that all exercises of shared agency involve shared intentions in some way.%
 \footnote{
-See \citet[p.\ 5]{Gilbert:2006wr}: `I take a collective action to involve a collective intention.'  
+On the latter claim, see \citet[p.\ 5]{Gilbert:2006wr}: `I take a collective action to involve a collective intention.'  
 Related claims are made by 
@@
 
-Why might anyone hold, contrary to what we shall argue, that a full account of shared intention without appeal to ingredients other than shared intention? 
-Two considerations might initially appear favour this view.
+\section{An objection}
+At this point opponents sometimes attempt to dismiss our view out of hand along these lines.
+Philosophers are interested in shared \emph{intentional} agency. 
+If there are forms of shared agency other than this one, we might be right to claim that understanding these will require appeal to motor representation.
+But---so the objection---such forms of shared agency (if they exist at all) are not of interest to philosophers. 
+So we are clearly wrong to claim that the philosophical project of distinguishing shared from individual agency might require appeal to ingredients other than shared intentions.
 
-First, if someone thought that it is possible to distinguish systematically between the contrast cases by appeal to shared intention (and perhaps only by appeal to shared intention), this could motivate holding that for an event to be a joint action is for it to be appropriately related to a shared intention.
-We shall argue, however, that distinguishing systematically between the contrast cases cannot be done entirely in terms of shared intention. 
-[*CV: this argument is missing from this draft.]
+One possible response to this line of opposition would be to weaken our claim. 
+We could become neutral on whether a full account of the sort of shared agency philosophers have focussed on can be given without appeal to motor representation,
+but insist that there is a form of shared agency about which parallel philosophical questions arise. 
+This conciliatory response would preserve most of the key claims in what follows, but a stronger response is available.
 
-{\label{text:second_consideration} A second, related consideration} which might appear to favour this view is a parallel between shared agency and ordinary, individual agency.
-If a full account of individual agency could be given by appeal to intention and other propositional attitudes with no appeal to motor representation,
-then it would seem unlikely that giving a full account of shared agency must somehow involve appeal to motor representation.
-However, as mentioned above, some have argued that fully characterising individual agency will require appeal to motor representation as well as intention \citep{pacherie:2000_content,butterfill:2012_intention}.
-Given these arguments, the parallel between individual and shared agency motivates considering the possibility that  there is a constitutive role for motor representation in shared agency. 
-We also note that some philosophers allow that there could be agents whose actions are purposive---and, in some cases, even intentional---although  the agents have no intentions at all.
+As mentioned above, some have argued that fully characterising individual agency requires appeal to motor representation as well as intention \citep{pacherie:2000_content,butterfill:2012_intention}. 
+These arguments concern intentional action.
+So if (as we suppose) these arguments are not known to be defective,
+then, assuming a parallel between individual and shared agency,  
+it should not be assumed at the outset that appealing to motor representation in a philosophical account of shared agency implies that the account does not apply to what opponents call `shared intentional agency'.
+We do not state our position in exactly the terms used by our opponent because of doubts about the qualifier `intentional' that are not specific to discussions of shared agency.\footnotemark \  \ 
+However, we do aim show in what follows that a philosophical account of the sort of shared agency that philosophers have been interested in will require appeal to structures of motor representation.
+
+\footnotetext{ 
+We avoid using the qualifier `intentional' for reasons not specific to questions about shared agency. 
+(We do use the term `shared intentional activity', but our use merely follows Bratman's stipulation that this term applies to an activity appropriately related to a shared intention.)
+Briefly, some philosophers offer reasons for holding that there could be agents whose actions are intentional although  the agents have no intentions at all.
 For instance, Bratman describes a creature who `acts on the basis of its beliefs and considered desires' only (not intentions) as doing things `intentionally' (\citeyear[p.\ 251]{bratman:2000_valuing}; see \citealp[pp.\ 137--8]{Bratman:1987xw} for further justification).
-Certainly, then, it is not unreasonable to question the claim that a full account of ordinary, individual agency can be given just in terms of intention.
+These and related considerations reveal a coherent and natural way of using the term `intentional' on which it would apply to actions in virtue of their being appropriately related to motor representations of outcomes to which they are directed.
+It is not obvious to us whether using the term in this way is incorrect.
+}
 
-\section{Motor representation}
-Since we aim to argue that a fully adequate account of shared agency must appeal to motor representation, it may be helpful first to provide some background about motor representation.
+\section{Shared motor representation}
+Our argument starts with an empirical premise about enabling conditions for shared agency: there are reciprocal, agent-neutral motor representations of outcomes whose obtaining would normally involve action on the part of each agent;
+moreover, these structures of motor representation sometimes facilitate coordination when agents exercise shared agency.
+This needs unpacking. 
+So before considering evidence for this premise (in the next section), let us first explain  it.
 
-A \textit{motor} representation is the sort of representation that enables us to reach for, grasp and transfer objects in a coordinated and fluid way.  
-%Imagine someone who, while remaining seated, moves an object from her right to her left, passing it between her hands as she does so. 
-%The transition between hands is not something she needs to think or reason about. 
-%But if the transition occurs without any pause---if she does not have to wait before releasing the object from her right hand into her left in order to check that the object is securely in her left hand---this is a sign that her action involves predicting the position of her left hand.
-%***continue this example?
-We follow several psychologists and philosophers in supposing that motor representations feature in planning and monitoring action \citep[e.g.][]{wolpert:1995internal, miall:1996_forward}.
-[*HY asks why focus here on forward models]
+A \textit{motor} representation is the sort of representation that enables agents to reach for, grasp and transfer objects in a coordinated and fluid way. 
+Motor representations play a key role in monitoring and planning actions  \citep[e.g.][]{wolpert:1995internal, miall:1996_forward,Wilson:2005qu}.
+Unlike intentions and related mental states, motor representations are usually involved in planning actions over milliseconds rather than minutes or even years.
 
-Motor representations are in some ways similar to intentions,
-as some have recently argued (\citealp[pp.\ 189-90]{pacherie:2008_action}; \citealp{butterfill:2012_intention}).
-First, like intentions, some motor representations represent outcomes and not merely kinematic features of action.
-Like intentions, some motor representations play a role in coordinating multiple  component activities by virtue of their role as elements in hierarchically structured plans.
-And, like intentions, some motor representations coordinate these activities in a way that would normally facilitate the outcome’s occurrence.
+Some motor representations resemble intentions  in 
+	representing outcomes (rather than merely kinematic or dynamic features of action), 
+	coordinating multiple  component activities by virtue of their role as elements in hierarchically structured plans,
+	and coordinating these activities in a way that would normally facilitate the represented outcome's occurrence (\citealp{hamilton_action_2008}; \citealp[pp.\ 189-90]{pacherie:2008_action}; \citealp{butterfill:2012_intention}).
+Despite these points of similarity, 
+motor representations can be distinguished from intentions and other action-related intentions,
+	 as well as from perceptual representations,
+	 by their representational format \citep{butterfill:2012_intention}.
 
-One major dissimilarity between intentions and motor representations is that motor representations lead a kind of double life.
+Motor representations lead a kind of double life.
 For motor representations are involved not only in producing actions but also in observing actions. 
-Indeed there seem to be some striking similarities between the sorts of processes and representations usually involved in performing a particular action and those which typically occur when observing someone else perform that action \citep{rizzolatti_functional_2010,rizzolatti_mirrors_2008}.%
+Indeed there are some striking similarities between the sorts of processes and representations usually involved in performing a particular action and those which typically occur when observing someone else perform that action.
+In some cases it is almost as if the observer were planning the observed action, only to stop just short of performing it herself.%
 \footnote{ 
+For reviews, see \citet{jeannerod_motor_2006,rizzolatti_mirrors_2008,rizzolatti_functional_2010}.
 If motor representations occur in action observation, then observing actions might sometimes facilitate performing compatible actions and interfere with performing incompatible actions.
@@
 }
+When motor representations of outcomes trigger a planning-like process in action observation, this may allow the observer to predict others' actions \citep{Flanagan:2003lm,ambrosini:2011_grasping,ambrosini:2012_tie,Costantini:2012fk}.
 
-[*HY: I'm still thinking about how best to introduce motor representation: any pointers?]
+A representation (motor or not) is \emph{agent-neutral} if its content does not specify any particular agent or agents.%
+\footnote{
+Our use of the term `agent-netural' to describe motor representations bears no relation to the use of the same term  to describe reasons \citep[on the latter, see][]{Parfit:1984fk}.
+}
+To illustrate, agent neutral representations are sometimes found at the early stages of planning.
+Imagine that you and some friends are tasked with preparing a holiday.  
+You might first write down a plan of action without specifying who will act; the plan simply describes what is to be done.
+The plan  will eventually be implemented by you and your friends
+ but this is not written in plan itself  and so it is agent-neutral.
+Of course the fact that this plan is your collective plan may be represented elsewhere; this fact may also be implicit in  the plan's being stapled to the door of your communal kitchen.
+The agent-neutrality of a representation does not require that the agents are nowhere specified, only that they are not specified in the content of the representation.%
+\footnote{
+Strictly the following argument does not hinge on the agent-neutrality of representations.
+It is sufficient for our purposes that there are reciprocal motor representations  of outcomes whose obtaining would normally involve action on the part of each of the reciprocating agents.
+In principle such representations could have contents which specify particular agents. 
+However, we focus on agent-neutral representations to show that our view is consistent with the possibility that reciprocal motor representations are agent-neutral. 
+}
 
-So much for the bare notion of motor representation.
-Why suppose that motor representation might be needed to explain shared agency?
+Two or more agents have \emph{reciprocal} representations (motor or not) just if there is a single outcome and each agent has a motor representation of that outcome. 
+Reciprocal representations can occur both when one agent observes another acting alone and also when two or more agents act together.
+Where one agent observes another acting alone,
+it is possible that both have motor representations of an outcome to which the agent's action is directed.
+For the agent, this representation plays a role in planning and monitoring; for the observer, this representation may facilitate prediction and recognition.
+This is one case of reciprocal motor representation---or, as it is more usually called, mirroring.
+But our main interest is in a different case,
+one where two or more agents are acting together,
+ there is a single outcome to which their actions are directed,
+ and the agents each have a motor representation of this outcome.
+In this case of reciprocal motor representation, the agents each represent an outcome whose obtaining would normally involve not only their own actions but also those of other agents.%
+\footnote{ 
+	Note that our having reciprocal motor representations would not by itself imply that we know that the represented outcome's obtaining would normally involve actions other than our own.
+}
+So when we act together, some of my motor representations may concern outcomes that are partly but not entirely to be realised by my actions; and likewise for you.
+To illustrate, suppose our task is to move an object from A to B, where you pick it up and pass it to me so that I can then place it.
+In this case I may represent the movement of the object from A to B and not only the component movements.
+I represent a collective outcome of our actions and not just outcomes to which each of our actions are individually directed.
+%
+%So it is not just that, when we act together, I represent the outcomes to which your actions are directed and the outcomes to which my actions are directed; I may also represent an outcome to which all of our actions are directed.
+
+We shall use the term \emph{shared} motor representation as an abbreviation for the reciprocal, agent-neutral motor representation of outcomes whose obtaining would normally require action on the part of each reciprocating agent.
+Note that shared motor representations are shared only in the sense in which two people can share a name.  
+(We do not mean to suggest that they are shared in the sense in which two people can share a parent: there need be no representation with two or more subjects.)
+Note also that a shared motor representation is not a special kind of motor representation: it is merely a structure of ordinary motor representations. 
+At this stage we are not strictly entitled use the term `shared motor representation' as an abbreviation.
+After all, we have yet to show that these structures of motor representation play a constitutive role in explaining shared agency, or even an enabling role; 
+we have yet to review evidence that they exist.
+So the choice of term merely reflects our aim.
 
@@
 
-\section{Social motor representation}
-We start from an empirical premise which is about enabling conditions for shared agency only: some joint actions are facilitated by reciprocal, agent-neutral motor representations of outcomes whose obtaining would normally involve action on the part of each agent.
-
-This needs unpacking. 
-As already mentioned, a \textit{motor} representation is the sort of representation that enables us to reach for, grasp and transfer objects in a coordinated and fluid way.
-To repeat, motor representations are not concerned with merely kinematic or dynamic features of actions only.
-Rather, some motor representations represent outcomes, such as the movement of a target object from one place to another.
-
-A representation (motor or not) is \emph{agent-neutral} if its content does not specify any particular agent or agents.%
-\footnote{
-Our use of the term `agent-netural' to describe motor representations bears no relation to the use of the same term  to describe reasons \citep[on the latter, see][]{Parfit:1984fk}.
-}
-To illustrate, agent neutral representations are sometimes found at the early stages of planning.
-Imagine that you and some friends are tasked with preparing a holiday.  
-You might first write down a plan of action without specifying who will act; the plan simply describes what is to be done.
-The plan  will eventually be implemented by you and your friends
- but this is not written in plan itself  and so it is agent-neutral.
-Of course the fact that this plan is your collective plan may be represented elsewhere; this fact may also be implicit in  the plan's being stapled to the door of your communal kitchen.
-The agent-neutrality of a representation does not require that the agents are nowhere specified, only that they are not specified in the content of the representation.%
-\footnote{
-Strictly the following argument does not hinge on the agent-neutrality of representations.
-It is sufficient for our purposes that there are reciprocal motor representations  of outcomes whose obtaining would normally involve action on the part of each of the reciprocating agents.
-In principle such representations could have contents which specify other agents or multiple agents. 
-We focus on agent-neutral representations to show that our view is consistent with the possibility that reciprocal motor representations are agent-neutral. 
-}
-
-Two or more agents have \emph{reciprocal} motor representations  just if there is a single outcome and each agent has a motor representation of that outcome. 
 It is hardly controversial that reciprocal motor representations exist, for their existence is suggested by by a large body of research on motor cognition in action observation.
@@
 
-Suppose that Aravinda runs the trains and Gerhard  the busses.
-Let us stipulate that we measure the extent to which the two services, train and bus, are coordinated is to be measured by the sum of all passengers' total journey times.
-So new timetables would result in better coordination than the current timetables just if the new timetables would decrease the sum of passengers' total journey times.
-To illustrate, improving coordination might involve having trains arrive at a station shortly before, rather than shortly after, busses depart from there.
-Now Aravinda and Gerhard each intend they, Aravinda and Gerhard, coordinate the trains with the busses  to the greatest extent possible given other constraints; and they intend to do this by way of these intentions and meshing subplans of them, and this is common knowledge between them.
+In small city state, Aravinda organises the trams and Gerhard the busses.
+Each is responsible for devising and updating the timetable for the service.
+Let us stipulate that the extent to which the tram and bus services are coordinated is to be measured by the sum of all passengers' total journey times.
+So a change to the timetables would result in better coordination  just if it would decrease the sum of passengers' total journey times.
+To illustrate, improving coordination might involve having trams arrive at a stop shortly before, rather than shortly after, busses depart from there.
+Now Aravinda and Gerhard each intend they, Aravinda and Gerhard, coordinate the trams with the busses  to the greatest extent possible given other constraints; and they intend to do this by way of these intentions and meshing subplans of them, and this is common knowledge between them.
 Aravinda and Gerhard thus meet Bratman's sufficient conditions for them to have a shared intention.
 Acting on this intention, 
-every January Aravinda updates the train timetables and sends Gerhard the changes.
+every January Aravinda updates the tram timetables and sends Gerhard the changes.
 Likewise, Gerhard  updates the bus timetables every June and sends Aravinda the changes.
@@
 First, it is unlikely to be optimal in the sense of resulting, eventually, in a combination of plans such that no other combination of plans would have been better for at least one service and no worse for either service.
-Second, it is unlikely to be efficient in the sense of allowing Gerhard and Aravinda to arrive at an optimal combination of plans for the two services, trains and buses, with the fewest iterations.
+Second, it is unlikely to be efficient in the sense of allowing Gerhard and Aravinda to arrive at an optimal combination of plans for the two services, trams and buses, with the fewest iterations.%
+% --- I removed this footnote because the case is a bit different
+% --- In Konvalinka, the subjects have to act simultaneously so they can't wait to see what the other does
+%\footnote{
+%The limits of this type of coordination, where each agent tries to predict the actions of the other, is nicely illustrated by the `two-way interaction' condition of an experiment by \citet{konvalinka:2010_follow}.
+%When subjects were required to synchronise the timing of finger taps with each other in this condition, 
+%%each had to predict the others' action and modify their own accordingly. 
+%they ended up with a pattern of errors `oscillating in opposite directions' as each attempted to move towards the other (p.\ 2228).
+%}
 
 How could they do better?
-One possibility may be to have a single person planning the busses and trains---perhaps, for example, Aravinda could buy Gerhard's franchise. 
+One possibility may be to have a single person planning the busses and trams---perhaps, for example, Aravinda could buy Gerhard's franchise. 
 But suppose that this is not possible, and that there are limits on how much information Aravinda and Gerhard can share. 
 Then another way that might improve things would be to have Aravinda and Gerhard each plan everything.
-So Aravinda would plan the best timetable for all the trains and busses, and Gerhard would do the same.
-Since Aravinda only controls the trains, she can only implement the train-related part of this master timetable.
+So Aravinda would plan the best timetable for all the trams and busses, and Gerhard would do the same.
+Since Aravinda only controls the trams, she can only implement the tram-related part of this master timetable.
 Likewise, Gerhard can only implement the bus part of his master timetable.
@@
 Then having each agent plan the whole joint action means that (i) each agent plans the other agent's action,
-(ii) each agent's plan for the other agent's action is approximately the same as that agent's plan for her own action,
+(ii) each agent's plan for the other agent's action is approximately the same as that agent's plan for her own action [***AMBIGUOUS: `that agent's plan' must refer to the other agent, not `each agent',
 and 

### 2012-10-28 counterexample to bratman on shared intention/a counterexample to bratman on shared intention.tex
on counterexample to bratman: just about to remove cooperatively neural section

 %!TEX TS-program = xelatex
%!TEX encoding = UTF-8 Unicode

\def \papersize {a4paper}

\documentclass[12pt,\papersize]{extarticle}
% extarticle is like article but can handle 8pt, 9pt, 10pt, 11pt, 12pt, 14pt, 17pt, and 20pt text

\def \ititle {Is This a Counterexample to Bratman on Shared Intention?}
\def \isubtitle {}
\def \iauthor {Stephen A.\ Butterfill}
\def \iemail{s.butterfill@warwick.ac.uk}
%\date{}

\input{$HOME/Documents/submissions/preamble_steve_paper3}
%\author{}
%\date{}

%\setromanfont[Mapping=tex-text]{Sabon LT Std}

\begin{document}

\setlength\footnotesep{1em}

\bibliographystyle{$HOME/Documents/submissions/mynewapa} %apalike

\maketitle
%\tableofcontents
\title{}

\begin{abstract}
\noindent
***

\end{abstract}

\section{Shared Intention}
Why, if at all, is a notion of shared intention needed? 
This question is standardly answered by appeal to contrast cases \citep[compare][p.\ 150]{Bratman:2009lv}.
Thus \citet{gilbert_walking_1990} contrasts our intentionally walking together with two people who happen to be walking side by side. 
The two pairs' movements may be the same and similarly coordinated (to avoid collision), so intentionally walking together cannot be only a matter of how we move our bodies or how our movements are coordinated. 
Also, in both cases each individual's walking is intentional, so our intentionally walking together cannot be a matter only of our each intentionally walking.%
\footnote{
Compare \citet{Pears:1971fk} who uses contrast cases to argue that whether something is an ordinary, individual action depends on its antecedents. 
}  
Perhaps, then, a notion of shared intention is needed to distinguish the two cases.
It is our acting on a shared intention that we walk together which distinguishes us from two strangers who happen to be walking side by side.%
\footnote{
Many philosophers agree that a notion of shared intention is useful for understanding acting together. 
Compare \citet[p.\ 5]{Gilbert:2006wr}: `I take a collective action to involve a collective intention.'  See also  
	\citet[p.\ 381]{Carpenter:2009wq}, 
	\citet[p.\ 369]{Call:2009fk}, 
	\citet{Kutz:2000si}, 
	\citet[p.\ 117]{rakoczy_pretend_2006} and 
	\citet{Tollefsen:2005vh}.
	}

But what could shared intention be?
In an influential series of papers,\footnote{ 
See \citet{Bratman:1992mi,Bratman:1993je,Bratman:1999fr,Bratman:2009lv}.
For influences beyond philosophy, see e.g.\ \citet{Tomasello:2005wx} and \citet{Knoblich:2008hy}. 
}
Bratman claims that the following are collectively sufficient\footnotemark \ conditions for you and I to have a shared intention that we J:
%
\footnotetext{
In \citet{Bratman:1992mi}, the following were offered as jointly sufficient \textit{and individually necessary} conditions; the retreat to sufficient conditions occurs in \citet[][pp.\ 143-4]{Bratman:1999fr} where he notes that `for all that I have said, shared intention might be multiply realizable.'
} 
%
\begin{quote}
\label{quote:bratman_account}
`1. (a) I intend that we J and (b) you intend that we J

`2. I intend that we J in accordance with and because of la, lb, and meshing subplans of la and lb; you intend that we J in accordance with and because of la, lb, and meshing subplans of la and lb

`3. 1 and 2 are common knowledge between us' \citep[][p.\ View 4]{Bratman:1993je}
\end{quote}
%
Care is needed in specifying the contents of the intentions concerning our J-ing in these clauses. 
As mentioned above, an appeal to shared intention is supposed to  characterise systematically a difference between two agents who intentionally J together (e.g.\ walk  together) and two agents who also  J together but do not do so intentionally (e.g.\ they happen to be walking side by side). 
So we must avoid tacitly appealing to this distinction by  restricting possible values of J to those which involve our intentionally doing something together. 
Rather the above  conditions, (1)--(3), must be sufficient for shared intention even for some values of J which are `neutral with respect to shared intentionality'.%
\footnote{
 \citet[p.\ 147]{Bratman:1999fr}.
 This refines Bratman's earlier view that some admissable values of J are cooperatively neutral 
 	where an  act-type is \emph{cooperatively neutral} just if `joint performance of an act of that type may be cooperative, but it need not be' \citep[p.\ 330]{Bratman:1992mi}. 
}

A consequence is that, in the right situations, one of us can rationally intend that we J unilaterally, that is without anyone else having this intention. 
If you know that I am going to Chicago via a certain route at a particular time
and if you can rationally intend that you go to Chicago in the same manner,
then you can rationally intend that we go to Chicago together---providing, of course, that you intend our going to Chicago together in a way that is neutral with respect to shared intentionality.
What follows depends on the premise that this is indeed possible.%
\footnote{ 
\citet{Bratman:1999fr}  defends this claim at length. 
Note also that a primary application of his account of shared intention depends on its truth.
}

In this paper we give a counterexample to Bratman's  claim that the above conditions, (1)--(3), are collectively sufficient conditions for shared intention. 
We will also suggest a revision which would enable the counterexample to be avoided.
Apart from improving our understanding of what shared intention is, 
this will also bear on the kinds of planning that are involved in acting together.

Before going further we must distinguish two versions of the claim that (1)--(3) are collectively sufficient for shared intention.
The \emph{weak claim} is that there is some J such that these conditions are sufficient for you and I to intend that we J.
The \emph{strong claim} is that for any J, these conditions are sufficient for you and I to intend that we J.
One of Bratman's aims is to show that it is possible to give an account of shared intention using concepts that `are available within the theory of individual planning agency' \citep[p.\ 163]{Bratman:2009lv}.  
Achieving this aim would require the strong claim, and it is to the strong claim that our counterexample is directed.

\section{Background: Cooperatively Neutral Act-types}
When Elisabeth and her daughters go to the park together, 
	they usually do so cooperatively 
	but sometimes she is coerced by threats and tantrums.  
(Bratman's example is going to Chicago `in the mafia sense', but we are  more familiar with domestic coercion.)
Observing them one day, we might 
	know that Elisabeth and her daughters are going to the park together 
	without knowing whether this particular trip is cooperative or coercive.
So we can think of this action in a way that is cooperatively neutral.
In general an act-type is \emph{cooperatively neutral} just if `joint performance of an act of that type may be cooperative, but it need not be' \citep[p.\ 330]{Bratman:1992mi}.

What an agent intends when then she intends that we J in the first of Bratman's sufficient conditions, (1), can be a cooperatively neutral act type. 
Bratman allows for this possibility in order to avoid potentially circular appeal to shared intention in specifying what we intend when we intend that we J.

\section{The Counterexample}
Although it is simple, introducing the counterexample will involve several steps.

Ayesha and Benji  are playing a simple video game which involves moving a cross around a two-dimensional space littered with barriers.
Ayesha can only move the cross backwards or forwards,
while Benji can only move it left or right.
Ayesha and Benji are given tasks independently. 
These tasks always involve making the cross touch a target within 100 seconds of starting. 
A player succeeds when the cross touches her target, regardless of what happens to the cross afterwards.  
(It does not have to stay touching the target.)
For example, 
	Ayesha's task might be to make the cross touch the red square
	while
	Benji's task might be to make the cross touch the blue circle. 
It is possible that either or both succeed, or that they both fail.
Ayesha and Benji always have common knowledge of the tasks they are assigned.
Each movement carries a small cost to the player who moves, so that Ayesha and Benji each attempt to minimize how much he or she moves the cross consistently with completing his or her task.
Ayesha and Benji are each neutral on whether the other succeeds or fails.
They are not opponents and do not seek to undermine each other's efforts, but each is entirely concerned  with his or her own task.

In the previous round, 
	Ayesha's task was to touch the red square
	and
	Benji's task was to touch the blue circle.
As it happened, the red square and the blue circle were positioned such that it was impossible to touch one without touching the other, as Ayesha and Benji knew.
Furthermore, Ayesha intended that they, Ayesha and Benji, move the cross to the red square; Benji intended that they, Ayesha and Benji, move the cross to the blue circle; and each intended that they do so in accordance with and because of these intention and meshing subplans of them; and this was all common knowledge between them.
In this way, Ayesha and Benji almost met Bratman's conditions for shared intention.
They would have met the conditions but for the fact that Ayesha intended that they J$_1$ whereas Benji intended that they J$_2$ where J$_1$ $\neq$ J$_2$.

At first glance this combination of intentions may seem incoherent or irrational. 
How could Ayesha rationally intend that they, Ayesha and Benji, J$_1$ knowing that Benji has no corresponding intention? 
To see that this is possible, note that, on Bratman's view,
 the act types which are intended can be \emph{cooperatively neutral}, that is such that `joint performance of an act of that type may be cooperative, but it need not be' \citep[p.\ 330]{Bratman:1992mi}.
(This enables Bratman to avoid circularity in specifying what a shared intention is.)  
So Ayesha's knowledge that Benji intends that they J$_2$ together with her knowledge that their J$_2$-ing is sufficient for their J$_1$-ing makes it possible for her to intend, unilaterally, that they J$_1$.%
\footnote{
These issues are discussed at length in \citet{Bratman:1999fr}. 
Note that Bratman's account requires that it be possible for Ayesha to intend, unilaterally, that they J$_1$.%
}

\section{Prior counterexamples}
***Must mention that Tollefsen and Gold \& Sugden counterexamples fail.

\section{The Counterexample}
Suppose that Aravinda runs the trains and Gerhard  the busses.
Let us stipulate that the extent to which the two services, train and bus, are coordinated is to be measured by the total time passengers spend waiting between a bus and a train.
So changes to the timetables would result in better coordination just if the changes would total waiting time.
To illustrate, improving coordination might involve having trains arrive at a station shortly before, rather than shortly after, busses depart from there.
Now Aravinda and Gerhard each intend that they, Aravinda and Gerhard, coordinate the trains with the busses to the greatest extent possible given other constraints; and they intend to do this by way of these intentions and meshing subplans of them, and this is common knowledge between them. 
Aravinda and Gerhard thus meet Bratman's sufficient conditions for them to have a shared intention.
Acting on this intention, 
every January Aravinda updates the train timetables and sends Gerhard the changes.
Likewise, Gerhard  updates the bus timetables every June and sends Aravinda the changes.
In this way  each is responsive to the other's intentions and subplans of these. 
Indeed, each may even try to predict changes in the other's subplans and modify their own accordingly.  
But from each individual's point of view, the other's plans are  merely a constraint.

***

So whether Gerhard succeeds relative to his intention depends on both his own and Aravinda's plans; and likewise for Aravinda.
%This is why each intends that they optimally coordinate the services in accordance with meshing subplans.
This is why each is responsive to the other's plans. 
Indeed, each may even try to predict changes in the other's plans and modify their own accordingly.  
But from each individual's point of view, the other's plans are  merely a constraint.
This approach to planning suffers from two defects.
First, it is unlikely to be optimal in the sense of resulting, eventually, in a combination of plans such that no other combination of plans would have been better for at least one service and no worse for either service.
Second, it is unlikely to be efficient in the sense of allowing Gerhard and Aravinda to arrive at an optimal combination of plans for the two services, trains and buses, with the fewest iterations.
How could they do better?
One possibility may be to have a single plan covering busses and trains---perhaps, for example, Aravinda could buy Gerhard's franchise. 
%But suppose that this is not possible, and that there are limits on how much information Aravinda and Gerhard can share. 
Now there is a single goal to which Aravinda and Gerhard's activities are both directed.
But suppose that this is not possible, and that they are unable to integrate their planning more tightly---practical constraints or regulation prevents them from opening up their planning to each other.
Can they still do better than treat each other's plans as a constraint?
Possibly.
Each plans the whole bus--train operation.

New contrast case: Aravinda and Gerhard planning the whole thing vs. merely responding to each other's plans.
Meet Bratman's sufficient conditions for shared intention in both cases.
(Specifically, in both cases they act on a shared intention that they coordinate the trains with the busses to the greatest extent possible given other constraints.)
But intuitively is a case of coordinating the trains with the busses together whereas the other case involves Aravinda and Gerhard merely doing this in parallel.
So shared intention is insufficient to fully explain the contrast cases.

I'm also inclined to think it is not necessary for solving the contrast cases (distributive goals ...\ collective goals).
The contrast cases do not provide a firm anchor for theorising about shared intention.

\bibliography{$HOME/endnote/phd_biblio}

\end{document}

### 2012-10-28 counterexample to bratman on shared intention/a counterexample to bratman on shared intention.tex
adding pure unshared intention

@@
 Thus \citet{gilbert_walking_1990} contrasts our intentionally walking together with two people who happen to be walking side by side. 
-The two pairs' movements may be the same and similarly coordinated (to avoid collision), so intentionally walking together cannot be only a matter of how we move our bodies or how our movements are coordinated.
-Also, in both cases each individual's walking is intentional, so our intentionally walking together cannot be a matter only of our each intentionally walking.%
+And \citet{Searle:1990em} contrasts park visitors who  simultaneously run to a central shelter in performing a dance with park visitors who likewise run to the central shelter but only because of an impending storm.
+These and other contrast cases invite the question, 
+How does our intentionally acting together differ from our merely acting in parallel? 
+The first contrast case, Gilbert's, shows that the difference can’t be just  a matter of coordination because people who are merely happen to be walking side by side each other also need to coordinate their actions in order to avoid colliding.
+Note also that in both cases each individual's walking is intentional, so our intentionally walking together cannot be a matter only of our each intentionally walking.
+The second contrast case, Searle's, shows that the difference can’t just be that the resulting actions have a common effect because merely parallel actions can have common effects too.%
 \footnote{
-Compare \citet{Pears:1971fk} who uses contrast cases to argue that whether something is an ordinary, individual action depends on its antecedents. 
+This use of contrast cases resembles \citet{Pears:1971fk}: he uses contrast cases to argue that whether something is an ordinary, individual action depends on its antecedents. 
 } 
 Perhaps, then, a notion of shared intention is needed to distinguish the two cases.  
-It is our acting on a shared intention that we walk together which distinguishes us from two strangers who happen to be walking side by side.%
+Perhaps it is our acting on a shared intention that we walk together which distinguishes us from two strangers who happen to be walking side by side.%
 \footnote{
@@
 
-A consequence is that, in the right situations, one of us can rationally intend that we J unilaterally, that is without anyone else having this intention. 
-If you know that I am going to Chicago via a certain route at a particular time
-and if you can rationally intend that you go to Chicago in the same manner,
-then you can rationally intend that we go to Chicago together---providing, of course, that you intend our going to Chicago together in a way that is neutral with respect to shared intentionality.
+A consequence is that, in the right situations, one of us can rationally intend that we J, and can intend this unilaterally, that is without depending on anyone else intending that we J. 
+Suppose you know that I am going to Chicago via a certain route at a particular time, 
+and that I will do this regardless of what you do.
+Suppose also that you can rationally intend that you go to Chicago in the same manner.
+Then you can rationally intend that we go to Chicago together, 
+	and you can intend this irrespective of whether I have any corresponding intention---providing, of course, that in so intending you are conceiving of our going to Chicago together in a way that is neutral with respect to shared intentionality.
 What follows depends on the premise that this is indeed possible.%
@@
 \citet{Bratman:1999fr}  defends this claim at length. 
-Note also that a primary application of his account of shared intention depends on its truth.
+Note also that this claim must be true if Bratman's account of shared intention is to provide an informative way of distinguishing between the contrast cases mentioned at the start.
 }
@@
 In this paper we give a counterexample to Bratman's  claim that the above conditions, (1)--(3), are collectively sufficient conditions for shared intention. 
-We will also suggest a revision which would enable the counterexample to be avoided.
+We shall also suggest a revision which would enable the counterexample to be avoided.
 Apart from improving our understanding of what shared intention is, 
@@
 
-\section{Background: Cooperatively Neutral Act-types}
-When Elisabeth and her daughters go to the park together, 
-	they usually do so cooperatively 
-	but sometimes she is coerced by threats and tantrums.  
-(Bratman's example is going to Chicago `in the mafia sense', but we are  more familiar with domestic coercion.)
-Observing them one day, we might 
-	know that Elisabeth and her daughters are going to the park together 
-	without knowing whether this particular trip is cooperative or coercive.
-So we can think of this action in a way that is cooperatively neutral.
-In general an act-type is \emph{cooperatively neutral} just if `joint performance of an act of that type may be cooperative, but it need not be' \citep[p.\ 330]{Bratman:1992mi}.
+\section{Unshared Intentions}
+Before we can introduce the counterexample we need to introduce what we shall call `unshared intentions'.
+We have an \emph{unshared intention} that we <J$_1$, J$_2$> where J$_1$$\neq$J$_2$ just if:
+%
+\begin{quote}
+\label{df:unshared_intention}
+1. (a) I intend that we J$_1$ and (b) you intend that we J$_2$
 
-What an agent intends when then she intends that we J in the first of Bratman's sufficient conditions, (1), can be a cooperatively neutral act type. 
-Bratman allows for this possibility in order to avoid potentially circular appeal to shared intention in specifying what we intend when we intend that we J.
+2.
+I intend that we J$_1$ in accordance with and because of la, lb, and meshing subplans of la and lb; you intend that we J$_2$ in accordance with and because of la, lb, and meshing subplans of la and lb
 
-\section{The Counterexample}
-Although it is simple, introducing the counterexample will involve several steps.
+3. 1 and 2 are common knowledge between us.
+\end{quote}
+In defining unshared intention we have used conditions exactly like Bratman's sufficient conditions for shared intention except that Bratman's conditions have J$_1$$=$J$_2$.
+At this point it might be natural for readers to suppose that agents could not have unshared intentions,
+or at least that they could not do so without irrationality.
+In this section we describe a possible situation in which two agents  have an unshared intention without irrationality, deception or even ignorance.
 
+Let us first introduce the activity we shall focus on.
 Ayesha and Benji  are playing a simple video game which involves moving a cross around a two-dimensional space littered with barriers.
-Ayesha can only move the cross backwards or forwards,
-while Benji can only move it left or right.
-Ayesha and Benji are given tasks independently. 
-These tasks always involve making the cross touch a target within 100 seconds of starting. 
-A player succeeds when the cross touches her target, regardless of what happens to the cross afterwards.  
-(It does not have to stay touching the target.)
-For example, 
-	Ayesha's task might be to make the cross touch the red square
+Ayesha can only accelerate the cross backwards or forwards,
+while Benji can only accelerate it left or right. 
+The cross moves around and interacts with the barriers in ways both players can predict.
+The players are given tasks independently. 
+These tasks always involve making the cross hit a target within two minutes of starting. 
+A player succeeds when the cross hits her target, regardless of what happens to the cross afterwards.  
+(It may go on to hit another target.)
+In this case, 
+	Ayesha's task is to make the cross hit the red square
 	while
-	Benji's task might be to make the cross touch the blue circle. 
-It is possible that either or both succeed, or that they both fail.
-Ayesha and Benji always have common knowledge of the tasks they are assigned.
+	Benji's task is make the cross hit the blue circle. 
+In general it is possible that either or both will succeed, or that they will both fail.
 Each movement carries a small cost to the player who moves, so that Ayesha and Benji each attempt to minimize how much he or she moves the cross consistently with completing his or her task.
@@
 They are not opponents and do not seek to undermine each other's efforts, but each is entirely concerned  with his or her own task.
+All of this is common knowledge for Ayesha and Benji.
+They both know who has which task, what constraints they face and what their motives are.
 
-In the previous round, 
-	Ayesha's task was to touch the red square
-	and
-	Benji's task was to touch the blue circle.
-As it happened, the red square and the blue circle were positioned such that it was impossible to touch one without touching the other, as Ayesha and Benji knew.
-Furthermore, Ayesha intended that they, Ayesha and Benji, move the cross to the red square; Benji intended that they, Ayesha and Benji, move the cross to the blue circle; and each intended that they do so in accordance with and because of these intention and meshing subplans of them; and this was all common knowledge between them.
-In this way, Ayesha and Benji almost met Bratman's conditions for shared intention.
-They would have met the conditions but for the fact that Ayesha intended that they J$_1$ whereas Benji intended that they J$_2$ where J$_1$ $\neq$ J$_2$.
+Consider the possibility of one player intending, unilaterally, that the two players do something.
+Suppose that one of the players---Ayesha, say---can knowledgeably predict that if she performs a certain sequence actions, <a$_1$, a$_2$, ...\ a$_n$>, then Benji will simultaneously perform certain other actions, <b$_1$, b$_2$, ...\ b$_n$>,
+ and the upshot will be that the cross hits the red square.
+Were this to happen, we could say, truthfully, that Ayesha and Benji made the cross hit the red square.
+Suppose, further, that Ayesha can intend to perform those actions <a$_1$, a$_2$, ...\ a$_n$>.
+Then Ayesha can intend, unilaterally, that they, Ayesha and Benji, make the cross hit the red square.
 
-At first glance this combination of intentions may seem incoherent or irrational. 
-How could Ayesha rationally intend that they, Ayesha and Benji, J$_1$ knowing that Benji has no corresponding intention? 
-To see that this is possible, note that, on Bratman's view,
- the act types which are intended can be \emph{cooperatively neutral}, that is such that `joint performance of an act of that type may be cooperative, but it need not be' \citep[p.\ 330]{Bratman:1992mi}.
-(This enables Bratman to avoid circularity in specifying what a shared intention is.)  
-So Ayesha's knowledge that Benji intends that they J$_2$ together with her knowledge that their J$_2$-ing is sufficient for their J$_1$-ing makes it possible for her to intend, unilaterally, that they J$_1$.%
-\footnote{
-These issues are discussed at length in \citet{Bratman:1999fr}. 
-Note that Bratman's account requires that it be possible for Ayesha to intend, unilaterally, that they J$_1$.%
-}
+Unshared intentions require a kind of symmetry.
+Let us suppose that the above sequences of actions,
+	Ayesha's <a$_1$, a$_2$, ...\ a$_n$>  and 
+	Benji's <b$_1$, b$_2$, ...\ b$_n$>,
+will also result in the cross hitting the blue circle. 
+(Since the cross has momentum, we can suppose that it will hit both the red square and the blue circle at some time after these actions are performed.)
+Then by the reasoning just offered, Benji could intend that they, Ayesha and Benji make the cross hit the blue circle.
+So Ayesha and Benji could meet the first condition, (1), for having an unshared intention.
+
+What about the second condition, (2)?
+Suppose that Ayesha knows two further things.
+First, that Benji intends that they, Ayesha and Benji, make the cross hit the blue circle.
+Second, that in acting on his intention Benji will perform actions <b$_1$, b$_2$, ...\ b$_n$>.
+Then Ayesha can intend that they, Ayesha and Benji, make the cross hit the red square in accordance with and because of her intention that they make the cross hit the red square and in accordance with and because of Benji's intention that they make the cross hit the blue circle.
+
+This is not quite enough to meet the second condition, (2), because there is also a requirement about meshing subplans. 
+To make sense of this requirement we need to generalise Bratman's definition of meshing:
+\begin{quote}
+`our individual subplans concerning our J-ing \emph{mesh} just in case there is some way we could J that would not violate either of our subplans but would, rather, involve the successful execution of those subplans' \citep[p.\ 106]{Bratman:1993je}.
+\end{quote}
+A natural generalisation is this:
+\begin{quote}
+our individual subplans concerning our <J$_1$, J$_2$>-ing \emph{mesh} just in case there is some way we could <J$_1$, J$_2$> that would not violate either of our subplans but would, rather, involve the successful execution of those subplans.
+\end{quote}
+%
+To illustrate, 
+there would be a failure to mesh if, in intending that they make the cross hit the red square,  Ayesha's plans had included pushing Benji out of the way and seizing his controls. 
+But in the case we have been describing there is no such failure to mesh.
+Each agent's subplans involve manipulating his or her own controls,
+and the successes each seeks in doing this depends on the other successfully carrying out their subplans.
+So Ayesha can rationally intend that they, Ayesha and Benji, make the cross hit the red square in accordance with and because of their intentions and meshing subplans of them.
+And Benji likewise for making the cross hit the blue circle.
+
+The only outstanding requirement for Ayesha and Benji to have an unshared intention is that their various intentions are common knowledge. 
+Assuming common knowledge is possible where agents have a shared intention, it is likewise possible in this case of unshared intention.
+So Ayesha and Benji can have an unshared intention that they <J$_1$, J$_2$> where J$_1$ is Ayesha and Benji's making the cross hit the red square and J$_2$ is their making the cross hit the blue circle.
+
+So far we have shown that it is possible for two agents to have an unshared intention without irrationality, deception or ignorance.
+Of course unshared intentions may be rare. 
+But what matters for our counterexample is just that they are possible.
+
+\section{Pure unshared intentions}
+An unshared intention is \emph{pure} if having that unshared intention does not involve having any shared intention. 
+It may initially appear obvious that all unshared intentions are pure. 
+This is less obvious than it seems, however.
+Since Bratman provides only sufficient conditions for shared intention, his account doesn't tell us that the unshared intention is not a shared intention.
+For all Bratman says, the conditions defining unshared intention might also be collectively sufficient conditions for shared intention. 
+And while we might be tempted to assume that for two agents to intentionally J together they must at least each have intentions concerning J, 
+this assumption might reasonably be rejected by those who, like Bratman, reject the Simple View according to which when an individual intentionally J-s she has an intention concerning her J-ing \citep{Bratman:1984jr}.
+So the purity of an unshared intention cannot be assumed without argument.
+
+How could we establish the purity of a particular unshared intention? 
+For reasons that will become clear later, we need to show that  Ayesha and Benji's unshared intention that they <J$_1$, J$_2$>  is pure.
+
+This question might initially appear frivolous but it is not straightforward to answer.
+Since Bratman provides only sufficient conditions for shared intention, his account doesn't tell us that the unshared intention is not a shared intention.
+And while we might be tempted to assume that for two agents to intentionally J together they must at least each have intentions concerning J, 
+this assumption might reasonably be rejected by those who, like Bratman, reject the Simple View according to which when an individual intentionally J-s she has an intention concerning her J-ing \citep{Bratman:1984jr}.
+
+\section{The Counterexample}
+*Need a beefier activity, something more than cooperatively neutral.
+
+In the situation just described,
+Ayesha and Benji are playing a game and have different tasks.
+Thanks to special features of the game environment, they both succeed by acting on an unshared intention.
+Now compare two further players, Yasmin and Zak, who are playing the same game. 
+Their situations, knowledge states, intentions and actions are as similar as possible to Ayesha's and Benji's except for one detail.
+Just by chance they have been assigned identical tasks: Yasmin's task is to make the cross hit the red square and Zak's task is the same.
+So where Ayesha and Benji have an unshared intention that they <J$_1$, J$_2$>,
+Yasmin and Zak meet Bratman's conditions (1)--(3) for having a shared intention that they J$_1$.  
+But these intentions, the unshared intention and the Bratman intention (as we might call the structure of intention and knowledge), result in them performing same actions (Yasmin does what Ayesha does, Zak what Benji does) in the same way.
+
+We shall argue that although Yasmin and Zak meet Bratman's conditions (1)--(3) for having a shared intention that they J$_1$, they do not have any such shared intention.
+This, then, is our counterexample.
+Of course it is probably not yet obvious that Yasmin and Zak do not have a shared intention.
+To see that they do not we need to introduce one more case.
+
+We started this paper by explaining the need for a notion of shared intention by appeal to contrast cases. 
+An account of shared intention should enable us to  distinguish systematically between two types of case in which we J together, one where we intentionally J together and the other where  our J-ing together involves  merely parallel actions.
+As noted, Gilbert's example of walking together shows that merely parallel actions can be mutually responsive, as when two strangers walking side by side avoid collision thanks to meshing subplans.
+
+*Yasmin and Zak: it's like we've tied the Gilbert's walking strangers' ankles together, so that they can only succeed together.
+
+***
+
+Our argument is:
+\begin{enumerate}[label=\roman*]
+\item Yasmin and Zak have a shared intention that they J$_1$ only if Ayesha and Benji have a shared intention in virtue of having the 
+ unshared intention that they <J$_1$, J$_2$>.
+\item Ayesha and Benji's unshared intention that they <J$_1$, J$_2$> is a not shared intention.
+\end{enumerate}
+%
+Therefore:
+%
+\begin{enumerate}[resume,label=\roman*]
+\item Yasmin and Zak do not have a shared intention that they J$_1$.
+\end{enumerate}
+%
+In the rest of this section we defend the two premises, starting with the second, (ii).
+
+Could Ayesha and Benji's unshared intention that they <J$_1$, J$_2$> be a  shared intention?
+This question might initially appear frivolous but it is not straightforward to answer.
+Since Bratman provides only sufficient conditions for shared intention, his account doesn't tell us that the unshared intention is not a shared intention.
+And while we might be tempted to assume that for two agents to intentionally J together they must at least each have intentions concerning J, 
+this assumption might reasonably be rejected by those who, like Bratman, reject the Simple View according to which when an individual intentionally J-s she has an intention concerning her J-ing \citep{Bratman:1984jr}.
+
+In defence of the second premise, (ii), consider that shared intentions are suppose to play a certain functional role.
+Among other things, they are supposed to coordinate
 

### 2012-10-28 counterexample to bratman on shared intention/a counterexample to bratman on shared intention.tex
sunday

@@
 
-\section{Pure unshared intentions}
+\section{Pure unshared intentions and imaginary agents [*cut?]}
 An unshared intention is \emph{pure} if having that unshared intention does not involve having any shared intention. 
 It may initially appear obvious that all unshared intentions are pure. 
-This is less obvious than it seems, however.
+This is not obvious, however.
 Since Bratman provides only sufficient conditions for shared intention, his account doesn't tell us that the unshared intention is not a shared intention.
@@
 
-How could we establish the purity of a particular unshared intention? 
-For reasons that will become clear later, we need to show that  Ayesha and Benji's unshared intention that they <J$_1$, J$_2$>  is pure.
+For reasons to be explained later, we need to show that  Ayesha and Benji's unshared intention that they <J$_1$, J$_2$>  is pure.
+How could we do this? 
+Ayesha sees Benji's actions as constraints on her own, or else as opportunities.
+She exploits Benji's intentions for her own ends.
+Of course the situation is reciprocal: Benji exploits Ayesha in equal measure.
+Each allows himself or herself to be exploited by the other because being exploited enables exploiting.
+The sort of shared agency that an account of shared intention is supposed to capture must involve more than this sort of reciprocal exploitation, where each agent sees the other's actions only as constraints or opportunities.
+Intuitively shared agency must somehow involve greater unity.
+But how can this intuition support an argument?
 
-This question might initially appear frivolous but it is not straightforward to answer.
-Since Bratman provides only sufficient conditions for shared intention, his account doesn't tell us that the unshared intention is not a shared intention.
-And while we might be tempted to assume that for two agents to intentionally J together they must at least each have intentions concerning J, 
-this assumption might reasonably be rejected by those who, like Bratman, reject the Simple View according to which when an individual intentionally J-s she has an intention concerning her J-ing \citep{Bratman:1984jr}.
+We started this paper by explaining the need for a notion of shared intention by appeal to contrast cases (following Bratman). 
+It is contrast cases that are supposed give us a pre-theoretical handle on shared intention.
+So to show that Ayesha and Benji's unshared intention is pure, we need to construct a contrasting case.
+
+***Lily and Isabel, 
+	like many people, 
+ 	sometimes attribute intentions and other states to imaginary agents, and perform actions on their behalf. 
+In some cases this enables them to further their own real-world objectives (as in `Teddy wants to go to the park now').	
+Some of these imaginary agents are toys. 
+But some are imaginary aggregate agents. 
+Sometimes, that is, Lily and Isabel imagine that there is an agent distinct from either of them and such that all its parts are parts of them.
+They attribute intentions and other states to this agent, and act on its behalf.
+
+An account of shared intention should enable us to  distinguish systematically between two types of case in which we J together, one where we intentionally J together and the other where  our J-ing together involves  merely parallel actions.
+As noted, Gilbert's example of walking together shows that merely parallel actions can be mutually responsive, as when two strangers walking side by side avoid collision thanks to meshing subplans.
 
 \section{The Counterexample}
-*Need a beefier activity, something more than cooperatively neutral.
+%*Need a beefier activity, something more than cooperatively neutral.
 
@@
 Yasmin and Zak meet Bratman's conditions (1)--(3) for having a shared intention that they J$_1$.  
-But these intentions, the unshared intention and the Bratman intention (as we might call the structure of intention and knowledge), result in them performing same actions (Yasmin does what Ayesha does, Zak what Benji does) in the same way.
+But these intentions, the unshared intention and the Bratman intention (as we might label the structure of intention and knowledge while leaving open whether it constitutes a shared intention), result in them performing same actions in the same way.
+That is, Yasmin reasons about Zak much as Ayesha reasons about Benji and Yasmin and does what Ayesha does, and likewise for Zak and Benji.
+
+We shall argue that Yasmin and Zak do not have a shared intention that they J$_1$ despite meeting meet Bratman's conditions (1)--(3).
+This is probably not yet obvious, however.
+Let us first attempt to provide some intuitive support for this claim.
+Return to Ayesha and Benji for a moment.
+Ayesha sees Benji's actions as constraints to work around and opportunities to exploit.
+She exploits Benji's intentions for her own ends.
+Of course the situation is reciprocal: Benji exploits Ayesha in equal measure.
+Each allows himself or herself to be exploited by the other because being exploited enables exploiting.
+The sort of shared agency that an account of shared intention is supposed to capture must involve more than this sort of reciprocal exploitation, where each agent merely exploits the other's intentions.
+But Yasmin and Zak are engaged in just the same sort of reciprocal exploitation as Ayesha and Benji.
+The only difference is that, quite by chance, Yasmin's task and Zak's task are the same.
+So, intuitively, Yasmin and Zak, like Ayesha and Benji, belong in the same category as two people who happen to be walking side by side and not in the category exemplified by two people intentionally walking together.
+
+Assuming for a moment that this is right,
+what might explain the difference between Yasmin and Zak on the one hand and cases like Gilbert's two people intentionally walking together on the other hand?
+To answer this question in the spirit of Bratman's planning-theoretic approach, we need to step back and think about planning generally.
+
+Imagine observing Ayesha and Benji as they play the game. 
+Suppose you decide to make a plan of action for one or both of them.
+In doing this you could either adopt one player's perspective or you could adopt a fictional shared perspective.
+Adopting one player's perspective, Ayesha's say, 
+	you think about what she might do to achieve her task,
+	predict how this will influence Benji's actions,
+	and then feed  these predictions back into the question of what Ayesha might do.
+From this perspective, Benji's actions are constraints to work around and opportunities to exploit.
+Alternatively you might adopt a fictional shared perspective, 
+	 mentally replacing Ayesha and Benji with a single, imaginary aggregate agent who has to perform both tasks alone.
+	 In this case, you think about what combination of actions would best achieve both tasks and then, right at the end, you assign these actions to Ayesha and Benji.
+From this perspective, it is as if there is a single mind behind the plan.
+
+Note that the number of agents involved does not dictate which perspective we can take.
+Just as we could take a fictional shared perspective where two or more agents are involved, 
+so also we could take a fictional split perspective on the actions of  a single agent, and even of ourselves.
+The newly elected  politician with a young family and working partner might see herself almost as if she were two people engaged in different projects while forced to shared a single body.
+
+Shared perspective of a fictional aggregate agent.
+Can be a matter of negotiation what the fictional aggregate agent's values, preferences and intentions are.
+
+Perhaps the reason why Yasmin and Zak fail to have a shared intention (if they do---this has not yet been argued) is that each plans entirely from her own perspective. 
+Maybe shared intention sometimes requires that agents be disposed to plan at least some of their actions from a shared perspective.
+
+Now consider a third (and final) pair of players, Lily and Isabel.
+Lily and Isabel each take the appropriate shared perspective, as if imagining a fictional aggregate agent who had been assigned their two tasks, 
+and they plan accordingly.
+And this is common knowledge.
+What matters is not what they do, however.
+It is that the J which they intend is characterised in terms of planning from a shared perspective.
+What they intend when they intend that they J does not leave it open whether they will adopt a shared perspective in planning their J-ing.
+J is neutral with respect to shared intention but it is not neutral with respect to the requirement that each agent individually adopt a shared perspective when planning how to J.
+***HERE
+
+Imagine two tasks need to be completed; perhaps you need to buy some shoes and get a haircut. 
+Which actions you perform in attempting the first task constrains what actions you can perform in attempting the second task, and conversely.
+For instance, if you shop for the shoes in one store, you won't have enough money left for the best hairdressers.
+There are two ways to approach planning these tasks.
+One approach involves treating the two tasks as separate.
+Perhaps you first think of a plan that is optimal for buying shoes.
+Then, taking this plan as a constraint, you attempt to plan the haircut.
+But as it happens you discover that the options for the haircut are unsatisfactory.
+So you revise the shoe-buying plan and try again to plan the haircut.
+A different approach to planning would be to treat the two tasks as part of a larger, single task.
+Perhaps you identify several possible combinations of shoe-buying and haircut-getting actions.
+You then ask yourself which combination is best overall and then perform the component actions.
+
+Take a situation like the game Ayesha and Benji are playing but now imagine that you are taking over from both players and intend to perform both tasks. 
+One way to plan would involve thinking of two separate tasks.
+First you think about the best way to make the cross hit the red square.
+
+%
+%The difference separating Ayesha and Benji, on the one hand, from Yasmin and Zak on the other 
+%	is unlike the difference separating two strangers who merely happen to be walking side by side from two people intentionally walking together.
+%So if Ayesha and Benji's do not intentionally act together, then nor do Yasmin and Zak.
+
+***Can get a parallel to the special case discussed here by considering walkers whose legs have been tied together in such a way that they are forced to accommodate each other if they are to get around.
+Such people may walk together and meet the Bratman conditions without actually
 

### 2012-10-29 intention motor representation goal ascription with corrado sinigaglia (paper 2)/Motor Representation in Goal Ascription.tex
working on it

@@
 \noindent
-Word count: 7 500
+Word count: *
 \end{abstract}
@@
 
-***HERE
-
-So far we have been arguing that some motor representations are about outcomes by contrasting this claim with the hypothesis that motor all representations concern merely kinematic features of action.
-[*make more vivid!]  This is not quite enough for a convincing argument because some people have claimed that some motor representations in some way concern sensory consequences of actions.
+So far we have been arguing that some motor representations are about outcomes by contrasting this claim with the hypothesis that all motor representations concern merely kinematic features of action.
+This is not quite enough for a convincing argument because some people have argued, convincingly, that some motor representations in some way concern sensory consequences of actions (*refs).
 To illustrate, grasping a small ball between your finger and thumb will have predictable tactile and visual consequences, and it is possible (at least in principle) that some motor representations involved in grasping the ball might somehow concern these sensory consequences.
@@
 Take grasping a ball again: although the outcome---the grasping of the ball---reliably has sensory consequences, these are neither necessary nor sufficient for the outcome to occur.  
-They are not necessary because the grasping could also be performed with a foot or a tool, which could substantially alter the sensory consequences of the action; and because grasping could also in principle occur in complete darkness with numb (or even deafferented) hands.
+They are not necessary because grasping could also be performed with a foot or a tool, which could substantially alter the sensory consequences of the action; and because grasping could also in principle occur in complete darkness with numb hands.%
+\footnote{
+Here we leave open the question of whether there are actions with certain  proprioceptive sensory consequences that are necessary for the action to occur. 
+} 
 The sensory consequences are also not sufficient because in principle the pattern of movements of finger and thumb in relation to the ball could occur without there being any event directed to the grasping of an object---such movements might in principle occur as part of an exercise in moving one's fingers in which the ball is an unwanted distraction rather than a target.
-So our claim that motor representations represent outcomes is conceptually quite distinct from the claim that some motor representations in some way concern sensory consequences of actions: in principle, both claims might be true, or one might be true while the other is false.
+So our claim that motor representations represent outcomes is conceptually quite distinct from the claim that some motor representations might in some way concern sensory consequences of actions: in principle, both claims might be true, or one might be true while the other is false.
 
 Suppose, at least for the sake of argument, that some motor representations do indeed somehow concern sensory consequences of actions. 
-Does we still have grounds for holding that there are motor representations of outcomes?
+Do we still have grounds for holding that there are motor representations of outcomes?
 Yes.
@@
 To return to the example, grasping a ball can be done with a hand, a foot or a tool.
-Using different effectors will alter how the action looks (its visual consequences) but not the goal to which it is directed. 
-So the fact that observing instances of all three types of action can involve a common motor representation shows that not all motor representations can concern sensory consequences and merely kinematic features only.
+Using different effectors will alter how the action looks (its visual consequences), sounds and feels but not the goal to which it is directed. 
+So the fact that observing instances of all three types of action can involve a common motor representation shows that not all motor representations concern either sensory consequences or merely kinematic features only.
+Motor representations of outcomes are also needed.
 
-***The claim that some motor representations represent outcomes raises two further questions.
-First, which outcomes can be represented motorically?
-And, second, what (if anything) ensures that an outcome represented motorically in action observation is an outcome to which the observed action is directed?
-
-Actions are typically directed to many outcomes.
+The claim that some motor representations represent outcomes raises a further question.  Which outcomes can be represented motorically?  
+This question arises because actions are typically directed to many outcomes.
 For instance, a single action might be directed to grasping the scissors, to cutting a newspaper and to removing a particularly useful article to file later.
@@
 
-[*So the plan is to end this section after arguing that mr don't represent directedness, and then have a new section break on 'Capturing without representing'.
-The Q.
-for that section might be: What sort of process links the outcome represented to the observed action and ensures that the outcome represented is one to which the action is directed?]
+\section{How to capture the directedness of a goal-directed action}
+Our overall question is how motor representation could facilitate judgements about the goals of observed actions.
+We have just argued that motor representations which occur in action observation can represent outcomes.
+Now it would be natural to suppose that these outcomes are sometimes outcomes to which the observed actions are directed, and that this is no accident.
+But the argument so far provides no clue as to what might ensure this, as it is silent on processes linking the outcome represented to the observed action.
+What (if anything) ensures that an outcome represented motorically in action observation is an outcome to which the observed action is directed?
 
-\subsection{Directedness}
-Our overall question is how motor representation could enable goal ascription judgements.
-The fact that motor representations can represent outcomes is a step towards answering this question.
-However we are not yet in a position to answer it.
-To see why, consider that someone might (at least in principle) observe an action and represent an outcome without yet relating the outcome to the action.
-Kate is observing Ludwig kick a ball.
-Kate has represented this outcome, the ball's being kicked, but she does not represent the outcome as in any way related to Ludwig's action.
-Rather, she simply finds herself somehow disposed to kick the ball.
-This is not because she intends to pass the ball back to Ludwig or to compete with him or because of any intention concerning Ludwig: it is just because representing the outcome has suggested to her the possibility of performing this action.
-So Kate represents the outcome to which Ludwig's action is directed but she does not identify that outcome as a goal to which his action is directed.
-(She does not relate the outcome to his action at all.)  More would be required for goal ascription.
-To ascribe a goal to an action involves not only representing the action and the outcome which is the goal but also relating the outcome to the action.
+To see the force of this question, note that there is a distinction between merely representing outcomes---which may in fact be outcomes to which certain actions are directed---and representing outcomes as outcomes to which a particular action is directed.
+This point applies to representations generally, including both motor representations and intentions.
+For instance, you might know that a friend intends to break an egg (and so represents an outcome, namely an egg's breaking) and see the friend handling an egg which breaks without yet knowing whether this breaking was appropriately related to that intention.
+To represent an outcome and an action does not necessarily involve representing any relation between them.
+Our first task, then, is to consider whether some motor representations might somehow represent, not merely outcomes, but also the directedness of particular actions to certain outcomes.
 
-It is important in what follows that we sometimes represent intentions without linking them to actions.
-For instance, you might know that a friend intends to break an egg and see the friend handling an egg which breaks without yet knowing whether this breaking was appropriately related to that intention.
-To represent an intention and an action does not necessarily involve representing any relation between the action and the intention.
-
-How might the directedness of actions to outcomes be represented or otherwise captured?
-In philosophy the notion of goal-directedness is standardly explained in terms of intention: for an action to be directed to a outcome is for the action to be appropriately related to an intention whose content specifies that outcome (or, on some views, an intention specifying an appropriately related outcome).
-Given this, it is perhaps natural to suppose that ascribing a particular goal to a given action might involve representing that action, representing an intention whose content specifies the goal, and representing the action as appropriately related to the intention.
-If so, motor representation could have at most a limited role in goal ascription.
+But how might the directedness of an action to an outcome be represented?
+The notion of goal-directedness is standardly explained in terms of intention: for an action to be directed to an outcome is for the action to be appropriately related to an intention whose content specifies that outcome (or, on some views, an intention specifying an appropriately related outcome---see \citealp{Bratman:1984jr}).
+Given this view, no motor representation could represent the directedness of an action to an outcome.
 After all, no motor representation represents an intention.
 
-The problem is not significantly different if we suppose that goal-directedness can be explained in terms of mental representations other than intentions.
-For instance, some have argued that the directedness of an action to an outcome can be explained in terms of motor representation and not only in terms of intention \citep{butterfill:2012_intention}.
-Suppose for the sake of argument that this is correct.
-We still face substantially the same puzzle about how motor representation in goal ascription could capture the directedness of an action to an outcome.
-For motor representations do not represent any mental representations at all.
+There are, of course, other attempts to explicate the notion of goal-directedness.
+For instance, some have argued that the directedness of some actions to particular outcomes can be explained in terms of motor representation and not only in terms of intention \citep{butterfill:2012_intention}.
+This changes nothing, however.
+For motor representations do not represent any representations at all.
 So the directedness of an action to an outcome cannot be represented motorically.
 
-One way to avoid our argument for this conclusion would be to claim that the directedness of an action to a goal can be understood in ways that do not involve intention or indeed any representation.
-For instance, an action's being directed to an outcome may consist in its having the function of bringing about that outcome.
-Here function could be construed teleologically.
+One way to avoid our argument for this conclusion would be to claim that the directedness of an action to a goal can be understood in ways that do not involve intention or any representation at all.
+For instance, consider the idea that an action's being directed to an outcome may consist in its having the function of bringing about that outcome, where function might be construed teleologically.%
+\footnote{
 Teleological accounts of function, and of the application of this notion to understanding goal-directed action, have been extensively developed (*see further Godfrey-Smith 1996; Millikan 1984; Neander 1991; Price 2001; Wright 1976).
-Suppose we accept for the sake of argument  that goal-directedness can be explained in terms of function rather than intention or any other kind of mental representation.
-Substantially the same puzzle remains.
-After all, motor representations no more represent functions than they represent intentions.
+} 
+Even if we accepted this idea, we would still have to conclude that motor representations do not represent goal-directedness.
+After all, motor representations no more represent functions than they represent representations.
 
-At this point it seems to us plausible that motor representations cannot represent the directedness of an action to an outcome.
-If we supposed that all goal ascription involved representing directedness, it would be hard to understand the role of motor representation in enabling goal ascription.
-But it would be a mistake to suppose that directedness must be represented in goal ascription.
-There are ways of capturing directedness which do not involve representing it.
-In the rest of this section we shall argue that in goal ascription the link between actions and outcomes to which they are directed can be captured not only by representing it but alternatively by a planning-like process which involves determining how an outcome is to be achieved and whether given movements are appropriate to achieving that outcome.
+If, as we have suggested, motor representations can represent outcomes but not the directness of actions to outcomes, what (if anything) might ensure that, when someone observes a particular action, outcomes she represents motorically are outcomes to which that action is directed?
+We must face this question because if it turned out that it was only ever a coincidence that outcomes represented motorically in action observation are outcomes to which the observed action is directed, then it would be hard to understand the role of motor representation in facilitating judgments.
+In the rest of this section we shall argue that the directedness of an action to an outcome could be captured motorically by means of a planning-like process which involves determining how an outcome is to be achieved and whether given movements are appropriate to achieving that outcome.
 
-To see how the directedness of an action to an outcome can be captured without being represented, let us first step back to consider cases of goal ascription which involve representing intentions only (so motor representation is not yet in the picture).
-Observing Ayesha gathering some papers during a long meeting that shows no sign of ending, you conjecture that she is intending to leave discretely.
-You use this conjecture about Ayesha's intention to determine one or more likely sequences of behaviours, almost as if you were planning from her point of view how to act on that intention.
+To see how this might work, let us step back, momentarily, from motor phenomena to introduce the idea with an illustration.
+Observing Ayesha gathering some papers during a long meeting that shows no sign of ending, you conjecture that a goal of these and her next actions will be to leave discretely.
+You use this conjecture about the outcome to which Ayesha's actions are directed to determine one or more likely sequences of behaviours, almost as if you were planning from her point of view how to achieve that outcome.
 This leaves you with expectations concerning how she will act.
 You might be waiting for Ayesha, who has no other friends present, to give you a nod before quietly pushing away from the table.
-If what you observe matches these expectations, or if there are only deviations that can be explained in ways consistent with your conjecture about Ayesha's intention, then you may come to regard the conjecture as confirmed.
-But if there is a mismatch---if, for example, Ayesha signals that she wants attention and looks around to see whether people are looking at her---then you might reject the conjecture about her intention to leave discretely.
-As this illustrates, goal ascription often (perhaps not always) involves determining how an action will or would unfold, and this process of determination may ensure that there is a reliable connection between the action's being directed to a certain goal and your ascribing that goal to the action.
-So the directedness of an action to a goal can be captured by a planning-like process in addition to being represented.
+If what you observe matches these expectations, or if there are only deviations that can be explained in ways consistent with your conjecture about the goal of Ayesha's actions, then you may come to regard the conjecture as confirmed.
+But if there is a mismatch---if, for example, Ayesha signals that she wants attention and looks around to check people are looking at her---then you might reject the conjecture about the goal of her actions being to leave discretely.
+As this illustrates, determining how an action will or would unfold and then monitoring how it does unfold can ensure that there is a reliable connection between the action's being directed to a certain outcome and your representing that outcome.
+So the directedness of an action to a goal can be captured by a planning-like process together with appropriate monitoring.
 
-You might suggest that where an observer represents another agent's intention, she would only ever determine how the agent will act if she also supposes the actions will be directed to fulfilling this intention.
-This may be right.
-The illustration is only supposed to show that goal-directedness could in principle be captured without being represented.
+In outline the link between an action and an outcome to which it is directed can be maintained like this.
+The representation of an outcome leads to a planning-like process which generates predictions about how the action will unfold, and this representation is weakened to the extent that the predictions are not met.
+This is how planning processes can be coopted to ensure that an outcome represented is likely to be an outcome to which the observed action is directed.
 
-Motor representations and processes may likewise capture the directedness of an action to an outcome.
+Could motor processes capture the directedness of an action to an outcome in this way?
 In action observation it is not just that there may be motor representations of outcomes.
-There is evidence that a motor representation of an outcome sometimes causes a determination of which movements are likely to be performed to achieve that outcome  \citep{kilner:2004_motor, urgesi:2010_simulating}.
+There is also evidence that a motor representation of an outcome sometimes causes a determination of which movements are likely to be performed to achieve that outcome  \citep{kilner:2004_motor, urgesi:2010_simulating}.
 The processes involved in determining how observed actions are likely to unfold given their outcomes are closely related, or identical, to processes involved in performing actions.
 This is suggested both by the involvement of regions of the brain associated with action (*refs), and by patterns of facilitation and interference which occur when simultaneously observing and performing actions \citep{brass:2000_compatibility, kilner:2003_interference, craighero:2002_hand}.
-It is almost as if the observer were planning from the agent's point of view how to act.
-These planning-like processes sometimes leave observers with expectations about how an action will unfold.
-For instance, if you were to observe just the early phases of a grasping movement your eyes may jump to its likely target, ignoring nearby objects \citep{ambrosini:2011_grasping}.
+It is almost as if the observer were planning the agent's actions.
+
+Planning-like processes in action observation have also been demonstrated by measuring observers’ predictive gaze.
+If you were to observe just the early phases of a grasping movement your eyes may jump to its likely target, ignoring nearby objects \citep{ambrosini:2011_grasping}.
 These proactive eye movements resemble those you would typically make if you were acting yourself \citep{Flanagan:2003lm}: in reaching for something, it is natural to look at the target rather than your hand.
-Importantly, the occurrence of such proactive eye movements in action observation depends both on representing the outcome of an action (even temporary interference in your ability to represent the outcome will interfere with the eye movements,  \citealp{Costantini:2012fk}) and also on planning-like processes (having to perform actions incongruent with those you are observing also interferes with proactive eye movements, \citealp{Costantini:2012uq}).
-Finally, going slightly beyond what there is currently direct evidence for, it is also plausible that whether the expectations about how the action will unfold are met or not will influence whether the motor representation of that outcome persists.
-This ensures that, where someone is observing an action, there is a reliable connection between the observer representing an outcome and the action being directed to that outcome.
-There is a sense, then, in which the directedness of an action to a goal can be captured by motor representations and processes: As might also happen in the case of intention, the directedness of the action to a goal is captured by planning-like processes.
+Importantly, the occurrence of such proactive eye movements in action observation depends both on representing the outcome of an action (even temporary interference in the observer's ability to represent the outcome will interfere with the eye movements, \citealp{Costantini:2012fk}) and also on planning-like processes (requiring the observer to perform actions incongruent with those she is observing also interferes with proactive eye movements, \citealp{Costantini:2012uq}).
+This, then, is further evidence that observers do something almost like planning the actions they observe.
 
-Let us stipulate that to capture the directedness of an action to an outcome is to (a) represent the action; and to (b) determine how the action is likely to unfold given that it would normally realise that outcome; where (c) this determination leaves the observer with expectations about how the action will unfold; and (d) whether the representation of the outcome persists depends on the expectations being met.
-We have been arguing that there is evidence that motor representations and processes can capture the directedness of actions.
+We can now see how motor processes might ensure that an outcome represented is likely to be an outcome to which the observed action is directed.
+As we have seen, there is evidence that observers represent outcomes motorically and that planning-like processes generate expectations about how the actions will unfold given these outcomes.
+To take a tiny step further, we conjecture that, in action observation, motor representations of outcomes are weakened to the extent that the expectations they generate are unmet \citep[compare][]{Fogassi:2005nf}.
+This is how the directedness of action to an outcome can be captured motorically even in the absence of any representation of intention or function.
 
-It is perhaps worth pausing here to stress a parallel between performing and observing actions.
-Start with performance.
-In virtue of what is an action related to the outcome or outcomes to which it is directed?
-Suppose that in performing an action there is a motor representation of an outcome.
-Now an action is not directed to an outcome just in virtue of there being a representation of that outcome, of course.
-Rather the outcome has to play some guiding role in determining how the action will unfold.
-So what makes it the case that the action is directed to the outcome is not just the content of the representation but also its role in determining how the action is to unfold.
-Now switch to action observation.
-How might an observer capture the directedness of an action to an outcome?
-As in performing an action, there is sometimes a motor representation of an outcome and a process of working out how the action will unfold, as we have seen.
-But unlike performing action, in observation the process of working out how the action will unfold does not normally directly cause of the action (of course).
-So if expectations about how the action will unfold are not met, it is not normally possible to alter the action; instead it may be necessary to link the action to an outcome different from the one whose representation generated these unmet expectations.
-This is how the representations and processes which ground the directedness of actions to outcomes in performing actions may also be involved in capturing the directedness in observing actions.
-%
-\begin{quote}
-[***\textit{A previous, rough attempt at the above paragraph}:  \\
-grounding /capturing: In performing an action there is a representation of an outcome.
-Now an action is not directed to an outcome just in virtue of there being a representation of that outcome, of course.
-Rather the outcome has to play some guiding role in working out how the action will unfold.
-So what determines that the action is directed to the outcome is both the representation and the process of working out how the action is to unfold.
-Now the case of action observation is similar.
-There is a representation of an outcome and a process of working out how the action will unfold.
-A difference is that the process of working out how the action will unfold is not (of course) normally a cause of the action.
-So the representations and processes which ground the directedness of actions to outcomes in performing actions: these representations and processes also capture the directedness in observing actions.]
-\end{quote}
+Let us sum up so far.
+Our overall question is how motor representations could ever facilitate judgements about the goals of actions.
+Our focus until now has been relations between the contents of judgements and the contents of motor representations.
+In the previous section we argued that motor representations do not all represent merely kinematic features of action; rather, some motor representations represent outcomes to which actions are directed.
+This led us to ask (in the present section), how the directedness of an action to an outcome might be captured motorically.
+We argued that motor representations probably do not represent the directedness of an action to an outcome, but explained how directedness can be captured by planning-like motor processes which occur in action observation.
+Potentially, then, there is a respect in which the content of a motor representation may resemble the content of judgement about the goal to which an action is directed: there may be an outcome which both represent, and both may  capture the directedness of an action to that outcome.
+We therefore conjecture that where motor representation facilitates a judgement about the goal to which an action is directed,
+there is a motor representation of this outcome, or of a related outcome, and a planning-like motor process which captures the directedness of the action to the outcome.
+Motor representations can facilitate judgements about the goals of actions because the goals can be represented motorically and because the directedness of actions to those goals can be captured motorically.
 
-In this section we have considered a first obstacle to understanding how judgements about the goals of actions sometimes depend on motor representations.
-The obstacle was that motor representations might appear not to concern the goals of actions at all, making their influence on judgements about goals difficult to understand.
-In removing this obstacle we made two points.
-The first was that motor representations do not all represent merely kinematic features of action; rather, some motor representations represent outcomes to which actions are directed.
-The second point was that although motor representations probably do not represent the directedness of an action to an outcome, reflection on the role of motor representations and processes in action observation shows that this directedness can be captured motorically.
-Potentially, then, there is a respect in which the content of a motor representation may resemble the content of judgement about the goal to which an action is directed: they may represent a single outcome and capture the directedness of an action to that outcome.
-We therefore propose that where a judgement about the goal to which an action is directed depends on motor representations, there is a motor representation of an outcome which captures the directedness of the action to the goal.%
+This conjecture---that motor representation not only facilitates judgements but also does so partly by virtue of the contents of motor representations matching those of judgements---is surprising to the extent that it is natural to suppose that motor representation is isolated from judgement.
+While the evidence all but forces us to accept the conjecture, this conjecture, far from answering our overall question, only serves to make more mysterious the facilitatory effect of motor representations on judgements about the goals of actions.
+How could motor representations have content-respecting influences on judgements?%
 \footnote{
-***Do we need to allow that the outcomes represented motorically and in judgement may not be identical?
+It may be tempting to make a radical move by supposing that there are direct inferential relations between motor representation and thought.
+However there are good arguments for denying that motor representations are inferentially integrated with the thoughts involved in explicit judgements about the goals of observed actions \citep{butterfill:2012_intention}.
 }
 
-\section{Action and Experience}
+\section{Experience of Action}
 \label{sec:processes}
-Our overall question is how judgements about the goals of actions could ever depend on motor representations.
-The previous section concerned the possibility of a match between the contents of motor representations and judgements.
 Now to say that two varieties of representations---motor representations and judgements---sometimes have nonaccidentally matching contents is not yet to explain how that match might come about.
@@
 MR<----> Action Judgment (action concept) <----> B \& D Judgment (b\&d concept)
-
-\bibliography{$HOME/endnote/phd_biblio}
-
-\end{document}

### 2012-11-01 intention motor representation goal ascription with corrado sinigaglia (paper 2)/Motor Representation in Goal Ascription.tex
working on it

@@
 \noindent
-Word count: *
+Word count: ***
 \end{abstract}
@@
 
-Motor representation sometimes occurs in action observation and facilitates judgements about the goals to which actions are directed (see section \vref{sec:evidence} and \citealp{rizzolatti_functional_2010} for a review).
+Motor processes sometimes occur in action observation and facilitate judgements about the goals to which actions are directed (see section \vref{sec:evidence} and \citealp{rizzolatti_functional_2010} for a review).
 Suppose you observe someone's hand moving towards a pair of scissors. 
 You may judge that she will shortly perform an action directed to cutting something up. 
-Making this judgement may depend on representing the other's actions in motor terms \citep{boria:2009_intention, ortigue:2010_understanding}.  
-This is not to say, of course, that you could not make a judgement of this type without representing the action in motor terms, only that making this particular judgement may depend on motor representation.
+Making this sort of judgement is sometimes facilitated by the occurrence of motor processes \citep{boria:2009_intention, ortigue:2010_understanding}.  
+This is not to say that it would be altogether impossible to make a judgement of this type without the involvement of motor processes. 
+What the evidence shows is rather that judgements about the goals of actions are sometimes faster or more accurate (or both) when motor processes occur in action observation than when they do not.
 How could this work? 
-How could some judgements about the goals of actions ever depend on motor representations?
+How could motor processes sometimes facilitate judgements about the goals to which actions are directed?
 
@@
 This is plausibly a consequence of increased motor expertise rather than any form of perceptual learning because the subjects were blindfolded during training (and the point-light stimuli on which their judgements were based are, of course, entirely visual).
-A related way to measure the effects of motor representation on judgements is to temporarily lesion part of the motor cortex using repetitive transcranial magnetic stimulation  \citealp{urgesi:2007_representation, moro:2008_neural}.
+A related way to measure the effects of motor representation on judgements is to temporarily lesion part of the motor cortex using repetitive transcranial magnetic stimulation  \citep{urgesi:2007_representation, moro:2008_neural}.
 \citet{urgesi:2007_representation} measured how long subjects took to make judgements about the type of action they were observing.
@@
 Our focus until now has been relations between the contents of judgements and the contents of motor representations.
-In the previous section we argued that motor representations do not all represent merely kinematic features of action; rather, some motor representations represent outcomes to which actions are directed.
-This led us to ask (in the present section), how the directedness of an action to an outcome might be captured motorically.
+In the previous section we argued that the contents of some motor representations cannot be fully characterised in terms of merely kinematic features and sensory consequences of action only; rather, some motor representations represent outcomes to which actions are directed.
+This led us to ask, in the present section, what if anything ensures that outcomes represented motorically in action observation are outcomes to which the observed action is directed.
 We argued that motor representations probably do not represent the directedness of an action to an outcome, but explained how directedness can be captured by planning-like motor processes which occur in action observation.
 Potentially, then, there is a respect in which the content of a motor representation may resemble the content of judgement about the goal to which an action is directed: there may be an outcome which both represent, and both may  capture the directedness of an action to that outcome.
-We therefore conjecture that where motor representation facilitates a judgement about the goal to which an action is directed,
-there is a motor representation of this outcome, or of a related outcome, and a planning-like motor process which captures the directedness of the action to the outcome.
+We therefore conjecture that where motor representation facilitates a judgement about the goal to which an action is directed, there is a motor representation of this outcome, or of a related outcome, and a planning-like process which captures the directedness of the action to the outcome.
 Motor representations can facilitate judgements about the goals of actions because the goals can be represented motorically and because the directedness of actions to those goals can be captured motorically.
 
-This conjecture---that motor representation not only facilitates judgements but also does so partly by virtue of the contents of motor representations matching those of judgements---is surprising to the extent that it is natural to suppose that motor representation is isolated from judgement.
-While the evidence all but forces us to accept the conjecture, this conjecture, far from answering our overall question, only serves to make more mysterious the facilitatory effect of motor representations on judgements about the goals of actions.
-How could motor representations have content-respecting influences on judgements?%
+This conjecture, however, far from fully answering our overall question, throws up a further challenge.
+For the conjecture, that motor representation facilitates judgements partly by virtue of the contents of motor representations matching those of judgements, implies that there are content-respecting influences of motor representation on judgement.
+The existence of such content-respecting influences is surprising given that motor representations are to a significant degree isolated from judgements.
+True, the evidence all but forces us to accept that such influences occur.
+But that leaves us with the question of how such influences are possible.
+How could motor representations sometimes have content-respecting influences on judgements?
+What might connect some privileged motor representations to judgements about goals in action observation?%
 \footnote{
-It may be tempting to make a radical move by supposing that there are direct inferential relations between motor representation and thought.
-However there are good arguments for denying that motor representations are inferentially integrated with the thoughts involved in explicit judgements about the goals of observed actions \citep{butterfill:2012_intention}.
-}
+At this point it may be tempting to make a radical move by supposing that there are direct inferential relations between motor representation and thought.
+But the evidence we have reviewed does not support this radical move, and there are compelling arguments for denying that motor representations are inferentially integrated with the thoughts involved in explicit judgements about the goals of observed actions (see e.g.\ \citealp{butterfill:2012_intention}).
+ Note also that someone who made this radical move would have to explain why motor representation and judgement appear to be inferentially isolated in many other cases.}
+
+DUSSELDORF ARGUMENT (rough, from Friday skype):
+(1) Motor imagery involves experience of action as such (e.g. of grasping as such) and not just of sensory consequences of actions
+(one relevant consideration is that you know not only which action you are imagining but when you have succeeded in imagining performing it);
+(corollary: there can be experiences of action without experiences of acting)
+(2) There are common elements between what you  experience when you imagine acting and what sometimes you experience when you observe an action; in particular, the experience of action is common to both
+Therefore
+(3) Sometimes observing action involves experiencing action as such
 
@@
 \label{sec:processes}
-Now to say that two varieties of representations---motor representations and judgements---sometimes have nonaccidentally matching contents is not yet to explain how that match might come about.
-Which processes could be responsible for the match?
-When judgements about the goals of observed actions depend on the observers' motor representations, which processes connect motor representations to judgements?
+Our aim in this section is to explain how motor representations could sometimes have content-respecting influences on judgements about the goals of observed actions.
+Broadly, our proposal will be that the influence of motor representations on judgement is like that of visual representations.
+Some some visual representations have content-respecting influences on judgement because they give rise to visual experiences of the objects and features represented.
+Similarly, some motor representations have content-respecting influences on judgement because they give rise to motor experiences of actions.
+Or so we shall argue.
 
-In broad terms, our will be proposal is that the dependence of observational judgements on motor representations goes via experience.
+TBE: content-respecting influences of (MR of G$_1$) onto (Jud about G$_2$) where G$_1$ matches G$_2$
+
+Explanation-1: MR of G captures directedness of action a -> experiences of action a as directed to G -> judgement that action is directed to goal G
+	Key component: Dusseldorf argument, imagination
+
+Explanation-2: MR of G captures directedness of action a -> experiences revelatory of action but experiences as of sensory consequences of actions -> association between experienced sensory consequences and action directed to goal G (how?
+for example, I know the sound of a newspaper being opened) -> judgement that action a is directed to goal G
+
+E.g. clapping:
+	representation of sound
+	representation of action
+	association: characteristic sounds<>action
+	Thanks to the association, experience of the sounds can result in a representation of the action
+
+Broadly, our will be proposal is that the influence of motor representation on observational judgements goes via experience.
 It is because motor representations can affect our experiences when we observe actions that those motor representations can facilitate judgements about these actions.
@@
 MR<----> Action Judgment (action concept) <----> B \& D Judgment (b\&d concept)
+
+\bibliography{$HOME/endnote/phd_biblio}
+
+\end{document}

### 2012-11-01 counterexample to bratman on shared intention/a counterexample to bratman on shared intention.tex
working on it

@@
 %
-Care is needed in specifying the contents of the intentions concerning our J-ing in these clauses. 
-As mentioned above, an appeal to shared intention is supposed to  characterise systematically a difference between two agents who intentionally J together (e.g.\ walk  together) and two agents who also  J together but do not do so intentionally (e.g.\ they happen to be walking side by side).
-So we must avoid tacitly appealing to this distinction by  restricting possible values of J to those which involve our intentionally doing something together. 
+How could we show that our meeting these conditions is not in fact sufficient for us to have a shared intention? 
+We seek a case where the conditions are met but we lack a shared intention.
+But, in advance of deciding whether Bratman's account is right, how could we determine whether we have a shared intention?
+As already mentioned, an appeal to shared intention is supposed to  characterise systematically a difference between two agents who intentionally J together (e.g.\ walk  together) and two agents who also  J together but do not do so intentionally (e.g.\ they happen to be walking side by side).
+Let's stipulate that two cases \emph{differ with respect to shared agency} when they differ in the way exemplified by the contrast between our intentionally walking together and our merely walking in parallel.
+Suppose, then, that we construct three cases, A, B and C.
+The three cases are as similar as possible, except that the above conditions are met in cases B and C but not in case A.
+Now if Bratman's conditions are sufficient for shared intention, A should differ from B with respect to shared agency, and B should not differ from C with respect to shared agency.
+But suppose that in fact A does not differ from B with respect to shared agency, but B does so differ from C.%
+\footnote{
+Strictly speaking only A and B are needed for the counterexample, of course.
+But including C makes it clearer that A and B do not differ with respect to shared agency.
+}
+Then we can conclude that Bratman's conditions are not sufficient for shared intention.
+This is how our counterexample will work.
+
+Care is needed in specifying the contents of the intentions concerning our J-ing in the above clauses, (1)--(3). 
+As we have stressed, 
+an appeal to shared intention is supposed to characterise systematically differences with respect to shared agency---to characterise, that is, a difference between two agents who intentionally J together and two agents who also  J together but do not do so intentionally.
+We must therefore avoid tacitly appealing to this distinction by  restricting possible values of J to those which involve our intentionally doing something together. 
 Rather the above  conditions, (1)--(3), must be sufficient for shared intention even for some values of J which are `neutral with respect to shared intentionality'.%
@@
 \citet{Bratman:1999fr}  defends this claim at length. 
-Note also that this claim must be true if Bratman's account of shared intention is to provide an informative way of distinguishing between the contrast cases mentioned at the start.
+Note also that this claim must be true if Bratman's account of shared intention is to provide an informative and systematic distinction between the contrast cases mentioned at the start.
 }
@@
 The \emph{strong claim} is that for any J, these conditions are sufficient for you and I to intend that we J.
-One of Bratman's aims is to show that it is possible to give an account of shared intention using concepts that `are available within the theory of individual planning agency' \citep[p.\ 163]{Bratman:2009lv}.
+One of Bratman's aims is to show that an account of shared intention can be \emph{conceptually conservative}:
+he aims, that is, to show that it is possible to give an account of shared intention using concepts that `are available within the theory of individual planning agency' \citep[p.\ 163]{Bratman:2009lv}.
 Achieving this aim would require the strong claim, and it is to the strong claim that our counterexample is directed.
@@
 \section{Unshared Intentions}
-Before we can introduce the counterexample we need to introduce what we shall call `unshared intentions'.
+Before we can introduce the counterexample we need to define what we shall call `unshared intentions'.
 We have an \emph{unshared intention} that we <J$_1$, J$_2$> where J$_1$$\neq$J$_2$ just if:
@@
 \begin{quote}
-our individual subplans concerning our <J$_1$, J$_2$>-ing \emph{mesh} just in case there is some way we could <J$_1$, J$_2$> that would not violate either of our subplans but would, rather, involve the successful execution of those subplans.
+our individual subplans concerning our <J$_1$, J$_2$>-ing \emph{mesh} just in case there is some way I could J$_1$ and you could J$_2$ that would not violate either of our subplans but would, rather, involve the successful execution of those subplans.
 \end{quote}

### 2012-11-02 intention motor representation goal ascription with corrado sinigaglia (paper 2)/Motor Representation in Goal Ascription.tex
various

@@
 \title{}
+
+% --- paste from google docs after here ---
 
@@
 
+\section{***Notes for Action Experience section}
+
 DUSSELDORF ARGUMENT (rough, from Friday skype):
@@
 
-\section{Experience of Action}
-\label{sec:processes}
-Our aim in this section is to explain how motor representations could sometimes have content-respecting influences on judgements about the goals of observed actions.
-Broadly, our proposal will be that the influence of motor representations on judgement is like that of visual representations.
-Some some visual representations have content-respecting influences on judgement because they give rise to visual experiences of the objects and features represented.
-Similarly, some motor representations have content-respecting influences on judgement because they give rise to motor experiences of actions.
-Or so we shall argue.
+------- there is evidence that motor representation might have content-respecting influences on judgment
+
+MR ----?----- JUD
+  experience: how to argue 
+….............how is possibile? experience revelatory of action
+------------------- parallel between colour and action experience
 
@@
 
+\section{Experience of Action}
+\label{sec:processes}
+Our aim in this section is to explain how motor representations could sometimes have content-respecting influences on judgements about the goals of observed actions.
 Broadly, our will be proposal is that the influence of motor representation on observational judgements goes via experience.
-It is because motor representations can affect our experiences when we observe actions that those motor representations can facilitate judgements about these actions.
-However this is consistent with two quite different hypotheses concerning the experiences associated with observing action.
+It is because motor representations can affect our experiences when we observe actions that those motor representations can facilitate judgements about the actions.
+However this is consistent with two quite different hypotheses concerning the experiences associated with motor representation in action observation.
 
-One hypothesis is that in observing action we experience not only movements and sounds but also goal-directed actions as such: it is this experience which enables judgements of action, and motor impairments affect judgements of action by preventing one from experiencing actions as such.
-On the other hypothesis, in observing actions presented as point-light stimuli or acoustically we experience movements and sounds only.
-Motor representations somehow make these particular experiences of movements and sounds possible, and these experiences in turn ground judgements about actions.
-On both hypotheses motor representations enable experiences which are revelatory of action in the sense that they can provide reasons for correct judgements about which action someone is performing.
-The difference concerns what these revelatory experiences are experiences of: whether some are experiences of goal-directed actions as such (the first hypothesis) or whether all are experiences of movements and sounds only (the second hypothesis).
+One hypothesis, call it the Indirect Hypothesis, is based on evidence that, in observing action, motor representation can influence visual and acoustic experience.
+(Perhaps this holds for other modalities too).
+On this hypothesis, these influences enable an observer to identify goals to which the observed action is directed because the observer is additionally able to recognise actions from their associated sensory effects.
+To illustrate, suppose that an observer can associate the characteristic sound of a newspaper being snapped open with the action of snapping open a newspaper.
+Suppose also that, when observing an action, motor representations can influence whether, when or how she experiences this characteristic sound.
+Then motor processes can indirectly influence the speed and accuracy of the observer's judgements about which actions are directed to this goal, the snapping open of a newspaper, by influencing whether, when or how she experiences the characteristic sound of this action.%
+\footnote{
+The Indirect Hypothesis is inspired by and consistent with views defended in \citet{Csibra:2007fy} and *Knoblich.
+However, to our knowledge these authors have not considered exactly our question about content-respecting influences of motor representation on judgement, and we are not suggesting that the would endorse the Indirect Hypothesis in its entirety.
+}
 
-In what follows we first consider evidence that the second hypothesis is plausible and then, in a further section, we consider arguments for the first hypothesis.
-Although we shall conclude that the first hypothesis is correct---there are experience of action as such---it is helpful to approach this by way of evidence relevant to the second hypothesis.
+The other hypothesis is in some ways more radical.
+This hypothesis, call it the Direct Hypothesis, is that in observing action we experience not only movements and sounds but also goal-directed actions as such.
+Such experiences---call the motor experiences---stand to motor representations roughly as visual experiences stand to visual representations.
+The motor experience of an action as directed to a particular outcome is made possible by a motor representation of that outcome.
+So motor representations can have a content-respecting influence on judgements because motor representations enable motor experiences on which judgements are based.%
+\footnote{
+In formulating the Direct Hypothesis we draw on ideas from \citet{rizzolatti_mirrors_2008} and \citet{Jeannerod:1999kj}.
+}
 
-\subsection{Motor representation can enable visual and acoustic experiences}
-For convenience, in what follows we shall use the term `perceptual experiences' to mean experiences of the visual, acoustic, tactile and other things commonly associated with perception.
-As a terminological stipulation (for which there are reasons, but they come later), no experiences of action as such are perceptual experiences.
-Later, when we have considered arguments that some experiences are experiences of actions as such, we shall contrast `perceptual experience' with `motor experience'.
+Let us say that an experience, motor or not, is revelatory of action just if it  can provide the subject of experience with reasons for correct judgements about which action someone is performing.
+On both hypotheses motor representations enable experiences which are revelatory of action.
+And both hypotheses use this broad idea to explain how motor representations could have content-respecting influences on judgements.
+The primary difference between the hypotheses concerns what these revelatory experiences are experiences of: whether all such are experiences of movements, sounds and the like only (the Indirect Hypothesis), or whether some are experiences of goal-directed actions as such (the Direct Hypothesis).
 
-The possibility that some motor representations enable visual experiences is suggested by research showing that motor plans, such as might be involved in producing a gesture with the hand, can influence visual processes \citep{bortoletto:2011_action}.
+In the next section we expand on and consider evidence for the first hypothesis before turning to the second hypothesis in the following section.
+
+\section{The Indirect Hypothesis}
+The Indirect Hypothesis is depends on the claim that motor representations can influence sensory experiences.
+This claim is supported by a variety of evidence.
+\citet{bortoletto:2011_action} show that motor plans, such as might be involved in producing a gesture with the hand, can influence visual processes \citep{bortoletto:2011_action}.
 The influence of motor representations on visual processes varies depending on both the gesture to be performed and the visual stimulus presented.

### 2012-11-02 counterexample to bratman on shared intention/a counterexample to bratman on shared intention.tex
working

@@
 How could we show that our meeting these conditions is not in fact sufficient for us to have a shared intention? 
-We seek a case where the conditions are met but we lack a shared intention.
-But, in advance of deciding whether Bratman's account is right, how could we determine whether we have a shared intention?
+We seek a case where the conditions are met although we lack a shared intention.
+But how could we determine whether we have a shared intention in advance of deciding whether Bratman's account is right?
 As already mentioned, an appeal to shared intention is supposed to  characterise systematically a difference between two agents who intentionally J together (e.g.\ walk  together) and two agents who also  J together but do not do so intentionally (e.g.\ they happen to be walking side by side).
@@
 Now if Bratman's conditions are sufficient for shared intention, A should differ from B with respect to shared agency, and B should not differ from C with respect to shared agency.
-But suppose that in fact A does not differ from B with respect to shared agency, but B does so differ from C.%
+But suppose that in fact A does not differ from B with respect to shared agency, whereas B does so differ from C.%
 \footnote{
@@
 
-We shall argue that Yasmin and Zak do not have a shared intention that they J$_1$ despite meeting meet Bratman's conditions (1)--(3).
-This is probably not yet obvious, however.
-Let us first attempt to provide some intuitive support for this claim.
-Return to Ayesha and Benji for a moment.
+We claim that Yasmin and Zak have a shared intention that they J$_1$
+only if 
+Ayesha and Benji have a shared intention. 
+This claim follows from the similarities of the two cases.
+The only difference is that Yasmin and Zak happen to be assigned the same task, whereas Ayesha and Beatrice are not.
+But neither Yasmin nor Zak considers the fact that they have the same task to be important.
+
+Let us pause to emphasise some ways in which the case of Yasmin and Zak is relevantly similar to that of  Ayesha and Benji. 
 Ayesha sees Benji's actions as constraints to work around and opportunities to exploit.
@@
 Each allows himself or herself to be exploited by the other because being exploited enables exploiting.
-The sort of shared agency that an account of shared intention is supposed to capture must involve more than this sort of reciprocal exploitation, where each agent merely exploits the other's intentions.
 But Yasmin and Zak are engaged in just the same sort of reciprocal exploitation as Ayesha and Benji.
 The only difference is that, quite by chance, Yasmin's task and Zak's task are the same.
-So, intuitively, Yasmin and Zak, like Ayesha and Benji, belong in the same category as two people who happen to be walking side by side and not in the category exemplified by two people intentionally walking together.
+So if we focus on how  
+		the case of Yasmin and Zak
+	differs from
+		 that of Ayesha and Benji,
+we can see that these differences do not plausibly amount to a difference with respect to shared agency.
 
+At this point it may be tempting to assume that Ayesha and Benji do not have a shared intention. 
+We might assume this on the grounds that
+ the sort of shared agency that an account of shared intention is supposed to capture must involve more than this sort of reciprocal exploitation, where each agent sees the other's actions only as constraints or opportunities.
+And given that Ayesha and Benji do not have a shared intention,
+	we could  infer that neither do Yasmin and Zak.
+This would give us the required counterexample.
+For, as already noted, Yasmin and Zak do meet the conditions Bratman claims are sufficient for shared intention.
+
+We must be careful here, however.
+The best reply to our counterexample in defence of Bratman's account probably involves claiming that Ayesha and Benji have a shared intention rather than denying that their case differs from that of Yasmin and Zak in ways relevant for shared agency.
+So we must ask, 
+Could Ayesha and Benji's unshared intention that they <J$_1$, J$_2$> be a  shared intention?
+This question might initially appear frivolous but it is not straightforward to answer.
+Since Bratman provides only sufficient conditions for shared intention, his account doesn't tell us that the unshared intention is not a shared intention.
+And while we might be tempted to assume that for two agents to intentionally J together they must at least each have intentions concerning J, 
+this assumption might reasonably be rejected by those who, like Bratman, reject the Simple View according to which when an individual intentionally J-s she has an intention concerning her J-ing \citep{Bratman:1984jr}.
+
+***Simple point: the fact that we have the same task should play a role in our planning.
+It is something that we should embrace, and perhaps even want.
+(Be careful: two people might intentionally act together although each would prefer to act alone, and would seize the opportunity to do so.)  Cordula's idea (modified) was that we aim for shared success (it's not a question of whether I care about \\emph{your} success, but I must care about \emph{our} success).
+
+\section{***OLD}
 Assuming for a moment that this is right,

### 2012-11-09 counterexample to bratman on shared intention/a counterexample to bratman on shared intention.tex
before removing old stuff

--- 2012-10-28 counterexample to bratman on shared intention/a counterexample to bratman on shared intention.tex
@@
 And \citet{Searle:1990em} contrasts park visitors who  simultaneously run to a central shelter in performing a dance with park visitors who likewise run to the central shelter but only because of an impending storm.
+%And Bratman (\citeyear{Bratman:1992mi}, p.\ 333; \citeyear{Bratman:1993je}, p.\ 103--4) contrasts two people who cooperatively plan to go to New York  together with two people who each have an intention that they go to New York together by way of coercing the other into travelling with her.
 These and other contrast cases invite the question, 
-How does our intentionally acting together differ from our merely acting in parallel? 
+How do cases involving shared agency differ from cases involving parallel agency only?
+
 The first contrast case, Gilbert's, shows that the difference can’t be just  a matter of coordination because people who are merely happen to be walking side by side each other also need to coordinate their actions in order to avoid colliding.
-Note also that in both cases each individual's walking is intentional, so our intentionally walking together cannot be a matter only of our each intentionally walking.
+Note also that in both cases each individual's walking is intentional, so our intentionally walking together cannot be  only a matter of our each intentionally walking.
 The second contrast case, Searle's, shows that the difference can’t just be that the resulting actions have a common effect because merely parallel actions can have common effects too.%
@@
 %
-Care is needed in specifying the contents of the intentions concerning our J-ing in these clauses. 
-As mentioned above, an appeal to shared intention is supposed to  characterise systematically a difference between two agents who intentionally J together (e.g.\ walk  together) and two agents who also  J together but do not do so intentionally (e.g.\ they happen to be walking side by side).
-So we must avoid tacitly appealing to this distinction by  restricting possible values of J to those which involve our intentionally doing something together. 
+In this paper we give a counterexample to Bratman's  claim that the above conditions, (1)--(3), are collectively sufficient conditions for shared intention. 
+We shall also suggest a revision to avoid the counterexample.
+%Apart from improving our understanding of what shared intention is, this will also bear on the kinds of planning that are involved in acting together.
+
+Before going further we must distinguish two versions of the claim that (1)--(3) are collectively sufficient for shared intention.
+The \emph{weak claim} is that there is some J such that these conditions are sufficient for you and I to intend that we J.
+The \emph{strong claim} is that for any J, these conditions are sufficient for you and I to intend that we J.
+One of Bratman's aims is to show that an account of shared intention can be \emph{conceptually conservative}:
+he aims, that is, to show that it is possible to give an account of shared intention using concepts that `are available within the theory of individual planning agency' \citep[p.\ 163]{Bratman:2009lv}.
+Achieving this aim would require the strong claim, and it is to the strong claim that our counterexample is directed.
+
+How could we show that our meeting Bratman's conditions, (1)--(3), is not in fact sufficient for us to have a shared intention? 
+We seek a case where the conditions are met although we lack a shared intention.
+But how could we determine that we lack a shared intention?
+As already mentioned, the notion of shared intention is supposed to 
+	make it possible to 
+	characterise systematically a difference between 
+		cases involving shared agency (such as our walking together)
+		and
+		cases involving parallel agency only (such as two strangers who happen to be walking the same route side-by-side). 
+Suppose, then, that we had a trio of cases, A, B and C, each involving two agents.
+Suppose, further, that A and B involved parallel agency only, whereas C involved shared agency.
+Then we could be sure that A and B do not involve shared intention.
+Now suppose 
+	that, for some relevant J, Bratman's conditions, (1)--(3) above, were met in cases B and C alike, 
+	that in each case the structure of intention and knowledge played an appropriate role in guiding the agents' actions,
+	and that in each case the agents thereby successfully J.
+Then case B would be our counterexample:
+Bratman's conditions are met but there is no shared intention.
+This is how our counterexample will work.
+
+Several preliminaries are necessary for the construction of our counterexample. 
+These preliminaries might easily give the impression that our counterexample depends on an artificial settings. 
+However, having introduced the primary counterexample in an artificial setting, we will  go on to show that counterexamples can also be constructed for mundane activities including walking together.
+
+%%
+%%two agents who intentionally J together (e.g.\ walk  together) and two agents who also  J together but do not do so intentionally (e.g.\ they happen to be walking side by side). 
+%%
+%This has two consequences.
+%First, it should be impossible to construct a pair cases, A and B,
+%meeting these criteria:
+%	(i) in each case, two agents act intentionally;
+%	(ii) in the first case, the two agents do not meet conditions (1)--(3) above;
+%	(iii) in the second case, the agents do meet conditions (1)--(3) above;
+%	(iv) the cases, A and B, are otherwise as similar as possible;
+%	and 
+%	(v) neither A nor B exemplifies the sort of shared agency gestured at by appeal to the contrast cases we started with.
+%Second, it should also be impossible to construct a pair cases, B and C,
+%meeting these criteria:
+%	(i$^\prime$) in each case, two agents act intentionally;
+%	(ii$^\prime$) in each case, the agents do meet conditions (1)--(3) above;
+%	and 
+%	(iii$^\prime$) C but not B exemplifies the sort of shared agency gestured at by appeal to the contrast cases we started with.
+%
+%
+%
+%
+%Let's stipulate that two cases \emph{differ with respect to shared agency} when they differ in the way exemplified by the contrast between our intentionally walking together and our merely walking in parallel.
+%Suppose, then, that we construct three cases, A, B and C.
+%The three cases are as similar as possible, except that the above conditions are met in cases B and C but not in case A.
+%Now if Bratman's conditions are sufficient for shared intention, A should differ from B with respect to shared agency, and B should not differ from C with respect to shared agency.
+%But suppose that in fact A does not differ from B with respect to shared agency, whereas B does so differ from C.%
+%\footnote{
+%Strictly speaking only A and B are needed for the counterexample, of course.
+%But including C makes it clearer that A and B do not differ with respect to shared agency.
+%}
+%Then we can conclude that Bratman's conditions are not sufficient for shared intention.
+%This is how our counterexample will work.
+
+\section{Neutral with Respect to Shared Intentionality}
+Before we can introduce the counterexample 
+we need to highlight a feature of Bratman's account 
+	which we shall be exploiting.
+
+Consider the contents of the intentions concerning our J-ing in the above clauses, (1)--(3). 
+What sort of activity can you intend when you intend that we J?
+We cannot restrict possible values of J to activities which involve shared agency.
+In imposing any such restriction we would be assuming the very notion that an account of shared intention is supposed to illuminate. 
+%
+%As we have stressed, 
+%	make it possible to 
+%	characterise systematically a difference between 
+%		cases involving shared agency (such as our walking together)
+%		and
+%		cases involving parallel agency only (such as two strangers who happen to be walking the same route side-by-side). 
+%We must therefore avoid tacitly appealing to this distinction by  restricting possible values of J to those which involve shared agency. 
 Rather the above  conditions, (1)--(3), must be sufficient for shared intention even for some values of J which are `neutral with respect to shared intentionality'.%
@@
 and that I will do this regardless of what you do.
-Suppose also that you can rationally intend that you go to Chicago in the same manner.
+Suppose also that you can rationally intend that you go to Chicago in the same manner,
+and that you know that if you act on this intention the upshot will be that we will go to Chicago together (although I may not know that we are going together---perhaps you will conceal your presence from me).
 Then you can rationally intend that we go to Chicago together, 
@@
 \citet{Bratman:1999fr}  defends this claim at length. 
-Note also that this claim must be true if Bratman's account of shared intention is to provide an informative way of distinguishing between the contrast cases mentioned at the start.
+Note also that this claim must be true if Bratman's account of shared intention is to provide an informative and systematic distinction between the contrast cases mentioned at the start.
 }
 
-In this paper we give a counterexample to Bratman's  claim that the above conditions, (1)--(3), are collectively sufficient conditions for shared intention. 
-We shall also suggest a revision which would enable the counterexample to be avoided.
-Apart from improving our understanding of what shared intention is, 
-this will also bear on the kinds of planning that are involved in acting together.
-
-Before going further we must distinguish two versions of the claim that (1)--(3) are collectively sufficient for shared intention.
-The \emph{weak claim} is that there is some J such that these conditions are sufficient for you and I to intend that we J.
-The \emph{strong claim} is that for any J, these conditions are sufficient for you and I to intend that we J.
-One of Bratman's aims is to show that it is possible to give an account of shared intention using concepts that `are available within the theory of individual planning agency' \citep[p.\ 163]{Bratman:2009lv}.
-Achieving this aim would require the strong claim, and it is to the strong claim that our counterexample is directed.
-
 \section{Unshared Intentions}
-Before we can introduce the counterexample we need to introduce what we shall call `unshared intentions'.
-We have an \emph{unshared intention} that we <J$_1$, J$_2$> where J$_1$$\neq$J$_2$ just if:
+As a further preliminary we need to introduce a definition.
+Let us stipulate that we have an \emph{unshared intention} that we <J$_1$, J$_2$> where J$_1$$\neq$J$_2$ just if:
 %
@@
 In this section we describe a possible situation in which two agents  have an unshared intention without irrationality, deception or even ignorance.
+This possible situation is not the promised counterexample, but it does form the basis for it.
 
@@
 Each movement carries a small cost to the player who moves, so that Ayesha and Benji each attempt to minimize how much he or she moves the cross consistently with completing his or her task.
+At the outset, 
 Ayesha and Benji are each neutral on whether the other succeeds or fails.
@@
  and the upshot will be that the cross hits the red square.
-Were this to happen, we could say, truthfully, that Ayesha and Benji made the cross hit the red square.
+Were this to happen, it would be true that Ayesha and Benji made the cross hit the red square.
 Suppose, further, that Ayesha can intend to perform those actions <a$_1$, a$_2$, ...\ a$_n$>.
@@
 will also result in the cross hitting the blue circle. 
-(Since the cross has momentum, we can suppose that it will hit both the red square and the blue circle at some time after these actions are performed.)
+(Since the cross has momentum, we can suppose that it will hit both the red square and the blue circle at some time after these action sequences have been performed.)
 Then by the reasoning just offered, Benji could intend that they, Ayesha and Benji make the cross hit the blue circle.
@@
 This is not quite enough to meet the second condition, (2), because there is also a requirement about meshing subplans. 
-To make sense of this requirement we need to generalise Bratman's definition of meshing:
+To make apply this requirement we need to generalise Bratman's definition of meshing:
 \begin{quote}
@@
 there would be a failure to mesh if, in intending that they make the cross hit the red square,  Ayesha's plans had included pushing Benji out of the way and seizing his controls. 
-But in the case we have been describing there is no such failure to mesh.
+There would also be a failure of mesh if Ayesha were planning to trick Benji into a situation where we would be unable to perform the actions he had been planning.
+But in the case we have been describing there are no such failures to mesh.
 Each agent's subplans involve manipulating his or her own controls,
@@
 
-\section{Pure unshared intentions and imaginary agents [*cut?]}
-An unshared intention is \emph{pure} if having that unshared intention does not involve having any shared intention. 
-It may initially appear obvious that all unshared intentions are pure. 
-This is not obvious, however.
-Since Bratman provides only sufficient conditions for shared intention, his account doesn't tell us that the unshared intention is not a shared intention.
-For all Bratman says, the conditions defining unshared intention might also be collectively sufficient conditions for shared intention. 
-And while we might be tempted to assume that for two agents to intentionally J together they must at least each have intentions concerning J, 
-this assumption might reasonably be rejected by those who, like Bratman, reject the Simple View according to which when an individual intentionally J-s she has an intention concerning her J-ing \citep{Bratman:1984jr}.
-So the purity of an unshared intention cannot be assumed without argument.
+%\section{Pure Unshared Intentions}
+%We stipulate that an unshared intention is \emph{pure} if having that unshared intention does not involve having any shared intention. 
+%It may be tempting to assume that all unshared intentions are pure. 
+%This is not obvious, however, and may not even be true.
+%Since Bratman provides only sufficient conditions for shared intention, his account doesn't tell us that an unshared intention is not a shared intention.
+%For all Bratman says, the conditions defining unshared intention might also be collectively sufficient conditions for shared intention. 
+%
+%But doesn't shared agency require that there be a single activity, J-ing, about which each the agent involved has an intention? 
+%This, too, is not obvious.
+It might reasonably be denied by those who, like Bratman, reject the Simple View according to which when an individual intentionally J-s she has an intention concerning her J-ing \citep{Bratman:1984jr}.
+%So the purity of an unshared intention cannot be assumed without argument. 
+%This is why we do not rely on this assumption in what follows.
+%
+%%
+%How can we show that Ayesha and Benji's unshared intention that they <J$_1$, J$_2$>  is pure?
+%Perhaps we can appeal to intuition.
+%Ayesha sees Benji's actions as constraints on her own, or else as opportunities.
+%She exploits Benji's intentions for her own ends.
+%Of course the situation is reciprocal: Benji exploits Ayesha in equal measure.
+%Each allows himself or herself to be exploited by the other because being exploited enables exploiting,
+%and this is the full extent of their cooperation. 
+%Perhaps it is clear enough that 
+%the sort of shared agency that an account of shared intention is supposed to capture must involve more than this sort of reciprocal exploitation, where each agent sees the other's actions only as constraints or opportunities.%
+%\footnote{
+%Note that we are not suggesting that reciprocal exploitation is incompatible with shared agency.  
+%The claim under consideration here is rather that shared agency requires more than reciprocal exploitation.
+%} 
+%If so, we can already claim that Ayesha and Benji have a pure unshared intention.
+%
+%
+%
+%
+%
+%We started this paper by explaining the need for a notion of shared intention by appeal to contrast cases (following Bratman). 
+%It is contrast cases that are supposed give us a pre-theoretical handle on shared intention.
+%So to show that Ayesha and Benji's unshared intention is pure, we need to construct a contrasting case.
+%
+%***Lily and Isabel, 
+%	like many people, 
+% 	sometimes attribute intentions and other states to imaginary agents, and perform actions on their behalf. 
+%In some cases this enables them to further their own real-world objectives (as in `Teddy wants to go to the park now').	
+%Some of these imaginary agents are toys. 
+%But some are imaginary aggregate agents. 
+%Sometimes, that is, Lily and Isabel imagine that there is an agent distinct from either of them and such that all its parts are parts of them.
+%They attribute intentions and other states to this agent, and act on its behalf.
+%
+%
+%
 
-For reasons to be explained later, we need to show that  Ayesha and Benji's unshared intention that they <J$_1$, J$_2$>  is pure.
-How could we do this? 
-Ayesha sees Benji's actions as constraints on her own, or else as opportunities.
-She exploits Benji's intentions for her own ends.
-Of course the situation is reciprocal: Benji exploits Ayesha in equal measure.
-Each allows himself or herself to be exploited by the other because being exploited enables exploiting.
-The sort of shared agency that an account of shared intention is supposed to capture must involve more than this sort of reciprocal exploitation, where each agent sees the other's actions only as constraints or opportunities.
-Intuitively shared agency must somehow involve greater unity.
-But how can this intuition support an argument?
-
-We started this paper by explaining the need for a notion of shared intention by appeal to contrast cases (following Bratman). 
-It is contrast cases that are supposed give us a pre-theoretical handle on shared intention.
-So to show that Ayesha and Benji's unshared intention is pure, we need to construct a contrasting case.
-
-***Lily and Isabel, 
-	like many people, 
- 	sometimes attribute intentions and other states to imaginary agents, and perform actions on their behalf. 
-In some cases this enables them to further their own real-world objectives (as in `Teddy wants to go to the park now').	
-Some of these imaginary agents are toys. 
-But some are imaginary aggregate agents. 
-Sometimes, that is, Lily and Isabel imagine that there is an agent distinct from either of them and such that all its parts are parts of them.
-They attribute intentions and other states to this agent, and act on its behalf.
-
-An account of shared intention should enable us to  distinguish systematically between two types of case in which we J together, one where we intentionally J together and the other where  our J-ing together involves  merely parallel actions.
-As noted, Gilbert's example of walking together shows that merely parallel actions can be mutually responsive, as when two strangers walking side by side avoid collision thanks to meshing subplans.
+%
+%An account of shared intention should enable us to  distinguish systematically between two types of case in which we J together, one where we intentionally J together and the other where  our J-ing together involves  merely parallel actions.
+%As noted, Gilbert's example of walking together shows that merely parallel actions can be mutually responsive, as when two strangers walking side by side avoid collision thanks to meshing subplans.
 
@@
 Yasmin and Zak meet Bratman's conditions (1)--(3) for having a shared intention that they J$_1$.  
-But these intentions, the unshared intention and the Bratman intention (as we might label the structure of intention and knowledge while leaving open whether it constitutes a shared intention), result in them performing same actions in the same way.
+But Yasmin, in planning and acting, does not rely on the coincidence of their intentions; and nor does Zak.
+(Yasmin relies on the fact Zak intends that they J$_1$, of course; but she does not rely on the fact that Zak intends what she intends.)
+Furthermore, due to an artefact of the way the game is structured,
+the unshared intention and the Bratman intention (as we might label the structure of intention and knowledge while leaving open whether it constitutes a shared intention), result in the two pairs performing same actions in the same way.
 That is, Yasmin reasons about Zak much as Ayesha reasons about Benji and Yasmin and does what Ayesha does, and likewise for Zak and Benji.
 
-We shall argue that Yasmin and Zak do not have a shared intention that they J$_1$ despite meeting meet Bratman's conditions (1)--(3).
-This is probably not yet obvious, however.
-Let us first attempt to provide some intuitive support for this claim.
-Return to Ayesha and Benji for a moment.
-Ayesha sees Benji's actions as constraints to work around and opportunities to exploit.
+We claim that Yasmin and Zak have a shared intention that they J$_1$
+only if 
+Ayesha and Benji have a shared intention.%
+\footnote{
+Strictly speaking, 
+	what matters for our argument is whether or not Ayesha and Benji have a shared intention \emph{in virtue of having the unshared intention that they <J$_1$, J$_2$>}.
+	This is because, strictly speaking, we need to show, not that Yasmin and Zak lack any shared intention whatsoever, but only that they lack a shared intention that they J$_1$ in virtue of meeting Bratman's conditions.
+	For ease of exposition this is not made explicit in the main text.
+} 
+This claim follows from the similarities of the two cases.
+The only difference is that Yasmin and Zak happen to be assigned the same task, whereas Ayesha and Benji are not.
+And neither Yasmin nor Zak makes use of the fact that they have the same task. 
+(This is not due to ignorance: it's just how they choose to approach their tasks.)
+So if we consider on how  
+		the case of Yasmin and Zak
+	differs from
+		 that of Ayesha and Benji,
+we can see that these differences do not plausibly amount to a difference with respect to shared agency.
+Shared intention cannot feature in one case but not the other.
+
+To show that the case of Yasmin and Zak do not have a shared intention it remains only for us to show that Ayesha and Benji do not have one. 
+Here we must be careful.
+First note that,
+	 since Bratman provides only sufficient conditions for shared intention, 
+	 his account doesn't tell us that an unshared intention is not a shared intention.
+For all Bratman says, the conditions defining unshared intention might  be sufficient for shared intention.
+
+But doesn't shared intention require at least this much,
+that there be a single activity  about which each the agent involved has an intention? 
+This might reasonably be doubted by those who, like Bratman, reject the Simple View according to which when an individual intentionally F-s she has an intention concerning her F-ing \citep{Bratman:1984jr}.
+So 
+	we shall not infer  that Ayesha and Zak lack a shared intention
+	just because (by construction) there is no F such that Ayesha and Zak's each intend that they, Ayesha and Zak, F.
+
+Can we then appeal directly to intuition to show that Ayesha and Zak lack a shared intention?
+Ayesha sees Benji's actions as constraints on her own, or else as opportunities.
 She exploits Benji's intentions for her own ends.
 Of course the situation is reciprocal: Benji exploits Ayesha in equal measure.
-Each allows himself or herself to be exploited by the other because being exploited enables exploiting.
-The sort of shared agency that an account of shared intention is supposed to capture must involve more than this sort of reciprocal exploitation, where each agent merely exploits the other's intentions.
-But Yasmin and Zak are engaged in just the same sort of reciprocal exploitation as Ayesha and Benji.
-The only difference is that, quite by chance, Yasmin's task and Zak's task are the same.
-So, intuitively, Yasmin and Zak, like Ayesha and Benji, belong in the same category as two people who happen to be walking side by side and not in the category exemplified by two people intentionally walking together.
+Each allows himself or herself to be exploited by the other because being exploited enables exploiting,
+and this is the full extent of their cooperation. 
+We don't suppose that reciprocal exploitation is incompatible with shared intention. 
+But Ayesha and Zak's interaction consists entirely in this sort reciprocal exploitation, where each agent sees the other's actions only as constraints or opportunities.
+Perhaps it is clear enough that 
+the sort of shared agency that an account of shared intention is supposed to capture must involve more than this. 
+If so, we can already claim that Ayesha and Benji are not acting on a shared intention.
+But philosophers' intuitions about shared agency may not be entirely, so it would be better if we could avoid such a blunt appeal to intuition.
 
+How else could we support the claim that Yasmin and Zak lack a shared intention? 
+As mentioned at the start, 
+the contrast cases are often used to anchor intuitions in theorising about shared agency.
+In the next section we shall further support our claim by contrasting Yasmin and Zak's case with a further case, one which is as similar as possible and which does seem to involve shared agency.
+The fact that Yasmin and Zak's case contrasts with this new case will support the claim that Yasmin and Zak lack shared intention.
+This is the aim of the following section.
+A first motive for introducing this new contrast case is that it provides an indication concerning what is missing from Bratman's account, and so may help us to understand why his conditions are not sufficient for shared intention.
+
+\section{Agent-neutral Plans}
+
+Above we introduced what will turn out to be a counterexample to Bratman's account, the case of Yasmin and Zak. 
+But we have yet to show that this case really is a counterexample. 
+For all we have said so far, a proponent of Bratman's view might insist that Yasmin and Zak do have a shared intention.
+To show that they do not,
+we shall contrast Yasmin and Zak's case with a third case that is as similar as possible but does involve shared intention.
+
+%---the following is a mistake (because Yasmin and Zak do take each other's task into account.
+%It is unlikely that humans would spontaneously behave as Yasmin and Zak do.%
+%\footnote{
+%This is nicely illustrated in a series of experiments by Sebanz and colleagues in which subject spontaneously take another's plans into account in planning their own actions, even where doing so is manifestly not required and can impair their performance  \citep{Sebanz:2003kf,Sebanz:2005fk,tsai:2011_groop_effect}.
+%}
+
+First we need to introduce the notion of an agent-neutral plan.
+We stipulate that a planning process, or a plan, is \emph{agent-neutral} just if it does not involve identifying any particular agents.  
+This sort of planning is quite common.
+For example, some housemates who have decided to take on an allotment to grow vegetables might sit down together to plan what needs doing without yet assigning roles to particular individuals. 
+In so planning, each housemate is thinking about what is to be done and not what she herself will do.  
+(This is an idealisation, of course.)
+At some point the planning is done  and they divide up the roles.
+Of course they may not find a way of dividing up roles that everyone is prepared to go along with---individuals' preferences, abilities and intentions may block the plan's adoption.
+But suppose the housemates do divide up the roles in a way that is acceptable to everyone, 
+	and that each implements her part in the plan.
+Then each conceives of her own and the others' actions as part of single plan directed to achieving a single outcome.
+
+In the above example of agent-neutral planning,
+	the housemates plan together and agree on a common plan.
+	Planning together is plausibly an activity which involves shared agency.
+	Note, however, that an individual can construct an agent-neutral plan by herself, even if its eventual execution will involve others.
+	In fact, two or more individuals who are assigned a task might each individually engage in agent-neutral planning in parallel.
+	The task demands and their planning strategies may conspire to ensure that they each come up with the same agent-neutral plan, or at least that their plans are compatible.%
+\footnote{
+Suppose that, for some outcome, two or more agents each have a plan for the realisation of that outcome (these plans are possibly but not necessarily agent-neutral).
+Then their plans are \emph{compatible} just if 
+	there is a way of assigning agents to any unspecified roles in each plan such that  each agent could rationally act on the intention that they realise the outcome by 	performing the role she is assigned in her own plan even though all facts about which agents are performing which roles in which plans were common knowledge to the agents.
+To illustrate, suppose that our task is to press a button simultaneously. 
+If your plan specifies that the agents start to move in exactly 60 seconds and press the button 5 seconds later whereas my plan specifies only that there is a leader and the agents are to press the button 5 seconds after the  leader starts moving, then our plans are compatible.
+}
+%
+Similarly, even if the agents continue to operate individually and without communicating or otherwise intentionally sharing information, the task demands and manifest properties of the agents, such as their distribution in space, may ensure that, having arrived at their plans, each agent assigns the same roles to the same individuals (or at least that each agent assigns roles in such a way that the resulting agent-specifying plans are compatible).
+%
+Finally, each agent may know enough about herself and the others to be able to determine, without communicating, whether the plan and role assignments will be acceptable to everyone. 
+	And all of this---that they engage in parallel agent-neutral planning resulting in compatible plans and role assignments, which are acceptable to all---may be common knowledge to the agents.
+	So it is possible, in principle at least, 
+	that several agents might each individually engage in agent-neutral planning and rationally perform their part in the resulting plan, knowing that the others will do likewise.
+	Parallel agent-neutral planning can rationally result in coordinated action without requiring shared agency.
+
+Now that we have some background on agent-neutral plans,
+let us introduce a third and final case.
+This case needs to be as similar as possible to that of Yasmin and Zak's while involving shared agency.
+
+Lily and Isabel start in the same situation as Yasmin and Zak. 
+Each is tasked with making the cross hit the red square (J$_1$). 
+Once again each cares only about her own success at the outset. 
+%---following version is for shortening, where the above background is skipped
+%Lily takes the view that the best way for her to succeed is to engage in a kind of agent-neutral planning.
+%A planning process is \emph{agent-neutral} when it does not involve identifying any particular agents.
+%This sort of planning is quite common; for example, some housemates, having decided to go camping, might sit down to plan what needs doing without yet assigning roles to particular individuals.
+%Similarly, Lily, knowing that she and Isabel have the same task,
+%plans how two agents in their situation could J$_1$.  
+Lily, knowing that she and Isabel have the same task, takes the view that the best way for her to succeed is plan how two agents in their situation could J$_1$.  
+So Lily ends up with an agent-neutral plan that specifies both her own actions and Isabel's actions without identifying her own actions as such (and without identifying Isabel's actions as Isabel's either).
+Having done this, Lily then assigns one role in the plan to herself and the to Isabel.
+At this point Lily considers whether she would be prepared to go along with the plan given her intentions, preferences and values, and she also considers whether Isabel would be prepared to go along with it too.
+In this case it happens that both would be prepared to go along with the plan.
+Lily then knowledgeably predicts that Isabel, who has similar planning abilities and has been approaching their task in a similar way, will have made a compatible plan.
+Lily therefore attempts to carry out her part in the plan she made; and Isabel does likewise.
+
+In short, then, Lily and Isabel are like Yasmin and Zak in nearly every respect. 
+Each pair has a Bratman intention that they J$_1$, each pair acts on this intention and each pair ends up performing the same sequence of actions.
+The difference is just that Yasmin and Zak make no use of the fact that they are performing the same task.
+	This fact does not show up in their planning. 
+	Rather, each plans her own actions only and treats the other's actions as constraints to work around or opportunities to exploit.
+By contrast, Lily and Isabel embrace the fact that they are performing the same task. 
+They may not like having to act together; in fact each may far prefer to act alone were that possible.
+And, like Yasmin and Zak, they are unconcerned with each other's success except insofar as their own success depends on it.
+But Lily and Isabel nevertheless make use of the fact that they have to perform a single task together by each constructing a single plan covering both of their actions and then carrying out their parts in these plans.
+So why does Lily and Isabel's case, but not Yasmin and Zak's case, plausibly involve shared agency?
+The reason is  this:
+	that at some stage of the planning which rationally guides and coordinates their actions,
+	Lily and Isabel each conceive of their actions as part of a single plan directed to achieving a single outcome.
+
+Given that Yasmin and Zak's case contrasts with Lily and Isabel's,
+we conclude that Yasmin and Zak do not have a shared intention.
+As explained above,
+ Yasmin and Zak do meet Bratman's conditions for shared intention, and do act appropriately on the corresponding intentions and knowledge. 
+ So Yasmin and Zak's case  is a counterexample to sufficiency of Bratman's conditions for shared intention.
+
+Note that in describing the case of Lily and Isabel we have not introduced contralateral commitments or other elements foreign to Bratman's account of shared agency. 
+Of course, we have not shown that these are not necessary for shared agency; 
+	but if we knew that they were necessary we would not need a counterexample. 
+Our counterexample is useful because it draws on the same planning resources Bratman's account draws on, with just one addition.
+The addition is the idea that agents can conceive, or fail to conceive, of their actions as part of a single plan.
+
+Given that Yasmin and Zak's case is a counterexample,
+there are also many further counterexamples involving mundane activities and less elaborate props.
+Here, for instance, is how Yasmin and Zak walk together.
+They are firmly tied at the ankle, and neither is strong enough to move without the other.
+Yasmin needs to get to the corner, and so does Zak.
+Further, they have a Bratman intention that they walk to the corner, and they act on this intention in walking to the corner.
+Again, this is not a case of shared agency.
+The problem is not that Yasmin and Zak are tied together against their will; 
+after all, many cases of shared agency are involuntary in this sense.
+The problem is rather that each conceives of the others' intentions and actions only as constraints and opportunities.
+They fail to make any use of the fact that they intend the same outcome,
+ and so they fail see their actions as part of a single plan.
+
+Yasmin and Zak's attitude may be unnatural.
+Perhaps humans in their situation would not normally ignore the fact that they have the same task in the way that these agents do.
+This may be why Bratman's conditions appear, misleadingly, to be sufficient for shared intention.
+
+\section{Revising Bratman's Account}
+How could Bratman's account be revised to avoid the counterexample?
+Perhaps we can simply add to Bratman's conditions.
+Consider the view that these are sufficient condition for us to have a shared intention that we J:
+%
+\begin{enumerate}
+\label{revised_account}
+\item (a) I intend that we J and (b) you intend that we J.
+ \item I intend that we J in accordance with and because of la, lb, and meshing subplans of la and lb; you intend that we J in accordance with and because of la, lb, and meshing subplans of la and lb.
+\item At some point in my planning, I produce a plan for all of our actions; and you do likewise.
+\item Our plans are compatible and we (partially) implement these plans.
+\item All of this, 1--4, is common knowledge between us.
+\end{enumerate}
+%
+This view avoids the counterexamples we constructed,
+but at the cost of a loss of generality.
+For these conditions are only met once agents have not only done some planning but also started to act.
+While the aim is not to provide necessary conditions for shared intention, we should aim for enough generality to illuminate some central cases.
+So we shall not consider this view further.
+
+For a different approach, consider activities which meet these conditions concerning some outcome G:
+\begin{enumerate}
+\item Each agent has a plan for how they, the agents, can G.  (This can be an agent-neutral plan plus an assignment of roles.)
+\item Each agent's plan is acceptable to every agent (it does not conflict with any individual's preferences, abilities or intentions).
+\item The agents' plans are compatible.
+\item Each agent performs the roles assigned to her in her own plan, and the agents thereby G. 
+\end{enumerate}
+
+%
+%*** each thinks what \emph{we} should do.  The fact that they have the same task shows up in their planning.
+%
+%
+%***Simple point: the fact that we have the same task should play a role in our planning.
+It is something that we should embrace, and perhaps even want.
+(Be careful: two people might intentionally act together although each would prefer to act alone, and would seize the opportunity to do so.)  Cordula's idea (modified) was that we aim for shared success (it's not a question of whether I care about \\emph{your} success, but I must care about \emph{our} success).
+
+\section{Conclusion}
+Bratman's conditions are not sufficient for shared intention because they leave open the possibility that  when two or more agents meet his conditions, each fails to rely on the fact that there is a single outcome to which all of their actions are directed and so conceives of the others' actions merely as constraints and opportunities.
+But Bratman's conditions appear to be sufficient for shared intention because it is natural that when two or more agents meet his conditions, they would rely on the fact that their is a single outcome to which all of their actions are directed and therefore be disposed to conceive of all their actions as parts of a single plan.
+The case of Yasmin and Zak shows that although this may be natural, it not necessary.
+
+\section{***OLD}
 Assuming for a moment that this is right,

### 2012-11-10 counterexample to bratman on shared intention/a counterexample to bratman on shared intention.tex
draft done, rough

@@
 
-\def \ititle {Is This a Counterexample to Bratman on Shared Intention?}
+\def \ititle {A Counterexample to Bratman on Shared Intention}
 \def \isubtitle {}
@@
 \noindent
-***
-
+At the centre of one leading account of shared agency, Michael Bratman's, is the claim that certain conditions are sufficient for us to have a shared intention that we J.
+This paper provides a counterexample to the sufficiency of these conditions and suggests a way of strengthening them to avoid the counterexample.
+%The counterexample exploits the possibility of two agents meeting Bratman's conditions for shared intention even though each conceives of the other's intentions and actions only as constraints to work around and opportunities to avoid.
+%Shared intention requires more than this: and it may require that the agents conceive of all of their actions as parts of a single plan.
 \end{abstract}
@@
 This question is standardly answered by appeal to contrast cases \citep[compare][p.\ 150]{Bratman:2009lv}.
-Thus \citet{gilbert_walking_1990} contrasts our intentionally walking together with two people who happen to be walking side by side. 
+Thus \citet{gilbert_walking_1990} contrasts friends intentionally walking together with two people who happen to be walking side by side. 
 And \citet{Searle:1990em} contrasts park visitors who  simultaneously run to a central shelter in performing a dance with park visitors who likewise run to the central shelter but only because of an impending storm.
@@
 \section{Neutral with Respect to Shared Intentionality}
+\label{sec:netural_wrt_shared_intentionality}
 Before we can introduce the counterexample 
@@
 \section{Agent-neutral Plans}
+\label{sec:distributed_plan}
 
@@
 In so planning, each housemate is thinking about what is to be done and not what she herself will do.  
-(This is an idealisation, of course.)
-At some point the planning is done  and they divide up the roles.
+%(This is an idealisation, of course.)
+At some point the housemates stop planning.
+(This does not necessarily mean that they have a fully worked out plan; like any other plans, agent-neutral plans can have gaps that may need filling in later.)
+They now divide up the roles.
 Of course they may not find a way of dividing up roles that everyone is prepared to go along with---individuals' preferences, abilities and intentions may block the plan's adoption.
@@
 	In fact, two or more individuals who are assigned a task might each individually engage in agent-neutral planning in parallel.
-	The task demands and their planning strategies may conspire to ensure that they each come up with the same agent-neutral plan, or at least that their plans are compatible.%
+	The task demands and their planning strategies may conspire to ensure that they each come up with the same agent-neutral plan, and task demands and manifest properties of the agents, such as their distribution in space, may ensure that each agent also assigns the same roles to the same individuals.
+(Strictly speaking it is not necessary for the plans and role assignments to be identical; it is enough if the resulting agent-specifying plans are, in a special sense, compatible.%
 \footnote{
-Suppose that, for some outcome, two or more agents each have a plan for the realisation of that outcome (these plans are possibly but not necessarily agent-neutral).
-Then their plans are \emph{compatible} just if 
-	there is a way of assigning agents to any unspecified roles in each plan such that  each agent could rationally act on the intention that they realise the outcome by 	performing the role she is assigned in her own plan even though all facts about which agents are performing which roles in which plans were common knowledge to the agents.
+\label{fn:df_compatible}
+Suppose that, for some outcome, two or more agents each have a plan for the realisation of that outcome. 
+(These plans may, but need not, specify roles for all of the agents; but the plans must be agent-specific, not agent-neutral.)
+By saying that these plans are \emph{compatible} we mean that:
+(i) 
+no agent would be prevented from performing the role she is assigned in her own plan by other agents performing the roles they are assigned in their plans;
+and
+(ii)
+if all facts about which agents have which roles in which plans were common knowledge to the agents,
+this would not affect the rationality of their each acting on the intention that they realise the outcome by performing the role she is assigned in her own plan.
+%	each agent could rationally act on the intention that they realise the outcome by performing the role she is assigned in her own plan,
+%	and she could do so even if all facts about which agents are performing which roles in which plans were common knowledge to the agents.
 To illustrate, suppose that our task is to press a button simultaneously. 
 If your plan specifies that the agents start to move in exactly 60 seconds and press the button 5 seconds later whereas my plan specifies only that there is a leader and the agents are to press the button 5 seconds after the  leader starts moving, then our plans are compatible.
-}
-%
-Similarly, even if the agents continue to operate individually and without communicating or otherwise intentionally sharing information, the task demands and manifest properties of the agents, such as their distribution in space, may ensure that, having arrived at their plans, each agent assigns the same roles to the same individuals (or at least that each agent assigns roles in such a way that the resulting agent-specifying plans are compatible).
+})
 %
@@
 
-\section{Revising Bratman's Account}
+\section{How to Fix Bratman's Account}
+\label{sec:fix}
+
 How could Bratman's account be revised to avoid the counterexample?
-Perhaps we can simply add to Bratman's conditions.
-Consider the view that these are sufficient condition for us to have a shared intention that we J:
+We might try strengthening Bratman's requirement about meshing subplans of intentions.
+This requirement is that, concerning our shared intention that we J, we each intend that we J in accordance with our intentions that we J and meshing subplans of them (see (2) \vpageref{quote:bratman_account}).
+The case of Yasmin and Zak suggests that this requirement is not strong enough.
+And the case of Lily and Isabel suggests that we might try strengthening it by appealing in some way to the idea that each agent can have a plan specifying roles for all of the agents.
+
+One way to strengthen Bratman's requirements is by constraining the possible ways in which agents who have a shared intention that they J may conceive of J in intending that they J.
+As we have seen, Bratman's account requires only that, in some cases, the agents conceive of J in a way that is neutral with respect to shared intentionality.
+We shall suggest a further requirement that is compatible with this one.
+
+We stipulate that an activity is \emph{distributed} just if: 
+	there is a single outcome, G,
+	and two or more agents each have a plan for G which specifies a role for all of the agents;	
+	these plans are compatible%
+\footnote{
+The notion of compatibility is defined in footnote \vref{fn:df_compatible}.
+}
 %
-\begin{enumerate}
-\label{revised_account}
-\item (a) I intend that we J and (b) you intend that we J.
- \item I intend that we J in accordance with and because of la, lb, and meshing subplans of la and lb; you intend that we J in accordance with and because of la, lb, and meshing subplans of la and lb.
-\item At some point in my planning, I produce a plan for all of our actions; and you do likewise.
-\item Our plans are compatible and we (partially) implement these plans.
-\item All of this, 1--4, is common knowledge between us.
-\end{enumerate}
+and each acceptable to all; 
+	each agent performs the roles assigned to her in her own plan;
+	the activity consists in nothing other than the agents performing these roles;
+	and
+	by performing these roles, the agents G.
+In specifying the class of distributed activities we have not made direct appeal to shared agency. 
+Nor is any covert appeal required, as our earlier discussion should  make clear (see Section \vref{sec:distributed_plan}).
+To use Bratman's term (see Section \vref{sec:netural_wrt_shared_intentionality}), distributed activities can be conceived of in ways that are neutral with respect to shared intentionality.
+So no circularity is involved in appealing to the notion of a distributed activity in explicating shared intention.
+
+Many ordinary interactions involve distributed activities.
+Two housemates have agreed to make a pizza and each start with an idea of what needs to be done.
+One starts by preparing the dough, so the other peels, washes and chops the vegetables.
+As the activity unfolds in this way, each gradually adds details to her plan and assigns roles.
+
+We should also note that distributed activities do not have to be very complex.
+In walking to a metro station together,
+you and I may each form and act on a plan that specifies some details of both of our movements.
+
+It is possible, then, that when two or more agents each intend that they do something, 
+	perhaps to paint a house,
+in having these intentions they conceive of the activity as a distributed activity.
+Where this happens, part of what the agents intend is, in effect, that they each construe their own actions as part of a single plan which also involves the others' plans.
+
+We can exploit this possibility in strengthening Bratman's conditions for shared intention.
+To avoid counterexamples like the case of Yasmin and Zak,
+we leave the conditions for shared intention exactly as Bratman specifies them (see \vpageref{quote:bratman_account})
+but add the further requirement that the agents must, in intending that they J, each conceive of J as a distributed activity.
+
+We have not shown, of course, that this modification to Bratman's account will enable it to avoid any other counterexamples.
+Perhaps further requirements are necessary.
+Or perhaps it is impossible to give informative sufficient conditions for shared intention.
+Our aim here has only been to show that there is a way of avoiding one type of counterexample without abandoning core features of Bratman's planning approach to shared intention.
+If, as we suppose, that general approach succeeds in illuminating at least one kind of shared agency, it needs to be modified in order to capture a simple intuition about shared agency.
+The intuition is this:
+	our intending to act in accordance with our intentions and meshing subplans of them is not sufficient for us to have a shared intention;
+	instead we must conceive of our actions as part of a single plan.
+A planning approach to shared agency can readily accommodate this intuition by invoking distributed activities.
+
 %
-This view avoids the counterexamples we constructed,
-but at the cost of a loss of generality.
-For these conditions are only met once agents have not only done some planning but also started to act.
-While the aim is not to provide necessary conditions for shared intention, we should aim for enough generality to illuminate some central cases.
-So we shall not consider this view further.
+%
+%Consider the view that these are sufficient condition for us to have a shared intention that we J:
+%%
+%\begin{enumerate}
+%\label{revised_account}
+%\item (a) I intend that we J and (b) you intend that we J.
+%% --- Bratman's condition
+%% \item I intend that we J in accordance with and because of la, lb, and meshing subplans of la and lb; you intend that we J in accordance with and because of la, lb, and meshing subplans of la and lb.
+%\item I intend that we J by way of our each having and partially implementing on a plan for J-ing which specifies roles for both of us, where these plans are compatible%
+%\footnote{
+%The notion of compatibility is defined in footnote \vref{fn:df_compatible}.
+%}
+%%
+%and each acceptable to both of us; and you intend likewise.
+%\item 1 and 2 are common knowledge between us.
+%\end{enumerate}
+%%
+%If the new second requirement, (2), is met then so is Bratman's.
+%(This is because our having compatible plans which specify a role for both us ensures that our subplans mesh given how 
+%
+%Requirements (1), (2) and (4) are Bratman's; the others are new.
+%This view would avoid the counterexamples,
+%but at a heavy cost.
+%For these conditions are only met once agents have not only done some planning but also started to act.
+%While the aim is not to provide necessary conditions for shared intention, we should aim to illuminate some central cases.
+%We can do better than this.
+%
+%For a different approach, 
+%
+%Consider two agents who meet Bratman's sufficient conditions for having a shared intention that they J, 
+%	(1)--(3) on page ***,
+%	and who each conceive of J as a distributed activity.
+%	So part of what they intend when they intend that they, the two agents, J is that they perform a distributed activity.
+%	For example, they might intend that they perform the distributed activity of making the cross hit the red square.
+%
 
-For a different approach, consider activities which meet these conditions concerning some outcome G:
-\begin{enumerate}
-\item Each agent has a plan for how they, the agents, can G.  (This can be an agent-neutral plan plus an assignment of roles.)
-\item Each agent's plan is acceptable to every agent (it does not conflict with any individual's preferences, abilities or intentions).
-\item The agents' plans are compatible.
-\item Each agent performs the roles assigned to her in her own plan, and the agents thereby G. 
-\end{enumerate}
+%	
+%	
+%
+%there are several agents 
+%
+%  which meet these conditions concerning some outcome, G:
+%%
+%\begin{enumerate}
+%\item Each agent has a plan for how they, the agents, can G.  (This can be an agent-neutral plan plus an assignment of roles.)
+%\item Each agent's plan is acceptable to every agent (it does not conflict with any individual's preferences, abilities or intentions).
+%\item The agents' plans are compatible.
+%\item Each agent performs the roles assigned to her in her own plan, and the agents thereby G. 
+%\end{enumerate}
+%%
+%
+%
+%***Can now get rid of the meshing subplans clause,
+%and probably the common knowledge clause too.
+%
 
@@
 \section{Conclusion}
+We started with some contrasts between shared agency and parallel individual agency.
+These provide one intuitive fix on the notion of shared intention: 
+ shared intention, whatever it is, is what distinguishes thinks like two friends walking somewhere together from things like two strangers who walk the same route side by side.
+By introducing new contrast cases,
+we provided a counterexample to the view that a certain interlocking structure of intention and knowledge is sufficient for shared intention. 
+Two agents might meet conditions
+Bratman offers as sufficient for shared intention---that is, it might be common knowledge to them that they might each intend that they J and intend that they J in accordance with and because of these intentions and meshing subplans of them---even though they lack a shared intention.
+
 Bratman's conditions are not sufficient for shared intention because they leave open the possibility that  when two or more agents meet his conditions, each fails to rely on the fact that there is a single outcome to which all of their actions are directed and so conceives of the others' actions merely as constraints and opportunities.
-But Bratman's conditions appear to be sufficient for shared intention because it is natural that when two or more agents meet his conditions, they would rely on the fact that their is a single outcome to which all of their actions are directed and therefore be disposed to conceive of all their actions as parts of a single plan.
-The case of Yasmin and Zak shows that although this may be natural, it not necessary.
+Bratman's conditions may appear to be sufficient for shared intention because it may be natural that when two or more agents meet his conditions, they would rely on the fact that there is a single outcome to which all of their actions are directed and therefore be disposed to conceive of all their actions as parts of a single plan.
+Our counterexample, the case of Yasmin and Zak, shows that although this may be natural, it not necessary.
 
-\section{***OLD}
-Assuming for a moment that this is right,
-what might explain the difference between Yasmin and Zak on the one hand and cases like Gilbert's two people intentionally walking together on the other hand?
-To answer this question in the spirit of Bratman's planning-theoretic approach, we need to step back and think about planning generally.
+There is a way to revise Bratman's account so as to avoid this sort of counterexample without departing from the spirit of his planning approach.
+The remedy we propose hinges on the possibility that sometimes an agent, in planning for an outcome, can specify both her own and other agent's actions.
+Further, in the right circumstances,
+two or more agents with the same task
+	can each individually specify roles for all of the agents in planning for that task.
+Even where such planning occurs in parallel (and so need not involve no shared agency), sometimes it can rationally result in coordinated action.
+These reflections suggest a way of appealing to the possibility that, in acting on a shared intention, each agent has a plan for all of their actions
+in order to capture the idea that agents acting on a shared intention conceive of all of their actions as part of a single plan.
+%One way to capture this idea, 
+%put roughly and without essential qualifications (see Section \vref{sec:fix}),
+% is by appeal to the idea that each agent has a plan for all of their actions.
 
-Imagine observing Ayesha and Benji as they play the game. 
-Suppose you decide to make a plan of action for one or both of them.
-In doing this you could either adopt one player's perspective or you could adopt a fictional shared perspective.
-Adopting one player's perspective, Ayesha's say, 
-	you think about what she might do to achieve her task,
-	predict how this will influence Benji's actions,
-	and then feed  these predictions back into the question of what Ayesha might do.
-From this perspective, Benji's actions are constraints to work around and opportunities to exploit.
-Alternatively you might adopt a fictional shared perspective, 
-	 mentally replacing Ayesha and Benji with a single, imaginary aggregate agent who has to perform both tasks alone.
-	 In this case, you think about what combination of actions would best achieve both tasks and then, right at the end, you assign these actions to Ayesha and Benji.
-From this perspective, it is as if there is a single mind behind the plan.
-
-Note that the number of agents involved does not dictate which perspective we can take.
-Just as we could take a fictional shared perspective where two or more agents are involved, 
-so also we could take a fictional split perspective on the actions of  a single agent, and even of ourselves.
-The newly elected  politician with a young family and working partner might see herself almost as if she were two people engaged in different projects while forced to shared a single body.
-
-Shared perspective of a fictional aggregate agent.
-Can be a matter of negotiation what the fictional aggregate agent's values, preferences and intentions are.
-
-Perhaps the reason why Yasmin and Zak fail to have a shared intention (if they do---this has not yet been argued) is that each plans entirely from her own perspective. 
-Maybe shared intention sometimes requires that agents be disposed to plan at least some of their actions from a shared perspective.
-
-Now consider a third (and final) pair of players, Lily and Isabel.
-Lily and Isabel each take the appropriate shared perspective, as if imagining a fictional aggregate agent who had been assigned their two tasks, 
-and they plan accordingly.
-And this is common knowledge.
-What matters is not what they do, however.
-It is that the J which they intend is characterised in terms of planning from a shared perspective.
-What they intend when they intend that they J does not leave it open whether they will adopt a shared perspective in planning their J-ing.
-J is neutral with respect to shared intention but it is not neutral with respect to the requirement that each agent individually adopt a shared perspective when planning how to J.
-***HERE
-
-Imagine two tasks need to be completed; perhaps you need to buy some shoes and get a haircut. 
-Which actions you perform in attempting the first task constrains what actions you can perform in attempting the second task, and conversely.
-For instance, if you shop for the shoes in one store, you won't have enough money left for the best hairdressers.
-There are two ways to approach planning these tasks.
-One approach involves treating the two tasks as separate.
-Perhaps you first think of a plan that is optimal for buying shoes.
-Then, taking this plan as a constraint, you attempt to plan the haircut.
-But as it happens you discover that the options for the haircut are unsatisfactory.
-So you revise the shoe-buying plan and try again to plan the haircut.
-A different approach to planning would be to treat the two tasks as part of a larger, single task.
-Perhaps you identify several possible combinations of shoe-buying and haircut-getting actions.
-You then ask yourself which combination is best overall and then perform the component actions.
-
-Take a situation like the game Ayesha and Benji are playing but now imagine that you are taking over from both players and intend to perform both tasks. 
-One way to plan would involve thinking of two separate tasks.
-First you think about the best way to make the cross hit the red square.
-
-%
-%The difference separating Ayesha and Benji, on the one hand, from Yasmin and Zak on the other 
-%	is unlike the difference separating two strangers who merely happen to be walking side by side from two people intentionally walking together.
-%So if Ayesha and Benji's do not intentionally act together, then nor do Yasmin and Zak.
-
-***Can get a parallel to the special case discussed here by considering walkers whose legs have been tied together in such a way that they are forced to accommodate each other if they are to get around.
-Such people may walk together and meet the Bratman conditions without actually
-
-We shall argue that although Yasmin and Zak meet Bratman's conditions (1)--(3) for having a shared intention that they J$_1$, they do not have any such shared intention.
-This, then, is our counterexample.
-Of course it is probably not yet obvious that Yasmin and Zak do not have a shared intention.
-To see that they do not we need to introduce one more case.
-
-We started this paper by explaining the need for a notion of shared intention by appeal to contrast cases. 
-An account of shared intention should enable us to  distinguish systematically between two types of case in which we J together, one where we intentionally J together and the other where  our J-ing together involves  merely parallel actions.
-As noted, Gilbert's example of walking together shows that merely parallel actions can be mutually responsive, as when two strangers walking side by side avoid collision thanks to meshing subplans.
-
-*Yasmin and Zak: it's like we've tied the Gilbert's walking strangers' ankles together, so that they can only succeed together.
-
-***
-
-Our argument is:
-\begin{enumerate}[label=\roman*]
-\item Yasmin and Zak have a shared intention that they J$_1$ only if Ayesha and Benji have a shared intention in virtue of having the 
- unshared intention that they <J$_1$, J$_2$>.
-\item Ayesha and Benji's unshared intention that they <J$_1$, J$_2$> is a not shared intention.
-\end{enumerate}
-%
-Therefore:
-%
-\begin{enumerate}[resume,label=\roman*]
-\item Yasmin and Zak do not have a shared intention that they J$_1$.
-\end{enumerate}
-%
-In the rest of this section we defend the two premises, starting with the second, (ii).
-
-Could Ayesha and Benji's unshared intention that they <J$_1$, J$_2$> be a  shared intention?
-This question might initially appear frivolous but it is not straightforward to answer.
-Since Bratman provides only sufficient conditions for shared intention, his account doesn't tell us that the unshared intention is not a shared intention.
-And while we might be tempted to assume that for two agents to intentionally J together they must at least each have intentions concerning J, 
-this assumption might reasonably be rejected by those who, like Bratman, reject the Simple View according to which when an individual intentionally J-s she has an intention concerning her J-ing \citep{Bratman:1984jr}.
-
-In defence of the second premise, (ii), consider that shared intentions are suppose to play a certain functional role.
-Among other things, they are supposed to coordinate
-
-\section{Prior counterexamples}
-***Must mention that Tollefsen and Gold \& Sugden counterexamples fail.
-
-\section{The Counterexample}
-Suppose that Aravinda runs the trains and Gerhard  the busses.
-Let us stipulate that the extent to which the two services, train and bus, are coordinated is to be measured by the total time passengers spend waiting between a bus and a train.
-So changes to the timetables would result in better coordination just if the changes would total waiting time.
-To illustrate, improving coordination might involve having trains arrive at a station shortly before, rather than shortly after, busses depart from there.
-Now Aravinda and Gerhard each intend that they, Aravinda and Gerhard, coordinate the trains with the busses to the greatest extent possible given other constraints; and they intend to do this by way of these intentions and meshing subplans of them, and this is common knowledge between them.
-Aravinda and Gerhard thus meet Bratman's sufficient conditions for them to have a shared intention.
-Acting on this intention, 
-every January Aravinda updates the train timetables and sends Gerhard the changes.
-Likewise, Gerhard  updates the bus timetables every June and sends Aravinda the changes.
-In this way  each is responsive to the other's intentions and subplans of these. 
-Indeed, each may even try to predict changes in the other's subplans and modify their own accordingly.  
-But from each individual's point of view, the other's plans are  merely a constraint.
-
-***
-
-So whether Gerhard succeeds relative to his intention depends on both his own and Aravinda's plans; and likewise for Aravinda.
-%This is why each intends that they optimally coordinate the services in accordance with meshing subplans.
-This is why each is responsive to the other's plans. 
-Indeed, each may even try to predict changes in the other's plans and modify their own accordingly.  
-But from each individual's point of view, the other's plans are  merely a constraint.
-This approach to planning suffers from two defects.
-First, it is unlikely to be optimal in the sense of resulting, eventually, in a combination of plans such that no other combination of plans would have been better for at least one service and no worse for either service.
-Second, it is unlikely to be efficient in the sense of allowing Gerhard and Aravinda to arrive at an optimal combination of plans for the two services, trains and buses, with the fewest iterations.
-How could they do better?
-One possibility may be to have a single plan covering busses and trains---perhaps, for example, Aravinda could buy Gerhard's franchise. 
-%But suppose that this is not possible, and that there are limits on how much information Aravinda and Gerhard can share. 
-Now there is a single goal to which Aravinda and Gerhard's activities are both directed.
-But suppose that this is not possible, and that they are unable to integrate their planning more tightly---practical constraints or regulation prevents them from opening up their planning to each other.
-Can they still do better than treat each other's plans as a constraint?
-Possibly.
-Each plans the whole bus--train operation.
-
-New contrast case: Aravinda and Gerhard planning the whole thing vs. merely responding to each other's plans.
-Meet Bratman's sufficient conditions for shared intention in both cases.
-(Specifically, in both cases they act on a shared intention that they coordinate the trains with the busses to the greatest extent possible given other constraints.)
-But intuitively is a case of coordinating the trains with the busses together whereas the other case involves Aravinda and Gerhard merely doing this in parallel.
-So shared intention is insufficient to fully explain the contrast cases.
-
-I'm also inclined to think it is not necessary for solving the contrast cases (distributive goals ...\ collective goals).
-The contrast cases do not provide a firm anchor for theorising about shared intention.
+No amount of acting on intentions about other's intentions is not sufficient for shared agency.  
+But conceiving of one's own and other's intentions and actions as part of a single plan might be.
 

### 2012-11-10 intention motor representation goal ascription with corrado sinigaglia (paper 2)/Motor Representation in Goal Ascription.tex
cleaning up

@@
 \label{sec:processes}
-Our aim in this section is to explain how motor representations could sometimes have content-respecting influences on judgements about the goals of observed actions.
-Broadly, our will be proposal is that the influence of motor representation on observational judgements goes via experience.
+Our aim in this and the following sections is to explore how motor representations could sometimes have content-respecting influences on judgements about the goals of observed actions.
+Broadly, our proposal is that the influence of motor representation on observational judgements goes via experience.
 It is because motor representations can affect our experiences when we observe actions that those motor representations can facilitate judgements about the actions.
-However this is consistent with two quite different hypotheses concerning the experiences associated with motor representation in action observation.
+However this broad idea is consistent with two quite different hypotheses concerning the experiences associated with motor representation in action observation.
 
-One hypothesis, call it the Indirect Hypothesis, is based on evidence that, in observing action, motor representation can influence visual and acoustic experience.
-(Perhaps this holds for other modalities too).
-On this hypothesis, these influences enable an observer to identify goals to which the observed action is directed because the observer is additionally able to recognise actions from their associated sensory effects.
+One hypothesis, call it the Indirect Hypothesis, is based on evidence that, in observing action, motor representation can influence visual and acoustic experiences (see figure \vref{fig_indirect_hypothesis}).
+(Perhaps this holds for other modalities too.) The relevant experiences are perceptual experiences of bodily configurations and movements as an action unfolds, and perceptual experiences of characteristic sensory effects of actions.
+Such experiences sometimes enable an observer to make judgements about the goals of actions, for she may associate particular patterns of bodily movement or particular sensory effects with particular goals.
+But in many cases of action observation, the agent's bodily movements and the sensory effects of her actions may be partially obscured or surrounded by irrelevant distractors, or otherwise difficult to identify.
+But suppose that, as some evidence suggests, motor representation can influence---and even enhance---such experiences of bodily configurations and movements, and of sensory effects.
+Then motor representations might thereby facilitate judgements about the goals of actions.
+So on the Indirect Hypothesis, motor representations sometimes enhance perceptual experiences of movements or sensory effects which in turn sometimes facilitate judgements about the goals of actions.
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=\textwidth]{fig_indirect_hypothesis.png}
+\caption{
+\label{fig_indirect_hypothesis}
+	The Indirect Hypothesis.
+}
+\end{center}
+\end{figure}
+
 To illustrate, suppose that an observer can associate the characteristic sound of a newspaper being snapped open with the action of snapping open a newspaper.
-Suppose also that, when observing an action, motor representations can influence whether, when or how she experiences this characteristic sound.
+Suppose also that, when observing an action, motor  representations can influence whether, when or how this observer experiences this characteristic sound.
 Then motor processes can indirectly influence the speed and accuracy of the observer's judgements about which actions are directed to this goal, the snapping open of a newspaper, by influencing whether, when or how she experiences the characteristic sound of this action.%
 \footnote{
-The Indirect Hypothesis is inspired by and consistent with views defended in \citet{Csibra:2007fy} and *Knoblich.
+The Indirect Hypothesis is inspired by and consistent with views defended in \citet{Csibra:2007fy} and \citet{Wilson:2005qu}.
 However, to our knowledge these authors have not considered exactly our question about content-respecting influences of motor representation on judgement, and we are not suggesting that the would endorse the Indirect Hypothesis in its entirety.
@@
 The other hypothesis is in some ways more radical.
-This hypothesis, call it the Direct Hypothesis, is that in observing action we experience not only movements and sounds but also goal-directed actions as such.
-Such experiences---call the motor experiences---stand to motor representations roughly as visual experiences stand to visual representations.
+This hypothesis, call it the Direct Hypothesis, is that in observing action we experience not only movements and sounds but also goal-directed actions as such (see figure \vref{fig_direct_hypothesis}).
+Such experiences---call them motor experiences---stand to motor representations somewhat as visual experiences stand to visual representations.
 The motor experience of an action as directed to a particular outcome is made possible by a motor representation of that outcome.
-So motor representations can have a content-respecting influence on judgements because motor representations enable motor experiences on which judgements are based.%
+So motor representations can have a content-respecting influence on judgements because motor representations enable motor experiences on which the judgements can be  based.
+Or so the Direct Hypothesis claims.%
 \footnote{
@@
 
-Let us say that an experience, motor or not, is revelatory of action just if it  can provide the subject of experience with reasons for correct judgements about which action someone is performing.
+\begin{figure}
+\begin{center}
+\includegraphics[width=\textwidth]{fig_direct_hypothesis.png}
+\caption{
+\label{fig_direct_hypothesis}
+	The Direct Hypothesis.
+}
+\end{center}
+\end{figure}
+
+Let us say that an experience, motor or not, is \emph{revelatory of action} just if it can provide the subject of experience with reasons for correct judgements about which action someone is performing.
 On both hypotheses motor representations enable experiences which are revelatory of action.
 And both hypotheses use this broad idea to explain how motor representations could have content-respecting influences on judgements.
-The primary difference between the hypotheses concerns what these revelatory experiences are experiences of: whether all such are experiences of movements, sounds and the like only (the Indirect Hypothesis), or whether some are experiences of goal-directed actions as such (the Direct Hypothesis).
+The primary difference between the hypotheses concerns what these revelatory experiences are experiences of: whether all are experiences of movements, sounds and the like only (the Indirect Hypothesis), or whether some are experiences of goal-directed actions as such (the Direct Hypothesis).
 
-In the next section we expand on and consider evidence for the first hypothesis before turning to the second hypothesis in the following section.
+In the next section we expand on and consider evidence for the Indirect Hypothesis before turning to the Direct Hypothesis in the following section.
 
 \section{The Indirect Hypothesis}
-The Indirect Hypothesis is depends on the claim that motor representations can influence sensory experiences.
-This claim is supported by a variety of evidence.
-\citet{bortoletto:2011_action} show that motor plans, such as might be involved in producing a gesture with the hand, can influence visual processes \citep{bortoletto:2011_action}.
-The influence of motor representations on visual processes varies depending on both the gesture to be performed and the visual stimulus presented.
-Importantly, these influences can be detected before any action occurs, showing that it is the motor representations rather than any consequent movements which are influencing visual processes.
-[*Some other evidence; stress that the effect can be found using different measures.]  As it may be possible to influence visual processes in ways that make no difference to experience, we do not take these findings alone to establish the claim that motor representations could enable visual experiences.
-But these findings do at least show that the claim is consistent with the neuropsychological facts.%
+The Indirect Hypothesis depends on the claim that motor representations can influence perceptual experiences.
+If this is right, we would expect to find that motor representation can have a content-respecting influence on perceptual processing.
+There is indeed evidence of such influences.
+For instance, \citet{bortoletto:2011_action} show that motor plans, such as might be involved in producing a gesture with the hand, can influence visual processes.
+Here the influence of motor representations on visual processes varies depending on both the gesture being planned and the visual stimulus presented.
+Importantly, these influences can be detected before any action occurs, showing that it is the motor representations rather than any consequent movements which are influencing visual processes.%
 \footnote{
@@
 }  
+This evidence shows that the claim about motor representations influencing experience is at least consistent with how motor and perceptual processes interact.
 
-What would establish that some motor representations enable perceptual experiences?
+Of course, this evidence does not show that motor representation influences perceptual experiences.
+After all, it may be possible to influence sensory processes in ways that make no difference to experience.
+But there is also evidence that motor representation can influence perceptual judgements.
 For some very special pairs of tones, the first is sometimes perceived as lower in pitch than the second and sometimes as higher in pitch.
-Repp and Knoblich (\citeyear{repp:2007_action}) had a group of expert pianists and a control group of non-experts press two keys in sequence, where the first key was sometimes left and sometimes right of the second key.
-They found that, for the expert pianists, the direction of the key presses modulated the experience: it influenced the perceived direction of the change in pitch.
-Could what modulates experience in this case be not a motor representation but the occurrence of a movement, or perhaps even the perception of a movement of the subject's own fingers?
-Against this, note that the effect was not observed in non-expert pianists: for them the direction of action did not measurably influence the experienced pitch direction.
-Since both groups moved in the same way, if the modulation were due to movement only we would expect it to occur irrespective of expertise.
-Instead it seems likely that differences in expertise between the two groups of subjects affected how the movements they performed were represented motorically, and these differences in motor representation are in turn what explains their experience of the change in pitch.
+Call these \emph{ambiguous tone pairs}.
+Repp and Knoblich (\citeyear{repp:2007_action}) had a group of expert pianists and a control group of non-experts press two keys in sequence, where the first key was sometimes to the left, and sometimes to the right, of the second key.
+The key presses produced an ambiguous tone pair, and the tones always occurred in the same order regardless of which key was pressed first.
+By asking subjects to report how they perceived the relative pitch of the tones, Repp and Knoblich found that, for the expert pianists, the direction of the key presses influenced the perceived direction of the change in pitch.
+Could what influences perceptual experience in this case be not a motor representation but the occurrence of a movement, or perhaps even the perception of a movement of the subject's own fingers?
+Against this possibility, note that the effect was not observed in non-expert pianists: for them the direction of action did not measurably influence the perceived pitch direction.
+Since both groups moved in the same way, if the influence were due to movement only we would expect it to occur irrespective of expertise.
+Instead it seems likely that differences in expertise between the two groups of subjects, expert and non-expert, affected how the movements they performed were represented motorically, and that these differences in motor representation are in turn what explains their perception of relative pitch.
 This study, and others like it,%
 \footnote{
-See also \citet{zwickel:2010_interference} who investigate effects of action on visual experience of motion.  *more
+See also \citet{zwickel:2010_interference} who investigate effects of action on visual experience of motion.  *more \citep{schutz-bosbach:2007_perceptual}
 } 
-provide relatively direct evidence that motor representations can enable acoustic experiences.
+provide relatively direct evidence that motor representations can influence acoustic experiences.
 
-How do these findings bear on the second hypothesis above, according to which motor representation in action observation enables experiences of sounds and movements that somehow facilitate the identification of actions?
-Perhaps the motor representations enable certain experiences of kinematic features which somehow enable subjects to better identify actions.
-We do not claim that this is entirely plausible, only that the second hypothesis cannot be rejected out of hand.
-This is enough to show that we cannot infer that there are experiences of action just from the fact that motor representation enables experiences which are revelatory of action.
-Stronger evidence for the first hypothesis is needed.
+How do these findings bear on the Indirect Hypothesis above, according to which motor representation in action observation enables experiences of sounds and movements that somehow facilitate the identification of actions?
+Consider observing someone playing two notes on a piano.
+Given that you know something about which sounds are produced by various piano playing actions, hearing the relative pitch of these notes might in principle assist you in recognizing that a goal of her action is to play these two notes.
+So if, as the findings indicate, motor representation can influence perceptions of pitch, it might thereby indirectly facilitate judgements about the goals of actions.
 
-\subsection{Experiences of action}
-Our focus is experiences involved in observing actions, but let us first consider experiences involved in performing actions.
+Of course the evidence we have considered leaves open the question of how extensive the influence of motor representations on perceptual experiences is.
+For all we have said, such influences may be rare.
+This is a potential problem for the Indirect Hypothesis, which requires such influences to occur in every case in which motor representation facilitates goal ascription.
+However our aim in this paper is not to establish the correctness of an explanation but, more modestly, only to understand how motor representation could in principle facilitate judgements about the goals of actions.
+Wilson and Knoblich (\citeyear{Wilson:2005qu}, p.\ 463) consider the possibility that motor representation may influence the representation of agents' bodies and their movements.
+They suggest that motor representation in action observation may help `to fill in missing or ambiguous information' and to project the likely course of an agent's bodily movements.
+This idea allows us to see that the Indirect Hypothesis is at least in principle capable of explaining how motor representation might facilitate goal ascription in a wide range of cases.
+Perhaps, then, motor representations of outcomes to which actions are directed influence---and even enhance---perceptual experiences of the agents' bodies and movements in ways that facilitate judgements about the goals of the actions.
+So the Indirect Hypothesis.
+
+MR--->PE of configuration \& movements ---->> Judg
+MR ---> PE of pitch -->Judg about goal
+Pitch: MR influence Perc [no action observation]
+
+\subsection{The Direct Hypothesis: Experiences of Action}
+Our aim is to understand how motor representations of outcomes can have a content-respecting influence on judgements about the goals of actions. 
+According to the Direct Hypothesis, a motor representation of an outcome can underpin an experience of an action as directed to a matching outcome, and this experience can in turn ground a judgement about the action's being directed to a matching outcome.
+This hypothesis hinges on two potentially controversial claims.
+The first is that in observing action we experience not only movements, sounds and other sensory effects of actions but also the goal-directed actions as such.
+And the second is that such experiences are underpinned by motor representations of corresponding outcomes.
+
+Is the first claim plausible?  Consider the following argument:
+%
+\begin{enumerate}
+\item
+There is a way of imagining acting which involves experiencing goal-directed actions as such.
+
+\item
+There are sometimes elements common to what you experience when you imagine performing a goal-directed action and what you experience when you observe someone else performing that action; in particular, the experience of action is common to both.
+\end{enumerate}
+Therefore:
+\begin{enumerate}[resume]
+\item
+Sometimes observing a goal-directed action involves experiencing it as such.
+\end{enumerate}
+%
+Without attempting to show that the premises are true, we shall offer some considerations which indicate that they are at least defensible.
+
+On the first premise, (1), let us start by observing that there are two ways of imagining acting.
+One, cognitive way of imagining involves something like thinking about performing an action.
+Another, phenomenologically action-like way of imagining involves something almost like acting except that such imaginings are not necessarily responsive to the features of actual objects and do not necessarily result in bodily movements.
+Suppose that you are about to perform a delicate operation with some scissors.
+You stand poised to act, but first pause to mentally pantomime performing the action.
+In some respects the experiences involved in this imaginative exercise may be barely distinguishable from experiences that might be involved in actually performing the operation (\citealp[p.\ 161]{currie:1997_mental}; \citealp[p.\ 727]{jeannerod:1995_mental}; \citealp[p.\ 638-9]{kosslyn:2001_neural}).
+In fact, imagining performing the operation in advance has some of the benefits of actually rehearsing it.
+For imagining performing an action in this phenomenologically action-like way sometimes resembles actually performing the action with respect to timing \citep{decety:1989_timing, decety:1996_imagined, Jeannerod:1994oz} and the sorts of factors that make acting more or less awkward \citep{parsons:1994_temporal, frak:2001_orientation}.
+The first premise, (1), concerns this phenomenologically action-like way of imagining.
+
+What do we experience when we imagine acting?  ***HERE
+a.  There are no movements or sensory effects to experience (but there is  no action produced either).
+b.  There is motor representation of action (but there may also be visual representation of sensory effects of the action not performed).
+c.
+I could be mistaken about the movements involved in an action and about its sensory effects but still know that I am imagining it (and not just because I intend to imagine it and then do some imaginative thing or other---if my mind wanders so that I start imaginig a different action, I could know this).
+(one relevant consideration is that you know not only which action you are imagining but when you have succeeded in imagining performing it); ()
+(corollary: there can be experiences of action without experiences of acting)
+
 We shall argue that acting sometimes involves experiences which can ground judgements about what one has done even though these are not experiences of movements or sounds.

### 2012-11-14 counterexample to bratman on shared intention/a counterexample to bratman on shared intention.tex
very rough
first draft done, proof read (sort of)

@@
 
-\def \ititle {A Counterexample to Bratman on Shared Intention}
+\def \ititle {Shared Agency Involves Changing Perspective: A Counterexample to Bratman}
 \def \isubtitle {}
@@
 \noindent
-At the centre of one leading account of shared agency, Michael Bratman's, is the claim that certain conditions are sufficient for us to have a shared intention that we J.
-This paper provides a counterexample to the sufficiency of these conditions and suggests a way of strengthening them to avoid the counterexample.
-%The counterexample exploits the possibility of two agents meeting Bratman's conditions for shared intention even though each conceives of the other's intentions and actions only as constraints to work around and opportunities to avoid.
-%Shared intention requires more than this: and it may require that the agents conceive of all of their actions as parts of a single plan.
+The leading, best developed account of shared agency, Michael Bratman's, hinges on the claim that, roughly,
+we have a shared intention if we each intend that we J and also that we J by way of these intentions and meshing subplans of them where this is all common knowledge.
+This paper provides a counterexample to the sufficiency of this condition for shared intention.
+The counterexample arises because it is possible for agents to meet the condition, and to act rationally on the specified knowledge and intentions,
+ while conceiving of each other's actions only as constraints to work around and opportunities to exploit.
+Shared agency requires more than this, or so we argue.
+We also suggest a way of revising Bratman's account  that is consistent with his general approach.
+In some or all cases, shared agency differs from individual agency in part because shared agency involves changing perspective so that one’s own and other’s actions are conceived of as parts of a single plan.
+
 \end{abstract}
@@
 These and other contrast cases invite the question, 
-How do cases involving shared agency differ from cases involving parallel agency only?
+How do cases involving shared agency differ from cases involving only  parallel individual agency?
 
@@
 
-`3. 1 and 2 are common knowledge between us' \citep[][p.\ View 4]{Bratman:1993je}
+`3. 1 and 2 are common knowledge between us' \citep[][p.\ View 4]{Bratman:1993je}.
 \end{quote}
@@
 The \emph{strong claim} is that for any J, these conditions are sufficient for you and I to intend that we J. 
-One of Bratman's aims is to show that an account of shared intention can be \emph{conceptually conservative}:
-he aims, that is, to show that it is possible to give an account of shared intention using concepts that `are available within the theory of individual planning agency' \citep[p.\ 163]{Bratman:2009lv}.
-Achieving this aim would require the strong claim, and it is to the strong claim that our counterexample is directed.
+Our initial counterexample will be directed to the strong claim.
+However, after we have constructed this counterexample it will become clear that there are also counterexamples involving a wide range mundane activities to which Bratman explicitly takes his account to apply.
+Even walking together%
+	\footnote{
+	See \citet[p.\ 150]{Bratman:2009lv}.
+	}
+	%
+and painting a house together%
+	\footnote{
+	See \citet[p.\ 331]{Bratman:1992mi}.
+	}
+	%
+can serve as counterexamples.
+%One of Bratman's aims is to show that an account of shared intention should be \emph{conceptually conservative}.
+%He aims, that is, to show that  it is possible to give an account of shared intention using only concepts that `are available within the theory of individual planning agency' \citep[p.\ 163]{Bratman:2009lv}.
+%Achieving this aim would require the strong claim, and it is to the strong claim that our counterexample is directed.%
+%\footnote{
+%We do not aim to attack the claim that it is possible to give a conceptually conservative account of shared intention. 
+%The point we are making here is just that if an account
+%}
 
@@
 		and
-		cases involving parallel agency only (such as two strangers who happen to be walking the same route side-by-side). 
+		cases involving only parallel  individual agency (such as two strangers who happen to be walking the same route side-by-side). 
 Suppose, then, that we had a trio of cases, A, B and C, each involving two agents.
@@
 This is how our counterexample will work.
+(Strictly speaking, case A is not necessary; but including it simplifies exposition.)
 
@@
 
+To illustrate, first consider Bratman's `mafia case' where two people go to New York together by virtue of one of them forcing the other into a car and driving off \citep[p.\ 333]{Bratman:1992mi}. 
+Next consider a paradigm case of where two friends cooperatively go to New York together.
+It's possible to conceive of an act type of going to New York together which is broad enough that both cases, the coercive and the cooperative, fall under it.
+And it's even possible to desire or intend to be involved in an act of this type.
+A gangster might intend to go to New York with a future victim without having yet decided whether they will do this cooperatively or coercively.
+
 A consequence is that, in the right situations, one of us can rationally intend that we J, and can intend this unilaterally, that is without depending on anyone else intending that we J. 
-Suppose you know that I am going to Chicago via a certain route at a particular time, 
+Suppose you know that I am going to New York via a certain route at a particular time, 
 and that I will do this regardless of what you do.
-Suppose also that you can rationally intend that you go to Chicago in the same manner,
-and that you know that if you act on this intention the upshot will be that we will go to Chicago together (although I may not know that we are going together---perhaps you will conceal your presence from me).
-Then you can rationally intend that we go to Chicago together, 
-	and you can intend this irrespective of whether I have any corresponding intention---providing, of course, that in so intending you are conceiving of our going to Chicago together in a way that is neutral with respect to shared intentionality.
+Suppose also that you can rationally intend that you go to New York in the same manner,
+and that you know that if you act on this intention the upshot will be that we will go to New York together (although I may not know that we are going together---perhaps you will conceal your presence from me).
+Then you can rationally intend that we go to New York together, 
+	and you can intend this irrespective of whether I have any corresponding intention---providing, of course, that in so intending you are conceiving of our going to New York together in a way that is neutral with respect to shared intentionality.
 What follows depends on the premise that this is indeed possible.%
@@
 \label{df:unshared_intention}
-1. (a) I intend that we J$_1$ and (b) you intend that we J$_2$
+1$^\prime$. (a) I intend that we J$_1$ and (b) you intend that we J$_2$
 
-2.
+2$^\prime$.
 I intend that we J$_1$ in accordance with and because of la, lb, and meshing subplans of la and lb; you intend that we J$_2$ in accordance with and because of la, lb, and meshing subplans of la and lb
 
-3. 1 and 2 are common knowledge between us.
+3$^\prime$. 1 and 2 are common knowledge between us.
 \end{quote}
@@
 Let us first introduce the activity we shall focus on.
-Ayesha and Benji  are playing a simple video game which involves moving a cross around a two-dimensional space littered with barriers.
+Ayesha and Ahmed  are playing a simple video game which involves moving a cross around a two-dimensional space littered with barriers.
 Ayesha can only accelerate the cross backwards or forwards,
-while Benji can only accelerate it left or right. 
+while Ahmed can only accelerate it left or right. 
 The cross moves around and interacts with the barriers in ways both players can predict.
@@
 	while
-	Benji's task is make the cross hit the blue circle. 
+	Ahmed's task is make the cross hit the blue circle. 
 In general it is possible that either or both will succeed, or that they will both fail.
-Each movement carries a small cost to the player who moves, so that Ayesha and Benji each attempt to minimize how much he or she moves the cross consistently with completing his or her task.
+Each movement carries a small cost to the player who moves, so that Ayesha and Ahmed each attempt to minimize how much he or she moves the cross consistently with completing his or her task.
 At the outset, 
-Ayesha and Benji are each neutral on whether the other succeeds or fails.
+Ayesha and Ahmed are each neutral on whether the other succeeds or fails.
 They are not opponents and do not seek to undermine each other's efforts, but each is entirely concerned  with his or her own task.
-All of this is common knowledge for Ayesha and Benji.
+All of this is common knowledge for Ayesha and Ahmed.
 They both know who has which task, what constraints they face and what their motives are.
@@
 Consider the possibility of one player intending, unilaterally, that the two players do something.
-Suppose that one of the players---Ayesha, say---can knowledgeably predict that if she performs a certain sequence actions, <a$_1$, a$_2$, ...\ a$_n$>, then Benji will simultaneously perform certain other actions, <b$_1$, b$_2$, ...\ b$_n$>,
+Suppose that one of the players---Ayesha, say---can knowledgeably predict that if she performs a certain sequence actions, <a$_1$, a$_2$, ...\ a$_n$>, then Ahmed will simultaneously perform certain other actions, <b$_1$, b$_2$, ...\ b$_n$>,
  and the upshot will be that the cross hits the red square.
-Were this to happen, it would be true that Ayesha and Benji made the cross hit the red square.
+Were this to happen, it would be true that Ayesha and Ahmed made the cross hit the red square.
 Suppose, further, that Ayesha can intend to perform those actions <a$_1$, a$_2$, ...\ a$_n$>.
-Then Ayesha can intend, unilaterally, that they, Ayesha and Benji, make the cross hit the red square.
+Then Ayesha can intend, unilaterally, that they, Ayesha and Ahmed, make the cross hit the red square (see Section \vref{sec:netural_wrt_shared_intentionality}).
 
@@
 	Ayesha's <a$_1$, a$_2$, ...\ a$_n$>  and 
-	Benji's <b$_1$, b$_2$, ...\ b$_n$>,
+	Ahmed's <b$_1$, b$_2$, ...\ b$_n$>,
 will also result in the cross hitting the blue circle. 
 (Since the cross has momentum, we can suppose that it will hit both the red square and the blue circle at some time after these action sequences have been performed.)
-Then by the reasoning just offered, Benji could intend that they, Ayesha and Benji make the cross hit the blue circle.
-So Ayesha and Benji could meet the first condition, (1), for having an unshared intention.
+Then by the reasoning just offered, Ahmed could intend that they, Ayesha and Ahmed make the cross hit the blue circle.
+So Ayesha and Ahmed could meet the first condition, (1), for having an unshared intention.
 
@@
 Suppose that Ayesha knows two further things.
-First, that Benji intends that they, Ayesha and Benji, make the cross hit the blue circle.
-Second, that in acting on his intention Benji will perform actions <b$_1$, b$_2$, ...\ b$_n$>.
-Then Ayesha can intend that they, Ayesha and Benji, make the cross hit the red square in accordance with and because of her intention that they make the cross hit the red square and in accordance with and because of Benji's intention that they make the cross hit the blue circle.
+First, that Ahmed intends that they, Ayesha and Ahmed, make the cross hit the blue circle.
+Second, that in acting on his intention Ahmed will perform actions <b$_1$, b$_2$, ...\ b$_n$>.
+Then Ayesha can intend that they, Ayesha and Ahmed, make the cross hit the red square in accordance with and because of her intention that they make the cross hit the red square and in accordance with and because of Ahmed's intention that they make the cross hit the blue circle.
 
 This is not quite enough to meet the second condition, (2), because there is also a requirement about meshing subplans. 
-To make apply this requirement we need to generalise Bratman's definition of meshing:
+To apply this requirement we need to generalise Bratman's definition of meshing:
 \begin{quote}
@@
 To illustrate, 
-there would be a failure to mesh if, in intending that they make the cross hit the red square,  Ayesha's plans had included pushing Benji out of the way and seizing his controls. 
-There would also be a failure of mesh if Ayesha were planning to trick Benji into a situation where we would be unable to perform the actions he had been planning.
-But in the case we have been describing there are no such failures to mesh.
+their subplans would fail to mesh if, in intending that they make the cross hit the red square,  Ayesha's plans had included pushing Ahmed out of the way and seizing his controls. 
+They would also fail to mesh if Ayesha were planning to trick Ahmed into a situation where we would be unable to perform the actions he had been planning.
+But in the case we have been describing, the agents subplans mesh perfectly.
 Each agent's subplans involve manipulating his or her own controls,
 and the successes each seeks in doing this depends on the other successfully carrying out their subplans.
-So Ayesha can rationally intend that they, Ayesha and Benji, make the cross hit the red square in accordance with and because of their intentions and meshing subplans of them.
-And Benji likewise for making the cross hit the blue circle.
+So Ayesha can rationally intend that they, Ayesha and Ahmed, make the cross hit the red square in accordance with and because of their intentions and meshing subplans of them.
+And Ahmed likewise for making the cross hit the blue circle.
 
-The only outstanding requirement for Ayesha and Benji to have an unshared intention is that their various intentions are common knowledge. 
-Assuming common knowledge is possible where agents have a shared intention, it is likewise possible in this case of unshared intention.
-So Ayesha and Benji can have an unshared intention that they <J$_1$, J$_2$> where J$_1$ is Ayesha and Benji's making the cross hit the red square and J$_2$ is their making the cross hit the blue circle.
+The only outstanding requirement for Ayesha and Ahmed to have an unshared intention is that their various intentions are common knowledge. 
+Assuming common knowledge is possible concerning Bratman's conditions (1)--(2) on page \pageref{quote:bratman_account}, 
+it is likewise possible for the corresponding conditions on unshared intention, (1$^\prime$)--(2$^\prime$) on page  \pageref{df:unshared_intention}.
+So Ayesha and Ahmed can have an unshared intention that they <J$_1$, J$_2$> where J$_1$ is Ayesha and Ahmed's making the cross hit the red square and J$_2$ is their making the cross hit the blue circle.
 
@@
 %%
-%How can we show that Ayesha and Benji's unshared intention that they <J$_1$, J$_2$>  is pure?
+%How can we show that Ayesha and Ahmed's unshared intention that they <J$_1$, J$_2$>  is pure?
 %Perhaps we can appeal to intuition.
-%Ayesha sees Benji's actions as constraints on her own, or else as opportunities.
-%She exploits Benji's intentions for her own ends.
-%Of course the situation is reciprocal: Benji exploits Ayesha in equal measure.
+%Ayesha sees Ahmed's actions as constraints on her own, or else as opportunities.
+%She exploits Ahmed's intentions for her own ends.
+%Of course the situation is reciprocal: Ahmed exploits Ayesha in equal measure.
 %Each allows himself or herself to be exploited by the other because being exploited enables exploiting,
@@
 %} 
-%If so, we can already claim that Ayesha and Benji have a pure unshared intention.
+%If so, we can already claim that Ayesha and Ahmed have a pure unshared intention.
 %
@@
 %It is contrast cases that are supposed give us a pre-theoretical handle on shared intention.
-%So to show that Ayesha and Benji's unshared intention is pure, we need to construct a contrasting case.
+%So to show that Ayesha and Ahmed's unshared intention is pure, we need to construct a contrasting case.
 %
-%***Lily and Isabel, 
+%***Caitlin and Ciaran, 
 %	like many people, 
@@
 %But some are imaginary aggregate agents. 
-%Sometimes, that is, Lily and Isabel imagine that there is an agent distinct from either of them and such that all its parts are parts of them.
+%Sometimes, that is, Caitlin and Ciaran imagine that there is an agent distinct from either of them and such that all its parts are parts of them.
 %They attribute intentions and other states to this agent, and act on its behalf.
@@
 In the situation just described,
-Ayesha and Benji are playing a game and have different tasks.
+Ayesha and Ahmed are playing a game and have different tasks.
 Thanks to special features of the game environment, they both succeed by acting on an unshared intention.
-Now compare two further players, Yasmin and Zak, who are playing the same game. 
-Their situations, knowledge states, intentions and actions are as similar as possible to Ayesha's and Benji's except for one detail.
-Just by chance they have been assigned identical tasks: Yasmin's task is to make the cross hit the red square and Zak's task is the same.
-So where Ayesha and Benji have an unshared intention that they <J$_1$, J$_2$>,
-Yasmin and Zak meet Bratman's conditions (1)--(3) for having a shared intention that they J$_1$.  
-But Yasmin, in planning and acting, does not rely on the coincidence of their intentions; and nor does Zak.
-(Yasmin relies on the fact Zak intends that they J$_1$, of course; but she does not rely on the fact that Zak intends what she intends.)
+Now compare two further players, Beatrice and Baldric, who are playing the same game. 
+Their situations, knowledge states, intentions and actions are as similar as possible to Ayesha's and Ahmed's except for one detail.
+Just by chance they have been assigned identical tasks: Beatrice's task is to make the cross hit the red square and Baldric's task is the same.
+So where Ayesha and Ahmed have an unshared intention that they <J$_1$, J$_2$>,
+Beatrice and Baldric meet Bratman's conditions (1)--(3) for having a shared intention that they J$_1$.  
+But Beatrice, in planning and acting, does not rely on the coincidence of their intentions; and nor does Baldric.
+(Beatrice relies on the fact Baldric intends that they J$_1$, of course; but she does not rely on the fact that Baldric intends what she intends.)
 Furthermore, due to an artefact of the way the game is structured,
-the unshared intention and the Bratman intention (as we might label the structure of intention and knowledge while leaving open whether it constitutes a shared intention), result in the two pairs performing same actions in the same way.
-That is, Yasmin reasons about Zak much as Ayesha reasons about Benji and Yasmin and does what Ayesha does, and likewise for Zak and Benji.
+the unshared intention and the 
+	Bratman intention 
+	\label{df:bratman_intention}
+	(as we might label the structure of intention and knowledge while leaving open whether it constitutes a shared intention), result in the two pairs performing same actions in the same way.
+That is, Beatrice reasons about Baldric much as Ayesha reasons about Ahmed and Beatrice and does what Ayesha does, and likewise for Baldric and Ahmed.
 
-We claim that Yasmin and Zak have a shared intention that they J$_1$
+We claim that Beatrice and Baldric have a shared intention that they J$_1$
 only if 
-Ayesha and Benji have a shared intention.%
+Ayesha and Ahmed have a shared intention.%
 \footnote{
 Strictly speaking, 
-	what matters for our argument is whether or not Ayesha and Benji have a shared intention \emph{in virtue of having the unshared intention that they <J$_1$, J$_2$>}.
-	This is because, strictly speaking, we need to show, not that Yasmin and Zak lack any shared intention whatsoever, but only that they lack a shared intention that they J$_1$ in virtue of meeting Bratman's conditions.
+	what matters for our argument is whether Ayesha and Ahmed have a shared intention \emph{in virtue of having the unshared intention that they <J$_1$, J$_2$>}.
+	This is because, strictly speaking, we need to show, not that Beatrice and Baldric lack any shared intention whatsoever, but only that they lack a shared intention that they J$_1$ just in virtue of meeting Bratman's conditions (see (1)--(3) on page  \pageref{quote:bratman_account}).
 	For ease of exposition this is not made explicit in the main text.
@@
 This claim follows from the similarities of the two cases.
-The only difference is that Yasmin and Zak happen to be assigned the same task, whereas Ayesha and Benji are not.
-And neither Yasmin nor Zak makes use of the fact that they have the same task. 
+The only difference is that Beatrice and Baldric happen to be assigned the same task, whereas Ayesha and Ahmed are not.
+And neither Beatrice nor Baldric makes use of the fact that they have the same task. 
 (This is not due to ignorance: it's just how they choose to approach their tasks.)
 So if we consider on how  
-		the case of Yasmin and Zak
+		the case of Beatrice and Baldric
 	differs from
-		 that of Ayesha and Benji,
+		 that of Ayesha and Ahmed,
 we can see that these differences do not plausibly amount to a difference with respect to shared agency.
@@
 
-To show that the case of Yasmin and Zak do not have a shared intention it remains only for us to show that Ayesha and Benji do not have one. 
+To show that Beatrice and Baldric do not have a shared intention it remains only for us to show that Ayesha and Ahmed do not have one. 
 Here we must be careful.
@@
 So 
-	we shall not infer  that Ayesha and Zak lack a shared intention
-	just because (by construction) there is no F such that Ayesha and Zak's each intend that they, Ayesha and Zak, F.
+	we shall not infer  that Ayesha and Ahmed lack a shared intention
+	just because (by construction) there is no F such that Ayesha and Ahmed each intend that they, Ayesha and Ahmed, F.
 
-Can we then appeal directly to intuition to show that Ayesha and Zak lack a shared intention?
-Ayesha sees Benji's actions as constraints on her own, or else as opportunities.
-She exploits Benji's intentions for her own ends.
-Of course the situation is reciprocal: Benji exploits Ayesha in equal measure.
+Can we then appeal directly to intuition to show that Ayesha and Ahmed lack a shared intention?
+Ayesha sees Ahmed's actions as constraints on her own, or else as opportunities.
+She exploits Ahmed's intentions for her own ends.
+Of course the situation is reciprocal: Ahmed exploits Ayesha in equal measure.
 Each allows himself or herself to be exploited by the other because being exploited enables exploiting,
@@
 We don't suppose that reciprocal exploitation is incompatible with shared intention. 
-But Ayesha and Zak's interaction consists entirely in this sort reciprocal exploitation, where each agent sees the other's actions only as constraints or opportunities.
+But Ayesha and Ahmed's interaction consists entirely in this sort reciprocal exploitation, where each agent sees the other's actions only as constraints or opportunities.
 Perhaps it is clear enough that 
 the sort of shared agency that an account of shared intention is supposed to capture must involve more than this. 
-If so, we can already claim that Ayesha and Benji are not acting on a shared intention.
+If so, we can already claim that Ayesha and Ahmed are not acting on a shared intention.
 But philosophers' intuitions about shared agency may not be entirely, so it would be better if we could avoid such a blunt appeal to intuition.
 
-How else could we support the claim that Yasmin and Zak lack a shared intention? 
+How else could we support the claim that Beatrice and Baldric lack a shared intention? 
 As mentioned at the start, 
 the contrast cases are often used to anchor intuitions in theorising about shared agency.
-In the next section we shall further support our claim by contrasting Yasmin and Zak's case with a further case, one which is as similar as possible and which does seem to involve shared agency.
-The fact that Yasmin and Zak's case contrasts with this new case will support the claim that Yasmin and Zak lack shared intention.
+In the next section we shall further support our claim by contrasting Beatrice and Baldric's case with a further case, one which is as similar as possible and which does seem to involve shared agency.
+The fact that Beatrice and Baldric's case contrasts with this new case will support the claim that Beatrice and Baldric lack shared intention.
 This is the aim of the following section.
@@
 
-Above we introduced what will turn out to be a counterexample to Bratman's account, the case of Yasmin and Zak. 
+Above we introduced what will turn out to be a counterexample to Bratman's account, the case of Beatrice and Baldric. 
 But we have yet to show that this case really is a counterexample. 
-For all we have said so far, a proponent of Bratman's view might insist that Yasmin and Zak do have a shared intention.
+For all we have said so far, a proponent of Bratman's view might insist that Beatrice and Baldric do have a shared intention that they J$_1$.
 To show that they do not,
-we shall contrast Yasmin and Zak's case with a third case that is as similar as possible but does involve shared intention.
+we shall contrast Beatrice and Baldric's case with a third case that is as similar as possible but does involve shared intention.
 
-%---the following is a mistake (because Yasmin and Zak do take each other's task into account.
-%It is unlikely that humans would spontaneously behave as Yasmin and Zak do.%
+%---the following is a mistake (because Beatrice and Baldric do take each other's task into account.
+%It is unlikely that humans would spontaneously behave as Beatrice and Baldric do.%
 %\footnote{
@@
 	In fact, two or more individuals who are assigned a task might each individually engage in agent-neutral planning in parallel.
-	The task demands and their planning strategies may conspire to ensure that they each come up with the same agent-neutral plan, and task demands and manifest properties of the agents, such as their distribution in space, may ensure that each agent also assigns the same roles to the same individuals.
+	The task demands and their planning strategies may conspire to ensure that they each come up with the same agent-neutral plan.
+	The task demands and manifest properties of the agents, such as their distribution in space, may also ensure that each agent also assigns the same roles to the same individuals.
 (Strictly speaking it is not necessary for the plans and role assignments to be identical; it is enough if the resulting agent-specifying plans are, in a special sense, compatible.%
@@
 (i) 
-no agent would be prevented from performing the role she is assigned in her own plan by other agents performing the roles they are assigned in their plans;
+no agent would normally be prevented from performing the role she is assigned in her own plan by other agents performing the roles they are assigned in their plans;
 and
@@
 To illustrate, suppose that our task is to press a button simultaneously. 
-If your plan specifies that the agents start to move in exactly 60 seconds and press the button 5 seconds later whereas my plan specifies only that there is a leader and the agents are to press the button 5 seconds after the  leader starts moving, then our plans are compatible.
+If your plan is that each of us will to start to move in exactly 60 seconds and press the button 5 seconds later whereas my plan specifies that you are the leader and we are each to press the button 5 seconds after you start moving, then our plans are compatible.
 })
@@
 Finally, each agent may know enough about herself and the others to be able to determine, without communicating, whether the plan and role assignments will be acceptable to everyone. 
-	And all of this---that they engage in parallel agent-neutral planning resulting in compatible plans and role assignments, which are acceptable to all---may be common knowledge to the agents.
+	And all of this---that they engage in parallel, agent-neutral planning resulting in identical (or compatible) plans and role assignments, which are acceptable to all---may be common knowledge to the agents.
 	So it is possible, in principle at least, 
 	that several agents might each individually engage in agent-neutral planning and rationally perform their part in the resulting plan, knowing that the others will do likewise.
-	Parallel agent-neutral planning can rationally result in coordinated action without requiring shared agency.
+	Parallel, agent-neutral planning can rationally result in coordinated action without presupposing shared agency.%
+\footnote{
+Note that we are claiming only that shared agency is not presupposed.
+Our view is consistent with (but does not depend on) the claim
+that
+	if some agents each engage in parallel, agent-neutral planning and then rationally perform their part in the resulting plan,
+	the upshot would be an exercise of shared agency.
+}
+
+Different perspectives are involved in
+	forming a plan for  your own actions  only
+and 
+	forming a plan for  your own and others actions.
+%\footnote{
+%Perhaps it is impossible, strictly speaking, to plan others' actions. 
+%In speaking loosely about planning your own and others actions, we mean the activity of constructing an agent-neutral plan and assigning roles as described above.
+%}
+  Of course, in planning your only your own actions, it may be necessary to consider how others' actions will affect, and be affected by, your won.
+But in this case, the others' actions feature in your planning as constraints and opportunities.
+By contrast, forming a plan for your own and others' actions 
+	involves conceiving of your own and the others' actions as elements in a single plan directed to a single outcome.
 
@@
 let us introduce a third and final case.
-This case needs to be as similar as possible to that of Yasmin and Zak's while involving shared agency.
+This case needs to be as similar as possible to Beatrice and Baldric's while involving shared agency.
 
-Lily and Isabel start in the same situation as Yasmin and Zak. 
+Caitlin and Ciaran start in the same situation as Beatrice and Baldric. 
 Each is tasked with making the cross hit the red square (J$_1$). 
@@
 %---following version is for shortening, where the above background is skipped
-%Lily takes the view that the best way for her to succeed is to engage in a kind of agent-neutral planning.
+%Caitlin takes the view that the best way for her to succeed is to engage in a kind of agent-neutral planning.
 %A planning process is \emph{agent-neutral} when it does not involve identifying any particular agents.
 %This sort of planning is quite common; for example, some housemates, having decided to go camping, might sit down to plan what needs doing without yet assigning roles to particular individuals.
-%Similarly, Lily, knowing that she and Isabel have the same task,
+%Similarly, Caitlin, knowing that she and Ciaran have the same task,
 %plans how two agents in their situation could J$_1$.  
-Lily, knowing that she and Isabel have the same task, takes the view that the best way for her to succeed is plan how two agents in their situation could J$_1$.  
-So Lily ends up with an agent-neutral plan that specifies both her own actions and Isabel's actions without identifying her own actions as such (and without identifying Isabel's actions as Isabel's either).
-Having done this, Lily then assigns one role in the plan to herself and the to Isabel.
-At this point Lily considers whether she would be prepared to go along with the plan given her intentions, preferences and values, and she also considers whether Isabel would be prepared to go along with it too.
+Caitlin, knowing that she and Ciaran have the same task, takes the view that the best way for her to succeed is plan how two agents in their situation could J$_1$.  
+So Caitlin ends up with an agent-neutral plan.
+She next assigns one role in the plan to herself and the other to Ciaran.
+At this point Caitlin considers whether she would be prepared to go along with the plan given her intentions, preferences and values, and she also considers whether Ciaran would be prepared to go along with it too.
 In this case it happens that both would be prepared to go along with the plan.
-Lily then knowledgeably predicts that Isabel, who has similar planning abilities and has been approaching their task in a similar way, will have made a compatible plan.
-Lily therefore attempts to carry out her part in the plan she made; and Isabel does likewise.
+Caitlin then knowledgeably predicts that Ciaran, who has similar planning abilities and has been approaching their task in a similar way, will have made an identical or compatible%
+\footnote{
+The notion of compatible is defined in footnote \vref{fn:df_compatible}.
+}
+%
+ plan which, like hers, is acceptable to both of them.
+Finally, Caitlin attempts to carry out her part in her plan; and Ciaran does likewise.
 
-In short, then, Lily and Isabel are like Yasmin and Zak in nearly every respect. 
-Each pair has a Bratman intention that they J$_1$, each pair acts on this intention and each pair ends up performing the same sequence of actions.
-The difference is just that Yasmin and Zak make no use of the fact that they are performing the same task.
+In short, then, Caitlin and Ciaran are like Beatrice and Baldric in nearly every respect. 
+Each pair has a Bratman intention that they J$_1$,%
+\footnote{
+As stipulated above (see page \pageref*{df:bratman_intention}), 
+two agents have a \emph{Bratman intention} that they J just if they meet conditions (1)--(3) on page \pageref{quote:bratman_account}.
+}
+%
+ each pair acts on this intention and each pair ends up performing the same sequence of actions.
+The difference is just that Beatrice and Baldric make no use of the fact that they are performing the same task.
 	This fact does not show up in their planning. 
 	Rather, each plans her own actions only and treats the other's actions as constraints to work around or opportunities to exploit.
-By contrast, Lily and Isabel embrace the fact that they are performing the same task. 
+By contrast, Caitlin and Ciaran embrace the fact that they are performing the same task. 
 They may not like having to act together; in fact each may far prefer to act alone were that possible.
-And, like Yasmin and Zak, they are unconcerned with each other's success except insofar as their own success depends on it.
-But Lily and Isabel nevertheless make use of the fact that they have to perform a single task together by each constructing a single plan covering both of their actions and then carrying out their parts in these plans.
-So why does Lily and Isabel's case, but not Yasmin and Zak's case, plausibly involve shared agency?
+And, like Beatrice and Baldric, they are unconcerned with each other's success except insofar as their own success depends on it.
+But Caitlin and Ciaran nevertheless make use of the fact that they have to perform a single task together by each constructing a single plan covering both of their actions and then carrying out their parts in these plans.
+So why does Caitlin and Ciaran's case, but not Beatrice and Baldric's, plausibly involve shared agency?
 The reason is  this:
-	that at some stage of the planning which rationally guides and coordinates their actions,
-	Lily and Isabel each conceive of their actions as part of a single plan directed to achieving a single outcome.
+	 at some stage of the planning which rationally guides and coordinates their actions,
+	Caitlin and Ciaran each conceive of their actions as part of a single plan directed to achieving a single outcome.
 
-Given that Yasmin and Zak's case contrasts with Lily and Isabel's,
-we conclude that Yasmin and Zak do not have a shared intention.
-As explained above,
- Yasmin and Zak do meet Bratman's conditions for shared intention, and do act appropriately on the corresponding intentions and knowledge. 
- So Yasmin and Zak's case  is a counterexample to sufficiency of Bratman's conditions for shared intention.
+Given the way Beatrice and Baldric's case contrasts with Caitlin and Ciaran's,
+we conclude that Beatrice and Baldric do not have a shared intention.
+But, by construction, 
+ Beatrice and Baldric do meet Bratman's conditions for shared intention, and they do act appropriately on the corresponding intentions and knowledge. 
+ So Beatrice and Baldric's case  is a counterexample to sufficiency of Bratman's conditions for shared intention (see (1)--(3) on page \pageref{quote:bratman_account}).
 
-Note that in describing the case of Lily and Isabel we have not introduced contralateral commitments or other elements foreign to Bratman's account of shared agency. 
-Of course, we have not shown that these are not necessary for shared agency; 
-	but if we knew that they were necessary we would not need a counterexample. 
-Our counterexample is useful because it draws on the same planning resources Bratman's account draws on, with just one addition.
+In describing Caitlin and Ciaran we have not introduced contralateral commitments or other elements foreign to Bratman's account of shared agency. 
+Our counterexample draws on the same planning resources Bratman's account draws on, with just one addition.
 The addition is the idea that agents can conceive, or fail to conceive, of their actions as part of a single plan. 
+In at least some cases, shared agency differs from individual agency in part because shared agency involves changing perspective so that one’s own and other’s actions are conceived of as parts of a single plan.
 
-Given that Yasmin and Zak's case is a counterexample,
+Given that Beatrice and Baldric's case is a counterexample,
 there are also many further counterexamples involving mundane activities and less elaborate props.
-Here, for instance, is how Yasmin and Zak walk together.
+Here, for instance, is how Beatrice and Baldric walk together.
 They are firmly tied at the ankle, and neither is strong enough to move without the other.
-Yasmin needs to get to the corner, and so does Zak.
+Beatrice needs to get to the corner, and so does Baldric.
 Further, they have a Bratman intention that they walk to the corner, and they act on this intention in walking to the corner.
-Again, this is not a case of shared agency.
-The problem is not that Yasmin and Zak are tied together against their will; 
+As before, each plans her own walking in the light of her predictions about how the other will walk, treating the other's actions as constraints and opportunities.
+This is not a case of shared agency.
+The reason is not that Beatrice and Baldric are tied together against their will; 
 after all, many cases of shared agency are involuntary in this sense.
-The problem is rather that each conceives of the others' intentions and actions only as constraints and opportunities.
+The reason their walking does not involve shared agency 
+is rather that each conceives of the others' intentions and actions only as constraints and opportunities.
 They fail to make any use of the fact that they intend the same outcome,
@@
 
-Yasmin and Zak's attitude may be unnatural.
-Perhaps humans in their situation would not normally ignore the fact that they have the same task in the way that these agents do.
+Beatrice and Baldric's attitude may seem unnatural. 
+Perhaps humans in their situation would not normally fail to exploit the fact that they have the same task in the way that these agents do.
 This may be why Bratman's conditions appear, misleadingly, to be sufficient for shared intention.
@@
 We might try strengthening Bratman's requirement about meshing subplans of intentions.
-This requirement is that, concerning our shared intention that we J, we each intend that we J in accordance with our intentions that we J and meshing subplans of them (see (2) \vpageref{quote:bratman_account}).
-The case of Yasmin and Zak suggests that this requirement is not strong enough.
-And the case of Lily and Isabel suggests that we might try strengthening it by appealing in some way to the idea that each agent can have a plan specifying roles for all of the agents.
+This requirement is that, concerning our shared intention that we J, we each intend that we J in accordance with our intentions that we J and meshing subplans of them (see (2) on page \pageref{quote:bratman_account}).
+The case of Beatrice and Baldric suggests that this requirement is not strong enough.
+And the case of Caitlin and Ciaran suggests that we might try strengthening it by appealing in some way to the idea that each agent can have a plan specifying roles for all of the agents.
 
-One way to strengthen Bratman's requirements is by constraining the possible ways in which agents who have a shared intention that they J may conceive of J in intending that they J.
+One way to strengthen Bratman's account would be by adding a condition to his three.
+An alternative, which we prefer, is to introduce a new restriction on how agents who have a shared intention may conceive of J in intending that they J.
 As we have seen, Bratman's account requires only that, in some cases, the agents conceive of J in a way that is neutral with respect to shared intentionality.
@@
 We stipulate that an activity is \emph{distributed} just if: 
-	there is a single outcome, G,
+\begin{enumerate}[label=\emph{\alph*})]
+\item 	there is a single outcome, G,
 	and two or more agents each have a plan for G which specifies a role for all of the agents;	
-	these plans are compatible%
+\item 	these plans are identical or compatible,%
 \footnote{
@@
 %
-and each acceptable to all; 
-	each agent performs the roles assigned to her in her own plan;
-	the activity consists in nothing other than the agents performing these roles;
+		and they are each acceptable to all; 
+\item	each agent performs the roles assigned to her in her own plan;
+\item	the activity consists in nothing other than the agents performing these roles;
 	and
-	by performing these roles, the agents G.
+\item	by performing these roles, the agents G.
+
+\end{enumerate}
+%
 In specifying the class of distributed activities we have not made direct appeal to shared agency. 
 Nor is any covert appeal required, as our earlier discussion should  make clear (see Section \vref{sec:distributed_plan}).
-To use Bratman's term (see Section \vref{sec:netural_wrt_shared_intentionality}), distributed activities can be conceived of in ways that are neutral with respect to shared intentionality.
-So no circularity is involved in appealing to the notion of a distributed activity in explicating shared intention.
+To use Bratman's term, distributed activities can be conceived of in ways that are neutral with respect to shared intentionality (see Section \vref{sec:netural_wrt_shared_intentionality}).
+So we can appeal  to the notion of a distributed activity in explicating shared intention without circularity.
 
 Many ordinary interactions involve distributed activities.
-Two housemates have agreed to make a pizza and each start with an idea of what needs to be done.
+Two housemates have agreed to make a pizza. 
+Each start with an idea of what needs doing.
 One starts by preparing the dough, so the other peels, washes and chops the vegetables.
@@
 
-We should also note that distributed activities do not have to be very complex.
+Distributed activities can be simple.
 In walking to a metro station together,
-you and I may each form and act on a plan that specifies some details of both of our movements.
+two friends may each form and act on a plan that specifies some details of both of their movements.
+This is enough for their activity to be distributed in the above sense.
 
-It is possible, then, that when two or more agents each intend that they do something, 
-	perhaps to paint a house,
-in having these intentions they conceive of the activity as a distributed activity.
-Where this happens, part of what the agents intend is, in effect, that they each construe their own actions as part of a single plan which also involves the others' plans.
+When two or more agents each intend that they do something, 
+they may conceive of the intended activity as a distributed activity.
+Where this happens, part of what the agents intend is, in effect, that they will each, at some stage of their planning, construe their own actions as part of a single plan which also involves the others' plans.
 
-We can exploit this possibility in strengthening Bratman's conditions for shared intention.
-To avoid counterexamples like the case of Yasmin and Zak,
-we leave the conditions for shared intention exactly as Bratman specifies them (see \vpageref{quote:bratman_account})
+We can exploit this possibility %of having intentions about distributed activities 
+to strengthen Bratman's conditions for shared intention.
+We leave the conditions for shared intention exactly as Bratman specifies them (see (1)--(3) on page \pageref{quote:bratman_account})
 but add the further requirement that the agents must, in intending that they J, each conceive of J as a distributed activity.
 
+Thus revised, the account provides conditions that are met in the case of Caitlin and Ciaran but not Beatrice and Baldric, as required.
 We have not shown, of course, that this modification to Bratman's account will enable it to avoid any other counterexamples.
@@
 Or perhaps it is impossible to give informative sufficient conditions for shared intention.
-Our aim here has only been to show that there is a way of avoiding one type of counterexample without abandoning core features of Bratman's planning approach to shared intention.
-If, as we suppose, that general approach succeeds in illuminating at least one kind of shared agency, it needs to be modified in order to capture a simple intuition about shared agency.
-The intuition is this:
-	our intending to act in accordance with our intentions and meshing subplans of them is not sufficient for us to have a shared intention;
-	instead we must conceive of our actions as part of a single plan.
-A planning approach to shared agency can readily accommodate this intuition by invoking distributed activities.
+Our aims here were only 
+		to show that Bratman's account faces counterexamples,
+	and 
+		to suggest a way of revising it without abandoning core features of Bratman's approach to shared intention.
 
@@
 \section{Conclusion}
-We started with some contrasts between shared agency and parallel individual agency.
+We started with some contrasts between shared agency and parallel but merely individual agency.
 These provide one intuitive fix on the notion of shared intention: 
@@
 
-Bratman's conditions are not sufficient for shared intention because they leave open the possibility that  when two or more agents meet his conditions, each fails to rely on the fact that there is a single outcome to which all of their actions are directed and so conceives of the others' actions merely as constraints and opportunities.
-Bratman's conditions may appear to be sufficient for shared intention because it may be natural that when two or more agents meet his conditions, they would rely on the fact that there is a single outcome to which all of their actions are directed and therefore be disposed to conceive of all their actions as parts of a single plan.
-Our counterexample, the case of Yasmin and Zak, shows that although this may be natural, it not necessary.
+Bratman's conditions are not sufficient for shared intention because they leave open the possibility that two or more agents could meet his conditions while each failing to make use of the fact that there is a single outcome to which all of their actions are directed.
+When this happens, each agent conceives of the others' actions merely as constraints and opportunities.
+Bratman's conditions may initially appear be sufficient for shared intention because it may be natural that when for agents meet his conditions to make use of the fact that there is a single outcome to which all of their actions are directed and  to conceive of all their actions as parts of a single plan.
+Our counterexample, the case of Beatrice and Baldric, shows that although this may be natural, it not necessary.
 
-There is a way to revise Bratman's account so as to avoid this sort of counterexample without departing from the spirit of his planning approach.
-The remedy we propose hinges on the possibility that sometimes an agent, in planning for an outcome, can specify both her own and other agent's actions.
+We identified a way to revise Bratman's account so as to avoid this sort of counterexample without departing substantially from his planning approach.
+The remedy we propose hinges on the possibility that an agent, in planning for an outcome, can sometimes plan both her own and other agent's actions.
 Further, in the right circumstances,
@@
 	can each individually specify roles for all of the agents in planning for that task.
-Even where such planning occurs in parallel (and so need not involve no shared agency), sometimes it can rationally result in coordinated action.
-These reflections suggest a way of appealing to the possibility that, in acting on a shared intention, each agent has a plan for all of their actions
-in order to capture the idea that agents acting on a shared intention conceive of all of their actions as part of a single plan.
+Even where such planning occurs in parallel (and so need not involve no shared agency),  it can sometimes rationally result in coordinated action.
+These reflections indicate that, in giving sufficient conditions for shared intention, we can appeal to the possibility that, in acting on a shared intention, each agent has a plan for all of their actions.
+This enables us to capture the idea that agents acting on a shared intention conceive of all of their actions as part of a single plan.
 %One way to capture this idea, 
@@
 
-No amount of acting on intentions about other's intentions is not sufficient for shared agency.  
-But conceiving of one's own and other's intentions and actions as part of a single plan might be.
+No amount of forming intentions about other's intentions and acting on knowledge of such intentions is sufficient, all by itself, for shared intention.  
+In many, perhaps all cases, shared agency requires changing perspective 
+to conceive of one's own and other's intentions and actions as part of a single plan.
 

### 2012-11-19 intention motor representation joint action (paper 3)/conf budapest philosophy 2012-11/shared agency and motor representation GOAL VERSION v2.tex
ceu talk in preparation

--- 2012-10-28 intention motor representation joint action (paper 3)/conf budapest philosophy 2012-11/shared agency and motor representation.tex
@@
 
-\section{***Discussion with Cordula Vesper}
-\begin{enumerate}
-\item Say more about what shared agency at the start: use the contrast cases instead of examples, and say that it involves some kind of fusing of agency.
-\item Note that in the jumping experiments, only the subject with the shorter jump changes what she does depending on the difference in jump-length between the two subjects.
+%
+%
+%\section{***Discussion with Cordula Vesper}
+%\begin{enumerate}
+%\item Say more about what shared agency at the start: use the contrast cases instead of examples, and say that it involves some kind of fusing of agency.
+%\item Note that in the jumping experiments, only the subject with the shorter jump changes what she does depending on the difference in jump-length between the two subjects.
 Vesper et al interpret this as showing that the subject with the shorter jump plans the whole activity.
@@
 And this makes sense: here they achieve coordination by first distributing roles (a strategy) and then having at least one subject plan both actions (so there is slightly less reliance on the two plans coinciding).
-Of course the appearance that they are distributing roles might just be an artefact of reciprocal planning where the upshot of the planning is that the person with the longer jump just does what she would normally do.
-
-\item {[done]}  Don't say `neither shared nor an intention' unless you explain this.
-\item {[done]} Can cite \citet{konvalinka:2010_follow} as example of what happens when two people each try to predict the other.
-\end{enumerate}
-
-\section{***Discussion with Hong Yu}
-`Once we move beyond the single agent there's no a priori reason to suppose that you have a single kind of agentive structure' (Hong Yu).
-This is slightly tricky for me because I don't think there's a single kind of agentive structure in individual action either.
-
-Dissimilarity between this paper and the paper on intention and motor representation: whereas it was maybe plausible to conjecture that all intentions bottom out in motor representation, it is not similarly plausible to claim that shared intentions bottom out in social motor representations (there are surely cases of shared agency where the sharing is entirely a matter of intention and does not involve motor representations at all).
-
-Hong Yu suggested that we need to include the argument about format in order to argue that motor representations have a role distinct from that of shared intention.
-
-In discussion: I need to shift this paper from saying:
-\begin{quote}
-There is a form of shared agency and that form cannot be characterised without reference to motor representation.
-\end{quote}
-to saying:
-\begin{quote}
-There are at least two forms of shared agency (at least two agentive structures) and some of these are such that characterising them involves appeal to motor representation.
-\end{quote}
-This shift reveals a new problem: we become more open to the objection that philosophy should concern forms of shared agency involving intention only.
-
-I need to stress the \emph{parity argument}.
-The argument is this:
-If 
-\begin{quote}
-
-	two or more agents' actions being appropriately related to a shared intention (as Bratman characterises it) is sufficient for shared agency, 
-\end{quote}
-then
-\begin{quote}
-	two or more agents' actions being appropriately related to a structure of social motor representation is also sufficient for shared 
-	agency.
-\end{quote}
+%Of course the appearance that they are distributing roles might just be an artefact of reciprocal planning where the upshot of the planning is that the person with the longer jump just does what she would normally do.
+%
+%\item {[done]}  Don't say `neither shared nor an intention' unless you explain this.
+%\item {[done]} Can cite \citet{konvalinka:2010_follow} as example of what happens when two people each try to predict the other.
+%\end{enumerate}
+%
+%
+%
+%\section{***Discussion with Hong Yu}
+%`Once we move beyond the single agent there's no a priori reason to suppose that you have a single kind of agentive structure' (Hong Yu).
+%This is slightly tricky for me because I don't think there's a single kind of agentive structure in individual action either.
+%
+%Dissimilarity between this paper and the paper on intention and motor representation: whereas it was maybe plausible to conjecture that all intentions bottom out in motor representation, it is not similarly plausible to claim that shared intentions bottom out in social motor representations (there are surely cases of shared agency where the sharing is entirely a matter of intention and does not involve motor representations at all).
+%
+%Hong Yu suggested that we need to include the argument about format in order to argue that motor representations have a role distinct from that of shared intention.
+%
+%In discussion: I need to shift this paper from saying:
+%\begin{quote}
+%There is a form of shared agency and that form cannot be characterised without reference to motor representation.
+%\end{quote}
+%to saying:
+%\begin{quote}
+%There are at least two forms of shared agency (at least two agentive structures) and some of these are such that characterising them involves appeal to motor representation.
+%\end{quote}
+%This shift reveals a new problem: we become more open to the objection that philosophy should concern forms of shared agency involving intention only.
+%
+%I need to stress the \emph{parity argument}.
+%The argument is this:
+%If 
+%\begin{quote}
+%
+%	two or more agents' actions being appropriately related to a shared intention (as Bratman characterises it) is sufficient for shared agency, 
+%\end{quote}
+%then
+%\begin{quote}
+%	two or more agents' actions being appropriately related to a structure of social motor representation is also sufficient for shared 
+%	agency.
+%\end{quote}
 
 \section{Introduction}
+Shared agency is a familiar feature of everyday life, but it's hard to get a good fix on it without assuming a big chunk of theory.
+I think the best available way to zoom in on shared agency is to think about contrast cases, that is pairs of cases which are as similar as possible except one involves shared agency and the other involves parallel but merely individual agency.
+
 When members of a flash mob in the Central Cafe respond to a pre-arranged cue by noisily opening their newspapers, they exercise shared agency. 
@@
 But two people who merely happened to be walking to the metro station side by side would not be exercising shared agency \citep{gilbert_walking_1990}.
-These sorts of case invite the question, 
-How do actions involving shared agency differ from merely parallel actions involving  only individual agency?
 
-The first example shows that the difference between shared agency and parallel individual agency can’t just be that the resulting actions have a common effect because merely parallel actions can have common effects too.
-And the second example shows that the difference can’t just be a matter of coordination, because people who are merely happen to be walking side by side each other also need to coordinate their actions in order to avoid colliding.
-Apparently, then, systematically distinguishing cases of shared agency from cases of parallel but merely individual agency  requires reflection on causes of actions.
+I'll use the term \emph{joint action} for an exercise of shared agency, as contrasted with an \emph{individual action} which is an exercise of individual agency.
 
-%
-%Shared agency is paradigmatically involved when two (or more) people 
-%	paint a house together \citep{Bratman:1992mi}, 
-%	lift a heavy sofa together \citep{Velleman:1997oo}, 
-%	prepare a hollandaise sauce together \citep{Searle:1990em}, 
-%	go to Chicago together \citep{Kutz:2000si}, 
-%	and walk together \citep{gilbert_walking_1990}.
-%In developmental psychology paradigm cases involving shared agency include  two people 
-%	tidying up the toys together \citep{Behne:2005qh},
-%	cooperatively pulling handles in sequence to make a dog-puppet sing \citep{Brownell:2006gu},
-%	bouncing a ball on a large trampoline together \citep{Tomasello:2007gl}.
+A basic question about joint action is,
+What is the relation between a joint action and the outcome or outcomes to which it is directed?
+Ayesha and Beatrice lift the table. ***
 
-Which representations %and processes 
-are needed to distinguish shared from individual agency?
-Philosophers have so far considered intentions and commitments but ignored the possibility that appeal to motor representations may also be needed. 
-%To considerations jointly motivate considering whether fully characterising shared agency may require appeal to motor representation.
-By contrast, concerning individual rather than shared agency, some have argued that motor representation may play a role in explaining agency \citep[e.g.][]{pacherie:2000_content,pacherie:2008_action,butterfill:2012_intention}.
-In addition, scientific research provides evidence that motor representations sometimes play a role in enabling agents to exercise shared agency, as we shall explain in detail below.
-So it is possible to ask whether appeal to motor representation  %might not only sometimes enable agents to exercise shared agency but also
-could also play a role in distinguishing shared from individual agency.
-In this paper we argue that it could.
-Shared agency is not always only a matter of our intentions or commitments: in some cases it also constitutively involves certain structures of motor representation.
-Or so we aim to show.
+The standard answer to this question involves shared intention.
+It's hard to say exactly what a shared intention is because each philosopher seems to take a different view.
+But for my purposes it's enough to say that shared intention, whatever exactly it is, stands to joint action as ordinary, individual intention stands to ordinary, individual action.
+Building on this parallel,
+we can say that when two or more agents act on a shared intention that they free they cat,
+\begin{enumerate}
+\item the shard intention involves a representation on the part of each agent of this outcome, the freeing of the cat
+\item the shared intention coordinates the several agents' actions
+\item and the shared intention coordinates their actions in such a way that, normally, the coordination would facilitate the occurrence of the represented outcome. 
+\end{enumerate}
+In this way we can explain the directedness of a joint action to an outcome by appeal to shared intention.
 
-Before going further,
-let us step back to ask 
-	why it might be useful 
-	to  distinguish systematically between shared and individual agency?
-Shared agency raises a tangle of scientific and philosophical questions.  Psychologically we want to know which mechanisms make it possible \citep{Sebanz:2006yq,vesper_minimal_2010}.  
-Developmentally we want to know when shared agency emerges, what it presupposes and whether it might somehow facilitate socio-cognitive, pragmatic or symbolic development \citep{Moll:2007gu,Hughes:2004zj,Brownell:2006gu}.
-Phenomenologically we want to characterise what (if anything) is special about experiences of action and agency when shared agency is involved \citep{Pacherie:2010fk}.  
-Metaphysically we want to know what kinds of entities and structures are implied by the existence of shared agency \citep{Gilbert:1992rs,Searle:1994lb}.  
-And normatively we want to know what kinds of commitments (if any) are entailed by shared agency and how these commitments arise \citep{Roth:2004ki}.
-%, plus a formal account of how practical reasoning for joint action differs (if at all) from individual practical reasoning \citep{Sugden:2000mw,Gold:2007zd}
-A philosophical account of shared agency 
-that made it possible to systematically distinguish shared from individual agency  
-may support investigation of these questions (as \citealp{Bratman:2009lv} suggests).  
-%Or, if there is more than one form of shared agency, the account should provide a principled way of distinguishing among forms of individual and shared agency.
-
-As already mentioned, our aim is to show that distinguishing individual from shared agency sometimes involves appeal to certain structures of motor representation.
-Our first step is to review standard approaches to shared agency.
-
-\section{Shared intention}
-On the leading, best developed accounts of shared agency in philosophy and psychology, 
-shared agency is explained  in terms of a special kind of intention or structure of intention, knowledge and commitment often called a shared intention.
-For two or more agents to exercise shared agency is for them to have a shared intention and for this shared intention to be appropriately related to their actions.
-Such actions comprise a \emph{shared intentional activity} (\citealp[p.\ 142]{Bratman:1999fr}).%
-%On this widely held view, the difference between our walking together and two strangers merely walking side-by-side is that one event is appropriately related to a shared intention whereas the other is not.
-
-Explaining shared agency in terms of shared intention pushes us back to the question, What is shared intention?
-Here is much disagreement.
-Some hold that it differs from
-ordinary intention with respect to the attitude involved (\citealp{Searle:1990em}). 
-Others have explored the notion that it differs from ordinary intention with respect to its subject, which is plural \citep{Gilbert:1992rs,helm_plural_2008}, 
-or that it differs from ordinary intention in the way it arises, namely through team reasoning \citep{Gold:2007zd}, 
-or that it involves distinctive obligations or commitments to others (\citealp{Gilbert:1992rs}; \citealp{Roth:2004ki}).
-Opposing all such views, \citet{Bratman:1992mi,Bratman:2009lv} argues that shared intention can be realised by multiple ordinary individual intentions and other attitudes whose contents interlock in a distinctive way.
-
-To avoid having to take sides on what shared intention is, 
-let us abstract the details
-and focus on features that everyone should agree on.
-Minimally, a shared intention stands to the actions comprising a shared intentional activity in roughly the way that an ordinary, individual intention stands to an ordinary, individual action.
-Now an ordinary, individual intention represents an outcome, coordinates an agent's actions, and coordinates the agent's actions in such a way that, normally, this coordination would facilitate the occurrence of the represented outcome.
-Similarly, 
-on almost any account,
-a shared intention involves there being a single outcome which each agent represents, where these representations coordinate the several agents' actions and coordinate them in such a way that, normally, the coordination would facilitate the occurrence of the represented outcome.
-
-There is another feature of shared intention that should be uncontroversial.  
+There is another feature of shared intention that should be uncontroversial on any account of it.  
 Shared intentions or their components feature in practical reasoning alongside ordinary, individual intentions,
@@
 
-So much for shared intention.
-Can we give a full account of shared agency by saying (roughly) that exercises of shared agency involve agents acting on a shared intention?
-Or are further ingredients necessary in systematically distinguishing shared from individual  agency?
-Our view is that further ingredients are necessary.
-In giving a full account of shared agency, it is also necessary to appeal to certain structures of motor representation. 
-Note, however, that our view is consistent with various weaker claims linking shared intention to shared agency.
-Our view is consistent with the claim that appeal shared intention is also needed in giving an account of shared agency,  
+In this talk I'm going to argue that in some cases of joint action,
+the directedness of the joint action to an outcome cannot be explained in terms of shared intention.
+In some cases, the directedness of a joint action to a goal needs to be explained in terms of a special structure of motor representation.
+I'll call this \emph{shared motor representation}.
+As we'll see,
+shared motor representation \textbf{resembles} shared intention in that if our actions are controlled by a shared motor representation, then:
+%
+\begin{enumerate}
+\item there is a single outcome which we each represent 
+\item and these representations coordinate our actions
+\item and coordinate them in such a way that, normally,  the coordination would facilitate the occurrence of the represented outcome. 
+\end{enumerate}
+%
+But shared motor representation \textbf{differs} from shared intention in not integrating in practical reasoning with ordinary individual intention.
+There are normative constraints linking ordinary intentions to shared intentions,
+but no  normative constraints link intention to shared motor representation.
+
+I should say, right at the start, that this speculative philosophy.
+The view inspired by some findings which provide some support for it, but the findings don't establish that the view is correct.
+
+This then is my thesis.
+To understand shared agency we need to understand the relation between joint actions and the  outcomes to which they are directed.
+And understanding this relation requires appeal, not only to structures of intention, commitment and knowledge, 
+but also to structures of motor representation.
+Shared agency is not always only a matter of our intentions or commitments: in some cases it also constitutively involves certain structures of motor representation.
+Or so I aim to show.
+
+This conflicts with the view that understanding shared agency requires understanding only shared intention.
+So I disagree with Facundo Alonso's statement that:
+%
+\begin{quote}
+`the key property of joint action lies in its internal component [...] in the participants' having a ``collective'' or ``shared'' intention' \citep[pp.\ 444-5]{alonso_shared_2009}.
+\end{quote}
+%
+But my disagreement is not terribly radical.
+I'm not denying that appeal shared intention is also needed in giving an account of shared agency,  
 %and with the claim that \emph{some forms} of shared agency can be fully characterised by appeal to shared intention alone.
-and with the claim that all exercises of shared agency involve shared intentions in some way.%
+and I'm not even disagreeing here with those who claim that all exercises of shared agency involve shared intentions in some way.
+The point is just that we can't fully understand shared agency in terms of intention, commitment and knowledge:
+instead, it's also necessary to invoke motor representations.
+
+Before I start on this,
+why should anyone care?
+Shared agency raises a tangle of scientific and philosophical questions.  
+%
+\begin{enumerate}
+\item Psychologically we want to know which mechanisms make it possible \citep{Sebanz:2006yq,vesper_minimal_2010}.  
+\item Developmentally we want to know when shared agency emerges, what it presupposes and whether it might somehow facilitate socio-cognitive, pragmatic or symbolic development \citep{Moll:2007gu,Hughes:2004zj,Brownell:2006gu}.
+\item Phenomenologically we want to characterise what (if anything) is special about experiences of action and agency when shared agency is involved \citep{Pacherie:2010fk}.  
+\item Metaphysically we want to know what kinds of entities and structures are implied by the existence of shared agency \citep{Gilbert:1992rs,Searle:1994lb}.  
+\item And normatively we want to know what kinds of commitments (if any) are entailed by shared agency and how these commitments arise \citep{Roth:2004ki}.
+%, plus a formal account of how practical reasoning for joint action differs (if at all) from individual practical reasoning \citep{Sugden:2000mw,Gold:2007zd}
+\end{enumerate}
+%
+Maybe identifying a constitutive role for motor representation in shared agency will, in a small way,
+ support investigation of these questions.  
+%Or, if there is more than one form of shared agency, the account should provide a principled way of distinguishing among forms of individual and shared agency.
+
+\section{The theory (non-motor)}
+I want to start by sketching the theory in a way that abstracts from motor representation,
+  so that you can see the general outlines.
+  This will involve introducing a series of ideas.  
+We'll come to the evidence only right at the end.
+
+\subsection{*distributive goal}
+Two or more agents' actions have a distributive goal just if ...
+
+Contrast cases: don't have distributive goal.
+
+But some actions involving parallel but merely individual agency do involve distributive goals.  
+One dark night two people each independently intend to paint a large bridge red.   
+More exactly, each intends that her painting grounds or partially grounds the bridge's being painted red.\footnote{
+Event $D$ \emph{partially grounds} event $E$ if there are events including $D$ which ground $E$.
+(So any event which grounds $E$ thereby also partially grounds $E$; 
+we nevertheless describe actions as `grounding or partially grounding' events for emphasis.)
+See the definition of \emph{plural grounding} *.
+}  
+(These intentions ensure that it is possible for both people to succeed in painting the bridge, as well as for either of them to succeed alone.)
+Because the bridge is large and they start from different ends, the two people have no idea of each other's involvement until they meet in the middle.
+Nor did they expect that anyone else would be involved in painting the bridge red.  
+On almost any account, this implies that they were not acting on a shared intention.
+Despite this, 
+they both succeed in painting the bridge red. 
+As this illustration suggests, 
+it is possible for two or more agents' actions to have a distributive goal without them thereby exercising shared agency.
+So distributive goals are not shared intentions.
+
+\subsection{*collective goal}
+Let an outcome, possible or actual, be a \emph{collective goal \label{df_collective_goal}} of a joint action, or of any collection of goal-directed actions, where three conditions are met: 
+	(a) this outcome is a distributive goal of the actions; 
+	(b) the actions are coordinated; and 
+	(c)  coordination of this type would normally  facilitate occurrences of outcomes of this type.  
+Examples of actions  that typically have collective goals include two people jointly sawing a log with a two-handled saw and  
+three people jointly lifting a heavy table.
+The bridge painters (from section *) are different: their actions do not have a collective goal because they are not coordinated.
+
+Where two or more agents' actions have a collective goal there is a sense in which, taken together, their actions are directed to the collective goal.  
+It is not just that each agent individually pursues the collective goal; in addition, there is coordination among their actions which plays a role in bringing about the collective goal.  
+We can put this in terms of the direction metaphor.  
+Any structure or mechanism providing this coordination is directing the agents' actions to the collective goal.  
+So the notion of a collective goal provides a schematic answer to the question about the relation between joint actions and the goals to which they are directed.
+
+The notion of a collective goal assumes that of coordination.  This should be understood in a broad sense.  
+When two agents between them lift a heavy block by means of each agent pulling on either end of a rope connected to the block via a system of pulleys, their pullings count as coordinated in this broad sense.
+In this case, the agents' actions are coordinated by a mechanism in their environment, the rope, and not necessarily by any psychological mechanism.  
+By invoking a broad notion of coordination 
+and invoking coordination of actions rather than of agents,
+the definition of collective goal avoids direct appeal to psychological states.
+
+%as well as non-psychological factors such as the dynamical properties of agents' bodies \citep[e.g.][]{schmidt_richardons:_2008}.
+To make a conjecture based on work with bees and ants, in some cases
+the coordination needed for a collective goal may even be supplied by 
+	 behavioural patterns \citep{seeley2010honeybee}  
+	 and 
+	 pheromonal signals \citep[pp.\ 178-83, 206-21]{hoelldobler2009superorganism}.
+
+This is not to say that collective goals never involve psychological states.
+%The notion of a collective goal is more abstract than shared intention and other notions that have been used in characterising joint action.
+%On many or all accounts, shared intentions function in part to coordinate actions in such a way as to facilitate realisation of the shared intention.
+In fact,
+one way for several actions to have a collective goal is for their agents to be acting on a shared intention; 
+a shared intention supplies the required coordination.
+
+But the possibility we are interested in is that the coordination required for a collective goal involves structures of motor representation rather than shared intention.
+How could this work in principle?
+
+*Aim: Describe reciprocal agent-neutral planning for outcomes whose obtaining would normally involve action on the part of each agent.
+
+\subsection{Agent-neutral}
+
+First we need to introduce the notion of an agent-neutral plan.
+By saying that something is \emph{agent-neutral} I mean just that it does not involve identifying any particular agents.
+
+%Having an agent-neutral plan makes perfect sense if you're stranded on a desert island because, after all, there it's obvious that you're going to do everything so there's no need to specify who will bring in the washing.
+
+Agent-neutral planning for outcomes whose realisation would normally involve action on the part of several agents is also quite common.
+For example, some housemates who have decided to take on an allotment to grow vegetables might sit down together to plan what needs doing without yet assigning roles to particular individuals. 
+In so planning, each housemate is thinking about what is to be done and not what she herself will do.  
+%(This is an idealisation, of course.)
+At some point the housemates stop planning.
+(This does not necessarily mean that they have a fully worked out plan; like any other plans, agent-neutral plans can have gaps that may need filling in later.)
+They now divide up the roles in a way that everyone is prepared to go along with,
+	and that each implements her part in the plan.
+
+In the above example of agent-neutral planning,
+	the housemates \textbf{plan together} and \textbf{agree} on a common plan.
+But an individual can construct an agent-neutral plan by herself, even if its eventual execution will involve others.
+	In fact, two or more individuals who are assigned a task might each individually engage in agent-neutral planning in parallel.
+
+Of course, for us to engage in reciprocal, parallel agent-neutral planning concerning a particular outcome only makes sense if the  plans we construct will be identical or similar enough that differences don't matter.
+And, of course, we must also all assign the same or similar roles to ourselves and each other.
+So reciprocal, parallel agent-neutral planning only makes sense where: 
+\begin{enumerate}
+\item 	we have relevantly similar planning strategies and expertise
+\item there is one salient way for us to achieve the outcome; and
+\item there is one salient way for us to assign roles to each other.
+\end{enumerate}
+
+	The task demands and their planning strategies may conspire to ensure that they each come up with the same agent-neutral plan.
+	The task demands and manifest properties of the agents, such as their distribution in space, may also ensure that each agent also assigns the same roles to the same individuals.
+(Strictly speaking it is not necessary for the plans and role assignments to be identical; it is enough if the resulting agent-specifying plans are, in a special sense, compatible.%
 \footnote{
-On the latter claim, see \citet[p.\ 5]{Gilbert:2006wr}: `I take a collective action to involve a collective intention.'  
-Related claims are made by 
-	\citet[p.\ 381]{Carpenter:2009wq}, 
-	\citet[p.\ 369]{Call:2009fk}, 
-	\citet{Kutz:2000si},  
-	\citet[p.\ 117]{rakoczy_pretend_2006},  and
-	\citet{Tollefsen:2005vh}.
+\label{fn:df_compatible}
+Suppose that, for some outcome, two or more agents each have a plan for the realisation of that outcome. 
+(These plans may, but need not, specify roles for all of the agents; but the plans must be agent-specific, not agent-neutral.)
+By saying that these plans are \emph{compatible} we mean that:
+(i) 
+no agent would normally be prevented from performing the role she is assigned in her own plan by other agents performing the roles they are assigned in their plans;
+and
+(ii)
+if all facts about which agents have which roles in which plans were common knowledge to the agents,
+this would not affect the rationality of their each acting on the intention that they realise the outcome by performing the role she is assigned in her own plan.
+%	each agent could rationally act on the intention that they realise the outcome by performing the role she is assigned in her own plan,
+%	and she could do so even if all facts about which agents are performing which roles in which plans were common knowledge to the agents.
+To illustrate, suppose that our task is to press a button simultaneously. 
+If your plan is that each of us will to start to move in exactly 60 seconds and press the button 5 seconds later whereas my plan specifies that you are the leader and we are each to press the button 5 seconds after you start moving, then our plans are compatible.
+})
+%
+Finally, each agent may know enough about herself and the others to be able to determine, without communicating, whether the plan and role assignments will be acceptable to everyone. 
+	And all of this---that they engage in parallel, agent-neutral planning resulting in identical (or compatible) plans and role assignments, which are acceptable to all---may be common knowledge to the agents.
+	So it is possible, in principle at least, 
+	that several agents might each individually engage in agent-neutral planning and rationally perform their part in the resulting plan, knowing that the others will do likewise.
+	Parallel, agent-neutral planning can rationally result in coordinated action without presupposing shared agency.%
+\footnote{
+Note that we are claiming only that shared agency is not presupposed.
+Our view is consistent with (but does not depend on) the claim
+that
+	if some agents each engage in parallel, agent-neutral planning and then rationally perform their part in the resulting plan,
+	the upshot would be an exercise of shared agency.
 }
 
-\section{An objection}
-At this point opponents sometimes attempt to dismiss our view out of hand along these lines.
-Philosophers are interested in shared \emph{intentional} agency. 
-If there are forms of shared agency other than this one, we might be right to claim that understanding these will require appeal to motor representation.
-But---so the objection---such forms of shared agency (if they exist at all) are not of interest to philosophers. 
-So we are clearly wrong to claim that the philosophical project of distinguishing shared from individual agency might require appeal to ingredients other than shared intentions.
+\subsection{*}
+Having each agent plan the whole joint action means that (i) each agent plans the other agent's action,
+(ii) each agent's plan for the other agent's action is approximately the same as that agent's plan for her own action [***AMBIGUOUS: `that agent's plan' must refer to the other agent, not `each agent',
+and 
+(iii) each  agent's plans for their own action are constrained by the plans for the other agent's action.
 
-One possible response to this line of opposition would be to weaken our claim. 
-We could become neutral on whether a full account of the sort of shared agency philosophers have focussed on can be given without appeal to motor representation,
-but insist that there is a form of shared agency about which parallel philosophical questions arise. 
-This conciliatory response would preserve most of the key claims in what follows, but a stronger response is available.
+\textbf{So what enables the two agents' plans to mesh is not that they represent each other's plans but more simply that they plan each other's actions as well as their own actions as if they were each about to do the whole thing themselves.}
 
-As mentioned above, some have argued that fully characterising individual agency requires appeal to motor representation as well as intention \citep{pacherie:2000_content,butterfill:2012_intention}. 
-These arguments concern intentional action.
-So if (as we suppose) these arguments are not known to be defective,
-then, assuming a parallel between individual and shared agency,  
-it should not be assumed at the outset that appealing to motor representation in a philosophical account of shared agency implies that the account does not apply to what opponents call `shared intentional agency'.
-We do not state our position in exactly the terms used by our opponent because of doubts about the qualifier `intentional' that are not specific to discussions of shared agency.\footnotemark \  \ 
-However, we do aim show in what follows that a philosophical account of the sort of shared agency that philosophers have been interested in will require appeal to structures of motor representation.
+Each agent is planning (and monitoring) both their actions almost as if a single agent were going to execute the whole action.
+And of course this is exactly what we want for small-scale joint action---we want two or more agents to act as one.
 
-\footnotetext{ 
-We avoid using the qualifier `intentional' for reasons not specific to questions about shared agency. 
-(We do use the term `shared intentional activity', but our use merely follows Bratman's stipulation that this term applies to an activity appropriately related to a shared intention.)
-Briefly, some philosophers offer reasons for holding that there could be agents whose actions are intentional although  the agents have no intentions at all.
-For instance, Bratman describes a creature who `acts on the basis of its beliefs and considered desires' only (not intentions) as doing things `intentionally' (\citeyear[p.\ 251]{bratman:2000_valuing}; see \citealp[pp.\ 137--8]{Bratman:1987xw} for further justification).
-These and related considerations reveal a coherent and natural way of using the term `intentional' on which it would apply to actions in virtue of their being appropriately related to motor representations of outcomes to which they are directed.
-It is not obvious to us whether using the term in this way is incorrect.
-}
+So what is the difference between the individual and the joint case?
+From the present point of view, the primary difference may be that in joint action there is a need to prevent execution of the parts of the action which are not one’s own.
+
+So far: shown theoretically how there could be directedness of joint action to outcome by appeal to reciprocal representations of outcomes which trigger agent-neutral planning.
+
+But unclear that any such representations exist, or at least that they plan any significant role in ordinary cases of joint action.
+What makes this theoretical possibility interesting is evidence that these representations occur in motor control of action.
+
+***section: say what motor processes are; examine what would be evidence for different points.
 
 \section{Shared motor representation}
-Our argument starts with an empirical premise about enabling conditions for shared agency: there are reciprocal, agent-neutral motor representations of outcomes whose obtaining would normally involve action on the part of each agent;
+Our argument starts with an empirical premise about enabling conditions for joint action: there are reciprocal, agent-neutral motor representations of outcomes whose obtaining would normally involve action on the part of each agent;
 moreover, these structures of motor representation sometimes facilitate coordination when agents exercise shared agency.
@@
 Unlike intentions and related mental states, motor representations are usually involved in planning actions over milliseconds rather than minutes or even years.
+
+\textbf{*** explain motor representation in terms of passing object from one hand to the other}
 

### 2012-11-19 intention motor representation joint action (paper 3)/conf budapest philosophy 2012-11/shared agency and motor representation.tex
ceu talk in preparation

@@
 These sorts of case invite the question, 
-How do actions involving shared agency differ from merely parallel actions involving  only individual agency?
+How do activities involving shared agency differ from activities involving parallel agency only?
+
+We should be motivated to answer this question because 
+shared agency raises a tangle of scientific and philosophical questions.  Psychologically we want to know which mechanisms make it possible \citep{Sebanz:2006yq,vesper_minimal_2010}.  
+Developmentally we want to know when shared agency emerges, what it presupposes and whether it might somehow facilitate socio-cognitive, pragmatic or symbolic development \citep{Moll:2007gu,Hughes:2004zj,Brownell:2006gu}.
+Phenomenologically we want to characterise what (if anything) is special about experiences of action and agency when shared agency is involved \citep{Pacherie:2010fk}.  
+Metaphysically we want to know what kinds of entities and structures are implied by the existence of shared agency \citep{Gilbert:1992rs,Searle:1994lb}.  
+And normatively we want to know what kinds of commitments (if any) are entailed by shared agency and how these commitments arise \citep{Roth:2004ki}.
+%, plus a formal account of how practical reasoning for joint action differs (if at all) from individual practical reasoning \citep{Sugden:2000mw,Gold:2007zd}
+A philosophical account of shared agency 
+that made it possible to systematically distinguish shared from individual agency  
+may support investigation of these questions (as \citealp{Bratman:2009lv} suggests).  
+%Or, if there is more than one form of shared agency, the account should provide a principled way of distinguishing among forms of individual and shared agency.
+
+So, to return to the contrast cases,
+how do activities involving shared agency differ from activities involving parallel agency only?
 
@@
 And the second example shows that the difference can’t just be a matter of coordination, because people who are merely happen to be walking side by side each other also need to coordinate their actions in order to avoid colliding.
-Apparently, then, systematically distinguishing cases of shared agency from cases of parallel but merely individual agency  requires reflection on causes of actions.
+Apparently, then, systematically distinguishing cases of shared agency from cases of parallel but merely individual agency  requires reflection on agents' mental states.
 
@@
 
-Which representations %and processes 
-are needed to distinguish shared from individual agency?
-Philosophers have so far considered intentions and commitments but ignored the possibility that appeal to motor representations may also be needed. 
+Which mental states might distinguish shared from individual agency?
+
+***Philosophers have so far considered intentions and commitments but ignored the possibility that appeal to motor representations may also be needed. 
 %To considerations jointly motivate considering whether fully characterising shared agency may require appeal to motor representation.

### 2012-11-19 intention motor representation goal ascription with corrado sinigaglia (paper 2)/Motor Representation in Goal Ascription.tex
working

@@
 
+\section{***NEW PLAN***}
+\begin{enumerate}
+\item There is experience revelatory of action
+	\begin{enumerate}
+	\item there is a phenomenal element which is common to : (i) performing, (ii) imagining and (iii) observing action
+	\item this common phenomenal element enables us to know (i) which action we are performing, (ii) which action we are imagining, and (iii) which action we are observing.
+It is an experience revelatory of action.
+	\item the common phenomenal element depends on motor representation of an outcome to which the action is directed.  (This what links the motor representation to the experience).
+	\end{enumerate}
+
+\item What is this experience revelatory of action, and how does it explain the existence of content-respecting influences of motor representation on judgements about the goals of actions?
+	\begin{enumerate}
+	\item Indirect hypothesis: the experience revelatory of action is an experience of the movements involve in acting, or of sensory consequence of action
+
+	\item Direct hypothesis: the experience revelatory of action is an experience of goal-directed action as such
+	\end{enumerate}
+
+\end{enumerate}
+
 \section{Experience of Action}
+\label{sec:processes}
+Our aim in this and the following sections is to explore how motor representations could sometimes have content-respecting influences on judgements about the goals of observed actions.
+Broadly, our proposal is that the influence of motor representation on observational judgements goes via experience.
+It is because motor representations can affect our experiences when we observe actions that those motor representations can facilitate judgements about the actions.
+
+Our argument will involve three steps.
+The first step is to argue that in performing, imagining and observing actions alike there are \emph{experiences revelatory of action}, that is, experiences which provide the subject of experience with reasons for correct judgements about which action someone is performing.
+(Note that we initially leave open the question of what experiences revelatory of action are experiences of.)  The second step is to show that these experiences depend on motor representations.
+And the third step is to consider what experiences revelatory of action could be experiences of, which is necessary in order to describe how all of this can be used to explain the existence of content-respecting influences of motor representation on judgements about the goals of actions.
+
+When acting it is often possible to know which action you are performing---to know, for example, whether you are grasping a particular mug or stroking the cat (or perhaps both).
+How could you know this?
+In some cases it might depend on knowing your intentions.
+However, there are cases in which people act on an intention but do something not compatible with fulfilling the intention (*ref).
+Suppose you are facing a mug and holding a wine glass for a friend and intend to grasp the mug but, perhaps due to being distracted or a manipulation of your attention, start to drink from the wine glass.
+In such a case it is possible that you will realise what you are doing, stop it and attempt once more to grasp the mug.
+So you know what you are doing even though, in this sort of case, such knowledge does not plausibly depend on knowing what you intend.
+A natural thought, then, is that in performing an action such as grasping, it is sometimes possible to know what you are doing on the basis of experience.
+That is to say, in acting there are (at least sometimes) experiences revelatory of action.
+This much should be almost uncontroversial.
+
+Against this conclusion, someone may insist that knowledge of what you are doing depends on knowledge of your intentions even when distracted.
+On this view, someone who intends to do one thing but absent mindedly starts to drink from a wine glass she is holding for a friend really intends to drink and knows that she intends this.
+But now consider the special case of a subject with anarchic hand syndrome who attempts to picks up a cup of tea with one hand while using the other hand to put the tea back onto the table.%
+\footnote{
+\citet[p.\ 1114]{della_sala:1991_anarchic}.
+Subjects with anarchic hand syndrome, which usually follows lesions of the anterior part of the corpus callosum and of the supplementary motor area, perform actions incompatible with their avowed intentions.
+These patients may also refer the their anarchic hand as having a mind of its own.
+he anarchic hand syndrome \citep{della_sala:1994fk}.
+} 
+%
+This subject knows which actions she is performing with which hands: this knowledge is guiding her actions insofar as she uses one hand to control the other.
+But it seems unlikely that this knowledge depends on knowing what she intends.
+For she does not intend one of the actions she is performing, as she herself claims.
+Or else, contrary to her own reports, she knowingly has an irrational combination of intentions; but such combinations of intentions do reliably generate detailed predictions about actions and so knowledge of these intentions could not explain the subject's knowledge of her own actions.
+Either way, then, experience is needed to explain the possibility of knowing which actions one is performing.
+This strengthens the conclusion that, at least sometimes, performing actions involves experiences revelatory of actions.
+
+\section{Experience of Action [*OLD]}
 \label{sec:processes}
@@
 
-\subsection{The Direct Hypothesis: Experiences of Action}
+\section{The Direct Hypothesis: Experiences of Action}
 Our aim is to understand how motor representations of outcomes can have a content-respecting influence on judgements about the goals of actions. 
-According to the Direct Hypothesis, a motor representation of an outcome can underpin an experience of an action as directed to a matching outcome, and this experience can in turn ground a judgement about the action's being directed to a matching outcome.
+According to the Direct Hypothesis, a motor representation of an outcome can underpin an experience of an action as directed to a matching outcome, and this experience can in turn ground a judgement about the goal of the action.
 This hypothesis hinges on two potentially controversial claims.
@@
 
-Is the first claim plausible?  Consider the following argument:
+Is the first claim plausible?
+Compare actually grasping a mug with merely imagining grasping it.
+There is a way of imagining grasping the mug which resembles actually grasping it with respect to timing \citep{decety:1989_timing, decety:1996_imagined, Jeannerod:1994oz} and the sorts of factors that make acting more or less awkward \citep{parsons:1994_temporal, frak:2001_orientation}.
+Further, imagining grasping the mug in advance would bring some of the benefits of actually rehearsing grasping it (*ref).
+The experience involved in imagining grasping the mug in this way has something in common with the experience involved in actually grasping it.
+But what is common to the two experiences?
+
+****
+
+Is the first claim plausible?
+First note that there is a phenomenologically action-like way of imagining, one which involves something almost like acting except that such imaginings are not necessarily responsive to the features of actual objects and do not necessarily result in bodily movements.
+Suppose that you are about to perform a delicate operation with some scissors.
+You stand poised to act, but first pause to mentally pantomime performing the action.
+In some respects the experiences involved in this imaginative exercise may be barely distinguishable from experiences that might be involved in actually performing the operation (\citealp[p.\ 161]{currie:1997_mental}; \citealp[p.\ 727]{jeannerod:1995_mental}; \citealp[p.\ 638-9]{kosslyn:2001_neural}).
+In fact, imagining performing the operation in advance has some of the benefits of actually rehearsing it (*ref).
+For imagining performing an action in this phenomenologically action-like way sometimes resembles actually performing the action with respect to timing \citep{decety:1989_timing, decety:1996_imagined, Jeannerod:1994oz} and the sorts of factors that make acting more or less awkward \citep{parsons:1994_temporal, frak:2001_orientation}.
+
+Given that an experience of actually performing an action can have elements in common with the experience involved in merely imagining performing that action, what could these common elements be ?
+
+Consider the following argument:
 %
@@
 \item
-There is a way of imagining acting which involves experiencing goal-directed actions as such.
+In actually performing a goal-directed action it is sometimes possible experience the action as a goal-directed action.
 
 \item
-There are sometimes elements common to what you experience when you imagine performing a goal-directed action and what you experience when you observe someone else performing that action; in particular, the experience of action is common to both.
+There is a way of imagining acting which involves something almost like acting, except that such imaginings are not necessarily responsive to the features of actual objects and do not necessarily result in bodily movements.
+
+Therefore: sometimes, when you imagine performing an action you may experience a goal-directed action as such.
+
+\item
+What you experience when you imagine performing an action has elements in common with what you experience when you observe someone else performing that action; in particular, the experience of goal-directed action is common to both.
 \end{enumerate}
@@
 
-On the first premise, (1), let us start by observing that there are two ways of imagining acting.
+On the first premise, (1), we first need to distinguish two ways of imagining acting.
 One, cognitive way of imagining involves something like thinking about performing an action.
@@
 
-What do we experience when we imagine acting?  ***HERE
+Step 1: phenomenologically action-like imagination involves motor representation.
+We know this because how actions unfold in this way of imagining reflects motor constraints linked to the goal of the action (so not only constraints arising from, for example, finger configuration) and timings.
+For instance ...
+
+Step 2: What do we experience when we imagine acting?  ***HERE
 a.  There are no movements or sensory effects to experience (but there is  no action produced either).
@@
 c.
-I could be mistaken about the movements involved in an action and about its sensory effects but still know that I am imagining it (and not just because I intend to imagine it and then do some imaginative thing or other---if my mind wanders so that I start imaginig a different action, I could know this).
+I could be mistaken about the movements involved in an action and about its sensory effects but still know that I am imagining it (and not just because I intend to imagine it and then do some imaginative thing or other---if my mind wanders so that I start imagining a different action, I could know this).
 (one relevant consideration is that you know not only which action you are imagining but when you have succeeded in imagining performing it); ()
@@
 Such subjects may continue to have some experiences characteristic of action; for instance, when asked to flex an index finger they can report having done so even in the absence of visual information \citep{kristeva:2006}.
-These subjects' sensory deficits make it unlikely that their experiences could be of movements or the like: it is more plausible that, concerning acting, their experiences are like those of subjects with anosognosia for hemiplegia.
+These subjects' sensory deficits make it unlikely that their experiences could be of movements or the like (*no: motor-perception associations; but what do they report?): it is more plausible that, concerning acting, their experiences are like those of subjects with anosognosia for hemiplegia.
 None experience movement: what they experience is action.

### 2012-11-19 intention motor representation joint action (paper 3)/conf budapest philosophy 2012-11/motor representation and shared intention HANDOUT.tex
ceu talk in preparation
handout done

@@
 
-\documentclass[11pt]{extarticle}
+\documentclass[12pt]{extarticle}
 % extarticle is like article but can handle 8pt, 9pt, 10pt, 11pt, 12pt, 14pt, 17pt, and 20pt text
@@
 {\Large
-\textbf{Shared Agency and Motor Representation}
+\textbf{Shared Agency \\and Motor Representation}
 }
 
-Stephen A. Butterfill \& Corrado Sinigaglia
-
 %<s.butterfill@warwick.ac.uk>
-<butterfillS@ceu.hu>
+butterfillS@ceu.hu / s.butterfill@warwick.ac.uk
 
 \end{center}
-%
-%
-%\textbf{Abstract}
-%On the assumption that social motor representation plays a role in explaining how effective joint action is possible, do we also need motor representation to explain what joint action is?
-Philosophers tend to assume that motor representation is only an enabling condition for joint action and of no direct interest to narrowly philosophical theories of joint action and shared intention.
-In this talk I shall argue that social motor representation and shared intention  have distinctive roles in explaining the purposiveness of joint action.
-This gives rise to a challenge.
-On the one hand, effective joint action---imagine two people erecting a tent in a gale together---sometimes requires both shared intentions and social motor representations plus a certain kind of harmony between the two.
-On the other hand, recognizing their distinctive roles precludes the existence of direct inferential links between shared intentions and social motor representations.
-The challenge is to explain how these two kinds of representation could sometimes harmoniously contribute to effective joint action despite the lack of inferential integration.
-%
 
-\section{Introduction}
-A representation (motor or not) is \emph{agent-neutral} if its content does not specify any particular agent or agents.
+\section{Question}
+\emph{Shared Agency}.
+Sisters exercise shared agency when they cycle to school together; in contrast,  strangers who happen to be cycling the same route are exercising parallel but merely individual agency.\citep{gilbert_walking_1990}
 
-%***CHECK DEFINITIONS***
+When members of a flash mob in the Central Cafe respond to a pre-arranged cue by noisily opening their newspapers, they exercise shared agency. 
+But when others happen to noisily open their newspapers in response to the same cue, they do not.\citep{Searle:1990em}
 
-Two or more motor representations are \emph{reciprocal} just if there is a single outcome which each motor representation represents.
+A \emph{joint action} is an exercise of shared agency (in contrast to an \emph{individual action}).
 
-\textbf{Premise} 
-Reciprocal agent-neutral motor representation enables joint action
+What is the relation between a purposive joint action and the goal or goals to which it is directed?
 
-\begin{quote}
-`the social relation between individuals modulates action simulation ...\ motor activation during action anticipation depends on the ...\ relation between the actor and the observer ...\ Simulation of another person’s action, as reﬂected in the activation of motor cortices, gets stronger the more the other is perceived as an interaction partner.’  \citep{kourtis:2010_favoritism}
-\end{quote}
-
-\textbf{Question} Does reciprocal agent-neutral motor representation also play a role in explaining what joint action is?
-
-\textbf{Challenge}  How could social motor representation and shared intention harmoniously contribute to joint action?
-
-\section{The possibility of purposive joint action}
-What is the relation between a purposive joint action and the outcome or outcomes to which it is directed?
-
-Reciprocal agent-neutral motor representations can
-(1) involve a representation, on the part of each agent, of an outcome; 
-(2) coordinate the several agents’ activities;
-and 
-(3) coordinate the several agents’ activities in such a that would normally facilitate the occurrence of the represented outcome.
+\section{The Standard View}
+‘the key property of joint action lies in its internal component [...] in the participants’ having a “collective” or “shared” intention.’ \citep{alonso_shared_2009} %[pp.\ 444-5]
 
@@
 
-\section{An objection}
-‘the key property of joint action lies in its internal component [...] in the participants’ having a “collective” or “shared” intention.’ \citep[pp.\ 444-5]{alonso_shared_2009}
+`each agent does not just intend that the group perform the […] joint action.
+Rather, each agent intends as well that the group perform this joint action in accordance with subplans (of the intentions in favor of the joint action) that mesh'\citep%[p.\ 332]
+{Bratman:1992mi}
 
-But could some reciprocal agent-neutral motor representations be shared intentions?  No ...
-\begin{enumerate}
-\item Only representations with a common format can be inferentially integrated.
+\section{Thesis}
+In some cases 
+	it is not a shared intention 
+	but a special structure of motor representation,
+	a `shared motor representation',
+	in virtue of which a joint action is related to its goal.
 
-\item Any two intentions can be inferentially integrated in practical reasoning.
+\section{Shared Motor Representation}
 
-\item My intention that I visit Paris on Friday is a propositional attitude.
+A \emph{goal} is an outcome to which actions are, or might be, directed.  (Contrast a \emph{goal-state}, an intention or other state of an agent linking an action to a goal to which it is directed.)
 
-\item All intentions are propositional attitudes.
+An outcome is a \emph{distributive goal} of two or more actions  just if 
+(a) each action is individually directed to this outcome; and 
+(b) it is possible that: all actions succeed relative to this outcome.
 
-\item No motor representations are propositional attitudes.
+An outcome is a \emph{collective goal} of two or more actions just if
+(a) this outcome is a distributive goal of the actions;
+(b) the actions are coordinated; and 
+(c) coordination of this type would normally  facilitate occurrences of outcomes of this type
 
-\item No motor representations are intentions.
+A representation or plan is \emph{agent-neutral} if its content does not specify any particular agent or agents; a planning process is agent-neutral if it involves only agent-neutral representations.
+
+Events $D_1$, ...\ $D_n$ \emph{ground} $E$, if: $D_1$, ...\ $D_n$ and $E$ occur; 
+$D_1$, ...\ $D_n$ are each (perhaps improper) parts of $E$; and 
+every event that is a proper part of $E$ but does not overlap  $D_1$, ...\ $D_n$ is caused by some or all of $D_1$, ...\ $D_n$.
+
+For an individual to be \emph{among the agents of an event} is for there to be actions $a_1$, ...\ $a_n$ which ground this event where the individual is an agent of some (one or more) of these actions.
+
+We have a \emph{shared motor representation} of an outcome just if 
+\begin{enumerate}[label=\emph{\alph*}),itemsep=0pt,topsep=0pt]
+\item we each have a motor representation of this outcome; 
+\item we are each disposed to inhibit some but not all of the planning or actions resulting from (a);
+\item  we each expect that if the outcome occurs, we will all be among the agents of its occurrence; and
+\item the truth of (a) and (b) depends on the truth of (c).
 \end{enumerate}
+
+%Two or more representations (motor or not) are \emph{reciprocal} just if there is a single outcome which each represents.
+%
+
+\section{Evidence that Shared Motor %Rep\textsuperscript{\underline{n}} 
+Representation Exists}
+
+In joint action, motor planning can occur for another's actions,\citep{kourtis:2012_predictive} and can inform planning for one's own actions.\citep{vesper:2012_jumping}  %*submitted imagining acting paper
+
+In joint action, it is sometimes necessary to inhibit planning or performing another's action.\citep{sebanz:2006_twin_peaks} 
+Whether this is necessary depends on one's beliefs about co-actors' agency.\citep{tsai:2008_action}
+
+In some joint actions, the agents have a single representation of the whole action (not only separate representations of each agent's part).\citep{tsai:2011_groop_effect}
+
+% Vesper says forthcoming EEG paper using piano playing paradigm on agent-neutral identification of error: one brain wave signals whether there is an error, and a different brain wave signals whose error it is (also tells you whether the overall harmonics are affected)
 
@@
 
-Some joint actions involve both shared  intention and reciprocal agent-neutral motor representation.
+A shared motor representation is in \emph{harmony} with a shared intention if they concern matching outcomes.
 
-How are non-accidental matches between the outcomes specified by shared intentions and by reciprocal agent-neutral motor representations possible?
+Some joint actions involve both shared intention and shared motor representation.
 
-`motor imagery could play a crucial role in bridging the gap'\citep{pacherie:2000_content}
+How is non-accidental harmony between shared intentions and shared motor representations?
+
+Proposal: `motor imagery could play a crucial role in bridging the gap'\citep{pacherie:2000_content}
 

### 2012-11-19 intention motor representation joint action (paper 3)/conf budapest philosophy 2012-11/shared agency and motor representation GOAL VERSION v2.tex
working

@@
 Having each agent plan the whole joint action means that (i) each agent plans the other agent's action,
-(ii) each agent's plan for the other agent's action is approximately the same as that agent's plan for her own action [***AMBIGUOUS: `that agent's plan' must refer to the other agent, not `each agent',
+(ii) each agent's plan for the other agent's action is approximately the same as that agent's plan for her own action [***AMBIGUOUS: `that agent's plan' must refer to the other agent, not `each agent'],
 and 
@@
 
-\section{Interpersonal coordination and motor representation}
+\section{The Evidence: Interpersonal coordination and motor representation}
 
@@
 
-\section{Conclusion}
-We have been considering how to provide an account of shared agency that might contribute to investigating a tangle philosophical and scientific questions.
+\section{*Interface Problem}
+
+[*Include this?]
+
+I think we can see a possible solution to this challenge by considering another problem.
+In constructing realisers of shared intention from ordinary individual intention, we need intentions \emph{that we J}.
+As has been much discussed, the contents of these intentions cannot all refer to actions involving shared intentions.
+For this reason Michael Bratman suggests that things we intend are  cooperatively neutral activities.
+It is then necessary to add further intentions in order to transform cooperatively neutral activities into joint actions.
+But it also seems possible that in some cases, what we intend when we intend that we J is not a cooperatively neutral activity but instead a joint action of the sort which involves social motor representation.
+
+So perhaps harmony between shared intention and social motor representation is sometimes achieved in this way: what we intend when we share an intention is the sort of joint action that involves social motor representation.
+
+\section{*Conclusion version 1}
+In conclusion,
+I have suggested that 
+shared motor representations
+can 
+ground purposive joint action
+for they
+can coordinate two or more agents' actions in virtue of representing a single outcome to which each agent's actions are directed.
+In this respect shared motor representation resembles shared intention.
+
+But there are also differences.
+Shared motor representation
+and
+shared intention  
+have \textbf{distinctive roles in explaining shared agency}.
+And, relatedly, shared motor representation is \textbf{unlike shared intention} in that its constituent representations are non-propositional in format and so it doesn't inferentially integrate with ordinary intentions and knowledge states.
+
+This has two consequences.
+One is that we cannot fully understand shared agency is if we focus only on shared intention.
+The other is that we face a challenge.
+The challenge is to explain how these two kinds of representation could sometimes harmoniously contribute to effective shared agency despite not being inferentially integrated in practical reasoning.
+
+\section{*Conclusion version 2}
+My question was,
+What's the relation between a joint action and the goal or goals to which it is directed?
+
+I have been arguing that we may need to appeal to shared intention to answer this question,
+but we can't answer it by appeal only to shared intention.
+This is because some joint actions are related to goals by special structures of motor representation which I've been calling shared motor representations.
+
+Shared motor representation is \textbf{like shared intention} in that: (i) represent outcome; (ii) coordinate action; (iii) coordinate action in such a way that ...
+
+But  shared motor representation is \textbf{unlike shared intention} in that its constituent representations are non-propositional in format and so it doesn't inferentially integrate with ordinary intentions and knowledge states.
+
+This led us to the interface problem ...
+
+So to understand the relation between joint actions and their goals, we need to appeal, 	
+	not only to structures of intention and knowledge,
+	but also to structures of motor representation.
+
+This may matter for understanding the role of joint action in evolution and development;
+it may have consequences  for the metaphysics of joint action, and for claims about norms governing shared agency.
+
+If it's true that some joint actions involve shard motor representation,
+
+ phenomenology
+
+Having each agent plan the whole joint action means that (i) each agent plans the other agent's action,
+(ii) each agent's plan for the other agent's action is approximately the same as that agent's plan for her own action [***AMBIGUOUS: `that agent's plan' must refer to the other agent, not `each agent'],
+and 
+(iii) each  agent's plans for their own action are constrained by the plans for the other agent's action.
+
+\textbf{So what enables the two agents' plans to mesh is not that they represent each other's plans but more simply that they plan each other's actions as well as their own actions as if they were each about to do the whole thing themselves.}
 

### 2012-11-20 intention motor representation joint action (paper 3)/conf budapest philosophy 2012-11/shared agency and motor representation GOAL VERSION v2.tex
just before send

--- 2012-11-19 intention motor representation joint action (paper 3)/conf budapest philosophy 2012-11/shared agency and motor representation GOAL VERSION v2.tex
@@
 \section{Introduction}
-Shared agency is a familiar feature of everyday life, but it's hard to get a good fix on it without assuming a big chunk of theory.
-I think the best available way to zoom in on shared agency is to think about contrast cases, that is pairs of cases which are as similar as possible except one involves shared agency and the other involves parallel but merely individual agency.
+Shared agency is a familiar feature of everyday life.
+One way to get an intuitive fix on it is to think about contrast cases, that is pairs of cases which are as similar as possible except one involves shared agency and the other involves parallel but merely individual agency.
 
@@
 
+Shared agency raises \textbf{a tangle of scientific and philosophical questions}.  
+%
+\begin{enumerate}
+\item Psychologically we want to know which mechanisms make it possible \citep{Sebanz:2006yq,vesper_minimal_2010}.  
+\item Developmentally we want to know when shared agency emerges, what it presupposes and whether it might somehow facilitate socio-cognitive, pragmatic or symbolic development \citep{Moll:2007gu,Hughes:2004zj,Brownell:2006gu}.
+\item Phenomenologically we want to characterise what (if anything) is special about experiences of action and agency when shared agency is involved \citep{Pacherie:2010fk}.  
+\item Metaphysically we want to know what kinds of entities and structures are implied by the existence of shared agency \citep{Gilbert:1992rs,Searle:1994lb}.  
+\item And normatively we want to know what kinds of commitments (if any) are entailed by shared agency and how these commitments arise \citep{Roth:2004ki}.
+%, plus a formal account of how practical reasoning for joint action differs (if at all) from individual practical reasoning \citep{Sugden:2000mw,Gold:2007zd}
+\end{enumerate}
+%
+In investigating these questions it is sometimes useful to have a conceptual framework which enables us to distinguish systematically between shared and individual agency.
+My aim in what follows is to see whether there is a constitutive role for motor representation in explaining shared agency.
+There is now an impressive and growing body of evidence about the role of motor representation in \emph{enabling} joint action.
+But it is perhaps natural to suppose that motor representation is merely an enabling factor, so that ...
+
+\textbf{Shared agency is not always only a matter of our intentions or commitments: in some cases it also constitutively involves certain structures of motor representation.}
+Or so I aim to show.
+
+\subsection{*slide}
 A basic question about joint action is,
-What is the relation between a joint action and the outcome or outcomes to which it is directed?
+\textbf{What is the relation between a joint action and the outcome or outcomes to which it is directed?}
 Ayesha and Beatrice lift the table. ***
@@
 In some cases, the directedness of a joint action to a goal needs to be explained in terms of a special structure of motor representation.
-I'll call this \emph{shared motor representation}.
-As we'll see,
-shared motor representation \textbf{resembles} shared intention in that if our actions are controlled by a shared motor representation, then:
-%
-\begin{enumerate}
-\item there is a single outcome which we each represent 
-\item and these representations coordinate our actions
-\item and coordinate them in such a way that, normally,  the coordination would facilitate the occurrence of the represented outcome. 
-\end{enumerate}
-%
-But shared motor representation \textbf{differs} from shared intention in not integrating in practical reasoning with ordinary individual intention.
-There are normative constraints linking ordinary intentions to shared intentions,
-but no  normative constraints link intention to shared motor representation.
 
-I should say, right at the start, that this speculative philosophy.
-The view inspired by some findings which provide some support for it, but the findings don't establish that the view is correct.
-
-This then is my thesis.
+It would be natural to assume that
+motor representation plays at most an enabling role in explaining shared agency,
+that we need motor representation to explain how people manage to coordinate their actions in practice,
+but not to understand what shared agency is.
+But my thesis is that we need to appeal to action-guiding representations that are more primitive than intention and desire in order to explain what shared agency is:
 To understand shared agency we need to understand the relation between joint actions and the  outcomes to which they are directed.
@@
 but also to structures of motor representation.
-Shared agency is not always only a matter of our intentions or commitments: in some cases it also constitutively involves certain structures of motor representation.
-Or so I aim to show.
 
@@
 %
-But my disagreement is not terribly radical.
+But \textbf{my disagreement is constructive}. 
+I want to build on insights about shared intention.
 I'm not denying that appeal shared intention is also needed in giving an account of shared agency,  
@@
 
-Before I start on this,
-why should anyone care?
-Shared agency raises a tangle of scientific and philosophical questions.  
-%
-\begin{enumerate}
-\item Psychologically we want to know which mechanisms make it possible \citep{Sebanz:2006yq,vesper_minimal_2010}.  
-\item Developmentally we want to know when shared agency emerges, what it presupposes and whether it might somehow facilitate socio-cognitive, pragmatic or symbolic development \citep{Moll:2007gu,Hughes:2004zj,Brownell:2006gu}.
-\item Phenomenologically we want to characterise what (if anything) is special about experiences of action and agency when shared agency is involved \citep{Pacherie:2010fk}.  
-\item Metaphysically we want to know what kinds of entities and structures are implied by the existence of shared agency \citep{Gilbert:1992rs,Searle:1994lb}.  
-\item And normatively we want to know what kinds of commitments (if any) are entailed by shared agency and how these commitments arise \citep{Roth:2004ki}.
-%, plus a formal account of how practical reasoning for joint action differs (if at all) from individual practical reasoning \citep{Sugden:2000mw,Gold:2007zd}
-\end{enumerate}
-%
-Maybe identifying a constitutive role for motor representation in shared agency will, in a small way,
- support investigation of these questions.  
-%Or, if there is more than one form of shared agency, the account should provide a principled way of distinguishing among forms of individual and shared agency.
+\section{The theory (non-motor)}
+In what follows, I'm simply going to take for granted a range of empirical claims about motor representation and its role in facilitating joint action.
+There is a growing body of evidence in support of these claims, some of which is listed on your handout.
+But I'm not going to discuss the evidence.
+Instead I want us to do speculative philosophy,
+to try to tease out the broad implications of the evidence for understanding the structures of agency involved in joint action.
 
-\section{The theory (non-motor)}
-I want to start by sketching the theory in a way that abstracts from motor representation,
-  so that you can see the general outlines.
-  This will involve introducing a series of ideas.  
-We'll come to the evidence only right at the end.
+To this end it will be helpful to fix some terminology.
 
@@
 Any structure or mechanism providing this coordination is directing the agents' actions to the collective goal.  
-So the notion of a collective goal provides a schematic answer to the question about the relation between joint actions and the goals to which they are directed.
+So the notion of a collective goal provides \textbf{a schematic answer} to the question about the relation between joint actions and the goals to which they are directed.
 
@@
 
-*Aim: Describe reciprocal agent-neutral planning for outcomes whose obtaining would normally involve action on the part of each agent.
-
 \subsection{Agent-neutral}
@@
 Agent-neutral planning for outcomes whose realisation would normally involve action on the part of several agents is also quite common.
+[*Change example so that can vary parameters: we have decided to make a pizza ...]
 For example, some housemates who have decided to take on an allotment to grow vegetables might sit down together to plan what needs doing without yet assigning roles to particular individuals. 
@@
 	In fact, two or more individuals who are assigned a task might each individually engage in agent-neutral planning in parallel.
+
+Suppose that the housemates each sit alone in their rooms and plan the trip to Balaton.
 
@@
 
-\subsection{*}
-Having each agent plan the whole joint action means that (i) each agent plans the other agent's action,
+\subsection{*Motor}
+To see how there might be agent-neutral motor planning for outcomes realising which will involve the agent and another's or others' actions, 
+let us return to a case involving a single agent.
+
+Consider  an individual action.
+Suppose an agent moves a mug from one place to another, passing it from her left hand to her right hand half way.
+It is a familiar idea that motor planning, like planning generally, involves starting with relatively abstract representations of outcomes and gradually filling in details.
+We can capture this by supposing that 
+motor representations for planning and monitoring action involve a hierarchical structure of representations.
+At the top we might find a relatively abstract representation of an outcome, in this case of the movement of the object from one location to another.
+Action-relevant details are progressively filled in by representations at lower stages of the hierarchy. 
+Now in the action we are considering there is a need, even for the single agent, to coordinate the exchange between her two hands.
+How is this achieved? 
+We suppose that part of the answer involves the fact that planning for the movements of each hand is not done entirely independently.
+Rather there is a plan for the whole action
+and plans for the movements of each hand are components of this larger plan.
+It is in part because they are parts of a larger plan that the plan for one hand constrains and is constrained by the plan for the other hand.
+
+How is this relevant to the case of joint action?
+In joint action the agents have the same goal, to move the object from one place to another.
+They also face a similar coordination problem, requiring a precisely timed swap from one hand to another.
+Now suppose, that the same planning is involved in the individual case (where one agent performs the whole action) and in the joint action case (where the action is distributed between two agents).
+The planning is the same almost up to the actual muscle contractions.
+
+How could this be helpful?
+Suppose the agents' planning processes are similar enough that, for a given context and problem, they will produce approximately the same plans.
+Then having each agent plan the whole joint action means that (i) each agent plans the other agent's action,
 (ii) each agent's plan for the other agent's action is approximately the same as that agent's plan for her own action [***AMBIGUOUS: `that agent's plan' must refer to the other agent, not `each agent',
@@
 
-\textbf{So what enables the two agents' plans to mesh is not that they represent each other's plans but more simply that they plan each other's actions as well as their own actions as if they were each about to do the whole thing themselves.}
+[***TODO: (a) contrast this case with a team of experts, each with different motor expertise (e.g.\ musicians playing together).  
+They can't plan each others' actions.
+(b) Discuss in how much detail each others' actions should be planned so as to enable coordination.
+]
+
+So what enables the two agents' plans to mesh is not that they represent each other's plans but more simply that they plan each other's actions as well as their own actions as if they were each about to do the whole thing themselves.
 
@@
 And of course this is exactly what we want for small-scale joint action---we want two or more agents to act as one.
+This may be why the performance of dyads in joint actions often resembles the performance of individuals tasked with performing the whole action alone \citep{Knoblich:2003nf}.
 
 So what is the difference between the individual and the joint case?
-From the present point of view, the primary difference may be that in joint action there is a need to prevent execution of the parts of the action which are not one’s own.
+From the point of view of motor representation, the primary difference may be that in joint action there is a need to prevent execution of the parts of the action which are not one’s own.
 
-So far: shown theoretically how there could be directedness of joint action to outcome by appeal to reciprocal representations of outcomes which trigger agent-neutral planning.
+%Here then is the basic idea I take to be guiding Kourtis and others.
+%Coordination is sometimes achieved by having each agent’s motor system plan all of their actions; 
+%given some assumptions, this could be a way of making it likely that each will execute their part in the joint action in a way that meshes with the way the other agents execute their parts.
 
-But unclear that any such representations exist, or at least that they plan any significant role in ordinary cases of joint action.
-What makes this theoretical possibility interesting is evidence that these representations occur in motor control of action.
-
-***section: say what motor processes are; examine what would be evidence for different points.
-
-\section{Shared motor representation}
+\section{[*cut] Shared motor representation}
 Our argument starts with an empirical premise about enabling conditions for joint action: there are reciprocal, agent-neutral motor representations of outcomes whose obtaining would normally involve action on the part of each agent;
@@
 
-\section{Interpersonal coordination and motor representation}
-
-These bits of evidence are relevant (***):
-%
-\begin{enumerate}
-
-\item \citet{kourtis:2012_predictive} shows that motor planning can occur for others' actions when we are engaged in joint action with them.
-
-\item Vesper's ESPP paper (on jumping together and imagining jumping together --- jumping is published \citep{vesper:2012_jumping}, imagining jumping is about to be submitted (as of August 2012).
-This shows that individuals are capable of running motor simulations of multiple roughly simultaneous actions.
-(The important point for me is that one can simulate roughly simultaneous actions, not that the simulations are simultaneous.)
-
-\item The GROOP effect shows that there are representations which specify each agent's task in relation to the other (so they are not simple representing the outcomes to which each of their actions are directed; they are representing an outcome to which their actions taken together are directed.)
-
-\item Vesper says forthcoming EEG paper using piano playing paradigm on agent-neutral identification of error: one brain wave signals whether there is an error, and a different brain wave signals whose error it is (also tells you whether the overall harmonics are affected)
-\end{enumerate}
-
-It is hardly controversial that reciprocal motor representations exist, for their existence is suggested by by a large body of research on motor cognition in action observation.
-[***Rest of this paragraph belongs in the evidence section?
-Should be clearer about transition:
-(a) I am observing and we reciprocally represent an outcome of your action (mirroring).
-(b) We are interacting and reciprocally represent an outcome of your action (mirroring in joint action).
-(c) We are interacting and reciprocally represent an outcome to which our actions are distributively directed.%
-] 
-It is more controversial that reciprocal motor representation occurs in joint action, but there is some evidence for this claim too.%
-\footnote{
-See \citet{kourtis:2012_predictive}: `the partner’s expected action is simulated at the motor level, which probably facilitates effective performance of the joint action.'
-\citet{kourtis:2010_favoritism} show that 
-reciprocal motor representation is more likely to occur in joint action than is mere observation. 
-See also \citet{Knoblich:2003nf}.
- }
-
-What follows is speculative philosophy: we now take for granted that  social motor representation sometimes facilitates the exercise of shared agency and ask whether this matters for a philosophical account of shared agency.
-As a first step, we shall ask how social motor representation might facilitate agents in coordinating their actions.
-Note that answering this question does not directly commit us to any view about shared agency.
-After all, social motor representation might sometimes facilitate coordination while being extraneous to a philosophical account of shared agency.
-
-\section{How could social motor representation coordinate actions?}
+\section{[*cut] How could social motor representation coordinate actions?}
 Suppose that social motor representation is sometimes present when two or more agents act together.
@@
 How could social motor representation play a role in coordinating action when two or more agents act together?
-
-As a first step towards answering  this question let us illustrate a general principle about planning by scaling up from motor action all the way to corporate action.
-
-In small city state, Aravinda organises the trams and Gerhard the busses.
-Each is responsible for devising and updating the timetable for the service.
-Let us stipulate that the extent to which the tram and bus services are coordinated is to be measured by the sum of all passengers' total journey times.
-So a change to the timetables would result in better coordination  just if it would decrease the sum of passengers' total journey times.
-To illustrate, improving coordination might involve having trams arrive at a stop shortly before, rather than shortly after, busses depart from there.
-Now Aravinda and Gerhard each intend they, Aravinda and Gerhard, coordinate the trams with the busses  to the greatest extent possible given other constraints; and they intend to do this by way of these intentions and meshing subplans of them, and this is common knowledge between them.
-Aravinda and Gerhard thus meet Bratman's sufficient conditions for them to have a shared intention.
-Acting on this intention, 
-every January Aravinda updates the tram timetables and sends Gerhard the changes.
-Likewise, Gerhard  updates the bus timetables every June and sends Aravinda the changes.
-In this way  each is responsive to the other's intentions and subplans of these. 
-Indeed, each may even try to predict changes in the other's subplans and modify their own accordingly.  
-But from each individual's point of view, the other's plans are  merely a constraint.
-
-This approach to planning suffers from two defects.
-First, it is unlikely to be optimal in the sense of resulting, eventually, in a combination of plans such that no other combination of plans would have been better for at least one service and no worse for either service.
-Second, it is unlikely to be efficient in the sense of allowing Gerhard and Aravinda to arrive at an optimal combination of plans for the two services, trams and buses, with the fewest iterations.%
-% --- I removed this footnote because the case is a bit different
-% --- In Konvalinka, the subjects have to act simultaneously so they can't wait to see what the other does
-%\footnote{
-%The limits of this type of coordination, where each agent tries to predict the actions of the other, is nicely illustrated by the `two-way interaction' condition of an experiment by \citet{konvalinka:2010_follow}.
-%When subjects were required to synchronise the timing of finger taps with each other in this condition, 
-%%each had to predict the others' action and modify their own accordingly. 
-%they ended up with a pattern of errors `oscillating in opposite directions' as each attempted to move towards the other (p.\ 2228).
-%}
-
-How could they do better?
-One possibility may be to have a single person planning the busses and trams---perhaps, for example, Aravinda could buy Gerhard's franchise. 
-But suppose that this is not possible, and that there are limits on how much information Aravinda and Gerhard can share. 
-Then another way that might improve things would be to have Aravinda and Gerhard each plan everything.
-So Aravinda would plan the best timetable for all the trams and busses, and Gerhard would do the same.
-Since Aravinda only controls the trams, she can only implement the tram-related part of this master timetable.
-Likewise, Gerhard can only implement the bus part of his master timetable.
-But if Aravinda's and Gerhard's separate planning processes are sufficiently similar, this process may result in coordinated services.
-In this new case, 
-Aravinda and Gerhard coordinate their actions 
-not by knowing or predicting each others' plans 
-but by thinking about the best overall plan.
-Instead of each viewing the others' plans and actions as constraint on their own,
-in planning actions they absorb the others' actions.
-
-The question was how social motor representation might play a role in coordinating action.
-We propose that it might do so in the way illustrated by Aravinda and Gerhard's final attempt at coordination,
-where each plans everything despite only being in a position to execute half of the plan.
-
-%Let us go very slowly here.
-%Suppose that, in moving the mug, you try to locate two separate plans:
-%a plan for your left hand and a plan for your right hand.
-%In the Left Plan (as we might call it), the movements of the right hand feature as a constraint.
-%And conversely for the Right Plan.
-%It is almost as if you two hands are under the control of two separate agents who periodically share information about plans.
-%So if you update the Right Plan, you will need to make compensating changes to the Left Plan.
-%And of course those compensating changes to the left plan may in turn trigger further changes to the Right Plan.
-%And so on.
-%This approach might work in some cases. 
-%But it imposes two costs.
-%First, it is unlikely to be efficient in the sense of minimising the range of possible combinations that are considered. 
-%Second, it is unlikely to be optimal in the sense of finding a combination of plans such that no other combination of plans would have been better for at least one hand and no worse for any hand.
-%How could you have done beter?
-%The mistake was to have two separate plans, Left and Right, in each of which the movements of the other hand are a constraint.
-%The better
-
-To see how this might work,
-let us step down to motor action but first consider only  an individual action.
-Suppose an agent moves a mug from one place to another, passing it from her left hand to her right hand half way.
-It is a familiar idea that motor planning, like planning generally, involves starting with relatively abstract representations of outcomes and gradually filling in details.
-We can capture this by supposing that 
-motor representations for planning and monitoring action involve a hierarchical structure of representations.
-At the top we might find a relatively abstract representation of an outcome, in this case of the movement of the object from one location to another.
-Action-relevant details are progressively filled in by representations at lower stages of the hierarchy. 
-Now in the action we are considering there is a need, even for the single agent, to coordinate the exchange between her two hands.
-How is this achieved? 
-We suppose that part of the answer involves the fact that planning for the movements of each hand is not done entirely independently.
-Rather there is a plan for the whole action
-and plans for the movements of each hand are components of this larger plan.
-It is in part because they are parts of a larger plan that the plan for one hand constrains and is constrained by the plan for the other hand.
-
-How is this relevant to the case of joint action?
-In joint action the agents have the same goal, to move the object from one place to another.
-They also face a similar coordination problem, requiring a precisely timed swap from one hand to another.
-Now suppose, that the same planning is involved in the individual case (where one agent performs the whole action) and in the joint action case (where the action is distributed between two agents).
-The planning is the same almost up to the actual muscle contractions.
-
-How could this be helpful?
-Suppose the agents' planning processes are similar enough that, for a given context and problem, they will produce approximately the same plans.
-Then having each agent plan the whole joint action means that (i) each agent plans the other agent's action,
-(ii) each agent's plan for the other agent's action is approximately the same as that agent's plan for her own action [***AMBIGUOUS: `that agent's plan' must refer to the other agent, not `each agent',
-and 
-(iii) each  agent's plans for their own action are constrained by the plans for the other agent's action.
-
-[***TODO: (a) contrast this case with a team of experts, each with different motor expertise (e.g.\ musicians playing together).  
-They can't plan each others' actions.
-(b) Discuss in how much detail each others' actions should be planned so as to enable coordination.
-]
-
-So what enables the two agents' plans to mesh is not that they represent each other's plans but more simply that they plan each other's actions as well as their own actions as if they were each about to do the whole thing themselves.
-
-Each agent is planning (and monitoring) both their actions almost as if a single agent were going to execute the whole action.
-And of course this is exactly what we want for small-scale joint action---we want two or more agents to act as one.
-This may be why the performance of dyads in joint actions often resembles the performance of individuals tasked with performing the whole action alone \citep{Knoblich:2003nf}.
-
-So what is the difference between the individual and the joint case?
-From the point of view of motor representation, the primary difference may be that in joint action there is a need to prevent execution of the parts of the action which are not one’s own.
-
-%Here then is the basic idea I take to be guiding Kourtis and others.
-%Coordination is sometimes achieved by having each agent’s motor system plan all of their actions; 
-%given some assumptions, this could be a way of making it likely that each will execute their part in the joint action in a way that meshes with the way the other agents execute their parts.
 
@@
 
-\section{Conclusion}
-We have been considering how to provide an account of shared agency that might contribute to investigating a tangle philosophical and scientific questions.
+\section{*Interface Problem}
 
-Whereas some have claimed that shared agency can be fully explained in terms of a notion of shared intention, we have argued that some events are joint actions by virtue of being appropriately related to a structure of motor representations we call social motor representation.
-We don’t mean to suggest that all joint actions involve social motor representation.
-The view we are aiming to establish is rather this: Some events are joint actions in virtue of being appropriately related to social motor representations which bind their components together and ensure that there is a single outcome to which these components are collectively directed.
-This is why 
-%Recap: the question was: Does social motor representation  play a role in explaining what joint is?
-%I have just been arguing for a positive answer.
-%My thesis is this:
-%\textbf{Reciprocal agent-neutral motor representations coordinate multiple agents’ actions around an outcome in part by virtue of representing that outcome.}
-%That is, reciprocal social motor representations can ground the purposiveness of joint action.
-%This is why I think that 
-understanding shared agency requires understanding not only shared intention but also
- the coordinating role of social motor representation.
+[*Include this?]
 
-None of this is to deny that shared intention is among the ingredients needed to characterise shared agency.
-Indeed, it may be that the notion of social motor representation has a role to play in explaining what shared intention is.
-In constructing realisers of shared intention from ordinary individual intention, we need intentions \emph{that we $J$}.
-As has been much discussed \citep[e.g.][]{petersson_collectivity_2007}, the contents of these intentions cannot all refer to actions involving shared intentions.
-For this reason Michael Bratman suggests that things we intend are  cooperatively neutral activities. %*ref
+I think we can see a possible solution to this challenge by considering another problem.
+In constructing realisers of shared intention from ordinary individual intention, we need intentions \emph{that we J}.
+As has been much discussed, the contents of these intentions cannot all refer to actions involving shared intentions.
+For this reason Michael Bratman suggests that things we intend are  cooperatively neutral activities.
 It is then necessary to add further intentions in order to transform cooperatively neutral activities into joint actions.
-But it also seems possible that in some cases, what we intend when we intend that we $J$ is not a cooperatively neutral activity but instead a joint action of the sort which involves social motor representation.
+But it also seems possible that in some cases, what we intend when we intend that we J is not a cooperatively neutral activity but instead a joint action of the sort which involves social motor representation.
 
@@
 
-In conclusion, two things.
-First, some events are joint actions in virtue of being appropriately related to social motor representations which bind their components together and ensure that there is a single outcome to which these components are collectively directed.
-Second, and much more tentatively, in some cases social motor representation may be among the ingredients that realise a shared intention.
+\section{Conclusion}
+In conclusion,
+I have been arguing that 
+in order to fully understand 
+the relation between joint actions and the goals to which they are directed
+we need 
+to invoke 
+shared motor representations.
+This is because shared motor representations
+involve a representation, on the part of each agent, of a single outcome,
+and coordinate the agents' actions in such a way that, normally, the occurrence of the represented outcome will be facilitated.
+In this respect shared motor representation resembles shared intention.
+
+But there are also differences.
+Shared motor representation
+and
+shared intention  
+have \textbf{distinctive roles in explaining shared agency}.
+And, relatedly, shared motor representation is \textbf{unlike shared intention} in that its constituent representations are non-propositional in format and so it doesn't inferentially integrate with ordinary intentions and knowledge states.
+
+This has two consequences.
+One is that we cannot fully understand shared agency is if we focus only on shared intention.
+The other is that we face a challenge.
+The challenge is to explain how these two kinds of representation could sometimes harmoniously contribute to effective shared agency despite not being inferentially integrated in practical reasoning.
 

### 2012-12-07 intention motor representation goal ascription with corrado sinigaglia (paper 2)/Motor Representation in Goal Ascription.tex
updated

--- 2012-11-02 intention motor representation goal ascription with corrado sinigaglia (paper 2)/Motor Representation in Goal Ascription.tex
@@
 When judgements about the goals of observed actions depend on the observers' motor representations, which processes connect motor representations to judgements?
-We shall argue (in section \vref{sec:processes} that motor representations are related to experiences.
+We shall argue (in section \vref{sec:processes}) that motor representations are related to experiences.
 There is a parallel between visual and motor representations.
@@
 
-\section{***Notes for Action Experience section}
-
-DUSSELDORF ARGUMENT (rough, from Friday skype):
-(1) Motor imagery involves experience of action as such (e.g. of grasping as such) and not just of sensory consequences of actions
-(one relevant consideration is that you know not only which action you are imagining but when you have succeeded in imagining performing it);
-(corollary: there can be experiences of action without experiences of acting)
-(2) There are common elements between what you  experience when you imagine acting and what sometimes you experience when you observe an action; in particular, the experience of action is common to both
-Therefore
-(3) Sometimes observing action involves experiencing action as such
-
-------- there is evidence that motor representation might have content-respecting influences on judgment
-
-MR ----?----- JUD
-  experience: how to argue 
-….............how is possibile? experience revelatory of action
-------------------- parallel between colour and action experience
-
-TBE: content-respecting influences of (MR of G$_1$) onto (Jud about G$_2$) where G$_1$ matches G$_2$
-
-Explanation-1: MR of G captures directedness of action a -> experiences of action a as directed to G -> judgement that action is directed to goal G
-	Key component: Dusseldorf argument, imagination
-
-Explanation-2: MR of G captures directedness of action a -> experiences revelatory of action but experiences as of sensory consequences of actions -> association between experienced sensory consequences and action directed to goal G (how?
-for example, I know the sound of a newspaper being opened) -> judgement that action a is directed to goal G
-
-E.g. clapping:
-	representation of sound
-	representation of action
-	association: characteristic sounds<>action
-	Thanks to the association, experience of the sounds can result in a representation of the action
-
 \section{Experience of Action}
 \label{sec:processes}
-Our aim in this section is to explain how motor representations could sometimes have content-respecting influences on judgements about the goals of observed actions.
-Broadly, our will be proposal is that the influence of motor representation on observational judgements goes via experience.
-It is because motor representations can affect our experiences when we observe actions that those motor representations can facilitate judgements about the actions.
-However this is consistent with two quite different hypotheses concerning the experiences associated with motor representation in action observation.
+How could motor representations sometimes have content-respecting influences on judgements about the goals of observed actions?
+Broadly, our proposal is that the influence of motor representation on judgement goes via experience.
+It is because motor representations sometimes influence our experiences when we observe actions that those motor representations can facilitate judgements about the actions.
 
-One hypothesis, call it the Indirect Hypothesis, is based on evidence that, in observing action, motor representation can influence visual and acoustic experience.
-(Perhaps this holds for other modalities too).
-On this hypothesis, these influences enable an observer to identify goals to which the observed action is directed because the observer is additionally able to recognise actions from their associated sensory effects.
-To illustrate, suppose that an observer can associate the characteristic sound of a newspaper being snapped open with the action of snapping open a newspaper.
-Suppose also that, when observing an action, motor representations can influence whether, when or how she experiences this characteristic sound.
-Then motor processes can indirectly influence the speed and accuracy of the observer's judgements about which actions are directed to this goal, the snapping open of a newspaper, by influencing whether, when or how she experiences the characteristic sound of this action.%
+Let us explain the idea in slightly more detail.
+Observing actions sometimes involves \emph{experiences revelatory of action}, that is, experiences which provide the subject of experience with reasons for judgements about which action someone is performing.
+Suppose, for instance, that you are observing a stranger who you have never communicated with.
+You may be able to judge, on the basis of observation that the stranger is walking a dog or gasping a cane.
+You are able to make such judgements about the stranger's actions because your experiences provide you with reasons for them.
+
+We claim that some experiences revelatory of action depend on motor representations of outcomes, and in such a way that which actions the experiences reveal is determined, at least in part, by which outcomes are represented motorically.
+This is what explains how motor representation sometimes has a content-respecting influence on judgement.
+Our aim in  this section is to defend this claim about how experience depends on motor representation.
+Defending this claim involves several steps.
+
+The first step is to note that motor processes can influence visual processes.
+For instance, \citet{bortoletto:2011_action} show that motor plans, such as might be involved in producing a gesture with the hand, can influence visual processes.
+Here the influence of motor representations on visual processes varies depending on both the gesture being planned and the visual stimulus presented.
+Importantly, these influences can be detected before any action occurs, showing that it is the motor representations rather than any consequent movements which are influencing visual processes.%
 \footnote{
-The Indirect Hypothesis is inspired by and consistent with views defended in \citet{Csibra:2007fy} and *Knoblich.
-However, to our knowledge these authors have not considered exactly our question about content-respecting influences of motor representation on judgement, and we are not suggesting that the would endorse the Indirect Hypothesis in its entirety.
+For a recent review of some of the evidence concerning motor influences on perceptual processing, see \citet{halasz:2012_unconscious}. (*refs)
+}  
+Of course, this evidence does not show that motor representation influences experience.
+After all, it may be possible to influence visual processes in ways that make no difference to experience.
+But the evidence is relevant to our argument because if motor and perceptual processes did not interact at all, it might be reasonable to doubt that what is experienced can sometimes be determined, at least partly, by the contents of motor representations.
+
+Next consider evidence that motor representation can influence perceptual judgement.
+For some very special pairs of tones, the first is sometimes perceived as lower in pitch than the second and sometimes as higher in pitch.
+Call these \emph{ambiguous tone pairs}.
+Repp and Knoblich (\citeyear{repp:2007_action}) had a group of expert pianists and a control group of non-experts press two keys in sequence, where the first key was sometimes to the left, and sometimes to the right, of the second key.
+The key presses produced an ambiguous tone pair, and the tones always occurred in the same order regardless of which key was pressed first.
+By asking subjects to report how they perceived the relative pitch of the tones, Repp and Knoblich found that, for the expert pianists, the direction of the key presses influenced the perceived direction of the change in pitch.
+Could what influences perceptual experience in this case be not a motor representation but the occurrence of a movement, or perhaps even the perception of a movement of the subject's own fingers?
+Against this possibility, note that the effect was not observed in non-expert pianists: for them the direction of action did not measurably influence the perceived pitch direction.
+Since both groups moved in the same way, if the influence were due to movement only we would expect it to occur irrespective of expertise.
+Instead it seems likely that differences in expertise between the two groups of subjects, expert and non-expert, affected how the movements they performed were represented motorically, and that these differences in motor representation are in turn what explains their perception of relative pitch.
+This study, and others like it,%
+\footnote{
+See also \citet{zwickel:2010_interference} who investigate effects of action on visual experience of motion.  *more \citep{schutz-bosbach:2007_perceptual}
+}
+provide relatively direct evidence that motor representation can influence perceptual experience.
+
+There is a view about the control of action that may explain how motor representation sometimes influences perceptual experience.
+In planning one's future actions it is sometimes useful to be able to predict the likely sensory consequences of actions currently being performed.
+One can then treat these predictions as constraints and opportunities in planning what to do next.
+Accordingly, motor representations of outcomes may be inputs to mechanisms that generate predicted sensory consequences.
+And these predictions may not only feed into planning but could also influence how the consequences of actions are experienced.
+So in Repp and Knoblich's experiment, it is possible that when the expert pianists plan sequences of key presses, it is predictions about the likely sensory consequences of their actions which bias them to perceive the ambiguous tone pair one way or another.
+
+So far we have seen that motor representations can influence what is experienced.
+Our aim is to show, further, that some of these experiences are revelatory of action, and that the influences are such that the contents of motor representations determine, at least partly, which actions the experiences reveal.
+How might this work?
+Consider observing someone playing two notes on a piano.
+Judging which action they are performing may be difficult as your view may be partially obscured and you may be too far away to see exactly what is happening.
+But given that you know something about which sounds are produced by various piano playing actions, hearing the relative pitches of these notes might in principle assist you in recognizing that a goal of her action is to play these two notes.
+So if, as the findings indicate, motor representation can influence perceptions of pitch, it might thereby indirectly facilitate judgements about the goals of actions.
+
+For another illustration, suppose that an observer can associate the characteristic sound of a newspaper being snapped open with the action of snapping open a newspaper.
+Suppose also that, when observing an action, motor  representations can influence whether, when or how this observer experiences this characteristic sound.
+Then motor processes can indirectly influence the speed and accuracy of the observer's judgements about which actions are directed to this goal, the snapping open of a newspaper, by influencing whether, when or how she experiences the characteristic sound of this action.
+
+So far we have been considering experiences of characteristic sensory effects of actions.
+But experiences revelatory of action also include experiences of bodily configurations and movements as an action unfolds.
+[Ref exp Prinz and point lights]  These experiences may also be influenced by motor representations of outcomes.
+As Wilson and Knoblich (\citeyear{Wilson:2005qu}, p.\ 463) suggest, occlusion and other factors mean that observers typically have limited perceptual information about agents' bodies and movements: motor representation in action observation may usefully enable observers `to fill in missing or ambiguous information'.
+
+Here, then, is a proposal---call it the Indirect Hypothesis---about how the contents of motor representations partially determine which actions are revealed in experience (see also figure \vref{fig_indirect_hypothesis}).
+In observing action, motor representation can influence perceptual experiences.
+The relevant experiences are of bodily configurations and movements as an action unfolds, and of sensory effects of characteristic particular actions.
+Such experiences sometimes enable an observer to make judgements about the goals of actions, for she may associate particular patterns of bodily movement or particular sensory effects with particular goals.
+But in many cases of action observation, the agent's bodily movements and the sensory effects of her actions are partially obscured or surrounded by irrelevant distractors, or otherwise difficult to identify.
+But, as we have seen, there is evidence that motor representation can influence such experiences of bodily configurations and movements, and of sensory effects.
+Motor representations could thereby facilitate judgements about the goals of actions.
+So it is possible that motor representations sometimes enhance perceptual experiences of movements or sensory effects which in turn sometimes facilitate judgements about the goals of actions.%
+\footnote{
+The Indirect Hypothesis is inspired by and consistent with views defended in \citet{Csibra:2007fy} and \citet{Wilson:2005qu}.
+However, to our knowledge these authors have not considered  our question about content-respecting influences of motor representation on judgement, and we are not suggesting that the would endorse the Indirect Hypothesis in its entirety.
 }
 
-The other hypothesis is in some ways more radical.
-This hypothesis, call it the Direct Hypothesis, is that in observing action we experience not only movements and sounds but also goal-directed actions as such.
-Such experiences---call the motor experiences---stand to motor representations roughly as visual experiences stand to visual representations.
-The motor experience of an action as directed to a particular outcome is made possible by a motor representation of that outcome.
-So motor representations can have a content-respecting influence on judgements because motor representations enable motor experiences on which judgements are based.%
+\begin{figure}
+\begin{center}
+\includegraphics[width=\textwidth]{fig_indirect_hypothesis.png}
+\caption{
+\label{fig_indirect_hypothesis}
+	The Indirect Hypothesis.
+}
+\end{center}
+\end{figure}
+
+A possible objection to the Indirect Hypothesis is that the evidence we have considered leaves open the question of how extensive the influence of motor representations on perceptual experiences of bodily configurations, movements and the sensory effects of actions is.
+For all we have said, such influences may be rare.
+This is a potential problem for our attempt to use the Indirect Hypothesis to explain how motor representations have content-respecting influences on judgements.
+The Indirect Hypothesis does provide a theoretically coherent and empirically motivated explanation of how such influences are possible in principle.
+But currently available evidence does not allow us to determine whether the Indirect Hypothesis can explain every case in which a motor representations has a content-respecting influence on a judgement.
+Perhaps there is more to the story.
+
+\section{Motor Experience}
+Our question is how motor representations could have a content-respecting influence on judgements about the goals of actions.
+The answer, we suppose, must involve a relation between motor representation and experience.
+In the previous section we defended the hypothesis that motor representations of outcomes to which actions are directed influence perceptual experiences of the agents' bodies and movements in ways that facilitate judgements about the goals of the actions.
+But this hypothesis may not be the whole story about experience of action.
+In this section we discuss a more radical hypothesis---call it the Direct Hypothesis---about how the contents of motor representations can influence what we experience when we observe actions.
+
+The Direct Hypothesis starts from the premise that in observing action we experience not only bodily configurations, movements, sounds and the rest but also goal-directed actions as such .
+Such experiences---call them motor experiences---stand to motor representations somewhat as visual experiences stand to visual representations.
+So motor representations can have a content-respecting influence on judgements because motor representations enable motor experiences on which the judgements can be  based.
+Or so the Direct Hypothesis claims
+(see figure \vref{fig_direct_hypothesis}).%
 \footnote{
@@
 
-Let us say that an experience, motor or not, is revelatory of action just if it  can provide the subject of experience with reasons for correct judgements about which action someone is performing.
-On both hypotheses motor representations enable experiences which are revelatory of action.
-And both hypotheses use this broad idea to explain how motor representations could have content-respecting influences on judgements.
-The primary difference between the hypotheses concerns what these revelatory experiences are experiences of: whether all such are experiences of movements, sounds and the like only (the Indirect Hypothesis), or whether some are experiences of goal-directed actions as such (the Direct Hypothesis).
-
-In the next section we expand on and consider evidence for the first hypothesis before turning to the second hypothesis in the following section.
-
-\section{The Indirect Hypothesis}
-The Indirect Hypothesis is depends on the claim that motor representations can influence sensory experiences.
-This claim is supported by a variety of evidence.
-\citet{bortoletto:2011_action} show that motor plans, such as might be involved in producing a gesture with the hand, can influence visual processes \citep{bortoletto:2011_action}.
-The influence of motor representations on visual processes varies depending on both the gesture to be performed and the visual stimulus presented.
-Importantly, these influences can be detected before any action occurs, showing that it is the motor representations rather than any consequent movements which are influencing visual processes.
-[*Some other evidence; stress that the effect can be found using different measures.]  As it may be possible to influence visual processes in ways that make no difference to experience, we do not take these findings alone to establish the claim that motor representations could enable visual experiences.
-But these findings do at least show that the claim is consistent with the neuropsychological facts.%
-\footnote{
-For a recent review of some of the evidence concerning motor influences on perceptual processing, see \citet{halasz:2012_unconscious}. (*refs)
+\begin{figure}
+\begin{center}
+\includegraphics[width=\textwidth]{fig_direct_hypothesis.png}
+\caption{
+\label{fig_direct_hypothesis}
+	The Direct Hypothesis.
 }
-
-What would establish that some motor representations enable perceptual experiences?
-For some very special pairs of tones, the first is sometimes perceived as lower in pitch than the second and sometimes as higher in pitch.
-Repp and Knoblich (\citeyear{repp:2007_action}) had a group of expert pianists and a control group of non-experts press two keys in sequence, where the first key was sometimes left and sometimes right of the second key.
-They found that, for the expert pianists, the direction of the key presses modulated the experience: it influenced the perceived direction of the change in pitch.
-Could what modulates experience in this case be not a motor representation but the occurrence of a movement, or perhaps even the perception of a movement of the subject's own fingers?
-Against this, note that the effect was not observed in non-expert pianists: for them the direction of action did not measurably influence the experienced pitch direction.
-Since both groups moved in the same way, if the modulation were due to movement only we would expect it to occur irrespective of expertise.
-Instead it seems likely that differences in expertise between the two groups of subjects affected how the movements they performed were represented motorically, and these differences in motor representation are in turn what explains their experience of the change in pitch.
-This study, and others like it,%
-\footnote{
-See also \citet{zwickel:2010_interference} who investigate effects of action on visual experience of motion.  *more
-} 
-provide relatively direct evidence that motor representations can enable acoustic experiences.
-
-How do these findings bear on the second hypothesis above, according to which motor representation in action observation enables experiences of sounds and movements that somehow facilitate the identification of actions?
-Perhaps the motor representations enable certain experiences of kinematic features which somehow enable subjects to better identify actions.
-We do not claim that this is entirely plausible, only that the second hypothesis cannot be rejected out of hand.
-This is enough to show that we cannot infer that there are experiences of action just from the fact that motor representation enables experiences which are revelatory of action.
-Stronger evidence for the first hypothesis is needed.
-
-\subsection{Experiences of action}
-Our focus is experiences involved in observing actions, but let us first consider experiences involved in performing actions.
-We shall argue that acting sometimes involves experiences which can ground judgements about what one has done even though these are not experiences of movements or sounds.
-In effect, the idea will be that you can strip away perceptual experiences and still be left with experiences revelatory of action.
-These can only be experiences of action, or so we claim.
-Two phenomena are involved in making this argument.
-The first is anosognosia for hemiplegia, which shows that there can be experiences revelatory of action even in the absence of any actual movements.
-The second is deafference, which confirms that these experiences are unlikely to be illusory experiences of movement.
-Together these phenomena support the view that, in performing actions, we have experiences of actions and not only of movements, sounds and the like.
-
-The argument sketched so far concerns experiences associated with performing actions, not observing them.
-Note, however, that the basis of these experiences of action is motor representation and planning.
-Since these are the very things which, in action observation, underlie the experiences enabling one to identify actions, it is reasonable to conclude that there is a phenomenal element common to performing an action and observing an action performed, namely the action itself.
-Roughly, what you experience in performing an action is what you would experience if you were observing it.
-But this is all to come; let us start by introducing some background.
-
-Anosognosia for hemiplegia is a clinical condition in which patients deny, and appear largely unaware of, a severe paralysis of one or more limbs on one side of their body.
-It typically follows a stroke involving damage to motor areas in the right hemisphere of the brain (*refs).
-To illustrate the effects of this condition, consider a subject with anosognosia for hemiplegia who was asked to brush her hair holding a brush in her contralesional hand.
-Although she was unable to move the hand, she proceeded to move her head as if her hair were being brushed, and then reported having successfully brushed her hair \citep{berti:2008_motor}.
-This form of anosognosia cannot be explained as entirely due to neglect or complete ignorance of part of one's body \citep[p.\ 165]{berti:2008_motor}.
-
-Could anosognosia for hemiplegia be explained as mere  confabulation or a top-down effect of judgement on experience?
-This is incompatible with the performance of anosognosics on bimanual tasks \citep{berti:2008_motor, garbarini:2012_moving}.
-To illustrate, consider an anosognosic who has been asked to wash with both hands.
-Although she can only actually move one hand, she may move this hand in much the way that she would do if she were actually washing with both hands (*ref: personal communication?).
-This is hard for ordinary subjects to reproduce (try it).
-For a more carefully controlled illustration, consider anosognosic subjects who were asked to draw simultaneously with both hands, one hand was supposed to continuously draw a vertical line and the other hand to continuously draws a circle.
-Anosognosic subjects can of course only actually move one hand, but they show interference as if both hands were actually being moved.
-That is, if you were to perform this task, your vertical line would become somewhat oval; this interference is an effect of bimanual coordination.
-The striking finding is that anosognosic subjects show the same pattern of interference although not actually drawing a circle.
-By contrast, subjects with motor neglect and subjects with hemiplegia but no anosognosia do not show this pattern of interference \citep{garbarini:2012_moving}.
-This suggests that anosognosia involves relatively detailed action-related experiences and cannot be entirely explained as confabulation.
-
-On the leading, best supported explanation, anosognosia for hemiplegia arises from defects in monitoring action.
-When a subject with anosognosia for hemiplegia is asked to perform an action involving a hemiparetic limb, motor planning occurs as it might do in ordinary subjects.
-Now in ordinary subjects, and in hemiplegic subjects without anosognosia, the predicted effects of a planned action are compared with motor representations of action outcomes.
-Consequently if such subjects tried and failed to initiate an action, they would normally be aware of their failure even without seeing or touching their limbs.
-However, in anosognosic subjects the comparator is damaged and no signal of failure is generated.
-This is why it seems to them that they have acted despite failing even to initiate the action.
-Of course, anosognosic subjects can perceive that their objectives have not been met: after apparently opening a bottle, for instance, they can see that the bottle remains shut; they would not normally proceed to drink from it.
-And these subjects do not necessarily have visual, proprioceptive or tactile sensory impairments.
-But the experience associated with motor planning dominates, so that it seems to them as if they have acted.
-From their point of view, it seems that the bottle is peculiarly difficult to open rather than that they are unable to move a limb \citep[pp.\ 173-4]{berti:2008_motor}.
-
-How is anosognosia for hemiplegia relevant to the issue of whether there are experiences of action as such?
-It seems that anosognosic subjects have some experiences which are revelatory of action---revelatory in the sense that they provide reasons for judgements about actions---despite not actually moving a limb.
-What could these experiences be experiences of?
-Consider the claim that these are experiences of movement only.
-While it is arguably not generally correct to infer an absence of experience of movement from the absence of actual movement, in this particular case it does seems unlikely that there could be experience of movement.
-After all, some anosognosic subjects have intact perceptual systems: if they experience movement at all, they should experience an absence of movement.
-So any experiences of movement will not be the sort of experiences that would provide reasons for judgements about successful action.
-Given that (as argued above) the experiences on which anosognosic subjects' judgements of action are based are also not entirely a product of confabulation, it is plausible to accept that they are experiences of action as such.
-
-This view is further supported by research on deafferented subjects.
-These subjects lack tactile and proprioceptive sensory feedback from their actions but, unlike subjects with anosognosia for hemiplegia, can actually move their limbs.
-Such subjects may continue to have some experiences characteristic of action; for instance, when asked to flex an index finger they can report having done so even in the absence of visual information \citep{kristeva:2006}.
-These subjects' sensory deficits make it unlikely that their experiences could be of movements or the like: it is more plausible that, concerning acting, their experiences are like those of subjects with anosognosia for hemiplegia.
-None experience movement: what they experience is action.
-
-Anosognosia and deafference comprise a useful pair of cases: some anosognosics we have preserved perceptual senses and an absence of actual bodily movement, whereas individuals suffering deafference have damaged senses and actual bodily movement.
-But in both cases there are experiences revelatory of action.
-The existence of these disorders indicates that motor representation is sufficient for experiences capable of grounding judgements about actions even in the absence of any perceptual experiences of movements, sounds or the like.
-It is plausible to infer that these experiences are experiences of action as such.
-
-***HERE: todo.
-Two things.
-First, explain the timing experiment---it's not all about subjects with brain damage, there is converging evidence from ordinary subjects.
-Second, explain the comparitor model.
-
-***Anosognosia for hemiplegia : experience of action without acting (cannot be an experience of movement, \&c).
-Or could this be an illusory experience of one's body moving?
-\citep{garbarini:2012_moving}%
-\footnote{
-See also \citet{gallese:2010_bodily}: "Brugger, Kollias, Muri, Crelier, and Hepp-Reymond (2000) described one of these cases.
-fMRI imaging of their patient during phantom limb sensations of hand movements showed activation of premotor and posterior parietal cortex.
-Aplasic phantoms can therefore be explained as the phenomenal correlate of planning/monitoring action of an absent limb."}
-and hand-washing: \citep{berti:2008_motor}
-
-***parallel with speech (short)
-
-\section{Conclusion}
-Motor representation enables judgements about the goals of actions by making possible experiences of action as such.
-
-\subsection{Reasons}
-In making judgements about the goals of actions, it appears (perhaps veridically) to the thinker that she has reasons for those judgements, much as it appears to her that she has when making visual judgements about the particular colours of objects.
-Or, if we are sceptical about thinkers having reasons for visual judgements, at least we can say that it does not seem to the thinker that she is simply landed with the judgement.
-Contrast someone who `knew at a glance' that a vase is genuine Ding ware, or someone on a blind date who, on first catching sight of her partner for the night, `just knows' that this is the one she wants to spend the rest of her life with.
-Action judgements often seem different from these: rather than simply being landed with them, it may seem to a thinker that they bear a straightforward relation to her experience.
-
-Our view is consistent with the claim that it appears to a thinker that she has reasons for making judgements about the goals of actions even when  those judgements depend on motor representations.
-  Motor representation stands to experience of action as visual representation stands to experience of particular colours.
-
-\subsection{Goal ascription without metarepresentation}
-Having noted above that goal ascription often or always involves representing the directedness of an action to an outcome, we argued that this is something no motor representation can represent.
-Equally, however, we noted that the directedness of actions to outcomes can be captured by planning-like processes, and we suggested that motor representations can feature in such processes.
-One possibility, then, is that there might be a kind of goal ascription---or something resembling goal ascription---where the only representations involved are motor representations.
-This enables us to see how there could be goal ascription, or something very like it (we are not insisting that the term 'goal ascription' be used) could occur without metarepresentation.
-This gives us a first role for motor representation in goal ascription which distinguishes motor representation from representations of intentions or other mental states.
-Where goal ascription involves capturing the directedness of actions to goals by representations of intentions it thereby involves metarepresentation; but where the only representations involved in goal ascription are motor representations, there is no metarepresentation at all.
-
-\section{* PLAN *}
-[Corrado's Challenge: Why is this not just about implementation details?]
-\begin{enumerate}
-\item What roles, if any, might motor representation play in goal ascription?
-\item Some motor representations occur in action observation and sometimes facilitate goal ascription (Serino; basketball; \& others?) (otherwise the question would be bizarre).
-This is what makes the question pressing (timely).
-\item The Question: How is this possible?  How could motor representation facilitate goal ascription?
-\item Philosophical motivation: A familiar view in philosophy has it that goal ascription depends on the ascription of intentions, which in turn is interdependent with ascriptions of belief and desire (*refs).
-On this view goal ascription is possible only as part of a larger mindreading project.
-By contrast, developmental psychologists have assumed that goal ascription is possible even without knowing anything of an agent's mental states (*refs).
-Our concern is with this pure form of goal ascription, goal ascription that occurs independently of knowledge of mental states.
-\item There are two problems.  (a) adequacy of motor representation; (b) link between motor representation and judgement about the action and its goal
-\item Problem (a).  Motor representations may represent outcomes but not the directedness of actions to outcomes?
-\item Solution to problem (a): Capturing vs. representing the directedness of actions.
-\item Problem (b): If we thought of motor representation as cut off from experience, it would seem impossible that differences in motor representation could make a difference to the reasons available from the observer's point of view.
-\item However, there is evidence that motor representation influences experience (e.g. acoustic experience of pitch direction).
-\item One possibility: motor representation influences visual and acoustic experiences of movements and sounds, and these somehow facilitate goal ascription
-\item This does not seem to be a fully adequate explanation because ?
-\item Furthermore, motor representation enables experience which are (i) revelatory of action and (ii) distinct from visual or auditory experiences (Berti \&c).
-\item So maybe: motor representation enables experiences of action as such
-\item Conclusion: motor representation enables goal ascription by making possible experiences of action as such.
-The role of motor representation in goal ascription is like the role of visual representation in judgements of motion.
-\end{enumerate}
-
-MR----> | acquisition| Action Judgment (action concept) <----> B \& D Judgment (b\&d concept)
-tracking goal -----> || ascribing goal <----------------- > ascribing beliefs and desires
-
-MR<----> Action Judgment (action concept) <----> B \& D Judgment (b\&d concept)
+\end{center}
+\end{figure}
 

### 2012-12-07 counterexample to bratman on shared intention/a counterexample to bratman on shared intention.tex
working on it

@@
 %\date{}
+
+%avoid overhang
+\tolerance=5000
 
@@
 
-The first contrast case, Gilbert's, shows that the difference can’t be just  a matter of coordination because people who are merely happen to be walking side by side each other also need to coordinate their actions in order to avoid colliding.
+The first contrast case, Gilbert's, shows that the difference can’t be just  a matter of coordination because people who merely happen to be walking side by side each other also need to coordinate their actions in order to avoid colliding.
 Note also that in both cases each individual's walking is intentional, so our intentionally walking together cannot be  only a matter of our each intentionally walking.
@@
 \footnote{
-Many philosophers agree that a notion of shared intention is useful for understanding acting together. 
+Many philosophers hold that a notion of shared intention is useful for understanding acting together. 
 Compare \citet[p.\ 5]{Gilbert:2006wr}: `I take a collective action to involve a collective intention.'  See also  
@@
 The \emph{weak claim} is that there is some J such that these conditions are sufficient for you and I to intend that we J.
-The \emph{strong claim} is that for any J, these conditions are sufficient for you and I to intend that we J. 
+The \emph{strong claim\label{strong_claim}} is that for any J, these conditions are sufficient for you and I to intend that we J. 
 Our initial counterexample will be directed to the strong claim.
@@
 
-How could we show that our meeting Bratman's conditions, (1)--(3), is not in fact sufficient for us to have a shared intention? 
-We seek a case where the conditions are met although we lack a shared intention.
+To show that  Bratman's conditions, (1)--(3), are not in fact sufficient for shared intention we need 
+a case where the conditions are met and we lack a shared intention.
 But how could we determine that we lack a shared intention?
@@
 Now suppose 
-	that, for some relevant J, Bratman's conditions, (1)--(3) above, were met in cases B and C alike, 
-	that in each case the structure of intention and knowledge played an appropriate role in guiding the agents' actions,
-	and that in each case the agents thereby successfully J.
+	that, for some J, Bratman's conditions, (1)--(3) above, were met in cases B and C alike.
+Suppose also that in each case, B and C, the structure of intention and knowledge mention specified by Bratman's conditions 
+ played an appropriate role in guiding the agents' actions,
+	and that in each case the agents did thereby  successfully J.
+Finally, suppose that cases A, B and C are as similar as possible except for the differences mentioned.
 Then case B would be our counterexample:
@@
 And it's even possible to desire or intend to be involved in an act of this type.
-A gangster might intend to go to New York with a future victim without having yet decided whether they will do this cooperatively or coercively.
+A gangster might intend to go to New York with someone without having yet decided whether they will do this cooperatively or coercively.
 
@@
 \section{Unshared Intentions}
+\label{sec:unshared_intentions}
 As a further preliminary we need to introduce a definition.
@@
 Consider the possibility of one player intending, unilaterally, that the two players do something.
-Suppose that one of the players---Ayesha, say---can knowledgeably predict that if she performs a certain sequence actions, <a$_1$, a$_2$, ...\ a$_n$>, then Ahmed will simultaneously perform certain other actions, <b$_1$, b$_2$, ...\ b$_n$>,
+Suppose that one of the players---Ayesha, say---can knowledgeably predict that if she performs a certain sequence of actions, <a$_1$, a$_2$, ...\ a$_n$>, then Ahmed will simultaneously perform certain other actions, <b$_1$, b$_2$, ...\ b$_n$>,
  and the upshot will be that the cross hits the red square.
@@
 Suppose, further, that Ayesha can intend to perform those actions <a$_1$, a$_2$, ...\ a$_n$>.
-Then Ayesha can intend, unilaterally, that they, Ayesha and Ahmed, make the cross hit the red square (see Section \vref{sec:netural_wrt_shared_intentionality}).
+Then Ayesha can intend, unilaterally, that they, Ayesha and Ahmed, make the cross hit the red square.
+(This intention concerns an action conceived of as neutral with respect to shared intentionality; see  Section \vref{sec:netural_wrt_shared_intentionality}).
 
@@
 To illustrate, 
-their subplans would fail to mesh if, in intending that they make the cross hit the red square,  Ayesha's plans had included pushing Ahmed out of the way and seizing his controls. 
-They would also fail to mesh if Ayesha were planning to trick Ahmed into a situation where we would be unable to perform the actions he had been planning.
-But in the case we have been describing, the agents subplans mesh perfectly.
+Ayesha's and Ahmed's subplans would fail to mesh if, in intending that they make the cross hit the red square,  Ayesha's plans had included pushing Ahmed out of the way and seizing his controls. 
+They would also fail to mesh if Ayesha were planning to trick Ahmed into a situation where he would be unable to perform the actions he had been planning.
+But in the case we have been describing, the agents' subplans mesh perfectly.
 Each agent's subplans involve manipulating his or her own controls,
@@
 (Beatrice relies on the fact Baldric intends that they J$_1$, of course; but she does not rely on the fact that Baldric intends what she intends.) 
-Furthermore, due to an artefact of the way the game is structured,
+Let us further stipulate that, due to an artefact of the way the game is structured,
 the unshared intention and the 
@@
 	\label{df:bratman_intention}
-	(as we might label the structure of intention and knowledge while leaving open whether it constitutes a shared intention), result in the two pairs performing same actions in the same way.
-That is, Beatrice reasons about Baldric much as Ayesha reasons about Ahmed and Beatrice and does what Ayesha does, and likewise for Baldric and Ahmed.
+	(as we might label the structure of intention and knowledge 
+	%specified in conditions (1)--(3) on page \pageref{quote:bratman_account}
+	while leaving open whether it constitutes a shared intention) result in the two pairs performing same actions in the same way.
+That is, Beatrice reasons about Baldric much as Ayesha reasons about Ahmed and Beatrice and does what Ayesha does; and likewise for Baldric and Ahmed.
 
@@
 (This is not due to ignorance: it's just how they choose to approach their tasks.)
-So if we consider on how  
-		the case of Beatrice and Baldric
+So if we consider  how  
+		 Beatrice and Baldric's case
 	differs from
@@
 
-To show that Beatrice and Baldric do not have a shared intention it remains only for us to show that Ayesha and Ahmed do not have one. 
+To show that Beatrice and Baldric do not have a shared intention it remains only to show that Ayesha and Ahmed do not have one. 
 Here we must be careful.
@@
 that there be a single activity  about which each the agent involved has an intention? 
-This might reasonably be doubted by those who, like Bratman, reject the Simple View according to which when an individual intentionally F-s she has an intention concerning her F-ing \citep{Bratman:1984jr}.
+This might reasonably be doubted by those who, like Bratman, reject the Simple View according to which when an individual intentionally J-s she has an intention concerning her J-ing \citep{Bratman:1984jr}.
 So 
 	we shall not infer  that Ayesha and Ahmed lack a shared intention
-	just because (by construction) there is no F such that Ayesha and Ahmed each intend that they, Ayesha and Ahmed, F.
+	just because (by construction) there is no J such that Ayesha and Ahmed each intend that they, Ayesha and Ahmed, J.
 
@@
 If so, we can already claim that Ayesha and Ahmed are not acting on a shared intention.
-But philosophers' intuitions about shared agency may not be entirely, so it would be better if we could avoid such a blunt appeal to intuition.
+But philosophers' intuitions about shared agency may vary, so it would be better if we could avoid such a blunt appeal to intuition.
 
@@
 This is the aim of the following section.
-A first motive for introducing this new contrast case is that it provides an indication concerning what is missing from Bratman's account, and so may help us to understand why his conditions are not sufficient for shared intention.
+A further motive for introducing this new contrast case is that it provides an indication concerning what is missing from Bratman's account, and so may help us to understand why his conditions are not sufficient for shared intention.
 
@@
 This sort of planning is quite common.
-For example, some housemates who have decided to take on an allotment to grow vegetables might sit down together to plan what needs doing without yet assigning roles to particular individuals. 
+For example, some housemates who have decided to make a pizza might sit down together to plan what needs doing without yet assigning roles to particular individuals. 
 In so planning, each housemate is thinking about what is to be done and not what she herself will do.  
@@
 
-In the above example of agent-neutral planning,
+In this example
 	the housemates plan together and agree on a common plan.
@@
 	Note, however, that an individual can construct an agent-neutral plan by herself, even if its eventual execution will involve others. 
-	In fact, two or more individuals who are assigned a task might each individually engage in agent-neutral planning in parallel.
-	The task demands and their planning strategies may conspire to ensure that they each come up with the same agent-neutral plan.
-	The task demands and manifest properties of the agents, such as their distribution in space, may also ensure that each agent also assigns the same roles to the same individuals.
+	In fact, two or more individuals who are assigned the same task and who will eventually collaborate might first each individually engage in agent-neutral planning.
+
+Suppose, then, that our pizza-making housemates each individually devise an agent-neutral plan and come up with an assignment of agents to roles.  
+What could ensure that their parallel planning results in identical plans?
 (Strictly speaking it is not necessary for the plans and role assignments to be identical; it is enough if the resulting agent-specifying plans are, in a special sense, compatible.%
 \footnote{
+For an example of compatible plans, suppose that our task is to press a button simultaneously. 
+If your plan is that each of us will to start to act in exactly 60 seconds whereas my plan specifies that you are the leader and we will start to act when you do, then our plans are compatible.
+Consider also a case in which our task is to transport some apples and pears to a warehouse. 
+My plan involves you taking care of the apples while I put the pears into the yellow crates and load them into the van through the side door.
+If your plan involves me taking care of the pears while you put the apples into the green crates and load them into the van through the rear doors, our plans are compatible.
+Here is one way to capture this intuitive notion of compatibility.
 \label{fn:df_compatible}
@@
 if all facts about which agents have which roles in which plans were common knowledge to the agents,
-this would not affect the rationality of their each acting on the intention that they realise the outcome by performing the role she is assigned in her own plan.
+this would not affect the instrumental rationality of their each acting on the intention that they realise the outcome by performing the role she is assigned in her own plan.
 %	each agent could rationally act on the intention that they realise the outcome by performing the role she is assigned in her own plan,
 %	and she could do so even if all facts about which agents are performing which roles in which plans were common knowledge to the agents.
-To illustrate, suppose that our task is to press a button simultaneously. 
-If your plan is that each of us will to start to move in exactly 60 seconds and press the button 5 seconds later whereas my plan specifies that you are the leader and we are each to press the button 5 seconds after you start moving, then our plans are compatible.
 })
+Communication but not necessary ***here
+
+		The task demands and their planning strategies may conspire to ensure that they each come up with the same agent-neutral plan.
+	The task demands and manifest properties of the agents, such as their distribution in space, may also ensure that each agent  assigns the same roles to the same individuals.
+
 %
@@
   Of course, in planning your only your own actions, it may be necessary to consider how others' actions will affect, and be affected by, your won.
-But in this case, the others' actions feature in your planning as constraints and opportunities.
+But in this case,  others' actions feature in your planning as constraints and opportunities.
 By contrast, forming a plan for your own and others' actions 
@@
 The difference is just that Beatrice and Baldric make no use of the fact that they are performing the same task.
-	This fact does not show up in their planning. 
+	This fact does not feature in their planning. 
 	Rather, each plans her own actions only and treats the other's actions as constraints to work around or opportunities to exploit.
@@
  Beatrice and Baldric do meet Bratman's conditions for shared intention, and they do act appropriately on the corresponding intentions and knowledge. 
- So Beatrice and Baldric's case  is a counterexample to sufficiency of Bratman's conditions for shared intention (see (1)--(3) on page \pageref{quote:bratman_account}).
+ So Beatrice and Baldric's case  is a counterexample to sufficiency of Bratman's conditions for shared intention.
 
-In describing Caitlin and Ciaran we have not introduced contralateral commitments or other elements foreign to Bratman's account of shared agency. 
+In describing Caitlin and Ciaran we have not introduced contralateral commitments or any other element foreign to Bratman's account of shared agency. 
 Our counterexample draws on the same planning resources Bratman's account draws on, with just one addition.
 The addition is the idea that agents can conceive, or fail to conceive, of their actions as part of a single plan. 
-In at least some cases, shared agency differs from individual agency in part because shared agency involves changing perspective so that one’s own and other’s actions are conceived of as parts of a single plan.
+In at least some cases, shared agency differs from individual agency in part because shared agency involves changing perspective so that one’s own and others' actions are conceived of as parts of a single plan.
 
+\section{Mundane Counterexamples}
+Beatrice and Baldric's case is a counterexample to the claim that for any J, Bratman's conditions, (1)--(3) on page \pageref{quote:bratman_account}, are collectively sufficient for us to have a shared intention that we J.
+(We called this the `strong claim' earlier, \vpageref{strong_claim}.)
+If this counterexample were an isolated case, it might be tempting to suppose that it can provide little insight into shared agency.
+%After all, the sorts of activity that Bratman focuses on are things like walking together and painting the house together.
+Could proponents of Bratman's view avoid counterexamples by claiming that the his conditions are sufficient for mundane exercises of shared agency such as those involved in walking together or painting a house together?
+In this section we show that they could not.
 Given that Beatrice and Baldric's case is a counterexample,
 there are also many further counterexamples involving mundane activities and less elaborate props.
-Here, for instance, is how Beatrice and Baldric walk together.
+
+Here is how Beatrice and Baldric walk together.
 They are firmly tied at the ankle, and neither is strong enough to move without the other.
@@
 This is not a case of shared agency.
-The reason is not that Beatrice and Baldric are tied together against their will; 
-after all, many cases of shared agency are involuntary in this sense.
+The reason is not that Beatrice and Baldric are tied together against their wills; 
+after all, many cases of shared agency involve agents who are unwillingly bound to each other.
 The reason their walking does not involve shared agency 
 is rather that each conceives of the others' intentions and actions only as constraints and opportunities.
-They fail to make any use of the fact that they intend the same outcome,
- and so they fail see their actions as part of a single plan.
+Beatrice has a plan for what she, Beatrice, will do to ensure that they arrive at the corner; and Baldric likewise plans what he only will do.
+Each agent's plan relies on predictions about how the other will act and contains subplans which are conditional on the other's subplans and actions.
+But this involves no more than the sort of interdependence that might hold between two plans directed at different ends, as in cases of unshared intention.
+Beatrice and Baldric fail to make any use of the fact that they each intend the same thing, namely that they walk to the corner.
+
+For another counterexample, consider house painting.  
+First imagine that Ayesha and Ahmed are tasked with painting the outsides of adjacent houses on the same day,
+which happens to be windy.
+Each wants their house to be painted a single colour, and would prefer that it not be painted at all rather than that it be painted in several colours.
+But because they are spraying the paint onto the houses in blustery winds, neither can avoid spraying a significant quantity of paint onto the other's house as well as their own house. 
+In one respect this is fortunate. 
+For they will each lose so much paint from spraying in windy weather that, normally, neither would have enough paint to cover her house.
+But if they each make use of wind-borne spray from the other, both can succeed.
+They therefore form an unshared intention concerning their house painting projects. 
+(The notion of  unshared intention is explained in Section \vref{sec:unshared_intentions}.)
+In acting on this unshared intention,
+Ayesha starts painting first, using a colour which she knows is acceptable and available to Ahmed.
+Ayesha and Ahmed then each plan a route around their own house which will keep them out of the path of the other painter while also saving paint by taking advantage of foreseeable  wind-borne paint from the other.
+
+Ayesha and Ahmed's house painting involves only unshared intention and is therefore not a counterexample.
+But now imagine that there is a single house which Beatrice and Baldric are each motivated to paint.
+Because neither has sufficient paint for the job, they need to work together if they are to succeed.
+They therefore form a Bratman intention that they paint the house.
+But in acting on this intention, they plan and act much as Ayesha and Ahmed do.
+Each agent plans what she will do to ensure that they paint the house, treating the other's intentions and actions as constraints and opportunities. 
+Although the agents' plans and intentions interlock, 
+the agents do not conceive of their actions as parts of a single plan.
+From either agent's point of view, their intentions and actions are no more intimately related than are Ayesha and Ahmed's.
+This is not shared agency.
 
 Beatrice and Baldric's attitude may seem unnatural. 
-Perhaps humans in their situation would not normally fail to exploit the fact that they have the same task in the way that these agents do.
+Perhaps it is hard to believe that ordinary people in their situation would  fail to exploit the fact that they have the same task.
 This may be why Bratman's conditions appear, misleadingly, to be sufficient for shared intention.
@@
 
-One way to strengthen Bratman's account would be by adding a condition to his three.
+One way to strengthen Bratman's account would be by adding a fourth condition his three conditions.
 An alternative, which we prefer, is to introduce a new restriction on how agents who have a shared intention may conceive of J in intending that they J.
-As we have seen, Bratman's account requires only that, in some cases, the agents conceive of J in a way that is neutral with respect to shared intentionality.
+As we have seen, Bratman's account requires only that, in some cases, the agents conceive of J in a way that is neutral with respect to shared intentionality (see Section \vref{sec:netural_wrt_shared_intentionality}).
 We shall suggest a further requirement that is compatible with this one.
@@
 In walking to a metro station together,
-two friends may each form and act on a plan that specifies some details of both of their movements.
+two friends may each form and act on a plan that specifies some details of both of their actions.
 This is enough for their activity to be distributed in the above sense.
@@
 We started with some contrasts between shared agency and parallel but merely individual agency.
-These provide one intuitive fix on the notion of shared intention: 
+These provide an intuitive fix on the notion of shared intention: 
  shared intention, whatever it is, is what distinguishes thinks like two friends walking somewhere together from things like two strangers who walk the same route side by side.
@@
 we provided a counterexample to the view that a certain interlocking structure of intention and knowledge is sufficient for shared intention. 
-Two agents might meet conditions
-Bratman offers as sufficient for shared intention---that is, it might be common knowledge to them that they might each intend that they J and intend that they J in accordance with and because of these intentions and meshing subplans of them---even though they lack a shared intention.
+Two agents might lack a shared intention even though they meet conditions Bratman offers as sufficient for shared intention
+That is, they might lack a shared intention even though it is common knowledge to them that they  each intend that they J and intend that they J in accordance with and because of these intentions and meshing subplans of them.
 
@@
 When this happens, each agent conceives of the others' actions merely as constraints and opportunities.
-Bratman's conditions may initially appear be sufficient for shared intention because it may be natural that when for agents meet his conditions to make use of the fact that there is a single outcome to which all of their actions are directed and  to conceive of all their actions as parts of a single plan.
+Bratman's conditions may initially appear be sufficient for shared intention because it may be natural for agents who meet his conditions to make use of the fact that there is a single outcome to which all of their actions are directed and  to conceive of all their actions as parts of a single plan.
 Our counterexample, the case of Beatrice and Baldric, shows that although this may be natural, it not necessary.
@@
 We identified a way to revise Bratman's account so as to avoid this sort of counterexample without departing substantially from his planning approach.
-The remedy we propose hinges on the possibility that an agent, in planning for an outcome, can sometimes plan both her own and other agent's actions.
+The remedy we propose hinges on the possibility that an agent, in planning for an outcome, can sometimes plan both her own and other agents' actions.
 Further, in the right circumstances,
@@
 No amount of forming intentions about other's intentions and acting on knowledge of such intentions is sufficient, all by itself, for shared intention.  
-In many, perhaps all cases, shared agency requires changing perspective 
-to conceive of one's own and other's intentions and actions as part of a single plan.
+In many or all cases, shared agency requires changing perspective 
+to conceive of one's own and others' intentions and actions as parts of a single plan.
 

### 2012-12-08 counterexample to bratman on shared intention/a counterexample to bratman on shared intention.tex
workinhg on it added moran velleman

@@
 This paper provides a counterexample to the sufficiency of this condition for shared intention.
-The counterexample arises because it is possible for agents to meet the condition, and to act rationally on the specified knowledge and intentions,
+The counterexample arises because it is possible for agents to meet the condition, and to act rationally on the specified  intentions and knowledge,
  while conceiving of each other's actions only as constraints to work around and opportunities to exploit.
@@
 \section{The Counterexample}
-%*Need a beefier activity, something more than cooperatively neutral.
+\label{sec:the_counterexample}
 
@@
 
-\section{Agent-neutral Plans}
+\section{Individually planning all of  the agents' actions}
 \label{sec:distributed_plan}
@@
 
-First we need to introduce the notion of an agent-neutral plan.
+First we need to introduce the notion of an agent-neutral plan and some  observations.
 We stipulate that a planning process, or a plan, is \emph{agent-neutral} just if it does not involve identifying any particular agents.  
@@
 For example, some housemates who have decided to make a pizza might sit down together to plan what needs doing without yet assigning roles to particular individuals. 
+Someone needs to prepare the dough, 
+another person should prepare toppings,
+while a third might mix some salad.
 In so planning, each housemate is thinking about what is to be done and not what she herself will do.  
@@
 	Note, however, that an individual can construct an agent-neutral plan by herself, even if its eventual execution will involve others. 
-	In fact, two or more individuals who are assigned the same task and who will eventually collaborate might first each individually engage in agent-neutral planning.
+	In fact, two or more individuals who are assigned the same task and who will eventually collaborate might first each individually engage in agent-neutral planning and then individually assign roles.
+	And they might do this in isolation from each other, without discussion or communication.
 
-Suppose, then, that our pizza-making housemates each individually devise an agent-neutral plan and come up with an assignment of agents to roles.  
-What could ensure that their parallel planning results in identical plans?
-(Strictly speaking it is not necessary for the plans and role assignments to be identical; it is enough if the resulting agent-specifying plans are, in a special sense, compatible.%
+Where there is an outcome and some agents each individually,
+without discussion or communication,
+plan for the outcome 
+and where each agent has a single plan specifying roles for herself and all of the other agents,
+%When there is an outcome and some agents each individually,
+%without any discussion or communication,
+%construct an agent-neutral plan for this outcome and then assign roles to themselves and the other agents,
+we shall say that these agents each \emph{individually plan all of the agents' actions}.%
 \footnote{
-For an example of compatible plans, suppose that our task is to press a button simultaneously. 
-If your plan is that each of us will to start to act in exactly 60 seconds whereas my plan specifies that you are the leader and we will start to act when you do, then our plans are compatible.
-Consider also a case in which our task is to transport some apples and pears to a warehouse. 
+This phrase is ambiguous (in a way we know  no way concise of avoiding).
+The idea is that each agent has a single plan in which she and  all other agents are assigned roles.
+There is no requirement that any agent's plan must specify all the actions which the agents between them will perform.
+}
+Agents can do this by first constructing agent-neutral plans and then assigning roles to themselves and the other agents.
+So, to return to our earlier example, 
+ if the housemates, having decided to make a pizza, had each individually devised  agent-neutral plans for the construction of the pizza and then each individually assigned roles in their own plans to themselves and the others,
+then they would have each individually planed all of the agents' actions.
+
+For all we have said,  agent-neutral plans might not be involved in all cases where some agents each individually plan all of the agents' actions.
+We invoke agent-neutral plans only as an aid to showing that the notion of some 
+agents each individually planning all of the agents' actions is coherent.
+
+If some agents each individually plan all of the agents' actions, 
+could they nevertheless know  that they will form plans that are sufficiently similar or jointly viable?%
+\footnote{
+\label{fn:df_jointly_viable}
+Suppose that, for some outcome, two or more agents each have a plan for the realisation of that outcome. 
+%(These plans may, but need not, specify roles for all of the agents; but the plans must be agent-specific, not agent-neutral.)
+We stipulate that these plans are  \emph{jointly viable} just if
+the plan which involves each agent performing the roles she is assigned in her own plan is acceptable to all the agents and a viable means of realising the outcome.
+For an example of jointly viable but significantly different plans, suppose that our task is to transport some apples and pears to a warehouse. 
 My plan involves you taking care of the apples while I put the pears into the yellow crates and load them into the van through the side door.
-If your plan involves me taking care of the pears while you put the apples into the green crates and load them into the van through the rear doors, our plans are compatible.
-Here is one way to capture this intuitive notion of compatibility.
-\label{fn:df_compatible}
-Suppose that, for some outcome, two or more agents each have a plan for the realisation of that outcome. 
-(These plans may, but need not, specify roles for all of the agents; but the plans must be agent-specific, not agent-neutral.)
-By saying that these plans are \emph{compatible} we mean that:
-(i) 
-no agent would normally be prevented from performing the role she is assigned in her own plan by other agents performing the roles they are assigned in their plans;
-and
-(ii)
-if all facts about which agents have which roles in which plans were common knowledge to the agents,
-this would not affect the instrumental rationality of their each acting on the intention that they realise the outcome by performing the role she is assigned in her own plan.
-%	each agent could rationally act on the intention that they realise the outcome by performing the role she is assigned in her own plan,
-%	and she could do so even if all facts about which agents are performing which roles in which plans were common knowledge to the agents.
-})
-Communication but not necessary ***here
+If your plan involves me taking care of the pears while you put the apples into the green crates and load them into the van through the rear doors, then our plans are compatible.
+Consider also a case in which our task is to press a button simultaneously. 
+If your plan is that each of us will to start to act in exactly 60 seconds whereas my plan specifies that you are the leader and we will start to act when you do, then our plans are jointly viable.
+}
+%Where the agents could bring about an outcome in several different, equally good ways,
+%having some agents each individually plan all of the agents' actions is probably not useful.
+Consider housemates
+	making a pizza together, where everyone knows roughly the same recipe and differences are unlikely to matter.
+Or consider two agents 
+ moving  a table from one room to another together, where there is an obviously optimal route and the agents' positions manifestly determine which ends of the tables they should hold.
+For these outcomes and situations,
+ the agents' expertise and planning strategies 
+ may mean that each could knowledgeably predict that they will form sufficiently similar agent-neutral plans.
+And sometimes commonly known properties of the agents, such as their distribution in space or their sizes, mean that there is exactly one most obvious way of dividing up roles.
+%If  each agent also  knows enough about herself and the others to be able to determine, without communicating, whether her plan and role assignments will be acceptable to everyone, then
+% it need not be irrational for the agents each individually to plan all of the agents' actions. 
+In any such case, the agents who each individually plan all of the agents' actions may form sufficiently similar plans and know that they will do so.
 
-		The task demands and their planning strategies may conspire to ensure that they each come up with the same agent-neutral plan.
-	The task demands and manifest properties of the agents, such as their distribution in space, may also ensure that each agent  assigns the same roles to the same individuals.
-
-%
-Finally, each agent may know enough about herself and the others to be able to determine, without communicating, whether the plan and role assignments will be acceptable to everyone. 
-	And all of this---that they engage in parallel, agent-neutral planning resulting in identical (or compatible) plans and role assignments, which are acceptable to all---may be common knowledge to the agents.
-	So it is possible, in principle at least, 
-	that several agents might each individually engage in agent-neutral planning and rationally perform their part in the resulting plan, knowing that the others will do likewise.
-	Parallel, agent-neutral planning can rationally result in coordinated action without presupposing shared agency.%
+Agents who will collaborate in some way, such as by making a pizza together, can coordinate their actions by each individually planning all of the agents' actions.
+Suppose some agents 
+	 are about to act together in order to bring about some outcome.
+They know (we stipulate) that if they each individually plan all of the agents' actions, they will form sufficiently similar plans.
+Each agent also knows enough about herself and the others to be able to determine, without communicating, that the resulting plans will be acceptable to everyone. 
+And all of this is common knowledge to the agents.  
+Then these agents 
+might individually plan all of the agents' actions,
+and each might rationally perform her part in her plan, knowing that the others will do likewise.
+In this way, 
+ some agents' individually planning all of the agents' actions can rationally coordinate their plans and actions.%
 \footnote{
-Note that we are claiming only that shared agency is not presupposed.
-Our view is consistent with (but does not depend on) the claim
-that
-	if some agents each engage in parallel, agent-neutral planning and then rationally perform their part in the resulting plan,
-	the upshot would be an exercise of shared agency.
+Note that this coordination is achieved without presupposing that the agents exercise shared agency.
+Of course,
+	if some agents individually plan all of the agents' actions
+	and then each rationally performs her part in her own plan,
+	the upshot might be an exercise of shared agency.
+But this shared agency is a consequence of, not a requirement on, their individually planning all of the agents' actions.
 }
 
-Different perspectives are involved in
-	forming a plan for  your own actions  only
+But why would some agents each individually plan all of the agents' actions?
+After all, no agent gets  to perform  others' actions
+(or, if they do, this is exceptional).
+So why would anyone individually plan others' actions, knowing that she won't perform them?
+In many cases, 
+a plan for one agent's actions would need to be constrained by, and to constrain, plans for other agents' actions.
+When, for example,
+two people are about to move a heavy table from one room to the next,
+two separate plans for their individual actions would have to involve compatible timings, forces and routes.
+In principle,  each table-lifter might form an initial plan for her own actions only.
+Each could use her initial plan to make predictions about the  other's likely actions.
+She could then revise the plan for her own actions in the light of these predictions.
+This cycle of prediction and revision might be repeated.
+The plans might also involve conditional elements to accommodate unpredicted actions.
+So when an agent plans only her own actions,
+ she will sometimes need to rely on predictions about what others will do,
+ and these predictions will sometimes depend on what the agent plans to do.
+Another 
+ way to proceed
+	would be for the agent to plan all of the agents' actions,
+perhaps by first making an agent-neutral plan and then assigning roles.
+So the reason why some agents might each plan all of the agents' actions is that doing so provides one way to solve a coordination problem.
+We don't claim that this is the only or best way to coordinate.
+Our claim is just that some agents sometimes achieve coordination in part by means of each agent's individually planning all of the agents' actions.%
+\footnote{
+As an aside, we note that 
+\citet{tsai:2011_groop_effect} provide evidence indicating that planning for one's own and and others' actions happens not only in conscious practical reasoning but also in planning that we may not ordinarily be directly aware of.
+}
+
+Now recall Beatrice and Baldric (from Section \vref{sec:the_counterexample}).
+We said earlier that  they each see the other's actions  as constraints to work around or opportunities to exploit only.
+This is because each agent plans only her own actions.
+Neither agent plans for both of their actions and so neither conceives of both of their actions as parts of a single plan.
+In constructing a case that contrasts with Beatrice and Baldric's we shall rely on the distinction between
+a plan that includes both one's own and others' actions
 and 
-	forming a plan for  your own and others actions.
-%\footnote{
-%Perhaps it is impossible, strictly speaking, to plan others' actions. 
-%In speaking loosely about planning your own and others actions, we mean the activity of constructing an agent-neutral plan and assigning roles as described above.
-%}
-  Of course, in planning your only your own actions, it may be necessary to consider how others' actions will affect, and be affected by, your won.
-But in this case,  others' actions feature in your planning as constraints and opportunities.
-By contrast, forming a plan for your own and others' actions 
-	involves conceiving of your own and the others' actions as elements in a single plan directed to a single outcome.
+a plan that merely exploits, and is constrained by, predictions about the others' actions.
+This distinction matters independently of whether two such plans would result in different actions.
+It matters because 
+forming a plan for your own and others' actions 
+	involves changing perspective.
+That is, it involves conceiving of the others' actions not only as constraints and opportunities, but as on a par with your own actions insofar as your actions and theirs are elements in a single plan directed to a single outcome.
 
-Now that we have some background on agent-neutral plans,
-let us introduce a third and final case.
+In the case of ordinary, individual agency,
+there is a familiar and  important distinction between merely predicting one's own actions and planning them.%
+%
+\footnote{
+Even Velleman, who argues that `[d]eciding on an action … is just a peculiar way of predicting it' (\citeyear[p.\ 48]{Velleman:1985mc}), can accommodate a distinction between merely predicting and planning actions \citep[see][p.\ 195]{Velleman:2000fq}.
+%“How can one regard a proposition as true in such a way as to make it true?
+Well, when one accepts a proposition in response to its truth, one registers the influence of evidence and other reasons for belief, thereby manifesting an incli-nation to conform one’s acceptance to the facts.
+Accepting a proposition in such a way as to make it true would simply require a converse inclination, to conform the facts to one’s acceptance.
+And one can indeed be inclined to conform the facts to one’s acceptance, if the proposition accepted is about one’s own behavior.
+One need only be inclined to do what one accepts that one will do.
+If one has this inclination, then accepting that one will do something can be a way of making this proposition true, and it can therefore be an attempt at accepting the truth.”  (195)
+}
+%
+It is possible and perhaps sometimes useful to make predictions about one's future actions, of course.
+But to rely on predictions about one's own actions in planning can involve a kind of estrangement from oneself \citep[compare][]{Moran:2001hr}.
+We have just been arguing that 
+in the case of shared agency,
+the distinction between predicting and planning actions also applies to the actions of agents other than oneself.
+
+Applying this distinction in cases of shared agency requires additional subtlety, of course.
+After all,
+one agent cannot rationally include another's actions in a plan she knows she will act on unless she could predict the other's actions.
+(Or unless she knows the other agent will act directly on the first's intentions---see \citet{roth_shared_agency}.) 
+This fact should not obscure the distinct ways in which one conceives of another agent's actions in planning and predicting them, however.
+For, as we have argued, in some cases planning another agent's actions does not require actually predicting them.
+In some cases, planning another's actions does not depend on having predicted them but rather provides a ground for prediction.
+This is why there is a difference in perspective between merely predicting another's actions and conceiving of them as on a par with one's own in being elements of a single plan.
+
+Now that we have some background on how some agents might each individually plan all of the agents' actions,
+we can introduce a third and final case.
 This case needs to be as similar as possible to Beatrice and Baldric's while involving shared agency.
@@
 In this case it happens that both would be prepared to go along with the plan.
-Caitlin then knowledgeably predicts that Ciaran, who has similar planning abilities and has been approaching their task in a similar way, will have made an identical or compatible%
-\footnote{
-The notion of compatible is defined in footnote \vref{fn:df_compatible}.
-}
-%
- plan which, like hers, is acceptable to both of them.
+Caitlin then knowledgeably predicts that Ciaran, who has similar planning abilities and has been approaching their task in a similar way, will have made a sufficiently similar plan.
 Finally, Caitlin attempts to carry out her part in her plan; and Ciaran does likewise.
@@
 And, like Beatrice and Baldric, they are unconcerned with each other's success except insofar as their own success depends on it.
-But Caitlin and Ciaran nevertheless make use of the fact that they have to perform a single task together by each constructing a single plan covering both of their actions and then carrying out their parts in these plans.
+But Caitlin and Ciaran nevertheless make use of the fact that they have to perform a single task together by each constructing a single plan including both of their actions and then carrying out their parts in these plans.
 So why does Caitlin and Ciaran's case, but not Beatrice and Baldric's, plausibly involve shared agency?
@@
 	 at some stage of the planning which rationally guides and coordinates their actions,
-	Caitlin and Ciaran each conceive of their actions as part of a single plan directed to achieving a single outcome.
+	Caitlin and Ciaran each conceive of both of their actions as parts of a single plan directed to achieving a single outcome.
 
@@
 Our counterexample draws on the same planning resources Bratman's account draws on, with just one addition.
-The addition is the idea that agents can conceive, or fail to conceive, of their actions as part of a single plan. 
-In at least some cases, shared agency differs from individual agency in part because shared agency involves changing perspective so that one’s own and others' actions are conceived of as parts of a single plan.
+The addition is the idea that an agent 
+	might plan  her own actions only 
+	or else 
+	might plan both her own and others' actions.
+Agents often conceive of others' actions as external to their own plans, and so as either entirely irrelevant or constraints and opportunities only.
+But shared agency sometimes or always involves agents who each individually plan all of the agents' actions and so conceive of their own and the others' actions  as parts of a single plan.
+In at least some cases, shared agency can be distinguished from individual agency in part because it involves changing perspective in this way.
 
@@
 Beatrice and Baldric's case is a counterexample to the claim that for any J, Bratman's conditions, (1)--(3) on page \pageref{quote:bratman_account}, are collectively sufficient for us to have a shared intention that we J.
-(We called this the `strong claim' earlier, \vpageref{strong_claim}.)
+(We called this the `strong claim'  \vpageref{strong_claim}.)
 If this counterexample were an isolated case, it might be tempting to suppose that it can provide little insight into shared agency.

### 2012-12-10 counterexample to bratman on shared intention/a counterexample to bratman on shared intention.tex
in progress

@@
 
+%screws up word count for some reason:
 \bibliographystyle{$HOME/Documents/submissions/mynewapa} %apalike
@@
 We also suggest a way of revising Bratman's account  that is consistent with his general approach.
-In some or all cases, shared agency differs from individual agency in part because shared agency involves changing perspective so that one’s own and other’s actions are conceived of as parts of a single plan.
+In some or all cases, shared agency differs from individual agency in part because shared agency involves 
+changing perspective 
+to conceive of your own and another's actions  as parts of a single plan.
 
@@
 the contrast cases are often used to anchor intuitions in theorising about shared agency.
-In the next section we shall further support our claim by contrasting Beatrice and Baldric's case with a further case, one which is as similar as possible and which does seem to involve shared agency.
+In the next sections we shall further support our claim by contrasting Beatrice and Baldric's case with a further case, one which is as similar as possible and which does seem to involve shared agency.
 The fact that Beatrice and Baldric's case contrasts with this new case will support the claim that Beatrice and Baldric lack shared intention.
-This is the aim of the following section.
+This is the aim of the following sections.
 A further motive for introducing this new contrast case is that it provides an indication concerning what is missing from Bratman's account, and so may help us to understand why his conditions are not sufficient for shared intention.
@@
 To show that they do not,
-we shall contrast Beatrice and Baldric's case with a third case that is as similar as possible but does involve shared intention.
+we shall contrast Beatrice and Baldric's case with a third case that is as similar as possible but does involve shared intention (see Section \vref{sec:new_contrast_case}). 
+Before introducing this third case,
+we first need to introduce the notion of an agent-neutral plan and some background on planning.
 
@@
 
-First we need to introduce the notion of an agent-neutral plan and some  observations.
 We stipulate that a planning process, or a plan, is \emph{agent-neutral} just if it does not involve identifying any particular agents.  
@@
 If some agents each individually plan all of the agents' actions, 
-could they nevertheless know  that they will form plans that are sufficiently similar or jointly viable?%
+could they nevertheless know  that their various plans will be  sufficiently similar or jointly viable?%
 \footnote{
@@
  the agents' expertise and planning strategies 
- may mean that each could knowledgeably predict that they will form sufficiently similar agent-neutral plans.
+ may mean that each could knowledgeably predict that they will make sufficiently similar agent-neutral plans.
 And sometimes commonly known properties of the agents, such as their distribution in space or their sizes, mean that there is exactly one most obvious way of dividing up roles.
@@
 % it need not be irrational for the agents each individually to plan all of the agents' actions. 
-In any such case, the agents who each individually plan all of the agents' actions may form sufficiently similar plans and know that they will do so.
+In any such case, the agents who each individually plan all of the agents' actions may end up with sufficiently similar plans and know that they will do so.
 
@@
 	 are about to act together in order to bring about some outcome.
-They know (we stipulate) that if they each individually plan all of the agents' actions, they will form sufficiently similar plans.
+They know (we stipulate) that if they each individually plan all of the agents' actions, they will end up with sufficiently similar plans.
 Each agent also knows enough about herself and the others to be able to determine, without communicating, that the resulting plans will be acceptable to everyone. 
@@
 two separate plans for their individual actions would have to involve compatible timings, forces and routes.
-In principle,  each table-lifter might form an initial plan for her own actions only.
+In principle,  each table-lifter might make an initial plan for her own actions only.
 Each could use her initial plan to make predictions about the  other's likely actions.
@@
 As an aside, we note that 
-\citet{tsai:2011_groop_effect} provide evidence indicating that planning for one's own and and others' actions happens not only in conscious practical reasoning but also in planning that we may not ordinarily be directly aware of.
+\citet{tsai:2011_groop_effect} provide evidence indicating that planning for your own and and others' actions happens not only in conscious practical reasoning but also in planning that you may not ordinarily be directly aware of.
 }
@@
 Now recall Beatrice and Baldric (from Section \vref{sec:the_counterexample}).
-We said earlier that  they each see the other's actions  as constraints to work around or opportunities to exploit only.
-This is because each agent plans only her own actions.
-Neither agent plans for both of their actions and so neither conceives of both of their actions as parts of a single plan.
+We said earlier that  they each see the other's intentions and actions  as constraints to work around or opportunities to exploit only.
+This is because neither agent plans all of their actions and so neither conceives of both her own and the other's actions as parts of a single plan.
 In constructing a case that contrasts with Beatrice and Baldric's we shall rely on the distinction between
-a plan that includes both one's own and others' actions
+a plan that includes both your own and another's actions
 and 
-a plan that merely exploits, and is constrained by, predictions about the others' actions.
+a plan that merely exploits, and is constrained by, predictions about the other's actions.
 This distinction matters independently of whether two such plans would result in different actions.
 It matters because 
-forming a plan for your own and others' actions 
+making a plan for your own and another's actions 
 	involves changing perspective.
-That is, it involves conceiving of the others' actions not only as constraints and opportunities, but as on a par with your own actions insofar as your actions and theirs are elements in a single plan directed to a single outcome.
+That is, it involves conceiving of the other's actions not only as constraints and opportunities, but as on a par with your own actions insofar as your actions and their actions are elements in a single plan directed to a single outcome.
 
-In the case of ordinary, individual agency,
-there is a familiar and  important distinction between merely predicting one's own actions and planning them.%
+Note that our key distinction is not between merely predicting and planning another agent's actions.
+Beatrice and Baldric each intend to make the cross hit the red square (that is, to J${_1}$) in accordance with, and in part because of, each other's intentions.
+So when Beatrice predicts what Baldric will do, she can rely on facts about his intentions.
+She may even make predictions by planning his actions,
+and Baldric may likewise predict Beatrice's actions by planning them.
+So we can suppose that each plans the other's actions as well as separately planning her own actions,
+and that these two planning activities constrain each other.
+But---and this is our key distinction---neither makes a single plan for both her own and the other's actions.
+
+\section{One Plan or Two?}
+The distinction between making a single plan and making two separate plans may initially seem too subtle to be useful, 
+especially as 
+the two separate plans may be interdependent.
+Further, if two separate plans could both be executed (not just each plan but both together), then they can be combined into a single plan.
+The combined plan is simply to implement each of the two formerly separate plans. 
+Given these apparent grounds for scepticism,
+in this section we shall explain why the distinction between making a single plan and making two separate plans matters.
+
+***First,
+the fact that two separate \emph{plans} can be combined doesn't undermine the distinction between \emph{making} a single plan and \emph{making} two separate plans.
+To see why consider structure ...
+[... but the really deep point is not about structure: it's about keeping plans separate meaning adopting a predictive rather than a deliberative perspective towards the actions ...]
+
+Let us step back from shared agency  
+ to consider how 
+ the distinction between making a single plan and making two separate plans
+  applies in cases of ordinary, individual agency.
+Suppose you have to organise two events, a workshop and a wedding.
+If the two events were largely independent of each other,
+it would be simplest to make two separate plans for them.
+But in this case there are all kinds of constraints linking the two events.
+Some participants at the workshop are also wedding guests,
+and it happens that many choices of transport, entertainment and catering for one event constrain possible choices for the other event.
+In principle you could handle these by making two separate plans where, at any stage of your planning, making one plan involves treating the other plan as providing constraints to work around and opportunities to exploit.
+But envisaging the complex interdependence of these two plans  
+might reasonably incline you to make a single plan for the wedding and workshop.
+Making a single plan can be simpler because it allows for structuring the requirements differently.
+For instance, if you have two separate plans, you will need subplans for transport in each plan and so are forced to divide the problem of transporting participants into the problem of transporting them for the workshop and the problem of transporting them for the wedding.
+Having a single plan for both events allows you to treat transport as a single problem (but it does not require you to do so, of course).
+As this indicates, the distinction between making a single plan and making two separate but interdependent plans is not merely notional.
+Agents can have reasons for making a single plan rather than two separate ones.
+
+This claim needs qualification, however.
+It might be held that single agents always combine all plans for their own actions ... unity of their plans ... leads to the Moran point ***
+
+There is a familiar and  important distinction between merely predicting your own actions and planning them.%
 %
 \footnote{
-Even Velleman, who argues that `[d]eciding on an action … is just a peculiar way of predicting it' (\citeyear[p.\ 48]{Velleman:1985mc}), can accommodate a distinction between merely predicting and planning actions \citep[see][p.\ 195]{Velleman:2000fq}.
+Even Velleman, who argues that `[d]eciding on an action … is just a peculiar way of predicting it' (\citeyear[p.\ 48]{Velleman:1985mc}), can accommodate a distinction between merely predicting and planning actions.
+In his terms, we might say that 
+planning involves not merely predicting but also being inclined to make the prediction true \citep[see][p.\ 195]{Velleman:2000fq}.
 %“How can one regard a proposition as true in such a way as to make it true?
-Well, when one accepts a proposition in response to its truth, one registers the influence of evidence and other reasons for belief, thereby manifesting an incli-nation to conform one’s acceptance to the facts.
+Well, when one accepts a proposition in response to its truth, one registers the influence of evidence and other reasons for belief, thereby manifesting an inclination to conform one’s acceptance to the facts.
 Accepting a proposition in such a way as to make it true would simply require a converse inclination, to conform the facts to one’s acceptance.
@@
 %
-It is possible and perhaps sometimes useful to make predictions about one's future actions, of course.
-But to rely on predictions about one's own actions in planning can involve a kind of estrangement from oneself \citep[compare][]{Moran:2001hr}.
-We have just been arguing that 
-in the case of shared agency,
-the distinction between predicting and planning actions also applies to the actions of agents other than oneself.
+It is possible and perhaps sometimes useful to make predictions about your future actions, of course.
+%*nb 'your ... oneself' issue in this sentence:
+But to rely on mere predictions about your own actions in planning can involve a kind of estrangement from oneself \citep[compare][]{Moran:2001hr}.
+
+We are arguing that 
+something similar applies in the case of shared agency, 
+except that here the distinction between predicting and planning actions also applies to others' actions.
 
@@
 After all,
-one agent cannot rationally include another's actions in a plan she knows she will act on unless she could predict the other's actions.
-(Or unless she knows the other agent will act directly on the first's intentions---see \citet{roth_shared_agency}.) 
+you cannot rationally include another's actions in a plan you know you will act on unless you could predict the other's actions.%
+%
+\footnote{
+This statement requires qualification if, as \citet{roth_shared_agency} argues, 
+another agent might act directly on your intentions. 
+As the cases we are considering involve agents in symmetric roles, we can ignore that possibility here.
+We acknowledge, however, that Roth's idea has inspired the less radical view defended in this section.
+}
+%
 This fact should not obscure the distinct ways in which one conceives of another agent's actions in planning and predicting them, however.
@@
 In some cases, planning another's actions does not depend on having predicted them but rather provides a ground for prediction.
-This is why there is a difference in perspective between merely predicting another's actions and conceiving of them as on a par with one's own in being elements of a single plan.
+This is why 
+different perspectives are involved in 
+	merely predicting another's actions 
+	and 
+	conceiving of her actions and  your own actions as elements of a single plan.
 
+\section{A New Contrast Case}
+\label{sec:new_contrast_case}
 Now that we have some background on how some agents might each individually plan all of the agents' actions,
@@
 \item 	there is a single outcome, G,
-	and two or more agents each have a plan for G which specifies a role for all of the agents;	
+	and two or more agents each have a plan for G which specifies  roles for all of the agents;	
 \item 	these plans are identical or compatible,%
@@
 In walking to a metro station together,
-two friends may each form and act on a plan that specifies some details of both of their actions.
+two friends may each make and act on a plan that specifies some details of both of their actions.
 This is enough for their activity to be distributed in the above sense.
@@
 In many or all cases, shared agency requires changing perspective 
-to conceive of one's own and others' intentions and actions as parts of a single plan.
+to conceive of your own and another's intentions and actions as parts of a single plan.
 

### 2012-12-10 counterexample to bratman on shared intention/a counterexample to bratman on shared intention.tex
in progress

@@
 The fact that Beatrice and Baldric's case contrasts with this new case will support the claim that Beatrice and Baldric lack shared intention.
-This is the aim of the following sections.
-A further motive for introducing this new contrast case is that it provides an indication concerning what is missing from Bratman's account, and so may help us to understand why his conditions are not sufficient for shared intention.
+This is the proximal aim of the following sections.
+A more distal aim 
+is to understand why Bratman's conditions are not sufficient for shared intention
+and, relatedly, to provide materials for fixing his account.
 
@@
 we shall contrast Beatrice and Baldric's case with a third case that is as similar as possible but does involve shared intention (see Section \vref{sec:new_contrast_case}). 
-Before introducing this third case,
-we first need to introduce the notion of an agent-neutral plan and some background on planning.
+Before we can introduce this third case,
+we first need some background on planning.
 
@@
 while a third might mix some salad.
-In so planning, each housemate is thinking about what is to be done and not what she herself will do.  
+In so planning, each housemate is thinking about what 
+agents in their situation should do 
+and not what she herself or anyone in particular will do.  
 %(This is an idealisation, of course.)
@@
 We invoke agent-neutral plans only as an aid to showing that the notion of some 
-agents each individually planning all of the agents' actions is coherent.
+agents each individually planning all of the agents' actions is theoretically coherent.
 
-If some agents each individually plan all of the agents' actions, 
-could they nevertheless know  that their various plans will be  sufficiently similar or jointly viable?%
+But why would some agents ever each individually plan all of the agents' actions?
+One might suppose that this would occur only when each agent thinks that she in charge, or is fantasizing about controlling the others.
+If you know you aren't the boss and aren't going to decide what another agent will do, why  make a plan for that agent's actions as well as your own?
+Why not just plan your own actions?
+As we shall now explain, 
+one reason why some agents might each plan all of the agents' actions 
+	even when there is manifestly no leader and the agents all have the same status
+	is that doing so provides one way to solve a coordination problem.
+%%having agents each individually plan all of the agents' actions
+%can be a way of achieving coordination even where there is no leader and the agents all have the same status.
+
+%How could  this work?
+Sometimes a plan for your own actions needs to be constrained by, and to constrain, plans for other agents' actions.
+When, for example,
+you and someone else are about to move a heavy table from one room to another,
+there are constraints 
+on relations between
+	your actions
+	and
+	the other's actions.
+These relational constraints concern timings, forces, choices of route and more besides.
+Now meeting such relational constraints does not in principle require you to plan actions other than your own.
+In principle,  you  might make an initial plan for your own actions only.
+You could use you initial plan to make predictions about the  other's likely actions,
+and then revise this initial plan for your actions in the light of these predictions about the other's actions.
+This cycle of prediction and revision might be repeated.
+Your plan might also involve conditional elements to accommodate unpredicted actions.
+So if you plan your own actions only,
+ you will need to rely on predictions about what the other will do,
+ and these predictions will sometimes depend on what you plan to do.
+But there is another way to proceed,
+one which avoids the need for a plan that depends on predictions which in turn depend on the plan.
+For you can use the same planning processes which enable you to meet constraints on relations between two of your own actions
+	in order to meet constraints on relations between your own and the other's actions.
+That is, you could make a single plan for both your own and the other's actions.
+(You might do this by first making an agent-neutral plan and then assigning roles.)
+Making a single plan for your actions and the other's actions is a way of using your planning abilities to meet relational constraints.
+Of course, 
+this is only rational  if you are in a position to know that your plan for the other's actions will be sufficiently similar to her own plan for her actions.
+And it is unlikely that you could know this if 
+	you and the other agent differed too much in expertise,
+	if there were two or more equally good ways to achieve the outcome you are planning,
+	or if there were two or more equally good ways to assign roles.
+But suppose that you and the other person have similar expertise in moving furniture, 
+	that there is obviously exactly one most optimal route,
+	 and that your positions in space manifestly determine which ends of the table you should each hold and so which role each of you should play. 
+Then it is likely that you could  be in a position to know,
+	even without communication,
+	that your plan for the other agent's actions will nearly enough coincide with her own plan for her actions.
+So  making a single plan for your actions and the other's actions can be a way of 
+	working out how to 
+	meet constraints on the relation between your actions and hers and so of achieving coordination.  
+(Note that, as already mentioned,  we don't claim that this  is the only or best way to achieve coordination.)	
+And, clearly, this applies equally to the 
+ other person you are lifting the table with.
+ She  could work out how to coordinate her actions with yours by likewise making a single plan for her actions and yours.%
 \footnote{
-\label{fn:df_jointly_viable}
-Suppose that, for some outcome, two or more agents each have a plan for the realisation of that outcome. 
-%(These plans may, but need not, specify roles for all of the agents; but the plans must be agent-specific, not agent-neutral.)
-We stipulate that these plans are  \emph{jointly viable} just if
-the plan which involves each agent performing the roles she is assigned in her own plan is acceptable to all the agents and a viable means of realising the outcome.
-For an example of jointly viable but significantly different plans, suppose that our task is to transport some apples and pears to a warehouse. 
-My plan involves you taking care of the apples while I put the pears into the yellow crates and load them into the van through the side door.
-If your plan involves me taking care of the pears while you put the apples into the green crates and load them into the van through the rear doors, then our plans are compatible.
-Consider also a case in which our task is to press a button simultaneously. 
-If your plan is that each of us will to start to act in exactly 60 seconds whereas my plan specifies that you are the leader and we will start to act when you do, then our plans are jointly viable.
+As an aside, we note that 
+\citet{tsai:2011_groop_effect} provide evidence indicating that planning for your own and and others' actions happens not only in conscious practical reasoning but also in planning that you may not ordinarily be directly aware of.
 }
-%Where the agents could bring about an outcome in several different, equally good ways,
-%having some agents each individually plan all of the agents' actions is probably not useful.
-Consider housemates
-	making a pizza together, where everyone knows roughly the same recipe and differences are unlikely to matter.
-Or consider two agents 
- moving  a table from one room to another together, where there is an obviously optimal route and the agents' positions manifestly determine which ends of the tables they should hold.
-For these outcomes and situations,
- the agents' expertise and planning strategies 
- may mean that each could knowledgeably predict that they will make sufficiently similar agent-neutral plans.
-And sometimes commonly known properties of the agents, such as their distribution in space or their sizes, mean that there is exactly one most obvious way of dividing up roles.
-%If  each agent also  knows enough about herself and the others to be able to determine, without communicating, whether her plan and role assignments will be acceptable to everyone, then
-% it need not be irrational for the agents each individually to plan all of the agents' actions. 
-In any such case, the agents who each individually plan all of the agents' actions may end up with sufficiently similar plans and know that they will do so.
 
-Agents who will collaborate in some way, such as by making a pizza together, can coordinate their actions by each individually planning all of the agents' actions.
-Suppose some agents 
-	 are about to act together in order to bring about some outcome.
-They know (we stipulate) that if they each individually plan all of the agents' actions, they will end up with sufficiently similar plans.
-Each agent also knows enough about herself and the others to be able to determine, without communicating, that the resulting plans will be acceptable to everyone. 
-And all of this is common knowledge to the agents.  
-Then these agents 
-might individually plan all of the agents' actions,
-and each might rationally perform her part in her plan, knowing that the others will do likewise.
-In this way, 
- some agents' individually planning all of the agents' actions can rationally coordinate their plans and actions.%
+The question was why some agents would each individually plan all of the agents' actions.
+We have just argued that  
+one reason is that
+doing this would sometimes enable the agents to solve a coordination problem.
+Agents who will collaborate in an activity, such as moving a table or making a pizza, 
+can sometimes individually plan all of the agents' actions,
+knowing that each agent's plan is sufficiently similar to the other agents' plans.
+And sometimes each can rationally perform her part in her plan for all of the agent's actions, knowing that the others will do likewise.%
 \footnote{
@@
 But this shared agency is a consequence of, not a requirement on, their individually planning all of the agents' actions.
-}
-
-But why would some agents each individually plan all of the agents' actions?
-After all, no agent gets  to perform  others' actions
-(or, if they do, this is exceptional).
-So why would anyone individually plan others' actions, knowing that she won't perform them?
-In many cases, 
-a plan for one agent's actions would need to be constrained by, and to constrain, plans for other agents' actions.
-When, for example,
-two people are about to move a heavy table from one room to the next,
-two separate plans for their individual actions would have to involve compatible timings, forces and routes.
-In principle,  each table-lifter might make an initial plan for her own actions only.
-Each could use her initial plan to make predictions about the  other's likely actions.
-She could then revise the plan for her own actions in the light of these predictions.
-This cycle of prediction and revision might be repeated.
-The plans might also involve conditional elements to accommodate unpredicted actions.
-So when an agent plans only her own actions,
- she will sometimes need to rely on predictions about what others will do,
- and these predictions will sometimes depend on what the agent plans to do.
-Another 
- way to proceed
-	would be for the agent to plan all of the agents' actions,
-perhaps by first making an agent-neutral plan and then assigning roles.
-So the reason why some agents might each plan all of the agents' actions is that doing so provides one way to solve a coordination problem.
-We don't claim that this is the only or best way to coordinate.
-Our claim is just that some agents sometimes achieve coordination in part by means of each agent's individually planning all of the agents' actions.%
-\footnote{
-As an aside, we note that 
-\citet{tsai:2011_groop_effect} provide evidence indicating that planning for your own and and others' actions happens not only in conscious practical reasoning but also in planning that you may not ordinarily be directly aware of.
 }
@@
 
-\section{One Plan or Two?}
+\section{Why Making One Plan or Two Matters}
 The distinction between making a single plan and making two separate plans may initially seem too subtle to be useful, 

### 2012-12-10 intention motor representation joint action (paper 3)/conf budapest philosophy 2012-11/shared agency and motor representation.tex
in progress

@@
 The first example shows that the difference between shared agency and parallel individual agency can’t just be that the resulting actions have a common effect because merely parallel actions can have common effects too.
-And the second example shows that the difference can’t just be a matter of coordination, because people who are merely happen to be walking side by side each other also need to coordinate their actions in order to avoid colliding.
+And the second example shows that the difference can’t just be a matter of coordination, because people who merely happen to be walking side by side each other also need to coordinate their actions in order to avoid colliding.
 Apparently, then, systematically distinguishing cases of shared agency from cases of parallel but merely individual agency  requires reflection on agents' mental states.

### 2012-12-12 counterexample to bratman on shared intention/a counterexample to bratman on shared intention.tex
working draft done

@@
 %\tableofcontents
-\title{}
+
+\begin{center}
+Status: unsubmitted draft, work in progress
+\end{center}
 
@@
 
-The first contrast case, Gilbert's, shows that the difference can’t be just  a matter of coordination because people who merely happen to be walking side by side each other also need to coordinate their actions in order to avoid colliding.
+The first contrast case, Gilbert's, shows that the difference can’t be just  a matter of coordination because people who merely happen to be walking side by side  also need to coordinate their actions in order to avoid colliding.
 Note also that in both cases each individual's walking is intentional, so our intentionally walking together cannot be  only a matter of our each intentionally walking.
@@
 \footnote{
-Many philosophers hold that a notion of shared intention is useful for understanding acting together. 
-Compare \citet[p.\ 5]{Gilbert:2006wr}: `I take a collective action to involve a collective intention.'  See also  
+On the idea that a notion of shared intention is useful for understanding acting together, compare  
 	\citet[p.\ 381]{Carpenter:2009wq}, 
@@
 	\citet{Kutz:2000si}, 
-	\citet[p.\ 117]{rakoczy_pretend_2006} and 
-	\citet{Tollefsen:2005vh}.
+	\citet[p.\ 175]{pacherie_framing_2011},
+	\citet[p.\ 117]{rakoczy_pretend_2006}, 
+	\citet{Tollefsen:2005vh},
+and \citet[p.\ 5]{Gilbert:2006wr}: `I take a collective action to involve a collective intention.'  
 	}
@@
 
-To show that  Bratman's conditions, (1)--(3), are not in fact sufficient for shared intention we need 
+To show that  Bratman's conditions, (1)--(3) \vpageref{quote:bratman_account}, are not in fact sufficient for shared intention we need 
 a case where the conditions are met and we lack a shared intention.
@@
 Now suppose 
-	that, for some J, Bratman's conditions, (1)--(3) above, were met in cases B and C alike.
+	that, for some J, Bratman's conditions, (1)--(3), were met in cases B and C alike.
 Suppose also that in each case, B and C, the structure of intention and knowledge mention specified by Bratman's conditions 
@@
 
-Several preliminaries are necessary for the construction of our counterexample. 
-These preliminaries might easily give the impression that our counterexample depends on an artificial settings. 
-However, having introduced the primary counterexample in an artificial setting, we will  go on to show that counterexamples can also be constructed for mundane activities including walking together.
+%Several preliminaries are necessary for the construction of our counterexample. 
+%These preliminaries might easily give the impression that our counterexample depends on an artificial settings. 
+%However, having introduced the primary counterexample in an artificial setting, we will  go on to show that counterexamples can also be constructed for mundane activities including walking together.
 
@@
 
-Consider the contents of the intentions concerning our J-ing in the above clauses, (1)--(3). 
+Consider the contents of the intentions concerning our J-ing in conditions (1)--(3) \vpageref{quote:bratman_account}. 
 What sort of activity can you intend when you intend that we J?
-We cannot restrict possible values of J to activities which involve shared agency.
+Possible values of J cannot be restricted to activities which involve shared agency.
 In imposing any such restriction we would be assuming the very notion that an account of shared intention is supposed to illuminate. 
@@
 %We must therefore avoid tacitly appealing to this distinction by  restricting possible values of J to those which involve shared agency. 
-Rather the above  conditions, (1)--(3), must be sufficient for shared intention even for some values of J which are `neutral with respect to shared intentionality'.%
+Rather the above  conditions, (1)--(3) \vpageref{quote:bratman_account}, must be sufficient for shared intention even for some values of J which are `neutral with respect to shared intentionality.'%
 \footnote{
- \citet[p.\ 147]{Bratman:1999fr}.
+See \citet[p.\ 147]{Bratman:1999fr}.
  This refines Bratman's earlier view that some admissable values of J are cooperatively neutral 
@@
 And it's even possible to desire or intend to be involved in an act of this type.
-A gangster might intend to go to New York with someone without having yet decided whether they will do this cooperatively or coercively.
+As Bratman suggests,
+a gangster might intend to go to New York with someone without having yet come to a view on whether this will be cooperative or coercive.
 
@@
 \end{quote}
-In defining unshared intention we have used conditions exactly like Bratman's sufficient conditions for shared intention except that Bratman's conditions have J$_1$$=$J$_2$.
+In defining unshared intention we have used conditions almost exactly like Bratman's sufficient conditions for shared intention except that Bratman's conditions have J$_1$$=$J$_2$.
 At this point it might be natural for readers to suppose that agents could not have unshared intentions,
@@
 Each agent's subplans involve manipulating his or her own controls,
-and the successes each seeks in doing this depends on the other successfully carrying out their subplans.
+and the success each seeks in doing this depends on the other successfully carrying out their subplans.
 So Ayesha can rationally intend that they, Ayesha and Ahmed, make the cross hit the red square in accordance with and because of their intentions and meshing subplans of them.
@@
 	%specified in conditions (1)--(3) on page \pageref{quote:bratman_account}
-	while leaving open whether it constitutes a shared intention) result in the two pairs performing same actions in the same way.
-That is, Beatrice reasons about Baldric much as Ayesha reasons about Ahmed and Beatrice and does what Ayesha does; and likewise for Baldric and Ahmed.
+	while leaving open whether it constitutes a shared intention) result in the two pairs performing the same actions in the same way.
+That is, Beatrice reasons about Baldric much as Ayesha reasons about Ahmed, and Beatrice does what Ayesha does; and likewise for Baldric and Ahmed.
 
@@
 This claim follows from the similarities of the two cases.
-The only difference is that Beatrice and Baldric happen to be assigned the same task, whereas Ayesha and Ahmed are not.
+The only difference is that Beatrice and Baldric happen to have same task, whereas Ayesha and Ahmed have different tasks.
 And neither Beatrice nor Baldric makes use of the fact that they have the same task. 
-(This is not due to ignorance: it's just how they choose to approach their tasks.)
+%(This is not due to ignorance: it's just how they approach their tasks.)
 So if we consider  how  
@@
 But doesn't shared intention require at least this much,
-that there be a single activity  about which each the agent involved has an intention? 
+that there be a single activity  about which each agent involved has an intention? 
 This might reasonably be doubted by those who, like Bratman, reject the Simple View according to which when an individual intentionally J-s she has an intention concerning her J-ing \citep{Bratman:1984jr}.
@@
 Can we then appeal directly to intuition to show that Ayesha and Ahmed lack a shared intention?
-Ayesha sees Ahmed's actions as constraints on her own, or else as opportunities.
-She exploits Ahmed's intentions for her own ends.
+Ayesha sees Ahmed's actions as constraints to work around or  opportunities to exploit.
+%She exploits Ahmed's intentions for her own ends.
 Of course the situation is reciprocal: Ahmed exploits Ayesha in equal measure.
@@
 the contrast cases are often used to anchor intuitions in theorising about shared agency.
-In the next sections we shall further support our claim by contrasting Beatrice and Baldric's case with a further case, one which is as similar as possible and which does seem to involve shared agency.
-The fact that Beatrice and Baldric's case contrasts with this new case will support the claim that Beatrice and Baldric lack shared intention.
+In the following sections we shall further support our claim by contrasting Beatrice and Baldric's case with a further case, one which is as similar as possible and which does seem to involve shared agency.
+The way that Beatrice and Baldric's case contrasts with this new case will support the claim that Beatrice and Baldric lack shared intention.
 This is the proximal aim of the following sections.
@@
 
-\section{Individually planning all of  the agents' actions}
+\section{Individually planning all  the agents' actions}
 \label{sec:distributed_plan}
@@
 Where there is an outcome and some agents each individually,
-without discussion or communication,
+without discussion, communication or prior arrangement,
 plan for the outcome 
-and where each agent has a single plan specifying roles for herself and all of the other agents,
+and where each agent makes a single plan specifying roles for herself and all the other agents,
 %When there is an outcome and some agents each individually,
@@
 %construct an agent-neutral plan for this outcome and then assign roles to themselves and the other agents,
-we shall say that these agents each \emph{individually plan all of the agents' actions}.%
+we shall say that these agents each 
+\label{df:plan_all_actions}
+\emph{individually make a plan for all the agents' actions}.%
 \footnote{
 This phrase is ambiguous (in a way we know  no way concise of avoiding).
-The idea is that each agent has a single plan in which she and  all other agents are assigned roles.
+The idea is that each agent makes a single plan in which she and  all other agents are assigned roles.
 There is no requirement that any agent's plan must specify all the actions which the agents between them will perform.
 }
-Agents can do this by first constructing agent-neutral plans and then assigning roles to themselves and the other agents.
-So, to return to our earlier example, 
+%Agents can do this by first constructing agent-neutral plans and then assigning roles to themselves and the other agents.
+
+To illustrate by returning to our earlier example, 
  if the housemates, having decided to make a pizza, had each individually devised  agent-neutral plans for the construction of the pizza and then each individually assigned roles in their own plans to themselves and the others,
-then they would have each individually planed all of the agents' actions.
+then they would have each individually made a plan for all the agents' actions.%
+\footnote{
+Note that, for all we have said,  agent-neutral plans might not be involved in all cases where some agents each individually make a plan for plan all the agents' actions.
+We invoke agent-neutral plans only as an aid to showing that the notion of some 
+agents each individually making a plan for all the agents' actions is theoretically coherent.
+}
+%
 
-For all we have said,  agent-neutral plans might not be involved in all cases where some agents each individually plan all of the agents' actions.
-We invoke agent-neutral plans only as an aid to showing that the notion of some 
-agents each individually planning all of the agents' actions is theoretically coherent.
-
-But why would some agents ever each individually plan all of the agents' actions?
+But why would some agents ever each individually make a plan for all the agents' actions?
 One might suppose that this would occur only when each agent thinks that she in charge, or is fantasizing about controlling the others.
-If you know you aren't the boss and aren't going to decide what another agent will do, why  make a plan for that agent's actions as well as your own?
-Why not just plan your own actions?
+Fantasy aside, 
+if you know you aren't in charge, why  make a plan for your own and another agent's actions?
+Why not plan your own actions only?
 As we shall now explain, 
-one reason why some agents might each plan all of the agents' actions 
-	even when there is manifestly no leader and the agents all have the same status
+one reason why some agents might each make a plan for all  the agents' actions 
+	even when there is manifestly no one in charge and the agents all have the same status
 	is that doing so provides one way to solve a coordination problem.
-%%having agents each individually plan all of the agents' actions
+%%having agents each individually make a plan for all the agents' actions
 %can be a way of achieving coordination even where there is no leader and the agents all have the same status.
@@
 In principle,  you  might make an initial plan for your own actions only.
-You could use you initial plan to make predictions about the  other's likely actions,
+You could use your initial plan to make predictions about the  other's likely actions,
 and then revise this initial plan for your actions in the light of these predictions about the other's actions.
@@
  and these predictions will sometimes depend on what you plan to do.
-But there is another way to proceed,
-one which avoids the need for a plan that depends on predictions which in turn depend on the plan.
-For you can use the same planning processes which enable you to meet constraints on relations between two of your own actions
+But there is another way to proceed.
+%one which avoids the need for a plan that depends on predictions which in turn depend on the plan.
+You can use the same planning processes which enable you to meet constraints on relations between two of your own actions
 	in order to meet constraints on relations between your own and the other's actions.
@@
 (You might do this by first making an agent-neutral plan and then assigning roles.)
-Making a single plan for your actions and the other's actions is a way of using your planning abilities to meet relational constraints.
+
+Making a single plan for your own and the other's actions is a way of using your planning abilities to meet relational constraints.
 Of course, 
@@
 As an aside, we note that 
-\citet{tsai:2011_groop_effect} provide evidence indicating that planning for your own and and others' actions happens not only in conscious practical reasoning but also in planning that you may not ordinarily be directly aware of.
+\citet{tsai:2011_groop_effect} provide evidence indicating that making a plan for your own and and others' actions can happen not only in conscious practical reasoning but also in planning that you may not ordinarily be directly aware of.
 }
 
-The question was why some agents would each individually plan all of the agents' actions.
+The question was why some agents would each individually make a plan for all the agents' actions.
 We have just argued that  
@@
 Agents who will collaborate in an activity, such as moving a table or making a pizza, 
-can sometimes individually plan all of the agents' actions,
+can sometimes individually make a  plan for all the agents' actions,
 knowing that each agent's plan is sufficiently similar to the other agents' plans.
-And sometimes each can rationally perform her part in her plan for all of the agent's actions, knowing that the others will do likewise.%
+And sometimes each can rationally perform her part in her plan for all the agent's actions, knowing that the others will do likewise.%
 \footnote{
-Note that this coordination is achieved without presupposing that the agents exercise shared agency.
+As already mentioned, coordinating in this way does not require exercising shared agency because the planning can occur in parallel.
 Of course,
-	if some agents individually plan all of the agents' actions
+	if some agents individually make a plan for all the agents' actions
 	and then each rationally performs her part in her own plan,
 	the upshot might be an exercise of shared agency.
-But this shared agency is a consequence of, not a requirement on, their individually planning all of the agents' actions.
+But any such exercises of shared agency are a consequence of, not a requirement on, their individually making a plan for all the agents' actions.
 }
@@
 We said earlier that  they each see the other's intentions and actions  as constraints to work around or opportunities to exploit only.
-This is because neither agent plans all of their actions and so neither conceives of both her own and the other's actions as parts of a single plan.
+This is because neither agent makes a plan for both agents' actions and so neither conceives of both her own and the other's actions as parts of a single plan.%
+\footnote{
+We also stipulate that neither Ayesha nor Ahmed makes a plan for both of the agents' actions.
+(We have not shown that Ayesha and Ahmed could not have done this; we merely stipulate that, as is perhaps natural to assume anyway, they do not.)
+}
 In constructing a case that contrasts with Beatrice and Baldric's we shall rely on the distinction between
@@
 a plan that merely exploits, and is constrained by, predictions about the other's actions.
-This distinction matters independently of whether two such plans would result in different actions.
+This distinction matters independently of whether two such plans would result in different actions, as we shall argue in the next section.
 It matters because 
@@
 	involves changing perspective.
-That is, it involves conceiving of the other's actions not only as constraints and opportunities, but as on a par with your own actions insofar as your actions and their actions are elements in a single plan directed to a single outcome.
+That is, it involves conceiving of the other's actions not only as constraints to work around or opportunities to exploit, but as on a par with your own actions insofar as your actions and their actions are elements in a single plan directed to a single outcome.
 
@@
 \section{Why Making One Plan or Two Matters}
-The distinction between making a single plan and making two separate plans may initially seem too subtle to be useful, 
+The distinction between making a single plan and making two separate plans may initially seem too subtle to matter, 
 especially as 
-the two separate plans may be interdependent.
-Further, if two separate plans could both be executed (not just each plan but both together), then they can be combined into a single plan.
+the separate planning processes may constrain each other.
+Further, if any two separate plans can both be executed (not just each plan but both together), then they can be combined into a single plan.
 The combined plan is simply to implement each of the two formerly separate plans. 
-Given these apparent grounds for scepticism,
-in this section we shall explain why the distinction between making a single plan and making two separate plans matters.
+These reflections might misleadingly seem to 
+indicate that there is no significant difference between making a single plan and making two separate plans.
+However in this section we shall explain why the distinction between making a single plan and making two separate plans matters.
 
-***First,
-the fact that two separate \emph{plans} can be combined doesn't undermine the distinction between \emph{making} a single plan and \emph{making} two separate plans.
-To see why consider structure ...
-[... but the really deep point is not about structure: it's about keeping plans separate meaning adopting a predictive rather than a deliberative perspective towards the actions ...]
-
-Let us step back from shared agency  
- to consider how 
- the distinction between making a single plan and making two separate plans
-  applies in cases of ordinary, individual agency.
-Suppose you have to organise two events, a workshop and a wedding.
+As this distinction applies in cases of  shared and individual agency alike, we first focus on individual agency.
+It is true, as noted, that  two separate {plans} can be combined.
+But this doesn't undermine the distinction between {making} a single plan and {making} two separate plans.
+To see why not, 
+suppose you have to organise two events, a workshop and a wedding.
 If the two events were largely independent of each other,
@@
 and it happens that many choices of transport, entertainment and catering for one event constrain possible choices for the other event.
-In principle you could handle these by making two separate plans where, at any stage of your planning, making one plan involves treating the other plan as providing constraints to work around and opportunities to exploit.
-But envisaging the complex interdependence of these two plans  
-might reasonably incline you to make a single plan for the wedding and workshop.
-Making a single plan can be simpler because it allows for structuring the requirements differently.
-For instance, if you have two separate plans, you will need subplans for transport in each plan and so are forced to divide the problem of transporting participants into the problem of transporting them for the workshop and the problem of transporting them for the wedding.
-Having a single plan for both events allows you to treat transport as a single problem (but it does not require you to do so, of course).
-As this indicates, the distinction between making a single plan and making two separate but interdependent plans is not merely notional.
-Agents can have reasons for making a single plan rather than two separate ones.
+In principle you could handle this situation by making two separate plans where, at any stage of your planning, making one plan involves treating the other plan as providing constraints to work around and opportunities to exploit.
+But envisaging the complex interdependence of these two plans,
+you might reasonably be inclined to make a single plan for the wedding and workshop.
+Making a single plan can be simpler because it allows for greater flexibility in structuring subplans than would making separate plans for the workshop and wedding and then combining them.
+For instance, if you make separate plans for the wedding and workshop, you will need subplans for transport in each plan and so are forced to divide the problem of transporting participants into the problem of transporting them for the workshop and the problem of transporting them for the wedding.
+Making a single plan for both events allows you to treat transport as a single problem (but it does not require you to do so, of course).
+As this indicates, the difference between making a single plan and making two separate but interdependent plans is sometimes significant.
+People can have reasons for making a single plan rather than two separate ones.
 
-This claim needs qualification, however.
-It might be held that single agents always combine all plans for their own actions ... unity of their plans ... leads to the Moran point ***
-
-There is a familiar and  important distinction between merely predicting your own actions and planning them.%
+For our purposes it is necessary to focus on a further, less obvious way in which making a single plan can differ significantly from making two separate plans. 
+There is a familiar and  important distinction between planning  your own actions and merely predicting them.%
 %
@@
 %
-It is possible and perhaps sometimes useful to make predictions about your future actions, of course.
+It is possible and perhaps sometimes useful to make predictions about your future actions.
+%*people are typically very bad at predicting their own actions.
 %*nb 'your ... oneself' issue in this sentence:
-But to rely on mere predictions about your own actions in planning can involve a kind of estrangement from oneself \citep[compare][]{Moran:2001hr}.
+But to rely too much on merely predicting your own actions in planning can sometimes involve a kind of estrangement from oneself \citep[compare][]{Moran:2001hr}.
+Now imagine someone who has a strong commitment to keeping two or more areas of her life apart, so that she tries plan separately for each area of her life.
+When concerned with planning in one area, intentions and actions from other areas of her life feature in her planning only as 
+constraints to work around or opportunities to exploit.
+Of course, many or even all of these predictions might have their source in other episodes of planning.
+But 
+she systematically avoids conceiving of any two plans for different areas of her life as even potentially parts of a single, larger plan.
+So there is a kind of unity that she fails to conceive of the actions which make up her life as having.
+When planning for one area of her life, she adopts the perspective of an outsider on other areas of her life.
+%Echoing Moran, we might put the idea roughly by saying that she is constantly estranged from the actions making up one or another area of her life.
 
-We are arguing that 
-something similar applies in the case of shared agency, 
-except that here the distinction between predicting and planning actions also applies to others' actions.
+Our earlier discussion about making a plan for your own and others' actions 
+(see Section \vref{sec:distributed_plan})
+ indicates that 
+  the distinction between predicting and planning actions is also linked to how you conceive of others' actions, not just your own.
+To make a plan for your own and others' actions is to conceive of them as parts of a single plan and so as having a kind of unity.
+By contrast, 
+to make a plan for your own actions only would be to adopt the perspective of an outsider on the others' actions.
+And, similarly, to make a plan for the others' actions only would involve adopting the perspective of an outsider on your own actions.
+So even when you make a plan for others' actions,
+if you avoid conceiving of this plan as even potentially part of a larger plan involving your own actions as well,
+then
+at any point in time you will be adopting the perspective of an outsider on your actions or on their actions.
+A special feature of making a plan for your own and others' actions is that you do not adopt the perspective of an outsider with respect to any of these actions, yours or theirs.
+%When shared agency is not involved, it is probably right to take the perspective of an outsider on others' actions when not adopting their perspective.
+%When not involved in others' actions it is perhaps right to feel estranged from them.
+In at least some cases, shared agency may involve conceiving of both your own and others' actions not as constraints and opportunities but as elements of a single plan.
+%simultaneously not taking the perspective of an outsider on co-actors' actions.
 
-Applying this distinction in cases of shared agency requires additional subtlety, of course.
+The case of shared agency requires additional subtlety, of course.
 After all,
+for the examples we are considering in this paper,
 you cannot rationally include another's actions in a plan you know you will act on unless you could predict the other's actions.%
@@
 another agent might act directly on your intentions. 
-As the cases we are considering involve agents in symmetric roles, we can ignore that possibility here.
-We acknowledge, however, that Roth's idea has inspired the less radical view defended in this section.
+As the cases we are considering involve agents in symmetric roles, we do not consider that possibility here.
+We acknowledge being inspired by Roth's view,  however.
 }
 %
-This fact should not obscure the distinct ways in which one conceives of another agent's actions in planning and predicting them, however.
-For, as we have argued, in some cases planning another agent's actions does not require actually predicting them.
-In some cases, planning another's actions does not depend on having predicted them but rather provides a ground for prediction.
+This fact should not obscure the distinct ways in which you conceive of another agent's actions in planning and predicting them, however.
+For, as we have argued, in some circumstances
+	 planning another's actions does not require a distinct process of predicting them; rather, prediction can be a side-effect of planning (see Section \vref{sec:distributed_plan}).
 This is why 
@@
 \label{sec:new_contrast_case}
-Now that we have some background on how some agents might each individually plan all of the agents' actions,
+Now that we have some background on how some agents might each individually make a plan for all the agents' actions,
 we can introduce a third and final case.
@@
 	This fact does not feature in their planning. 
-	Rather, each plans her own actions only and treats the other's actions as constraints to work around or opportunities to exploit.
+	Rather, each makes a plan for her own actions only, and in so doing conceives of the other's actions as constraints to work around or opportunities to exploit.
 By contrast, Caitlin and Ciaran embrace the fact that they are performing the same task. 
@@
 The addition is the idea that an agent 
-	might plan  her own actions only 
+	might make a plan for her own actions only 
 	or else 
-	might plan both her own and others' actions.
+	might make a single plan for both her own and others' actions.
 Agents often conceive of others' actions as external to their own plans, and so as either entirely irrelevant or constraints and opportunities only.
-But shared agency sometimes or always involves agents who each individually plan all of the agents' actions and so conceive of their own and the others' actions  as parts of a single plan.
+But shared agency sometimes or always involves agents who each individually make a plan for all the agents' actions and so conceive of their own and the others' actions  as parts of a single plan.
 In at least some cases, shared agency can be distinguished from individual agency in part because it involves changing perspective in this way.
@@
 The reason their walking does not involve shared agency 
-is rather that each conceives of the others' intentions and actions only as constraints and opportunities.
-Beatrice has a plan for what she, Beatrice, will do to ensure that they arrive at the corner; and Baldric likewise plans what he only will do.
-Each agent's plan relies on predictions about how the other will act and contains subplans which are conditional on the other's subplans and actions.
-But this involves no more than the sort of interdependence that might hold between two plans directed at different ends, as in cases of unshared intention.
-Beatrice and Baldric fail to make any use of the fact that they each intend the same thing, namely that they walk to the corner.
+is rather that each
+makes a plan for his or her actions only.
+Beatrice makes a plan for what she, Beatrice, will do to ensure that they arrive at the corner; and Baldric likewise plans what he only will do.
+In doing this, each conceives of the other's intentions and actions as merely constraints to work around or opportunities to exploit.
+%Beatrice and Baldric thereby fail to make use of the fact that they each intend the same thing, namely that they walk to the corner.
+By contrast,
+when Caitlin and Ciaran walk together, each makes a plan for both agents' actions and so conceives of the other's actions and her own as on a par in being elements in a single plan.
 
@@
 Each wants their house to be painted a single colour, and would prefer that it not be painted at all rather than that it be painted in several colours.
-But because they are spraying the paint onto the houses in blustery winds, neither can avoid spraying a significant quantity of paint onto the other's house as well as their own house. 
+But because they are spraying the paint onto the houses in blustery winds, neither can avoid spraying a significant quantity of paint onto the other's house. 
 In one respect this is fortunate. 
-For they will each lose so much paint from spraying in windy weather that, normally, neither would have enough paint to cover her house.
+For they will each lose so much paint from spraying in windy weather that, normally, neither would have enough paint to cover his or her house.
 But if they each make use of wind-borne spray from the other, both can succeed.
@@
 Ayesha starts painting first, using a colour which she knows is acceptable and available to Ahmed.
-Ayesha and Ahmed then each plan a route around their own house which will keep them out of the path of the other painter while also saving paint by taking advantage of foreseeable  wind-borne paint from the other.
+Ayesha and Ahmed then each plan a route around his or her own house which will keep them out of the path of the other  while also saving paint by taking advantage of foreseeable  wind-borne paint from the other.
 
@@
 From either agent's point of view, their intentions and actions are no more intimately related than are Ayesha and Ahmed's.
-This is not shared agency.
+This is not shared agency, 
+as we can see by contrasting Caitlin and Ciaran's approach to painting the house which involves each making a plan for both agents' actions.
 
@@
 The case of Beatrice and Baldric suggests that this requirement is not strong enough.
-And the case of Caitlin and Ciaran suggests that we might try strengthening it by appealing in some way to the idea that each agent can have a plan specifying roles for all of the agents.
+And the case of Caitlin and Ciaran suggests that we might try strengthening it by appealing in some way to the idea that each agent can make a plan for all the agents' actions.%
+\footnote{
+The phrase `make a plan for all the agents' actions' is used the special sense explained \vpageref{df:plan_all_actions}.
+}
 
@@
 We stipulate that an activity is \emph{distributed} just if: 
+%
 \begin{enumerate}[label=\emph{\alph*})]
+%
 \item 	there is a single outcome, G,
-	and two or more agents each have a plan for G which specifies  roles for all of the agents;	
-\item 	these plans are identical or compatible,%
-\footnote{
-The notion of compatibility is defined in footnote \vref{fn:df_compatible}.
-}
+	and two or more agents each make a plan for G which specifies  roles for all of the agents;	
 %
-		and they are each acceptable to all; 
-\item	each agent performs the roles assigned to her in her own plan;
-\item	the activity consists in nothing other than the agents performing these roles;
-	and
-\item	by performing these roles, the agents G.
-
+\item 	these plans are identical or sufficiently close that if each agent performed the role assigned to her in her own plan, any differences between the agents' plans would not significantly affect their chances of success;
+%
+\item the plan which requires each agent to perform the role assigned to her in her own plan is acceptable to all the agents, or would be if they knew what it involved; 
+\item	each agent performs the roles assigned to her in her own plan; and
+%
+\item	the activity consists in nothing other than the agents performing these roles.
+%
 \end{enumerate}
@@
 
-Many ordinary interactions involve distributed activities.
-Two housemates have agreed to make a pizza. 
-Each start with an idea of what needs doing.
-One starts by preparing the dough, so the other peels, washes and chops the vegetables.
-As the activity unfolds in this way, each gradually adds details to her plan and assigns roles.
-
-Distributed activities can be simple.
+Many ordinary interactions involve distributed activities, and distributed activities can be simple.
 In walking to a metro station together,
@@
 to strengthen Bratman's conditions for shared intention.
-We leave the conditions for shared intention exactly as Bratman specifies them (see (1)--(3) on page \pageref{quote:bratman_account})
+We leave the conditions for shared intention exactly as Bratman specifies them (see (1)--(3) on page \pageref{quote:bratman_account}),
 but add the further requirement that the agents must, in intending that they J, each conceive of J as a distributed activity.
@@
 	and 
-		to suggest a way of revising it without abandoning core features of Bratman's approach to shared intention.
+		to suggest a way of revising it without abandoning  Bratman's planning approach to shared intention.
 
@@
 we provided a counterexample to the view that a certain interlocking structure of intention and knowledge is sufficient for shared intention. 
-Two agents might lack a shared intention even though they meet conditions Bratman offers as sufficient for shared intention
+Two agents might lack a shared intention even though they meet conditions Bratman offers as sufficient for shared intention. 
 That is, they might lack a shared intention even though it is common knowledge to them that they  each intend that they J and intend that they J in accordance with and because of these intentions and meshing subplans of them.
@@
 When this happens, each agent conceives of the others' actions merely as constraints and opportunities.
-Bratman's conditions may initially appear be sufficient for shared intention because it may be natural for agents who meet his conditions to make use of the fact that there is a single outcome to which all of their actions are directed and  to conceive of all their actions as parts of a single plan.
+Bratman's conditions may initially appear be sufficient for shared intention because it may be natural for two people who meet his conditions to make use of the fact that there is a single outcome to which both of their actions are directed and  to conceive of both agents' actions as parts of a single plan.
 Our counterexample, the case of Beatrice and Baldric, shows that although this may be natural, it not necessary.
 
-We identified a way to revise Bratman's account so as to avoid this sort of counterexample without departing substantially from his planning approach.
-The remedy we propose hinges on the possibility that an agent, in planning for an outcome, can sometimes plan both her own and other agents' actions.
-Further, in the right circumstances,
-two or more agents with the same task
-	can each individually specify roles for all of the agents in planning for that task.
-Even where such planning occurs in parallel (and so need not involve no shared agency),  it can sometimes rationally result in coordinated action.
-These reflections indicate that, in giving sufficient conditions for shared intention, we can appeal to the possibility that, in acting on a shared intention, each agent has a plan for all of their actions.
-This enables us to capture the idea that agents acting on a shared intention conceive of all of their actions as part of a single plan.
+We proposed a way to fix Bratman's account to avoid this sort of counterexample without departing substantially from his planning approach.
+This fix hinges on the possibility that  an agent can sometimes  make a plan for  both her own and other agents' actions.
+Even where such planning itself requires no shared agency because it occurs without communication or prior agreement, 
+  it can sometimes rationally result in coordinated action.
+This suggests that, in giving sufficient conditions for shared intention, we can appeal to the possibility that, in acting on a shared intention, each agent individually makes a plan for all of the agents' actions (see Section \vref{sec:distributed_plan} and Section \vref{sec:fix}).
+This enables us to capture the idea that agents acting on a shared intention sometimes or always conceive of all of their actions as parts of a single plan.
 %One way to capture this idea, 
@@
 
-No amount of forming intentions about other's intentions and acting on knowledge of such intentions is sufficient, all by itself, for shared intention.  
-In many or all cases, shared agency requires changing perspective 
+Sometimes no amount of forming intentions about other's intentions and acting on knowledge of such intentions is sufficient, all by itself, for shared intention.  
+In some or all cases, shared agency requires changing perspective 
 to conceive of your own and another's intentions and actions as parts of a single plan.
